### 2010-03-23 analysis.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{Implementation and Analysis}
\label{SECIII}\label{sec:analysis}

\begin{figure*}[h!]
\begin{center}
\includegraphics[width=1.0\textwidth]{figures/flow_chart.pdf}
\caption{\label{f:flowchart} 
}
\end{center}
\end{figure*}

### 2010-03-23 appendix.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{better noise model}

The assumption that the interferometer noise is well-modeled by a multivariate normal distribution is convenient, but false.  The presence of `glitches' in the interferometer, where the noise statistics change dramatically, is well documented.  Current methods, including ours in the form proposed in the paper, are easily fooled by these bursts of excess power, simply because the analyses assume that the only way that extra power can be introduced to the interferometer is by a gravitational wave.  The gravitational wave hypothesis $H_\mathrm{signal}$ will do a very poor job of explaining temporally coincident incoherent bursts of noise power in the interferometer, but the noise hypothesis $H_\mathrm{signal}$ in its simple form does even worse; the gravitational wave explanation is thus preferred.

We can generalize the noise hypothesis to cope with glitches by creating a model for glitches and adding that hypothesis to the set under consideration.  Like gravitational waves, glitches are infrequent, have poorly known waveforms, and poorly known power.  Unlike gravitational waves, they will not be correlated between instruments.

%The Gursel-Tinto method is not robust against interferometer glitches (nor does it claim to be; real interferometer data doesn't follow the normal distribution assumed in the derivation of the method.)  This is a problem in naive attempts to use the Gursel-Tinto algorithm as a search.  Excess energy of any almost any kind will be interpreted as a evidence of a gravitational wave.  An \emph{ad hoc} addition to the method was propsed in \cite{us}.  A better way forward is to use a noise model that more accurately reflects the 'bursty' or 'glitchy' nature of the data, by replacing the normal noise distribution with a long-tailed distribution.  An immediate problem is that the marginalization integral does not in general have a closed form solution for such distributions.  If we model the long-tailed distribution using two normal distributions the marginalization integral remains tractable.

One first attempt at such a hypothesis is to propose that an interferometer is either `quiet' with some probability $p(H_\mathrm{quiet}|H_\mathrm{noise})$ and has a unit normal noise distribution, or is `glitching' with probability $p(H_\mathrm{glitch}|H_\mathrm{noise})$ and has an increased standard deviation $\sigma_g$

\begin{equation}
P(\mathbf{x}|H_\mathrm{noise})
=
\prod_{i=1}^{N}\left[p(H_\mathrm{quiet}|H_\mathrm{noise})
(2\pi)^{-n/2}
\exp(-\frac{1}{2}\sum_{j=1}^n x_{ij}^2)
+p(H_\mathrm{glitch}|H_\mathrm{noise})
(2\pi)^{-n/2}\sigma_g^{-n}
\exp(-\frac{1}{2\sigma_g^2}\sum_{j=1}^n x_{ij}^2)
\right]\label{eq:glitchy}
\end{equation}

If there is excess energy in only one detector, the new noise hypothesis will readily explain it.  If there is excess energy in three detectors, the noise hypothesis must invoke three coincident glitches and is penalized by $p(H_\mathrm{glitch}|H_\mathrm{noise})^3$ reflecting our belief that triple-coincidence glitches are rare, and the prediction that the glitches are incoherent thinly spreads the hypothesis over a higher-dimensional space than that of the signal hypothesis, which is concentrated around $\mathrm{span}\,\mathbf{F}$. These factors make it possible for the gravitational wave hypothesis to be preferred for some data.

%Instrumental glitches (hereafter, just `glitches') correspond to noise components with a greater variance than the stationary instrumental noise.

%\label{marg} The evidence for model 1 is most easily calculated in
%the original basis, where the detectors are separable.  In this
%model we assume the glitch or noise outputs of the detectors are
%uncorrelated, so that the overall evidence is simply the product of
%the evidences from each:
%
% \begin{align}
% p(d_i|M_1) &= \int p_i(\sns|M_1)p(d_i|\sns,M_1)\,{\rm d}\sns\\
%            &= \frac{1}{(2\pi)^{1/2}}\left[(1-\alpha_i)\exp(-d_i^2/2) + \frac{\alpha_i}{g_i}\exp(-d_i^2/2g^2)
%            \right],
% \end{align}
%so that
% \begin{align}
% p(\vec{d}|M_1) &= \prod_i p(d_i|M_1)\\
%                &= \frac{1}{(2\pi)^{3/2}}\prod_i \left[(1-\alpha_i)\exp(-d_i^2/2) + \frac{\alpha_i}{g_i}\exp(-d_i^2/2g^2)
%                \right].
% \end{align}

%\begin{widetext}

%Similarly the evidence for model 2 comes straight from
%Eqn.~(\ref{lik2}) by setting $\shs$ to $s^2$,  and $\sns$ to unity,
%so that the overall Bayes factor for a signal to be present is
% \be
% B_{21}=\frac{\exp\left\{ - (\vfph.\,\vec{d})^2/[2(1+s^2|\vfp|^2)]
%                          - (\vfch.\,\vec{d})^2/[2(1+s^2|\vfc|^2)]
%                          - (\vkh\,.\,\vec{d})^2/2 \right\}}
%{[(1+s^2|\vfp|^2)(1+s^2|\vfc|^2)]^{1/2} \prod_i
%\left[(1-\alpha_i)\exp(-d_i^2/2) +
%\frac{\alpha_i}{g_i}\exp(-d_i^2/2g^2)\right]}.
% \label{bayes}
% \ee
%Using the orthogonality relation
% \be
% |\vec{d}|^2 = (\vfph.\,\vec{d})^2 + (\vfch.\,\vec{d})^2 +
% (\vkh\,.\,\vec{d})^2,
% \ee
%this same result may be written in terms of power components in the
%$(\fp,\fc)$-plane and in the total power $ |\vec{d}|^2$:
% \be
% B_{21}=\frac{\displaystyle\exp\left\{ \frac{1}{2}\frac{s^2|\vfp|^2}{(1+s^2|\vfp|^2)}(\vfph.\,\vec{d})^2
%                         +\frac{1}{2}\frac{s^2|\vfc|^2}{(1+s^2|\vfc|^2)}(\vfch.\,\vec{d})^2
%                          \right\}}
%{[(1+s^2|\vfp|^2)(1+s^2|\vfc|^2)]^{1/2} \displaystyle\prod_i
%\left\{(1-\alpha_i)\displaystyle\exp\left(-\frac{d_i^2}{2}\right) +
%\displaystyle\frac{\alpha_i}{g_i}\displaystyle\exp\left[\frac{1}{2}\left(1-\frac{1}{g_i^2}\right)d_i^2\right]\right\}}.
% \ee
%
% Consider the simplified scenario in which model 1 assumes all three
%detectors are currently glitching, so that $\alpha=1$, and where the
%characteristic glitch power is the same in each, so that
%$g_i^2=g^2$. The Bayes factor is now comparing the idea that (1) we
%are seeing a triple-coincidence glitch with (2) we are seeing a
%gravitational wave, and Eqn.~(\ref{bayes}) reduces to
% \be
% B_{21}=\frac{g^3\exp\left\{\displaystyle \frac{|\vec{d}|^2}{2g^2} -
% \left[\frac{(\vfph.\,\vec{d})^2}{2(1+s^2|\vfp|^2)}
%                          +
%                          \frac{(\vfch.\,\vec{d})^2}{2(1+s^2|\vfc|^2)}
%                          + \frac{(\vkh\,.\,\vec{d})^2}{2}\right] \right\}}
%{[(1+s^2|\vfp|^2)(1+s^2|\vfc|^2)]^{1/2} }.
% \ee
%The exponent is simply the difference between the chisquared
%statistic for a glitch of mean power $g^2$ and that for a
%gravitational wave signal of mean power $s^2$.  The other terms
%represent an Occam factor, reflecting the difference in the sizes of
%the competing hypothesis spaces.
%
%The same expression can be written as
% \be
% B_{21}=\frac{g^3\exp\left[ %\displaystyle
%              \frac{1}{2}\left(\frac{1}{g^2}-\frac{1}{1+s^2|\vfp|^2}\right)(\vfph.\,\vec{d})^2
%             +\frac{1}{2}\left(\frac{1}{g^2}-\frac{1}{1+s^2|\vfc|^2}\right)(\vfch.\,\vec{d})^2
%             -\frac{1}{2}\left(1-\frac{1}{g^2}\right) (\vkh\,.\,\vec{d})^2 \right]}
%         {[(1+s^2|\vfp|^2)(1+s^2|\vfc|^2)]^{1/2} },
% \ee
%to highlight the fact that power in the null-stream direction $\vkh$
%always reduces $B$.

%\end{widetext}

### 2010-03-23 conclusions.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{Conclusions}
\label{SECV}\label{sec:conclusions}

### 2010-03-23 egsm.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Si vous voulez plus petit passez en 11pt voire encore plus petit en 10pt
\documentclass[11pt]{report}
\author{Romain Cariou}
\date{Juin 2009}
\title{rapport résumé}

\usepackage{floatrow}

%Symboles divers dont euro ...
\usepackage{marvosym}

%% Je suis francophone !
\usepackage[francais]{babel}\usepackage[T1]{fontenc}

%% Je veux utiliser néanmoins des fontes qui « paraissent bien » en PDF
\usepackage[cyr]{aeguill}

%% J'aime bien pouvoir contrôler mes hauts de page !
\usepackage{fancyhdr}

\fancyhead[R]{\leftmark} % Nom du chapitre en cours

%% Je veux pouvoir inclure des figures...
%% À commenter si vous voulez faire du DVI :
\usepackage[pdftex]{graphicx}

%% ... des figures ``jpeg'' ou ``pdf''
%% À commenter si vous voulez faire du DVI :
\DeclareGraphicsExtensions{.jpg,.pdf,.png}

%%position des images un peu partout
\usepackage{wrapfig}

%% Je veux créer des Hyperdocuments
%% À commenter si vous voulez faire du DVI :
\usepackage[pdftex,colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

%% Je contrôle la taille de ma zone imprimée...
\usepackage{anysize}
%% ...en définissants les marges {gauche}{droite}{haute}{basse}
\marginsize{16mm}{16mm}{-10mm}{5mm}

%pour la numerotation des pages
\usepackage{lastpage}
\makeatletter
\renewcommand{\@evenfoot}%
	{\hfil \upshape {\thepage} / \pageref{LastPage}}
\renewcommand{\@oddfoot}{\@evenfoot}
\makeatother

%pour changer le style des notes de bas de page
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

%pour changer les marges d'une page
\usepackage{chngpage}

% Usage:
% ------
%
%%%%%%%%%%%%%
% \changetext
%%%%%%%%%%%%%
%     The \changetext command is for changing the size and horizontal position
% of the text block on a page. The command takes 5 arguments, each of which
% is a length or is empty. i.e.,
%
% \changetext{textheight}{textwidth}{evensidemargin}{oddsidemargin}{columnsep}
%
% The given lengths are added to the corresponding current lengths and
% the remainder of the current page is typeset using the changed text block
% layout. The new layout remains in effect until another \change... command
% is issued.
%
%%%%%%%%%%%%%
% \changepage
%%%%%%%%%%%%%
%     The \changepage command is for changing the general layout of
% a page. The command takes 9 arguments, each of which is a length or is empty.
% The first 5 arguments are the same as for \changetext and have the same effect.
% The last four arguments are:
%
% \changepage{5 args}{topmargin}{headheight}{headsep}{footskip}
%
% These lengths are added to the corresponding current lengths and
% thus modify the vertical positions of the elements of the page. The
% remainder of the current page is typeset using the changed text block
% and page layout. The new layout remains in effect until another
% \change... command is issued.

%pour mettre quelquechose en filigrane
\usepackage{tikz}

%modifier les marges
\usepackage[francais]{layout}

%pour ecrire sur 2 colonnes
\usepackage{multicol}
\setlength{\columnseprule}{0pt} %largeur trait séparateur
\setlength{\columnsep}{22pt}

%pour ecrire a coté des images
\usepackage{floatflt}

\newenvironment{narrow}[2]{%
\begin{list}{}{%
\setlength{\topsep}{0pt}%
\setlength{\leftmargin}{#1}%
\setlength{\rightmargin}{#1}%
\setlength{\listparindent}{\parindent}%
\setlength{\itemindent}{\parindent}%
\setlength{\parsep}{\parskip}}%
\item[]}{\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Je peux définir mes propres « commandes »...
%Pour changer le titre de la bibliographie
\renewcommand{\bibname}{R\'ef\'erences}

%% pour citer une figure :
\newcommand{\figref}[1]{figure~\ref{#1}}

%% pour citer une équation :
\newcommand{\Ref}[1]{(\ref{#1})}

%% pour mettre l'emphase sur un mot ou un groupe de mots :
\newcommand{\empha}[1]{\textit{\textbf{#1}}}

%% opérateurs de dérivation :
\newcommand{\D}{\partial}
\newcommand{\Dt}{\partial_t}
\newcommand{\Dx}{\partial_x}
\newcommand{\Dy}{\partial_y}

\begin{document}
%filigrane
\begin{tikzpicture}[remember picture,overlay]
	\node[rotate=60,scale=15,text opacity=0.1]
	  at (current page.center) {Confidential};
\end{tikzpicture}

%pour changer le titre de la Bibliographie
\renewcommand{\bibname}{R\'ef\'erences}

\author{Romain Cariou}
\title{Fiche résumé début de stage}

%% Voilà mes légendes de figures comme je les aime
\makeatletter
\def\figurename{{\protect\sc \protect\small\bfseries Fig.}}
\def\f@ffrench{\protect\figurename\space{\protect\small\bf \thefigure}\space}
\let\fnum@figure\f@ffrench%
\let\captionORI\caption
\def\caption#1{\captionORI{\rm\small #1}}
\makeatother

%Pour les numerotations
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\thesection .\Alph{subsection}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Couverture :
%\thispagestyle{plain}
{\Large

\begin{center}
%\vskip1cm
%% Le \vphantom{\int_\int} sert à introduire de l'espace entre les deux lignes
%% (essayez donc de le commenter)
\textbf{{\LARGE Bias in geometric sample mean}}
%$ \vphantom{\int_\int}$
 %\\
%\textbf{{\LARGE de l'expérience LIGO : Lloid}}
\begin{center}
\rule[1ex]{5cm}{0.15ex}
\end{center}
\end{center}
\vspace*{-0.8cm}
\begin{center}
\textsl{ Romain Cariou }
\end{center}

\normalsize
\vspace*{0.5cm}
\begin{multicols}{2}

Here we compute the bias factor $\beta$ of the sample median of a chi-square distributed variables relative to the sample geometric mean.\\ Let's define n chi-square distributed variable : $x_{l}$ l$\in\left[\hspace{-1ex}\left[\hspace{0.5ex}1,n\right]\kern-0.15em\right]$. The probability density function for each $x_{l}$ is given by : $f(x)=\frac{1}{2} e^{-x/2}$, where $$2=\langle x \rangle = \int_{0}^{\infty}xf(x)dx$$ is the \textit{expected value} of the random variable $x$. The median, $x_{1/2}$ is defined by $$ \frac{1}{2}=\int_{0}^{x_{1/2}}f(x)dx$$ which yields $x_{1/2}=2\,ln2$. Thus the bias of the \textit{expected value} is $ln2$.\\
Let's $\rho$ be the geometric sample mean of the chi-square variables : $$\rho=\left\langle(\prod_{l=1}^{n}x_{l})^{1/n}\right\rangle=\left\langle exp \left[\frac{1}{n}\sum_{l=1}^{n} \ln x_{l}\right] \right\rangle$$ so $\ln \rho$ is the arithmetic mean from the logarithm of the discret distribution and thus, for a continuous variable as $x_{l}$ we have : $$ \rho=exp \left(\int_{0}^{\infty}\ln x\,f(x)\,dx \right)$$ $$=exp \left(\frac{1}{2} \int_{0}^{\infty}e^{-x/2}\,\ln x dx \right)$$
The integral above can be calculated\footnote{\label{1}\textit{Table of integrals, series and products}, I.S. Gradshteyn/I.M. Ryzhik, BI ((256))(2).} and we find : $$ \rho\,\,=\,\frac{2}{e^{C}}= 1.122918967...\,. $$ where $C\,= \lim\limits_{s\to \infty}\left(\sum_{m=1}^{s}\frac{1}{m}-\ln s \right)$ $= 0.577215...\,.$\\
As the expected value of uncorrelated random variable is the product of the expected values (E(XY)=E(X)E(Y)), we can calculate the expected value of the geometric sample mean :$\langle \rho \rangle = \langle x_{l}^{1/n} \rangle^{n}$. The probability density function of the new random variable $Y=x^{1/n}$ is : $$P(Y)=\frac{n}{2}.e^{-Y^{n}/2}.Y^{n-1}$$ and so $$
\langle \rho \rangle = \left[\frac{n}{2}\int_{0}^{\infty}Y^{n}e^{-Y^{n}/2}dY \right]^{n}$$ $$=\frac{1}{2^{n}}\left[\int_{0}^{\infty}z^{1/n}e^{-z/2}dz\right]^{n}$$ with $z=Y^{n}$. This integral can be easily calculated\footnote{\label{2}\textit{Table of integrals, series and products},  I.S. Gradshteyn/I.M. Ryzhik, FI 2 779.} using the GAMMA function we have : $$ \langle \rho \rangle= \frac{2}{n^{n}}\left[\Gamma(\frac{1}{n})\right]^{n}$$
This result makes sense : we can calculate the exact limit as n $\rightarrow \infty$ : $\lim\limits_{n\to \infty}\langle \rho \rangle=\rho$ as illustrated on the figure below, which represent the expected value of the geometric sample mean versus n :

\vspace*{0.2cm}
\includegraphics[scale=0.19]{geosamplemean.png}

Therefore, assuming that expected value of the sample median can is \footnote{FINDCHIRP : an algorithm for detection of gravitational waves from inspiraling compact binaries, Appendix B.}: $$\langle x_{med}\rangle = 2 \times \sum_{l=1}^{n}\frac{(-1)^{l+1}}{l}$$
we have the following relation : $$\langle x_{med}\rangle\langle \rho\rangle=\frac{4}{n^{n}}\left[\Gamma(\frac{1}{n})\right]^{n}\sum_{l=1}^{n}\frac{(-1)^{l+1}}{l}$$
Thus the bias factor is : $$ \beta=\frac{2\sum_{l=1}^{n}\frac{(-1)^{l+1}}{l}}{\frac{2}{n^{n}}\left[\Gamma(\frac{1}{n})\right]^{n}}$$

\end{multicols}

\end{document}

### 2010-03-23 inspiral_svd.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass[aps,prd,twocolumn,showpacs,groupedaddress,showkeys,preprintnumbers]{revtex4}
\input{packages}
\input{macros}

% $Id:$

\begin{document}

\title{Low-latency search for gravitational-waves: prospects for near real time detection of compact binary coalescence}

\date{\today}

\author{Kipp Cannon} 
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Adrian Chapman} 
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Chad Hanna}  
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Drew Keppel}  
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Antony C. Searle}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Alan J. Weinstein}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\begin{abstract}

\end{abstract}

\pacs{}

\keywords{Gravitational Waves, Laser Interferometry}

\preprint{}

\maketitle

\input{introduction}

\input{method}

\input{analysis}

\input{results}

\input{conclusions}

\begin{acknowledgments}
LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the National Science Foundation and operates under cooperative agreement PHY-0107417. This paper has LIGO Document Number LIGO-P0900004-v1

\end{acknowledgments}

\appendix

\bibliography{references}

\end{document}

### 2010-03-23 introduction.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{Introduction}
\label{sec:introduction}
Modern gravitational-wave observatories are likely to deliver on the promise 
that their name implies by observing multiple gravitational-wave signals from
compact binary coalescence within the coming decade ~\cite{kopparapu2008}. 
There is now a world-wide network of gravitational-wave observatories 
including the laser interferometers
LIGO, VIRGO, GEO, and TAMA, which should assure that coincident detections 
will be possible among at least three separate sites lending confidence
to any detections.
~\cite{Abbott:2005qm, beauville2008, beauville2006, blackburn2005, LIGOS1instpaper}.  
However, having an electromagnetic or neutrino counterpart to a 
gravitational-wave detection would not only increase the confidence in 
the detection but would also greatly improve the astrophysical information
available from the event.

Work has already commenced to trigger gravitational-wave searches from 
electromagnetic observations ~\cite{triggeredsearches2008}. LIGO has recently
ruled out the possibility of GRB070201 originating from the merger of a
neutron star / neutron star or neutron star / black hole system in the
Andromeda galaxy ~\cite{GRB070201}.  The still open question is, will 
gravitational-wave observations be able to trigger electromagnetic followup?

This question has been investigated and it seems  at least
possible that gravitational-wave detections could provide a region of the
sky to prompt electromagnetic observation followups~\cite{sylvestre2003}.
Much work is currently underway in providing the best possible source 
localization from gravitational-wave detector networks
~\cite{markowitz2008, raymond2008, cavalier2006}.  Colloborations are
currently forming to provide infrastructure for the gravitational-wave
Astronomers to provide targets of opportunity for electromagnetic 
Astronomers ~\cite{kanner2008}.  Most models that predict simultaneous
gravitational-wave and electromagnetic observations also predict that the
peak amplitude of electromagetic radiation will occur soon after 
gravitational-wave emmission ~\cite{sylvestre2003}.  Thus in order to maximize
the chance of a successful electromagnetic followup the latencey of 
gravitational-wave signal analysis must be made minimal.

The parameter space of compact binary coalescence signals is large
~\cite{Owen:1995tm, Owen:1998dk} .  It is a
computationally burdensome task to analyze these signals with even moderate
latency ~\cite{Abbott:2007xi}.  
This work will describe how to exploit degeneracy in the signal
parameter space to answer more quickly whether or not a gravitational-wave 
is present, with what confidence, and where it is coming from. Specifically
we will explore using the singular value decomposition (SVD) to reduce the
effective number of filters required to search the data.  We note that others
have applied the use of SVD to gravitational wave data analysis to analyze
optimal gravitational-wave burst detection 
~\cite{bradyraymajumder2004, heng2008} and coherent networks 
of detectors ~\cite{wen2008}

The paper is organized as follows.  First we provide an overview of the 
standard method for detecting compact binary coalescence signals and
describe how it can be modified to accomodate low latency analysis.  We
then describe the pipeline we have constructed to implement these changes.
To validate the approach we present results of simulations and finish
with some concluding remarks.

### 2010-03-23 method.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{Method}
\label{SECII}\label{sec:method}
\subsection{The standard approach: matched filtering}
Compact binary coalescence searches typically employ matched filtering as the
first step to locating a gravitational wave signal ~\cite{findchirppaper}.  
A template waveform is convolved with the whitened data, and if 
properly normalized it produces a signal-to-noise ratio (SNR) that indicates the
probability of a signal being present ~\cite{wainstein:1962}.  
Since the parameters of the signal are
not known apriori, multiple templates must be filtered for a wide range of
parameters.  Let us assume that the waveforms are simply parameterized by
component mass and denote the $i$th template waveform of $N$ templates necessary
to search a given parameter space as $h_i(t) = f(m1,m2,t)$.
The SNR, $\rho_i$ for that template is
\begin{equation}
\label{eq:time_domain_snr}
\rho_i(t) = \int h_i(t) s(t-\tau) \, d\tau,
\end{equation}
where $s(t)$ is the detector data stream suitably whitened having mean zero
and unity variance.  Time domain convolutions don't have any inherent latency. 
The template waveform can be ``slid onto" the detector data as it is produced
giving the possibility of measuring a peak SNR at the time of inspiral
coalescence.  Typically it is not tractable to compute SNR via a
time domain convolution as indicated in (\ref{eq:time_domain_snr}) because
it involves an $\mathcal{O}(N^2)$ operation per template. Instead the
convolution theorem is applied and the SNR is obtained via an inverse Fourier
transform
\begin{equation}
\label{eq:freq_domain_snr}
\rho_i(t) = \int \tilde{h}_i(f)\,^*\,\tilde{s}(f)\,e^{-2\pi i f t}\, df,
\end{equation}
where the $\tilde{h}_i(f)$ and $\tilde{s}(f)$ denote the Fourier transforms
of the template and detector data respectively.  This operation is far
faster to compute with $\mathcal{O}(N\,{\rm log}\,N)$ operations per template.  
However, the acausal nature of the Fourier transform introduces an inherent
latency.  One must wait until they have collected data as long as their 
signal and for minimally overlapping computations that means having a latency
about as long as the template waveform.  The latency can be reduced by having
redudant overlapping transforms, but at an additional cost.

Thus we have that low latency searches are inherently more computationally
costly than moderate latency searches that employ frequency domain matched
filtering.  The next section discusses a method to mitigate the cost of
low latency searches.

\subsection{Reducing the number of filters with singular value decomposition}
It is clear that in order to reduce the cost of a low latency search for
compact binaries of unknown mass we need to reduce the number of convolutions
necessary to find the gravitational-wave signal.  Rather than filter the
$i$ templates corresponding to binaries of different component masses it is
possible to find a basis set of fewer filter $u_j$ that can be linearly combined
to reproduce $\rho_i$ to some accuracy.  For the remainder of this section
we will drop the explicit time dependence of the convolution and recognize
that the snr for any point in time is simply the innerproduct between a
template waveform vector ${\bf \hat{h}_i}$ and a vector of the detector data 
${\bf \hat{s}}$, $\rho_i = ({\bf \hat{h}_i}|{\bf \hat{s}})$.  
The goal is to have
\begin{equation}
\label{eq:linear_combination_of_snr}
\rho_i = \sum_{j=1}^{N'} M_{ij} ({\bf \hat{u}_j}|{\bf \hat{s}})
\end{equation}
where the number of inner products is reduced from $N$ to $N'$. In order to
find the basis set ${\bf u_j}$ we use the singular value decomposition of
the matrix 
${H_{ik}} = \{{\bf \hat{h}_1}, {\bf \hat{h}_2}, ...{\bf \hat{h}_N}\}$. Where
$i$ is the index over the template number and $k$ is the index over sample 
points. The singular value decomposition factors a matrix such that
\begin{equation}
\label{eq:svd}
{H_{ik}} = {A_{ij}\,B_{jl}\,U_{lk}}
\end{equation}
The matrix ${B_{jl}}$ is a diagonal matrix of singular values ranked in
order of importance of reconstructing the original matrix ${H_{ik}}$.
However since a search for binary inspirals only needs  waveform accuracy to a
few percent to be successful it is possible to make an approximate 
reconstruction of ${H_{ik}}$ 
\begin{equation}
\label{eq:svdapprox}
{H_{ik}} \approx {A_{ij}\,B'_{jl}\,U_{lk}}
\end{equation}
where the rank of ${B'_{jl}}$ is less than the rank of ${B_{jl}}$.
This reduces the number of rows in ${U_{lk}}$.  Now we have a 
new basis of template waveforms derived from ${U_{lk}} = \{{\bf \hat{u}_1}, {\bf \hat{u}_2}, ...{\bf \hat{u}_{N'}}\}$ and can write 
(\ref{eq:linear_combination_of_snr})
\begin{eqnarray}
\label{eq:svdsnr}
\rho_i &=& \sum_{j=1}^{N'} A_{il}\,B'_{lj} ({\bf \hat{u}_j}|{\bf \hat{s}}) \nonumber \\
       &=& \sum_{j=1}^{N'} M_{ij} ({\bf \hat{u}_j}|{\bf \hat{s}})
\end{eqnarray}

This result differs from other work that models gravitational-wave chirp
signals in approximate ways 
~\cite{Chassande-Mottin2006, Candes2008, BuonannoChenVallisneri:2003a} 
by starting with an exact representation of the desired template family and
producing a rigorous approximation with a tunable accuracy.

\subsection{Detection statistic}
It is not necessary to reconstruct the SNR of all templates in order to 
determine if a signal is present.  By thresholding on the sum of squares
of the orthoganal decomposition described above it is possible to answer
whether or not an inspiral like event is in the data.  Then only 
potentially interesting times can be reconstructed on a sample by sample 
basis.  Following the ideas in ~\cite{Anderson:2000yy} we employ a detection
statistic as
\begin{equation}
\label{eq:stat}
\Gamma^2 = \sum_{l} \sum_{j} {B'_{lj}}^2 {({\bf \hat{u}_j}|{\bf \hat{s}})}^2
\end{equation}

\subsection{Signal consistency check}
After reconstructing the SNR of an individual template it is possible to 
to check that the components along the SVD basis were consistent with 
expectation by performing a $\chi^2$ fit to the decomposition.  
\begin{equation}
\label{eq:chisq}
\chi^2_i = \sum_j {[({\bf \hat{u}_j}|{\bf \hat{s}}) - \rho_i / M_{ij}]}^2
\end{equation}

\subsection{Selectively reducing the sample rate of the data and template
waveforms}
The computational cost of filtering can be further reduced by selectively
downsampling certain intervals of the template waveform.  Note that the
frequency of gravitational radiation resulting from the inspiral of two
compact object is given by the quadrupole aproximation as ~\cite{finn1993}
\begin{equation}
\label{eq:fgw}
f_{\rm GW} = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t_c} \right]^{3/8}
\end{equation}
where $\mathcal{M}$ is the chirpmass in units of $G=c=1$ and $t_c$ is the time
before the coalescence of the binary.  Typically the filter is integrated to
a finite frequency, $f_e$ which corresponds to a time, $t_e$ before the 
coalescence~\cite{kidder1992, blanchet2002, findchirppaper, hanna2009}.  Given
the relationship $f/f_e = (t_e/t)^{3/8}$ it is possible to reduce a
filter waveform into critically sampled bands by dividing them up in time.
This is possible regardless of the time-frequency relationship as long as
it can be computed rigorously and thus the method doesn't depend on 
(\ref{eq:fgw}).  As a concrete example consider a waveform obeying 
(\ref{eq:fgw}) with a chirp mass of 1 $M_{\odot}$ for which we decide
to cut the filter at $f_e = 2048$Hz (corresponding to $t_e = 0.00093$s).  
It is then possible to compute sub bands which are each critically sampled
according to table \ref{table1}  
\begin{table}[!h]
\caption{
Example critically sampled sub bands for a 1 $M_{\odot}$ filter with a 
time frequency structure given by ($\ref{eq:fgw})$.  Most of the filter is
low frequency and it is possible to save computation by breaking up the filter
into bands each of which has restricted frequency content. 
}
\begin{center}
\label{table1}
\begin{tabular}{@{} lll @{}}
\hline
\hline
Sub band & Sample rate (Hz) &Interval (s) \\
\hline
1 & 2048 & [0.00093, 0.0059] \\
2 & 1024 & [0.0059, 0.037]  \\
3 & 512  & [0.037, 0.24]   \\
4 & 256  & [0.24, 1.51]   \\
5 & 128  & [1.51, 9.60]   \\
6 & 64   & [9.60, 60.95]  \\
7 & 32   & [60.95, 387.0]  \\
\hline
\end{tabular}
\end{center}
\end{table}
where it is assumed that the filter is cut to some finite length, 387s 
for this example.  This idea has been demonstrated by the Virgo Collaborations
MBTA pipeline~\cite{beauville2006, beauville2008}.

\subsection{Comparison of computational costs and latency}
Convolutions are best done with the use of the Fast Fourier Transform (FFT).
However there is an inherent latency in performing convolutions this way. This
section describes the computational cost and latency of FFT convolutions, time 
domain (TD) convolutions and TD convolutions with SVD.

Convolving one filter with data using FFTs involves 
$\mathcal{O}[10 N_s \log_2(N_s)]$
floating point operations where $N_s$ is the number of sample points in
the data being convolved and results in $N_s - N_f$ samples of usable output.  
In comparison time domain convolutions require
$\mathcal{O}[N_s N_f]$ where $N_f$ is the number of sample points in the
filter.  The FFT is faster to compute when $N_f > 10 \log_2(N_s)$.  However
the FFT has an inherent latency of $\mathcal{O}[N_s/f_s]$ seconds, where
$f_s$ is the sample rate when minimally overlapping FFTs are computed. It is
possible to reduce the latency of the FFT by performing vastly overlapping
FFTs.  The cost of such a procedure would be 
$\mathcal{O}[10 N_o N_s \log_2(N_s)]$.  Thus for comparable computational
cost where $N_f \approx 10 N_o \log_2(N_s)$ the latency of the FFT method 
would be $\mathcal{O}[N_s/f_s/N_o]$ seconds.  Depending on the task at hand,
latency requirements and computational costs can dictate what procedure to
use.

\subsection{Data Whitening}
Matched filtering the SVD basis vectors has motivated a decoupling of the
whitening routine from the matched filtering engine. This was necessary in
order weight templates appropriately by whitening them before the SVD
calculation, and, as such, we have also moved the data whitening outside the
match filter engine. We have chosen to whiten the data using a running
geometric average of the PSD computed by Hann windowing 8 second buffers of
data and using 50\% overlapping buffers. The running geometric average is
updated once per buffer from a median of the recent PSDs. We find that this
algorithm is extremely fast to converge to an accurate PSD and also very robust
against glitches. This is demonstrated below where we have injected
band-limited white noise burst with a central frequency of 150 Hz, a bandwidth
of 100 Hz, a time duration of 0.2 seconds, and a injection density of one
injection every $100/\pi \approx 32$ seconds. At this density we find there
should only ever be one injection affecting the median history so the glitches
should minimally affect PSD estimation as can be seen by comparing the local
RMS amplitude between injections to that without injections.

\begin{center}
\resizebox{0.9\textwidth}{!}{\includegraphics{figures/whiten_burst_step_100_semilogy.png}}
\end{center}

An inspiral, to leading order, has gravitational-wave frequency evolution given
by
\begin{equation}
f(t) = \frac{1}{8 \pi \mathcal{M}} \left( \frac{t_c - t}{5 \mathcal{M}} \right)^{-3/8}\;,
\end{equation}
whose time derivative is given by
\begin{equation}
\frac{df}{dt} = \frac{3}{320 \pi \mathcal{M}^2} \left( \frac{t_c - t}{5 \mathcal{M}} \right)^{-11/8}\;.
\end{equation}
Combining these, we find the frequency evolution as a function of frequency to
be
\begin{equation}
\frac{df}{dt}(f_0) = \frac{3}{320 \pi \mathcal{M}^2} \left( 8 \pi \mathcal{M} f_0 \right)^{3/11}\;,
\end{equation}
which can be inverted to get the minimum frequency which has a given frequency
derivative,
\begin{equation}
f_0 = \frac{1}{8 \pi \mathcal{M}} \left( \frac{320 \pi \mathcal{M}^2}{3} \frac{df}{dt} \right)^{11/3}\;.
\end{equation}

If we allow frequency bins to be affected by an injection only once within the
median history, we need to calculate the corresponding $df/dt$ for our PSD
calculation. An FFT of a buffer with length $T$ will have a frequency
resolution $df = 1/(2T)$. Hann windowing the data and overlapping buffers by
50\% introduces correlations in neighboring frequency bins of the FFT. This
means that we actually want the frequency to change by $df = 3/(2T)$ before the
next PSD calculation, which happens a $dt = T/2$ later, resulting in a minimum
$df/dt = 3/T^2$.

Combining these results we find the lower frequency bound for our injections,
assuming a chirp mass for a binary with $m_1 = m_2 = 1 M_{\odot}$ and $T = 8$
s, is 23 Hz.

We want to compute the equivalent injection density which would be biased as
much as lalapps\_inspiral's PSD estimator, which allows $\sim3$ injections per
2048 seconds. It computes the PSD by breaking the segment up into 16 256 second
chunks. These chunks are then combined into to sets of 8 chunks each. The
median of each set is then averaged across the sets for each frequency bin to
produce the PSD estimate for that 2048 second segment. This procedure results
in an average injection density of 1.5 per 8 PSDs in the median history, which
is comparable to LLOID's 1 per 7 PSDs in the median history. Since LLOID has an
injection history of 7 PSDs which span 32 seconds, this means LLOID can perform
one injection every 32 seconds, or 64 injections per 2048 second segment, more
than an order of magnitude increase in density over lalapps\_inspiral.

### 2010-03-23 outline.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass[aps,prd,showpacs,groupedaddress,showkeys,preprintnumbers]{revtex4}
\input{packages}
\input{macros}

\begin{document}

\title{(OUTLINE DRAFT)\\ Low-latency search for gravitational-waves: prospects for near real time detection of compact binary coalescence}

\date{\today}

\author{Kipp Cannon}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Adrian Chapman}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Chad Hanna}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Drew Keppel}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Antony C. Searle}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Alan J. Weinstein}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\begin{abstract}
The future of gravitational-wave astronomy permits the possibility that 
gravitational-wave detections from inspiralling compact objects could direct
electromagnetic followup searches. In order to be practical the latency of
gravitational-wave searches informing other astronomical instruments 
would have to be small to catch the peak electromagnetic radiation output. 
It is desirable to consider time domain filters for gravitational-wave data
to reduce the latency.   Typically the cost of time domain convolutions 
is prohibitive for any realizable analysis.  Here we present a method to
make the cost of time domain convolutions reasonable with the hope that it
could be employed for near real time detection of gravitational-wave signals
from compact binary mergers.  
\end{abstract}

\pacs{}

\keywords{Gravitational Waves, Laser Interferometry}

\preprint{}

\maketitle

\begin{enumerate}
\item{Introduction}
\begin{itemize}
\item{Introduce how GW searches could direct EM searches and do literature review}
\item{Introduce need for low latency searches and setup the discussion of our
method as well as a literature review for similar ideas}
\item{Outline the rest of the paper}
\end{itemize}
\item{Method}
\begin{itemize}
\item{Describe SVD decomposition of template banks}
\item{Describe selective downsampling of waveform portions}
\item{Describe computational cost / latency of analysis and compare to the 
standard method (theoretical)}
\end{itemize}
\item{Analysis}
\begin{itemize}
\item{Describe the input to the simulations we will do, i.e. the template bank,
the match conditions, the whitening, all the gory details...}
\end{itemize}
\item{Results}
\begin{itemize}
\item{Show results of our ability to reconstruct waveforms and compare to 
``theory"}
\item{Show results of the latency and computational costs and compare to ``theory"}
\end{itemize}
\item{Conclusions}
\begin{itemize}
\item{Sum up the status and describe the future work}
\end{itemize}
\end{enumerate}
\begin{acknowledgments}
LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the National Science Foundation and operates under cooperative agreement PHY-0107417. This paper has LIGO Document Number LIGO-???
\end{acknowledgments}

\end{document}

### 2010-03-23 results.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{Results}
\label{SECIV}\label{sec:results}

### 2010-03-23 wbayesian.tex
Moving directories a little
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@369 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass{article}
\usepackage{amssymb}
\begin{document}
\title{Bayesian search for sine-Gaussian bursts}
\author{Antony C. Searle}
\maketitle
\section{What are we computing?}

Suppose we have a white noise detector that makes a vector of observations $\mathbf{x}$.  These are distributed according to
\begin{eqnarray}
p_\mathbf{0}(\mathbf{x})&\propto&\exp(-\mathbf{x}^2/2)
\end{eqnarray}
If there is some additive signal $\mathbf{y}$ as well as the noise, then $\mathbf{x}-\mathbf{y}$ will have the same distribution.
\begin{eqnarray}
p_\mathbf{y}(\mathbf{x})&\propto&\exp(-(\mathbf{x-y})^2/2)
\end{eqnarray}
The way to decide between these situations is to look at the ratio
\begin{eqnarray}
\frac{p_\mathbf{y}(\mathbf{x})}{p_\mathbf{0}(\mathbf{x})}=\frac{\exp(-(\mathbf{x-y})^2/2)}{\exp(-\mathbf{x}^2/2)}=\exp(\mathbf{xy}-\mathbf{y}^2/2)
\end{eqnarray}
Bayesian and non-Bayesian statistics agree thus far.
How do we deal with an unknown $\mathbf{y}$?  The Bayesian method is to average this ratio over $\mathbf{y}$.  We might do this by computing
\begin{eqnarray}
\frac{ \frac{1}{V}\int_V p_\mathbf{y}(\mathbf{x}) d\mathbf{y}}{p_\mathbf{0}(\mathbf{x})}
\end{eqnarray}
if $\mathbf{y}$ is constrained to lie in some region of parameter space with volume $V$.  More generally, we can define some function $q(\mathbf{y})$ with
\begin{eqnarray}
\int_{\mathbb{R}^n} q(\mathbf{y}) d\mathbf{y} = 1
\end{eqnarray}
and computing the $q$-weighted average
\begin{eqnarray}
\frac{ \int_{\mathbb{R}^n} q(\mathbf{y}) p_\mathbf{y}(\mathbf{x}) d\mathbf{y}}{p_\mathbf{0}(\mathbf{x})}
\end{eqnarray}
The search computes this expression, slightly generalized to colored noise with covariance $\mathbf{\Sigma}$ by replacing $\mathbf{x}^2$ with $\mathbf{x}^T\mathbf{\Sigma}^{-1}\mathbf{x}$ (and likewise replacing $(\mathbf{x-y})^2$ with $(\mathbf{x-y})^T\mathbf{\Sigma}^{-1}(\mathbf{x-y})$

By judicious definition of $q$, we can restrict the analysis to certain populations of signal (such as $\mathbf{y}$ always a sine-gaussian) and encode our expectations about the relative abundance of signals (such as small signals being more common than large signals).  $q$ is the Bayesian prior plausibility distribution for the signal population, and the quantity the algorithm computes is the Bayes factor.  $q$ is a subjective choice, but as any $q$ makes definite statements about the anticipated signal population, it is constrained by the need to make reasonable statements.

If we choose a $q$ that is Gaussian (or Gaussian on some of its parameters), we can solve the averaging integral analytically (or solve it for some of its parameters).

%How does the non-Bayesian statistic deal with unknown $\mathbf{y}$?  One method is to maximize the statistic for individual $\mathbf{y}$
%\begin{eqnarray}
%\max_\mathbf{y}\frac{p_\mathbf{y}(\mathbf{x})}{p_\mathbf{0}(\mathbf{x})}
%\end{eqnarray}
%or maximize with $\mathbf{y}$ constrained to some subset of parameter space $S$
%\begin{eqnarray}
%\max_{\mathbf{y}\in S}\frac{p_\mathbf{y}(\mathbf{x})}{p_\mathbf{0}(\mathbf{x})}
%\end{eqnarray}
%or maximize the expression biased by some regulating function $\alpha(\mathbf{y})$
%\begin{eqnarray}
%\max_\mathbf{y}\frac{p_\mathbf{y}(\mathbf{x})}{p_\mathbf{0}(\mathbf{x})}\alpha(\mathbf{y})
%\end{eqnarray}

\section{Notation}

For an explanation of notation, refer to the draft Bayesian paper by browsing the lscsoft repository

\begin{verbatim}
http://www.lsc-group.phys.uwm.edu/cgi-bin/cvs/viewcvs.cgi/?cvsroot=lscsoft
\end{verbatim}

for

\begin{verbatim}
/matapps/src/searches/burst/coherent-network/doc/bayesian/
\end{verbatim}

or at the link below

\begin{verbatim}
http://www.lsc-group.phys.uwm.edu/cgi-bin/cvs/viewcvs.cgi/
*checkout*/matapps/src/searches/burst/coherent-network/doc/
bayesian/bayesian.pdf?rev=1.42&cvsroot=lscsoft&content-type
=application/pdf
\end{verbatim}

\section{Analysis}
The Bayes factor for a burst search with multivariate normal distribution priors on noise and signal waveform is
\begin{eqnarray}
\frac{
p(\mathbf{x}|H_\textrm{signal},I)
}{
p(\mathbf{x}|H_\textrm{noise},I)
}
&=&
\int_{-\infty}^{+\infty}\ldots\int_{-\infty}^{+\infty}
p(\mathbf{\rho},\tau,\theta,\phi|H_\textrm{signal},I)
\sqrt{\det(\mathbf{I}-\mathbf{\Sigma}\mathbf{K})}
\exp(\frac{1}{2}\mathbf{x}^T\mathbf{K}\mathbf{x})
\mathrm{d}\mathbf{\rho}\mathrm{d}\tau\mathrm{d}\theta\mathrm{d}\phi\nonumber
\end{eqnarray}
where the kernel is
\begin{eqnarray}
\mathbf{K}(\mathbf{\rho},\tau,\theta,\phi)&=&
(\mathbf{\Sigma}^{-1}\mathbf{F}\mathbf{W})
(
(\mathbf{F}\mathbf{W})^T
\mathbf{\Sigma}^{-1}
\mathbf{F}\mathbf{W}
+
\mathbf{A}^{-1}
)^{-1}
(\mathbf{\Sigma}^{-1}\mathbf{F}\mathbf{W})^T.\nonumber
\end{eqnarray}
The matrix $\mathbf{W}$ contains the sine- and cosine-Gaussian waveforms for each polarization.  The matrix $\mathbf{A}$ specifies the distribution of the amplitudes of those waveforms.  The matrix $\mathbf{F}$ is the detector response, mapping the waveforms onto the output by considering the effects of directional sensitivity and time-delay.

In this case, our signal model's extra variables bundled up in $\rho$ are central frequency $f$, quality factor $Q$ and amplitude $\sigma$.

For a system of detectors with independent noise,
\begin{eqnarray}
\mathbf{\Sigma}&=&\left[
\begin{array}{cccc}
\mathbf{\Sigma}_1 & \mathbf{0} & \cdots & \mathbf{0} \\
\mathbf{0} & \mathbf{\Sigma}_2 & \cdots & \mathbf{0} \\
\vdots & \vdots & \ddots & \vdots \\
\mathbf{0} & \mathbf{0} & \cdots & \mathbf{\Sigma}_N \\
\end{array}
\right].
\end{eqnarray}
For data that is appropriately timeshifted before analysis,
\begin{eqnarray}
\mathbf{F}&=&\left[
\begin{array}{cc}
F_1^+\mathbf{I} & F_1^\times\mathbf{I} \\
F_2^+\mathbf{I} & F_2^\times\mathbf{I} \\
\vdots & \vdots \\
F_N^+\mathbf{I} & F_N^\times\mathbf{I} \\
\end{array}
\right].
\end{eqnarray}
For sine-Gaussians, the waveforms are a vector $\mathbf{w}_1$ sampling the cosine-Gaussian and a vector $\mathbf{w}_2$ sampling the sine-Gaussian, so the model is
\begin{eqnarray}
\mathbf{W}&=&\left[
\begin{array}{cccc}
\mathbf{w}_1 & \mathbf{w}_2 & \mathbf{0} & \mathbf{0} \\
\mathbf{0} & \mathbf{0} & \mathbf{w}_1 & \mathbf{w}_2
\end{array}
\right]
\end{eqnarray}
and the amplitude prior covariance matrix
\begin{eqnarray}
\mathbf{A}&=&\left[
\begin{array}{cccc}
\sigma^2 & 0 & 0 & 0 \\
0 & \sigma^2 & 0 & 0 \\
0 & 0 & \sigma^2 & 0 \\
0 & 0 & 0 & \sigma^2 \\
\end{array}
\right].
\end{eqnarray}
We do not rely on any properties of sine-Gaussians, so the derivation holds for any two waveforms, and generalizes to any number of waveforms (though for many waveforms it will no longer be fast.)

Therefore,
\begin{eqnarray}
\mathbf{x}^T\mathbf{\Sigma}^{-1}\mathbf{FW}&=&
\mathbf{x}^T\left[
\begin{array}{cccc}
\mathbf{\Sigma}_1^{-1} & \mathbf{0} & \cdots & \mathbf{0} \\
\mathbf{0} & \mathbf{\Sigma}_2^{-1} & \cdots & \mathbf{0} \\
\vdots & \vdots & \ddots & \vdots \\
\mathbf{0} & \mathbf{0} & \cdots & \mathbf{\Sigma}_N^{-1} \\
\end{array}
\right]
\left[
\begin{array}{cc}
F_1^+\mathbf{I} & F_1^\times\mathbf{I} \\
F_2^+\mathbf{I} & F_2^\times\mathbf{I} \\
\vdots & \vdots \\
F_N^+\mathbf{I} & F_N^\times\mathbf{I} \\
\end{array}
\right]\mathbf{W}
\\
&=&
\mathbf{x}^T\left[
\begin{array}{cc}
F_1^+\mathbf{\Sigma}_1^{-1} & F_1^\times\mathbf{\Sigma}_1^{-1} \\
F_2^+\mathbf{\Sigma}_2^{-1} & F_2^\times\mathbf{\Sigma}_2^{-1} \\
\vdots & \vdots \\
F_3^+\mathbf{\Sigma}_3^{-1} & F_3^\times\mathbf{\Sigma}_3^{-1}
\end{array}
\right]\left[
\begin{array}{cccc}
\mathbf{w}_1 & \mathbf{w}_2 & \mathbf{0} & \mathbf{0} \\
\mathbf{0} & \mathbf{0} & \mathbf{w}_1 & \mathbf{w}_2
\end{array}
\right]\\
&=&
\left[
\begin{array}{c}
\mathbf{x}_1 \\
\mathbf{x}_2 \\
\vdots\\
\mathbf{x}_N
\end{array}
\right]^T
\left[
\begin{array}{cccc}
F_1^+\mathbf{\Sigma}_1^{-1}\mathbf{w}_1 & F_1^+\mathbf{\Sigma}_1^{-1}\mathbf{w}_2 & F_1^\times\mathbf{\Sigma}_1^{-1}\mathbf{w}_1 & F_1^\times\mathbf{\Sigma}_1^{-1}\mathbf{w}_2
\\
F_2^+\mathbf{\Sigma}_2^{-1}\mathbf{w}_1 & F_2^+\mathbf{\Sigma}_2^{-1}\mathbf{w}_2 & F_2^\times\mathbf{\Sigma}_2^{-1}\mathbf{w}_1 & F_2^\times\mathbf{\Sigma}_2^{-1}\mathbf{w}_2
\\
\vdots & \vdots & \vdots & \vdots
\\
F_N^+\mathbf{\Sigma}_N^{-1}\mathbf{w}_1 & F_N^+\mathbf{\Sigma}_N^{-1}\mathbf{w}_2 & F_N^\times\mathbf{\Sigma}_N^{-1}\mathbf{w}_1 & F_N^\times\mathbf{\Sigma}_N^{-1}\mathbf{w}_2
\end{array}
\right]\\
&=&
\left[
\begin{array}{c}
\sum_{i=1}^N F_i^+\mathbf{x}_i^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 \\
\sum_{i=1}^N F_i^+\mathbf{x}_i^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 \\
\sum_{i=1}^N F_i^\times\mathbf{x}_i^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 \\
\sum_{i=1}^N F_i^\times\mathbf{x}_i^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2
\end{array}
\right]^T,
\end{eqnarray}
a 4-vector.

Note that
\begin{eqnarray}
\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_i\mathbf{w}_j
\end{eqnarray}
is a scalar value, and is the (unnormalized) matched filter of the data from a detector with one of the two waveforms.  There are $2N$ such values, and they are expensive to compute (a Q-transform computes 2 at a time: the real and imaginary components of the complex amplitude).  However, the quantity does not depend on $\mathbf{F}$, so the values are not functions of direction.

The middle  of the kernel,
\begin{eqnarray}
(
(\mathbf{F}\mathbf{W})^T
\mathbf{\Sigma}^{-1}
\mathbf{F}\mathbf{W}
+
\mathbf{A}^{-1}
)^{-1}
\end{eqnarray}
is a function of direction, but it is only a 4-by-4 matrix.  It can be computed once and used for many arrival times (so long as the noise is stationary).

We can therefore cheaply compute the data-dependent portion of the Bayes factor by forming the matched filters for each detector individually, and for each direction on the sky performing the small number of operations required to combine the matched filter values.  If the detectors are not co-located, the matched filters will have to be computed at direction dependent time shifts, \emph{but} \texttt{wtransform} already produces an oversampled time series of filter outputs from which which we can cheaply look-up and/or interpolate the required values, so long as we are searching over arrival time as well as direction.

The size of the middle of the kernel is determined by the number of parameters in the signal model.  For white noise bursts, the number of basis waveforms and parameters is large and the implementation above will be slower than a more straightforward one.

The derivation above applies to any basis.  The Fourier basis is obviously a convenient one; it diagonalizes $\mathbf{\Sigma}^{-1}$ (for ``circular'' stationary noise).  The matched filters can be computed by doubly-whitening the data before applying a modified \texttt{wtransform} that returns the complex amplitude rather than the normalized energy.

The implementation performs only as many Q-transforms as the corresponding incoherent search.  The cost of marginalizing over arrival time, direction, frequency, Q (and possibly amplitude) will be significant, but should be much faster than a signal and null stream based search that has to perform as many Q transforms for \emph{every} direction considered.

We can also accelerate the computation of the middle part of the kernel
\begin{eqnarray}
\mathbf{FW}&=&
\left[
\begin{array}{cc}
F_1^+\mathbf{I} & F_1^\times\mathbf{I} \\
F_2^+\mathbf{I} & F_2^\times\mathbf{I} \\
\vdots & \vdots \\
F_N^+\mathbf{I} & F_N^\times\mathbf{I} \\
\end{array}
\right]\left[
\begin{array}{cccc}
\mathbf{w}_1 & \mathbf{w}_2 & \mathbf{0} & \mathbf{0} \\
\mathbf{0} & \mathbf{0} & \mathbf{w}_1 & \mathbf{w}_2
\end{array}
\right]\\
&=&
\left[
\begin{array}{cccc}
F_1^+\mathbf{w}_1 & F_1^+\mathbf{w}_2 & F_1^\times\mathbf{w}_1 & F_1^\times\mathbf{w}_2 \\
F_2^+\mathbf{w}_1 & F_2^+\mathbf{w}_2 & F_2^\times\mathbf{w}_1 & F_2^\times\mathbf{w}_2 \\
\vdots & \vdots & \vdots & \vdots \\
F_N^+\mathbf{w}_1 & F_N^+\mathbf{w}_2 & F_N^\times\mathbf{w}_1 & F_N^\times\mathbf{w}_2
\end{array}
\right]\\
\mathbf{(FW)}^T\mathbf{\Sigma}^{-1}\mathbf{FW}&=&
\left[
\begin{array}{cccc}
F_1^+\mathbf{w}_1^T & F_2^+\mathbf{w}^T_1 & \cdots & F_N^+\mathbf{w}^T_1 \\
F_1^+\mathbf{w}_2^T & F_2^+\mathbf{w}^T_2 & \cdots & F_N^+\mathbf{w}^T_2 \\
F_1^\times\mathbf{w}^T_1 & F_2^\times\mathbf{w}^T_1 & \cdots & F_N^\times\mathbf{w}^T_1 \\
F_1^\times\mathbf{w}^T_2 & F_2^\times\mathbf{w}^T_2 & \cdots & F_N^\times\mathbf{w}^T_2
\end{array}
\right]\\
&&\cdot
\left[
\begin{array}{cccc}
F_1^+\mathbf{\Sigma}_1^{-1}\mathbf{w}_1 & F_1^+\mathbf{\Sigma}_1^{-1}\mathbf{w}_2 & F_1^\times\mathbf{\Sigma}_1^{-1}\mathbf{w}_1 & F_1^\times\mathbf{\Sigma}_1^{-1}\mathbf{w}_2
\\
F_2^+\mathbf{\Sigma}_2^{-1}\mathbf{w}_1 & F_2^+\mathbf{\Sigma}_2^{-1}\mathbf{w}_2 & F_2^\times\mathbf{\Sigma}_2^{-1}\mathbf{w}_1 & F_2^\times\mathbf{\Sigma}_2^{-1}\mathbf{w}_2
\\
\vdots & \vdots & \vdots & \vdots
\\
F_N^+\mathbf{\Sigma}_N^{-1}\mathbf{w}_1 & F_N^+\mathbf{\Sigma}_N^{-1}\mathbf{w}_2 & F_N^\times\mathbf{\Sigma}_N^{-1}\mathbf{w}_1 & F_N^\times\mathbf{\Sigma}_N^{-1}\mathbf{w}_2
\end{array}
\right]\\
&=&
\end{eqnarray}
\begin{eqnarray}
\left[
\begin{array}{cccc}
\sum_{i=1}^N F_i^+F_i^+\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^+F_i^+\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 &
\sum_{i=1}^N F_i^+F_i^\times\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^+F_i^\times\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 \\
\sum_{i=1}^N F_i^+F_i^+\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^+F_i^+\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 &
\sum_{i=1}^N F_i^+F_i^\times\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^+F_i^\times\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 \\
\sum_{i=1}^N F_i^\times F_i^+\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^\times F_i^+\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 &
\sum_{i=1}^N F_i^\times F_i^\times\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^\times F_i^\times\mathbf{w}_1^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 \\
\sum_{i=1}^N F_i^\times F_i^+\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^\times F_i^+\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2 &
\sum_{i=1}^N F_i^\times F_i^\times\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_1 &
\sum_{i=1}^N F_i^\times F_i^\times\mathbf{w}_2^T\mathbf{\Sigma}_i^{-1}\mathbf{w}_2
\end{array}
\right]
\end{eqnarray}
so we can rapidly compute the middle of the kernel from the $4N$ direction-independent scalars \begin{eqnarray}
\mathbf{w}_i^T\mathbf{\Sigma}_j^{-1}\mathbf{w}_k
\end{eqnarray}
which seem to be the normalization terms for the matched filter.  (Symmetry reduces it to $3N$, with orthogonal waveforms reducing to $2N$)

In fact, the definition of the matched filter is
\begin{eqnarray}
\frac{\mathbf{x}^T\mathbf{\Sigma}^{-1}\mathbf{w}}{\sqrt{\mathbf{w}^T\mathbf{\Sigma}^{-1}\mathbf{w}}}
\end{eqnarray}
and we reduce to (the square of) it for one detector, one waveform, $\mathbf{F} = \mathbf{I}$, and $\sigma=\infty$.
\begin{eqnarray}
\mathbf{x}^T\mathbf{\Sigma}^{-1}\mathbf{w}(\mathbf{w}^T\mathbf{\Sigma}^{-1}\mathbf{w})^{-1}\mathbf{w}^T\mathbf{\Sigma}^{-1}\mathbf{x}
\end{eqnarray}

To compute the determinant,
\begin{eqnarray}
|\mathbf{I}-\mathbf{\Sigma K}|&=&|\mathbf{I}-\mathbf{\Sigma}\mathbf{\Sigma}^{-1}\mathbf{FW}((\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}+\mathbf{A}^{-1})^{-1}(\mathbf{\Sigma}^{-1}\mathbf{FW})^T|\\
&=&|\mathbf{I}-\mathbf{FW}((\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}+\mathbf{A}^{-1})^{-1}(\mathbf{\Sigma}^{-1}\mathbf{FW})^T|\\
&=&|\mathbf{I}-(\mathbf{\Sigma}^{-1}\mathbf{FW})^T\mathbf{FW}((\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}+\mathbf{A}^{-1})^{-1}|\\
&=&|\mathbf{I}-(\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}((\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}+\mathbf{A}^{-1})^{-1}|
\end{eqnarray}
(by Sylvester's theorem).  The matrices $(\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}$ and $((\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}+\mathbf{A}^{-1})^{-1}$ are only 4-by-4 and already need to be computed for the kernel.

Straightforward implementation was found to run 14,000 slower than real time.  Despite this, the operation is quite efficient: it computes the Bayes factor for $10^6$ hypotheses per second.  The problem is that we explicitly consider too many hypotheses.

For small $N$ we may want to recast these even further to get the kernel, an $N$-by-$N$ matrix, noting that we have already found
\begin{eqnarray}
\mathbf{\Sigma}^{-1}\mathbf{FW}
\end{eqnarray}
Since the expression involves an inverse, there's no point expanding it here or in the code, except to note that explicitly computing $\mathbf{K}$ is possible and desirable in some cases.

\section{Integration}

The straightforward implementation of the algorithm above is still prohibitively expensive because of the enormity of the parameter space.  The solution is to use a smarter integration algorithm.

One way to achieve this is to perform \emph{importance sampling}.  If we can find a cheap function that approximates the posterior, we can Monte-Carlo integrate it with much fewer, more intelligently distributed samples.  There is of course no unique right way of doing this.

One approach is to try a separable importance function.  This means choosing parameters that the posterior is separable in.  The existence of ring structures on the sky maps imply that $(\tau_1, \tau_2, \tau_3)$ will be a more separable basis than $(\tau, \theta, \phi)$ (we will concentrate on the $N=3$ case for now).
\begin{eqnarray}
p(\mathbf{x}|\tau_1,\tau_2,\tau_3,\rho,H_\textrm{signal}) \approx \prod_{i=1}^3 g(\mathbf{x},\tau_i)
\end{eqnarray}

The Bayes factor can be expressed as
\begin{eqnarray}
f(\tau_2-\tau_1,\tau_3-\tau_2)
\sqrt{|\mathbf{I}-\mathbf{\Sigma K}|}\exp(\sum_{ijkl}C_{ijkl}(\tau_2-\tau_1,\tau_3-\tau_1)(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_i\mathbf{w}_j(\tau_i))(\mathbf{x}_k^T\mathbf{\Sigma}^{-1}_k\mathbf{w}_l(\tau_k)))
\end{eqnarray}
suggesting an separable importance function of the form
\begin{eqnarray}
\prod_{i=1}^3\exp(\sum_{jk} b_{ijk} (\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_i\mathbf{w}_j(\tau_i))(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_i\mathbf{w}_k(\tau_i)))
\end{eqnarray}
where we could obtain $b_{ijk}$ in some tedious fashion.  However, we can also fairly easily obtain a well-motivated expression very like the above.

Consider
\begin{eqnarray}
\frac{p(\mathbf{x}_i|\tau,\rho,H_\textrm{signal},I)}{p(\mathbf{x}_i|H_\textrm{noise},I)}&=&
\int_{-\pi}^\pi\int_0^\pi
p(\theta,\phi|H_\textrm{signal},I)
\sqrt{|\mathbf{I}-\mathbf{\Sigma}_i\mathbf{K}|}
\exp(\frac{1}{2}\mathbf{x}_i^T\mathbf{K}\mathbf{x}_i)
\mathrm{d}\theta\mathrm{d}\phi\nonumber
\end{eqnarray}
where we adopt the convention that $\tau$ is measured at the detector.  We can also adopt a polarization coordinate system such that $F_i^\times=0$, and the simplifying assumptions that
\begin{eqnarray}
\mathbf{w}_1\mathbf{\Sigma}^{-1}_i\mathbf{w}_1&=&\mathbf{w}_2\mathbf{\Sigma}^{-1}_i\mathbf{w}_2\\
\mathbf{w}_1\mathbf{\Sigma}^{-1}_i\mathbf{w}_2&=&0
\end{eqnarray}
We then get
\begin{eqnarray}
\mathbf{x}^T_i{\Sigma}^{-1}_i\mathbf{FW}&=&F_i^+\left[
\begin{array}{c}
\mathbf{x}^T_i{\Sigma}^{-1}_i\mathbf{w}_1 \\
\mathbf{x}^T_i{\Sigma}^{-1}_i\mathbf{w}_2 \\
0 \\
0
\end{array}
\right]
\end{eqnarray}
and
\begin{eqnarray}
(\mathbf{FW})^T\mathbf{\Sigma}^{-1}\mathbf{FW}&=&
(F_i^+)^2\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1
\left[
\begin{array}{cccc}
1 & 0 & 0 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{array}
\right]
\end{eqnarray}
and
\begin{eqnarray}
((\mathbf{FW})^T\mathbf{\Sigma}_1^{-1}\mathbf{FW}+\mathbf{A}^{-1})^{-1}&=&
\left[
\begin{array}{cccc}
\frac{1}{(F_i^+)^2\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1 + \sigma^{-2}} & 0 & 0 & 0 \\
0 & \frac{1}{(F_i^+)^2\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1 + \sigma^{-2}} & 0 & 0 \\
0 & 0 & \sigma^2 & 0 \\
0 & 0 & 0 & \sigma^2 \\
\end{array}
\right]
\end{eqnarray}
and
\begin{eqnarray}
\mathbf{x}_1^T\mathbf{K}\mathbf{x}&=&\frac{(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1)^2+(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_2)^2}{\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1 + (F_i^+)^{-2}\sigma^{-2}}
\end{eqnarray}
and
\begin{eqnarray}
\sqrt{|\mathbf{I}-\mathbf{\Sigma}_1\mathbf{K}|}&=&
\frac{1}{\sigma^2(F_i^+)^2\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1+1}
\end{eqnarray}
%\frac{1}{\sigma^2(F_i^+)^2\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1}

so we want
\begin{eqnarray}
\int_{0}^1
p(F_i^+)
\frac{1}{\sigma^2(F_i^+)^2\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1+1}
\exp(\frac{1}{2}\frac{(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1)^2+(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_2)^2}{\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1+(F_i^+)^{-2}\sigma^{-2}})
\mathrm{d}F_i^+
\end{eqnarray}
In the large $\mathbf{x}$ and $\sigma$ regime, the integral is dominated by the contribution from the largest $F_i^+$; in other regimes it isn't strongly affected by $F_i^+$? (Check)  So to get a crude approximation we can just set $F_i^+ = 1$; the result is the Bayesian matched filter for a signal of size $\sigma$.
\begin{eqnarray}
&\approx&
\frac{1}{\sigma^2\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1+1}
\exp(\frac{1}{2}\frac{(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1)^2+(\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_2)^2}{\mathbf{w}_1^T\mathbf{\Sigma}^{-1}_1\mathbf{w}_1+\sigma^{-2}})
\end{eqnarray}
Why not just use the matched filter expression and dispense with $\sigma$ too? Because it allows us to weight the interferometers properly.  If one of the interferometers is much less sensitive than the others, the use of $\sigma$ flattens out the importance function; we will essentially ignore that interferometer's contribution.

The straightforward Monte-Carlo integration strategy is now: draw $\tau_i$ from the distributions defined by normalizing the above function for each $i$.  Sample the full Bayes factor function for these parameters (marginalizing over multiple directions that are consistent with a particular time, such as mirror directions for three detectors or rings for two detectors), and weight them by the inverse of the product of the sampling functions.  The average of many such samplings is the integral.  If the importance function is well chosen, it will converge much more quickly than a simple integration.

It is important to remember that, though the data contains noise, the integral we wish to perform is for a particular noise realization and is entirely deterministic.  There is no notion of losing degrees of freedom or introducing correlations when we use functions of the data to recast the integral like this.  However, the fact we are dealing with noise realizations will put a limit on the useful accuracy to compute the integral to: we want the integrator to contribute less uncertainty to the result than the uncertainty due to the instrumental noise.

The vast majority of $\tau$ triplets are invalid (no sky positions correspond to them) with the integral zero there.  This isn't a problem (the integral is zero there), but we must make a special effort to sample the the valid region.
The valid region is cylindrical, enclosing the $\tau_1=\tau_2=\tau_3$ diagonal, but only a very small part of parameter space.  (the observation is tens is seconds long, but the samples must be within tens of milliseconds of each other, so roughly one in a million samples will be valid if uniformly distributed). To sample the separable distribution only within this cylinder is not trivial.  It is not practical to discard samples outside, since the interior is a tiny fraction of the whole volume.

Instead, we must make new distributions, with marginalization:
\begin{eqnarray}
p(\tau_1)&=&g(\tau_1)\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty} g(\tau_2) g(\tau_3) h(\tau_1,\tau_2,\tau_3) d\tau_2 d\tau_3\\
p(\tau_1)&=&g(\tau_1)\int_{\tau_1-\delta_{12}}^{\tau_1+\delta_{12}}g(\tau_2)\int_{\tau_1-\delta_{13}}^{\tau_1+\delta_{13}}g(\tau_3)h(\tau_1,\tau_2,\tau_3)d\tau_3d\tau_2
\end{eqnarray}
If we throw away the detail of the cylinder described by $h$,
\begin{eqnarray}
p(\tau_1)&=&g(\tau_1)\int_{\tau_1-\delta_{12}}^{\tau_1+\delta_{12}}g(\tau_2)d\tau_2\int_{\tau_1-\delta_{13}}^{\tau_1+\delta_{13}}g(\tau_3)d\tau_3
\end{eqnarray}
If we sample $\tau_1$ from this distribution, and $\tau_2, \tau_3$ from their distributions constrained around $\tau_1$, the result will be identically distributed inside the valid region to the original distribution.  Some samples (about half?) will still lie outside the valid region, but this is more acceptable.  The use of cumulative sums makes it quite efficient to compute all the above.

Once we have a set of $\tau_i$ that may be valid, we need to compute the posterior for each.  The first step is to consider the set of source directions compatible with the time delays.  For three sites, there are zero or two directions consistent with each time delay.  (For two sites, which we aren't considering yet, a whole ring on the sky is consistent.)  To recover the source direction, we need to find the solutions to
\begin{eqnarray}
\left[\begin{array}{c}\tau_2-\tau_1\\\tau_3-\tau_1\end{array}\right]
&=&
\left[\begin{array}{c}\mathbf{c}^T_1-\mathbf{c}^T_2\\\mathbf{c}^T_1-\mathbf{c}^T_3\end{array}\right]
\mathbf{n}
\end{eqnarray}
where $\mathbf{c}_i$ are the coordinates of the beamsplitters in Earth-Cartesian lightseconds and the direction of the source is normalized $|\mathbf{n}|=1$.  It is presumably cheaper to compute these in advance for all representable sky directions, except when the number of samples is very small.  Via the SVD of the matrix, we can compute the portion of $\mathbf{n}$ in the row space for every relative-time delay coordinate and store the results in a cell array indexed by delay times.  For each cell, we check if the norm of partial $\mathbf{n}$ is less than one, and then compute the two directions by adding an orthogonal component to achieve the normalization.  \texttt{wresponse} is then used to get the antenna patterns.  The kernel matrix and normalization are then computed and stored in the cell array.

We also need to compute the sky area factor otherwise we end up biased against sources in the plane of the network.  The appropriate weighting is simply $z^{-1}$, i.e. the inverse of the coefficient for the out-of-plane direction, unless we want to get really fancy and deal with the exact tile shapes.

The time resolution of the output of the Q-transform enforce a 95\% overlap in their envelopes, not in the enclosed oscillations.  We have to regenerate them at higher resolutions to provide decent sky coverage.

This makes it even more important to have an efficient sampler.  Unfortunately, the random sampling and lookup tables make this difficult to implement efficiently in \textsc{Matlab}, as it can't be vectorized.  We have implemented this part of the code in \emph{C}, as \texttt{randh}.  This has eliminated a substantial contributor to the cost of executing the function.  The remaining code is dominated by the costs of inner-loop, unvectorizable accesses to the lookup table. These cannot be eliminated without reimplementing the bulk of the algorithm in \emph{C}.

\subsection{Refinements and special cases}

The system is complicated and difficult to debug.  Rearranging it to make it more modular and with better-defined intermediate products will be important to making it maintainable.

The use of the individual energy functions is strongly motivated, but might not be the most robust.  It does not take much energy for the integrator to disregard everything but the peak.  Could a weaker importance function have benefits?  This can be phrased in terms of us not wanting to reduce variance, but something else.  Given the strongly asymmetric distributions involved, are we actually in the central limit theorem regime?  Should we be working with log-Gaussians or something?

It might be fairly cheap to sample only valid directions, again using the \texttt{cumsum} tables.

To generate sky maps (for whatever reason; will be good for debugging) we'll need to switch to a different and in many ways easier integrator, that importance samples on only the one parameter (t1) using the product of the importances.  This integrator seems very nice; could this be a hint that there is a better decomposition than the current one?  It might even be more amenable to variable numbers of detectors; the downside will be the inability to disregard directions.

\section{Implementation notes}
The importance function seems to be off by a square/square-root.  This has been corrected, but it is unknown if this is due to a bug in the code, a mistake in the derivation, or a mistake in reasoning.

The standard Monte-Carlo error estimates seem to be far too rosy; we need better error estimates.  The distribution of estimates of the mean doesn't seem to be remotely Gaussian until approximately 1000 samples, and is much larger than the $\sigma/\sqrt{N}$ estimate would suggest.  Can we derive an alternative error estimate, based (for example) on the posterior for the sum of  lognormal distributed variables.

\section{Glitch model}
The weight of evidence for a glitch in any given detector is the same as the importance function:

\begin{eqnarray}
\ln \frac{p(\mathbf{x}_i|\mathbf{w},\sigma,\mathbf{\Sigma}_i)}
{p(\mathbf{x}_i|\mathbf{\Sigma}_i)}
&=&
-
\ln(\sigma^2\mathbf{w}_1\mathbf{\Sigma}_1^{-1}\mathbf{w}_1+1)
+
\frac{1}{2}
(
(x_i\Sigma_i w_1)^2 + (x_i\Sigma_i w_2)^2
)
/(w_1\Sigma_iw_1+\sigma^{-2})
\end{eqnarray}

Now,

\begin{eqnarray}
&&\ln\frac{p(x|\sigma,\Sigma)}{p(x|\Sigma)}\\
&=&
\ln
\frac{
\prod_{i=1}^3(\frac{1}{2}p(x_i|\sigma,\Sigma_i)+\frac{1}{2}p(x_i|\Sigma_i))}
{\prod_{i=1}^3 p(x_i|\Sigma_i)}
\\
&=&
\ln\frac{1}{8}\prod_{i=1}^3(\frac{p(x_i|\sigma,\Sigma_i)}{p(x_i|\Sigma_i)}+1)
\\
&=&-\ln{8}+\sum_{i=1}^3\ln(\frac{p(x_i|\sigma,\Sigma_i)}{p(x_i|\Sigma_i)}+1)
\end{eqnarray}

\end{document}

However, we want to make a slightly different implementation to cope with the fact that we only have the matched filter data for finite times, and we only want to consider those finite times to draw from.

It is cheap to compute the independent matched filters
\begin{eqnarray}
\mathbf{x}_i^T\mathbf{\Sigma}^{-1}_i\mathbf{w}_j(\tau_i)
\end{eqnarray}

We have considerable information in the expressions
\begin{eqnarray}
\mathbf{x}\mathbf{\Sigma}_i^{-1}\mathbf{w}_j(\tau_i)
\end{eqnarray}
We expect that
\begin{eqnarray}
\frac{p(\mathbf{x}|\rho,\tau,\theta,\phi)}{\prod_{i=1}^N p(\mathbf{x}_1|\rho,\tau,\theta,\phi)}
\end{eqnarray}
will be a much flatter function than the Bayes factor.  If we change variables to the times of arrival $\tau_i$ we can use the product of the individual matched filter results for each detector as an importance sampling function and Monte-Carlo integrate the result.  This means we are preferentially sampling at times when there is energy in the detector.  A coincidence analysis with coherent followup is a approximation to this strategy: identify energetic times, then compute the coherent statistic for them.

&=&|\mathbf{I}-\mathbf{UDU}^T(\mathbf{UDU}^T+\mathbf{A}^{-1})^{-1}|
&=&|\mathbf{I}-(\mathbf{UDU}^T(\mathbf{UDU}^T)^{-1}+\mathbf{A}^{-1}(\mathbf{UDU}^T)^{-1})^{-1}|\\
&=&|\mathbf{I}-(\mathbf{I}+\mathbf{A}^{-1}(\mathbf{UDU}^T)^{-1})^{-1}|

### 2010-03-30 egsm.tex
an edit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@398 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Ryzhik, FI 2 779.} using the GAMMA function we have : $$ \langle \rho \rangle= \frac{2}{n^{n}}\left[\Gamma(\frac{1}{n})\right]^{n}$$
-This result makes sense : we can calculate the exact limit as n $\rightarrow \infty$ : $\lim\limits_{n\to \infty}\langle \rho \rangle=\rho$ as illustrated on the figure below, which represent the expected value of the geometric sample mean versus n :
-
-\vspace*{0.2cm}
-\includegraphics[scale=0.19]{geosamplemean.png}
+This result makes sense : we can calculate the exact limit as n $\rightarrow \infty$ : $\lim\limits_{n\to \infty}\langle \rho \rangle=\rho$ as illustrated on the figure below, which represent the expected value of the geometric sample mean versus n:
+\resizebox{\linewidth}{!}{\includegraphics{geosamplemean.png}}
 

### 2011-02-04 analysis.tex
Huge set of changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@598 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{SECIII}\label{sec:analysis}
+
+\editorial{Should I mention \texttt{GStreamer} at all?} The detection method described above is much more efficient in terms of floating point operations than the traditional matched filter bank method.
+However, time slices and conditional reconstruction greatly complicate queueing, synchronizing, and bookkeeping of intermediate signals.
+A low latency implementation capable of recruiting more than one \textsc{cpu} core would be difficult to achieve within the familiar serial programming framework because of the nontrivial time-delay relationships between samples.
+Due to these complications, we chose to prototype the search using an open source signal processing environment called \texttt{GStreamer}.
+\editorial{Citation for \texttt{GStreamer}?}  Primarily used for playing, authoring, or streaming media on \texttt{Linux} systems, \texttt{GStreamer} is an integral component of the popular \texttt{Gnome} desktop.
+
+\editorial{It would be good to illustrate the layout of this particular template bank: masses spanned, time slice layout \dots}
+\begin{table}[!h]
+\caption{Operation counts per sample for six different detection methods.
+The operation counts for \textsc{lloid} assume a reconstruction duty cycle of 5\%.
+Note that the \textsc{fir} method with \textsc{lloid} is almost 10 times faster than the conventional \textsc{fft} method, despite having substantially lower latency.}
+\begin{center}
+\input{flop_budget}
+\end{center}
+\end{table}
+
+\begin{comment}
+
+% Ugly diagram!
 
@@
 \end{figure*}
+
+\end{comment}

### 2011-02-04 appendix.tex
Huge set of changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@598 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
+\section{Floating point operation counts}
+
+\editorial{I just yanked this from the methods section.
+It needs to be made more concise, maybe turned into a table of CPU budgets for various components of LLOID.} The filter bank can be implemented using finite impulse response (\textsc{fir}) filters, which are just sliding window dot products.
+If there are $M$ templates of length $n$, and the data stream contains $N$ samples, then applying the filter bank requires $2 M N n$ operations.
+
+More commonly, the matched filters are implemented using the \textsc{fft} convolution.
+This entails applying \textsc{fft}s to blocks of $D$ samples, with $2 n \leq D$, each block overlapping the previous one by $n$ samples.
+There are $N/(D-n)$ such blocks.
+Modern implementations of the Cooley-Tukey \textsc{fft}, such as the ubiquitous \texttt{fftw}, require about $4 N \lg N$ operations to evaluate a \textsc{dft} of size $N$~\cite{Johnson:2007p9654}.
+\editorial{This is more commonly known as ``overlap-save''.
+We should find someone else's operation count and cite it.}  A $D$ sample cross-correlation consists of a forward \textsc{fft}, an $D$ sample dot product, and an inverse \textsc{fft}, totaling $8 D \lg N + 2 D$ operations per block.
+Per sample, this is $(8 \lg D + 2) / (1 - n/D)$ operations.
+
+The \textsc{fir} filter implementation has the advantage that it has no intrinsic latency, whereas the \textsc{fft} convolution has at least the latency of the \textsc{fft} block size $D \geq 2 n$.
+For example, for a $1.4 - 1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the \textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.
+However, the \textsc{fir} filter implementation has the disadvantage of much greater overhead per sample than the \textsc{fft} convolution.
+For a $1\,\mathrm{ks}$ template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation requires about about $n / 8 \lg 2 n = 2.2 \times 10^4$ times more operations per sample than the \textsc{fft} implementation.
+
+\begin{comment}
+% Let's make glitch rejection in LLOID a separate paper.
+
 \section{better noise model}
@@
 These factors make it possible for the gravitational wave hypothesis to be preferred for some data.
+
+\end{comment}
 

### 2011-02-04 inspiral_svd.tex
Huge set of changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@598 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\documentclass[aps,prd,twocolumn,showpacs,groupedaddress,showkeys,preprintnumbers]{revtex4}
+\documentclass[showpacs,groupedaddress,showkeys,preprintnumbers]{iopart}
 \input{packages}
@@
 
-% $Id:$
-
 \begin{document}
 
-\title{Low-latency search for gravitational-waves: prospects for near real time detection of compact binary coalescence}
+\title[LLOID]{A novel method for detecting coalescing binaries in near realtime with Advanced LIGO and beyond \editorial{Submit to PRD or CQG?}}
 
@@
 
-\author{Kipp Cannon} 
-% \email{antony.searle@anu.edu.au}
-\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}
+\author{Kipp Cannon$^1$, Adrian Chapman$^2$, Chad Hanna$^3$, Drew Keppel$^4$, Antony C. Searle$^2$, Alan J. Weinstein$^2$}
 
-\author{Adrian Chapman} 
-% \email{antony.searle@anu.edu.au}
-\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}
-
-\author{Chad Hanna}  
-% \email{antony.searle@anu.edu.au}
-\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}
-
-\author{Drew Keppel}  
-% \email{antony.searle@anu.edu.au}
-\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}
-
-\author{Antony C. Searle}
-% \email{antony.searle@anu.edu.au}
-\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}
-
-\author{Alan J. Weinstein}
-% \email{antony.searle@anu.edu.au}
-\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}
+\address{$^1$ Canadian Institute for Theoretical Astrophysics, ON, Canada}
+\address{$^2$ LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA}
+\address{$^3$ Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
+\address{$^3$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
 
 \begin{abstract}
-
+Conventional matched filter bank methods for the detection of gravitational waves from the inspiral of compact binaries are computationally expensive, have hundreds of seconds of unavoidable intrinsic latency, and require arrays of largely redundant matched filters.
+Novel detection methods that are more computationally efficient and have lower latency will be required to realize the full potential of advanced of gravitational wave detectors that are currently under construction.
+In this paper, we describe a new detection method that exploits the properties of inspiral waveforms using multi-rate filtering, principal component analysis, and hierarchical detection.
+We provide receiver operating characteristics from a prototype search pipeline that is capable of low-latency or near-realtime detection with greatly reduced computational requirements in comparison with previously described methods.
 \end{abstract}
 
-\pacs{}
+\pacs{95.55.Ym, 84.30.Vn, 95.75.Wx}
 
-\keywords{Gravitational Waves, Laser Interferometry}
+%\keywords{Gravitational Waves, Laser Interferometry}
 
-\preprint{}
-
-\maketitle
-
+%\preprint{}
 \input{introduction}
@@
 
-\begin{acknowledgments}
-LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the National Science Foundation and operates under cooperative agreement PHY-0107417.
-This paper has LIGO Document Number LIGO-P0900004-v1
-
-\end{acknowledgments}
+\ack{\textsc{ligo} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the National Science Foundation and operates under cooperative agreement \textsc{phy}-\oldstylenums{0107417}.
+This paper has \textsc{ligo} Document Number \textsc{ligo}-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{1}.
 
 \appendix
+\input{appendix}
 
+\section*{References}
+
+\bibliographystyle{unsrt}
 \bibliography{references}

### 2011-02-04 introduction.tex
Huge set of changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@598 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{sec:introduction}
-Modern gravitational-wave observatories are likely to deliver on the promise 
+
+% I would like to say:
+%  - Low mass inspirals are the most promising source of gravitational radiation for advanced LIGO
+% - Mention S6 and VSR2, online running, EM followup
+% - Observational applications of near-realtime detection
+% - Point out challenges of advanced detectors: more templates, longer templates
+% - Advanced LIGO timetable
+% - Einstein telescope
+% - Layout of paper: description of method, procedure followed
+
+\begin{comment}
+
+However, the computational requirements of conventional detection strategies --- especially those that are well suited to low latency applications --- will be greatly increased by this upgrade for two reasons.
+First, advanced detectors will have much improved low-frequency sensitivity, so chirp signals from coalescing binaries will remain in the detectors' band for longer.
+This will demand that matched filter based searches use longer templates, requiring more memory and more cycles to process.
+Second, background noise for advanced detectors will be more spectrally uniform, so it will be possible to resolve the intrinsic parameters of a source that determine the time-frequency structure of the gravitational wave signal with much greater accuracy.
+This will require that matched filter bank based searches employ more matched filters, again demanding more memory and more cycles.
+
+\end{comment}
+
+\begin{comment}
+
+\editorial{This paragraph has too many citations.} Modern gravitational-wave observatories are likely to deliver on the promise 
 that their name implies by observing multiple gravitational-wave signals from
-compact binary coalescence within the coming decade ~\cite{kopparapu2008}. 
+compact binary coalescence within the coming decade \cite{kopparapu2008}. 
 There is now a world-wide network of gravitational-wave observatories 
 including the laser interferometers
-LIGO, VIRGO, GEO, and TAMA, which should assure that coincident detections 
+\textsc{ligo}, Virgo, \textsc{geo}, and \textsc{tama}, which should assure that coincident detections 
 will be possible among at least three separate sites lending confidence
-to any detections.
-~\cite{Abbott:2005qm, beauville2008, beauville2006, blackburn2005, LIGOS1instpaper}.  
+to any detections~\cite{Abbott:2005qm, beauville2008, beauville2006, blackburn2005, LIGOS1instpaper}.
+\editorial{This is a lot of citations just to say that the 4 detectors exist.
+Maybe if they were scattered throughout the paragraph?}
 However, having an electromagnetic or neutrino counterpart to a 
@@
 
-Work has already commenced to trigger gravitational-wave searches from 
-electromagnetic observations ~\cite{triggeredsearches2008}. LIGO has recently
-ruled out the possibility of GRB070201 originating from the merger of a
-neutron star / neutron star or neutron star / black hole system in the
-Andromeda galaxy ~\cite{GRB070201}.  The still open question is, will 
-gravitational-wave observations be able to trigger electromagnetic followup?
+\end{comment}
+
+In a series of detector upgrades that are now under way, \textsc{ligo} and Virgo will become the first advanced gravitational wave detectors, gaining a tenfold improvement in amplitude sensitivity \citeneeded{} and a corresponding thousandfold increase in observable volume in the local universe.
+It is anticipated that these detectors will be able to sense as many as 40 events per year \cite{Abadie:2010p10836}.
+
+As the gravitational wave detection horizon pushes outward, the ability to detect signals in near realtime will become valuable.  Having an electromagnetic or neutrino counterpart to a 
+gravitational-wave detection would not only increase the confidence in 
+the detection but will also greatly improve the astrophysical information
+available from the event.  Most models that predict simultaneous
+gravitational-wave and electromagnetic observations also predict that the
+peak amplitude of electromagetic radiation will occur soon after 
+gravitational-wave emmission~\cite{sylvestre2003}.  Thus in order to maximize
+the chance of a successful electromagnetic followup the latency of 
+gravitational-wave signal analysis must be made minimal.
+
+\editorial{This is old news, and doesn't seem relevant to me: EM-directed gravitational wave searches don't benefit from low-latency detection pipelines.} Work has already commenced to trigger gravitational-wave searches from
+electromagnetic observations~\cite{triggeredsearches2008}. \textsc{ligo} observations
+ruled out the possibility of GRB70201 originating from the merger of a
+\textsc{nsns} or \textsc{nsbh} system in the Andromeda galaxy~\cite{GRB070201}.
+
+In \oldstylenums{2010}, \textsc{ligo} and Virgo completed a period of joint data taking during which several all-sky detection pipelines operated in a low-latency configuration \citeneeded.
+Promising detection candidates were promptly sent for electromagnetic followup \citeneeded{} to several telescopes including Swift, \textsc{rotse}, \textsc{tarot}, and Zadko.
 
@@
 sky to prompt electromagnetic observation followups~\cite{sylvestre2003}.
-Much work is currently underway in providing the best possible source 
-localization from gravitational-wave detector networks
-~\cite{markowitz2008, raymond2008, cavalier2006}.  Colloborations are
-currently forming to provide infrastructure for the gravitational-wave
+\editorial{Cite an in-preparation paper of Larry's?} Much work is currently underway in providing the best possible source 
+localization from gravitational-wave detector networks~\cite{markowitz2008, raymond2008, cavalier2006}.  Colloborations are currently forming to provide infrastructure for the gravitational-wave
 Astronomers to provide targets of opportunity for electromagnetic 
-Astronomers ~\cite{kanner2008}.  Most models that predict simultaneous
+astronomers~\cite{kanner2008}.  Most models that predict coincident
 gravitational-wave and electromagnetic observations also predict that the
-peak amplitude of electromagetic radiation will occur soon after 
-gravitational-wave emmission ~\cite{sylvestre2003}.  Thus in order to maximize
-the chance of a successful electromagnetic followup the latencey of 
+peak electromagetic fluence will occur soon after 
+gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
+the chance of a successful electromagnetic followup the latency of 
 gravitational-wave signal analysis must be made minimal.
 
-The parameter space of compact binary coalescence signals is large
-~\cite{Owen:1995tm, Owen:1998dk} .  It is a
+The parameter space of compact binary coalescence signals is large~\cite{Owen:1995tm, Owen:1998dk}.  It is a
 computationally burdensome task to analyze these signals with even moderate
-latency ~\cite{Abbott:2007xi}.  
+latency~\cite{Abbott:2007xi}.  
 This work will describe how to exploit degeneracy in the signal
 parameter space to answer more quickly whether or not a gravitational-wave 
-is present, with what confidence, and where it is coming from. Specifically
-we will explore using the singular value decomposition (SVD) to reduce the
+is present. Specifically we will explore using the singular value decomposition (\textsc{svd}) to reduce the
 effective number of filters required to search the data.  We note that others
-have applied the use of SVD to gravitational wave data analysis to analyze
-optimal gravitational-wave burst detection 
-~\cite{bradyraymajumder2004, heng2008} and coherent networks 
-of detectors ~\cite{wen2008}
+have applied the use of \textsc{svd} to gravitational wave data analysis to analyze
+optimal gravitational-wave burst detection~\cite{bradyraymajumder2004, heng2008} and coherent networks 
+of detectors~\cite{wen2008} .
 

### 2011-02-04 method.tex
Huge set of changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@598 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \subsection{The standard approach: matched filtering}
-Compact binary coalescence searches typically employ matched filtering as the
-first step to locating a gravitational wave signal ~\cite{findchirppaper}.  
-A template waveform is convolved with the whitened data, and if 
-properly normalized it produces a signal-to-noise ratio (SNR) that indicates the
-probability of a signal being present ~\cite{wainstein:1962}.  
-Since the parameters of the signal are
-not known apriori, multiple templates must be filtered for a wide range of
-parameters.  Let us assume that the waveforms are simply parameterized by
-component mass and denote the $i$th template waveform of $N$ templates necessary
-to search a given parameter space as $h_i(t) = f(m1,m2,t)$.
-The SNR, $\rho_i$ for that template is
+
+Searches for gravitational waves from compact binary coalescences typically employ matched filter banks \cite{findchirppaper}.
+Potential inspiral signals are continuously parameterized by time, amplitude, phase, and a set of intrinsic source parameters $\theta$, which in this paper we shall take to consist of the two component masses of a binary, $\theta = (m_1, m_2)$.
+Let $h_+(\theta)$ and $h_\times(\theta)$ be, respectively, the `$+$' and `$\times$' polarization gravitational wave signals that would arise from a fiducial face-on binary at some distance.
+Because, for inspiral signals, $h_+$ and $h_\times$ are nearly in quadrature, they are generally combined into a single complex-valued template $h = h_+ + i h_\times$.
+
+The detection procedure for just one set of intrinsic source parameters $\theta$ starts by {\em whitening} the measurement data stream $x$.
+This involves finding a linear filter that renders the detector's noise \textsc{iid} and Gaussian.
+This filter is applied to the measurement, yielding the whitened data stream $x^\mathrm W$.
+The same linear filter is applied to the template $h(\theta)$, yielding the whitened template $h^\mathrm W (\theta)$.
+The matched filter is the normalized cross-correlation of $h^\mathrm W$ and $x^\mathrm W$, \editorial{After introducing the $\star$ notation for cross correlation, must we define it for continuous and discrete time?}
+$$
+\rho(\theta) = \frac{h^\mathrm W (\theta) \star x^\mathrm W}{|h^\mathrm W (\theta)|}.
+$$
+This is called the signal to noise ratio, or \textsc{snr}.
+The detection statistic is the modulus of this, $|\rho(\theta)|$, which has a $\chi^2$ distribution with 2 degrees of freedom in the absence of signal.%, and it indicates the probability of a signal being present \cite{wainstein:1962}\editorial{Comment about probability doesn't seem relevant.}.
+
+To construct a filter bank, matched filters are realized for discrete signal parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$, such that any possible signal will have a maximum cross-correlation of at least 0.97 with at least one template.
+\editorial{Citation needed for template placement procedure?}  Such a template bank is said to have a 97\% {\em minimum match}.
+This technique is designed so that an inspiral signal can be detected without any prior knowledge of its intrinsic parameters: at most 3\% of the \textsc{snr} is lost by a signal's parameters not exactly coinciding with a template's.
+A trigger is reported for the template parameters $\theta_i$ and time $t$ for which $|\rho|$ is a maximum over some moving interval in $\theta$ and $t$.
+
+\subsubsection{Latency and overhead}
+
+The matched filter bank can be implemented with finite impulse response (\textsc{fir}) filters.
+\textsc{fir} filters are perfectly suited for realtime detection, because they do not introduce any latency at all.
+However, \textsc{fir} filters are very expensive: for $N$ templates of $M$ samples each, a \textsc{fir} matched filter bank costs $\mathcal O(M N)$ operations per sample.
+
+Much more commonly, the matched filter bank is implemented using \textsc{fft} convolutions, costing only $\mathcal O(M \lg N)$ per sample but having a latency of at least $M$ samples.
+For example, for a bank of $1\,\mathrm{ks}$ templates sampled at $4096\,\mathrm{Hz}$, the \textsc{fft} implementation requires about about $2.2 \times 10^4$ times fewer floating point operations per sample than the \textsc{fir} implementation.
+However, the \textsc{fft} implementation has a latency of at least $1\,\mathrm{ks}$.
+
+This presents a dilemma: it seems that low latency detection using the \textsc{fir} implementation is prohibitively expensive, whereas computational cheap detection with the \textsc{fft} method comes with minutes to hours of latency.
+
+In the rest of the section, we will describe a detection strategy that makes use of some very general properties of inspiral template banks in order to evaluate a matched filter bank with far lower computational cost than the \textsc{fir} method and far less latency than the \textsc{fft} method.
+
+\subsection{Selectively reducing the sample rate of the data and template waveforms}
+
+Our first innovation is based upon the common knowledge that gravitational waveforms from the inspiral stage of compact binary coalescences are chirps: they are quasi-monochromatic and slowly evolving in frequency.
+
+\editorial{The time slice boundary notation is a little awkward and different identifiers are used in a few places in the source code.
+It would be nice to clean that up.}  It is possible to strike a balance between the low latency of the \textsc{fir} filter method and the low overhead of the \textsc{fft} method by breaking the templates into disjoint time intervals, or {\em time slices}.
+The outputs of the filters for all of the time slices are delayed by the appropriate numbers of samples and then summed to produce the \textsc{snr}s for original templates in the template bank.
+Each time slice may be implemented using either \textsc{fir} filters or \textsc{fft} convolution.
+If the $i$th time slice spans the times $[-t_\mathrm{end}^i, -t_\mathrm{start}^i)$ relative to the time of coalescence, then the latency is $\max \, (2 (t_\mathrm{end}^i - t_\mathrm{start}^i) - t_\mathrm{start}^i)$, where the maximum is taken over all time slices that are implemented using \textsc{fft} convolution.
+
+We can exploit the time-frequency structure of the templates by processing each time slice at a reduced sample rate.
+In the quadrupole approximation~\cite{finn1993}, the frequency of gravitational radiation resulting from the inspiral of two compact objects evolves according to
 \begin{equation}
-\label{eq:time_domain_snr}
-\rho_i(t) = \int h_i(t) s(t-\tau) \, d\tau,
+\label{eq:fgw}
+f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{-t} \right]^{3/8}
 \end{equation}
-where $s(t)$ is the detector data stream suitably whitened having mean zero
-and unity variance.  Time domain convolutions don't have any inherent latency. 
-The template waveform can be ``slid onto" the detector data as it is produced
-giving the possibility of measuring a peak SNR at the time of inspiral
-coalescence.  Typically it is not tractable to compute SNR via a
-time domain convolution as indicated in (\ref{eq:time_domain_snr}) because
-it involves an $\mathcal{O}(N^2)$ operation per template. Instead the
-convolution theorem is applied and the SNR is obtained via an inverse Fourier
-transform
+where $\mathcal{M}$ is the chirp mass and $t$ is the time relative to the coalescence of the binary~\cite{kidder1992, blanchet2002, findchirppaper, hanna2009}.
+\editorial{Way too many citations here.} Typically the template is truncated at some time before coalescence, corresponding to a finite frequency which is often chosen to be the frequency at the innermost stable circular orbit, $f_\mathrm{final} = f_\mathrm{ISCO} = 4400 \, M_\odot / M$.
+
+Applying the relationship $f/f_\mathrm{final} = (t_\mathrm{final}/t)^{3/8}$ it is possible to design time slices such that the templates are critically sampled at monotonically decreasing sample rates.
+The data stream is downsampled to many different sample rates, then filtered in each time slice, then upsampled to a high common sample rate, and summed.
+In order to work efficiently with radix-2 \textsc{fft}s, sample rates and sample counts for time slices are both constrained to be powers of 2.
+
+\editorial{I didn't mention the upsampling and downsampling overhead, but I included it in my calculation.
+Should I describe this here, in an appendix, or in a publicly available tech note in the \textsc{dcc}?} An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices} below.
+For this example, the latency for this time slice design is just $\input{time_slice_latency.tex}$ even if all of the time slices are implemented with the \textsc{fft} method.
+This set of time slices will require $\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir} filter method without time slices.
+
+\begin{table}[!h]
+\caption{Example of critically sampled, power-of-2 time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from $f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$ with a time frequency structure given by ($\ref{eq:fgw})$.}
+\label{table:time_slices}
+\begin{center}
+\input{time_slices.tex}
+\end{center}
+\end{table}
+
+\editorial{\textsc{mbta} should be cited in the introduction.}  This idea has been demonstrated by the Virgo Collaboration's \textsc{mbta} pipeline~\cite{beauville2006,beauville2008}, which operated in a low-latency mode during \textsc{ligo}'s sixth science run and Virgo's second science run in 2010.
+
+A similar procedure can be applied for any signal family that is piecewise band-limited, as long as the time-frequency evolution is understood, either symbolically as in equation~(\ref{eq:fgw}), or numerically.
+
+\subsection{Reducing the number of filters with the singular value decomposition}
+
+Our second innovation exploits the fact that the templates in inspiral template banks are, by design, highly correlated.
+It is possible to greatly reduce the number of matched filters required to achieve a particular minimum match by designing an appropriate set of orthonormal {\em basis templates}.
+A purely numerical technique based on the singular value decomposition (\textsc{svd}) is demonstrated by the authors in~\cite{Cannon:2010p10398}.
+
+\editorial{I'm sweeping the complex nature of the templates under the rug here.
+Is it OK to regard our packing of the real and imaginary parts of the \textsc{svd} as an implementation detail?
+It makes the math a little more concise.}  One may regard a bank of $N$ discretely sampled templates of length $M$ samples, $h^\mathrm W(\theta_i; t_j) = [\mathbf H]_{ij} = H_{ij}$, as a matrix.
+The singular value decomposition is an exact factorization that exists for any matrix such that
 \begin{equation}
-\label{eq:freq_domain_snr}
-\rho_i(t) = \int \tilde{h}_i(f)\,^*\,\tilde{s}(f)\,e^{-2\pi i f t}\, df,
+\label{eq:svd-exact}
+H_{ij} = [\mathbf {V \Sigma U}]_{ij} = \sum_{k=1}^{N} v_{ik} \sigma_k u_{kj},
 \end{equation}
-where the $\tilde{h}_i(f)$ and $\tilde{s}(f)$ denote the Fourier transforms
-of the template and detector data respectively.  This operation is far
-faster to compute with $\mathcal{O}(N\,{\rm log}\,N)$ operations per template.  
-However, the acausal nature of the Fourier transform introduces an inherent
-latency.  One must wait until they have collected data as long as their 
-signal and for minimally overlapping computations that means having a latency
-about as long as the template waveform.  The latency can be reduced by having
-redudant overlapping transforms, but at an additional cost.
+where $\mathbf V$ and $\mathbf U$ are both unitary matrices and $\mathbf \Sigma$ is a diagonal matrix.
+For our purposes, we associate the rows of $\mathbf U$ with a minimal set of basis templates, which become the kernels of basis filters.
+These filters give rise to the orthogonal \textsc{snr}s, $[\mathbf \rho^\perp]_k = \rho_k^\perp = \mathbf{U}_k \star x^\mathsf{W}$.
+The rows of $\mathbf{V \Sigma}$ become reconstruction coefficients that map linear combinations of the orthogonal \textsc{snr}s from the basis filters back onto \textsc{snr}s for the original templates of interest: $\rho_i = \sum_k [\mathbf V \mathbf \Sigma]_{ik} \rho_k^\perp$.
 
-Thus we have that low latency searches are inherently more computationally
-costly than moderate latency searches that employ frequency domain matched
-filtering.  The next section discusses a method to mitigate the cost of
-low latency searches.
+In many applications of the \textsc{svd}, including ours, the matrix can be well approximated by truncating the summation in equation~(\ref{eq:svd-exact}) at $L \ll N$:
+\begin{equation}
+H_{ij}^\prime = \sum_{k=1}^{L} v_{ik} \sigma_k u_{kj}
+\end{equation}
+The cumulative sum of squares of the {\em singular values}, $\sigma_k$, measures the Frobenius norm of the approximation, such that
+\begin{equation}
+\frac{\| \mathbf H^\prime \|}{\| \mathbf H \|} = \left(\sum_{k=1}^{L} |\sigma_k|^2\right)^{1/2} \left(\sum_{k=1}^{N} |\sigma_k|^2\right)^{-1/2}.
+\end{equation}
+This is also called the \textsc{svd} tolerance.  In our application, it relates to how much \textsc{snr} is lost by discarding $N - L$ of the basis filters with the lowest singular values.
 
-\subsection{Reducing the number of filters with singular value decomposition}
+\begin{comment}
+
+% Rewritten above.
+
 It is clear that in order to reduce the cost of a low latency search for
@@
 
-This result differs from other work that models gravitational-wave chirp
-signals in approximate ways 
-~\cite{Chassande-Mottin2006, Candes2008, BuonannoChenVallisneri:2003a} 
-by starting with an exact representation of the desired template family and
-producing a rigorous approximation with a tunable accuracy.
+\end{comment}
 
-\subsection{Detection statistic}
+\editorial{This passage, from the original draft, sounds confrontational to me, but these are good citations.
+Can we tone it down or move it elsewhere?}  This result differs from other work that models gravitational-wave chirp signals in approximate ways~\cite{Chassande-Mottin2006, Candes2008, BuonannoChenVallisneri:2003a} by starting with an exact representation of the desired template family and producing a rigorous approximation with a tunable accuracy.
+
+\subsection{Composite detection statistic and hierarchical detection}
+
+Although the \textsc{svd} allows us to reduce the number of matched filters, this comes at the price of having to perform a matrix multiplication by the reconstruction matrix $\mathbf {V \Sigma}$ at every sample.
+In some circumstances, this matrix multiplication may be more expensive than applying the orthogonal matched filters.
+
+\editorial{We have an old block diagram of our pipeline, but it takes up a lot of space.
+A new, more compact one that uses conventional signal-flow diagram symbols would be nice.}  Further speedup may be gained in a hierarchical detection scheme.
+In general, the orthogonal \textsc{snr}s alone provide some indication of whether any template in the original template bank is likely to have a large \textsc{snr}.
+Consider some composite detection statistic that is a scalar function of the orthogonal \textsc{snr}s, $\Gamma(\rho_1^\perp, \rho_2^\perp, \cdots, \rho_L^\perp)$.
+Suppose that we can infer the distribution of $\Gamma$ for a data stream with the signal present or with the signal absent.
+Then we may use a threshold crossing of $\Gamma$ to trigger the conditional application of the expensive reconstruction matrix only during the times when a signal is likely to be present.
+
+The composite detection statistic that we employ is a weighted sum of squares, $\Gamma = \sum_{k=1}^L w_k (\rho_k^\perp) ^ 2$, with the particular weights
+\begin{equation}
+w_k = \frac{\sigma_k^2}{\sigma_k^2 + N / A^2}.
+\end{equation}
+\editorial{We now have a pretty good description of time slices, the SVD, and conditional reconstruction.
+We still need to tie it all together.}Here, $A$ is a desired \textsc{snr} scale that is set by the analyst.
+As the authors show in ~\cite{svd-compdetstat}, this choice of weights has a better receiver operating characteristic for a signal of \textsc{snr}~=~$A$ than some other obvious choices, $w_k = \sigma_k^2$ or $w_k = 1$.
+
+\begin{comment}
+
+% Fold this into SVD discussion.
+
+%\subsection{Detection statistic}
 It is not necessary to reconstruct the SNR of all templates in order to 
@@
 potentially interesting times can be reconstructed on a sample by sample 
-basis.  Following the ideas in ~\cite{Anderson:2000yy} we employ a detection
+basis.  Following the ideas in~\cite{Anderson:2000yy} we employ a detection
 statistic as
@@
 \end{equation}
+
+\end{comment}
+
+\begin{comment}
+
+% We aren't using this chi square test.
 
@@
 
-\subsection{Selectively reducing the sample rate of the data and template
-waveforms}
-The computational cost of filtering can be further reduced by selectively
-downsampling certain intervals of the template waveform.  Note that the
-frequency of gravitational radiation resulting from the inspiral of two
-compact object is given by the quadrupole aproximation as ~\cite{finn1993}
-\begin{equation}
-\label{eq:fgw}
-f_{\rm GW} = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t_c} \right]^{3/8}
-\end{equation}
-where $\mathcal{M}$ is the chirpmass in units of $G=c=1$ and $t_c$ is the time
-before the coalescence of the binary.  Typically the filter is integrated to
-a finite frequency, $f_e$ which corresponds to a time, $t_e$ before the 
-coalescence~\cite{kidder1992, blanchet2002, findchirppaper, hanna2009}.  Given
-the relationship $f/f_e = (t_e/t)^{3/8}$ it is possible to reduce a
-filter waveform into critically sampled bands by dividing them up in time.
-This is possible regardless of the time-frequency relationship as long as
-it can be computed rigorously and thus the method doesn't depend on 
-(\ref{eq:fgw}).  As a concrete example consider a waveform obeying 
-(\ref{eq:fgw}) with a chirp mass of 1 $M_{\odot}$ for which we decide
-to cut the filter at $f_e = 2048$Hz (corresponding to $t_e = 0.00093$s).  
-It is then possible to compute sub bands which are each critically sampled
-according to table \ref{table1}  
-\begin{table}[!h]
-\caption{
-Example critically sampled sub bands for a 1 $M_{\odot}$ filter with a 
-time frequency structure given by ($\ref{eq:fgw})$.  Most of the filter is
-low frequency and it is possible to save computation by breaking up the filter
-into bands each of which has restricted frequency content. 
-}
-\begin{center}
-\label{table1}
-\begin{tabular}{@{} lll @{}}
-\hline
-\hline
-Sub band & Sample rate (Hz) &Interval (s) \\
-\hline
-1 & 2048 & [0.00093, 0.0059] \\
-2 & 1024 & [0.0059, 0.037]  \\
-3 & 512  & [0.037, 0.24]   \\
-4 & 256  & [0.24, 1.51]   \\
-5 & 128  & [1.51, 9.60]   \\
-6 & 64   & [9.60, 60.95]  \\
-7 & 32   & [60.95, 387.0]  \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-where it is assumed that the filter is cut to some finite length, 387s 
-for this example.  This idea has been demonstrated by the Virgo Collaborations
-MBTA pipeline~\cite{beauville2006, beauville2008}.
+\end{comment}
+
+\begin{comment}
+
+% I think that this part can be swalled into the computational costs discussion above.
 
@@
 use.
+
+\end{comment}
+
+\begin{comment}
+
+% Do we want to spend this much copy on the whitener?
+% What matters is that we have picked a low-latency whitening procedure.
 
@@
 than an order of magnitude increase in density over lalapps\_inspiral.
+\end{comment}

### 2011-02-04 gstlal_flop_budget.py

@@ -1,0 +1,289 @@
+#!/usr/bin/env python
+"""Floating point operation count budgets for LLOID."""
+__author__ = "Leo Singer <leo.singer@ligo.org>"
+# FIXME: are we counting a multiply-accumulate as one operation or two?
+
+
+import math
+
+
+#
+# Miscellaneous utility functions.
+#
+
+
+def lg(x):
+	"""Logarithm base 2."""
+	return math.log(x) / math.log(2.)
+
+
+#
+# Operation counts for convolution (or cross correlation) filters.
+#
+
+
+def td_conv_ops(kernel_len, count_channels=1):
+	"""Calculate the number of floating point operations per sample required to
+	apply a convolution filter with a kernel of kernel_len samples and
+	count_channels channels, using the FIR (time domain) implementation."""
+
+	# One multiply and one add per sample.
+	return 2 * kernel_len * count_channels
+
+
+def fd_conv_ops(kernel_len, block_len_factor=2, count_channels=1):
+	"""Calculate the number of floating point operations per sample required to
+	apply a convolution filter with a kernel of kernel_len samples and
+	count_channels channels, using the FFT (frequency domain) implementation and
+	a block size of (block_len) or (block_len_factor * kernel_len)."""
+
+	# 1 forward FFT and count_channels inverse FFTs
+	ops = 4 * lg(kernel_len * block_len_factor) * (count_channels + 1)
+
+	# count_channels dot products
+	ops += 2 * count_channels
+
+	# (1/block_len_factor) of the data gets reused from one block to the next
+	ops /= (1 - 1. / block_len_factor)
+
+	return ops
+
+
+#
+# Operation counts for miscellaneious signal processing operations.
+#
+
+
+def resample_filter_ops(kernel_len, in_rate, out_rate, count_channels=1):
+	"""Calculate the number of floating point operations per sample require to
+	apply an FIR resampling filter."""
+
+	# no-op if in_rate == out_rate
+	if in_rate == out_rate:
+		return 0
+
+	# FIR filter...
+	ops = td_conv_ops(kernel_len, count_channels)
+
+	# ... but (min(rates) / max(rates)) of the samples are zero.
+	ops *= float(min(in_rate, out_rate)) / max(in_rate, out_rate)
+
+	return ops
+
+
+#
+# Operation counts for LLOID methods.
+#
+
+
[…]

### 2011-02-04 time_slices.py
Forgot to commit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@600 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -1,0 +1,62 @@
+#!/usr/bin/env python
+
+# Imports
+
+from optparse import OptionParser, Option
+from gstlal.templates import time_slices
+import numpy as np
+
+# Command line interface
+
+opts, args = OptionParser(description = __doc__, option_list = [
+	Option("--mass1", type="float", metavar="solar masses", help="component mass 1"),
+	Option("--mass2", type="float", metavar="solar masses", help="component mass 2"),
+	Option("--flow", type="float", metavar="Hz", help="low frequency cutoff"),
+]).parse_args()
+
+if len(args) > 0:
+	raise ValueError("Too many arguments")
+
+for key in ("mass1", "mass2", "flow"):
+	if getattr(opts, key) is None:
+		raise ValueError("Required argument --%s is missing" % key)
+
+# Place time slices
+
+fhigh = 4400. / (opts.mass1 + opts.mass2)
+mass_pairs = ((opts.mass1, opts.mass2), )
+slices = time_slices(mass_pairs, opts.flow, fhigh)
+
+# Generate output table
+
+print r"\begin{tabular}{rr@{, }ll}"
+print r"\hline\hline"
+print r"$f_s$ (Hz) & [$t_\mathrm{end}$ & $t_\mathrm{start}$) & Samples \\"
+print r"\hline"
+old_rate = None
+for slice in slices:
+	begin = slice['begin']
+	if begin != 0.: # arrgh! floats have both negative and positive 0!
+		begin = -begin
+	if slice['rate'] == old_rate:
+		print r"\dots & [%(end)g&%(begin)g) & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
+	else:
+		print r"%(rate)d & [%(end)g&%(begin)g) & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'rate': slice['rate'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
+		old_rate = slice['rate']
+print r"\hline"
+print r"\end{tabular}"
+
+# Write latency
+
+print >>open("time_slice_latency.tex", "w"), (r"%g~\mathrm{s}" % max(2 * (slice['end'] - slice['begin']) - slice['begin'] for slice in slices)),
+
+# Write operation count
+
+rank_reduction = 0.1
+block_length_factor = 2.
+resample_kernel_length = 64
[…]

### 2011-02-05 introduction.tex
Small fixes and comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@601 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 peak amplitude of electromagetic radiation will occur soon after 
-gravitational-wave emmission~\cite{sylvestre2003}.  Thus in order to maximize
+gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
 the chance of a successful electromagnetic followup the latency of 
@@
 electromagnetic observations~\cite{triggeredsearches2008}. \textsc{ligo} observations
-ruled out the possibility of GRB70201 originating from the merger of a
+ruled out the possibility of GRB070201 originating from the merger of a
 \textsc{nsns} or \textsc{nsbh} system in the Andromeda galaxy~\cite{GRB070201}.
@@
 \editorial{Cite an in-preparation paper of Larry's?} Much work is currently underway in providing the best possible source 
-localization from gravitational-wave detector networks~\cite{markowitz2008, raymond2008, cavalier2006}.  Colloborations are currently forming to provide infrastructure for the gravitational-wave
+localization from gravitational-wave detector networks~\cite{markowitz2008, raymond2008, cavalier2006}.  Collaborations are currently forming to provide infrastructure for the gravitational-wave
 Astronomers to provide targets of opportunity for electromagnetic 

### 2011-02-05 method.tex
Small fixes and comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@601 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Much more commonly, the matched filter bank is implemented using \textsc{fft} convolutions, costing only $\mathcal O(M \lg N)$ per sample but having a latency of at least $M$ samples.
+Much more commonly, the matched filter bank is implemented using \textsc{fft} convolutions, costing only $\mathcal O(M \lg N)$ \editorial{$\lg$ is $\log_2$.
+Explain nomenclature?
+Switch to $\log_2$?} per sample but having a latency of at least $M$ samples.
 For example, for a bank of $1\,\mathrm{ks}$ templates sampled at $4096\,\mathrm{Hz}$, the \textsc{fft} implementation requires about about $2.2 \times 10^4$ times fewer floating point operations per sample than the \textsc{fir} implementation.
@@
 If the $i$th time slice spans the times $[-t_\mathrm{end}^i, -t_\mathrm{start}^i)$ relative to the time of coalescence, then the latency is $\max \, (2 (t_\mathrm{end}^i - t_\mathrm{start}^i) - t_\mathrm{start}^i)$, where the maximum is taken over all time slices that are implemented using \textsc{fft} convolution.
+\editorial{It may be worth pointing out that time-slicing buys you latency even
+without exploiting time-frequency structure. The downsampling really just buys
+you throughput. It may even be worth renaming the section.}
 

### 2011-02-06 time_slices.py
added time slice diagram
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@602 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -4,7 +4,11 @@
 
 from optparse import OptionParser, Option
 from gstlal.templates import time_slices
+from pylal import spawaveform
 import numpy as np
+import pylab
+from itertools import groupby, izip
+
 
 # Command line interface
 
@@ -26,6 +30,31 @@ for key in ("mass1", "mass2", "flow"):
 fhigh = 4400. / (opts.mass1 + opts.mass2)
 mass_pairs = ((opts.mass1, opts.mass2), )
 slices = time_slices(mass_pairs, opts.flow, fhigh)
+mc = spawaveform.chirpmass(opts.mass1, opts.mass2)
+
+
+# Generate plot
+tmin = min(slice['begin'] for slice in slices)
+tmax = max(slice['end'] for slice in slices)
+legend_artists = []
+legend_labels = []
+rate_key = lambda x: x['rate']
+for color, (rate, more_slices) in izip(pylab.linspace(1., 0., 5), groupby(sorted(slices, key = rate_key), key = rate_key)):
+	legend_artists += [pylab.Rectangle((0, 0), 1, 1, facecolor = str(color))]
+	legend_labels += ['%d Hz' % rate]
+	for slice in more_slices:
+		t = pylab.arange(-slice['end'], -slice['begin'])
+		a = (-0.2 * t / mc) ** (-1./4)
+		pylab.fill_between(t, -a, a, facecolor = str(color))
+		pylab.axvline(-slice['end'], color = 'k', linestyle = ':')
+pylab.gca().get_yaxis().set_visible(False)
+pylab.legend(legend_artists, legend_labels, loc = 'lower left')
+pylab.xlim(-tmax, 0.)
+pylab.ylim(-.15, .15)
+pylab.xlabel('time relative to coalescence')
+pylab.title(r'Time slices for a %g $\emdash$ %g M$_\odot$ inspiral' % (opts.mass1, opts.mass2))
+pylab.savefig('time_slices.pdf')
+
 
 # Generate output table
 

### 2011-02-06 time_slices.py
added y-axis label to plot
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@603 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -47,11 +47,12 @@ for color, (rate, more_slices) in izip(pylab.linspace(1., 0., 5), groupby(sorted
 		a = (-0.2 * t / mc) ** (-1./4)
 		pylab.fill_between(t, -a, a, facecolor = str(color))
 		pylab.axvline(-slice['end'], color = 'k', linestyle = ':')
-pylab.gca().get_yaxis().set_visible(False)
+pylab.yticks([], [])
 pylab.legend(legend_artists, legend_labels, loc = 'lower left')
 pylab.xlim(-tmax, 0.)
 pylab.ylim(-.15, .15)
 pylab.xlabel('time relative to coalescence')
+pylab.ylabel(r'gravitational wave strain amplitude')
 pylab.title(r'Time slices for a %g $\emdash$ %g M$_\odot$ inspiral' % (opts.mass1, opts.mass2))
 pylab.savefig('time_slices.pdf')
 

### 2011-02-07 inspiral_svd.tex
Updating format of authors and removing pointers to non-existent files
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@606 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\documentclass[showpacs,groupedaddress,showkeys,preprintnumbers]{iopart}
+\documentclass[showpacs,groupedaddress,showkeys,preprintnumbers]{revtex4-1}
+
 \input{packages}
@@
 
+% Institutions
+\newcommand{\LIGOCaltech}{LIGO Laboratory, California Institute of Technology, 
+Pasadena, CA 91125, USA}
+\newcommand{\CITA}{Canadian Institute for Theoretical Astrophysics, 60 St.
+George Street, University of Toronto, Toronto, ON M5S 3H8, Canada}
+\newcommand{\Perimeter}{Perimeter Institute for Theoretical Physics, Waterloo,
+Ontario N2L 2Y5, Canada}
+\newcommand{\AEI}{Albert-Einstein-Institut, Max-Planck-Institut f\"ur
+Gravitationsphysik, D-30167 Hannover, Germany}
+\newcommand{\Leibniz}{Leibniz Universit\"at Hannover, D-30167 Hannover, Germany}
+
 \begin{document}
 
-\title[LLOID]{A novel method for detecting coalescing binaries in near realtime with Advanced LIGO and beyond \editorial{Submit to PRD or CQG?}}
+\title[LLOID]{A novel method for detecting coalescing binaries in near realtime with Advanced LIGO and beyond}
 
@@
 
-\author{Kipp Cannon$^1$, Adrian Chapman$^2$, Chad Hanna$^3$, Drew Keppel$^4$, Antony C. Searle$^2$, Alan J. Weinstein$^2$}
+\author{Kipp~Cannon}
+\email{kipp.cannon@ligo.org}
+\affiliation{\CITA}
 
-\address{$^1$ Canadian Institute for Theoretical Astrophysics, ON, Canada}
-\address{$^2$ LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA}
-\address{$^3$ Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
-\address{$^3$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
+\author{Adrian~Chapman}
+\affiliation{\LIGOCaltech}
+
+\author{Chad~Hanna}
+\email{chad.hanna@ligo.org}
+\affiliation{\Perimeter}
+
+\author{Drew~Keppel}
+\email{drew.keppel@ligo.org}
+\affiliation{\AEI}
+\affiliation{\Leibniz}
+
+\author{Antony~C.~Searle}
+\email{antony.searle@ligo.org}
+\affiliation{\LIGOCaltech}
+
+\author{Alan~J.~Weinstein}
+\email{alan.weinstein@ligo.org}
+\affiliation{\LIGOCaltech}
 
 \begin{abstract}
-Conventional matched filter bank methods for the detection of gravitational waves from the inspiral of compact binaries are computationally expensive, have hundreds of seconds of unavoidable intrinsic latency, and require arrays of largely redundant matched filters.
-Novel detection methods that are more computationally efficient and have lower latency will be required to realize the full potential of advanced of gravitational wave detectors that are currently under construction.
-In this paper, we describe a new detection method that exploits the properties of inspiral waveforms using multi-rate filtering, principal component analysis, and hierarchical detection.
-We provide receiver operating characteristics from a prototype search pipeline that is capable of low-latency or near-realtime detection with greatly reduced computational requirements in comparison with previously described methods.
+%
+\editorial{Submit to PRD or CQG?}
+%
+Conventional matched filter bank methods for the detection of gravitational
+waves from the inspiral of compact binaries are computationally expensive, have
+hundreds of seconds of unavoidable intrinsic latency, and require arrays of
+largely redundant matched filters.  Novel detection methods that are more
+computationally efficient and have lower latency will be required to realize
+the full potential of advanced of gravitational wave detectors that are
+currently under construction.  In this paper, we describe a new detection
+method that exploits the properties of inspiral waveforms using multi-rate
+filtering, principal component analysis, and hierarchical detection.  We
+provide receiver operating characteristics from a prototype search pipeline
+that is capable of low-latency or near-realtime detection with greatly reduced
+computational requirements in comparison with previously described methods.
+%
 \end{abstract}
@@
 \pacs{95.55.Ym, 84.30.Vn, 95.75.Wx}
+
+\maketitle
 
@@
 
-\ack{\textsc{ligo} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the National Science Foundation and operates under cooperative agreement \textsc{phy}-\oldstylenums{0107417}.
-This paper has \textsc{ligo} Document Number \textsc{ligo}-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{1}.
+\acknowledgments
+\textsc{ligo} was constructed by the California Institute of Technology
+and Massachusetts Institute of Technology with funding from the National
+Science Foundation and operates under cooperative agreement
+\textsc{phy}-\oldstylenums{0107417}. This paper has \textsc{ligo} Document
+Number \textsc{ligo}-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{1}.
 
@@
 \input{appendix}
-
-\section*{References}
 

### 2011-02-07 method.tex
added time slice diagram
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@602 ccdd05d5-56c1-461a-bf98-85f6876523b4
Updating format of authors and removing pointers to non-existent files
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@606 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Should I describe this here, in an appendix, or in a publicly available tech note in the \textsc{dcc}?} An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices} below.
-For this example, the latency for this time slice design is just $\input{time_slice_latency.tex}$ even if all of the time slices are implemented with the \textsc{fft} method.
-This set of time slices will require $\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir} filter method without time slices.
+For this example, the latency for this time slice design is just
+%
+%$\input{time_slice_latency.tex}$
+%
+even if all of the time slices are implemented with the \textsc{fft} method.  This set of time slices will require
+%
+%$\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$
+%
+for pure \textsc{fft} cross-correlation without time slices, or
+%
+%$\input{time_slice_ops_td.tex}$
+%
+for the \textsc{fir} filter method without time slices.
 
@@
 \label{table:time_slices}
-\begin{center}
-\input{time_slices.tex}
-\end{center}
+\begin{minipage}[c]{0.5\textwidth}
+\centering
+%\includegraphics[trim = 3cm 0cm 1.5cm 0cm, scale=0.4]{time_slices.pdf}
+\end{minipage}
+\begin{minipage}[c]{0.3\textwidth}
+\centering
+%\input{time_slices.tex}
+\end{minipage}
 \end{table}
@@
 
-\editorial{This passage, from the original draft, sounds confrontational to me, but these are good citations.
-Can we tone it down or move it elsewhere?}  This result differs from other work that models gravitational-wave chirp signals in approximate ways~\cite{Chassande-Mottin2006, Candes2008, BuonannoChenVallisneri:2003a} by starting with an exact representation of the desired template family and producing a rigorous approximation with a tunable accuracy.
+\editorial{This sounds confrontational to me, but these are good citations.
+Can we tone it down?}  This result differs from other work that models gravitational-wave chirp signals in approximate ways~\cite{Chassande-Mottin2006, Candes2008, BuonannoChenVallisneri:2003a} by starting with an exact representation of the desired template family and producing a rigorous approximation with a tunable accuracy.
 
@@
 
-Although the \textsc{svd} allows us to reduce the number of matched filters, this comes at the price of having to perform a matrix multiplication by the reconstruction matrix $\mathbf {V \Sigma}$ at every sample.
+\editorial{We should cite \cite{Scharf:1994p12497} here.} Although the \textsc{svd} allows us to reduce the number of matched filters, this comes at the price of having to perform a matrix multiplication by the reconstruction matrix $\mathbf {V \Sigma}$ at every sample.
 In some circumstances, this matrix multiplication may be more expensive than applying the orthogonal matched filters.
 
-\editorial{We have an old block diagram of our pipeline, but it takes up a lot of space.
-A new, more compact one that uses conventional signal-flow diagram symbols would be nice.}  Further speedup may be gained in a hierarchical detection scheme.
+Further speedup may be gained in a hierarchical detection scheme.
 In general, the orthogonal \textsc{snr}s alone provide some indication of whether any template in the original template bank is likely to have a large \textsc{snr}.

### 2011-02-08 method.tex
Reverting 606
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@607 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Should I describe this here, in an appendix, or in a publicly available tech note in the \textsc{dcc}?} An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices} below.
-For this example, the latency for this time slice design is just
-%
-%$\input{time_slice_latency.tex}$
-%
-even if all of the time slices are implemented with the \textsc{fft} method.  This set of time slices will require
-%
-%$\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$
-%
-for pure \textsc{fft} cross-correlation without time slices, or
-%
-%$\input{time_slice_ops_td.tex}$
-%
-for the \textsc{fir} filter method without time slices.
+For this example, the latency for this time slice design is just $\input{time_slice_latency.tex}$ even if all of the time slices are implemented with the \textsc{fft} method.
+This set of time slices will require $\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir} filter method without time slices.
 
@@
 \centering
-%\includegraphics[trim = 3cm 0cm 1.5cm 0cm, scale=0.4]{time_slices.pdf}
+\includegraphics[trim = 3cm 0cm 1.5cm 0cm, scale=0.4]{time_slices.pdf}
 \end{minipage}
@@
 \centering
-%\input{time_slices.tex}
+\input{time_slices.tex}
 \end{minipage}

### 2011-02-08 results.tex
Huge set of changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@598 ccdd05d5-56c1-461a-bf98-85f6876523b4
added outline of injection population description
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@608 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{SECIV}\label{sec:results}
+
+\begin{itemize}
+\item 1 day of simulated $h(t)$.
+\item $h(t)$ has power spectrum \sout{prescribed for ``zero detuning, high power'' model in \cite{Shoemaker:2009p9770}} that somewhat resembles initial \textsc{ligo} noise models
+\item $h(t)$ generated by passing white, Gaussian noise through a bank of \textsc{iir} filters
+\item 1 noninjections run
+\item 10 injections runs
+\item injections are distributed uniformly in log distance, uniformly in sky location and binary orientation
+\item injections are reweighted to be uniformly distributed in volume
+\item injections are 80$\pm$20 seconds apart
+\item there are $\approx$ 10k injections
+\end{itemize}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.4]{figures/roc_99.pdf}
+\includegraphics[scale=0.4]{figures/roc_9995.pdf}
+\caption{Receiver operating characteristic (\textsc{roc}) curve of detection efficiency (\textsc{eff}) versus false alarm probability (\textsc{fap}).}
+\label{fig:roc}
+\end{center}
+\end{figure}

### 2011-02-11 appendix.tex
Adding comments and references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@613 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Per sample, this is $(8 \lg D + 2) / (1 - n/D)$ operations.
+\editorial{Drew: Why don't we change this to an overlap of $m$ samples so we can see what happens as we increase the overlap to reduce latency.}
 
 The \textsc{fir} filter implementation has the advantage that it has no intrinsic latency, whereas the \textsc{fft} convolution has at least the latency of the \textsc{fft} block size $D \geq 2 n$.
-For example, for a $1.4 - 1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the \textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.
+\editorial{Drew: Should the latency be D-n?} For example, for a $1.4 - 1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the \textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.
 However, the \textsc{fir} filter implementation has the disadvantage of much greater overhead per sample than the \textsc{fft} convolution.

### 2011-02-11 inspiral_svd.tex
Reverting 606
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@607 ccdd05d5-56c1-461a-bf98-85f6876523b4
Adding comments and references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@613 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-02-04 inspiral_svd.tex
@@
 
-\author{Kipp Cannon$^1$, Adrian Chapman$^2$, Chad Hanna$^3$, Drew Keppel$^4$, Antony C. Searle$^2$, Alan J. Weinstein$^2$}
+\author{Kipp Cannon$^1$, Adrian Chapman$^2$, Chad Hanna$^3$, Drew Keppel$^{4,5}$, Antony C. Searle$^2$, Leo Singer$^2$, Alan J. Weinstein$^2$}
 
@@
 \address{$^3$ Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
-\address{$^3$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
+\address{$^4$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
+\address{$^5$ Leibniz Universit\"{a}t Hannover, Hannover, Germany}
 
@@
 We provide receiver operating characteristics from a prototype search pipeline that is capable of low-latency or near-realtime detection with greatly reduced computational requirements in comparison with previously described methods.
+Help me!
 \end{abstract}

### 2011-02-11 introduction.tex
Adding comments and references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@613 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-As the gravitational wave detection horizon pushes outward, the ability to detect signals in near realtime will become valuable.  Having an electromagnetic or neutrino counterpart to a 
-gravitational-wave detection would not only increase the confidence in 
-the detection but will also greatly improve the astrophysical information
-available from the event.  Most models that predict simultaneous
-gravitational-wave and electromagnetic observations also predict that the
-peak amplitude of electromagetic radiation will occur soon after 
-gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
-the chance of a successful electromagnetic followup the latency of 
+As the gravitational wave detection horizon pushes outward, the ability to
+detect signals in near realtime will become valuable.  Having an
+electromagnetic or neutrino counterpart to a gravitational-wave detection would
+not only increase the confidence in the detection but will also greatly improve
+the astrophysical information available from the event.  Most models that
+predict simultaneous gravitational-wave and electromagnetic observations also
+predict that the peak amplitude of electromagetic radiation will occur soon
+after gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to
+maximize the chance of a successful electromagnetic followup the latency of
 gravitational-wave signal analysis must be made minimal.
 
-\editorial{This is old news, and doesn't seem relevant to me: EM-directed gravitational wave searches don't benefit from low-latency detection pipelines.} Work has already commenced to trigger gravitational-wave searches from
-electromagnetic observations~\cite{triggeredsearches2008}. \textsc{ligo} observations
-ruled out the possibility of GRB070201 originating from the merger of a
-\textsc{nsns} or \textsc{nsbh} system in the Andromeda galaxy~\cite{GRB070201}.
+\editorial{This is old news, and doesn't seem relevant to me: EM-directed
+gravitational wave searches don't benefit from low-latency detection
+pipelines.; Drew: I agree} Work has already commenced to trigger
+gravitational-wave searches from electromagnetic
+observations~\cite{triggeredsearches2008}. \textsc{ligo} observations ruled out
+the possibility of GRB070201 originating from the merger of a \textsc{nsns} or
+\textsc{nsbh} system in the Andromeda galaxy~\cite{GRB070201}.
 
@@
 
-This question has been investigated and it seems  at least
-possible that gravitational-wave detections could provide a region of the
-sky to prompt electromagnetic observation followups~\cite{sylvestre2003}.
-\editorial{Cite an in-preparation paper of Larry's?} Much work is currently underway in providing the best possible source 
-localization from gravitational-wave detector networks~\cite{markowitz2008, raymond2008, cavalier2006}.  Collaborations are currently forming to provide infrastructure for the gravitational-wave
-Astronomers to provide targets of opportunity for electromagnetic 
-astronomers~\cite{kanner2008}.  Most models that predict coincident
-gravitational-wave and electromagnetic observations also predict that the
-peak electromagetic fluence will occur soon after 
-gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
-the chance of a successful electromagnetic followup the latency of 
-gravitational-wave signal analysis must be made minimal.
+This question has been investigated and it seems  at least possible that
+gravitational-wave detections could provide a region of the sky to prompt
+electromagnetic observation followups~\cite{sylvestre2003}.  \editorial{Cite an
+in-preparation paper of Larry's? Drew: When it is on arXiv, yes.} Much work is
+currently underway in providing the best possible source localization from
+gravitational-wave detector networks~\cite{markowitz2008, raymond2008,
+cavalier2006}.  Collaborations are currently forming to provide infrastructure
+for the gravitational-wave Astronomers to provide targets of opportunity for
+electromagnetic astronomers~\cite{kanner2008}.  Most models that predict
+coincident gravitational-wave and electromagnetic observations also predict
+that the peak electromagetic fluence will occur soon after gravitational-wave
+emission~\cite{sylvestre2003}.  Thus in order to maximize the chance of a
+successful electromagnetic followup the latency of gravitational-wave signal
+analysis must be made minimal.
 

### 2011-02-11 method.tex
Adding comments and references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@613 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \end{equation}
-where $\mathcal{M}$ is the chirp mass and $t$ is the time relative to the coalescence of the binary~\cite{kidder1992, blanchet2002, findchirppaper, hanna2009}.
+where $\mathcal{M}$ is the chirp mass and $t$ is the time relative to the coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002, hanna2009}.
 \editorial{Way too many citations here.} Typically the template is truncated at some time before coalescence, corresponding to a finite frequency which is often chosen to be the frequency at the innermost stable circular orbit, $f_\mathrm{final} = f_\mathrm{ISCO} = 4400 \, M_\odot / M$.
@@
 
-\editorial{\textsc{mbta} should be cited in the introduction.}  This idea has been demonstrated by the Virgo Collaboration's \textsc{mbta} pipeline~\cite{beauville2006,beauville2008}, which operated in a low-latency mode during \textsc{ligo}'s sixth science run and Virgo's second science run in 2010.
+\editorial{\textsc{mbta} should be cited in the introduction.}  This idea has been demonstrated by the Virgo Collaboration's \textsc{mbta} pipeline~\cite{Marion2004,beauville2006,beauville2008,Buskulic2010}, which operated in a low-latency mode during \textsc{ligo}'s sixth science run and Virgo's second science run in 2010.
 

### 2011-02-11 time_slices.py

@@ -32,14 +32,14 @@ mass_pairs = ((opts.mass1, opts.mass2), )
 slices = time_slices(mass_pairs, opts.flow, fhigh)
 mc = spawaveform.chirpmass(opts.mass1, opts.mass2)
 
-
 # Generate plot
 tmin = min(slice['begin'] for slice in slices)
 tmax = max(slice['end'] for slice in slices)
 legend_artists = []
 legend_labels = []
 rate_key = lambda x: x['rate']
-for color, (rate, more_slices) in izip(pylab.linspace(1., 0., 5), groupby(sorted(slices, key = rate_key), key = rate_key)):
+num_rates = len(set([x[0] for x in sorted(slices, key = rate_key)]))
+for color, (rate, more_slices) in izip(pylab.linspace(1., 0., num_rates), groupby(sorted(slices, key = rate_key), key = rate_key)):
 	legend_artists += [pylab.Rectangle((0, 0), 1, 1, facecolor = str(color))]
 	legend_labels += ['%d Hz' % rate]
 	for slice in more_slices:
@@ -53,7 +53,7 @@ pylab.xlim(-tmax, 0.)
 pylab.ylim(-.15, .15)
 pylab.xlabel('time relative to coalescence')
 pylab.ylabel(r'gravitational wave strain amplitude')
-pylab.title(r'Time slices for a %g $\emdash$ %g M$_\odot$ inspiral' % (opts.mass1, opts.mass2))
+#pylab.title(r'Time slices for a %g $\emdash$ %g M$_\odot$ inspiral' % (opts.mass1, opts.mass2))
 pylab.savefig('time_slices.pdf')
 
 

### 2011-02-11 analysis.tex
Adding comments and references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@613 ccdd05d5-56c1-461a-bf98-85f6876523b4
Add comment on flops
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@614 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \editorial{It would be good to illustrate the layout of this particular template bank: masses spanned, time slice layout \dots}
+\editorial{Drew: I think this should say the FFT version of LLOID is almost 10 times faster than the conventional FFT method.}
+\editorial{Drew: It would be good to have these numbers for with SVD only, with SVD + TIME SLICES, and SVD+COMPDETSTAT. Then we can show which combination allows you to save time.}
 \begin{table}[!h]

### 2011-02-14 gstlal_flop_budget.py
Adding additional flop computations
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@616 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -5,6 +5,7 @@ __author__ = "Leo Singer <leo.singer@ligo.org>"
 
 
 import math
+import sys
 
 
 #
@@ -260,6 +261,15 @@ if __name__ == '__main__':
 				s = "%d%s" % (r, s)
 		return s
 
+	def prettyprint_sci(f):
+		f = r"%.1e"%f
+		s = r"$"
+		s += f.split('e')[0]
+		s += r" \times 10^{"
+		s += r"%i"%int(f.split('e')[-1])
+		s += r"}$"
+		return s
+
 	from optparse import Option, OptionParser
 	opts, args = OptionParser(usage = '%prog file').parse_args()
 
@@ -273,17 +283,28 @@ if __name__ == '__main__':
 	N = frags[0].mix_matrix.shape[1]
 	raw_kernel_length = (max(frag.end for frag in frags) - min(frag.start for frag in frags)) * max(rates)
 
-	print r"\begin{tabular}{r l}"
+	latencies = {}
+	latencies['fir'] = 1./max(rates)
+	latencies['fft'] = max([frag.end for frag in frags])
+	latencies['fft_slice'] = min([frag.end for frag in frags])
+
+	print r"\begin{tabular}{r c l}"
 	print r"\hline\hline"
-	print r"operations/sample & method \\"
+	print r"operations/sample & latency (s.) & method \\"
 	print r"\hline"
-	print r"%16s & %s \\" % (prettyprint_int(int(round(td_conv_ops(raw_kernel_length, count_channels=N)))), r"conventional \textsc{fir} method")
-	print r"%16s & %s \\" % (prettyprint_int(int(round(fd_conv_ops(raw_kernel_length, count_channels=N)))), r"conventional \textsc{fft} method")
-	print r"%16s & %s \\" % (prettyprint_int(int(round(td_conv_time_slice_ops(M, rates, count_channels=N)))), r"\textsc{fir} method with time slices")
-	print r"%16s & %s \\" % (prettyprint_int(int(round(fd_conv_time_slice_ops(M, rates, count_channels=N)))), r"\textsc{fft} method with time slices")
-	print r"%16s & %s \\" % (prettyprint_int(int(round(td_lloid_ops(L, M, N, rates)))), r"\textsc{fir} method with \textsc{lloid}")
[…]

### 2011-02-14 method.tex
fixed typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@617 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-This presents a dilemma: it seems that low latency detection using the \textsc{fir} implementation is prohibitively expensive, whereas computational cheap detection with the \textsc{fft} method comes with minutes to hours of latency.
+This presents a dilemma: it seems that low latency detection using the \textsc{fir} implementation is prohibitively expensive, whereas computationally cheap detection with the \textsc{fft} method comes with minutes to hours of latency.
 

### 2011-02-15 method.tex
Reworked time slice section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@618 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Our first innovation is based upon the common knowledge that gravitational waveforms from the inspiral stage of compact binary coalescences are chirps: they are quasi-monochromatic and slowly evolving in frequency.
+Our first innovation is to split each template into disjoint intervals, or \emph{time slices}.
+A matched filter is constructed for each time slice, whose outputs form an ensemble of partial \textsc{snr} streams.
+By linearity, these partial \textsc{snr} streams can be suitably time delayed and summed to reproduce exactly the \textsc{snr} with respect to the original template.
+By exploiting some general properties of inspiral waveforms, we shall see that time slices permit a reduction in latency relative to the \textsc{fft} matched filter, but with comparable computational overhead.
 
-\editorial{The time slice boundary notation is a little awkward and different identifiers are used in a few places in the source code.
-It would be nice to clean that up.}  It is possible to strike a balance between the low latency of the \textsc{fir} filter method and the low overhead of the \textsc{fft} method by breaking the templates into disjoint time intervals, or {\em time slices}.
-The outputs of the filters for all of the time slices are delayed by the appropriate numbers of samples and then summed to produce the \textsc{snr}s for original templates in the template bank.
-Each time slice may be implemented using either \textsc{fir} filters or \textsc{fft} convolution.
-If the $i$th time slice spans the times $[-t_\mathrm{end}^i, -t_\mathrm{start}^i)$ relative to the time of coalescence, then the latency is $\max \, (2 (t_\mathrm{end}^i - t_\mathrm{start}^i) - t_\mathrm{start}^i)$, where the maximum is taken over all time slices that are implemented using \textsc{fft} convolution.
-\editorial{It may be worth pointing out that time-slicing buys you latency even
-without exploiting time-frequency structure. The downsampling really just buys
-you throughput. It may even be worth renaming the section.}
+\editorial{\textsc{mbta} should be cited in the introduction.}  A similar idea has been demonstrated by the Virgo Collaboration's \textsc{mbta} pipeline~\cite{Marion2004,beauville2006,beauville2008,Buskulic2010}, which operated in a low-latency mode during \textsc{ligo}'s sixth science run and Virgo's second science run in 2010.
 
-We can exploit the time-frequency structure of the templates by processing each time slice at a reduced sample rate.
-In the quadrupole approximation~\cite{finn1993}, the frequency of gravitational radiation resulting from the inspiral of two compact objects evolves according to
+Each partial matched filter can be realized with either the \textsc{fir} or the \textsc{fft} method.
+If all of the time slices make use of the \textsc{fft} method, the resulting filter will have lower latency than a conventional \textsc{fft} matched filter, and also less computational overhead than an \textsc{fir} matched filter.
+A further benefit is that all of the time slices may be processed in parallel for additional speedup.
+
+If the template is known to be quasi-bandlimited, then the partial matched filter for each time slice may be processed at a reduce sample rate.
+Compact binary inspiral waveforms fit this description superbly: they are chirps whose frequency and amplitude rise according to power laws of time.
+They are not only quasi-bandlimited, but quasi-monochromatic.
+By exploiting this property, we can implement a matched filter for an inspiral waveform using a multirate filter bank that requires far fewer floating point operations.
+
+For concreteness and simplicity, let us consider an inspiral waveform in the quadrupole approximation, for which the time-frequency relation is
 \begin{equation}
 \label{eq:fgw}
-f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{-t} \right]^{3/8}
+f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t} \right]^{3/8}.
 \end{equation}
-where $\mathcal{M}$ is the chirp mass and $t$ is the time relative to the coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002, hanna2009}.
-\editorial{Way too many citations here.} Typically the template is truncated at some time before coalescence, corresponding to a finite frequency which is often chosen to be the frequency at the innermost stable circular orbit, $f_\mathrm{final} = f_\mathrm{ISCO} = 4400 \, M_\odot / M$.
+Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G M_\odot / c^3 \approx 5 \mu\mathrm{s}$) and $t$ is the time remaining until the coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002, hanna2009}.
+\editorial{Way too many citations here.} Usually the template is truncated at some prescribed time $t = t_0$ before coalescence, often chosen to correspond to the innermost stable circular orbit at which $f = f_\mathrm{ISCO} = 4400 \, M_\odot / M$.
+The full template will extend to some low frequency cutoff, $f = f_\mathrm{low}$ corresponding to a time $t_\mathrm{low}$.
+This template has a duration of $t_\mathrm{low} - t_\mathrm{0}$ and is critically sampled at a rate of $2 f_\mathrm{isco}$.
 
-Applying the relationship $f/f_\mathrm{final} = (t_\mathrm{final}/t)^{3/8}$ it is possible to design time slices such that the templates are critically sampled at monotonically decreasing sample rates.
-The data stream is downsampled to many different sample rates, then filtered in each time slice, then upsampled to a high common sample rate, and summed.
-In order to work efficiently with radix-2 \textsc{fft}s, sample rates and sample counts for time slices are both constrained to be powers of 2.
+The time slices for this template consist of the $K$ intervals $(t_K, t_{K-1}], \dots, (t_2, t_1], (t_1, t_0]$ sampled at frequencies $f_\mathrm{K-1}, \dots, f_1, f_0$ where $f_0 = 2 f_\mathrm{ISCO}$, $t_0 = t_\mathrm{ISCO}$, $f_{K-1} \geqslant 2 f_\mathrm{low}$, and $t_K \leqslant t_\mathrm{low}$.
 
-\editorial{I didn't mention the upsampling and downsampling overhead, but I included it in my calculation.
-Should I describe this here, in an appendix, or in a publicly available tech note in the \textsc{dcc}?} An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices} below.
+If all of these time slices are implemented using the \textsc{fft}, then the latency of this filter bank is $\max_j \left( 2(t_j - t_{j-1}) - t_{j-1} \right) = \max_j \left( 2 t_j - 3 t_{j-1} \right)$.
+
+An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices} below.
 For this example, the latency for this time slice design is just $\input{time_slice_latency.tex}$ even if all of the time slices are implemented with the \textsc{fft} method.
@@
 
-\begin{table}[!h]
+\begin{table}[h!]
 \caption{Example of critically sampled, power-of-2 time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from $f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$ with a time frequency structure given by ($\ref{eq:fgw})$.}
@@
 \end{table}
-
-\editorial{\textsc{mbta} should be cited in the introduction.}  This idea has been demonstrated by the Virgo Collaboration's \textsc{mbta} pipeline~\cite{Marion2004,beauville2006,beauville2008,Buskulic2010}, which operated in a low-latency mode during \textsc{ligo}'s sixth science run and Virgo's second science run in 2010.
-
-A similar procedure can be applied for any signal family that is piecewise band-limited, as long as the time-frequency evolution is understood, either symbolically as in equation~(\ref{eq:fgw}), or numerically.
 

### 2011-02-15 time_slices.py

@@ -61,18 +61,13 @@ pylab.savefig('time_slices.pdf')
 
 print r"\begin{tabular}{rr@{, }ll}"
 print r"\hline\hline"
-print r"$f_s$ (Hz) & [$t_\mathrm{end}$ & $t_\mathrm{start}$) & Samples \\"
+print r"$f_k$ (Hz) & ($t_{k+1}$ & $t_k$] & Samples \\"
 print r"\hline"
-old_rate = None
 for slice in slices:
 	begin = slice['begin']
 	if begin != 0.: # arrgh! floats have both negative and positive 0!
 		begin = -begin
-	if slice['rate'] == old_rate:
-		print r"\dots & [%(end)g&%(begin)g) & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
-	else:
-		print r"%(rate)d & [%(end)g&%(begin)g) & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'rate': slice['rate'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
-		old_rate = slice['rate']
+	print r"%(rate)d & [%(end)g&%(begin)g) & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'rate': slice['rate'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
 print r"\hline"
 print r"\end{tabular}"
 

### 2011-02-15 method.tex
reversed signs of time slice boundaries
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@619 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{eq:fgw}
-f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t} \right]^{3/8}.
+f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{-t} \right]^{3/8}.
 \end{equation}
-Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G M_\odot / c^3 \approx 5 \mu\mathrm{s}$) and $t$ is the time remaining until the coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002, hanna2009}.
-\editorial{Way too many citations here.} Usually the template is truncated at some prescribed time $t = t_0$ before coalescence, often chosen to correspond to the innermost stable circular orbit at which $f = f_\mathrm{ISCO} = 4400 \, M_\odot / M$.
-The full template will extend to some low frequency cutoff, $f = f_\mathrm{low}$ corresponding to a time $t_\mathrm{low}$.
-This template has a duration of $t_\mathrm{low} - t_\mathrm{0}$ and is critically sampled at a rate of $2 f_\mathrm{isco}$.
+Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G M_\odot / c^3 \approx 5 \mu\mathrm{s}$) and $t$ is the time relative to the coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002, hanna2009}.
+\editorial{Way too many citations here.} Usually the template is truncated at some prescribed time $t = t_0$, often chosen to correspond to the innermost stable circular orbit at which $f = f_\mathrm{ISCO} \approx 4400 \, M_\odot / M$.
+An inspiral signal will enter the \textsc{ligo} band at a low frequency cutoff, $f = f_\mathrm{low}$ corresponding to a time $t_\mathrm{low}$.
+This template has a duration of $t_\mathrm{0} - t_\mathrm{low}$ and is critically sampled at a rate of $2 f_\mathrm{isco}$.
 
@@
 
-If all of these time slices are implemented using the \textsc{fft}, then the latency of this filter bank is $\max_j \left( 2(t_j - t_{j-1}) - t_{j-1} \right) = \max_j \left( 2 t_j - 3 t_{j-1} \right)$.
-
-An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices} below.
-For this example, the latency for this time slice design is just $\input{time_slice_latency.tex}$ even if all of the time slices are implemented with the \textsc{fft} method.
-This set of time slices will require $\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir} filter method without time slices.
+If all of these time slices are implemented using the \textsc{fft}, then the latency of this filter bank is $\max_j {\left[-\left( 2(t_j - t_{j-1}) - t_{j-1} \right)\right]} = \max_j \left( 3 t_{j-1} - 2 t_j \right)$.
 
@@
 \end{table}
+
+An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices}.
+For this example, the latency for this time slice design is just $\input{time_slice_latency.tex}$ even if all of the time slices are implemented with the \textsc{fft} method.
+This set of time slices will require $\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir} filter method without time slices.
 

### 2011-02-15 time_slices.py
fixed some notation mismatch issues
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@620 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -61,13 +61,13 @@ pylab.savefig('time_slices.pdf')
 
 print r"\begin{tabular}{rr@{, }ll}"
 print r"\hline\hline"
-print r"$f_k$ (Hz) & ($t_{k+1}$ & $t_k$] & Samples \\"
+print r"$f_k$ (Hz) & ($t_{k+1}$ & $t_k$] (s) & Samples \\"
 print r"\hline"
 for slice in slices:
 	begin = slice['begin']
 	if begin != 0.: # arrgh! floats have both negative and positive 0!
 		begin = -begin
-	print r"%(rate)d & [%(end)g&%(begin)g) & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'rate': slice['rate'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
+	print r"%(rate)d & $(%(end)g$&$%(begin)g]$ & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'rate': slice['rate'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
 print r"\hline"
 print r"\end{tabular}"
 

### 2011-02-15 method.tex
some refinements to svd language
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@621 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 It is possible to greatly reduce the number of matched filters required to achieve a particular minimum match by designing an appropriate set of orthonormal {\em basis templates}.
-A purely numerical technique based on the singular value decomposition (\textsc{svd}) is demonstrated by the authors in~\cite{Cannon:2010p10398}.
+A purely numerical technique based on the application of the singular value decomposition (\textsc{svd}) to inspiral waveforms is demonstrated by the authors in~\cite{Cannon:2010p10398}.
 

### 2011-02-15 inspiral_svd.tex
added Nick to author list
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@622 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added Toronto to CITA byline
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@623 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\author{Kipp Cannon$^1$, Adrian Chapman$^2$, Chad Hanna$^3$, Drew Keppel$^{4,5}$, Antony C. Searle$^2$, Leo Singer$^2$, Alan J. Weinstein$^2$}
+\author{Kipp Cannon$^1$, Adrian Chapman$^2$, Nickolas Fotopoulos$^2$, Chad Hanna$^3$, Drew Keppel$^{4,5}$, Antony C. Searle$^2$, Leo Singer$^2$, Alan J. Weinstein$^2$}
 
-\address{$^1$ Canadian Institute for Theoretical Astrophysics, ON, Canada}
+\address{$^1$ Canadian Institute for Theoretical Astrophysics, Toronto, ON, Canada}
 \address{$^2$ LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA}

### 2011-02-15 introduction.tex
lloid_paper intro: reworked introduction quite a bit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@625 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-02-05 introduction.tex
@@
 
-% I would like to say:
-%  - Low mass inspirals are the most promising source of gravitational radiation for advanced LIGO
+% Outline:
+%  - Low-mass inspirals are the most promising source of gravitational radiation for advanced \textsc{ligo}
+% - Motivation for high throughput
+%   - Point out challenges of advanced detectors: more templates, longer templates
+% - Motivation for low latency
+%   - Observational applications of near-realtime detection
 %   - Mention S6 and VSR2, online running, EM followup
-% - Observational applications of near-realtime detection
-% - Point out challenges of advanced detectors: more templates, longer templates
-% - Advanced LIGO timetable
-% - Einstein telescope
 % - Layout of paper: description of method, procedure followed
 
-\begin{comment}
+Low-mass inspiraling binaries are the most promising sources of gravitational
+radiation for Advanced \textsc{ligo}. Observation demonstrates that they must
+exist in the local Universe in some abundance~\cite{Abadie:2010p10836}.
+[science] In the
+final stages of inspiral, the neutron star is tidally disrupted, providing the
+matter required to produce bright electromagnetic
+counterparts~\cite{shibata:2007}. Conveniently, the gravitational waveforms are
+understood and implemented to the level required by Advanced \textsc{ligo} to
+detect~\cite{BuonannoIyerOchsnerYiSathya2009}, allowing the sensitive technique
+of matched filtering.
 
-However, the computational requirements of conventional detection strategies --- especially those that are well suited to low latency applications --- will be greatly increased by this upgrade for two reasons.
-First, advanced detectors will have much improved low-frequency sensitivity, so chirp signals from coalescing binaries will remain in the detectors' band for longer.
-This will demand that matched filter based searches use longer templates, requiring more memory and more cycles to process.
-Second, background noise for advanced detectors will be more spectrally uniform, so it will be possible to resolve the intrinsic parameters of a source that determine the time-frequency structure of the gravitational wave signal with much greater accuracy.
-This will require that matched filter bank based searches employ more matched filters, again demanding more memory and more cycles.
+The parameter space of compact binary coalescence signals is
+large~\cite{Owen:1995tm, Owen:1998dk} leading to large filter banks and
+significant computational cost. Following the series of detector upgrades that
+are now under way, \textsc{ligo} and Virgo will attain their Advanced
+configurations, gaining a tenfold improvement in amplitude sensitivity over
+Initial configurations~\cite{advLIGOrefdesign,advVirgorefdesign} and a
+corresponding thousandfold increase in observable volume in the local Universe.
+The computational requirements of conventional detection strategies will be
+greatly increased by this upgrade for two reasons. First, advanced detectors
+will have much improved low-frequency sensitivity, so chirp signals from
+coalescing binaries will remain in the detectors' band for a significantly
+longer time. This will demand that matched filter based searches use longer
+templates, requiring more memory and more cycles to process. Second, background
+noise for advanced detectors will be more spectrally uniform, so it will be
+possible to resolve the intrinsic parameters of a source that determine the
+time-frequency structure of the gravitational wave signal with much greater
+accuracy. This will require that matched filter bank based searches employ
+more matched filters, again demanding more memory and more cycles.
 
-\end{comment}
+Beyond raw throughput, the ability to detect signals in near realtime will
+become increasingly valuable as the gravitational-wave detection horizon pushes
+outward.  Having an electromagnetic or neutrino counterpart to a
+gravitational-wave detection would not only increase the confidence in the
+detection but will also greatly increase the astrophysical information
+available from the event. Most models that predict simultaneous
+gravitational-wave and electromagnetic observations also predict that the peak
+amplitude of electromagetic radiation will occur soon after gravitational-wave
+emission~\cite{sylvestre2003}. Thus in order to maximize the chance of a
+successful electromagnetic followup the latency of gravitational-wave signal
+analysis must be made minimal.
 
-\begin{comment}
+In \oldstylenums{2010}, \textsc{ligo} and Virgo completed S6/VSR3,
+a short period of
+joint data taking during which several all-sky detection pipelines operated in
+a low-latency configuration \citeneeded. A few candidates of moderate
+significance were promptly sent for electromagnetic followup \citeneeded{} to
+several telescopes including Swift, \textsc{rotse}, \textsc{tarot}, and Zadko.
+They achieved latencies of BLAH, but required human vetting of each candidate
+before alerting the telescope partners. As confidence in the infrastructure
+increases, humans can be removed from the loop, setting a demonstrated previous
+best latency at BLAH.
 
-\editorial{This paragraph has too many citations.} Modern gravitational-wave observatories are likely to deliver on the promise 
-that their name implies by observing multiple gravitational-wave signals from
-compact binary coalescence within the coming decade \cite{kopparapu2008}. 
-There is now a world-wide network of gravitational-wave observatories 
-including the laser interferometers
-\textsc{ligo}, Virgo, \textsc{geo}, and \textsc{tama}, which should assure that coincident detections 
-will be possible among at least three separate sites lending confidence
-to any detections~\cite{Abbott:2005qm, beauville2008, beauville2006, blackburn2005, LIGOS1instpaper}.
-\editorial{This is a lot of citations just to say that the 4 detectors exist.
-Maybe if they were scattered throughout the paragraph?}
-However, having an electromagnetic or neutrino counterpart to a 
-gravitational-wave detection would not only increase the confidence in 
-the detection but would also greatly improve the astrophysical information
-available from the event.
-
-\end{comment}
-
-In a series of detector upgrades that are now under way, \textsc{ligo} and Virgo will become the first advanced gravitational wave detectors, gaining a tenfold improvement in amplitude sensitivity \citeneeded{} and a corresponding thousandfold increase in observable volume in the local universe.
-It is anticipated that these detectors will be able to sense as many as 40 events per year \cite{Abadie:2010p10836}.
-
-As the gravitational wave detection horizon pushes outward, the ability to detect signals in near realtime will become valuable.  Having an electromagnetic or neutrino counterpart to a 
-gravitational-wave detection would not only increase the confidence in 
-the detection but will also greatly improve the astrophysical information
-available from the event.  Most models that predict simultaneous
-gravitational-wave and electromagnetic observations also predict that the
-peak amplitude of electromagetic radiation will occur soon after 
-gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
-the chance of a successful electromagnetic followup the latency of 
-gravitational-wave signal analysis must be made minimal.
-
-\editorial{This is old news, and doesn't seem relevant to me: EM-directed gravitational wave searches don't benefit from low-latency detection pipelines.} Work has already commenced to trigger gravitational-wave searches from
-electromagnetic observations~\cite{triggeredsearches2008}. \textsc{ligo} observations
-ruled out the possibility of GRB070201 originating from the merger of a
-\textsc{nsns} or \textsc{nsbh} system in the Andromeda galaxy~\cite{GRB070201}.
-
-In \oldstylenums{2010}, \textsc{ligo} and Virgo completed a period of joint data taking during which several all-sky detection pipelines operated in a low-latency configuration \citeneeded.
-Promising detection candidates were promptly sent for electromagnetic followup \citeneeded{} to several telescopes including Swift, \textsc{rotse}, \textsc{tarot}, and Zadko.
-
-This question has been investigated and it seems  at least
-possible that gravitational-wave detections could provide a region of the
-sky to prompt electromagnetic observation followups~\cite{sylvestre2003}.
-\editorial{Cite an in-preparation paper of Larry's?} Much work is currently underway in providing the best possible source 
-localization from gravitational-wave detector networks~\cite{markowitz2008, raymond2008, cavalier2006}.  Collaborations are currently forming to provide infrastructure for the gravitational-wave
-Astronomers to provide targets of opportunity for electromagnetic 
-astronomers~\cite{kanner2008}.  Most models that predict coincident
-gravitational-wave and electromagnetic observations also predict that the
-peak electromagetic fluence will occur soon after 
-gravitational-wave emission~\cite{sylvestre2003}.  Thus in order to maximize
-the chance of a successful electromagnetic followup the latency of 
-gravitational-wave signal analysis must be made minimal.
-
-The parameter space of compact binary coalescence signals is large~\cite{Owen:1995tm, Owen:1998dk}.  It is a
-computationally burdensome task to analyze these signals with even moderate
-latency~\cite{Abbott:2007xi}.  
-This work will describe how to exploit degeneracy in the signal
-parameter space to answer more quickly whether or not a gravitational-wave 
-is present. Specifically we will explore using the singular value decomposition (\textsc{svd}) to reduce the
-effective number of filters required to search the data.  We note that others
-have applied the use of \textsc{svd} to gravitational wave data analysis to analyze
-optimal gravitational-wave burst detection~\cite{bradyraymajumder2004, heng2008} and coherent networks 
-of detectors~\cite{wen2008} .
+This work will describe how to exploit degeneracy in the signal parameter space
+to answer more quickly and economically whether or not a gravitational wave is
+present.
+\editorial{nvf: I use some rather different words here to summarize the
+techniques. I don't want to simply repeat words from later, but I don't want
+to jar a reader with changing jargon either. Feel free to edit.}
+Specifically, we will exploit the time-frequency structure of chirp
+signals to downsample different parts of the waveform, use the singular value
+decomposition (\textsc{svd}) factors to identify the redundancy of the bank and
+reduce the effective number of filters required to search the data, and use
+matched subspace filters built from \textsc{svd} to identify when computation
+can be avoided.  We note that others have applied the use of \textsc{svd} to
+gravitational wave data analysis to analyze optimal gravitational-wave burst
+detection~\cite{bradyraymajumder2004, heng2008} and coherent networks of
+detectors~\cite{wen2008}.
 

### 2011-02-15 introduction.tex
lloid_paper intro: add more science and speculate on negative latency
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@630 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 exist in the local Universe in some abundance~\cite{Abadie:2010p10836}.
-[science] In the
+A detection would strongly constrain mass and
+spin~\cite{finn1993,Poisson:1995ef}, and with sufficient signal strength, the
+neutron star equation of state~\cite{flanagan:021502,Read:2009}, general
+relativity in the strong-field regime~\cite{Will:2005va}, the graviton
+mass~\cite{PhysRevD.80.044002,Keppel:2010qu}, and the Lorentz Invariance
+principle~\cite{Ellis2006402}. In the
 final stages of inspiral, the neutron star is tidally disrupted, providing the
@@
 will have much improved low-frequency sensitivity, so chirp signals from
-coalescing binaries will remain in the detectors' band for a significantly
+coalescing binaries will remain in the detectors' bands for a significantly
 longer time. This will demand that matched filter based searches use longer
@@
 
+\editorial{More speculative and no reference. Thoughts? Move to discussion? Worth adding any plots?}
+More tantalizingly, these long waveforms enable us to accumulate a significant
+fraction of the signal-to-noise ratio before coalescence. With enough
+signal-to-noise ratio (and a lot of work!) it should be possible to reconfigure
+the detectors on-the-fly to more sensitively capture the later merger part of
+the signal, where neutron star equation of state information can be extracted.
+With similar caveats, it is certainly possible to send alerts to telescopes
+several seconds before prompt emission. There are no theoretical hurdles to
+negative latency, but careful attention must be paid to the algorithms and
+implementation of search pipelines to take advantage of a fortuitously loud
+signal.
+
+\editorial{If we publish before the LUMIN papers, we have to redo this section.
+We can't make claims about those private communications with the telescopes.
+Else, cite the pre-S6 LUMIN paper \cite{kanner2008}.}
 In \oldstylenums{2010}, \textsc{ligo} and Virgo completed S6/VSR3,
@@
 before alerting the telescope partners. As confidence in the infrastructure
-increases, humans can be removed from the loop, setting a demonstrated previous
-best latency at BLAH.
+increases, humans can be removed from the loop, demonstrating a best latency
+of BLAH.
 

### 2011-02-17 gstlal_flop_budget.py
Adding latency column
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@632 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -294,10 +294,10 @@ if __name__ == '__main__':
 	print r"\hline"
 	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(td_conv_ops(raw_kernel_length, count_channels=N)))), prettyprint_sci(latencies['fir']), r"conventional \textsc{fir} method")
 	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(fd_conv_ops(raw_kernel_length, count_channels=N)))), prettyprint_sci(latencies['fft']), r"conventional \textsc{fft} method")
-	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(td_lloid_ops([int(N/7.)], [raw_kernel_length], N, [max(rates)], reconstruction_duty_cycle=1.)))), prettyprint_sci(latencies['fir']), r"\textsc{fir} method with \textsc{svd}")
-	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(fd_lloid_ops([int(N/7.)], [raw_kernel_length], N, [max(rates)], reconstruction_duty_cycle=1.)))), prettyprint_sci(latencies['fft']), r"\textsc{fft} method with \textsc{svd}")
-	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(td_lloid_ops([int(N/7.)], [raw_kernel_length], N, [max(rates)])))), prettyprint_sci(latencies['fir']), r"\textsc{fir} method with \textsc{lloid} and no time slices")
-	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(fd_lloid_ops([int(N/7.)], [raw_kernel_length], N, [max(rates)])))), prettyprint_sci(latencies['fft']), r"\textsc{fft} method with \textsc{lloid} and no time slices")
+	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(td_lloid_ops([int(N/3.)], [raw_kernel_length], N, [max(rates)], reconstruction_duty_cycle=1.)))), prettyprint_sci(latencies['fir']), r"\textsc{fir} method with \textsc{svd}")
+	print r"%16s & %s & %s \\" % (prettyprint_int(int(round(fd_lloid_ops([int(N/3.)], [raw_kernel_length], N, [max(rates)], reconstruction_duty_cycle=1.)))), prettyprint_sci(latencies['fft']), r"\textsc{fft} method with \textsc{svd}")
[…]

### 2011-02-17 introduction.tex
Adding comments about EM followup
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@633 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 detection but will also greatly increase the astrophysical information
-available from the event. Most models that predict simultaneous
-gravitational-wave and electromagnetic observations also predict that the peak
-amplitude of electromagetic radiation will occur soon after gravitational-wave
-emission~\cite{sylvestre2003}. Thus in order to maximize the chance of a
-successful electromagnetic followup the latency of gravitational-wave signal
-analysis must be made minimal.
+available from the event. \editorial{dk: I don't think this is true for
+neutrino counterparts. Just as there is no need to point GW detectors, there is
+no need to point neutrino detectors so the association can easily be done
+offline.} Most models that predict simultaneous gravitational-wave and
+electromagnetic observations also predict that the peak amplitude of
+electromagetic radiation will occur soon after gravitational-wave
+emission~\cite{sylvestre2003}. \editorial{dk: This is not true for all EM
+bands.  Radio afterglows from short GRBs are not expected to be seen until 1
+month after observation and then observable for several months.} Thus in order
+to maximize the chance of a successful electromagnetic followup the latency of
+gravitational-wave signal analysis must be made minimal.
 

### 2011-02-17 method.tex
Making method section 80 col width
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@634 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Searches for gravitational waves from compact binary coalescences typically employ matched filter banks \cite{findchirppaper}.
-Potential inspiral signals are continuously parameterized by time, amplitude, phase, and a set of intrinsic source parameters $\theta$, which in this paper we shall take to consist of the two component masses of a binary, $\theta = (m_1, m_2)$.
-Let $h_+(\theta)$ and $h_\times(\theta)$ be, respectively, the `$+$' and `$\times$' polarization gravitational wave signals that would arise from a fiducial face-on binary at some distance.
-Because, for inspiral signals, $h_+$ and $h_\times$ are nearly in quadrature, they are generally combined into a single complex-valued template $h = h_+ + i h_\times$.
+Searches for gravitational waves from compact binary coalescences typically
+employ matched filter banks \cite{findchirppaper}.  Potential inspiral signals
+are continuously parameterized by time, amplitude, phase, and a set of
+intrinsic source parameters $\theta$, which in this paper we shall take to
+consist of the two component masses of a binary, $\theta = (m_1, m_2)$.  Let
+$h_+(\theta)$ and $h_\times(\theta)$ be, respectively, the `$+$' and `$\times$'
+polarization gravitational wave signals that would arise from a fiducial
+face-on binary at some distance.  Because, for inspiral signals, $h_+$ and
+$h_\times$ are nearly in quadrature, they are generally combined into a single
+complex-valued template $h = h_+ + i h_\times$.
 
-The detection procedure for just one set of intrinsic source parameters $\theta$ starts by {\em whitening} the measurement data stream $x$.
-This involves finding a linear filter that renders the detector's noise \textsc{iid} and Gaussian.
-This filter is applied to the measurement, yielding the whitened data stream $x^\mathrm W$.
-The same linear filter is applied to the template $h(\theta)$, yielding the whitened template $h^\mathrm W (\theta)$.
-The matched filter is the normalized cross-correlation of $h^\mathrm W$ and $x^\mathrm W$, \editorial{After introducing the $\star$ notation for cross correlation, must we define it for continuous and discrete time?}
+The detection procedure for just one set of intrinsic source parameters
+$\theta$ starts by {\em whitening} the measurement data stream $x$.  This
+involves finding a linear filter that renders the detector's noise \textsc{iid}
+and Gaussian.  This filter is applied to the measurement, yielding the whitened
+data stream $x^\mathrm W$.  The same linear filter is applied to the template
+$h(\theta)$, yielding the whitened template $h^\mathrm W (\theta)$.  The
+matched filter is the normalized cross-correlation of $h^\mathrm W$ and
+$x^\mathrm W$, \editorial{After introducing the $\star$ notation for cross
+correlation, must we define it for continuous and discrete time?}
+%
 $$
-\rho(\theta) = \frac{h^\mathrm W (\theta) \star x^\mathrm W}{|h^\mathrm W (\theta)|}.
+%
+\rho(\theta) = \frac{h^\mathrm W (\theta) \star x^\mathrm W}{|h^\mathrm W
+(\theta)|}.
+%
 $$
-This is called the signal to noise ratio, or \textsc{snr}.
-The detection statistic is the modulus of this, $|\rho(\theta)|$, which has a $\chi^2$ distribution with 2 degrees of freedom in the absence of signal.%, and it indicates the probability of a signal being present \cite{wainstein:1962}\editorial{Comment about probability doesn't seem relevant.}.
+%
+This is called the signal to noise ratio, or \textsc{snr}.  The detection
+statistic is the modulus of this, $|\rho(\theta)|$, which has a $\chi^2$
+distribution with 2 degrees of freedom in the absence of signal.%, and it
+indicates the probability of a signal being present
+\cite{wainstein:1962}\editorial{Comment about probability doesn't seem
+relevant.}.
 
-To construct a filter bank, matched filters are realized for discrete signal parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$, such that any possible signal will have a maximum cross-correlation of at least 0.97 with at least one template.
-\editorial{Citation needed for template placement procedure?}  Such a template bank is said to have a 97\% {\em minimum match}.
-This technique is designed so that an inspiral signal can be detected without any prior knowledge of its intrinsic parameters: at most 3\% of the \textsc{snr} is lost by a signal's parameters not exactly coinciding with a template's.
-A trigger is reported for the template parameters $\theta_i$ and time $t$ for which $|\rho|$ is a maximum over some moving interval in $\theta$ and $t$.
+To construct a filter bank, matched filters are realized for discrete signal
+parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$, such that any possible
+signal will have a maximum cross-correlation of at least 0.97 with at least one
+template.  \editorial{Citation needed for template placement procedure?}  Such
+a template bank is said to have a 97\% {\em minimum match}.  This technique is
+designed so that an inspiral signal can be detected without any prior knowledge
+of its intrinsic parameters: at most 3\% of the \textsc{snr} is lost by a
+signal's parameters not exactly coinciding with a template's.  A trigger is
+reported for the template parameters $\theta_i$ and time $t$ for which $|\rho|$
+is a maximum over some moving interval in $\theta$ and $t$.
 
@@
 
-The matched filter bank can be implemented with finite impulse response (\textsc{fir}) filters.
-\textsc{fir} filters are perfectly suited for realtime detection, because they do not introduce any latency at all.
-However, \textsc{fir} filters are very expensive: for $N$ templates of $M$ samples each, a \textsc{fir} matched filter bank costs $\mathcal O(M N)$ operations per sample.
+The matched filter bank can be implemented with finite impulse response
+(\textsc{fir}) filters.  \textsc{fir} filters are perfectly suited for realtime
+detection, because they do not introduce any latency at all.  However,
+\textsc{fir} filters are very expensive: for $N$ templates of $M$ samples each,
+a \textsc{fir} matched filter bank costs $\mathcal O(M N)$ operations per
+sample.
 
-Much more commonly, the matched filter bank is implemented using \textsc{fft} convolutions, costing only $\mathcal O(M \lg N)$ \editorial{$\lg$ is $\log_2$.
-Explain nomenclature?
-Switch to $\log_2$?} per sample but having a latency of at least $M$ samples.
-For example, for a bank of $1\,\mathrm{ks}$ templates sampled at $4096\,\mathrm{Hz}$, the \textsc{fft} implementation requires about about $2.2 \times 10^4$ times fewer floating point operations per sample than the \textsc{fir} implementation.
-However, the \textsc{fft} implementation has a latency of at least $1\,\mathrm{ks}$.
+Much more commonly, the matched filter bank is implemented using \textsc{fft}
+convolutions, costing only $\mathcal O(M \lg N)$ \editorial{$\lg$ is $\log_2$.
+Explain nomenclature? Switch to $\log_2$?} per sample but having a latency of
+at least $M$ samples.  For example, for a bank of $1\,\mathrm{ks}$ templates
+sampled at $4096\,\mathrm{Hz}$, the \textsc{fft} implementation requires about
+about $2.2 \times 10^4$ times fewer floating point operations per sample than
+the \textsc{fir} implementation.  However, the \textsc{fft} implementation has
+a latency of at least $1\,\mathrm{ks}$.
 
-This presents a dilemma: it seems that low latency detection using the \textsc{fir} implementation is prohibitively expensive, whereas computationally cheap detection with the \textsc{fft} method comes with minutes to hours of latency.
+This presents a dilemma: it seems that low latency detection using the
+\textsc{fir} implementation is prohibitively expensive, whereas computationally
+cheap detection with the \textsc{fft} method comes with minutes to hours of
+latency.
 
-In the rest of the section, we will describe a detection strategy that makes use of some very general properties of inspiral template banks in order to evaluate a matched filter bank with far lower computational cost than the \textsc{fir} method and far less latency than the \textsc{fft} method.
+In the rest of the section, we will describe a detection strategy that makes
+use of some very general properties of inspiral template banks in order to
+evaluate a matched filter bank with far lower computational cost than the
+\textsc{fir} method and far less latency than the \textsc{fft} method.
 
@@
 
-Our first innovation is to split each template into disjoint intervals, or \emph{time slices}.
-A matched filter is constructed for each time slice, whose outputs form an ensemble of partial \textsc{snr} streams.
-By linearity, these partial \textsc{snr} streams can be suitably time delayed and summed to reproduce exactly the \textsc{snr} with respect to the original template.
-By exploiting some general properties of inspiral waveforms, we shall see that time slices permit a reduction in latency relative to the \textsc{fft} matched filter, but with comparable computational overhead.
+Our first innovation is to split each template into disjoint intervals, or
+\emph{time slices}.  A matched filter is constructed for each time slice, whose
+outputs form an ensemble of partial \textsc{snr} streams.  By linearity, these
+partial \textsc{snr} streams can be suitably time delayed and summed to
+reproduce exactly the \textsc{snr} with respect to the original template.  By
+exploiting some general properties of inspiral waveforms, we shall see that
+time slices permit a reduction in latency relative to the \textsc{fft} matched
+filter, but with comparable computational overhead.
 
-\editorial{\textsc{mbta} should be cited in the introduction.}  A similar idea has been demonstrated by the Virgo Collaboration's \textsc{mbta} pipeline~\cite{Marion2004,beauville2006,beauville2008,Buskulic2010}, which operated in a low-latency mode during \textsc{ligo}'s sixth science run and Virgo's second science run in 2010.
+\editorial{\textsc{mbta} should be cited in the introduction.}  A similar idea
+has been demonstrated by the Virgo Collaboration's \textsc{mbta}
+pipeline~\cite{Marion2004,beauville2006,beauville2008,Buskulic2010}, which
+operated in a low-latency mode during \textsc{ligo}'s sixth science run and
+Virgo's second science run in 2010.
 
-Each partial matched filter can be realized with either the \textsc{fir} or the \textsc{fft} method.
-If all of the time slices make use of the \textsc{fft} method, the resulting filter will have lower latency than a conventional \textsc{fft} matched filter, and also less computational overhead than an \textsc{fir} matched filter.
-A further benefit is that all of the time slices may be processed in parallel for additional speedup.
+Each partial matched filter can be realized with either the \textsc{fir} or the
+\textsc{fft} method.  If all of the time slices make use of the \textsc{fft}
+method, the resulting filter will have lower latency than a conventional
+\textsc{fft} matched filter, and also less computational overhead than an
+\textsc{fir} matched filter.  A further benefit is that all of the time slices
+may be processed in parallel for additional speedup.
 
-If the template is known to be quasi-bandlimited, then the partial matched filter for each time slice may be processed at a reduce sample rate.
-Compact binary inspiral waveforms fit this description superbly: they are chirps whose frequency and amplitude rise according to power laws of time.
-They are not only quasi-bandlimited, but quasi-monochromatic.
-By exploiting this property, we can implement a matched filter for an inspiral waveform using a multirate filter bank that requires far fewer floating point operations.
+If the template is known to be quasi-bandlimited, then the partial matched
+filter for each time slice may be processed at a reduce sample rate.  Compact
+binary inspiral waveforms fit this description superbly: they are chirps whose
+frequency and amplitude rise according to power laws of time.  They are not
+only quasi-bandlimited, but quasi-monochromatic.  By exploiting this property,
+we can implement a matched filter for an inspiral waveform using a multirate
+filter bank that requires far fewer floating point operations.
 
-For concreteness and simplicity, let us consider an inspiral waveform in the quadrupole approximation, for which the time-frequency relation is
-\begin{equation}
-\label{eq:fgw}
-f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{-t} \right]^{3/8}.
+For concreteness and simplicity, let us consider an inspiral waveform in the
+quadrupole approximation, for which the time-frequency relation is
+%
+\begin{equation} \label{eq:fgw}
+%
+f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{-t}
+\right]^{3/8}.
+%
 \end{equation}
-Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G M_\odot / c^3 \approx 5 \mu\mathrm{s}$) and $t$ is the time relative to the coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002, hanna2009}.
-\editorial{Way too many citations here.} Usually the template is truncated at some prescribed time $t = t_0$, often chosen to correspond to the innermost stable circular orbit at which $f = f_\mathrm{ISCO} \approx 4400 \, M_\odot / M$.
-An inspiral signal will enter the \textsc{ligo} band at a low frequency cutoff, $f = f_\mathrm{low}$ corresponding to a time $t_\mathrm{low}$.
-This template has a duration of $t_\mathrm{0} - t_\mathrm{low}$ and is critically sampled at a rate of $2 f_\mathrm{isco}$.
+%
+Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
+M_\odot / c^3 \approx 5 \mu\mathrm{s}$) and $t$ is the time relative to the
+coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002,
+hanna2009}. \editorial{Way too many citations here.} Usually the template is
+truncated at some prescribed time $t = t_0$, often chosen to correspond to the
+innermost stable circular orbit at which $f = f_\mathrm{ISCO} \approx 4400 \,
+M_\odot / M$.  An inspiral signal will enter the \textsc{ligo} band at a low
+frequency cutoff, $f = f_\mathrm{low}$ corresponding to a time
+$t_\mathrm{low}$.  This template has a duration of $t_\mathrm{0} -
+t_\mathrm{low}$ and is critically sampled at a rate of $2 f_\mathrm{isco}$.
 
-The time slices for this template consist of the $K$ intervals $(t_K, t_{K-1}], \dots, (t_2, t_1], (t_1, t_0]$ sampled at frequencies $f_\mathrm{K-1}, \dots, f_1, f_0$ where $f_0 = 2 f_\mathrm{ISCO}$, $t_0 = t_\mathrm{ISCO}$, $f_{K-1} \geqslant 2 f_\mathrm{low}$, and $t_K \leqslant t_\mathrm{low}$.
+The time slices for this template consist of the $K$ intervals $(t_K, t_{K-1}],
+\dots, (t_2, t_1], (t_1, t_0]$ sampled at frequencies $f_\mathrm{K-1}, \dots,
+f_1, f_0$ where $f_0 = 2 f_\mathrm{ISCO}$, $t_0 = t_\mathrm{ISCO}$, $f_{K-1}
+\geqslant 2 f_\mathrm{low}$, and $t_K \leqslant t_\mathrm{low}$.
 
-If all of these time slices are implemented using the \textsc{fft}, then the latency of this filter bank is $\max_j {\left[-\left( 2(t_j - t_{j-1}) - t_{j-1} \right)\right]} = \max_j \left( 3 t_{j-1} - 2 t_j \right)$.
+If all of these time slices are implemented using the \textsc{fft}, then the
+latency of this filter bank is $\max_j {\left[-\left( 2(t_j - t_{j-1}) -
+t_{j-1} \right)\right]} = \max_j \left( 3 t_{j-1} - 2 t_j \right)$.
 
 \begin{table}[h!]
-\caption{Example of critically sampled, power-of-2 time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from $f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$ with a time frequency structure given by ($\ref{eq:fgw})$.}
+\caption{Example of critically sampled, power-of-2 time slices for a $1.4 - 1.4
+\, M_{\odot}$ template extending from $f_\mathrm{low} = 10 \, \mathrm{Hz}$ to
+$f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$ with a time frequency structure given by
+($\ref{eq:fgw})$.}
 \label{table:time_slices}
@@
 
-An example time slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ is shown in table~\ref{table:time_slices}.
-For this example, the latency for this time slice design is just $\input{time_slice_latency.tex}$ even if all of the time slices are implemented with the \textsc{fft} method.
-This set of time slices will require $\input{time_slice_ops_firslice.tex}$ operations per sample per template, compared with $\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir} filter method without time slices.
+An example time slice design satisfying these constraints for a $1.4 - 1.4 \,
+M_{\odot}$ is shown in table~\ref{table:time_slices}.  For this example, the
+latency for this time slice design is just $\input{time_slice_latency.tex}$
+even if all of the time slices are implemented with the \textsc{fft} method.
+This set of time slices will require $\input{time_slice_ops_firslice.tex}$
+operations per sample per template, compared with
+$\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation
+without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir}
+filter method without time slices.
 
@@
 
-Our second innovation exploits the fact that the templates in inspiral template banks are, by design, highly correlated.
-It is possible to greatly reduce the number of matched filters required to achieve a particular minimum match by designing an appropriate set of orthonormal {\em basis templates}.
-A purely numerical technique based on the application of the singular value decomposition (\textsc{svd}) to inspiral waveforms is demonstrated by the authors in~\cite{Cannon:2010p10398}.
+Our second innovation exploits the fact that the templates in inspiral template
+banks are, by design, highly correlated.  It is possible to greatly reduce the
+number of matched filters required to achieve a particular minimum match by
+designing an appropriate set of orthonormal {\em basis templates}.  A purely
+numerical technique based on the application of the singular value
+decomposition (\textsc{svd}) to inspiral waveforms is demonstrated by the
+authors in~\cite{Cannon:2010p10398}.
 
 \editorial{I'm sweeping the complex nature of the templates under the rug here.
-Is it OK to regard our packing of the real and imaginary parts of the \textsc{svd} as an implementation detail?
-It makes the math a little more concise.}  One may regard a bank of $N$ discretely sampled templates of length $M$ samples, $h^\mathrm W(\theta_i; t_j) = [\mathbf H]_{ij} = H_{ij}$, as a matrix.
-The singular value decomposition is an exact factorization that exists for any matrix such that
+Is it OK to regard our packing of the real and imaginary parts of the
+\textsc{svd} as an implementation detail?  It makes the math a little more
+concise.}  One may regard a bank of $N$ discretely sampled templates of length
+$M$ samples, $h^\mathrm W(\theta_i; t_j) = [\mathbf H]_{ij} = H_{ij}$, as a
+matrix.  The singular value decomposition is an exact factorization that exists
+for any matrix such that
+%
+\begin{equation} \label{eq:svd-exact}
+%
+H_{ij} = [\mathbf {V \Sigma U}]_{ij} = \sum_{k=1}^{N} v_{ik} \sigma_k u_{kj},
+%
+\end{equation}
+%
+where $\mathbf V$ and $\mathbf U$ are both unitary matrices and $\mathbf
+\Sigma$ is a diagonal matrix.  For our purposes, we associate the rows of
+$\mathbf U$ with a minimal set of basis templates, which become the kernels of
+basis filters.  These filters give rise to the orthogonal \textsc{snr}s,
+$[\mathbf \rho^\perp]_k = \rho_k^\perp = \mathbf{U}_k \star x^\mathsf{W}$.  The
+rows of $\mathbf{V \Sigma}$ become reconstruction coefficients that map linear
+combinations of the orthogonal \textsc{snr}s from the basis filters back onto
+\textsc{snr}s for the original templates of interest: $\rho_i = \sum_k [\mathbf
+V \mathbf \Sigma]_{ik} \rho_k^\perp$.
+
+In many applications of the \textsc{svd}, including ours, the matrix can be
+well approximated by truncating the summation in equation~(\ref{eq:svd-exact})
+at $L \ll N$:
+%
 \begin{equation}
-\label{eq:svd-exact}
-H_{ij} = [\mathbf {V \Sigma U}]_{ij} = \sum_{k=1}^{N} v_{ik} \sigma_k u_{kj},
+%
+H_{ij}^\prime = \sum_{k=1}^{L} v_{ik} \sigma_k u_{kj}
+%
 \end{equation}
-where $\mathbf V$ and $\mathbf U$ are both unitary matrices and $\mathbf \Sigma$ is a diagonal matrix.
-For our purposes, we associate the rows of $\mathbf U$ with a minimal set of basis templates, which become the kernels of basis filters.
-These filters give rise to the orthogonal \textsc{snr}s, $[\mathbf \rho^\perp]_k = \rho_k^\perp = \mathbf{U}_k \star x^\mathsf{W}$.
-The rows of $\mathbf{V \Sigma}$ become reconstruction coefficients that map linear combinations of the orthogonal \textsc{snr}s from the basis filters back onto \textsc{snr}s for the original templates of interest: $\rho_i = \sum_k [\mathbf V \mathbf \Sigma]_{ik} \rho_k^\perp$.
-
-In many applications of the \textsc{svd}, including ours, the matrix can be well approximated by truncating the summation in equation~(\ref{eq:svd-exact}) at $L \ll N$:
+%
+The cumulative sum of squares of the {\em singular values}, $\sigma_k$,
+measures the Frobenius norm of the approximation, such that
+%
 \begin{equation}
-H_{ij}^\prime = \sum_{k=1}^{L} v_{ik} \sigma_k u_{kj}
+%
+\frac{\| \mathbf H^\prime \|}{\| \mathbf H \|} = \left(\sum_{k=1}^{L}
+|\sigma_k|^2\right)^{1/2} \left(\sum_{k=1}^{N} |\sigma_k|^2\right)^{-1/2}.
+%
 \end{equation}
-The cumulative sum of squares of the {\em singular values}, $\sigma_k$, measures the Frobenius norm of the approximation, such that
-\begin{equation}
-\frac{\| \mathbf H^\prime \|}{\| \mathbf H \|} = \left(\sum_{k=1}^{L} |\sigma_k|^2\right)^{1/2} \left(\sum_{k=1}^{N} |\sigma_k|^2\right)^{-1/2}.
-\end{equation}
-This is also called the \textsc{svd} tolerance.  In our application, it relates to how much \textsc{snr} is lost by discarding $N - L$ of the basis filters with the lowest singular values.
+%
+This is also called the \textsc{svd} tolerance.  In our application, it relates
+to how much \textsc{snr} is lost by discarding $N - L$ of the basis filters
+with the lowest singular values.
 

### 2011-02-17 results.tex
added iir filter bank description
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@638 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{SECIV}\label{sec:results}
+
+We tested the new detection method with mock Advanced \textsc{ligo} having a power spectrum prescribed by the ``zero detuning, high power'' noise model in \cite{Shoemaker:2009p9770}.
+Colored Gaussian noise is generated by passing 5 independent realizations of white Gaussian noise sampled at 16384 Hz through \textsc{iir} filters with coefficients shown in equation~(\ref{eq:iirbank}) below, then summing the filters' outputs.
+This carefully designed filter bank reproduces the noise model very faithfully, but since it is composed of \textsc{iir} filters it can produce mock data in realtime very cheaply.
+
+\begin{eqnarray}
+\label{eq:iirbank}
+x_i &[n] \sim& \mathcal{N}[0, 1] \quad \forall \, i, n \nonumber\\\noalign{\vskip 2mm}
+y_1 &[n] =& \left(4 \times 10^{-28}\right) x_1[n] - y_1[n-1] + 2 \cdot .99995 \cos \left(\frac{2\pi \cdot 9.103}{16384}\right) y_1[n-2] - .99995^2 y_1[n-3] \nonumber\\\noalign{\vskip 2mm}
+y_2 &[n] =& \left(1.1 \times 10^{-23}\right) (x_2[n] - x_2[n-1]) \nonumber\\\noalign{\vskip 2mm}
+y_3 &[n] =& \left(4 10^{-27}\right) x_3[n] - y_3[n-1] + 2 \cdot .999 y_3[n-2] - .999^2 y_3[n-3] \nonumber\\\noalign{\vskip 2mm}
+y_4 &[n] =& \left(4 \times 10^{-26}\right) x_4[n] - y_4[n-1] + 2 \cdot .87 \cos \left(\frac{2\pi \cdot 50}{16384}\right) y_4[n-2] - .87 ^ 2 y_4[n-3] \nonumber\\\noalign{\vskip 2mm}
+y_5 &[n] =& \left(6.5 \times 10^{-24}\right) x_5[n] - y_5[n-1] - .45 \cdot 2 y_5[n-2] - .45 ^ 2 y_5[n-3] \nonumber\\\noalign{\vskip 2mm}
+\; y &[n] =& \frac{3\sqrt{16384}}{4} \left(y_1[n] + y_2[n] + y_3[n] + y_4[n] + y_5[n]\right)
+\end{eqnarray}
 

### 2011-02-18 appendix.tex
Added mock aLIGO noise description
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@639 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
+\section{Filter bank for generating mock Advanced {\sc ligo} strain data}
+\label{appendix:mock-data}
+
+The filter bank described below reproduces the ``zero detuning, high power'' Advanced \textsc{ligo} noise model of  \cite{Shoemaker:2009p9770} very faithfully.
+Since it is composed of a small number of third order or lower linear filters, a digital implementation of it can produce mock data in realtime with negligibly few floating point operations.
+
+First, generate 5 independent streams of white Gaussian noise, $x_1, \dots, x_5$, sampled at 16384 Hz.
+Next, the apply the (\textsc{f}/\textsc{i})\textsc{ir} filters described in equation~(\ref{eq:iirbank}) to generate $y_1, \dots, y_5$ from $x_1, \dots, x_5$ respectively.
+Finally, sum and scale all of the $y_1, \dots, y_5$ together to obtain the final output $y$.
+
+The power spectrum of $y$ is the sum of squares of the magnitudes of the transfer functions of all of the filters.
+The output of the filter bank is compared with the noise model in Figure~\ref{fig:mock-psd} below.
+
+The mock Advanced \textsc{ligo} noise generator is implemented by the GStreamer element \texttt{lal\_fakeadvligosrc}, which is included with the analysis code.
+
+\begin{eqnarray}
+\label{eq:iirbank}
+x_i &[n] \sim& \mathcal{N}[0, 1] \quad \forall \, i, n \nonumber\\\noalign{\vskip 2mm}
+y_1 &[n] =& \left(4 \times 10^{-28}\right) x_1[n] - y_1[n-1] + 2 \cdot .99995 \cos \left(\frac{2\pi \cdot 9.103}{16384}\right) y_1[n-2] - .99995^2 y_1[n-3] \nonumber\\\noalign{\vskip 2mm}
+y_2 &[n] =& \left(1.1 \times 10^{-23}\right) (x_2[n] - x_2[n-1]) \nonumber\\\noalign{\vskip 2mm}
+y_3 &[n] =& \left(10^{-27}\right) x_3[n] - y_3[n-1] + 2 \cdot .999 y_3[n-2] - .999^2 y_3[n-3] \nonumber\\\noalign{\vskip 2mm}
+y_4 &[n] =& \left(4 \times 10^{-26}\right) x_4[n] - y_4[n-1] + 2 \cdot .87 \cos \left(\frac{2\pi \cdot 50}{16384}\right) y_4[n-2] - .87 ^ 2 y_4[n-3] \nonumber\\\noalign{\vskip 2mm}
+y_5 &[n] =& \left(6.5 \times 10^{-24}\right) x_5[n] - y_5[n-1] - 2 \cdot .45 y_5[n-2] - .45 ^ 2 y_5[n-3] \nonumber\\\noalign{\vskip 2mm}
+\; y &[n] =& \frac{3\sqrt{16384}}{4} \left(y_1[n] + y_2[n] + y_3[n] + y_4[n] + y_5[n]\right)
+\end{eqnarray}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[scale=0.4]{mock_psd.pdf}
+\caption{Power spectrum of 100 s of output from mock data filter bank compared with the ``zero detuning, high power'' Advanced \textsc{ligo} noise model.}
+\label{fig:mock-psd}
+\end{center}
+\end{figure}
+
+Note that the \textsc{cpu} overhead of this procedure will be entirely dominated by drawing the pseudorandom numbers $x_1, \dots, x_5$.
+
 \section{Floating point operation counts}

### 2011-02-18 results.tex
Added mock aLIGO noise description
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@639 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-02-08 results.tex
@@
 \label{SECIV}\label{sec:results}
+
+We tested the new detection method with mock Advanced \textsc{ligo} data having a power spectrum prescribed by the ``zero detuning, high power'' noise model in \cite{Shoemaker:2009p9770}.
+Colored Gaussian noise is generated by passing 5 independent realizations of white Gaussian noise sampled at 16384 Hz through a bank of 5 third-order \textsc{iir} filters, then summing the filters' outputs.
+This carefully designed filter bank reproduces the noise model very faithfully, but since it is composed of \textsc{iir} filters it can produce mock data in realtime very cheaply.
+See \ref{appendix:mock-data} for implementation details.
 

### 2011-02-18 results.tex
added note about iligo versus aligo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@642 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added description of injection population
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@643 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{SECIV}\label{sec:results}
+
+We generated receiver operating characteristic (\textsc{roc}) curves for a selection of choices for the composite detection statistic threshold and the \textsc{svd} tolerance.
+For each \textsc{roc} curve, or `run', we analyzed mock data with injections to assess detection efficiencies and mock data without injections to evaluate false alarm rates.
+
+\subsection{Detector noise characteristics}
 
@@
 See \ref{appendix:mock-data} for implementation details.
+\editorial{In this section, we refer to aLIGO run parameters, but the results are actually using initial LIGO noise and initial LIGO parameters.
+This will be fixed.}
+
+\subsection{Injection population}
+
+For each run, we performed about $10^4$ injections into our mock dataset.
+Injections were 2PN post-Newtonian inspiral waveforms with component masses independently and identically distributed in $[1, 3]$ $M_\odot$.
+Injections were tapered \editorial{Check --taper-injection option with Drew.} at a frequency of 10~Hz where they enter the Advanced \textsc{ligo} detection band.
+This dictated that the longest injection waveform had a duration of {\color{red} 1~777~s}.
+In order to guarantee that no two injection waveforms overlapped in the data, the coalescence times of injections were spaced apart by {\color{red} 2~000~$\pm 100$~s}.
+For the sake of economizing disk space, we simultaneously analyzed {\color{red} 16 sets} different sets of about $10^3$ injections apiece in the same {\color{red} 14 day} stretch of mock strain data.
+
+Injections were distributed uniformly in log distance.
+We used the following heuristic to provide a reasonably balanced number of distant, marginally detectable signals, and nearby, easily detected signals:
+$$
+d_\mathrm{min} = \frac{1}{2} \cdot \frac{\rho_\mathrm{candle}}{8.0 \cdot \min {d_\mathrm{eff}}}, \quad
+d_\mathrm{max} = \frac{1}{4} \cdot \frac{\rho_\mathrm{candle}}{5.5 \cdot \max {d_\mathrm{eff}}}.
+$$
+Here, $\rho_\mathrm{candle}$ is a fiducially selected \textsc{snr}, and $d_\mathrm{eff}$ is the distance at which a face-on-system would produce a matched filter \textsc{snr} of $\rho_\mathrm{candle}$.
+
+Orientations of injection sources are drawn uniformly, such that the sky location is distributed uniformly in $4\pi$ and the cosine of the binary inclination angle $\iota$ is distributed uniformly in $[0, \pi]$.
 

### 2011-02-18 results.tex
Fixed typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@644 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In order to guarantee that no two injection waveforms overlapped in the data, the coalescence times of injections were spaced apart by {\color{red} 2~000~$\pm 100$~s}.
-For the sake of economizing disk space, we simultaneously analyzed {\color{red} 16 sets} different sets of about $10^3$ injections apiece in the same {\color{red} 14 day} stretch of mock strain data.
+For the sake of economizing disk space, we simultaneously analyzed {\color{red} 16 statistically independent sets} of about $10^3$ injections apiece in the same {\color{red} 14 day} stretch of mock strain data.
 

### 2011-02-18 results.tex
fixed spacing
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@645 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 This dictated that the longest injection waveform had a duration of {\color{red} 1~777~s}.
-In order to guarantee that no two injection waveforms overlapped in the data, the coalescence times of injections were spaced apart by {\color{red} 2~000~$\pm 100$~s}.
+In order to guarantee that no two injection waveforms overlapped in the data, the coalescence times of injections were spaced apart by {\color{red} 2~000~$\pm$~100~s}.
 For the sake of economizing disk space, we simultaneously analyzed {\color{red} 16 statistically independent sets} of about $10^3$ injections apiece in the same {\color{red} 14 day} stretch of mock strain data.

### 2011-02-18 appendix.tex
iir filter equations flushed full left
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@647 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{eq:iirbank}
-x_i &[n] \sim& \mathcal{N}[0, 1] \quad \forall \, i, n \nonumber\\\noalign{\vskip 2mm}
-y_1 &[n] =& \left(4 \times 10^{-28}\right) x_1[n] - y_1[n-1] + 2 \cdot .99995 \cos \left(\frac{2\pi \cdot 9.103}{16384}\right) y_1[n-2] - .99995^2 y_1[n-3] \nonumber\\\noalign{\vskip 2mm}
-y_2 &[n] =& \left(1.1 \times 10^{-23}\right) (x_2[n] - x_2[n-1]) \nonumber\\\noalign{\vskip 2mm}
-y_3 &[n] =& \left(10^{-27}\right) x_3[n] - y_3[n-1] + 2 \cdot .999 y_3[n-2] - .999^2 y_3[n-3] \nonumber\\\noalign{\vskip 2mm}
-y_4 &[n] =& \left(4 \times 10^{-26}\right) x_4[n] - y_4[n-1] + 2 \cdot .87 \cos \left(\frac{2\pi \cdot 50}{16384}\right) y_4[n-2] - .87 ^ 2 y_4[n-3] \nonumber\\\noalign{\vskip 2mm}
-y_5 &[n] =& \left(6.5 \times 10^{-24}\right) x_5[n] - y_5[n-1] - 2 \cdot .45 y_5[n-2] - .45 ^ 2 y_5[n-3] \nonumber\\\noalign{\vskip 2mm}
-\; y &[n] =& \frac{3\sqrt{16384}}{4} \left(y_1[n] + y_2[n] + y_3[n] + y_4[n] + y_5[n]\right)
+\fl x_i &[n] \sim& \mathcal{N}[0, 1] \quad \forall \, i, n \nonumber\\\noalign{\vskip 2mm}
+\fl y_1 &[n] =& \left(4 \times 10^{-28}\right) x_1[n] - y_1[n-1] + 2 \cdot .99995 \cos \left(\frac{2\pi \cdot 9.103}{16384}\right) y_1[n-2] - .99995^2 y_1[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl y_2 &[n] =& \left(1.1 \times 10^{-23}\right) (x_2[n] - x_2[n-1]) \nonumber\\\noalign{\vskip 2mm}
+\fl y_3 &[n] =& \left(10^{-27}\right) x_3[n] - y_3[n-1] + 2 \cdot .999 y_3[n-2] - .999^2 y_3[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl y_4 &[n] =& \left(4 \times 10^{-26}\right) x_4[n] - y_4[n-1] + 2 \cdot .87 \cos \left(\frac{2\pi \cdot 50}{16384}\right) y_4[n-2] - .87 ^ 2 y_4[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl y_5 &[n] =& \left(6.5 \times 10^{-24}\right) x_5[n] - y_5[n-1] - 2 \cdot .45 y_5[n-2] - .45 ^ 2 y_5[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl \; y &[n] =& \frac{3\sqrt{16384}}{4} \left(y_1[n] + y_2[n] + y_3[n] + y_4[n] + y_5[n]\right)
 \end{eqnarray}

### 2011-02-18 results.tex
Added description of noninjections
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@646 ccdd05d5-56c1-461a-bf98-85f6876523b4
removed some outline stuff that has been converted to prose
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@648 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\begin{itemize}
-\item 1 day of simulated $h(t)$.
-\item $h(t)$ has power spectrum \sout{prescribed for ``zero detuning, high power'' model in \cite{Shoemaker:2009p9770}} that somewhat resembles initial \textsc{ligo} noise models
-\item $h(t)$ generated by passing white, Gaussian noise through a bank of \textsc{iir} filters
-\item 1 noninjections run
-\item 10 injections runs
-\item injections are distributed uniformly in log distance, uniformly in sky location and binary orientation
-\item injections are reweighted to be uniformly distributed in volume
-\item injections are 80$\pm$20 seconds apart
-\item there are $\approx$ 10k injections
-\end{itemize}
+\subsection{Noninjection population}
+
+We completed one injection-free pass over the 14~day dataset. \editorial{And **why**?  This part is too terse.}
+
+\subsection{Trigger reduction procedure}
+
+{\color{red} Describe triggering, clustering, injection finding, injection re-weighting, and \textsc{roc} plotting here.}
 

### 2011-02-18 mock_psd.py
forgot to commit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@649 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -1,0 +1,66 @@
+import sys
+sys.path.append('../lloid_aligo')
+
+import noisemodels
+from scipy import signal
+from numpy import cos, pi, sqrt, random
+import pylab
+
+
+# Filter coefficients
+
+b1 = [4e-28]
+a1 = [1., -2 * .99995 * cos(2*pi * 9.103 / 16384), .99995**2]
+
+b2 = [1.1e-23, -1.1e-23]
+a2 = [1.]
+
+b3 = [1e-27]
+a3 = [1., -2 * .999, .999**2]
+
+b4 = [4e-26]
+a4 = [1., -2 * .87 * cos(2*pi * 50 / 16384), .87**2]
+
+b5 = [6.5e-24]
+a5 = [1, 2*.45, .45**2]
+
+b = [b1, b2, b3, b4, b5]
+a = [a1, a2, a3, a4, a5]
+
+
+# Calculate total filter bank response
+
+h = 0.
+for bb, aa in zip(b, a):
+	w, hi = signal.freqz(bb, aa, 16384)
+	h += abs(hi)**2
+h = sqrt(h) * sqrt(16384) * 3. / 4
+w *= 16384/(2*pi)
+
+
+# Generate some colored Gaussian noise
+
+x = random.randn(len(b), 16384*101)
+y = sum(signal.lfilter(bb, aa, xx) for bb, aa, xx in zip(b, a, x))
+Pxx, freqs = pylab.psd(y[16384:] * sqrt(16384) * 3. / 4, NFFT=16384, Fs=16384, noverlap=4096)
+
+# Read GWINC model
+
+gwinc_f, gwinc_h = dict(noisemodels.asds)['zero det., high power'].T
+
+
+# Plot
+
+pylab.clf()
+pylab.loglog(gwinc_f, gwinc_h, color='0.75', linewidth=10, label='aLIGO noise model')
+pylab.loglog(freqs, sqrt(Pxx), '-r', label='filter bank output')
+pylab.loglog(w, h / sqrt(16384) * sqrt(2), '-k', label='filter transfer func.')
+pylab.xlim(1, 8192)
+pylab.ylim(1e-24, 3e-21)
+pylab.grid(which='minor')
+pylab.grid(which='major')
+pylab.legend()
+pylab.xlabel('frequency (Hz)')
+pylab.ylabel(r'strain amplitude spectral density, $|h(f)|$ $\left(1/\sqrt{\mathrm{Hz}}\right)$')
[…]

### 2011-02-18 method.tex
Deleted commented out text that had been accidentally split onto multiple lines
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@652 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 statistic is the modulus of this, $|\rho(\theta)|$, which has a $\chi^2$
-distribution with 2 degrees of freedom in the absence of signal.%, and it
-indicates the probability of a signal being present
-\cite{wainstein:1962}\editorial{Comment about probability doesn't seem
-relevant.}.
+distribution with 2 degrees of freedom in the absence of signal.
 

### 2011-02-18 analysis.tex
added reference for GStreamer
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@653 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-02-04 analysis.tex
@@
 
-\editorial{Should I mention \texttt{GStreamer} at all?} The detection method described above is much more efficient in terms of floating point operations than the traditional matched filter bank method.
+\editorial{Should I mention GStreamer at all?} The detection method described above is much more efficient in terms of floating point operations than the traditional matched filter bank method.
 However, time slices and conditional reconstruction greatly complicate queueing, synchronizing, and bookkeeping of intermediate signals.
 A low latency implementation capable of recruiting more than one \textsc{cpu} core would be difficult to achieve within the familiar serial programming framework because of the nontrivial time-delay relationships between samples.
-Due to these complications, we chose to prototype the search using an open source signal processing environment called \texttt{GStreamer}.
-\editorial{Citation for \texttt{GStreamer}?}  Primarily used for playing, authoring, or streaming media on \texttt{Linux} systems, \texttt{GStreamer} is an integral component of the popular \texttt{Gnome} desktop.
+Due to these complications, we chose to prototype the search using an open source signal processing environment called GStreamer \cite{gstreamer}.
+Primarily used for playing, authoring, or streaming media on Linux systems, GStreamer is an integral component of the popular Gnome desktop.
 
 \editorial{It would be good to illustrate the layout of this particular template bank: masses spanned, time slice layout \dots}
+\editorial{Drew: I think this should say the FFT version of LLOID is almost 10 times faster than the conventional FFT method.}
 \begin{table}[!h]

### 2011-02-19 results.tex
added description of triggers
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@654 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsection{Trigger reduction procedure}
+\subsection{Triggering and data reduction}
 
-{\color{red} Describe triggering, clustering, injection finding, injection re-weighting, and \textsc{roc} plotting here.}
+For each template with parameters $\theta_i$, a trigger is recorded any time the absolute value of the \textsc{snr}, $|\rho(\theta_i)|$, is greater than a threshold $\rho^* = 5.5$, and is also a maximum over the 10~ms before and after.
+\editorial{This is a proposed trigger criterion that is much simpler than what we currently use.} A trigger is therefore an ordered triple $(t, i, \rho)$, recording the time, template number, and complex \textsc{snr} of an excursion in the outputs of the matched filter bank.
+
+Offline, triggers for all templates collectively are clustered by time, such that within a sliding window of 4 seconds, all triggers except the one with the highest $\rho$ is kept.
+
+For the injection-free passes, all clustered triggers are regarded as false alarms.  We accumulate the \textsc{snr}s of all of the false alarms into the set $\{|\rho|\}_\mathrm{false\,alarm}$.
+
+For each injection, we check whether there was any clustered trigger that was within 2 seconds of the injection's time.
+If there was, then the injection is considered to be recovered at the \textsc{snr} of that trigger.
+Otherwise, the injection was not recovered because the \textsc{snr} threshold $\rho^*$ was not met, so we consider the injection to be recovered only at an \textsc{snr} of $\infty$.
+From all found injections, we accumulate the injected distance and the recovered \textsc{snr} into the set $\{(d, |\rho|)\}_\mathrm{found}$.
 

### 2011-02-19 appendix.tex
added floating point operation count table
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@655 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\editorial{I just yanked this from the methods section.
-It needs to be made more concise, maybe turned into a table of CPU budgets for various components of LLOID.} The filter bank can be implemented using finite impulse response (\textsc{fir}) filters, which are just sliding window dot products.
+Each addition and each multiplication is counted as a single floating point operation.   We are not assuming that a multiply-accumulate is available as a single operation.
+
+\begin{table}[htdp]
+\caption{Number of floating point operations per sample (multiplications and divisions) required for a selection of signal processing operations used in \textsc{lloid}.}
+\begin{center}
+\setlength{\extrarowheight}{6pt}
+\begin{tabular}{l c}
+\hline
+Process & ops/sample \\
+\hline\hline
+\textsc{fir} matched filter, M templates of length $n$ & $2 M n$ \\\hline
+\textsc{fft} matched filter, M templates of length $n$, blocks of length $D$ & $\frac{4 (M + 1) \lg D + 2 M}{1 - n/D}$ \\\hline
+$n$ tap \textsc{fir} resampling filter, sample rates $f_1 < f_2$ & $2 M n f_1 / f_2$ \\\hline
+multiply $M \times L$ real matrix by $L\times1$ real vector & $2 M L$ \\\hline
+\end{tabular}
+\end{center}
+\label{table:flops}
+\end{table}%
+
+The filter bank can be implemented using finite impulse response (\textsc{fir}) filters, which are just sliding window dot products.
 If there are $M$ templates of length $n$, and the data stream contains $N$ samples, then applying the filter bank requires $2 M N n$ operations.

### 2011-02-21 analysis.tex
Implementing my changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@658 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\section{Implementation and Analysis}
+\section{Software Framework}
 \label{SECIII}\label{sec:analysis}
 
-\editorial{Should I mention GStreamer at all?} The detection method described above is much more efficient in terms of floating point operations than the traditional matched filter bank method.
-However, time slices and conditional reconstruction greatly complicate queueing, synchronizing, and bookkeeping of intermediate signals.
-A low latency implementation capable of recruiting more than one \textsc{cpu} core would be difficult to achieve within the familiar serial programming framework because of the nontrivial time-delay relationships between samples.
-Due to these complications, we chose to prototype the search using an open source signal processing environment called GStreamer \cite{gstreamer}.
-Primarily used for playing, authoring, or streaming media on Linux systems, GStreamer is an integral component of the popular Gnome desktop.
-
-\editorial{It would be good to illustrate the layout of this particular template bank: masses spanned, time slice layout \dots}
-\editorial{Drew: I think this should say the FFT version of LLOID is almost 10 times faster than the conventional FFT method.}
-\begin{table}[!h]
-\caption{Operation counts per sample for six different detection methods.
-The operation counts for \textsc{lloid} assume a reconstruction duty cycle of 5\%.
-Note that the \textsc{fir} method with \textsc{lloid} is almost 10 times faster than the conventional \textsc{fft} method, despite having substantially lower latency.}
-\begin{center}
-\input{flop_budget}
-\end{center}
-\end{table}
+The detection method described above is much more efficient in terms of
+floating point operations than the traditional matched filter bank method.
+However, time slices and conditional reconstruction greatly complicate
+queueing, synchronizing, and bookkeeping of intermediate signals.  A low
+latency implementation capable of recruiting more than one \textsc{cpu} core
+would be difficult to achieve within the familiar serial programming framework
+because of the nontrivial time-delay relationships between samples.  Due to
+these complications, we chose to prototype the search using an open source
+signal processing environment called GStreamer \cite{gstreamer}.  Primarily
+used for playing, authoring, or streaming media on Linux systems, GStreamer is
+an integral component of the popular Gnome desktop.
 

### 2011-02-21 appendix.tex
Adding 0's in from of decimal places for readability
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@656 ccdd05d5-56c1-461a-bf98-85f6876523b4
Adding comment on aLIGO noise
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@657 ccdd05d5-56c1-461a-bf98-85f6876523b4
Implementing my changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@658 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\section{Filter bank for generating mock Advanced {\sc ligo} strain data}
-\label{appendix:mock-data}
+\section{Floating point operation counts}
+\label{appendix:cost}
 
-The filter bank described below reproduces the ``zero detuning, high power'' Advanced \textsc{ligo} noise model of  \cite{Shoemaker:2009p9770} very faithfully.
-Since it is composed of a small number of third order or lower linear filters, a digital implementation of it can produce mock data in realtime with negligibly few floating point operations.
+Each addition and each multiplication is counted as a single floating point
+operation. We are not assuming that a multiply-accumulate is available as a
+single operation.
 
-First, generate 5 independent streams of white Gaussian noise, $x_1, \dots, x_5$, sampled at 16384 Hz.
-Next, the apply the (\textsc{f}/\textsc{i})\textsc{ir} filters described in equation~(\ref{eq:iirbank}) to generate $y_1, \dots, y_5$ from $x_1, \dots, x_5$ respectively.
-Finally, sum and scale all of the $y_1, \dots, y_5$ together to obtain the final output $y$.
+\begin{comment}
 
-The power spectrum of $y$ is the sum of squares of the magnitudes of the transfer functions of all of the filters.
-The output of the filter bank is compared with the noise model in Figure~\ref{fig:mock-psd} below.
+% I think that this part can be swalled into the computational costs discussion above.
 
-The mock Advanced \textsc{ligo} noise generator is implemented by the GStreamer element \texttt{lal\_fakeadvligosrc}, which is included with the analysis code.
+\subsection{Comparison of computational costs and latency}
+Convolutions are best done with the use of the Fast Fourier Transform (FFT).
+However there is an inherent latency in performing convolutions this way. This
+section describes the computational cost and latency of FFT convolutions, time 
+domain (TD) convolutions and TD convolutions with SVD.
 
-\begin{eqnarray}
-\label{eq:iirbank}
-\fl x_i &[n] \sim& \mathcal{N}[0, 1] \quad \forall \, i, n \nonumber\\\noalign{\vskip 2mm}
-\fl y_1 &[n] =& \left(4 \times 10^{-28}\right) x_1[n] - y_1[n-1] + 2 \cdot .99995 \cos \left(\frac{2\pi \cdot 9.103}{16384}\right) y_1[n-2] - .99995^2 y_1[n-3] \nonumber\\\noalign{\vskip 2mm}
-\fl y_2 &[n] =& \left(1.1 \times 10^{-23}\right) (x_2[n] - x_2[n-1]) \nonumber\\\noalign{\vskip 2mm}
-\fl y_3 &[n] =& \left(10^{-27}\right) x_3[n] - y_3[n-1] + 2 \cdot .999 y_3[n-2] - .999^2 y_3[n-3] \nonumber\\\noalign{\vskip 2mm}
-\fl y_4 &[n] =& \left(4 \times 10^{-26}\right) x_4[n] - y_4[n-1] + 2 \cdot .87 \cos \left(\frac{2\pi \cdot 50}{16384}\right) y_4[n-2] - .87 ^ 2 y_4[n-3] \nonumber\\\noalign{\vskip 2mm}
-\fl y_5 &[n] =& \left(6.5 \times 10^{-24}\right) x_5[n] - y_5[n-1] - 2 \cdot .45 y_5[n-2] - .45 ^ 2 y_5[n-3] \nonumber\\\noalign{\vskip 2mm}
-\fl \; y &[n] =& \frac{3\sqrt{16384}}{4} \left(y_1[n] + y_2[n] + y_3[n] + y_4[n] + y_5[n]\right)
-\end{eqnarray}
+Convolving one filter with data using FFTs involves 
+$\mathcal{O}[10 N_s \log_2(N_s)]$
+floating point operations where $N_s$ is the number of sample points in
+the data being convolved and results in $N_s - N_f$ samples of usable output.  
+In comparison time domain convolutions require
+$\mathcal{O}[N_s N_f]$ where $N_f$ is the number of sample points in the
+filter.  The FFT is faster to compute when $N_f > 10 \log_2(N_s)$.  However
+the FFT has an inherent latency of $\mathcal{O}[N_s/f_s]$ seconds, where
+$f_s$ is the sample rate when minimally overlapping FFTs are computed. It is
+possible to reduce the latency of the FFT by performing vastly overlapping
+FFTs.  The cost of such a procedure would be 
+$\mathcal{O}[10 N_o N_s \log_2(N_s)]$.  Thus for comparable computational
+cost where $N_f \approx 10 N_o \log_2(N_s)$ the latency of the FFT method 
+would be $\mathcal{O}[N_s/f_s/N_o]$ seconds.  Depending on the task at hand,
+latency requirements and computational costs can dictate what procedure to
+use.
 
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[scale=0.4]{mock_psd.pdf}
-\caption{Power spectrum of 100 s of output from mock data filter bank compared with the ``zero detuning, high power'' Advanced \textsc{ligo} noise model.}
-\label{fig:mock-psd}
-\end{center}
-\end{figure}
-
-Note that the \textsc{cpu} overhead of this procedure will be entirely dominated by drawing the pseudorandom numbers $x_1, \dots, x_5$.
-
-\section{Floating point operation counts}
-
-Each addition and each multiplication is counted as a single floating point operation.   We are not assuming that a multiply-accumulate is available as a single operation.
+\end{comment}
 
@@
 For a $1\,\mathrm{ks}$ template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation requires about about $n / 8 \lg 2 n = 2.2 \times 10^4$ times more operations per sample than the \textsc{fft} implementation.
+
+\section{Filter bank for generating mock Advanced {\sc ligo} strain data}
+\label{appendix:mock-data}
+
+The filter bank described below reproduces the ``zero detuning, high power'' Advanced \textsc{ligo} noise model of  \cite{Shoemaker:2009p9770} very faithfully.
+Since it is composed of a small number of third order or lower linear filters, a digital implementation of it can produce mock data in realtime with negligibly few floating point operations.
+\editorial{dk: We should add the seismic wall to this PSD.
+I know this is waiting on switching to gsteamer 0.10.32 but people will worry about this otherwise.}
+
+First, generate 5 independent streams of white Gaussian noise, $x_1, \dots, x_5$, sampled at 16384 Hz.
+Next, the apply the (\textsc{f}/\textsc{i})\textsc{ir} filters described in equation~(\ref{eq:iirbank}) to generate $y_1, \dots, y_5$ from $x_1, \dots, x_5$ respectively.
+Finally, sum and scale all of the $y_1, \dots, y_5$ together to obtain the final output $y$.
+
+The power spectrum of $y$ is the sum of squares of the magnitudes of the transfer functions of all of the filters.
+The output of the filter bank is compared with the noise model in Figure~\ref{fig:mock-psd} below.
+
+The mock Advanced \textsc{ligo} noise generator is implemented by the GStreamer element \texttt{lal\_fakeadvligosrc}, which is included with the analysis code.
+
+\begin{eqnarray}
+\label{eq:iirbank}
+\fl x_i &[n] \sim& \mathcal{N}[0, 1] \quad \forall \, i, n \nonumber\\\noalign{\vskip 2mm}
+\fl y_1 &[n] =& \left(4 \times 10^{-28}\right) x_1[n] - y_1[n-1] + 2 \cdot 0.99995 \cos \left(\frac{2\pi \cdot 9.103}{16384}\right) y_1[n-2] - 0.99995^2 y_1[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl y_2 &[n] =& \left(1.1 \times 10^{-23}\right) (x_2[n] - x_2[n-1]) \nonumber\\\noalign{\vskip 2mm}
+\fl y_3 &[n] =& \left(10^{-27}\right) x_3[n] - y_3[n-1] + 2 \cdot 0.999 y_3[n-2] - 0.999^2 y_3[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl y_4 &[n] =& \left(4 \times 10^{-26}\right) x_4[n] - y_4[n-1] + 2 \cdot 0.87 \cos \left(\frac{2\pi \cdot 50}{16384}\right) y_4[n-2] - 0.87 ^ 2 y_4[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl y_5 &[n] =& \left(6.5 \times 10^{-24}\right) x_5[n] - y_5[n-1] - 2 \cdot 0.45 y_5[n-2] - 0.45 ^ 2 y_5[n-3] \nonumber\\\noalign{\vskip 2mm}
+\fl \; y &[n] =& \frac{3\sqrt{16384}}{4} \left(y_1[n] + y_2[n] + y_3[n] + y_4[n] + y_5[n]\right)
+\end{eqnarray}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[scale=0.4]{mock_psd.pdf}
+\caption{Power spectrum of 100 s of output from mock data filter bank compared with the ``zero detuning, high power'' Advanced \textsc{ligo} noise model.}
+\label{fig:mock-psd}
+\end{center}
+\end{figure}
+
+Note that the \textsc{cpu} overhead of this procedure will be entirely dominated by drawing the pseudorandom numbers $x_1, \dots, x_5$.
 

### 2011-02-21 method.tex
Implementing my changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@658 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 evaluate a matched filter bank with far lower computational cost than the
-\textsc{fir} method and far less latency than the \textsc{fft} method.
+conventional \textsc{fir} method and far less latency than the conventional
+\textsc{fft} method.
 
@@
 If the template is known to be quasi-bandlimited, then the partial matched
-filter for each time slice may be processed at a reduce sample rate.  Compact
+filter for each time slice may be processed at a reduced sample rate.  Compact
 binary inspiral waveforms fit this description superbly: they are chirps whose
@@
 Our second innovation exploits the fact that the templates in inspiral template
-banks are, by design, highly correlated.  It is possible to greatly reduce the
-number of matched filters required to achieve a particular minimum match by
-designing an appropriate set of orthonormal {\em basis templates}.  A purely
-numerical technique based on the application of the singular value
-decomposition (\textsc{svd}) to inspiral waveforms is demonstrated by the
-authors in~\cite{Cannon:2010p10398}.
+banks are, by design, highly correlated.  This is due to the fact that points
+are chosen in parameter space to be templates such that any point from the
+region parameter space we cover has a match to the nearest template greater
+than our chosen minimum match. It is possible to greatly reduce the number of
+matched filters required to achieve a particular minimum match by designing an
+appropriate set of orthonormal {\em basis templates}.  A purely numerical
+technique based on the application of the singular value decomposition
+(\textsc{svd}) to inspiral waveforms is demonstrated by the authors
+in~\cite{Cannon:2010p10398}.
 
-\editorial{I'm sweeping the complex nature of the templates under the rug here.
-Is it OK to regard our packing of the real and imaginary parts of the
-\textsc{svd} as an implementation detail?  It makes the math a little more
-concise.}  One may regard a bank of $N$ discretely sampled templates of length
-$M$ samples, $h^\mathrm W(\theta_i; t_j) = [\mathbf H]_{ij} = H_{ij}$, as a
-matrix.  The singular value decomposition is an exact factorization that exists
-for any matrix such that
+One may regard a bank of $N/2$ discretely-sampled, complex-valued templates of
+length $M$ samples, $h^\mathrm W(\theta_i; t_j) = [\mathbf H]_{ij} = H_{ij}$,
+as a real matrix of shape $N \times M$, where we have unpacked each
+complex-valued template into two real-valued templates.  The singular value
+decomposition is an exact factorization that exists for any matrix such that
 %
@@
 basis filters.  These filters give rise to the orthogonal \textsc{snr}s,
-$[\mathbf \rho^\perp]_k = \rho_k^\perp = \mathbf{U}_k \star x^\mathsf{W}$.  The
-rows of $\mathbf{V \Sigma}$ become reconstruction coefficients that map linear
-combinations of the orthogonal \textsc{snr}s from the basis filters back onto
-\textsc{snr}s for the original templates of interest: $\rho_i = \sum_k [\mathbf
-V \mathbf \Sigma]_{ik} \rho_k^\perp$.
+$\rho_k^\perp = u_k \star x^\mathsf{W}$.  The rows of $\mathbf{V \Sigma}$
+become reconstruction coefficients that map linear combinations of the
+orthogonal \textsc{snr}s from the basis filters back onto \textsc{snr}s for the
+original templates of interest: $\rho_i = \sum_k v_{ik} \sigma_k \rho_k^\perp$.
 
@@
 well approximated by truncating the summation in equation~(\ref{eq:svd-exact})
-at $L \ll N$:
+at $N' \ll N$:
 %
@@
 %
-H_{ij}^\prime = \sum_{k=1}^{L} v_{ik} \sigma_k u_{kj}
+H_{ij}^\prime = \sum_{k=1}^{N'} v_{ik} \sigma_k u_{kj}
 %
@@
 %
-\frac{\| \mathbf H^\prime \|}{\| \mathbf H \|} = \left(\sum_{k=1}^{L}
+\frac{\| \mathbf H^\prime \|}{\| \mathbf H \|} = \left(\sum_{k=1}^{N'}
 |\sigma_k|^2\right)^{1/2} \left(\sum_{k=1}^{N} |\sigma_k|^2\right)^{-1/2}.
@@
 This is also called the \textsc{svd} tolerance.  In our application, it relates
-to how much \textsc{snr} is lost by discarding $N - L$ of the basis filters
+to how much \textsc{snr} is lost by discarding $N - N'$ of the basis filters
 with the lowest singular values.
 
-\begin{comment}
-
-% Rewritten above.
-
-It is clear that in order to reduce the cost of a low latency search for
-compact binaries of unknown mass we need to reduce the number of convolutions
-necessary to find the gravitational-wave signal.  Rather than filter the
-$i$ templates corresponding to binaries of different component masses it is
-possible to find a basis set of fewer filter $u_j$ that can be linearly combined
-to reproduce $\rho_i$ to some accuracy.  For the remainder of this section
-we will drop the explicit time dependence of the convolution and recognize
-that the snr for any point in time is simply the innerproduct between a
-template waveform vector ${\bf \hat{h}_i}$ and a vector of the detector data 
-${\bf \hat{s}}$, $\rho_i = ({\bf \hat{h}_i}|{\bf \hat{s}})$.  
-The goal is to have
-\begin{equation}
-\label{eq:linear_combination_of_snr}
-\rho_i = \sum_{j=1}^{N'} M_{ij} ({\bf \hat{u}_j}|{\bf \hat{s}})
-\end{equation}
-where the number of inner products is reduced from $N$ to $N'$. In order to
-find the basis set ${\bf u_j}$ we use the singular value decomposition of
-the matrix 
-${H_{ik}} = \{{\bf \hat{h}_1}, {\bf \hat{h}_2}, ...{\bf \hat{h}_N}\}$. Where
-$i$ is the index over the template number and $k$ is the index over sample 
-points. The singular value decomposition factors a matrix such that
-\begin{equation}
-\label{eq:svd}
-{H_{ik}} = {A_{ij}\,B_{jl}\,U_{lk}}
-\end{equation}
-The matrix ${B_{jl}}$ is a diagonal matrix of singular values ranked in
-order of importance of reconstructing the original matrix ${H_{ik}}$.
-However since a search for binary inspirals only needs  waveform accuracy to a
-few percent to be successful it is possible to make an approximate 
-reconstruction of ${H_{ik}}$ 
-\begin{equation}
-\label{eq:svdapprox}
-{H_{ik}} \approx {A_{ij}\,B'_{jl}\,U_{lk}}
-\end{equation}
-where the rank of ${B'_{jl}}$ is less than the rank of ${B_{jl}}$.
-This reduces the number of rows in ${U_{lk}}$.  Now we have a 
-new basis of template waveforms derived from ${U_{lk}} = \{{\bf \hat{u}_1}, {\bf \hat{u}_2}, ...{\bf \hat{u}_{N'}}\}$ and can write 
-(\ref{eq:linear_combination_of_snr})
-\begin{eqnarray}
-\label{eq:svdsnr}
-\rho_i &=& \sum_{j=1}^{N'} A_{il}\,B'_{lj} ({\bf \hat{u}_j}|{\bf \hat{s}}) \nonumber \\
-       &=& \sum_{j=1}^{N'} M_{ij} ({\bf \hat{u}_j}|{\bf \hat{s}})
-\end{eqnarray}
-
-\end{comment}
-
-\editorial{This sounds confrontational to me, but these are good citations.
-Can we tone it down?}  This result differs from other work that models gravitational-wave chirp signals in approximate ways~\cite{Chassande-Mottin2006, Candes2008, BuonannoChenVallisneri:2003a} by starting with an exact representation of the desired template family and producing a rigorous approximation with a tunable accuracy.
+This result differs from other work that models gravitational-wave chirp
+signals in approximate ways~\cite{Chassande-Mottin2006, Candes2008,
+BuonannoChenVallisneri:2003a} by starting with an exact representation of the
+desired template family and producing a rigorous approximation with a tunable
+accuracy.
 
@@
 
-\editorial{We should cite \cite{Scharf:1994p12497} here.} Although the \textsc{svd} allows us to reduce the number of matched filters, this comes at the price of having to perform a matrix multiplication by the reconstruction matrix $\mathbf {V \Sigma}$ at every sample.
-In some circumstances, this matrix multiplication may be more expensive than applying the orthogonal matched filters.
+\editorial{We should cite \cite{Scharf:1994p12497} here.} Although the
+\textsc{svd} allows us to reduce the number of matched filters, this comes at
+the price of having to perform a matrix multiplication by the reconstruction
+matrix $\mathbf {V \Sigma}$ at every sample.  In some circumstances, this
+matrix multiplication may be more expensive than applying the orthogonal
+matched filters.
 
-Further speedup may be gained in a hierarchical detection scheme.
-In general, the orthogonal \textsc{snr}s alone provide some indication of whether any template in the original template bank is likely to have a large \textsc{snr}.
-Consider some composite detection statistic that is a scalar function of the orthogonal \textsc{snr}s, $\Gamma(\rho_1^\perp, \rho_2^\perp, \cdots, \rho_L^\perp)$.
-Suppose that we can infer the distribution of $\Gamma$ for a data stream with the signal present or with the signal absent.
-Then we may use a threshold crossing of $\Gamma$ to trigger the conditional application of the expensive reconstruction matrix only during the times when a signal is likely to be present.
+Further speedup may be gained in a hierarchical detection scheme.  In general,
+the orthogonal \textsc{snr}s alone provide some indication of whether any
+template in the original template bank is likely to have a large \textsc{snr}.
+Consider some composite detection statistic that is a scalar function of the
+orthogonal \textsc{snr}s, $\Gamma(\rho_1^\perp, \rho_2^\perp, \cdots,
+\rho_L^\perp)$.  Suppose that we can infer the distribution of $\Gamma$ for a
+data stream with the signal present or with the signal absent.  Then we may use
+a threshold crossing of $\Gamma$ to trigger the conditional application of the
+expensive reconstruction matrix only during the times when a signal is likely
+to be present.
 
-The composite detection statistic that we employ is a weighted sum of squares, $\Gamma = \sum_{k=1}^L w_k (\rho_k^\perp) ^ 2$, with the particular weights
+Following the ideas in~\cite{Anderson:2000yy}, we employ is a weighted sum of
+squares, $\Gamma = \sum_{k=1}^{N'} w_k (\rho_k^\perp) ^ 2$, with the particular
+weights
+%
 \begin{equation}
-w_k = \frac{\sigma_k^2}{\sigma_k^2 + N / A^2}.
+%
+w_k = \frac{\sigma_k^2}{\sigma_k^2 + N / A^2},
+%
 \end{equation}
-\editorial{We now have a pretty good description of time slices, the SVD, and conditional reconstruction.
-We still need to tie it all together.}Here, $A$ is a desired \textsc{snr} scale that is set by the analyst.
-As the authors show in ~\cite{svd-compdetstat}, this choice of weights has a better receiver operating characteristic for a signal of \textsc{snr}~=~$A$ than some other obvious choices, $w_k = \sigma_k^2$ or $w_k = 1$.
-
-\begin{comment}
-
-% Fold this into SVD discussion.
-
-%\subsection{Detection statistic}
-It is not necessary to reconstruct the SNR of all templates in order to 
-determine if a signal is present.  By thresholding on the sum of squares
-of the orthoganal decomposition described above it is possible to answer
-whether or not an inspiral like event is in the data.  Then only 
-potentially interesting times can be reconstructed on a sample by sample 
-basis.  Following the ideas in~\cite{Anderson:2000yy} we employ a detection
-statistic as
-\begin{equation}
-\label{eq:stat}
-\Gamma^2 = \sum_{l} \sum_{j} {B'_{lj}}^2 {({\bf \hat{u}_j}|{\bf \hat{s}})}^2
-\end{equation}
-
-\end{comment}
-
-\begin{comment}
-
-% We aren't using this chi square test.
-
-\subsection{Signal consistency check}
-After reconstructing the SNR of an individual template it is possible to 
-to check that the components along the SVD basis were consistent with 
-expectation by performing a $\chi^2$ fit to the decomposition.  
-\begin{equation}
-\label{eq:chisq}
-\chi^2_i = \sum_j {[({\bf \hat{u}_j}|{\bf \hat{s}}) - \rho_i / M_{ij}]}^2
-\end{equation}
-
-\end{comment}
-
-\begin{comment}
-
-% I think that this part can be swalled into the computational costs discussion above.
+%
+where $A^2$ is a desired \textsc{snr} scale that is set by the analyst.  As the
+authors show in~\cite{svd-compdetstat}, this choice of weights has a better
+receiver operating characteristic for a signal of \textsc{snr}~=~$A$ than some
+other obvious choices, $w_k = \sigma_k^2$ or $w_k = 1$.
 
 \subsection{Comparison of computational costs and latency}
-Convolutions are best done with the use of the Fast Fourier Transform (FFT).
-However there is an inherent latency in performing convolutions this way. This
-section describes the computational cost and latency of FFT convolutions, time 
-domain (TD) convolutions and TD convolutions with SVD.
 
-Convolving one filter with data using FFTs involves 
-$\mathcal{O}[10 N_s \log_2(N_s)]$
-floating point operations where $N_s$ is the number of sample points in
-the data being convolved and results in $N_s - N_f$ samples of usable output.  
-In comparison time domain convolutions require
-$\mathcal{O}[N_s N_f]$ where $N_f$ is the number of sample points in the
-filter.  The FFT is faster to compute when $N_f > 10 \log_2(N_s)$.  However
-the FFT has an inherent latency of $\mathcal{O}[N_s/f_s]$ seconds, where
-$f_s$ is the sample rate when minimally overlapping FFTs are computed. It is
-possible to reduce the latency of the FFT by performing vastly overlapping
-FFTs.  The cost of such a procedure would be 
-$\mathcal{O}[10 N_o N_s \log_2(N_s)]$.  Thus for comparable computational
-cost where $N_f \approx 10 N_o \log_2(N_s)$ the latency of the FFT method 
-would be $\mathcal{O}[N_s/f_s/N_o]$ seconds.  Depending on the task at hand,
-latency requirements and computational costs can dictate what procedure to
-use.
+Using all of these innovations, we are able to greatly reduce the compuational
+cost of searching for these signals. An example of this can be seen in
+table~\ref{table:comp_cost} where we compute the computational cost of
+filtering 256 templates from a $97\%$ minimal match template bank whitened with
+an (FIXME: is this correct?) Advanced \textsc{ligo} \textsc{asd}. It is
+interesting to note that by themselves, both the \textsc{svd} and time-slice
+innovations \emph{increase} the overall computational cost when combined with
+\textsc{fft} filtering. It is only when combined with conditional
+reconstruction do these methods yield computational savings. Details of the
+computational cost of different operations can be found in \ref{appendix:cost}.
 
-\end{comment}
+\editorial{It would be good to illustrate the layout of this particular
+template bank: masses spanned, time slice layout \dots}
+\begin{table}[!h]
+\caption{Operation counts per sample for six different detection methods.  The
+operation counts for \textsc{lloid} assume a reconstruction duty cycle of 5\%.
+Note that the \textsc{fft} method with \textsc{lloid} is almost 10 times faster
+than the conventional \textsc{fft} method, despite having substantially lower
+latency.}
+\label{table:comp_cost}
+\begin{center}
+\input{flop_budget}
+\end{center}
+\end{table}
 

### 2011-02-21 appendix.tex
Fixing latency of FFT method in appendix
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@659 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{table}[htdp]
-\caption{Number of floating point operations per sample (multiplications and divisions) required for a selection of signal processing operations used in \textsc{lloid}.}
+\caption{Number of floating point operations per sample (multiplications and
+divisions) required for a selection of signal processing operations used in
+\textsc{lloid}.}
 \begin{center}
@@
 
-The filter bank can be implemented using finite impulse response (\textsc{fir}) filters, which are just sliding window dot products.
-If there are $M$ templates of length $n$, and the data stream contains $N$ samples, then applying the filter bank requires $2 M N n$ operations.
+The filter bank can be implemented using finite impulse response (\textsc{fir})
+filters, which are just sliding window dot products.  If there are $M$
+templates of length $n$, and the data stream contains $N$ samples, then
+applying the filter bank requires $2 M N n$ operations.
 
-More commonly, the matched filters are implemented using the \textsc{fft} convolution.
-This entails applying \textsc{fft}s to blocks of $D$ samples, with $2 n \leq D$, each block overlapping the previous one by $n$ samples.
-There are $N/(D-n)$ such blocks.
-Modern implementations of the Cooley-Tukey \textsc{fft}, such as the ubiquitous \texttt{fftw}, require about $4 N \lg N$ operations to evaluate a \textsc{dft} of size $N$~\cite{Johnson:2007p9654}.
-\editorial{This is more commonly known as ``overlap-save''.
-We should find someone else's operation count and cite it.}  A $D$ sample cross-correlation consists of a forward \textsc{fft}, an $D$ sample dot product, and an inverse \textsc{fft}, totaling $8 D \lg N + 2 D$ operations per block.
-Per sample, this is $(8 \lg D + 2) / (1 - n/D)$ operations.
-\editorial{Drew: Why don't we change this to an overlap of $m$ samples so we can see what happens as we increase the overlap to reduce latency.}
+More commonly, the matched filters are implemented using the \textsc{fft}
+convolution.  This entails applying \textsc{fft}s to blocks of $D$ samples,
+with $2 n \leq D$, each block overlapping the previous one by $n$ samples.
+There are $N/(D-n)$ such blocks.  Modern implementations of the Cooley-Tukey
+\textsc{fft}, such as the ubiquitous \texttt{fftw}, require about $4 N \lg N$
+operations to evaluate a \textsc{dft} of size $N$~\cite{Johnson:2007p9654}.
+\editorial{This is more commonly known as ``overlap-save''.  We should find
+someone else's operation count and cite it.}  A $D$ sample cross-correlation
+consists of a forward \textsc{fft}, an $D$ sample dot product, and an inverse
+\textsc{fft}, totaling $8 D \lg N + 2 D$ operations per block.  Per sample,
+this is $(8 \lg D + 2) / (1 - n/D)$ operations. \editorial{Drew: Why don't we
+change this to an overlap of $m$ samples so we can see what happens as we
+increase the overlap to reduce latency.}
 
-The \textsc{fir} filter implementation has the advantage that it has no intrinsic latency, whereas the \textsc{fft} convolution has at least the latency of the \textsc{fft} block size $D \geq 2 n$.
-\editorial{Drew: Should the latency be D-n?} For example, for a $1.4 - 1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the \textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.
-However, the \textsc{fir} filter implementation has the disadvantage of much greater overhead per sample than the \textsc{fft} convolution.
-For a $1\,\mathrm{ks}$ template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation requires about about $n / 8 \lg 2 n = 2.2 \times 10^4$ times more operations per sample than the \textsc{fft} implementation.
+The \textsc{fir} filter implementation has the advantage that it has no
+intrinsic latency, whereas the \textsc{fft} convolution has at least the
+latency of the \textsc{fft} block size $D-n \geq n$. For example, for a $1.4 -
+1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the
+\textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.  However, the
+\textsc{fir} filter implementation has the disadvantage of much greater
+overhead per sample than the \textsc{fft} convolution.  For a $1\,\mathrm{ks}$
+template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation
+requires about about $n / 8 \lg 2 n = 2.2 \times 10^4$ times more operations
+per sample than the \textsc{fft} implementation.
 

### 2011-02-21 method.tex
Fixing latency of FFT method in appendix
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@659 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 algorithm is extremely fast to converge to an accurate PSD and also very robust
-against glitches. This is demonstrated below where we have injected
-band-limited white noise burst with a central frequency of 150 Hz, a bandwidth
-of 100 Hz, a time duration of 0.2 seconds, and a injection density of one
-injection every $100/\pi \approx 32$ seconds. At this density we find there
-should only ever be one injection affecting the median history so the glitches
-should minimally affect PSD estimation as can be seen by comparing the local
-RMS amplitude between injections to that without injections.
-
-\begin{center}
-\resizebox{0.9\textwidth}{!}{\includegraphics{figures/whiten_burst_step_100_semilogy.png}}
-\end{center}
+against glitches.
 
@@
 by
+%
 \begin{equation}
-f(t) = \frac{1}{8 \pi \mathcal{M}} \left( \frac{t_c - t}{5 \mathcal{M}} \right)^{-3/8}\;,
+%
+f(t) = \frac{1}{8 \pi \mathcal{M}} \left( \frac{t_c - t}{5 \mathcal{M}}
+\right)^{-3/8}\;,
+%
 \end{equation}
+%
 whose time derivative is given by
+%
 \begin{equation}
-\frac{df}{dt} = \frac{3}{320 \pi \mathcal{M}^2} \left( \frac{t_c - t}{5 \mathcal{M}} \right)^{-11/8}\;.
+%
+\frac{df}{dt} = \frac{3}{320 \pi \mathcal{M}^2} \left( \frac{t_c - t}{5
+\mathcal{M}} \right)^{-11/8}\;.
+%
 \end{equation}
+%
 Combining these, we find the frequency evolution as a function of frequency to
 be
+%
 \begin{equation}
-\frac{df}{dt}(f_0) = \frac{3}{320 \pi \mathcal{M}^2} \left( 8 \pi \mathcal{M} f_0 \right)^{3/11}\;,
+%
+\frac{df}{dt}(f_0) = \frac{3}{320 \pi \mathcal{M}^2} \left( 8 \pi \mathcal{M}
+f_0 \right)^{3/11}\;,
+%
 \end{equation}
+%
 which can be inverted to get the minimum frequency which has a given frequency
 derivative,
+%
 \begin{equation}
-f_0 = \frac{1}{8 \pi \mathcal{M}} \left( \frac{320 \pi \mathcal{M}^2}{3} \frac{df}{dt} \right)^{11/3}\;.
+%
+f_0 = \frac{1}{8 \pi \mathcal{M}} \left( \frac{320 \pi \mathcal{M}^2}{3}
+\frac{df}{dt} \right)^{11/3}\;.
+%
 \end{equation}

### 2011-02-24 inspiral_svd.tex
removed strange text in abstract
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@624 ccdd05d5-56c1-461a-bf98-85f6876523b4
incremented dcc version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@631 ccdd05d5-56c1-461a-bf98-85f6876523b4
switch to iopart-num bibstyle to get urls
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@640 ccdd05d5-56c1-461a-bf98-85f6876523b4
bumped version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@660 ccdd05d5-56c1-461a-bf98-85f6876523b4
changed titl

@@
 
-\title[LLOID]{A novel method for detecting coalescing binaries in near realtime with Advanced LIGO and beyond \editorial{Submit to PRD or CQG?}}
+\title[LLOID]{Towards realtime detection of compact binary coalescence to
+improve the prospects for multimessenger astronomy\editorial{Submit to PRD or
+CQG?}}
 
@@
 
-\author{Kipp Cannon$^1$, Adrian Chapman$^2$, Nickolas Fotopoulos$^2$, Chad Hanna$^3$, Drew Keppel$^{4,5}$, Antony C. Searle$^2$, Leo Singer$^2$, Alan J. Weinstein$^2$}
+\author{Kipp Cannon$^{1}$, Romain Cariou$^{2}$, Adrian Chapman$^3$, M.
+Crispin-Ortuzar$^4$, Nickolas Fotopoulos$^3$, Melissa Frei $^5$, Chad
+Hanna$^{6}$, Erin Kara$^{7}$, Drew Keppel$^{8,9}$, Laura Liao$^{10}$, Stephen
+Privitera $^3$, Antony C.  Searle$^3$, Leo Singer$^3$, Alan J.  Weinstein$^3$}
 
-\address{$^1$ Canadian Institute for Theoretical Astrophysics, Toronto, ON, Canada}
-\address{$^2$ LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA}
-\address{$^3$ Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
-\address{$^4$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
-\address{$^5$ Leibniz Universit\"{a}t Hannover, Hannover, Germany}
+\address{$^1$ Canadian Institute for Theoretical Astrophysics, Toronto, ON,
+Canada}
+\address{$^2$  Département de physique, École Normale Supérieure de Cachan, 61
+Avenue du Président Wilson, 94235 Cachan Cedex, France}
+\address{$^3$ LIGO Laboratory - California Institute of Technology, Pasadena,
+CA, USA} 
+\address{$^4$ Facultat de Física, Universitat de València, E-46100 Burjassot,
+Spain} 
+\address{$^5$ The University of Texas at Austin, Austin, TX, USA}
+\address{$^6$ Perimeter Institute for Theoretical Physics, Waterloo, ON,
+Canada} 
+\address{$^7$ Department of Physics and Astronomy, Barnard College, Columbia
+University, New York, NY 10027, USA} 
+\address{$^8$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r
+Gravitationphysik, Hannover, Germany} 
+\address{$^9$ Leibniz Universit\"{a}t Hannover, Hannover, Germany}
+\address{$^{10}$ Ryerson University, Toronto, ON, Canada}
 
 \begin{abstract}
-Conventional matched filter bank methods for the detection of gravitational waves from the inspiral of compact binaries are computationally expensive, have hundreds of seconds of unavoidable intrinsic latency, and require arrays of largely redundant matched filters.
-Novel detection methods that are more computationally efficient and have lower latency will be required to realize the full potential of advanced of gravitational wave detectors that are currently under construction.
-In this paper, we describe a new detection method that exploits the properties of inspiral waveforms using multi-rate filtering, principal component analysis, and hierarchical detection.
-We provide receiver operating characteristics from a prototype search pipeline that is capable of low-latency or near-realtime detection with greatly reduced computational requirements in comparison with previously described methods.
-Help me!
+The rapid detection of compact binary coalescence with a network of advanced
+gravitational-wave detectors offers a unique opportunity for multimessenger
+Astronomy.  Prompt alerts to the Astronomical community will assure that all
+available instruments are observing the electromagnetic counterparts to the
+first gravitational waves from coalescing compact objects.   Rapid, near
+realtime alerts, will improve the chances of detecting weak, decaying
+electromagnetic signals and correlating such signals with the gravitational
+wave strain.  In this paper, we describe a new, near realtime, detection
+strategy that exploits the properties of inspiral waveforms using principal
+component analysis, multi-rate filtering, and hierarchical detection.  We
+provide receiver operating characteristics from a prototype search pipeline
+that is capable of near-realtime detection with acheivable computational
+requirements.  
+\editorial{CH: I wasn't happy with the tone of the abstract.  I want to
+focus on the science goal of multimessenger Astronomy.}
 \end{abstract}
@@
 
-\ack{\textsc{ligo} was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the National Science Foundation and operates under cooperative agreement \textsc{phy}-\oldstylenums{0107417}.
-This paper has \textsc{ligo} Document Number \textsc{ligo}-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{1}.
+\ack{\textsc{ligo} was constructed by the California Institute of Technology
+and Massachusetts Institute of Technology with funding from the National
+Science Foundation and operates under cooperative agreement
+\textsc{phy}-\oldstylenums{0107417}. This paper has \textsc{ligo} Document
+Number \textsc{ligo}-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{3}.}
 
@@
 
-\bibliographystyle{unsrt}
+\bibliographystyle{iopart-num}
 \bibliography{references}

### 2011-02-24 introduction.tex
changed intro
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@663 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-02-15 introduction.tex
@@
 
-Low-mass inspiraling binaries are the most promising sources of gravitational
-radiation for Advanced \textsc{ligo}. Observation demonstrates that they must
-exist in the local Universe in some abundance~\cite{Abadie:2010p10836}.
-[science] In the
-final stages of inspiral, the neutron star is tidally disrupted, providing the
-matter required to produce bright electromagnetic
-counterparts~\cite{shibata:2007}. Conveniently, the gravitational waveforms are
-understood and implemented to the level required by Advanced \textsc{ligo} to
-detect~\cite{BuonannoIyerOchsnerYiSathya2009}, allowing the sensitive technique
-of matched filtering.
+The coalescence of compact binary systems consisting of neutron stars and/or
+black holes is the most promising source of gravitational radiation for
+ground-based gravitational wave detectors such as Advanced
+\textsc{ligo}~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, GEO~\cite{GEOWeb}, and
+LCGT~\cite{LCGTWeb}. Radio observation of binary pulsar systems in the Galaxy
+show that such systems are losing energy according to the prediction of
+gravitational wave emission from general
+relativity~\cite{Taylor:1982}\citeneeded.  The binary periods are decreasing as
+they lose energy to gravitational waves and subsequently the separation of the
+component neutron stars is also decreasing. As this occurs, the gravitational
+wave amplitude increases causing a run-away reaction that will eventually lead
+to the merger of the pulsars within the age of our Universe.  One can
+extrapolate the expected coalescence rate from the electromagnetically observed
+systems.  Although this procedure has large uncertainty, tens of binary
+coalescence events are realistically expected to be observed in the advanced
+detector era later this decade~\cite{Abadie:2010p10836}.
 
-The parameter space of compact binary coalescence signals is
-large~\cite{Owen:1995tm, Owen:1998dk} leading to large filter banks and
-significant computational cost. Following the series of detector upgrades that
-are now under way, \textsc{ligo} and Virgo will attain their Advanced
-configurations, gaining a tenfold improvement in amplitude sensitivity over
-Initial configurations~\cite{advLIGOrefdesign,advVirgorefdesign} and a
-corresponding thousandfold increase in observable volume in the local Universe.
-The computational requirements of conventional detection strategies will be
-greatly increased by this upgrade for two reasons. First, advanced detectors
-will have much improved low-frequency sensitivity, so chirp signals from
-coalescing binaries will remain in the detectors' band for a significantly
-longer time. This will demand that matched filter based searches use longer
-templates, requiring more memory and more cycles to process. Second, background
-noise for advanced detectors will be more spectrally uniform, so it will be
-possible to resolve the intrinsic parameters of a source that determine the
-time-frequency structure of the gravitational wave signal with much greater
-accuracy. This will require that matched filter bank based searches employ
-more matched filters, again demanding more memory and more cycles.
+Only the final tens of minutes of the merger will produce a signal with
+sufficient amplitude that it will be observable with Advanced LIGO\citeneeded.
+Furthermore, due to the frequency response of Advanced ground based
+gravitational wave detectors the last tens of seconds produce the majority of
+the signal-to-noise ratio (SNR).  The gravitational-wave frequency must be
+above $\sim$10Hz  before any appreciable signal-to-noise ratio can be
+accumulated\footnote{Which implies an orbital frequency of 5 Hz assuming that
+the dominant gravitational wave frequency is twice the orbital frequency} and
+most of the SNR is accumulated near the minimum of the detector noise spectral
+density $\sim100$Hz.  The gravitational wave signal will increase in amplitude
+and frequency (i.e. it will chirp) until it reaches the innermost stable
+circular orbit (ISCO), which is $\approx 4.4\times
+10^3\,$Hz$\,\times(M/\Msun)^{-1}$ for a test particle orbiting a Schwarzschild
+black hole with mass $M$. After it reaches the ISCO the binary will merge, and
+if it forms a black hole will emit gravitational radiation according to the
+fundamental modes of the resulting Kerr black hole (which are exponentially
+decreasing in time)\citeneeded.  This entire process is sufficiently well
+modeled that matched filtering can be used to detect these systems by
+establishing a grid of filters in the binary parameter space in a way that
+assures minimal loss in signal-to-noise ratio for any binary signal within the
+parameter space boundaries\citeneeded.
 
-Beyond raw throughput, the ability to detect signals in near realtime will
-become increasingly valuable as the gravitational-wave detection horizon pushes
-outward.  Having an electromagnetic or neutrino counterpart to a
-gravitational-wave detection would not only increase the confidence in the
-detection but will also greatly increase the astrophysical information
-available from the event. Most models that predict simultaneous
-gravitational-wave and electromagnetic observations also predict that the peak
-amplitude of electromagetic radiation will occur soon after gravitational-wave
-emission~\cite{sylvestre2003}. Thus in order to maximize the chance of a
-successful electromagnetic followup the latency of gravitational-wave signal
-analysis must be made minimal.
+As the binary approaches the merging point, the neutron stars are tidally
+disrupted.  This may provide the matter required to produce bright
+electromagnetic counterparts in the moments before and during the
+merger~\cite{shibata:2007}.   \hl{List more mechanisms!! such as plasma and
+magnetic fields...}.  In order to observe the prompt and potentially brightest
+electromagnetic emission, telescopes will have to be pointing in the direction
+of the event during the seconds surrounding the merger.  In order to assure
+that this occurs more often than by random chance from various astronomical
+surveys operating independently, the gravitational-wave community has initiated
+a project to send alerts when potential gravitational wave transients are
+observed.  Unfortunately, at best, gravitational wave experiments can only hope
+to have accumulated sufficient SNR within seconds (or less) prior to the merger.
+This makes it challenging to alert telescopes promptly to catch as much of
+the electromagnetic signal as possible.
 
-In \oldstylenums{2010}, \textsc{ligo} and Virgo completed S6/VSR3,
-a short period of
-joint data taking during which several all-sky detection pipelines operated in
-a low-latency configuration \citeneeded. A few candidates of moderate
-significance were promptly sent for electromagnetic followup \citeneeded{} to
-several telescopes including Swift, \textsc{rotse}, \textsc{tarot}, and Zadko.
-They achieved latencies of BLAH, but required human vetting of each candidate
-before alerting the telescope partners. As confidence in the infrastructure
-increases, humans can be removed from the loop, setting a demonstrated previous
-best latency at BLAH.
+%
+\editorial{If we publish before the LUMIN papers, we have to redo this section.
+We can't make claims about those private communications with the telescopes.
+Else, cite the pre-S6 LUMIN paper \cite{kanner2008}. CH, the info is public,
+Brennan presented it at GWPAW, fair game then.  I have cited the presentation
+for now and spoke with Brennan about the upcoming paper (also cited as in
+preparation)} 
+%
+In \oldstylenums{2010}, LIGO completed its sixth science run (S6) and Virgo
+completed its third science run (VSR3).   When both LIGO and Virgo were
+operating, several all-sky detection pipelines operated in a low-latency
+configuration~\cite{HugheyGWPAW2011, S6lowlatency}. A few candidates were sent
+for electromagnetic followup to several telescopes including Swift,
+\textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008, HugheyGWPAW2011}.
+The S6 analyses achieved latencies of $~\sim$30 minutes, which were dominated by
+a human vetting process.  The intrinsic gravitational-wave trigger generator
+latency was $\sim$2--5 minutes.  In this work we will assume that the human
+vetting process can be automated in the advanced detector era, and that the
+dominating latency will be trigger generation.  From this perspective we will
+try to address the realistic expectations for reducing trigger latency from
+$\sim$minutes to $\lesssim$seconds.
 
-This work will describe how to exploit degeneracy in the signal parameter space
-to answer more quickly and economically whether or not a gravitational wave is
-present.
-\editorial{nvf: I use some rather different words here to summarize the
-techniques. I don't want to simply repeat words from later, but I don't want
-to jar a reader with changing jargon either. Feel free to edit.}
-Specifically, we will exploit the time-frequency structure of chirp
-signals to downsample different parts of the waveform, use the singular value
-decomposition (\textsc{svd}) factors to identify the redundancy of the bank and
-reduce the effective number of filters required to search the data, and use
-matched subspace filters built from \textsc{svd} to identify when computation
-can be avoided.  We note that others have applied the use of \textsc{svd} to
-gravitational wave data analysis to analyze optimal gravitational-wave burst
-detection~\cite{bradyraymajumder2004, heng2008} and coherent networks of
-detectors~\cite{wen2008}.
+The reduction in latency is a challenging task.  As one pushes typical
+finite-impulse-response (FIR) filtering techniques from high latency
+configurations with a computational complexity that scales as
+$\mathcal{O}(N\log{N})$, where $N$ is the number of sample points in the
+calculation, to low-latency configuration, the complexity goes to
+$\mathcal{O}(N^2)$.  Naively this would require orders of magnitude more
+computational resources.  However, there are aspects of searches for binary
+coalescence that can be exploited to offset some of the difference in
+computational complexity.  1) The time frequency properties of the binary
+waveforms are such that multirate filtering may be employed. 2) The redundancy
+in the filter banks used to cover a particular mass parameter space may be
+reduced via rank reduction schemes such as singular value
+decomposition~\cite{Cannon:2010p10398}.  We will use both aspects to
+demonstrate that a very low latency analysis of gravitational-wave data for
+compact binary sources is possible with the current computing resources
+available.  Assuming other aspects of gravitational-wave observation latency
+can be reduced significantly, this should allow the possibility for alerts to
+be sent within seconds of the binary merger.
 

### 2011-02-24 introduction.tex
lloid_paper intro: A polishing pass
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@666 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 black holes is the most promising source of gravitational radiation for
-ground-based gravitational wave detectors such as Advanced
+ground-based gravitational-wave detectors such as Advanced
 \textsc{ligo}~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, GEO~\cite{GEOWeb}, and
-LCGT~\cite{LCGTWeb}. Radio observation of binary pulsar systems in the Galaxy
-show that such systems are losing energy according to the prediction of
-gravitational wave emission from general
-relativity~\cite{Taylor:1982}\citeneeded.  The binary periods are decreasing as
-they lose energy to gravitational waves and subsequently the separation of the
-component neutron stars is also decreasing. As this occurs, the gravitational
-wave amplitude increases causing a run-away reaction that will eventually lead
-to the merger of the pulsars within the age of our Universe.  One can
-extrapolate the expected coalescence rate from the electromagnetically observed
-systems.  Although this procedure has large uncertainty, tens of binary
-coalescence events are realistically expected to be observed in the advanced
-detector era later this decade~\cite{Abadie:2010p10836}.
+LCGT~\cite{LCGTWeb}.
+A detection would strongly constrain mass and
+spin~\cite{finn1993,Poisson:1995ef}, and with sufficient signal strength, the
+neutron star equation of state~\cite{flanagan:021502,Read:2009}, general
+relativity in the strong-field regime~\cite{Will:2005va}, the graviton
+mass~\cite{PhysRevD.80.044002,Keppel:2010qu}, and the Lorentz Invariance
+principle~\cite{Ellis2006402}.
+Radio observation of binary pulsar systems in the Galaxy show the existence of
+systems losing energy according to the prediction of gravitational-wave
+emission from general relativity~\cite{Taylor:1982,Abadie:2010p10836}. As
+energy is lost to gravitational waves, a binary's orbital separation decreases,
+as does its orbital period.
+As this occurs, the gravitational-wave amplitude increases, causing
+a run-away inspiral that eventually leads to the merger of the compact objects.
+One can estimate the expected coalescence rate by extrapolating the orbital
+periods of systems observed electromagnetically. Although this procedure has
+large uncertainty, tens of binary coalescence events are realistically expected
+to be observed in the advanced detector era later this
+decade~\cite{Abadie:2010p10836}.
 
-Only the final tens of minutes of the merger will produce a signal with
-sufficient amplitude that it will be observable with Advanced LIGO\citeneeded.
-Furthermore, due to the frequency response of Advanced ground based
-gravitational wave detectors the last tens of seconds produce the majority of
-the signal-to-noise ratio (SNR).  The gravitational-wave frequency must be
-above $\sim$10Hz  before any appreciable signal-to-noise ratio can be
-accumulated\footnote{Which implies an orbital frequency of 5 Hz assuming that
-the dominant gravitational wave frequency is twice the orbital frequency} and
-most of the SNR is accumulated near the minimum of the detector noise spectral
-density $\sim100$Hz.  The gravitational wave signal will increase in amplitude
-and frequency (i.e. it will chirp) until it reaches the innermost stable
-circular orbit (ISCO), which is $\approx 4.4\times
-10^3\,$Hz$\,\times(M/\Msun)^{-1}$ for a test particle orbiting a Schwarzschild
-black hole with mass $M$. After it reaches the ISCO the binary will merge, and
-if it forms a black hole will emit gravitational radiation according to the
-fundamental modes of the resulting Kerr black hole (which are exponentially
-decreasing in time)\citeneeded.  This entire process is sufficiently well
-modeled that matched filtering can be used to detect these systems by
-establishing a grid of filters in the binary parameter space in a way that
-assures minimal loss in signal-to-noise ratio for any binary signal within the
-parameter space boundaries\citeneeded.
+The gravitational-wave inspiral signal starts at low frequencies, gradually
+increasing in amplitude and frequency (i.e. it will chirp) until it reaches the
+innermost stable circular orbit (\textsc{isco}) ($f_\mathrm{ISCO} \approx
+4.4\times 10^3\,$Hz$\,\times(M/\Msun)^{-1}$) \editorial{Is $f_\mathrm{ISCO}$
+the orbital frequency or GW frequency?} for a test particle orbiting a
+Schwarzschild black hole with the total mass of the system $M$.  The frequency
+response of Advanced ground-based gravitational-wave detectors is such that the
+gravitational-wave frequency must be above $\sim 10$Hz before any appreciable
+signal-to-noise ratio (\textsc{snr}) can be accumulated\footnote{Which implies
+an orbital frequency of 5 Hz assuming that the dominant gravitational-wave
+frequency is twice the orbital frequency}.  Most of the \textsc{snr} is
+accumulated near the minimum of the detector noise spectral density at $\sim
+100$Hz. The inspiral waveform of a binary neutron star system will spend $\sim
+10^3$s in band.  Conveniently, the gravitational waveforms of compact binary
+coalescence are understood to the level required by Advanced \textsc{ligo} to
+detect~\cite{BuonannoIyerOchsnerYiSathya2009} via matched filtering techniques.
+Matched filtering detection pipelines maximize signal-to-noise ratio over a
+bank of filters, which are template waveforms in the binary parameter space.
+The templates are spaced to bound the loss in signal-to-noise ratio for any
+binary signal within the parameter space
+boundaries~\cite{Owen:1995tm,Owen:1998dk}.
 
-As the binary approaches the merging point, the neutron stars are tidally
-disrupted.  This may provide the matter required to produce bright
-electromagnetic counterparts in the moments before and during the
-merger~\cite{shibata:2007}.   \hl{List more mechanisms!! such as plasma and
-magnetic fields...}.  In order to observe the prompt and potentially brightest
+As a binary containing a neutron star with a low-mass companion (a neutron star
+or black hole with mass less than $10\Msun$) approaches the merging point, the
+neutron star or neutron stars are tidally disrupted. This can provide
+the matter required to produce bright electromagnetic counterparts in the
+moments following the merger~\cite{shibata:2007}.
+\editorial{nvf: Chad thinks we need more mechanisms. I think this stands on
+its own without the acceleration mechanism.}
+In order to observe the prompt and potentially brightest
 electromagnetic emission, telescopes will have to be pointing in the direction
@@
 surveys operating independently, the gravitational-wave community has initiated
-a project to send alerts when potential gravitational wave transients are
-observed.  Unfortunately, at best, gravitational wave experiments can only hope
-to have accumulated sufficient SNR within seconds (or less) prior to the merger.
-This makes it challenging to alert telescopes promptly to catch as much of
-the electromagnetic signal as possible.
+a project to send alerts when potential gravitational-wave transients are
+observed. Unfortunately, at best, gravitational-wave experiments can only hope
+to have accumulated sufficient \textsc{snr} within seconds (or less) prior to
+the merger.  This makes it challenging to alert telescopes promptly to catch as
+much of the electromagnetic signal as possible.
 
-%
-\editorial{If we publish before the LUMIN papers, we have to redo this section.
-We can't make claims about those private communications with the telescopes.
-Else, cite the pre-S6 LUMIN paper \cite{kanner2008}. CH, the info is public,
-Brennan presented it at GWPAW, fair game then.  I have cited the presentation
-for now and spoke with Brennan about the upcoming paper (also cited as in
-preparation)} 
-%
-In \oldstylenums{2010}, LIGO completed its sixth science run (S6) and Virgo
-completed its third science run (VSR3).   When both LIGO and Virgo were
+In 2010, \textsc{ligo} completed its sixth science run (S6) and Virgo
+completed its third science run (VSR3). When both \textsc{ligo} and Virgo were
 operating, several all-sky detection pipelines operated in a low-latency
-configuration~\cite{HugheyGWPAW2011, S6lowlatency}. A few candidates were sent
+configuration~\cite{HugheyGWPAW2011, S6lowlatency}. Several candidates were sent
 for electromagnetic followup to several telescopes including Swift,
 \textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008, HugheyGWPAW2011}.
-The S6 analyses achieved latencies of $~\sim$30 minutes, which were dominated by
+The S6 analyses achieved latencies of $~\sim30$--$60$ minutes, which were
+dominated by
 a human vetting process. The intrinsic gravitational-wave trigger generator
 latency was $\sim$2--5 minutes. In this work we will assume that the human
-vetting process can be automated in the advanced detector era, and that the
+vetting process can be automated in the advanced detector era and that the
 dominating latency will be trigger generation. From this perspective we will
@@
 The reduction in latency is a challenging task.  As one pushes typical
-finite-impulse-response (FIR) filtering techniques from high latency
+finite-impulse-response (\textsc{fir}) filtering techniques from high latency
 configurations with a computational complexity that scales as
@@
 calculation, to low-latency configuration, the complexity goes to
-$\mathcal{O}(N^2)$.  Naively this would require orders of magnitude more
+$\mathcal{O}(N^2)$. Na\"ively this would require orders of magnitude more
 computational resources. However, there are aspects of searches for binary
 coalescence that can be exploited to offset some of the difference in
-computational complexity.  1) The time frequency properties of the binary
-waveforms are such that multirate filtering may be employed. 2) The redundancy
-in the filter banks used to cover a particular mass parameter space may be
-reduced via rank reduction schemes such as singular value
+computational complexity. 1) The time-frequency morphology of the binary
+waveforms is such that multi-rate filtering may be employed. 2) The redundancy
+in the filter banks used to cover a particular parameter space may be
+reduced via rank-reduction schemes such as the singular value
 decomposition~\cite{Cannon:2010p10398}. We will use both aspects to
 demonstrate that a very low latency analysis of gravitational-wave data for
-compact binary sources is possible with the current computing resources
-available.  Assuming other aspects of gravitational-wave observation latency
+compact binary sources is possible with the current computing resources.
+Assuming other aspects of gravitational-wave observation latency
 can be reduced significantly, this should allow the possibility for alerts to
@@
 standard method for detecting compact binary coalescence signals and
-describe how it can be modified to accomodate low latency analysis.  We
-then describe the pipeline we have constructed to implement these changes.
-To validate the approach we present results of simulations and finish
-with some concluding remarks.
+describe how it can be modified to accomodate low-latency analysis. We
+then describe the pipeline we have constructed that implements these changes.
+To validate the approach we present results of simulations and conclude
+with some remarks on what remains to prepare for the Advanced detector era.

### 2011-03-01 results.tex
results: double checked max duration, so un-highlighted it
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@669 ccdd05d5-56c1-461a-bf98-85f6876523b4
results: took out editorial about --taper-injections
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@670 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Injections were 2PN post-Newtonian inspiral waveforms with component masses independently and identically distributed in $[1, 3]$ $M_\odot$.
-Injections were tapered \editorial{Check --taper-injection option with Drew.} at a frequency of 10~Hz where they enter the Advanced \textsc{ligo} detection band.
-This dictated that the longest injection waveform had a duration of {\color{red} 1~777~s}.
+Injections were tapered at a frequency of 10~Hz where they enter the Advanced \textsc{ligo} detection band.
+This dictated that the longest injection waveform had a duration of 1~777~s.
 In order to guarantee that no two injection waveforms overlapped in the data, the coalescence times of injections were spaced apart by {\color{red} 2~000~$\pm$~100~s}.

### 2011-03-31 snr_in_time.py
accumulated SNR plot
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@700 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -1,0 +1,125 @@
+#!/usr/bin/python
+import numpy
+import matplotlib
+matplotlib.use('Agg')
+import pylab
+import scipy
+from scipy import interpolate
+import sys
+
+#
+# default plot paramaters
+#
+
+matplotlib.rcParams.update({
+	"font.size": 9.0,
+	"axes.titlesize": 9.0,
+	"axes.labelsize": 9.0,
+	"xtick.labelsize": 9.0,
+	"ytick.labelsize": 9.0,
+	"legend.fontsize": 9.0,
+	"figure.figsize": (3,3),
+	"figure.dpi": 600,
+	"subplots.left": 0.05,
+	"subplots.right": 0.95,
+	"subplots.bottom": 0.05,
+	"subplots.top": 0.95,
+	"savefig.dpi": 600,
+	"text.usetex": True     # render all text with TeX
+})
+
+#
+# Constants
+#
+
+LAL_MTSUN_SI = 4.9254909500000001e-06
+PI = 3.14159265
+
+#
+# Mchirp
+#
+
+def mchirp(m1, m2):
+        return (m1 * m2)**.6 / (m1+m2)**.2
+
+#
+# cumulative fractional snr computed over f
+#
+
+def snr(f, asd):
+	df = f[1] - f[0]
+	out = numpy.cumsum(df * f**(-7./3) / asd(f)**2)**.5
+	return out# / out[-1]
+
+#
+# frequency to time relationship assuming 0PN 
+#
+
+def freq_to_time(Mc,f):
+	Mc = Mc * LAL_MTSUN_SI
+	return 5. * Mc / 256. * (PI * Mc * f) ** (-8./3.)
+
+#
+# the number of events above fractional snr of SNR assuming number detectable
+#
+
+def snr_to_num(fracsnr, number=40):
+	return number * (fracsnr)**3
+
+# Start a new figure
+fig = pylab.figure(1, figsize=(5.5,2.25))
+
+markers = ['k-', 'k--', 'k-.']
+
+snrd = {}
+
+# loop over the noise curves
+# ZERO_DET_high_P.txt must be first since it normalizes the rate
+for i, cfile in enumerate(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt', 'data/BHBH_20deg.txt']):
+
+	# Load the data
+	A = numpy.loadtxt(cfile)
+
+	# interpolate the asd
+	asd = interpolate.interp1d(A[:,0], A[:,1])
+
[…]

### 2011-03-31 time_slices.py
polished up figure and table formatting
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@705 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -6,9 +6,25 @@ from optparse import OptionParser, Option
 from gstlal.templates import time_slices
 from pylal import spawaveform
 import numpy as np
+import matplotlib
 import pylab
 from itertools import groupby, izip
-
+matplotlib.rcParams.update({
+        "font.size": 9.0,
+        "axes.titlesize": 9.0,
+        "axes.labelsize": 9.0,
+        "xtick.labelsize": 9.0,
+        "ytick.labelsize": 9.0,
+        "legend.fontsize": 8.0,
+        "figure.figsize": (3,3),
+        "figure.dpi": 600,
+        "subplots.left": 0.1,
+        "subplots.right": 0.75,
+        "subplots.bottom": 0.25,
+        "subplots.top": 0.75,
+        "savefig.dpi": 600,
+        "text.usetex": True     # render all text with TeX
+})
 
 # Command line interface
 
@@ -33,6 +49,8 @@ slices = time_slices(mass_pairs, opts.flow, fhigh)
 mc = spawaveform.chirpmass(opts.mass1, opts.mass2)
 
 # Generate plot
+pylab.figure()
+ax = pylab.subplot(111)
 tmin = min(slice['begin'] for slice in slices)
 tmax = max(slice['end'] for slice in slices)
 legend_artists = []
@@ -51,24 +69,24 @@ pylab.yticks([], [])
 pylab.legend(legend_artists, legend_labels, loc = 'lower left')
 pylab.xlim(-tmax, 0.)
 pylab.ylim(-.15, .15)
-pylab.xlabel('time relative to coalescence')
+pylab.xlabel('time relative to coalescence (s)')
 pylab.ylabel(r'gravitational wave strain amplitude')
-#pylab.title(r'Time slices for a %g $\emdash$ %g M$_\odot$ inspiral' % (opts.mass1, opts.mass2))
+matplotlib.pyplot.subplots_adjust(left=0.06, right=0.88, top=0.95, bottom=0.12)
+
 pylab.savefig('time_slices.pdf')
 
 
 # Generate output table
 
-print r"\begin{tabular}{rr@{, }ll}"
-print r"\hline\hline"
-print r"$f_k$ (Hz) & ($t_{k+1}$ & $t_k$] (s) & Samples \\"
+print r"\begin{tabular}{|lrr}"
+#FIXME change the symbol for the number of sample points per slice if the macro changes
[…]

### 2011-03-31 inspiral_svd.tex
more focus on multimessenger astronomy and early warning, added acks
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@706 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \input{macros}
+\renewcommand{\arraystretch}{1.2}
 
@@
 
-\title[LLOID]{Towards realtime detection of compact binary coalescence to
-improve the prospects for multimessenger astronomy\editorial{Submit to PRD or
-CQG?}}
+\title[LLOID]{Improving the prospects for multimessenger astronomy with
+\earlywarning\ detection of compact binary coalescence.}
 
@@
 electromagnetic signals and correlating such signals with the gravitational
-wave strain.  In this paper, we describe a new, near realtime, detection
-strategy that exploits the properties of inspiral waveforms using principal
-component analysis, multi-rate filtering, and hierarchical detection.  We
-provide receiver operating characteristics from a prototype search pipeline
-that is capable of near-realtime detection with acheivable computational
-requirements.  
-\editorial{CH: I wasn't happy with the tone of the abstract.  I want to
-focus on the science goal of multimessenger Astronomy.}
+wave strain.  In this paper, we describe a new, realtime, detection strategy
+that exploits the properties of inspiral waveforms using principal component
+analysis and multi-rate filtering.  We provide detection efficiency
+characteristics from a prototype search pipeline that is capable of
+\earlywarning\ detection with acheivable computational requirements.  
 \end{abstract}
@@
 
-\ack{\textsc{ligo} was constructed by the California Institute of Technology
+\ack{\LIGO\ was constructed by the California Institute of Technology
 and Massachusetts Institute of Technology with funding from the National
 Science Foundation and operates under cooperative agreement
-\textsc{phy}-\oldstylenums{0107417}. This paper has \textsc{ligo} Document
-Number \textsc{ligo}-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{3}.}
+\textsc{phy}-\oldstylenums{0107417}. CH would like to thank Ilya Mandel
+for many discussions about rate estimates and the prospects of early
+detection.
 
-\appendix
-\input{appendix}
+This paper has \LIGO\ Document
+Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{3}.}
+
+%\appendix
+%\input{appendix}
 

### 2011-03-31 introduction.tex
reworking to emphasize early warning
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@707 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The coalescence of compact binary systems consisting of neutron stars and/or
-black holes is the most promising source of gravitational radiation for
-ground-based gravitational-wave detectors such as Advanced
-\textsc{ligo}~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, GEO~\cite{GEOWeb}, and
-LCGT~\cite{LCGTWeb}.
-A detection would strongly constrain mass and
-spin~\cite{finn1993,Poisson:1995ef}, and with sufficient signal strength, the
-neutron star equation of state~\cite{flanagan:021502,Read:2009}, general
-relativity in the strong-field regime~\cite{Will:2005va}, the graviton
+%
+The coalescence of compact binary systems consisting of neutron stars (NS)
+and/or black holes (BH) is the most promising source of gravitational radiation
+for ground-based gravitational-wave detectors such as Advanced
+\LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb}, and
+\LCGT~\cite{LCGTWeb}.  Detections will constrain the mass and spin of compact
+objects~\cite{finn1993,Poisson:1995ef}, the neutron star equation of
+state~\cite{flanagan:021502,Read:2009}, general relativity in the strong-field
+regime~\cite{Will:2005va}, the graviton
 mass~\cite{PhysRevD.80.044002,Keppel:2010qu}, and the Lorentz Invariance
 principle~\cite{Ellis2006402}.
-Radio observation of binary pulsar systems in the Galaxy show the existence of
-systems losing energy according to the prediction of gravitational-wave
-emission from general relativity~\cite{Taylor:1982,Abadie:2010p10836}. As
-energy is lost to gravitational waves, a binary's orbital separation decreases,
-as does its orbital period.
-As this occurs, the gravitational-wave amplitude increases, causing
-a run-away inspiral that eventually leads to the merger of the compact objects.
-One can estimate the expected coalescence rate by extrapolating the orbital
-periods of systems observed electromagnetically. Although this procedure has
-large uncertainty, tens of binary coalescence events are realistically expected
-to be observed in the advanced detector era later this
-decade~\cite{Abadie:2010p10836}.
 
-The gravitational-wave inspiral signal starts at low frequencies, gradually
-increasing in amplitude and frequency (i.e. it will chirp) until it reaches the
-innermost stable circular orbit (\textsc{isco}) ($f_\mathrm{ISCO} \approx
-4.4\times 10^3\,$Hz$\,\times(M/\Msun)^{-1}$) \editorial{Is $f_\mathrm{ISCO}$
-the orbital frequency or GW frequency?} for a test particle orbiting a
-Schwarzschild black hole with the total mass of the system $M$.  The frequency
-response of Advanced ground-based gravitational-wave detectors is such that the
-gravitational-wave frequency must be above $\sim 10$Hz before any appreciable
-signal-to-noise ratio (\textsc{snr}) can be accumulated\footnote{Which implies
-an orbital frequency of 5 Hz assuming that the dominant gravitational-wave
-frequency is twice the orbital frequency}.  Most of the \textsc{snr} is
-accumulated near the minimum of the detector noise spectral density at $\sim
-100$Hz. The inspiral waveform of a binary neutron star system will spend $\sim
-10^3$s in band.  Conveniently, the gravitational waveforms of compact binary
-coalescence are understood to the level required by Advanced \textsc{ligo} to
-detect~\cite{BuonannoIyerOchsnerYiSathya2009} via matched filtering techniques.
-Matched filtering detection pipelines maximize signal-to-noise ratio over a
-bank of filters, which are template waveforms in the binary parameter space.
-The templates are spaced to bound the loss in signal-to-noise ratio for any
-binary signal within the parameter space
-boundaries~\cite{Owen:1995tm,Owen:1998dk}.
+Radio observation of binary pulsar systems in the Galaxy provide indirect
+evidence that such systems are losing energy according to the prediction of
+gravitational-wave emission from general
+relativity~\cite{Taylor:1982,Abadie:2010p10836}.  One can estimate the expected
+coalescence rate by extrapolating the orbital periods of systems observed
+electromagnetically.  Although this procedure has large uncertainty, tens of
+binary coalescence events are realistically expected to be observed in the
+advanced detector era later this decade~\cite{Abadie:2010p10836}.
 
-As a binary containing a neutron star with a low-mass companion (a neutron star
-or black hole with mass less than $10\Msun$) approaches the merging point, the
-neutron star or neutron stars are tidally disrupted. This can provide
-the matter required to produce bright electromagnetic counterparts in the
-moments following the merger~\cite{shibata:2007}.
-\editorial{nvf: Chad thinks we need more mechanisms. I think this stands on
-its own without the acceleration mechanism.}
-In order to observe the prompt and potentially brightest
-electromagnetic emission, telescopes will have to be pointing in the direction
-of the event during the seconds surrounding the merger. In order to assure
-that this occurs more often than by random chance from various astronomical
-surveys operating independently, the gravitational-wave community has initiated
-a project to send alerts when potential gravitational-wave transients are
-observed. Unfortunately, at best, gravitational-wave experiments can only hope
-to have accumulated sufficient \textsc{snr} within seconds (or less) prior to
-the merger.  This makes it challenging to alert telescopes promptly to catch as
-much of the electromagnetic signal as possible.
+As a binary system loses energy to gravitational waves, its orbital separation
+decreases. This causes a run-away inspiral with the gravitational-wave
+amplitude and frequency increasing until the system eventually merges near the
+innermost stable circular orbit (\ISCO).
+%
+\footnote{$f_\mathrm{ISCO} \approx 4.4\times 10^3\,$Hz$\,\times(M/\Msun)^{-1}$. 
+for a test particle orbiting a Schwarzschild black hole with the total mass of
+the system $M$.}
+%
+If a neutron star is involved it may become tidally disrupted near the merger.
+This can provide the matter required to produce bright electromagnetic
+counterparts in the moments following the merger~\cite{shibata:2007}.
+\editorial{nvf: Chad thinks we need more mechanisms. I think this stands on its
+own without the acceleration mechanism.} In order to observe the prompt and
+potentially brightest electromagnetic emission, telescopes will have to be
+pointing in the direction of the event during the seconds surrounding the
+merger. In order to assure that this occurs more often than by random chance
+from various astronomical surveys operating independently, the
+gravitational-wave community has initiated a project to send alerts when
+potential gravitational-wave transients are observed.
 
@@
 operating, several all-sky detection pipelines operated in a low-latency
-configuration~\cite{HugheyGWPAW2011, S6lowlatency}. Several candidates were sent
-for electromagnetic followup to several telescopes including Swift,
+configuration~\cite{HugheyGWPAW2011, S6lowlatency}. Candidates were sent for
+electromagnetic followup to several telescopes including Swift,
 \textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008, HugheyGWPAW2011}.
 The S6 analyses achieved latencies of $~\sim30$--$60$ minutes, which were
-dominated by
-a human vetting process. The intrinsic gravitational-wave trigger generator
-latency was $\sim$2--5 minutes. In this work we will assume that the human
-vetting process can be automated in the advanced detector era and that the
-dominating latency will be trigger generation. From this perspective we will
-try to address the realistic expectations for reducing trigger latency from
-$\sim$minutes to $\lesssim$seconds.
+dominated by a human vetting process. The intrinsic gravitational-wave trigger
+generator latency was $\sim$2--5 minutes.  In this work we will assume that
+the human vetting process can be automated in the advanced detector era and
+that the dominating latency will be trigger generation. From this perspective
+we will try to address the realistic expectations for reducing trigger latency
+from $\sim$ minutes \emph{after} the merger to minutes \emph{before} the
+merger for sufficiently loud events.  Figure \ref{fig:earlywarning} presents
+the possibility for early detection of NS--NS binaries with Advanced \LIGO\
+assuming the rate predictions in \cite{Abadie:2010p10836}.
+%
+\begin{figure}
+\includegraphics{N_before_Tc.pdf}
+\caption{\label{fig:earlywarning}\FIXME{Drew has independently reproduced
+this calculation} Expected number of NS--NS sources that will be detectable
+$T_{bc}$ seconds before coalescence.  The solid line is the most likely yearly
+rate estimate $N_{\mathrm{re}}$ for Advanced \LIGO\ and the shaded region is
+the interval $N_{\mathrm{low}}$ to $N_{\mathrm{high}}$ from
+\cite{Abadie:2010p10836}.  Note that assuming \SNR\ 8 is sufficient for
+detection and that we observe $N_{\mathrm{re}} = 40$ events per year with a
+detector having the ZERO\_DET\_high\_P noise model described
+in~\cite{ALIGONoiseZERO_DET_high_P}, $\sim10$ sources may be detected within
+10 seconds of merger and $\sim1$ sources maybe detected within 100 seconds of
+merger.  Other noise models, ZERO\_DET\_low\_P~\cite{ALIGONoiseZERO_DET_low_P}
+and BHBH\_20deg~\cite{ALIGONoiseBHBH_20deg} provide better results for early detection
+but at the cost of fewer total events observed above SNR 8.}
+\end{figure}
 
-The reduction in latency is a challenging task.  As one pushes typical
-finite-impulse-response (\textsc{fir}) filtering techniques from high latency
-configurations with a computational complexity that scales as
-$\mathcal{O}(N\log{N})$, where $N$ is the number of sample points in the
-calculation, to low-latency configuration, the complexity goes to
-$\mathcal{O}(N^2)$. Na\"ively this would require orders of magnitude more
-computational resources. However, there are aspects of searches for binary
-coalescence that can be exploited to offset some of the difference in
-computational complexity. 1) The time-frequency morphology of the binary
-waveforms is such that multi-rate filtering may be employed. 2) The redundancy
-in the filter banks used to cover a particular parameter space may be
-reduced via rank-reduction schemes such as the singular value
-decomposition~\cite{Cannon:2010p10398}. We will use both aspects to
-demonstrate that a very low latency analysis of gravitational-wave data for
-compact binary sources is possible with the current computing resources.
-Assuming other aspects of gravitational-wave observation latency
-can be reduced significantly, this should allow the possibility for alerts to
-be sent within seconds of the binary merger.
+Although it is technically possible to detect NS--NS signals with the
+\earlywarning\ predicted in figure \ref{fig:earlywarning}, the reduction in
+latency is a challenging task.  As finite-impulse-response (\fir) filtering
+techniques are pushed from high latency configurations to low-latency
+configurations more computational resources are required.  A brute-force
+implementation of low lateny filtering may be too computationally costly to
+implement.  However, there are aspects of searches for binary coalescence that
+can be exploited to offset some of the intrinsic computational complexity: 1)
+The time-frequency evolution of the binary waveforms is such that multi-rate
+filtering may be employed. 2) The redundancy in the filter banks used to cover
+a particular parameter space may be reduced via rank-reduction schemes such as
+the singular value decomposition~\cite{Cannon:2010p10398}. We will use both
+aspects to demonstrate that a very low latency analysis with predictive
+detection of compact binary sources is possible with current computing
+resources.  Assuming other aspects of gravitational-wave observation latency
+can be reduced significantly, this should allow the possibility for prompt
+alerts to be sent to the Astronomical community.
 
-The paper is organized as follows. First we provide an overview of the
-standard method for detecting compact binary coalescence signals and
-describe how it can be modified to accomodate low-latency analysis. We
-then describe the pipeline we have constructed that implements these changes.
-To validate the approach we present results of simulations and conclude
-with some remarks on what remains to prepare for the Advanced detector era.
+The paper is organized as follows. First we provide an overview of our method
+for detecting compact binary coalescence signals in an \earlywarning\
+analysis. We then describe the pipeline we have constructed that implements
+our method.  To validate the approach we present results of simulations and
+conclude with some remarks on what remains to prepare for the Advanced
+detector era.

### 2011-03-31 time_slices.py
inserted macro for a symbol
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@708 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -80,7 +80,7 @@ pylab.savefig('time_slices.pdf')
 
 print r"\begin{tabular}{|lrr}"
 #FIXME change the symbol for the number of sample points per slice if the macro changes
-print r"$\mathbf{f_k}$ \bf{(Hz)} & $\mathbf{(t_{k+1}}$,$\mathbf{t_k]}$ \bf{(s)} & $\mathbf{M_{\mathrm{k}}}$ \\"
+print r"$\mathbf{f_k}$ \bf{(Hz)} & $\mathbf{(t_{k+1}}$,$\mathbf{t_k]}$ \bf{(s)} & $\mathbf{\slicessamps}$ \\"
 print r"\hline"
 for slice in slices:
 	begin = slice['begin']

### 2011-03-31 results.tex
put a fixme at the top
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@709 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{SECIV}\label{sec:results}
+
+\FIXME{
+I would like to see this section changed to emphasize the low-latency time domain filtering
+1. Use a live simulated white noise source (this ignores the latency of whitening, but that goes
+beyond the scope of this paper, and we mention this and perhaps suggest some exporation
+2. Use TD filtering of N templates
+3. Present the performance and latency and provide estimates for number of cores required for realtime ALIGO search based on current infrastructure
+4. Compute the impulse responses of the templates and histogram the SNR loss for the SVD and time/slice/resampling
+5. Put  a tee after some of the low frequency stages and perhaps test the possibility of
+predictive filtering (also with latency measurments)
+This will remove the complications of running a full pipeline.  We need to sort things out before
+we do that, and this paper should not be delayed any further.  The above tests will make the point we need to make
+}
 

### 2011-03-31 method.tex
substantial rework, moved comp. cost portion of appendix into main body
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@711 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\section{Method}
+\section{Early warning searches for compact binary coalescence}
 \label{SECII}\label{sec:method}
-\subsection{The standard approach: matched filtering}
 
-Searches for gravitational waves from compact binary coalescences typically
-employ matched filter banks \cite{findchirppaper}.  Potential inspiral signals
-are continuously parameterized by time, amplitude, phase, and a set of
-intrinsic source parameters $\theta$, which in this paper we shall take to
-consist of the two component masses of a binary, $\theta = (m_1, m_2)$.  Let
-$h_+(\theta)$ and $h_\times(\theta)$ be, respectively, the `$+$' and `$\times$'
-polarization gravitational wave signals that would arise from a fiducial
-face-on binary at some distance.  Because, for inspiral signals, $h_+$ and
-$h_\times$ are nearly in quadrature, they are generally combined into a single
-complex-valued template $h = h_+ + i h_\times$.
+In this section we describe a decomposition of the compact binary parameter
+space that reduces low latency filtering cost sufficiently to allow for the
+possibility of \earlywarning\ detection with modest computing requirements.  We
+expand on the ideas of~\cite{Marion2004, Buskulic2010} that describe a
+multiband decomposition of the compact binary parameter space that resulted in
+search with $\sim$minutes latency in LIGO's S6 and Virgo's VSR2 science runs.
+We combine this with the orthogonal decomposition described
+in~\cite{Cannon:2010p10398} that exploits the redundancy of the template banks.
 
-The detection procedure for just one set of intrinsic source parameters
-$\theta$ starts by {\em whitening} the measurement data stream $x$.  This
-involves finding a linear filter that renders the detector's noise \textsc{iid}
-and Gaussian.  This filter is applied to the measurement, yielding the whitened
-data stream $x^\mathrm W$.  The same linear filter is applied to the template
-$h(\theta)$, yielding the whitened template $h^\mathrm W (\theta)$.  The
-matched filter is the normalized cross-correlation of $h^\mathrm W$ and
-$x^\mathrm W$, \editorial{After introducing the $\star$ notation for cross
-correlation, must we define it for continuous and discrete time?}
+Inspiral signals are parameterized by a set of intrinsic parameters
+$\bar{\theta}$ that determine the amplitude and phase evolution of the observed
+binary signals.  Searches for gravitational waves from compact binary
+coalescence typically employ matched filter banks, called template banks,
+\cite{findchirppaper} that discretely sample the possible intrinsic parameters.
+The filter models of the waveforms, known as templates are weighted by the
+detector noise amplitude spectral density, $x(t,\bar{\theta})$ and filtered
+against the whitened detector data.  A sufficient number of templates are
+chosen to assure a minimal loss of SNR~\cite{Owen:1995tm,Owen:1998dk}.  To
+construct a template bank, matched filters are chosen with discrete signal
+parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$, such that any possible
+signal will have a cross-correlation of $\geq0.97$ with at least one template.
+Such a template bank is said to have a 97\% {\em minimum match}. For systems
+where the effects of spin can be ignored, the intrinsic parameters are the
+component masses of the binary, $\bar{\theta} = (m_1, m_2)$.
+
+The procedure for choosing a bank of matched filters for gravitational wave
+searches is known~\cite{Owen:1995tm}. Therefore, here we will assume that a
+bank of filters has already been chosen based on the targetted parameter space
+and we will denote the $i^{\mathrm th}$ filter with parameters $\bar{\theta_i}$
+as a function of time $x_i(\tau)$.%
 %
-$$
+\footnote{There are two gravitational wave polarizations, $+$ and $\times$. A
+given detector will observe a combination of these polarizations that will
+largely be degenerate with an overall unknown constant phase.  This can be
+maximized over by filtering for quadrature phases and taking the magnitude of
+the result. For simplicity we will ignore that aspect in this work as it is
+straightforward to generalize, but not necessary for understanding any of the
+points that will be made.}
 %
-\rho(\theta) = \frac{h^\mathrm W (\theta) \star x^\mathrm W}{|h^\mathrm W
-(\theta)|}.
+In this work we will discuss transformations to a set of filters
+$\{x_i(\tau)\}$.  Some of these transformations are not useful or practical
+over the entire parameter space.  For that reason we assume from here onward
+that the set of filters $\{x_i(\tau)\}$ refer to a set of near neighbor filters
+that can be chosen as a subset of the full parameter space.  Several such
+subsets can be chosen until all of the filters are a member of one local set.
+
+Filtering the detector data $h(t)$ involves a convolution of the data with the
+filter.  For a unit-normalized filter, and whitened detector data, the result
+can be interpreted as the signal-to-noise ratio, $\rho_i(t)$ and is defined as
 %
-$$
+% Filtering equations
 %
-This is called the signal to noise ratio, or \textsc{snr}.  The detection
-statistic is the modulus of this, $|\rho(\theta)|$, which has a $\chi^2$
-distribution with 2 degrees of freedom in the absence of signal.
+\begin{eqnarray}
+\rho_i(t) &=& \int x_i(\tau) \, h(t-\tau) \, d \tau \label{eq:SNRTD} \\
+          &=& \int \tilde{x}_i(f) \, \tilde{h}(f) \, \e^{-2\pi\i ft} \label{eq:SNRFD} df,
+\end{eqnarray}
+%
+%
+where the second line is a result of the convolution theorem and
+$\tilde{x}_i(f)$ is the Fourier transform of $x_i(\tau)$ as is $\tilde{h}(f)$ the
+Fourier transform of $h(t)$.
 
-To construct a filter bank, matched filters are realized for discrete signal
-parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$, such that any possible
-signal will have a maximum cross-correlation of at least 0.97 with at least one
-template.  \editorial{Citation needed for template placement procedure?}  Such
-a template bank is said to have a 97\% {\em minimum match}.  This technique is
-designed so that an inspiral signal can be detected without any prior knowledge
-of its intrinsic parameters: at most 3\% of the \textsc{snr} is lost by a
-signal's parameters not exactly coinciding with a template's.  A trigger is
-reported for the template parameters $\theta_i$ and time $t$ for which $|\rho|$
-is a maximum over some moving interval in $\theta$ and $t$.
+The evaluation of the integrals in \eqref{eq:SNRTD} and \eqref{eq:SNRFD} are
+implemented as sums over sample points for the digitized gravitational wave
+detector output.  Discrete Fourier transforms can be computed efficiently
+numerically.  For that reason \eqref{eq:SNRFD} is typically far faster,
+computationally.  To evaluate \eqref{eq:SNRTD} requires \order{\tmpsamps
+\hoftsamps} floating point operations per filter $x_i$, where \tmpsamps\ is the
+number of sample points in the filter $x_i$ and \hoftsamps\ is the number of
+sample points in the data $h$.  Assuming \numtmps\ filters are required, the
+total cost is \order{\numtmps \tmpsamps \hoftsamps} However, \eqref{eq:SNRFD}
+requires only \order{\hoftsamps \log \hoftsamps} operations per filter assuming
+transform lengths that are longer than the filter (i.e. \hoftsamps\ $>$
+\tmpsamps), resulting in a total cost of \order{\numtmps \hoftsamps \log
+\hoftsamps}.  In most cases \tmpsamps\ $\gg$ $\log$ \hoftsamps\ and the
+computational savings by choosing the frequency domain integral form
+\eqref{eq:SNRFD} is clear.  However, to take full advantage of the
+computational efficiency of \eqref{eq:SNRFD} requires an acausal knowledge of
+the detector data $h(t)$, which implies an inherent latency.  In contrast,
+\eqref{eq:SNRTD} can be updated every time a new sample point of detector data
+is taken.
 
-\subsubsection{Latency and overhead}
-
-The matched filter bank can be implemented with finite impulse response
-(\textsc{fir}) filters.  \textsc{fir} filters are perfectly suited for realtime
-detection, because they do not introduce any latency at all.  However,
-\textsc{fir} filters are very expensive: for $N$ templates of $M$ samples each,
-a \textsc{fir} matched filter bank costs $\mathcal O(M N)$ operations per
-sample.
-
-Much more commonly, the matched filter bank is implemented using \textsc{fft}
-convolutions, costing only $\mathcal O(M \lg N)$ \editorial{$\lg$ is $\log_2$.
-Explain nomenclature? Switch to $\log_2$?} per sample but having a latency of
-at least $M$ samples.  For example, for a bank of $1\,\mathrm{ks}$ templates
-sampled at $4096\,\mathrm{Hz}$, the \textsc{fft} implementation requires about
-about $2.2 \times 10^4$ times fewer floating point operations per sample than
-the \textsc{fir} implementation.  However, the \textsc{fft} implementation has
-a latency of at least $1\,\mathrm{ks}$.
-
-This presents a dilemma: it seems that low latency detection using the
-\textsc{fir} implementation is prohibitively expensive, whereas computationally
-cheap detection with the \textsc{fft} method comes with minutes to hours of
-latency.
-
-In the rest of the section, we will describe a detection strategy that makes
-use of some very general properties of inspiral template banks in order to
-evaluate a matched filter bank with far lower computational cost than the
-conventional \textsc{fir} method and far less latency than the conventional
-\textsc{fft} method.
+In order to minimize latency we propose using the time domain convolution
+presented in \eqref{eq:SNRTD}.  However, because the brute force evaluation of
+\eqref{eq:SNRTD} is far too costly to be useful, we will consider an
+approximation to \eqref{eq:SNRTD} that can reduce substantially the cost of
+\realtime\ filtering. This approximation has the form
+%
+% orthogonal decomposition filtering
+%
+\begin{equation}
+\rho_i(t) \approx \sum_k^{\numslices} \sum_j^{\numsvdtmps} 
+	\int_{\tau_{k}}^{\tau_{k+1}} v_{ijk} \sigma_{jk} u_{jk}\,(\tau) \, h(t-\tau) \, d \tau \label{eq:decomp}
+\end{equation} 
+%
+%
+where $u_{jk}(\tau)$ is an orthogonal basis set of filters spanning the space
+of $\{x_i(\tau)\}$ and $\sigma_{jk} v_{ijk}$ is a tensor relating the filters
+$u_{jk}(\tau)$ to the original filter set $\{x_i(\tau)\}$.  We claim that with
+a suitable choice of filters $u_{jk}(\tau)$ one can reduce the computational
+cost of \eqref{eq:SNRTD} sufficiently to feasibly search for
+gravitational-waves from compact binary coalescence in \realtime.  This
+requires 1) exploiting the redundancy of the template bank and 2) exploiting
+the time-frequency characteristics of the binary waveforms.  We describe our
+procedure for producing the decompostion in \eqref{eq:decomp} in the remainder
+of this section.
 
@@
 
-Our first innovation is to split each template into disjoint intervals, or
-\emph{time slices}.  A matched filter is constructed for each time slice, whose
-outputs form an ensemble of partial \textsc{snr} streams.  By linearity, these
-partial \textsc{snr} streams can be suitably time delayed and summed to
-reproduce exactly the \textsc{snr} with respect to the original template.  By
-exploiting some general properties of inspiral waveforms, we shall see that
-time slices permit a reduction in latency relative to the \textsc{fft} matched
-filter, but with comparable computational overhead.
+The first step of the orthogonal decomposition described in \eqref{eq:decomp}
+is to divide the templates into \emph{time slices}.  This is a time domain
+analogue to the frequency domain decomposition described in ~\cite{Marion2004,
+Buskulic2010, beauville2006, beauville2008}.  A matched filter is constructed
+for each time slice.  The outputs form an ensemble of partial \textsc{snr}
+streams.  By linearity, these partial \SNR\ streams can be suitably time
+delayed and summed to reproduce the \SNR\ of the full template.  We will show
+in the next section that this, combined with the singular value docomposition,
+is sufficient to enable a computationally efficient time-domain search and
+furthermore is an essential part of an \earlywarning\ detection scheme.
 
-\editorial{\textsc{mbta} should be cited in the introduction.}  A similar idea
-has been demonstrated by the Virgo Collaboration's \textsc{mbta}
-pipeline~\cite{Marion2004,beauville2006,beauville2008,Buskulic2010}, which
-operated in a low-latency mode during \textsc{ligo}'s sixth science run and
-Virgo's second science run in 2010.
-
-Each partial matched filter can be realized with either the \textsc{fir} or the
-\textsc{fft} method.  If all of the time slices make use of the \textsc{fft}
-method, the resulting filter will have lower latency than a conventional
-\textsc{fft} matched filter, and also less computational overhead than an
-\textsc{fir} matched filter.  A further benefit is that all of the time slices
-may be processed in parallel for additional speedup.
-
-If the template is known to be quasi-bandlimited, then the partial matched
-filter for each time slice may be processed at a reduced sample rate.  Compact
-binary inspiral waveforms fit this description superbly: they are chirps whose
-frequency and amplitude rise according to power laws of time.  They are not
-only quasi-bandlimited, but quasi-monochromatic.  By exploiting this property,
-we can implement a matched filter for an inspiral waveform using a multirate
-filter bank that requires far fewer floating point operations.
-
-For concreteness and simplicity, let us consider an inspiral waveform in the
+For concreteness and simplicity, we will consider an inspiral waveform in the
 quadrupole approximation, for which the time-frequency relation is
@@
 M_\odot / c^3 \approx 5 \mu\mathrm{s}$) and $t$ is the time relative to the
-coalescence of the binary~\cite{findchirppaper, kidder1992, blanchet2002,
-hanna2009}. \editorial{Way too many citations here.} Usually the template is
-truncated at some prescribed time $t = t_0$, often chosen to correspond to the
-innermost stable circular orbit at which $f = f_\mathrm{ISCO} \approx 4400 \,
-M_\odot / M$.  An inspiral signal will enter the \textsc{ligo} band at a low
-frequency cutoff, $f = f_\mathrm{low}$ corresponding to a time
-$t_\mathrm{low}$.  This template has a duration of $t_\mathrm{0} -
-t_\mathrm{low}$ and is critically sampled at a rate of $2 f_\mathrm{isco}$.
+coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
+template is truncated at some prescribed time $t_0$, or equivalently frequency \fmax.
+This is often chosen to correspond to the \ISCO\ defined previously. An inspiral signal
+will enter the detection band at a low frequency, $f = f_\mathrm{low}$
+corresponding to a time $t_\mathrm{low}$.  The template is assumed to be zero
+outside the interval $[t_\mathrm{low}, t_\mathrm{0})$ and is said to have  a
+duration of $t_\mathrm{0} - t_\mathrm{low}$. It is critically sampled at a
+rate of $2 \fmax$.
 
-The time slices for this template consist of the $K$ intervals $(t_K, t_{K-1}],
-\dots, (t_2, t_1], (t_1, t_0]$ sampled at frequencies $f_\mathrm{K-1}, \dots,
-f_1, f_0$ where $f_0 = 2 f_\mathrm{ISCO}$, $t_0 = t_\mathrm{ISCO}$, $f_{K-1}
-\geqslant 2 f_\mathrm{low}$, and $t_K \leqslant t_\mathrm{low}$.
-
-If all of these time slices are implemented using the \textsc{fft}, then the
-latency of this filter bank is $\max_j {\left[-\left( 2(t_j - t_{j-1}) -
-t_{j-1} \right)\right]} = \max_j \left( 3 t_{j-1} - 2 t_j \right)$.
-
+The monotonic time-frequency relationship of \eqref{eq:fgw} allows us to choose
+time slice boundaries that require substantially less bandwidth at early times
+in the inspiral.  Our goal is to reduce the filtering cost of a large fraction
+of the waveform by computing part of the filter integral at a lower sample
+rate.  Specifically we consider here time slice boundaries with the next
+highest power-of-two sample rate that critically samples the time sliced
+filter.  The time slices for this template consist of the $k$ intervals $(t_k,
+t_{k-1}], \dots, (t_2, t_1], (t_1, t_0]$ sampled at frequencies
+$f_\mathrm{k-1}, \dots, f_1, f_0$ where $f_0 \geq 2 f_\mathrm{ISCO}$, $t_0 =
+t_\mathrm{ISCO}$, $f_{k-1} \geqslant 2 f_\mathrm{low}$, and $t_k \leqslant
+t_\mathrm{low}$. An example time slice design satisfying these constraints for
+a $1.4 - 1.4 \, M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
+%
 \begin{table}[h!]
-\caption{Example of critically sampled, power-of-2 time slices for a $1.4 - 1.4
-\, M_{\odot}$ template extending from $f_\mathrm{low} = 10 \, \mathrm{Hz}$ to
-$f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$ with a time frequency structure given by
-($\ref{eq:fgw})$.}
-\label{table:time_slices}
-\begin{minipage}[c]{0.5\textwidth}
+\begin{minipage}[c]{0.52\textwidth}
 \centering
-\includegraphics[trim = 3cm 0cm 1.5cm 0cm, scale=0.4]{time_slices.pdf}
+\includegraphics{time_slices.pdf}
 \end{minipage}
@@
 \end{minipage}
+\caption{\label{table:time_slices} Example of nearly critically sampled,
+power-of-two time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from
+$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
+with a time frequency structure given by ($\ref{eq:fgw})$. $f_k$ is the sample
+rate of the time slice, $(t_{k+1}, t_k]$ are the boundaries in seconds
+preceeding coalescence and \slicessamps\ are the number of sample points in the
+$k^{\mathrm{th}}$ filter.}
+\end{table}
+%
+
+Rather than applying a unique time slice decomposition to each template
+waveform, we find a decomposition that is adequate for the entire set
+$\{x_i\}$.  This is facilated by choosing templates in the set $\{x_i\}$ with
+similar chirp masses, $\mathcal{M}$.  The time slice decompositions of the
+filters $x_i(\tau)$ lead to new filters $y_{ik}(\tau)$ satisfying
+%
+\begin{equation}
+\label{eq:time_slices}
+x_i(\tau) \approx \sum_k^{\numslices} y_{ik}(\tau)
+\end{equation}
+where \numslices\ is the number of time slices required.  Note that we write
+the relationship as approximate since the resampling implementation may have
+some inherent loss in quality.  We note that by construction these filters are
+orthogonal over the index $k$ since they are disjoint in time.  In the next
+section we examine how to reduce the number of filters $y_{ik}(\tau)$ via
+singular value decomposition to construct a set of filters that is also
+orthogonal within a given time slice.
+
+\subsection{Reducing the number of filters with the singular value
+decomposition}
+
+As described previously, the template banks are, by design, highly correlated.
+It is possible to greatly reduce the number of filters required to achieve a
+particular minimum match by designing an appropriate set of orthonormal {\em
+basis templates}.  A purely numerical technique based on the application of the
+singular value decomposition (\SVD) to inspiral waveforms is demonstrated
+in~\cite{Cannon:2010p10398}.  Using the results of ~\cite{Cannon:2010p10398} we
+establish that the filters of the previous section can be approximated to high
+accuracy by the expansion in the singular value basis
+%
+\begin{equation}
+y_{ik}(\tau) \approx \sum_j^{\numsvdtmps} \sigma_{jk} v_{ijk} u_{j}(\tau)
+\label{eq:svddecomp}
+\end{equation}
+%
+where $\sigma_{jk}$ is the $j^{\mathrm{th}}$ singular value for the
+$k^{\mathrm{th}}$ time slice, $v_{ijk}$ is an orthogonal matrix for the
+$k^{\mathrm{th}}$ time slice  and $u_{jk}$ is a new orthogonal basis filter set
+for the $k^{\mathrm{th}}$ time slice.  The authors of \cite{Cannon:2010p10398}
+showed that to high reconstruction accuracy far fewer filters are needed than
+were in the original template bank. We find that when combined with the time
+slice decomposition, the number of \SVD\ filters, \numsvdtmps\ is much smaller
+than the original number of filters \numtmps.  We combine \eqref{eq:svddecomp}
+with \eqref{eq:time_slices} to arrive at \eqref{eq:decomp}.  In the next
+section we compute the expected computational cost scaling of this
+decomposition and compare it with the brute force implementation in
+\eqref{eq:SNRTD} and higher latency \fft\ methods.
+
+\subsection{Comparison of computational costs}
+
+We now examine the computational cost scaling of the approximate implementation
+of \eqref{eq:SNRTD} as \eqref{eq:decomp}.  An actual implementation of this
+decomposition in a working analysis pipeline is discussed in the next section
+along with measured computational requirements.  For convenience, table
+\ref{tab:recap} is a recap of the meaning of various symbols used in this
+calculation.
+%
+% symbols used in FLOPs calculations table
+%
+\begin{table}
+\begin{tabular}{rl}
+\bf{Symbol}	& \bf{Definition} \\
+\hline
+\hoftsamps	& The number of sample points in the data \\
+\tmpsamps	& The number of sample points in template $x_i$ \\
+\numtmps	& The number of templates in the set $\{x_i\}$ \\
+\numslices	& The number of time slices \\
+\numsvdtmps	& The number of orthogonal filters in time slice $k$ \\
+\slicessamps	& The number of samples in time slice $k$ \\
+\fmax		& The maximum frequency of a filter \\
+$f_k$		& The sample frequency of the $k^{\mathrm{th}}$ filter \\
+\resampsamps	& The number of sample points in the resample filter
+\end{tabular}
+\caption{\label{tab:recap} Notation used to describe filtering.  This table
+provides a quick reference for symbols used.}
 \end{table}
 
-An example time slice design satisfying these constraints for a $1.4 - 1.4 \,
-M_{\odot}$ is shown in table~\ref{table:time_slices}.  For this example, the
-latency for this time slice design is just $\input{time_slice_latency.tex}$
-even if all of the time slices are implemented with the \textsc{fft} method.
-This set of time slices will require $\input{time_slice_ops_firslice.tex}$
-operations per sample per template, compared with
-$\input{time_slice_ops_conv.tex}$ for pure \textsc{fft} cross-correlation
-without time slices, or $\input{time_slice_ops_td.tex}$ for the \textsc{fir}
-filter method without time slices.
+In table \ref{table:flops} we present the computational cost scaling in
+floating point operations per sample for common tasks in the pipeline.  
+%
+% common tasks FLOPS table
+%
+\begin{table}[htdp]
+\begin{center}
+\begin{tabular}{l l}
+\bf{Process} & \bf{ops/sample} \\
+\hline
+% FIR
+\fir\ matched filter, \numtmps\ templates of length \tmpsamps\ & $2 \numtmps \tmpsamps$ \\
+% FFT
+\fft\ matched filter, \numtmps\ templates of length \tmpsamps\, & \multirow{2}{*}{$\mbox{\fontsize{14}{17}\selectfont $\frac{4 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock}$}$} \\
+% FFT contd...
+$\,\,\,\,\,\,$ blocks of length \fftblock & \\
+% Resampling
+\fir\ resampling filter, length \resampsamps\ for each of \numtmps\ templates& \multirow{2}{*}{$2 \numtmps \resampsamps\ f_1 / f_2$} \\
+% Resampling contd...
+$\,\,\,\,\,\,$ and sample rates $f_1 < f_2$ & \\
+% Matrix multiply
+multiply $M \times L$ real matrix by $L\times1$ real vector & $2 M L$ \\
+%
+\end{tabular}
+\end{center}
+\caption{Number of floating point operations per sample (multiplications and
+divisions) required for a selection of signal processing operations used in
+\textsc{lloid}.}
+\label{table:flops}
+\end{table}
 
-\subsection{Reducing the number of filters with the singular value decomposition}
+The filter bank can be implemented using finite impulse response \fir\ filters,
+which are just sliding window dot products.  If there are \numtmps\ templates
+of length \tmpsamps\, and the data stream contains \hoftsamps\ samples, then
+applying the filter bank requires $2 \numtmps \tmpsamps \hoftsamps$ operations.
 
-Our second innovation exploits the fact that the templates in inspiral template
-banks are, by design, highly correlated.  This is due to the fact that points
-are chosen in parameter space to be templates such that any point from the
-region parameter space we cover has a match to the nearest template greater
-than our chosen minimum match. It is possible to greatly reduce the number of
-matched filters required to achieve a particular minimum match by designing an
-appropriate set of orthonormal {\em basis templates}.  A purely numerical
-technique based on the application of the singular value decomposition
-(\textsc{svd}) to inspiral waveforms is demonstrated by the authors
-in~\cite{Cannon:2010p10398}.
+More commonly, the matched filters are implemented using the \fft\ convolution.
+This entails applying {\fft}s to blocks of \fftblock\ samples, with $\tmpsamps
+\leq \fftblock$, each block overlapping the previous one by $ \fftblock -
+\tmpsamps $ samples.  There are $\hoftsamps/(\fftblock-\tmpsamps)$ such blocks
+required to filter \hoftsamps\ samples of data.  Modern implementations of the
+Cooley-Tukey \fft, such as the ubiquitous \texttt{fftw}, require about $4 N \lg
+N$ operations to evaluate a \textsc{dft} of size $N$~\cite{Johnson:2007p9654}.
+%
+%
+\editorial{This is more commonly known as ``overlap-save''.  We should find
+someone else's operation count and cite it.}  
+%
+%
+A \fftblock\ sample cross-correlation consists of a forward \fft, a \fftblock
+sample dot product, and an inverse \fft\, totaling $8 \fftblock \lg \fftblock +
+2 D$ operations per block.  Per sample, this is $(8 \lg \fftblock + 2) / (1 -
+\tmpsamps/\fftblock)$ operations.  As this expression indicates the number
+of operations increases as the block size \fftblock\ approaches the
+filter length \tmpsamps.
+%
+%
+\editorial{Drew: Why don't we change this to an overlap of $m$ samples so we
+can see what happens as we increase the overlap to reduce latency.}
 
-One may regard a bank of $N/2$ discretely-sampled, complex-valued templates of
-length $M$ samples, $h^\mathrm W(\theta_i; t_j) = [\mathbf H]_{ij} = H_{ij}$,
-as a real matrix of shape $N \times M$, where we have unpacked each
-complex-valued template into two real-valued templates.  The singular value
-decomposition is an exact factorization that exists for any matrix such that
+The \fir\ filter implementation has the advantage that it has no intrinsic
+latency, whereas the \fft\ convolution has latency of $\fftblock-\tmpsamps$. 
 %
-\begin{equation} \label{eq:svd-exact}
 %
-H_{ij} = [\mathbf {V \Sigma U}]_{ij} = \sum_{k=1}^{N} v_{ik} \sigma_k u_{kj},
+%For example, for a $1.4 -
+%1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the
+%\textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.  
 %
+However, the \fir\ filter implementation has the disadvantage of much greater
+overhead per sample than the \fft\ convolution.  For a $1\,\mathrm{ks}$
+template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation
+requires about about $\tmpsamps / 8 \lg 2 \tmpsamps = 2.2 \times 10^4$ times
+more operations per sample than the \fft implementation.  We will now consider
+the computational cost of the \fir\ filter implementation described in the
+previous sections.
+
+It is convenient to express the computational cost of the entire filtering
+procedure in floating point operations per second {\flops}.  The cost will be
+the sum of the cost of the \fir\ filtering for the orthogonal filter in each
+time slice plus the cost of reconstructing the original waveforms with matrix
+operations and resampling.  Using the formulas in table \ref{table:flops} we
+arrive at
+%
+\begin{equation}
+%\mathrm{Computational\, cost} 
+\mathrm{C.C.} \propto 2 \sum_k^{\numslices} (\slicessamps + \numtmps) \numsvdtmps f_k \nonumber
+	r \sum_{k, f_k \neq f_{\fmax}}^{\numslices} \numtmps \resampsamps f_k 
+\label{eq:TDcost}
 \end{equation}
 %
-where $\mathbf V$ and $\mathbf U$ are both unitary matrices and $\mathbf
-\Sigma$ is a diagonal matrix.  For our purposes, we associate the rows of
-$\mathbf U$ with a minimal set of basis templates, which become the kernels of
-basis filters.  These filters give rise to the orthogonal \textsc{snr}s,
-$\rho_k^\perp = u_k \star x^\mathsf{W}$.  The rows of $\mathbf{V \Sigma}$
-become reconstruction coefficients that map linear combinations of the
-orthogonal \textsc{snr}s from the basis filters back onto \textsc{snr}s for the
-original templates of interest: $\rho_i = \sum_k v_{ik} \sigma_k \rho_k^\perp$.
+where the first sum is the total cost for filtering and reconstructing the
+orthogonal filters $u_{jk}$.  The second sum is the cost of resampling the
+reconstructed time slice outputs to the original sample rate {\fmax}.  It
+assumes a \fir\ filter with \resampsamps\ sample points.  Note that the cost
+for resampling only occurs for the downsampled time slices. The resampling cost
+largely depends on {\resampsamps}.  The computational cost of \eqref{eq:TDcost}
+is dominated by the highest frequency terms in the sum.  Comparing just the
+highest frequency term of \eqref{eq:TDcost} with \eqref{eq:SNRTD} shows \order{
+(\slicesamps{\mathrm{max}} + \numtmps) \svdtmps{\mathrm{max}} \fmax} \flops\
+versus \order{\numtmps \tmpsamps \fmax} {\flops}.  The full calculation for a
+particular patch of parameter space is shown in table \ref{table:comp_cost}.
+\hl{FIXME we need to actually say what went into this}
 
-In many applications of the \textsc{svd}, including ours, the matrix can be
-well approximated by truncating the summation in equation~(\ref{eq:svd-exact})
-at $N' \ll N$:
-%
-\begin{equation}
-%
-H_{ij}^\prime = \sum_{k=1}^{N'} v_{ik} \sigma_k u_{kj}
-%
-\end{equation}
-%
-The cumulative sum of squares of the {\em singular values}, $\sigma_k$,
-measures the Frobenius norm of the approximation, such that
-%
-\begin{equation}
-%
-\frac{\| \mathbf H^\prime \|}{\| \mathbf H \|} = \left(\sum_{k=1}^{N'}
-|\sigma_k|^2\right)^{1/2} \left(\sum_{k=1}^{N} |\sigma_k|^2\right)^{-1/2}.
-%
-\end{equation}
-%
-This is also called the \textsc{svd} tolerance.  In our application, it relates
-to how much \textsc{snr} is lost by discarding $N - N'$ of the basis filters
-with the lowest singular values.
-
-This result differs from other work that models gravitational-wave chirp
-signals in approximate ways~\cite{Chassande-Mottin2006, Candes2008,
-BuonannoChenVallisneri:2003a} by starting with an exact representation of the
-desired template family and producing a rigorous approximation with a tunable
-accuracy.
-
-\subsection{Composite detection statistic and hierarchical detection}
-
-\editorial{We should cite \cite{Scharf:1994p12497} here.} Although the
-\textsc{svd} allows us to reduce the number of matched filters, this comes at
-the price of having to perform a matrix multiplication by the reconstruction
-matrix $\mathbf {V \Sigma}$ at every sample.  In some circumstances, this
-matrix multiplication may be more expensive than applying the orthogonal
-matched filters.
-
-Further speedup may be gained in a hierarchical detection scheme.  In general,
-the orthogonal \textsc{snr}s alone provide some indication of whether any
-template in the original template bank is likely to have a large \textsc{snr}.
-Consider some composite detection statistic that is a scalar function of the
-orthogonal \textsc{snr}s, $\Gamma(\rho_1^\perp, \rho_2^\perp, \cdots,
-\rho_L^\perp)$.  Suppose that we can infer the distribution of $\Gamma$ for a
-data stream with the signal present or with the signal absent.  Then we may use
-a threshold crossing of $\Gamma$ to trigger the conditional application of the
-expensive reconstruction matrix only during the times when a signal is likely
-to be present.
-
-Following the ideas in~\cite{Anderson:2000yy}, we employ is a weighted sum of
-squares, $\Gamma = \sum_{k=1}^{N'} w_k (\rho_k^\perp) ^ 2$, with the particular
-weights
-%
-\begin{equation}
-%
-w_k = \frac{\sigma_k^2}{\sigma_k^2 + N / A^2},
-%
-\end{equation}
-%
-where $A^2$ is a desired \textsc{snr} scale that is set by the analyst.  As the
-authors show in~\cite{svd-compdetstat}, this choice of weights has a better
-receiver operating characteristic for a signal of \textsc{snr}~=~$A$ than some
-other obvious choices, $w_k = \sigma_k^2$ or $w_k = 1$.
-
-\subsection{Comparison of computational costs and latency}
-
-Using all of these innovations, we are able to greatly reduce the compuational
-cost of searching for these signals. An example of this can be seen in
-table~\ref{table:comp_cost} where we compute the computational cost of
-filtering 256 templates from a $97\%$ minimal match template bank whitened with
-an (FIXME: is this correct?) Advanced \textsc{ligo} \textsc{asd}. It is
-interesting to note that by themselves, both the \textsc{svd} and time-slice
-innovations \emph{increase} the overall computational cost when combined with
-\textsc{fft} filtering. It is only when combined with conditional
-reconstruction do these methods yield computational savings. Details of the
-computational cost of different operations can be found in \ref{appendix:cost}.
+\editorial{From here down introduce an example, but don't go through the FFT
+methods in lloid focus just on the comparison of the TD method without lloid
+tricks, the FD method without lloid tricks, and the TD lloid method (no comp
+detection statistic though), remember that the goal is near realtime detection
+not computational cost improvements on FD methods}
 
@@
 \begin{table}[!h]
-\caption{Operation counts per sample for six different detection methods.  The
-operation counts for \textsc{lloid} assume a reconstruction duty cycle of 5\%.
-Note that the \textsc{fft} method with \textsc{lloid} is almost 10 times faster
-than the conventional \textsc{fft} method, despite having substantially lower
-latency.}
 \label{table:comp_cost}
 \begin{center}
-\input{flop_budget}
+\input{flop_budget_example}
 \end{center}
+\caption{\hl{FIXME: Operation counts per sample for four different detection methods.}}
 \end{table}

### 2011-05-04 inspiral_svd.tex
added flop budget
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@740 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 electromagnetic signals and correlating such signals with the gravitational
-wave strain.  In this paper, we describe a new, realtime, detection strategy
-that exploits the properties of inspiral waveforms using principal component
-analysis and multi-rate filtering.  We provide detection efficiency
-characteristics from a prototype search pipeline that is capable of
-\earlywarning\ detection with acheivable computational requirements.  
+wave strain.  In this paper, we describe a detection strategy that exploits the
+properties of inspiral waveforms using principal component analysis and
+multi-rate filtering.  We provide detection efficiency characteristics from a
+prototype search pipeline that is capable of \earlywarning\ detection with
+acheivable computational requirements.  
 \end{abstract}
@@
 
-\ack{\LIGO\ was constructed by the California Institute of Technology
-and Massachusetts Institute of Technology with funding from the National
-Science Foundation and operates under cooperative agreement
-\textsc{phy}-\oldstylenums{0107417}. CH would like to thank Ilya Mandel
-for many discussions about rate estimates and the prospects of early
-detection.
+\ack{\LIGO\ was constructed by the California Institute of Technology and
+Massachusetts Institute of Technology with funding from the National Science
+Foundation and operates under cooperative agreement
+\textsc{phy}-\oldstylenums{0107417}. CH would like to thank Ilya Mandel for
+many discussions about rate estimates and the prospects of early detection.
 

### 2011-05-04 introduction.tex
added flop budget
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@740 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \includegraphics{N_before_Tc.pdf}
-\caption{\label{fig:earlywarning}\FIXME{Drew has independently reproduced
-this calculation} Expected number of NS--NS sources that will be detectable
-$T_{bc}$ seconds before coalescence.  The solid line is the most likely yearly
-rate estimate $N_{\mathrm{re}}$ for Advanced \LIGO\ and the shaded region is
-the interval $N_{\mathrm{low}}$ to $N_{\mathrm{high}}$ from
+\caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
+be detectable $T_{bc}$ seconds before coalescence.  The solid line is the most
+likely yearly rate estimate $N_{\mathrm{re}}$ for Advanced \LIGO\ and the
+shaded region is the interval $N_{\mathrm{low}}$ to $N_{\mathrm{high}}$ from
 \cite{Abadie:2010p10836}.  Note that assuming \SNR\ 8 is sufficient for
@@
 detector having the ZERO\_DET\_high\_P noise model described
-in~\cite{ALIGONoiseZERO_DET_high_P}, $\sim10$ sources may be detected within
-10 seconds of merger and $\sim1$ sources maybe detected within 100 seconds of
+in~\cite{ALIGONoiseZERO_DET_high_P}, $\sim10$ sources may be detected within 10
+seconds of merger and $\sim1$ sources maybe detected within 100 seconds of
 merger.  Other noise models, ZERO\_DET\_low\_P~\cite{ALIGONoiseZERO_DET_low_P}
-and BHBH\_20deg~\cite{ALIGONoiseBHBH_20deg} provide better results for early detection
-but at the cost of fewer total events observed above SNR 8.}
+and BHBH\_20deg~\cite{ALIGONoiseBHBH_20deg} provide better results for early
+detection but at the cost of fewer total events observed above SNR 8.}
 \end{figure}

### 2011-05-04 inspiral_svd.tex
Revised abstract
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@742 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\title[LLOID]{Improving the prospects for multimessenger astronomy with
-\earlywarning\ detection of compact binary coalescence.}
+\title[LLOID]{Improving the prospects for multimessenger astronomy with \earlywarning\ detection of compact binary coalescence.}
+
+% Backup title.
+%\title[LLOID]{Method for early-warning detection of compact binary coalescence.}
 
@@
 Canada}
-\address{$^2$  Département de physique, École Normale Supérieure de Cachan, 61
-Avenue du Président Wilson, 94235 Cachan Cedex, France}
+\address{$^2$  D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, 61
+Avenue du Pr\'{e}sident Wilson, 94235 Cachan Cedex, France}
 \address{$^3$ LIGO Laboratory - California Institute of Technology, Pasadena,
@@
 The rapid detection of compact binary coalescence with a network of advanced
-gravitational-wave detectors offers a unique opportunity for multimessenger
-Astronomy.  Prompt alerts to the Astronomical community will assure that all
-available instruments are observing the electromagnetic counterparts to the
-first gravitational waves from coalescing compact objects.   Rapid, near
-realtime alerts, will improve the chances of detecting weak, decaying
-electromagnetic signals and correlating such signals with the gravitational
-wave strain.  In this paper, we describe a detection strategy that exploits the
-properties of inspiral waveforms using principal component analysis and
-multi-rate filtering.  We provide detection efficiency characteristics from a
-prototype search pipeline that is capable of \earlywarning\ detection with
-acheivable computational requirements.  
+gravitational-wave detectors will offer a unique opportunity for multimessenger astronomy.
+Prompt detection alerts to the astronomical community may make it possible to observe the onset of electromagnetic emission from compact binary coalescence.
+We demonstrate a computationally practical analysis strategy that produces early warning triggers even before gravitational radiation from the final merger has arrived at the detectors.
+With current rate estimates for the Advanced \textsc{ligo} design configuration, we should detect $\sim$10 sources earlier than 10 seconds before merger in 1 year of livetime.
 \end{abstract}

### 2011-05-05 introduction.tex
First pass on introduction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@743 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
+\begin{comment}
 The coalescence of compact binary systems consisting of neutron stars (NS)
 and/or black holes (BH) is the most promising source of gravitational radiation
-for ground-based gravitational-wave detectors such as Advanced
-\LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb}, and
-\LCGT~\cite{LCGTWeb}.  Detections will constrain the mass and spin of compact
-objects~\cite{finn1993,Poisson:1995ef}, the neutron star equation of
-state~\cite{flanagan:021502,Read:2009}, general relativity in the strong-field
-regime~\cite{Will:2005va}, the graviton
-mass~\cite{PhysRevD.80.044002,Keppel:2010qu}, and the Lorentz Invariance
-principle~\cite{Ellis2006402}.
-
-Radio observation of binary pulsar systems in the Galaxy provide indirect
-evidence that such systems are losing energy according to the prediction of
-gravitational-wave emission from general
-relativity~\cite{Taylor:1982,Abadie:2010p10836}.  One can estimate the expected
-coalescence rate by extrapolating the orbital periods of systems observed
-electromagnetically.  Although this procedure has large uncertainty, tens of
-binary coalescence events are realistically expected to be observed in the
+for Advanced \LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb}, and
+\LCGT~\cite{LCGTWeb}.  Tens of
+binary coalescence events are expected to be observed in the
 advanced detector era later this decade~\cite{Abadie:2010p10836}.
+\end{comment}
 
@@
 amplitude and frequency increasing until the system eventually merges near the
-innermost stable circular orbit (\ISCO).
-%
-\footnote{$f_\mathrm{ISCO} \approx 4.4\times 10^3\,$Hz$\,\times(M/\Msun)^{-1}$. 
-for a test particle orbiting a Schwarzschild black hole with the total mass of
-the system $M$.}
-%
-If a neutron star is involved it may become tidally disrupted near the merger.
-This can provide the matter required to produce bright electromagnetic
-counterparts in the moments following the merger~\cite{shibata:2007}.
-\editorial{nvf: Chad thinks we need more mechanisms. I think this stands on its
-own without the acceleration mechanism.} In order to observe the prompt and
-potentially brightest electromagnetic emission, telescopes will have to be
-pointing in the direction of the event during the seconds surrounding the
-merger. In order to assure that this occurs more often than by random chance
-from various astronomical surveys operating independently, the
-gravitational-wave community has initiated a project to send alerts when
-potential gravitational-wave transients are observed.
+innermost stable circular orbit (\ISCO).  If a neutron star is involved it may become tidally disrupted near the merger.
+This disrupted matter can fuel a bright electromagnetic
+counterpart in the system's final moments as a binary~\cite{shibata:2007}.
+In order to observe the prompt and
+potentially most intense electromagnetic emission, telescopes must point during the seconds surrounding the
+merger.
 
-In 2010, \textsc{ligo} completed its sixth science run (S6) and Virgo
-completed its third science run (VSR3). When both \textsc{ligo} and Virgo were
+\editorial{Citation needed for LOOC-UP} To make this possible, the gravitational-wave community initiated a project to send alerts when
+potential gravitational-wave transients are observed.  In October 2010, \textsc{ligo} completed its sixth science run (S6) and Virgo
+completed its third science run (VSR3).  While both \textsc{ligo} detectors and Virgo were
 operating, several all-sky detection pipelines operated in a low-latency
-configuration~\cite{HugheyGWPAW2011, S6lowlatency}. Candidates were sent for
-electromagnetic followup to several telescopes including Swift,
-\textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008, HugheyGWPAW2011}.
-The S6 analyses achieved latencies of $~\sim30$--$60$ minutes, which were
-dominated by a human vetting process. The intrinsic gravitational-wave trigger
-generator latency was $\sim$2--5 minutes.  In this work we will assume that
-the human vetting process can be automated in the advanced detector era and
-that the dominating latency will be trigger generation. From this perspective
-we will try to address the realistic expectations for reducing trigger latency
-from $\sim$ minutes \emph{after} the merger to minutes \emph{before} the
-merger for sufficiently loud events.  Figure \ref{fig:earlywarning} presents
-the possibility for early detection of NS--NS binaries with Advanced \LIGO\
-assuming the rate predictions in \cite{Abadie:2010p10836}.
+configuration, namely \textsc{mbta}, ihope, Coherent WaveBurst, and Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
+\editorial{Get references for these low-latency pipelines.} The S6 analysis achieved latencies of 30--60 minutes, which were
+dominated by a human vetting process. Candidates were sent for
+electromagnetic followup to several telescopes; Swift,
+\textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008, HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved the best gravitational-wave trigger
+generation latencies of 2--5 minutes.  We assume that
+in the advanced detector era the vetting process will be automated, so current
+trigger generation and telescope actuation would then dominate latency.
+
+To this end, we have the ambition of reporting candidates not minutes \emph{after} the merger, but seconds \emph{before}.
+By looking for threshold crossings before the signal leaves the detection band, it is possible to trade some signal to noise ratio (\textsc{snr}) for latency.
+Figure \ref{fig:earlywarning} shows projected early trigger rates for NS--NS binaries in Advanced \LIGO\
+assuming the event rate predictions in \cite{Abadie:2010p10836}.
 %

### 2011-05-05 introduction.tex
Put first paragraph back in
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@744 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-%
-\begin{comment}
-The coalescence of compact binary systems consisting of neutron stars (NS)
+\editorial{Would the introduction be more effective without this paragraph?} The coalescence of compact binary systems consisting of neutron stars (NS)
 and/or black holes (BH) is the most promising source of gravitational radiation
@@
 advanced detector era later this decade~\cite{Abadie:2010p10836}.
-\end{comment}
 
-As a binary system loses energy to gravitational waves, its orbital separation
+As a compact binary system loses energy to gravitational waves, its orbital separation
 decreases. This causes a run-away inspiral with the gravitational-wave

### 2011-05-05 introduction.tex
Finished revising introduction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@747 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Although it is technically possible to detect NS--NS signals with the
-\earlywarning\ predicted in figure \ref{fig:earlywarning}, the reduction in
-latency is a challenging task.  As finite-impulse-response (\fir) filtering
-techniques are pushed from high latency configurations to low-latency
-configurations more computational resources are required.  A brute-force
-implementation of low lateny filtering may be too computationally costly to
-implement.  However, there are aspects of searches for binary coalescence that
-can be exploited to offset some of the intrinsic computational complexity: 1)
-The time-frequency evolution of the binary waveforms is such that multi-rate
-filtering may be employed. 2) The redundancy in the filter banks used to cover
-a particular parameter space may be reduced via rank-reduction schemes such as
-the singular value decomposition~\cite{Cannon:2010p10398}. We will use both
-aspects to demonstrate that a very low latency analysis with predictive
-detection of compact binary sources is possible with current computing
-resources.  Assuming other aspects of gravitational-wave observation latency
+Predictive detection of CBCs will require striking a balance between latency and throughput.
+CBC searches consist of banks of matched filters, or cross-correlations between the data stream and a bank of nominal ``template'' signals.
+There are many different implementations of matched filters, but most have high throughput at the cost of high latency, or low latency at the cost of low throughput.
+The former are epitomized by the overlap-save algorithm for \textsc{fft} convolution, currently the preferred method in gravitational wave searches.
+The most obvious example of the latter is the time domain (\textsc{td}) convolution, which is has no latency at all.
+However, its computational complexity is quadratic in the length of the templates, so it is prohibitively expensive for long templates.
+
+Fortunately, the morphology of inspiral signals can be exploited to offset some of the computational complexity of low latency algorithms.
+First, the signals evolve slowly in frequency, so that they can be broken into contiguous bandlimited time intervals and processed at possibly lower sample rates.
+Second, inspiral filter banks consist of highly similar templates, admitting principal component analysis to reduce the number of templates.
+We described a rank reduction scheme based on singular value decomposition in \cite{Cannon:2010p10398}.
+We will use both aspects to demonstrate that a very low latency analysis with predictive detection of compact binary sources is possible with current computing resources.
+Assuming other aspects of gravitational-wave observation latency
 can be reduced significantly, this should allow the possibility for prompt
-alerts to be sent to the Astronomical community.
+alerts to be sent to the astronomical community.
 

### 2011-05-16 results.tex
Cleaned out old results
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@751 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-We generated receiver operating characteristic (\textsc{roc}) curves for a selection of choices for the composite detection statistic threshold and the \textsc{svd} tolerance.
-For each \textsc{roc} curve, or `run', we analyzed mock data with injections to assess detection efficiencies and mock data without injections to evaluate false alarm rates.
-
 \subsection{Detector noise characteristics}
@@
 We tested the new detection method with mock Advanced \textsc{ligo} data having a power spectrum prescribed by the ``zero detuning, high power'' noise model in \cite{Shoemaker:2009p9770}.
-Colored Gaussian noise is generated by passing 5 independent realizations of white Gaussian noise sampled at 16384 Hz through a bank of 5 third-order \textsc{iir} filters, then summing the filters' outputs.
-This carefully designed filter bank reproduces the noise model very faithfully, but since it is composed of \textsc{iir} filters it can produce mock data in realtime very cheaply.
-See \ref{appendix:mock-data} for implementation details.
-\editorial{In this section, we refer to aLIGO run parameters, but the results are actually using initial LIGO noise and initial LIGO parameters.
-This will be fixed.}
-
-\subsection{Injection population}
-
-For each run, we performed about $10^4$ injections into our mock dataset.
-Injections were 2PN post-Newtonian inspiral waveforms with component masses independently and identically distributed in $[1, 3]$ $M_\odot$.
-Injections were tapered at a frequency of 10~Hz where they enter the Advanced \textsc{ligo} detection band.
-This dictated that the longest injection waveform had a duration of 1~777~s.
-In order to guarantee that no two injection waveforms overlapped in the data, the coalescence times of injections were spaced apart by {\color{red} 2~000~$\pm$~100~s}.
-For the sake of economizing disk space, we simultaneously analyzed {\color{red} 16 statistically independent sets} of about $10^3$ injections apiece in the same {\color{red} 14 day} stretch of mock strain data.
-
-Injections were distributed uniformly in log distance.
-We used the following heuristic to provide a reasonably balanced number of distant, marginally detectable signals, and nearby, easily detected signals:
-$$
-d_\mathrm{min} = \frac{1}{2} \cdot \frac{\rho_\mathrm{candle}}{8.0 \cdot \min {d_\mathrm{eff}}}, \quad
-d_\mathrm{max} = \frac{1}{4} \cdot \frac{\rho_\mathrm{candle}}{5.5 \cdot \max {d_\mathrm{eff}}}.
-$$
-Here, $\rho_\mathrm{candle}$ is a fiducially selected \textsc{snr}, and $d_\mathrm{eff}$ is the distance at which a face-on-system would produce a matched filter \textsc{snr} of $\rho_\mathrm{candle}$.
-
-Orientations of injection sources are drawn uniformly, such that the sky location is distributed uniformly in $4\pi$ and the cosine of the binary inclination angle $\iota$ is distributed uniformly in $[0, \pi]$.
-
-\subsection{Noninjection population}
-
-We completed one injection-free pass over the 14~day dataset. \editorial{And **why**?  This part is too terse.}
-
-\subsection{Triggering and data reduction}
-
-For each template with parameters $\theta_i$, a trigger is recorded any time the absolute value of the \textsc{snr}, $|\rho(\theta_i)|$, is greater than a threshold $\rho^* = 5.5$, and is also a maximum over the 10~ms before and after.
-\editorial{This is a proposed trigger criterion that is much simpler than what we currently use.} A trigger is therefore an ordered triple $(t, i, \rho)$, recording the time, template number, and complex \textsc{snr} of an excursion in the outputs of the matched filter bank.
-
-Offline, triggers for all templates collectively are clustered by time, such that within a sliding window of 4 seconds, all triggers except the one with the highest $\rho$ is kept.
-
-For the injection-free passes, all clustered triggers are regarded as false alarms.  We accumulate the \textsc{snr}s of all of the false alarms into the set $\{|\rho|\}_\mathrm{false\,alarm}$.
-
-For each injection, we check whether there was any clustered trigger that was within 2 seconds of the injection's time.
-If there was, then the injection is considered to be recovered at the \textsc{snr} of that trigger.
-Otherwise, the injection was not recovered because the \textsc{snr} threshold $\rho^*$ was not met, so we consider the injection to be recovered only at an \textsc{snr} of $\infty$.
-From all found injections, we accumulate the injected distance and the recovered \textsc{snr} into the set $\{(d, |\rho|)\}_\mathrm{found}$.
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=0.4]{figures/roc_99.pdf}
-\includegraphics[scale=0.4]{figures/roc_9995.pdf}
-\caption{Receiver operating characteristic (\textsc{roc}) curve of detection efficiency (\textsc{eff}) versus false alarm probability (\textsc{fap}).}
-\label{fig:roc}
-\end{center}
-\end{figure}

### 2011-05-16 method.tex
Rearranged some section headings
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@752 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 multiband decomposition of the compact binary parameter space that resulted in
-search with $\sim$minutes latency in LIGO's S6 and Virgo's VSR2 science runs.
+search with $\sim$minutes latency in \LIGO's S6 and Virgo's VSR2 science runs.
 We combine this with the orthogonal decomposition described
 in~\cite{Cannon:2010p10398} that exploits the redundancy of the template banks.
+
+\subsection{Conventional \CBC{} matched filter searches}
 
@@
 against the whitened detector data.  A sufficient number of templates are
-chosen to assure a minimal loss of SNR~\cite{Owen:1995tm,Owen:1998dk}.  To
+chosen to assure a minimal loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}.  To
 construct a template bank, matched filters are chosen with discrete signal
@@
 
+\subsection{Proposed method}
+
 In order to minimize latency we propose using the time domain convolution
@@
 
-\subsection{Selectively reducing the sample rate of the data and template waveforms}
+\subsubsection{Selectively reducing the sample rate of the data and template waveforms}
 
@@
 
-\subsection{Reducing the number of filters with the singular value
+\subsubsection{Reducing the number of filters with the singular value
 decomposition}

### 2011-05-17 introduction.tex
Added a new paragraph on optical flashes. Still needs some work.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@754 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 counterpart in the system's final moments as a binary~\cite{shibata:2007}.
-In order to observe the prompt and
-potentially most intense electromagnetic emission, telescopes must point during the seconds surrounding the
-merger.
 
-\editorial{Citation needed for LOOC-UP} To make this possible, the gravitational-wave community initiated a project to send alerts when
+Prompt electromagnetic emission can arise as shells of relativistically
+outflowing matter collide in the inner shock. Such an inner shock from a
+compact binary coalescence is believed to be the mechanism of short gamma-ray
+bursts~\cite{nakar07}. This prompt emission is a probe into the extreme initial
+conditions of the outflow. A very bright optical flash should accompany the
+x-rays and gamma-rays~\cite{Sari99}.
+\editorial{Should mention that optical flashes have been seen for a handful of long GRBs by very fast slews.
+TAROT has seen many within ~20 seconds.
+\cite{Klotz09}} In contrast, an
+afterglow is produced when the outflow impacts the interstellar medium.
+Afterglows represent lower Lorentz factors and the conditions of the
+environment far outside of the central engine, and are relatively insensitive
+to initial conditions. In order to observe the most intense electromagnetic
+emission and provide an unbroken chain of observations between the progenitor
+system and the expanding fireball, telescopes must point during the seconds
+surrounding the merger.
+
+\editorial{Citation needed for LOOC-UP} The gravitational-wave community initiated a project to send alerts when
 potential gravitational-wave transients are observed.  In October 2010, \textsc{ligo} completed its sixth science run (S6) and Virgo
@@
 \caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
-be detectable $T_{bc}$ seconds before coalescence.  The solid line is the most
+be detectable $T_\mathrm{bc}$ seconds before coalescence.  The solid line is the most
 likely yearly rate estimate $N_{\mathrm{re}}$ for Advanced \LIGO\ and the
@@
 
-Predictive detection of CBCs will require striking a balance between latency and throughput.
-CBC searches consist of banks of matched filters, or cross-correlations between the data stream and a bank of nominal ``template'' signals.
+Predictive detection of \CBC{}s will require striking a balance between latency and throughput.
+\CBC{} searches consist of banks of matched filters, or cross-correlations between the data stream and a bank of nominal ``template'' signals.
 There are many different implementations of matched filters, but most have high throughput at the cost of high latency, or low latency at the cost of low throughput.

### 2011-05-22 conclusions.tex
Added conclusion outline
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@745 ccdd05d5-56c1-461a-bf98-85f6876523b4
added some text to the conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@755 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{SECV}\label{sec:conclusions}
+
+We have demonstrated a computationally feasible procedure for the rapid
+detection of gravitational waves emitted during the coalescence of neutron
+stars and stellar mass black holes.  These sources are expected to produce
+prompt electromagnetic signals and may be the progenitors of some short-hard
+gamma ray bursts.  Rapid alerts to the broader astronomical community may
+improve the chances of detecting an electromagnetic counterpart in bands from
+X-ray down to radio.  We antipate requiring no more than \numcpus\ modern
+computer cores to analyze a four detector network of gravitational wave data
+for such systems assuming spin effects can be ignored in parameterizing the
+expected signals ~\cite{the papers that say that is okay}.  This is within the
+current computing capabilities of the LSC data grid\cite{LDG}.
+
+The algorithm we described has no intrinsic latency.  However, there are
+fundamental and practical latencies associated with the analysis and detection
+procedure. For example, the \LIGO detectors, data acquisition is synchronized
+to a 1/16 Hz cadence introducing an up-front latency of 125ms.  Data
+aggregation from the observatories will travel over various networks, each
+capable of high bandwith but perhaps only modest latency.  This could amount to
+a similar latency of $\sim$ 100ms.  Lastly, unless a realtime infrastructure is
+adopted post data acquisition, it is likely that there will be an inherent
+latency introduced by such infrastructure.  We have shown a prototype
+implementation using \gstlal\ that is capable of ~$\sim$ 1s latency. In our
+opinion, significant work would have to be done in order to improve upon this
+number. However, it should be considered for third generation detector design.
+For example, a tighter integration of analysis and data acquisition would be
+beneficial.
+
+We have ommitted discussion of source localization.  Localization is known to
+be poor for signals that are detected before coalescence~\cite{fairhurst}.
+However, one should not immediately dismiss the practical usefullness of a
+poorly localized source.  Even with poor localization, it should be possible to
+begin downselecting what observatories could view a potential signal and for
+such observatories to begin any necessary prerequisite activities.  In future
+works we wil explore more rigorously the pointing prospects with realistic
+simulations using the infrastructure and techniques described in this work.
+
+Latency budget, including `before' and `after' quotes for:
+
+\begin{itemize}
+\item Data acquisition
+\item Calibration
+\item Data aggregation
+\item Analysis
+\item Localization
+\item Alert
+\item Telescope actuation
+\item Total
+\end{itemize}
+
+Future work:
+
+\begin{itemize}
+\item Sub-solar mass search
+\item Hierarchical detection
+\end{itemize}

### 2011-05-22 conclusions.tex
fixed broken citation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@756 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 for such systems assuming spin effects can be ignored in parameterizing the
-expected signals ~\cite{the papers that say that is okay}.  This is within the
+expected signals ~\cite{the-papers-that-say-that-is-okay}.  This is within the
 current computing capabilities of the LSC data grid\cite{LDG}.

### 2011-05-22 introduction.tex
formatting to 80 columns for easier diffs
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@757 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\editorial{Would the introduction be more effective without this paragraph?} The coalescence of compact binary systems consisting of neutron stars (NS)
+\editorial{Would the introduction be more effective without this paragraph?}
+The coalescence of compact binary systems consisting of neutron stars (NS)
 and/or black holes (BH) is the most promising source of gravitational radiation
-for Advanced \LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb}, and
-\LCGT~\cite{LCGTWeb}.  Tens of
-binary coalescence events are expected to be observed in the
-advanced detector era later this decade~\cite{Abadie:2010p10836}.
+for Advanced \LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb},
+and \LCGT~\cite{LCGTWeb}.  Tens of binary coalescence events are expected to be
+observed in the advanced detector era later this
+decade~\cite{Abadie:2010p10836}.
 
-As a compact binary system loses energy to gravitational waves, its orbital separation
-decreases. This causes a run-away inspiral with the gravitational-wave
-amplitude and frequency increasing until the system eventually merges near the
-innermost stable circular orbit (\ISCO).  If a neutron star is involved it may become tidally disrupted near the merger.
-This disrupted matter can fuel a bright electromagnetic
-counterpart in the system's final moments as a binary~\cite{shibata:2007}.
+As a compact binary system loses energy to gravitational waves, its orbital
+separation decreases. This causes a run-away inspiral with the
+gravitational-wave amplitude and frequency increasing until the system
+eventually merges near the innermost stable circular orbit (\ISCO).  If a
+neutron star is involved it may become tidally disrupted near the merger.  This
+disrupted matter can fuel a bright electromagnetic counterpart in the system's
+final moments as a binary~\cite{shibata:2007}.
 
@@
 conditions of the outflow. A very bright optical flash should accompany the
-x-rays and gamma-rays~\cite{Sari99}.
-\editorial{Should mention that optical flashes have been seen for a handful of long GRBs by very fast slews.
-TAROT has seen many within ~20 seconds.
-\cite{Klotz09}} In contrast, an
-afterglow is produced when the outflow impacts the interstellar medium.
-Afterglows represent lower Lorentz factors and the conditions of the
-environment far outside of the central engine, and are relatively insensitive
-to initial conditions. In order to observe the most intense electromagnetic
-emission and provide an unbroken chain of observations between the progenitor
-system and the expanding fireball, telescopes must point during the seconds
-surrounding the merger.
+x-rays and gamma-rays~\cite{Sari99}. \editorial{Should mention that optical
+flashes have been seen for a handful of long GRBs by very fast slews. TAROT has
+seen many within ~20 seconds. \cite{Klotz09}} In contrast, an afterglow is
+produced when the outflow impacts the interstellar medium.  Afterglows
+represent lower Lorentz factors and the conditions of the environment far
+outside of the central engine, and are relatively insensitive to initial
+conditions. In order to observe the most intense electromagnetic emission and
+provide an unbroken chain of observations between the progenitor system and the
+expanding fireball, telescopes must point during the seconds surrounding the
+merger.
 
-\editorial{Citation needed for LOOC-UP} The gravitational-wave community initiated a project to send alerts when
-potential gravitational-wave transients are observed.  In October 2010, \textsc{ligo} completed its sixth science run (S6) and Virgo
-completed its third science run (VSR3).  While both \textsc{ligo} detectors and Virgo were
-operating, several all-sky detection pipelines operated in a low-latency
-configuration, namely \textsc{mbta}, ihope, Coherent WaveBurst, and Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
-\editorial{Get references for these low-latency pipelines.} The S6 analysis achieved latencies of 30--60 minutes, which were
-dominated by a human vetting process. Candidates were sent for
-electromagnetic followup to several telescopes; Swift,
-\textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008, HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved the best gravitational-wave trigger
-generation latencies of 2--5 minutes.  We assume that
-in the advanced detector era the vetting process will be automated, so current
-trigger generation and telescope actuation would then dominate latency.
+\editorial{Citation needed for LOOC-UP} The gravitational-wave community
+initiated a project to send alerts when potential gravitational-wave transients
+are observed.  In October 2010, \textsc{ligo} completed its sixth science run
+(S6) and Virgo completed its third science run (VSR3).  While both
+\textsc{ligo} detectors and Virgo were operating, several all-sky detection
+pipelines operated in a low-latency configuration, namely \textsc{mbta}, ihope,
+Coherent WaveBurst, and Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
+\editorial{Get references for these low-latency pipelines.} The S6 analysis
+achieved latencies of 30--60 minutes, which were dominated by a human vetting
+process. Candidates were sent for electromagnetic followup to several
+telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008,
+HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved
+the best gravitational-wave trigger generation latencies of 2--5 minutes.  We
+assume that in the advanced detector era the vetting process will be automated,
+so current trigger generation and telescope actuation would then dominate
+latency.
 
-To this end, we have the ambition of reporting candidates not minutes \emph{after} the merger, but seconds \emph{before}.
-By looking for threshold crossings before the signal leaves the detection band, it is possible to trade some signal to noise ratio (\textsc{snr}) for latency.
-Figure \ref{fig:earlywarning} shows projected early trigger rates for NS--NS binaries in Advanced \LIGO\
-assuming the event rate predictions in \cite{Abadie:2010p10836}.
+To this end, we have the ambition of reporting candidates not minutes
+\emph{after} the merger, but seconds \emph{before}.  By looking for threshold
+crossings before the signal leaves the detection band, it is possible to trade
+some signal to noise ratio (\textsc{snr}) for latency.  Figure
+\ref{fig:earlywarning} shows projected early trigger rates for NS--NS binaries
+in Advanced \LIGO\ assuming the event rate predictions in
+\cite{Abadie:2010p10836}.
 %
@@
 
-Predictive detection of \CBC{}s will require striking a balance between latency and throughput.
-\CBC{} searches consist of banks of matched filters, or cross-correlations between the data stream and a bank of nominal ``template'' signals.
-There are many different implementations of matched filters, but most have high throughput at the cost of high latency, or low latency at the cost of low throughput.
-The former are epitomized by the overlap-save algorithm for \textsc{fft} convolution, currently the preferred method in gravitational wave searches.
-The most obvious example of the latter is the time domain (\textsc{td}) convolution, which is has no latency at all.
-However, its computational complexity is quadratic in the length of the templates, so it is prohibitively expensive for long templates.
+Predictive detection of \CBC{}s will require striking a balance between latency
+and throughput.  \CBC{} searches consist of banks of matched filters, or
+cross-correlations between the data stream and a bank of nominal ``template''
+signals.  There are many different implementations of matched filters, but most
+have high throughput at the cost of high latency, or low latency at the cost of
+low throughput.  The former are epitomized by the overlap-save algorithm for
+\textsc{fft} convolution, currently the preferred method in gravitational wave
+searches.  The most obvious example of the latter is the time domain
+(\textsc{td}) convolution, which is has no latency at all.  However, its
+computational complexity is quadratic in the length of the templates, so it is
+prohibitively expensive for long templates.
 
-Fortunately, the morphology of inspiral signals can be exploited to offset some of the computational complexity of low latency algorithms.
-First, the signals evolve slowly in frequency, so that they can be broken into contiguous bandlimited time intervals and processed at possibly lower sample rates.
-Second, inspiral filter banks consist of highly similar templates, admitting principal component analysis to reduce the number of templates.
-We described a rank reduction scheme based on singular value decomposition in \cite{Cannon:2010p10398}.
-We will use both aspects to demonstrate that a very low latency analysis with predictive detection of compact binary sources is possible with current computing resources.
-Assuming other aspects of gravitational-wave observation latency
-can be reduced significantly, this should allow the possibility for prompt
-alerts to be sent to the astronomical community.
+Fortunately, the morphology of inspiral signals can be exploited to offset some
+of the computational complexity of low latency algorithms.  First, the signals
+evolve slowly in frequency, so that they can be broken into contiguous
+bandlimited time intervals and processed at possibly lower sample rates.
+Second, inspiral filter banks consist of highly similar templates, admitting
+principal component analysis to reduce the number of templates.  We described a
+rank reduction scheme based on singular value decomposition in
+\cite{Cannon:2010p10398}.  We will use both aspects to demonstrate that a very
+low latency analysis with predictive detection of compact binary sources is
+possible with current computing resources.  Assuming other aspects of
+gravitational-wave observation latency can be reduced significantly, this
+should allow the possibility for prompt alerts to be sent to the astronomical
+community.
 
 The paper is organized as follows. First we provide an overview of our method
-for detecting compact binary coalescence signals in an \earlywarning\
-analysis. We then describe the pipeline we have constructed that implements
-our method.  To validate the approach we present results of simulations and
-conclude with some remarks on what remains to prepare for the Advanced
-detector era.
+for detecting compact binary coalescence signals in an \earlywarning\ analysis.
+We then describe the pipeline we have constructed that implements our method.
+To validate the approach we present results of simulations and conclude with
+some remarks on what remains to prepare for the Advanced detector era.

### 2011-05-22 introduction.tex
added reference, editorial
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@758 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 outflowing matter collide in the inner shock. Such an inner shock from a
-compact binary coalescence is believed to be the mechanism of short gamma-ray
-bursts~\cite{nakar07}. This prompt emission is a probe into the extreme initial
-conditions of the outflow. A very bright optical flash should accompany the
-x-rays and gamma-rays~\cite{Sari99}. \editorial{Should mention that optical
-flashes have been seen for a handful of long GRBs by very fast slews. TAROT has
-seen many within ~20 seconds. \cite{Klotz09}} In contrast, an afterglow is
-produced when the outflow impacts the interstellar medium.  Afterglows
-represent lower Lorentz factors and the conditions of the environment far
-outside of the central engine, and are relatively insensitive to initial
-conditions. In order to observe the most intense electromagnetic emission and
-provide an unbroken chain of observations between the progenitor system and the
-expanding fireball, telescopes must point during the seconds surrounding the
-merger.
+compact binary coalescence is believed to be a mechanism for short gamma-ray
+bursts~\cite{Lee:2005, nakar07}. This prompt emission is a probe into the
+extreme initial conditions of the outflow. A very bright optical flash should
+accompany the x-rays and gamma-rays~\cite{Sari99}. \editorial{Should mention
+that optical flashes have been seen for a handful of long GRBs by very fast
+slews. TAROT has seen many within ~20 seconds. \cite{Klotz09}} In contrast, an
+afterglow is produced when the outflow impacts the interstellar medium.
+Afterglows represent lower Lorentz factors and the conditions of the
+environment far outside of the central engine, and are relatively insensitive
+to initial conditions. In order to observe the most intense electromagnetic
+emission and provide an unbroken chain of observations between the progenitor
+system and the expanding fireball, telescopes must point during the seconds
+surrounding the merger.
 
@@
 Coherent WaveBurst, and Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
-\editorial{Get references for these low-latency pipelines.} The S6 analysis
+\editorial{Get references for these low-latency pipelines.} The S6 analyses
 achieved latencies of 30--60 minutes, which were dominated by a human vetting
@@
 To this end, we have the ambition of reporting candidates not minutes
-\emph{after} the merger, but seconds \emph{before}.  By looking for threshold
-crossings before the signal leaves the detection band, it is possible to trade
-some signal to noise ratio (\textsc{snr}) for latency.  Figure
-\ref{fig:earlywarning} shows projected early trigger rates for NS--NS binaries
-in Advanced \LIGO\ assuming the event rate predictions in
+\emph{after} the merger, but seconds \emph{before}.  \editorial{I think we
+should say also something about how we can also get the full SNR within moments
+after the coalescence, it is obvious maybe, but we should be explicit}  By
+looking for threshold crossings before the signal leaves the detection band, it
+is possible to trade some signal to noise ratio (\textsc{snr}) for latency.
+Figure \ref{fig:earlywarning} shows projected early trigger rates for NS--NS
+binaries in Advanced \LIGO\ assuming the event rate predictions in
 \cite{Abadie:2010p10836}.

### 2011-05-23 conclusions.tex
Small cleanups and added Alessandra to acknowledgements
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@760 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 detection of gravitational waves emitted during the coalescence of neutron
-stars and stellar mass black holes.  These sources are expected to produce
-prompt electromagnetic signals and may be the progenitors of some short-hard
-gamma ray bursts.  Rapid alerts to the broader astronomical community may
+stars and stellar-mass black holes.  These sources are expected to produce
+prompt electromagnetic signals and may be the progenitors of some short hard
+gamma-ray bursts.  Rapid alerts to the broader astronomical community may
 improve the chances of detecting an electromagnetic counterpart in bands from
 X-ray down to radio.  We antipate requiring no more than \numcpus\ modern
-computer cores to analyze a four detector network of gravitational wave data
+computer cores to analyze a four-detector network of gravitational-wave data
 for such systems assuming spin effects can be ignored in parameterizing the
-expected signals ~\cite{the-papers-that-say-that-is-okay}.  This is within the
-current computing capabilities of the LSC data grid\cite{LDG}.
+expected signals~\cite{the-papers-that-say-that-is-okay}.  This is within the
+current computing capabilities of the \textsc{lsc} Data Grid~\cite{LDG}.
 
@@
 fundamental and practical latencies associated with the analysis and detection
-procedure. For example, the \LIGO detectors, data acquisition is synchronized
-to a 1/16 Hz cadence introducing an up-front latency of 125ms.  Data
+procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
+to a 1/16 Hz cadence introducing an up-front latency of 125 ms.  Data
 aggregation from the observatories will travel over various networks, each
 capable of high bandwith but perhaps only modest latency.  This could amount to
-a similar latency of $\sim$ 100ms.  Lastly, unless a realtime infrastructure is
+a similar latency of $\sim$ 100 ms.  Lastly, unless a realtime infrastructure is
 adopted post data acquisition, it is likely that there will be an inherent
 latency introduced by such infrastructure.  We have shown a prototype
-implementation using \gstlal\ that is capable of ~$\sim$ 1s latency. In our
+implementation using \gstlal\ that is capable of $\sim$1 s latency. In our
 opinion, significant work would have to be done in order to improve upon this
@@
 
-We have ommitted discussion of source localization.  Localization is known to
-be poor for signals that are detected before coalescence~\cite{fairhurst}.
-However, one should not immediately dismiss the practical usefullness of a
+We have omitted discussion of source localization. Localization is known to
+be poor for signals of low \SNR~\cite{Fairhurst2009}.
+However, one should not immediately dismiss the practical usefulness of a
 poorly localized source. Even with poor localization, it should be possible to

### 2011-05-23 inspiral_svd.tex
Small cleanups and added Alessandra to acknowledgements
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@760 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We demonstrate a computationally practical analysis strategy that produces early warning triggers even before gravitational radiation from the final merger has arrived at the detectors.
-With current rate estimates for the Advanced \textsc{ligo} design configuration, we should detect $\sim$10 sources earlier than 10 seconds before merger in 1 year of livetime.
+With current rate estimates for the Advanced \LIGO{} design configuration, we should detect $\sim$10 sources earlier than 10 seconds before merger in 1 year of livetime.
 \end{abstract}
@@
 many discussions about rate estimates and the prospects of early detection.
+NF would like to thank Alessandra Corsi for illuminating discussions on
+astronomical motivations.
 

### 2011-05-27 method.tex
misc section 2 cleanups and rewordings
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@761 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In this section we describe a decomposition of the compact binary parameter
-space that reduces low latency filtering cost sufficiently to allow for the
+space that reduces low-latency filtering cost sufficiently to allow for the
 possibility of \earlywarning\ detection with modest computing requirements.  We
@@
 multiband decomposition of the compact binary parameter space that resulted in
-search with $\sim$minutes latency in \LIGO's S6 and Virgo's VSR2 science runs.
+search with $\sim$minutes latency in \LIGO{}'s S6 and Virgo's VSR2 science runs.
 We combine this with the orthogonal decomposition described
@@
 $\bar{\theta}$ that determine the amplitude and phase evolution of the observed
-binary signals.  Searches for gravitational waves from compact binary
-coalescence typically employ matched filter banks, called template banks,
-\cite{findchirppaper} that discretely sample the possible intrinsic parameters.
-The filter models of the waveforms, known as templates are weighted by the
-detector noise amplitude spectral density, $x(t,\bar{\theta})$ and filtered
-against the whitened detector data.  A sufficient number of templates are
-chosen to assure a minimal loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}.  To
-construct a template bank, matched filters are chosen with discrete signal
-parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$, such that any possible
-signal will have a cross-correlation of $\geq0.97$ with at least one template.
-Such a template bank is said to have a 97\% {\em minimum match}. For systems
-where the effects of spin can be ignored, the intrinsic parameters are the
-component masses of the binary, $\bar{\theta} = (m_1, m_2)$.
-
-The procedure for choosing a bank of matched filters for gravitational wave
-searches is known~\cite{Owen:1995tm}. Therefore, here we will assume that a
-bank of filters has already been chosen based on the targetted parameter space
-and we will denote the $i^{\mathrm th}$ filter with parameters $\bar{\theta_i}$
-as a function of time $x_i(\tau)$.%
+binary signals. For systems where the effects of spin can be ignored, the
+intrinsic parameters are the component masses of the binary, $\bar{\theta} =
+(m_1, m_2)$. Searches for gravitational waves from compact binary coalescence
+typically employ matched filter banks, called template banks, that discretely
+sample the possible intrinsic parameters~\cite{findchirppaper}.  The filters
+for the waveforms $x(t,\bar{\theta})$, known as templates, are the waveforms
+weighted by the inverse detector noise amplitude spectral density, so as to
+whiten them. To construct a template bank, templates are chosen with discrete
+signal parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$ to assure a
+bounded loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}. That is, any possible
+signal within the search space will have a cross-correlation of $\geq0.97$ with
+at least one template. Such a template bank is said to have a 97\% {\em
+minimum match}. Data are filtered against each template to produce an \SNR
+time-series. Local peak-finding across time and templates determines detection
+candidates.
 %
-\footnote{There are two gravitational wave polarizations, $+$ and $\times$. A
+\footnote{There are two gravitational-wave polarizations, $+$ and $\times$. A
 given detector will observe a combination of these polarizations that will
@@
 maximized over by filtering for quadrature phases and taking the magnitude of
-the result. For simplicity we will ignore that aspect in this work as it is
-straightforward to generalize, but not necessary for understanding any of the
+the result. For simplicity we will ignore that aspect in this work, as it is
+straightforward to generalize but not necessary for understanding any of the
 points that will be made.}
-%
+
+We will denote the $i^\mathrm{th}$ filter with parameters $\bar{\theta_i}$
+as a function of time $x_i(\tau)$.
 In this work we will discuss transformations to a set of filters
@@
 over the entire parameter space.  For that reason we assume from here onward
-that the set of filters $\{x_i(\tau)\}$ refer to a set of near neighbor filters
+that the set of filters $\{x_i(\tau)\}$ refer to a set of near-neighbor filters
 that can be chosen as a subset of the full parameter space.  Several such
 subsets can be chosen until all of the filters are a member of one local set.
+\editorial{nvf: I don't think that the splitting of the bank is a crucial point here. Perhaps later if we discuss the up-front cost of the SVD.}
 
@@
 The evaluation of the integrals in \eqref{eq:SNRTD} and \eqref{eq:SNRFD} are
-implemented as sums over sample points for the digitized gravitational wave
+implemented as sums over sample points for the digitized gravitational-wave
 detector output.  Discrete Fourier transforms can be computed efficiently
@@
 \hoftsamps}.  In most cases \tmpsamps\ $\gg$ $\log$ \hoftsamps\ and the
-computational savings by choosing the frequency domain integral form
+computational savings by choosing the frequency-domain integral form
 \eqref{eq:SNRFD} is clear.  However, to take full advantage of the
@@
 
-In order to minimize latency we propose using the time domain convolution
-presented in \eqref{eq:SNRTD}.  However, because the brute force evaluation of
+In order to minimize latency we propose using the time-domain convolution
+presented in \eqref{eq:SNRTD}.  However, because the brute-force evaluation of
 \eqref{eq:SNRTD} is far too costly to be useful, we will consider an
@@
 cost of \eqref{eq:SNRTD} sufficiently to feasibly search for
-gravitational-waves from compact binary coalescence in \realtime.  This
-requires 1) exploiting the redundancy of the template bank and 2) exploiting
-the time-frequency characteristics of the binary waveforms.  We describe our
+gravitational waves from compact binary coalescence in \realtime.  This
+requires 1) exploiting
+the time-frequency characteristics of the binary waveforms and 2) exploiting the redundancy of the template bank. We describe our
 procedure for producing the decompostion in \eqref{eq:decomp} in the remainder
@@
 The first step of the orthogonal decomposition described in \eqref{eq:decomp}
-is to divide the templates into \emph{time slices}.  This is a time domain
-analogue to the frequency domain decomposition described in ~\cite{Marion2004,
+is to divide the templates into \emph{time slices}.  This is a time-domain
+analogue to the frequency-domain decomposition described in ~\cite{Marion2004,
 Buskulic2010, beauville2006, beauville2008}.  A matched filter is constructed
-for each time slice.  The outputs form an ensemble of partial \textsc{snr}
+for each time slice.  The outputs form an ensemble of partial \SNR{}
 streams.  By linearity, these partial \SNR\ streams can be suitably time
@@
 Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
-M_\odot / c^3 \approx 5 \mu\mathrm{s}$) and $t$ is the time relative to the
+M_\odot / c^3 \approx 5 \umu\mathrm{s}$) and $t$ is the time relative to the
 coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
@@
 The monotonic time-frequency relationship of \eqref{eq:fgw} allows us to choose
-time slice boundaries that require substantially less bandwidth at early times
+time-slice boundaries that require substantially less bandwidth at early times
 in the inspiral.  Our goal is to reduce the filtering cost of a large fraction
@@
 t_\mathrm{ISCO}$, $f_{k-1} \geqslant 2 f_\mathrm{low}$, and $t_k \leqslant
-t_\mathrm{low}$. An example time slice design satisfying these constraints for
+t_\mathrm{low}$. An example time-slice design satisfying these constraints for
 a $1.4 - 1.4 \, M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
@@
 
-Rather than applying a unique time slice decomposition to each template
+Rather than applying a unique time-slice decomposition to each template
 waveform, we find a decomposition that is adequate for the entire set
 $\{x_i\}$.  This is facilated by choosing templates in the set $\{x_i\}$ with
-similar chirp masses, $\mathcal{M}$.  The time slice decompositions of the
+similar chirp masses, $\mathcal{M}$.  The time-slice decompositions of the
 filters $x_i(\tau)$ lead to new filters $y_{ik}(\tau)$ satisfying
@@
 showed that to high reconstruction accuracy far fewer filters are needed than
-were in the original template bank. We find that when combined with the time
-slice decomposition, the number of \SVD\ filters, \numsvdtmps\ is much smaller
+were in the original template bank. We find that when combined with the
+time-slice decomposition, the number of \SVD\ filters, \numsvdtmps\ is much smaller
 than the original number of filters \numtmps.  We combine \eqref{eq:svddecomp}
@@
 section we compute the expected computational cost scaling of this
-decomposition and compare it with the brute force implementation in
+decomposition and compare it with the brute-force implementation in
 \eqref{eq:SNRTD} and higher latency \fft\ methods.
@@
 
-The filter bank can be implemented using finite impulse response \fir\ filters,
+The filter bank can be implemented using finite impulse response (\fir{}) filters,
 which are just sliding window dot products.  If there are \numtmps\ templates
@@
 orthogonal filters $u_{jk}$.  The second sum is the cost of resampling the
-reconstructed time slice outputs to the original sample rate {\fmax}.  It
+reconstructed time-slice outputs to the original sample rate {\fmax}.  It
 assumes a \fir\ filter with \resampsamps\ sample points.  Note that the cost
@@
 \editorial{It would be good to illustrate the layout of this particular
-template bank: masses spanned, time slice layout \dots}
+template bank: masses spanned, time-slice layout \dots}
 \begin{table}[!h]

### 2011-05-27 figures/diagram.tex
Added diagram
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@763 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass{amsart}
\usepackage[usenames]{color}
\usepackage{tikz}
\usepackage{signalflowdiagram}
\usetikzlibrary{fit,calc,chains,scopes}
\begin{document}

\tikzstyle{junction}=[draw, fill, circle, line width=0pt, inner sep=0pt, minimum size=3pt]
\tikzstyle{signal}=[draw, fill=white, circle, inner sep=0pt, minimum size=4pt]
\tikzstyle{resample}=[draw, circle, line width=1pt, inner sep=0pt, minimum size=0.3cm]

\begin{figure}
	\begin{center}
		\begin{tikzpicture}[
			node distance=2mm,
			every path/.style={line width=1pt},
			every node/.style={line width=1pt}
		]
			\foreach \x in {0,1,2} {
				\node[draw] at (0, {\x == 0 ? 0cm : (\x == 1 ? -3cm : -7.5cm)}) (u0) {};
				\node[draw, below=of u0] (u1) {};
				\node[below=of u1] (uellip\x) {$\vdots$};
				\node[draw, below=of uellip\x] (u2) {};

				\node[junction] at ($0.5*(u0) + 0.5*(u2) - (0.5, 0)$) (in\x tee) {};
				\node[resample, left=1.5cm of in\x tee] (in\x) {\tiny$\downarrow$};

				\foreach \i in {0,1,2} {
					\draw (in\x tee) |- (u\i);
					\node[junction, right=1cm of u\i] (tee u\i) {};
					\draw (u\i) -- (tee u\i);
				}

				\node[adder, right=2cm of tee u0] (v0\x) {};
				\node[adder, below=of v0\x] (v1\x) {};
				\node[adder, below=of v1\x] (v2\x) {};
				\node[right=0.85cm of uellip\x] {$\vdots$};
				\node[adder, right=2cm of tee u2] (v3\x) {};
				\node[above=0cm of v3\x] {$\vdots$};

				\foreach \i in {0,1,2}
					\foreach \j in {0,1,2,3}
						\draw (tee u\i) -- (v\j\x);

				\node[draw,dashed,line width=0.5pt,fit=(in\x tee) (u0) (u2)] (ortho\x) {};
				\node[draw,dashed,line width=0.5pt,fit=(tee u0) (tee u2) (v0\x) (v3\x)] (reconstruct\x) {};
			}

			\foreach \x in {1,2} {
				\node[resample, right=1cm of v0\x, yshift=0.5cm] (resample0\x) {\tiny$\uparrow$};
				\node[resample, right=1.5cm of v0\x, yshift=0.5cm] (resample1\x) {\tiny$\uparrow$};
				\node[resample, right=2cm of v0\x, yshift=0.5cm] (resample2\x) {\tiny$\uparrow$};
				\node[right=2.5cm of v0\x, yshift=0.5cm] (dots) {$\cdots$};
				\node[below=of dots] (dots) {$\cdots$};
				\node[below=of dots] (dots) {$\cdots$};
				\node[resample, right=3.5cm of v0\x, yshift=0.5cm] (resample3\x) {\tiny$\uparrow$};
			}

			\foreach \x in {0,1} {
				\node[adder, right=1cm of v0\x] (adder0\x) {};
				\node[adder, right=1.5cm of v1\x] (adder1\x) {};
				\node[adder, right=2cm of v2\x] (adder2\x) {};
				\node[adder, right=3.5cm of v3\x] (adder3\x) {};
				\foreach \i in {0,1,2,3} {
					\draw[->] (v\i \x) -- (adder\i \x);
				}
			}

			\foreach \i in {0,1,2,3} {
				\node[signal, right=4.5cm of v\i 0, label=right:$\rho_\i$] (out\i) {};
				\draw[->] (v\i 2) -| (resample\i 2);
				\draw[->] (resample\i 2) -- (adder\i 1) node[above=0.5cm of resample\i 2,fill=white] {$\vdots$};
				\draw[->] (adder\i 1) -- (resample\i 1);
				\draw[->] (resample\i 1) -- (adder\i 0);
				\draw[->] (adder\i 0) -- (out\i);
			}

			\node[draw, left=0.5cm of in0] (whiten) {\emph{whiten}};
			\node[signal, label=left:$x$, label=below:{\small 16384 Hz}, left=0.75cm of whiten] (x) {};
			\node[below=1.5cm of in1] (ellip) {$\vdots$};
			\node[right=1.6cm of ellip] {$\vdots$};
			\node[right=4.1cm of ellip] {$\vdots$};
			\node[above=1mm of out3] {\vdots};
			\draw[->] (x) -- (whiten);
			\draw[->] (whiten) -- (in0);
			\draw[->] (in0) -- (in1);
			\draw (in1) -- (ellip);
			\draw[->] (ellip) -- (in2);

			\draw (in0) -- (in0tee) node[midway,above] {\small 4096 Hz};
			\draw (in1) -- (in1tee) node[midway,above] {\small 512 Hz};
			\draw (in2) -- (in2tee) node[midway,above] {\small 32 Hz};

			\node[above=0.5cm of ortho0,align=center] (fir label) {Orthogonal \\ FIR filters \\ $\overbrace{\hspace{1cm}}^{}$};
			\node[right=0.1cm of fir label,align=center] {Reconstruction \\ matrices \\ $\overbrace{\hspace{2.75cm}}^{}$};
			\node[right=3.5cm of fir label,align=center] {Upsampling and \\ SNR accumulation \\ $\overbrace{\hspace{2.75cm}}^{}$};
			\node[left=0.1cm of fir label,align=center] {Down-\\ sampling \\ $\overbrace{\hspace{1cm}}^{}$};
			\node[left=1.75cm of fir label,align=center] {Data acquisition \\ and whitening \\ $\overbrace{\hspace{2cm}}^{}$};

		\end{tikzpicture}
		\caption{Schematic of LLOID pipeline illustrating signal flow.  Circles with arrows represent upsampling \protect\tikz{\protect \node[resample] {\tiny$\uparrow$};} or downsampling \protect\tikz{\protect \node[resample] {\tiny$\downarrow$};}.  Circles with plus signs represent summing junctions \protect\tikz{\protect \node[line width=1pt,adder] {\tiny$\uparrow$};}.  Squares \protect\tikz{\protect \node[line width=1pt,draw] {};} stand for FIR filters.  Sample rate decreases from the top of the diagram to the bottom.}
	\end{center}
\end{figure}

This diagram is a work in progress.  Some corrections to make:

{\color{red}
\begin{itemize}
\item Add delay lines.
\item Switch first downsampler and whitener.
\item Resamplers are connected in series, not parallel.
\item Translate figure to the left to align it visually with the caption.
\end{itemize}
}

\end{document}

### 2011-05-27 analysis.tex
fixme at the top
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@710 ccdd05d5-56c1-461a-bf98-85f6876523b4
caption
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@717 ccdd05d5-56c1-461a-bf98-85f6876523b4
added flop budget
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@740 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added new diagram to paper
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@764 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \label{SECIII}\label{sec:analysis}
+
+\FIXME{
+I would like to see this section have more of a description of the
+pipeline and a figure representing it.
+}
 
@@
 
-\begin{comment}
-
-% Ugly diagram!
-
-\begin{figure*}[h!]
-\begin{center}
-\includegraphics[width=1.0\textwidth]{figures/flow_chart.pdf}
-\caption{\label{f:flowchart} 
-}
-\end{center}
-\end{figure*}
-
-\end{comment}
+\begin{figure}
+	\includegraphics{figures/lloid-diagram.pdf}
+	\caption{Schematic of LLOID pipeline illustrating signal flow.
+Circles with arrows represent upsampling \protect\includegraphics{figures/upsample-symbol.pdf} or downsampling \protect\includegraphics{figures/downsample-symbol.pdf}.
+Circles with plus signs represent summing junctions \protect\includegraphics{figures/adder-symbol.pdf}.
+Squares \protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.
+Sample rate decreases from the top of the diagram to the bottom.}
+\end{figure}

### 2011-05-27 figures/diagram.tex
Added new diagram to paper
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@764 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \documentclass{amsart}
-\usepackage[usenames]{color}
 \usepackage{tikz}
 \usepackage{signalflowdiagram}
-\usetikzlibrary{fit,calc,chains,scopes}
+\usetikzlibrary{external,fit,calc,chains,scopes}
+\tikzexternalize
 \begin{document}
@@
 
-\begin{figure}
-	\begin{center}
+\tikzsetnextfilename{lloid-diagram}
 \begin{tikzpicture}[
@@
 \end{tikzpicture}
-		\caption{Schematic of LLOID pipeline illustrating signal flow.
-Circles with arrows represent upsampling \protect\tikz{\protect \node[resample] {\tiny$\uparrow$};} or downsampling \protect\tikz{\protect \node[resample] {\tiny$\downarrow$};}.
-Circles with plus signs represent summing junctions \protect\tikz{\protect \node[line width=1pt,adder] {\tiny$\uparrow$};}.
-Squares \protect\tikz{\protect \node[line width=1pt,draw] {};} stand for FIR filters.
-Sample rate decreases from the top of the diagram to the bottom.}
-	\end{center}
-\end{figure}
 
-This diagram is a work in progress.  Some corrections to make:
+\tikzsetnextfilename{upsample-symbol}
+\begin{tikzpicture}
+	\node[resample] {\tiny$\uparrow$};
+\end{tikzpicture}
 
-{\color{red}
-\begin{itemize}
-\item Add delay lines.
-\item Switch first downsampler and whitener.
-\item Resamplers are connected in series, not parallel.
-\item Translate figure to the left to align it visually with the caption.
-\end{itemize}
-}
+\tikzsetnextfilename{downsample-symbol}
+\begin{tikzpicture}
+	\node[resample] {\tiny$\downarrow$};
+\end{tikzpicture}
+
+\tikzsetnextfilename{adder-symbol}
+\begin{tikzpicture}
+	\node[line width=1pt,adder] {};
+\end{tikzpicture}
+
+\tikzsetnextfilename{fir-symbol}
+\begin{tikzpicture}
+	\node[line width=1pt,draw] {};
+\end{tikzpicture}
 

### 2011-05-27 introduction.tex
Small cleanups and added Alessandra to acknowledgements
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@760 ccdd05d5-56c1-461a-bf98-85f6876523b4
Updated astrophysical motivation; added many citations to observations.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@765 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 compact binary coalescence is believed to be a mechanism for short gamma-ray
-bursts~\cite{Lee:2005, nakar07}. This prompt emission is a probe into the
-extreme initial conditions of the outflow. A very bright optical flash should
-accompany the x-rays and gamma-rays~\cite{Sari99}. \editorial{Should mention
-that optical flashes have been seen for a handful of long GRBs by very fast
-slews. TAROT has seen many within ~20 seconds. \cite{Klotz09}} In contrast, an
-afterglow is produced when the outflow impacts the interstellar medium.
-Afterglows represent lower Lorentz factors and the conditions of the
-environment far outside of the central engine, and are relatively insensitive
-to initial conditions. In order to observe the most intense electromagnetic
-emission and provide an unbroken chain of observations between the progenitor
-system and the expanding fireball, telescopes must point during the seconds
-surrounding the merger.
+bursts (short GRBs)~\cite{Lee:2005, nakar07}. The same inner shocks, or
+potentially reverse shocks, can produce a bright accompanying optical flash~\cite{Sari99}. Prompt emission is a probe into the extreme
+initial conditions of the outflow, in contrast with afterglows, which are
+relatively insensitive to initial conditions. Optical flashes
+have only been observed for a handful of long
+GRBs~\cite{1999Natur.398..400A,2003Natur.422..284F,2006A&A...454L.119J,0004-637X-660-1-489,2008Natur.455..183R,2011A&A...528A..15G}
+by telescopes with extremely rapid response or, in the case of GRB 080319B, by
+pure serendipity, where several telescopes were observing a previous GRB in the
+same field~\cite{2008Natur.455..183R}. Short GRBs, on the other hand, typically
+fade too quickly to observe past the initial burst of gamma rays and hard
+x-rays. Rapid gravitational-wave transient alerts could enable the observation
+of optical flashes from short GRBs. An optical counterpart would vastly boost
+confidence in the gravitational-wave detection and provide the tight sky
+localization necessary to allow determination of the source's host galaxy,
+which leads to a redshift measurement. With both redshift and a coincident
+gravitational-wave observation, we can produce precision measurements of the
+Hubble constant~\cite{2010ApJ...725..496N}.
+
+To this end, we have the ambition of reporting candidates not minutes
+\emph{after} the merger, but seconds \emph{before}.  \editorial{I think we
+should say also something about how we can also get the full \SNR{} within moments
+after the coalescence, it is obvious maybe, but we should be explicit}  By
+looking for threshold crossings before the gravitational-wave signal leaves
+the detection band, it is possible to trade some signal to noise ratio (\SNR{})
+for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
+for NS--NS binaries in Advanced \LIGO\ assuming the event rate predictions
+in~\cite{Abadie:2010p10836}.
+%
+\begin{figure}
+\includegraphics{N_before_Tc.pdf}
+\caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
+be detectable $T_\mathrm{bc}$ seconds before coalescence.  The solid line is the most
+likely yearly rate estimate $N_{\mathrm{re}}$ for Advanced \LIGO\ and the
+shaded region is the interval $N_{\mathrm{low}}$ to $N_{\mathrm{high}}$ from
+\cite{Abadie:2010p10836}.  Note that assuming \SNR\ 8 is sufficient for
+detection and that we observe $N_{\mathrm{re}} = 40$ events per year with a
+detector having the ZERO\_DET\_high\_P noise model described
+in~\cite{ALIGONoiseZERO_DET_high_P}, $\sim10$ sources detectable within 10
+seconds of merger and $\sim1$ sources detectable within 100 seconds of
+merger.  Other noise models, ZERO\_DET\_low\_P~\cite{ALIGONoiseZERO_DET_low_P}
+and BHBH\_20deg~\cite{ALIGONoiseBHBH_20deg} provide better results for early
+detection but at the cost of fewer total events observed above \SNR{} 8.}
+\end{figure}
 
@@
 initiated a project to send alerts when potential gravitational-wave transients
-are observed.  In October 2010, \textsc{ligo} completed its sixth science run
+are observed.  In October 2010, \LIGO{} completed its sixth science run
 (S6) and Virgo completed its third science run (VSR3).  While both
@@
 latency.
-
-To this end, we have the ambition of reporting candidates not minutes
-\emph{after} the merger, but seconds \emph{before}.  \editorial{I think we
-should say also something about how we can also get the full SNR within moments
-after the coalescence, it is obvious maybe, but we should be explicit}  By
-looking for threshold crossings before the signal leaves the detection band, it
-is possible to trade some signal to noise ratio (\textsc{snr}) for latency.
-Figure \ref{fig:earlywarning} shows projected early trigger rates for NS--NS
-binaries in Advanced \LIGO\ assuming the event rate predictions in
-\cite{Abadie:2010p10836}.
-%
-\begin{figure}
-\includegraphics{N_before_Tc.pdf}
-\caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
-be detectable $T_\mathrm{bc}$ seconds before coalescence.  The solid line is the most
-likely yearly rate estimate $N_{\mathrm{re}}$ for Advanced \LIGO\ and the
-shaded region is the interval $N_{\mathrm{low}}$ to $N_{\mathrm{high}}$ from
-\cite{Abadie:2010p10836}.  Note that assuming \SNR\ 8 is sufficient for
-detection and that we observe $N_{\mathrm{re}} = 40$ events per year with a
-detector having the ZERO\_DET\_high\_P noise model described
-in~\cite{ALIGONoiseZERO_DET_high_P}, $\sim10$ sources may be detected within 10
-seconds of merger and $\sim1$ sources maybe detected within 100 seconds of
-merger.  Other noise models, ZERO\_DET\_low\_P~\cite{ALIGONoiseZERO_DET_low_P}
-and BHBH\_20deg~\cite{ALIGONoiseBHBH_20deg} provide better results for early
-detection but at the cost of fewer total events observed above SNR 8.}
-\end{figure}
 
@@
 principal component analysis to reduce the number of templates.  We described a
-rank reduction scheme based on singular value decomposition in
+rank-reduction scheme based on singular value decomposition in
 \cite{Cannon:2010p10398}.  We will use both aspects to demonstrate that a very

### 2011-05-28 figures/diagram.tex
diagram: added delay lines
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@767 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \usetikzlibrary{external,fit,calc,chains,scopes}
-\tikzexternalize
+\tikzexternalize[force remake=true]
 \begin{document}
@@
 		\node[junction] at ($0.5*(u0) + 0.5*(u2) - (0.5, 0)$) (in\x tee) {};
-		\node[resample, left=1.5cm of in\x tee] (in\x) {\tiny$\downarrow$};
+		\node[resample, left=1.75cm of in\x tee] (in\x) {\tiny$\downarrow$};
 
@@
 	\draw (in0) -- (in0tee) node[midway,above] {\small 4096 Hz};
-	\draw (in1) -- (in1tee) node[midway,above] {\small 512 Hz};
-	\draw (in2) -- (in2tee) node[midway,above] {\small 32 Hz};
+	\draw (in1) -- (in1tee) node[midway,draw,fill=white,label=above:{\small delay}, label=below:{\small 512 Hz}] {$z^{-t_1 f_1}$};
+	\draw (in2) -- (in2tee) node[midway,draw,fill=white,label=above:{\small delay},label=below:{\small 32 Hz}] {$z^{-\cdots}$};
 

### 2011-05-28 time_slices.py
Moved some settings to Matplotlibrc file to make plots more uniform
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@768 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -10,20 +10,11 @@ import matplotlib
 import pylab
 from itertools import groupby, izip
 matplotlib.rcParams.update({
-        "font.size": 9.0,
-        "axes.titlesize": 9.0,
-        "axes.labelsize": 9.0,
-        "xtick.labelsize": 9.0,
-        "ytick.labelsize": 9.0,
-        "legend.fontsize": 8.0,
         "figure.figsize": (3,3),
-        "figure.dpi": 600,
         "subplots.left": 0.1,
         "subplots.right": 0.75,
         "subplots.bottom": 0.25,
         "subplots.top": 0.75,
-        "savefig.dpi": 600,
-        "text.usetex": True     # render all text with TeX
 })
 
 # Command line interface

### 2011-05-28 time_slices.py
Tweaked table 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@769 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -10,7 +10,7 @@ import matplotlib
 import pylab
 from itertools import groupby, izip
 matplotlib.rcParams.update({
-        "figure.figsize": (3,3),
+        "figure.figsize": (3,2.8),
         "subplots.left": 0.1,
         "subplots.right": 0.75,
         "subplots.bottom": 0.25,
@@ -69,9 +69,9 @@ pylab.savefig('time_slices.pdf')
 
 # Generate output table
 
-print r"\begin{tabular}{|lrr}"
+print r"\begin{tabular}{lrr}"
 #FIXME change the symbol for the number of sample points per slice if the macro changes
-print r"$\mathbf{f_k}$ \bf{(Hz)} & $\mathbf{(t_{k+1}}$,$\mathbf{t_k]}$ \bf{(s)} & $\mathbf{\slicessamps}$ \\"
+print r"$f_k$ (Hz) & $(t_{k+1}$,$t_k]$ (s) & $\slicessamps$ \\"
 print r"\hline"
 for slice in slices:
 	begin = slice['begin']

### 2011-05-28 snr_in_time.py
Tweaked figure 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@772 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -1,32 +1,9 @@
 #!/usr/bin/python
 import numpy
-import matplotlib
-matplotlib.use('Agg')
 import pylab
 import scipy
 from scipy import interpolate
 import sys
-
-#
-# default plot paramaters
-#
-
-matplotlib.rcParams.update({
-	"font.size": 9.0,
-	"axes.titlesize": 9.0,
-	"axes.labelsize": 9.0,
-	"xtick.labelsize": 9.0,
-	"ytick.labelsize": 9.0,
-	"legend.fontsize": 9.0,
-	"figure.figsize": (3,3),
-	"figure.dpi": 600,
-	"subplots.left": 0.05,
-	"subplots.right": 0.95,
-	"subplots.bottom": 0.05,
-	"subplots.top": 0.95,
-	"savefig.dpi": 600,
-	"text.usetex": True     # render all text with TeX
-})
 
 #
 # Constants
@@ -67,7 +44,7 @@ def snr_to_num(fracsnr, number=40):
 	return number * (fracsnr)**3
 
 # Start a new figure
-fig = pylab.figure(1, figsize=(5.5,2.25))
+fig = pylab.figure(figsize=(5.5,2.3))
 
 markers = ['k-', 'k--', 'k-.']
 
@@ -110,16 +87,15 @@ for i, cfile in enumerate(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt'
 
 	# generate the figure
 	ax1 = fig.add_subplot(1,1,1, adjustable='box', aspect=0.5)
-	pylab.loglog(t, num, markers[i], lw=2, label=cfile.replace('.txt','').replace('data/','').replace('_','\\_'))
+	pylab.loglog(t, num, markers[i], lw=2, label=cfile.replace('.txt','').replace('data/','').replace('_', ' '))
 	#pylab.hold(1)
 	pylab.fill_between(t, numminus, numplus, color='k', alpha=0.15)
 
 pylab.grid()
 pylab.legend(loc='lower left')
-pylab.ylabel('$\\langle N \\rangle$ yr$^{-1}$')
-pylab.xlabel('$T_{bc}$')
-pylab.subplots_adjust(bottom=0.15,top=0.95,left=0.08,right=0.95)
+pylab.ylabel(r'$\langle N \rangle$ yr$^{-1}$')
+pylab.xlabel(r'Time before coalescence, $T_{bc}$')
[…]

### 2011-05-30 figures/diagram.tex
Removed whitener
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@776 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 		\node[junction] at ($0.5*(u0) + 0.5*(u2) - (0.5, 0)$) (in\x tee) {};
-		\node[resample, left=1.75cm of in\x tee] (in\x) {\tiny$\downarrow$};
 
@@
 
+	\node[junction, left=1.9cm of in0tee] (in0) {};
+	\node[resample, left=1.75cm of in1tee] (in1) {\tiny$\downarrow$};
+	\node[resample, left=1.75cm of in2tee] (in2) {\tiny$\downarrow$};
+
 	\foreach \x in {0,1} {
@@
 
-	\node[draw, left=0.5cm of in0] (whiten) {\emph{whiten}};
-	\node[signal, label=left:$x$, label=below:{\small 16384 Hz}, left=0.75cm of whiten] (x) {};
+	\node[signal, label=left:$x$, left=0.75cm of in0] (x) {};
+	%\node[draw, left=0.5cm of in0] (whiten) {\emph{whiten}};
+	%\node[signal, label=left:$x$, label=below:{\small 16384 Hz}, left=0.75cm of whiten] (x) {};
 	\node[below=1.5cm of in1] (ellip) {$\vdots$};
@@
 	\node[above=1mm of out3] {\vdots};
-	\draw[->] (x) -- (whiten);
-	\draw[->] (whiten) -- (in0);
+	\draw[->] (x) -- (in0);
+	%\draw[->] (x) -- (whiten);
+	%\draw[->] (whiten) -- (in0);
 	\draw[->] (in0) -- (in1);
@@
 	\node[right=0.1cm of fir label,align=center] {Reconstruction \\ matrices \\ $\overbrace{\hspace{2.75cm}}^{}$};
-	\node[right=3.5cm of fir label,align=center] {Upsampling and \\ SNR accumulation \\ $\overbrace{\hspace{2.75cm}}^{}$};
-	\node[left=0.1cm of fir label,align=center] {Down-\\ sampling \\ $\overbrace{\hspace{1cm}}^{}$};
-	\node[left=1.75cm of fir label,align=center] {Data acquisition \\ and whitening \\ $\overbrace{\hspace{2cm}}^{}$};
+	\node[right=3.5cm of fir label,align=center] {Interpolation and \\ SNR accumulation \\ $\overbrace{\hspace{2.75cm}}^{}$};
+	\node[left=0.3cm of fir label,align=center] {Decimation \\ of input \\ $\overbrace{\hspace{1cm}}^{}$};
+	%\node[left=1.75cm of fir label,align=center] {Data acquisition \\ and whitening \\ $\overbrace{\hspace{2cm}}^{}$};
 

### 2011-05-30 method.tex
Tweaked table 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@769 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed run-on
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@773 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fix a few macros without following spaces.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@778 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 at least one template. Such a template bank is said to have a 97\% {\em
-minimum match}. Data are filtered against each template to produce an \SNR
+minimum match}. Data are filtered against each template to produce an \SNR{}
 time-series. Local peak-finding across time and templates determines detection
@@
 \centering
+\vspace{0.8cm}
 \includegraphics{time_slices.pdf}
@@
 requires about about $\tmpsamps / 8 \lg 2 \tmpsamps = 2.2 \times 10^4$ times
-more operations per sample than the \fft implementation.  We will now consider
+more operations per sample than the \fft{} implementation.  We will now consider
 the computational cost of the \fir\ filter implementation described in the
@@
 highest frequency term of \eqref{eq:TDcost} with \eqref{eq:SNRTD} shows \order{
-(\slicesamps{\mathrm{max}} + \numtmps) \svdtmps{\mathrm{max}} \fmax} \flops\
+(\slicesamps{\mathrm{max}} + \numtmps) \svdtmps{\mathrm{max}} \fmax} \flops{}
 versus \order{\numtmps \tmpsamps \fmax} {\flops}.  The full calculation for a

### 2011-05-30 analysis.tex
Moved figure around a litle bit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@770 ccdd05d5-56c1-461a-bf98-85f6876523b4
WIP
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@779 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\section{Software Framework}
-\label{SECIII}\label{sec:analysis}
+\section{Implementation}
 
-\FIXME{
-I would like to see this section have more of a description of the
-pipeline and a figure representing it.
-}
+Our method involves an offline planning stage and an online, low latency filtering stage.
 
-The detection method described above is much more efficient in terms of
-floating point operations than the traditional matched filter bank method.
-However, time slices and conditional reconstruction greatly complicate
-queueing, synchronizing, and bookkeeping of intermediate signals.  A low
-latency implementation capable of recruiting more than one \textsc{cpu} core
-would be difficult to achieve within the familiar serial programming framework
-because of the nontrivial time-delay relationships between samples.  Due to
-these complications, we chose to prototype the search using an open source
-signal processing environment called GStreamer \cite{gstreamer}.  Primarily
-used for playing, authoring, or streaming media on Linux systems, GStreamer is
-an integral component of the popular Gnome desktop.
+\subsection{Planning stage}
 
-\begin{figure}
+The planning stage proceeds as follows.
+First, the templates are chosen by covering the space of mass parameters with a hexagonal grid \cite{PhysRevD.76.102004} in order to satisfy the minimum match criterion.
+Next, the templates are subdivided into groups of neighbors called ``sub-banks'' that are appropriately sized so that each bank can be efficiently handled by a single machine.
+(In this paper, we will only study a single sub-bank.)  Using our understanding of the time-frequency evolution of the templates, we choose time slice boundaries such that all of the templates within this sub-bank are sub-critically sampled at progressively lower sample rates.
+Next, the templates within this sub-bank are realized as \textsc{fir} filter coefficients.
+For each time slice, the templates are downsampled to the appropriate sample rate.
+Finally, the \textsc{svd} is applied to each time slice in order to produce a set of orthogonal \textsc{fir} filters and a reconstruction matrix that maps them back to the original templates.
+The downsampled orthogonal \textsc{fir} filter coefficients, the reconstruction matrix, and the time slice boundaries are all saved to disk.
+
+\subsection{Filtering stage}
+
+We have implemented a prototype of the low latency filtering stage using an open source signal processing environment called GStreamer \cite{gstreamer}.
+GStreamer is a vital component of many Linux systems, providing media playback, authoring, and streaming on devices from cell phones to desktop computers to streaming media servers.
+It turns out that it is also uniquely suited to gravitational wave data analysis.
+In our application, GStreamer excels at queueing, synchronizing, adding, and bookkeeping many different signals at different sample rates.
+It also provides some useful signal processing primitives such as decimators, \textsc{fir} filters, and interpolators.
+Most importantly, it permits us to recruit all of the host system's \textsc{cpu}s without having to write complicated and error-prone multithreaded code.
+
+\begin{figure}[htbp]
 	\includegraphics{figures/lloid-diagram.pdf}
@@
 \end{figure}
+
+The filter pipeline consists of six distinct stages.
+
+\subsubsection{Decimation}
+
+First, the sample rate of the whitened detector data is reduced to successively lower sample rates by decimation.
+Decimation involves applying an antialiasing filter to the data, and then downsampling by deleting samples.
+We use a 192-tap \textsc{fir} decimator provided by GStreamer.
+
+\subsubsection{Time delays}
+
+Each decimated detector data stream becomes the input for one time slice, but it must be appropriately delayed.
+
+\subsubsection{Orthogonal \textsc{fir} filters
+
+\subsubsection{Reconstruction}
+
+\subsubsection{Interpolation}
+
+\subsubsection{SNR accumulation}
+
+%Our new method consists of six distinct stages:
+%\begin{enumerate}
+%\item\label{item:decimation} Decimation of the detector data to successively lower sample rates
+%\item\label{item:delay} Delay lines to appropriately synchronize of all of the time slices
+%\item\label{item:ortho} Application of orthogonal \textsc{fir} filters to the decimated detector data
+%\item\label{item:reconstruct} Mixing of orthogonal filter output using reconstruction matrices
+%\item\label{item:interpolation} Interpolation of reconstructed output to a common sample rate
+%\item\label{item:accumulation} Accumulation of \textsc{snr} from each time slice
+%\end{enumerate}

### 2011-05-30 method.tex
'called template banks' sounds awkward
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@781 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 (m_1, m_2)$. Searches for gravitational waves from compact binary coalescence
-typically employ matched filter banks, called template banks, that discretely
+typically employ matched filter banks that discretely
 sample the possible intrinsic parameters~\cite{findchirppaper}.  The filters

### 2011-05-30 method.tex
Some small changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@782 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Inspiral signals are parameterized by a set of intrinsic parameters
-$\bar{\theta}$ that determine the amplitude and phase evolution of the observed
-binary signals. For systems where the effects of spin can be ignored, the
+$\bar{\theta}$%
+%
+\editorial{Change this to a harpoon or vector symbol.}
+%
+that determine the amplitude and phase evolution of the gravitational wave strain. For systems where the effects of spin can be ignored, the
 intrinsic parameters are the component masses of the binary, $\bar{\theta} =
@@
 sample the possible intrinsic parameters~\cite{findchirppaper}.  The filters
-for the waveforms $x(t,\bar{\theta})$, known as templates, are the waveforms
-weighted by the inverse detector noise amplitude spectral density, so as to
-whiten them. To construct a template bank, templates are chosen with discrete
-signal parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$ to assure a
+for the waveforms $h(t,\bar{\theta})$, known as templates, are the waveforms
+weighted by the inverse detector noise amplitude spectral density.
+To construct a template bank, templates are chosen with discrete signal parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$ to assure a
 bounded loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}. That is, any possible
 signal within the search space will have a cross-correlation of $\geq0.97$ with
-at least one template. Such a template bank is said to have a 97\% {\em
-minimum match}. Data are filtered against each template to produce an \SNR{}
+at least one template. Such a template bank is said to have a {\em
+minimum match} of 0.97. Data are filtered against each template to produce an \SNR{}
 time-series. Local peak-finding across time and templates determines detection

### 2011-05-30 snr_in_time.py
snr_in_time.py: \mathrm{bc}
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@783 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -94,7 +94,7 @@ for i, cfile in enumerate(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt'
 pylab.grid()
 pylab.legend(loc='lower left')
 pylab.ylabel(r'$\langle N \rangle$ yr$^{-1}$')
-pylab.xlabel(r'Time before coalescence, $T_{bc}$')
+pylab.xlabel(r'Time before coalescence, $T_\mathrm{bc}$')
 pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.08,right=0.95)
 pylab.xlim([0.01,1000])
 pylab.ylim([.1, 1000])

### 2011-05-31 notation.tex
added standalone notation.tex for rapid iteration
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@784 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass[letterpaper,11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}

\title{Notation proposal}
\author{Leo and Nick}

\begin{document}
\maketitle

Definition of matched filter:
\begin{equation}
\rho_i [k] = \sum_{n=0}^{N-1} h_{i n} x [k-n]
\end{equation}

Definition of time slices:
\begin{align}
h_{i k} &= \sum_s h_{i k}^{(s)} \\
h_{i k}^{(s)} &= \left\{ \begin{aligned}
    h_{ik} & \; \text{if } t^{(s+1)} < \frac{k}{f} \leq t^{(s)} \\
    0 & \; \text{, otherwise }
    \end{aligned} \right.
\end{align}

Definition of decimation:
\begin{equation}
x^{(s+1)}[k] = \sum_{n=-N^\downarrow/2}^{N^\downarrow/2-1} b_n^\downarrow x^{(s)}\!\left[k \frac{f^{(s)}}{f^{(s+1)}} - n\right]
\end{equation}

\textsc{svd}:
\begin{equation}
h^{(s)}_{i\left(\frac{f}{f^{(s)}} k\right)} = \sum_{m=0}^{M-1} v^{(s)}_{im} \sigma^{(s)}_m u^{(s)}_{mk}
\end{equation}

Early-warning \textsc{snr}:
\begin{equation}
    \rho^{(s)}_i[k] = \underbrace{\sum_{l=0}^{L^{(s)}-1} v^{(s)}_{il} \sigma^{(s)}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
    \underbrace{\sum_{n=0}^{N^{(s)} - 1} u^{(s)}_{ln} x^{(s)}[k-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + \underbrace{\sum_{n=-N^\uparrow/2}^{N^\uparrow/2-1} b^\uparrow_n \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{(s)}}{f^{(s-1)}}\right]}_\text{interpolation filter} \underbrace{\rho^{(s+1)}_i[q]}_{\substack{\text{accumulated {\sc snr}}\\ \text{from prior time slices}}}
\end{equation}

nvf commentary: I think it would be more natural to use $h_i[k]$ everywhere rather than $h_{ik}$. Accordingly, I'd notate $u^{(s)}_m[k]$ as the basis templates. Leo disagreed.

\begin{table}
\begin{tabular}{rl}
\hline
\textbf{Symbol} & \textbf{Definition} \\\hline
$f^{(s)}$ & The sample rate for time-slice $s$\\
$t^{(s)}$ & The time of ``late'' boundary for time-slice $s$\\
$N^{(s)}$ & The number of samples in time-slice $s$\\
$L^{(s)}$ & The number of principal components in time-slice $s$\\
$N^\uparrow$ & The number of sample points in the interpolation filter\\
$N^\downarrow$ & The number of sample points in the decimation filter\\
$M$ & The number of templates\\
\hline
\end{tabular}
\caption{constants}
\end{table}

\end{document}

### 2011-05-31 notation.tex
notation.tex: Updates from Leo's comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@785 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \usepackage{amsmath}
+\usepackage{stmaryrd}
 
@@
 \begin{equation}
-\rho_i [k] = \sum_{n=0}^{N-1} h_{i n} x [k-n]
+\rho_i [k] = \sum_{n=0}^{N-1} h_{i}[n] x [k-n]
 \end{equation}
@@
 \begin{align}
-h_{i k} &= \sum_s h_{i k}^{(s)} \\
-h_{i k}^{(s)} &= \left\{ \begin{aligned}
-    h_{ik} & \; \text{if } t^{(s+1)} < \frac{k}{f} \leq t^{(s)} \\
+h_{i}[k] &= \sum_s h_{i}^{s}[k] \\
+h_{i}^{s}[k] &= \left\{ \begin{aligned}
+    h_{i}[k] & \; \text{if } t^{s+1} < \frac{k}{f} \leq t^{s} \\
     0 & \; \text{, otherwise }
@@
 \begin{equation}
-x^{(s+1)}[k] = \sum_{n=-N^\downarrow/2}^{N^\downarrow/2-1} b_n^\downarrow x^{(s)}\!\left[k \frac{f^{(s)}}{f^{(s+1)}} - n\right]
+x^{s+1}[k] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] x^{s}\!\left[k \frac{f^{s}}{f^{s+1}} - n\right]
 \end{equation}
@@
 \begin{equation}
-h^{(s)}_{i\left(\frac{f}{f^{(s)}} k\right)} = \sum_{m=0}^{M-1} v^{(s)}_{im} \sigma^{(s)}_m u^{(s)}_{mk}
+h^{s}_{i}\!\left[\frac{f}{f^{s}} k\right] = \sum_{m=0}^{M-1} v^{s}_{im} \sigma^{s}_m u^{s}_{m}[k] \approx \sum_{l=0}^{L^s-1} v^{s}_{il} \sigma^{s}_l u^{s}_{l}[k]
 \end{equation}
@@
 \begin{equation}
-    \rho^{(s)}_i[k] = \underbrace{\sum_{l=0}^{L^{(s)}-1} v^{(s)}_{il} \sigma^{(s)}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
-    \underbrace{\sum_{n=0}^{N^{(s)} - 1} u^{(s)}_{ln} x^{(s)}[k-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + \underbrace{\sum_{n=-N^\uparrow/2}^{N^\uparrow/2-1} b^\uparrow_n \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{(s)}}{f^{(s-1)}}\right]}_\text{interpolation filter} \underbrace{\rho^{(s+1)}_i[q]}_{\substack{\text{accumulated {\sc snr}}\\ \text{from prior time slices}}}
+    \rho^{s}_i[k] = \underbrace{\sum_{l=0}^{L^{s}-1} v^{s}_{il} \sigma^{s}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
+    \underbrace{\sum_{n=0}^{N^{s} - 1} u^{s}_{l}[n] x^{s}[k-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + \underbrace{\sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{s}}{f^{s+1}}\right]}_\text{interpolation filter} \underbrace{\rho^{s+1}_i[q]}_{\substack{\text{accumulated {\sc snr}}\\ \text{from prior time slices}}}
 \end{equation}
-
-nvf commentary: I think it would be more natural to use $h_i[k]$ everywhere rather than $h_{ik}$. Accordingly, I'd notate $u^{(s)}_m[k]$ as the basis templates. Leo disagreed.
 
@@
 \textbf{Symbol} & \textbf{Definition} \\\hline
-$f^{(s)}$ & The sample rate for time-slice $s$\\
-$t^{(s)}$ & The time of ``late'' boundary for time-slice $s$\\
-$N^{(s)}$ & The number of samples in time-slice $s$\\
-$L^{(s)}$ & The number of principal components in time-slice $s$\\
-$N^\uparrow$ & The number of sample points in the interpolation filter\\
-$N^\downarrow$ & The number of sample points in the decimation filter\\
+$f^{s}$ & The sample rate for time-slice $s$\\
+$t^{s}$ & The time of ``late'' boundary for time-slice $s$\\
+$N^{s}$ & The number of samples in time-slice $s$\\
+$L^{s}$ & The number of principal components in time-slice $s$\\
+$N^\shortuparrow$ & The number of sample points in the interpolation filter\\
+$N^\shortdownarrow$ & The number of sample points in the decimation filter\\
 $M$ & The number of templates\\

### 2011-05-31 analysis.tex
Fixed tex syntax error
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@780 ccdd05d5-56c1-461a-bf98-85f6876523b4
added intro paragraph and formatted to 80 columns
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@786 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Our method involves an offline planning stage and an online, low latency filtering stage.
+In this section we describe an implementation of the method described in
+section \ref{label{sec:method}} suitable for rapid gravitational-wave searches
+for compact binary coalescence.  The previous method requires several
+computations that can be done before the analysis is actually underway.  Thus
+we divide the procedure into two stages 1) an offline planning stage and 2) an
+online, low latency filtering stage.  The offline stage can be done before the
+analysis is started an updated asychronously whereas the online stage must keep
+up with the detector output and produce search results as rapidly as possible.
+In the next two subsections we describe what these stages entail.
 
@@
 
-The planning stage proceeds as follows.
-First, the templates are chosen by covering the space of mass parameters with a hexagonal grid \cite{PhysRevD.76.102004} in order to satisfy the minimum match criterion.
-Next, the templates are subdivided into groups of neighbors called ``sub-banks'' that are appropriately sized so that each bank can be efficiently handled by a single machine.
-(In this paper, we will only study a single sub-bank.)  Using our understanding of the time-frequency evolution of the templates, we choose time slice boundaries such that all of the templates within this sub-bank are sub-critically sampled at progressively lower sample rates.
-Next, the templates within this sub-bank are realized as \textsc{fir} filter coefficients.
-For each time slice, the templates are downsampled to the appropriate sample rate.
-Finally, the \textsc{svd} is applied to each time slice in order to produce a set of orthogonal \textsc{fir} filters and a reconstruction matrix that maps them back to the original templates.
-The downsampled orthogonal \textsc{fir} filter coefficients, the reconstruction matrix, and the time slice boundaries are all saved to disk.
+The planning stage proceeds as follows.  First, the templates are chosen by
+covering the space of mass parameters with a hexagonal grid
+\cite{PhysRevD.76.102004} in order to satisfy the minimum match criterion.
+Next, the templates are subdivided into groups of neighbors called
+``sub-banks'' that are appropriately sized so that each bank can be efficiently
+handled by a single machine.  (In this paper, we will only study a single
+sub-bank.)  Using our understanding of the time-frequency evolution of the
+templates, we choose time slice boundaries such that all of the templates
+within this sub-bank are sub-critically sampled at progressively lower sample
+rates.  Next, the templates within this sub-bank are realized as \textsc{fir}
+filter coefficients.  For each time slice, the templates are downsampled to the
+appropriate sample rate.  Finally, the \textsc{svd} is applied to each time
+slice in order to produce a set of orthogonal \textsc{fir} filters and a
+reconstruction matrix that maps them back to the original templates.  The
+downsampled orthogonal \textsc{fir} filter coefficients, the reconstruction
+matrix, and the time slice boundaries are all saved to disk.
 
@@
 
-We have implemented a prototype of the low latency filtering stage using an open source signal processing environment called GStreamer \cite{gstreamer}.
-GStreamer is a vital component of many Linux systems, providing media playback, authoring, and streaming on devices from cell phones to desktop computers to streaming media servers.
-It turns out that it is also uniquely suited to gravitational wave data analysis.
-In our application, GStreamer excels at queueing, synchronizing, adding, and bookkeeping many different signals at different sample rates.
-It also provides some useful signal processing primitives such as decimators, \textsc{fir} filters, and interpolators.
-Most importantly, it permits us to recruit all of the host system's \textsc{cpu}s without having to write complicated and error-prone multithreaded code.
+We have implemented a prototype of the low latency filtering stage using an
+open source signal processing environment called GStreamer \cite{gstreamer}.
+GStreamer is a vital component of many Linux systems, providing media playback,
+authoring, and streaming on devices from cell phones to desktop computers to
+streaming media servers.  It turns out that it is also uniquely suited to
+gravitational wave data analysis.  In our application, GStreamer excels at
+queueing, synchronizing, adding, and bookkeeping many different signals at
+different sample rates.  It also provides some useful signal processing
+primitives such as decimators, \textsc{fir} filters, and interpolators.  Most
+importantly, it permits us to recruit all of the host system's \textsc{cpu}s
+without having to write complicated and error-prone multithreaded code.
 
@@
 
-First, the sample rate of the whitened detector data is reduced to successively lower sample rates by decimation.
-Decimation involves applying an antialiasing filter to the data, and then downsampling by deleting samples.
-We use a 192-tap \textsc{fir} decimator provided by GStreamer.
+First, the sample rate of the whitened detector data is reduced to successively
+lower sample rates by decimation.  Decimation involves applying an antialiasing
+filter to the data, and then downsampling by deleting samples.  We use a
+192-tap \textsc{fir} decimator provided by GStreamer.
 
@@
 
-Each decimated detector data stream becomes the input for one time slice, but it must be appropriately delayed.
+Each decimated detector data stream becomes the input for one time slice, but
+it must be appropriately delayed.
 
-\subsubsection{Orthogonal \textsc{fir} filters
+\subsubsection{Orthogonal \textsc{fir} filters}
 

### 2011-05-31 notation.tex
notation: fix underbrace spacing
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@787 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \documentclass[letterpaper,11pt]{article}
-\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
-\usepackage{graphicx}
 \usepackage{amssymb}
 \usepackage{amsmath}
-\usepackage{stmaryrd}
+\usepackage{stmaryrd}  % gives \shortuparrow and \shortdownarrow
 
@@
 Early-warning \textsc{snr}:
+\newcommand{\mystrut}{\rule[-2em]{0pt}{0pt}}
 \begin{equation}
-    \rho^{s}_i[k] = \underbrace{\sum_{l=0}^{L^{s}-1} v^{s}_{il} \sigma^{s}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
-    \underbrace{\sum_{n=0}^{N^{s} - 1} u^{s}_{l}[n] x^{s}[k-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + \underbrace{\sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{s}}{f^{s+1}}\right]}_\text{interpolation filter} \underbrace{\rho^{s+1}_i[q]}_{\substack{\text{accumulated {\sc snr}}\\ \text{from prior time slices}}}
+    \rho^{s}_i[k] = \underbrace{\mystrut\sum_{l=0}^{L^{s}-1} v^{s}_{il} \sigma^{s}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
+    \underbrace{\mystrut\sum_{n=0}^{N^{s} - 1} u^{s}_{l}[n] x^{s}[k-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + \underbrace{\mystrut\sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{s}}{f^{s+1}}\right]}_\text{interpolation filter} \underbrace{\mystrut\rho^{s+1}_i[q]}_{\substack{\text{accum.}\\ \text{prior {\sc snr}}}}
 \end{equation}

### 2011-05-31 notation.tex
notation.tex: try new k notation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@788 ccdd05d5-56c1-461a-bf98-85f6876523b4
notation.tex: try new k notation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@789 ccdd05d5-56c1-461a-bf98-85f6876523b4
added a few sentences
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@790 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \end{equation}
+where $h_i[k]$ denotes the $k^{\mathrm{th}}$ sample of the $i^{\mathrm{th}}$
+template waveform and $x[k-n]$ represents the $(k-n)^{\mathrm{th}}$ sample of
+the data.
 
@@
     0 & \; \text{, otherwise }
-    \end{aligned} \right.
+    \end{aligned} \right. \\
+h_{i}^{s}[k^{s}] &\equiv h_{i}^{s}\!\left[k\frac{f}{f^{s}}\right]
 \end{align}
+
+Throughout, we will use the convention that $k$ indexes time-series sampled at
+the full sample rate $f$ and $k^s$ indexes time-series sampled at a reduced
+sample rate $f^s$.  For the case of band limited functions it is sufficient to
+simply remove sample points to define the downsampled version as is shown in
+the previous equation.
 
@@
 \begin{equation}
-x^{s+1}[k] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] x^{s}\!\left[k \frac{f^{s}}{f^{s+1}} - n\right]
+x^{s+1}[k^{s+1}] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] x^{s}\!\left[k^{s} - n\right]
 \end{equation}
@@
 \begin{equation}
-h^{s}_{i}\!\left[\frac{f}{f^{s}} k\right] = \sum_{m=0}^{M-1} v^{s}_{im} \sigma^{s}_m u^{s}_{m}[k] \approx \sum_{l=0}^{L^s-1} v^{s}_{il} \sigma^{s}_l u^{s}_{l}[k]
+h^{s}_{i}\!\left[k^{s}\right] = \sum_{m=0}^{M-1} v^{s}_{im} \sigma^{s}_m u^{s}_{m}[k^{s}] \approx \sum_{l=0}^{L^s-1} v^{s}_{il} \sigma^{s}_l u^{s}_{l}[k^{s}]
 \end{equation}
@@
 Early-warning \textsc{snr}:
-\newcommand{\mystrut}{\rule[-2em]{0pt}{0pt}}
+\newcommand{\mystrut}{\rule[-2em]{0pt}{0pt}}  % line up the underbraces
 \begin{equation}

### 2011-05-31 notation.tex
a few minor changes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@791 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Definition of decimation:
+Definition of decimation for a function $f$:
 \begin{equation}
-x^{s+1}[k^{s+1}] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] x^{s}\!\left[k^{s} - n\right]
+f[k^{s+1}] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] f\!\left[k^{s} - n\right]
+\end{equation}
+
+Definition of interpolation for a function $f$:
+\begin{equation}
+f[k^{s}] = \sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] f\!\left[k^{s+1} - n\right]
 \end{equation}
@@
 \begin{equation}
-    \rho^{s}_i[k] = \underbrace{\mystrut\sum_{l=0}^{L^{s}-1} v^{s}_{il} \sigma^{s}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
-    \underbrace{\mystrut\sum_{n=0}^{N^{s} - 1} u^{s}_{l}[n] x^{s}[k-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + \underbrace{\mystrut\sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{s}}{f^{s+1}}\right]}_\text{interpolation filter} \underbrace{\mystrut\rho^{s+1}_i[q]}_{\substack{\text{accum.}\\ \text{prior {\sc snr}}}}
+    \rho^{s}_i[k^{s}] = \underbrace{\mystrut\sum_{l=0}^{L^{s}-1} v^{s}_{il} \sigma^{s}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
+    \underbrace{\mystrut\sum_{n=0}^{N^{s} - 1} u^{s}_{l}[n] x[k^{s}-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + 
+    % + 
+    % \underbrace{\mystrut\sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{s}}{f^{s+1}}\right]}_\text{interpolation filter} 
+    \underbrace{\mystrut\rho^{s+1}_i[k^{s}]}_{\substack{\text{accum.}\\ \text{prior {\sc snr}}}}
 \end{equation}

### 2011-05-31 notation.tex
fixed interp decimate functions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@792 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the full sample rate $f$ and $k^s$ indexes time-series sampled at a reduced
-sample rate $f^s$.  For the case of band limited functions it is sufficient to
-simply remove sample points to define the downsampled version as is shown in
-the previous equation.
+sample rate $f^s$, ($k^{s} \equiv k f \ f^{s}$).  For the case of band limited
+functions it is sufficient to simply remove sample points to define the
+downsampled version as is shown in the previous equation.
 
-Definition of decimation for a function $f$:
+Definition of decimation for a function $p$:
 \begin{equation}
-f[k^{s+1}] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] f\!\left[k^{s} - n\right]
+p[k^{s+1}] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] \, p\!\left[k^{s}-n\right]
 \end{equation}
 
-Definition of interpolation for a function $f$:
+Definition of interpolation for a function $p$:
 \begin{equation}
-f[k^{s}] = \sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] f\!\left[k^{s+1} - n\right]
+p[k^{s}] = \sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] \, p\!\left[k^{s+1}-n\right]
+%p[k^{s}] = \sum_{n=k^{s}-N^\shortuparrow/2}^{k^{s} + N^\shortuparrow/2-1} b^\shortuparrow[k^{s} - n] \, p\!\left[\frac{n f^{s+1}}{f^{s}}\right]
+%f[k^{s}] = \sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] f\!\left[n \frac{f^{s}}{f^{s+1}}  \right]
 \end{equation}

### 2011-05-31 notation.tex
fixed interp decimate functions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@793 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the full sample rate $f$ and $k^s$ indexes time-series sampled at a reduced
-sample rate $f^s$, ($k^{s} \equiv k f \ f^{s}$).  For the case of band limited
+sample rate $f^s$, ($k^{s} \equiv k f / f^{s}$).  For the case of band limited
 functions it is sufficient to simply remove sample points to define the

### 2011-05-31 analysis.tex
added some more text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@795 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 online, low latency filtering stage.  The offline stage can be done before the
-analysis is started an updated asychronously whereas the online stage must keep
+analysis is started an updated asychronously, whereas the online stage must keep
 up with the detector output and produce search results as rapidly as possible.
@@
 
-The planning stage proceeds as follows.  First, the templates are chosen by
-covering the space of mass parameters with a hexagonal grid
-\cite{PhysRevD.76.102004} in order to satisfy the minimum match criterion.
-Next, the templates are subdivided into groups of neighbors called
-``sub-banks'' that are appropriately sized so that each bank can be efficiently
-handled by a single machine.  (In this paper, we will only study a single
-sub-bank.)  Using our understanding of the time-frequency evolution of the
-templates, we choose time slice boundaries such that all of the templates
-within this sub-bank are sub-critically sampled at progressively lower sample
-rates.  Next, the templates within this sub-bank are realized as \textsc{fir}
-filter coefficients.  For each time slice, the templates are downsampled to the
-appropriate sample rate.  Finally, the \textsc{svd} is applied to each time
-slice in order to produce a set of orthogonal \textsc{fir} filters and a
-reconstruction matrix that maps them back to the original templates.  The
-downsampled orthogonal \textsc{fir} filter coefficients, the reconstruction
-matrix, and the time slice boundaries are all saved to disk.
+The choice of filter waveforms and singular value decomposition can be done in
+advance and will be valid as long as the detector noise spectrum remains
+roughly constant.  New filter waveforms can be computed asynchronously using
+updated spectrum estimates as they are available.
+
+The planning stage begins with choosing templates that cover the space of mass
+parameters with a hexagonal grid \cite{PhysRevD.76.102004} in order to satisfy
+the minimum match criterion, which assures a user specified maximum loss in SNR
+for signals that fall in-between the chosen templates.  Next, the templates are
+subdivided into groups of neighbors called ``sub-banks'' that are appropriately
+sized so that each bank can be efficiently handled by a single computer.
+Dividing the mass space into smaller sub-banks also aids in the computational
+cost of the singular value decomposition.  Using our understanding of the
+time-frequency evolution of the templates, we choose time slice boundaries as
+in \eqref{FIXME} such that all of the templates within a sub-bank are
+sub-critically sampled at progressively lower sample rates.  Next, the
+templates within the sub-bank are realized as \textsc{fir} filter coefficients.
+For each time slice, the templates are downsampled to the appropriate sample
+rate.  Finally, the \textsc{svd} is applied to each time slice in order to
+produce a set of orthogonal \textsc{fir} filters and a reconstruction matrix
+that maps them back to the original templates.  The downsampled orthogonal
+\textsc{fir} filter coefficients, the reconstruction matrix, and the time slice
+boundaries are all saved to disk.
 

### 2011-06-01 figures/diagram.tex
Updated diagram, using matrix for layout and adding early-warning SNR output
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@796 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-05-27 figures/diagram.tex
@@
 \usepackage{tikz}
-\usepackage{signalflowdiagram}
-\usetikzlibrary{external,fit,calc,chains,scopes}
-\tikzexternalize
+\usetikzlibrary{external,decorations.pathreplacing,matrix,fit,calc,chains,scopes}
+\tikzexternalize[force remake=true]
+
+% Copied from signalflowdiagram.sty
+
+%
+% Colors
+%
+\newcommand{\pathdrawcolor}{\blockdrawcolor}     % draw color of signal paths
+\newcommand{\pathfillcolor}{\blockfillcolor}     % fill color of signal paths
+\newcommand{\operatordrawcolor}{\blockdrawcolor} % draw color of operators
+\newcommand{\operatorfillcolor}{\blockfillcolor} % fill color of operators
+\newcommand{\blockdrawcolor}{black}              % draw color of blocks
+\newcommand{\blockfillcolor}{white}              % fill color of blocks
+
+%
+% line widths of
+%
+\newcommand{\diagramlinewidth}{0.5mm}  % signal flow diagram frames
+\newcommand{\pathlinewidth}{0.3mm}     % signal paths
+\newcommand{\operatorlinewidth}{0.3mm} % operator frames
+\newcommand{\blocklinewidth}{0.5mm}    % building block styles
+
+%
+% line widths, sizes, etc.
+%
+\newcommand{\nodesize}{1.5mm}         % size of nodes
+\newcommand{\terminalsize}{\nodesize} % size of terminals
+\newcommand{\operatorsize}{3mm}       % size of circular shaped operator symbols
+\newcommand{\delaysize}{6mm}          % minimum size of delay elements
+\newcommand{\blockwidth}{24mm}        % minimum width of blocks
+\newcommand{\blockheight}{12mm}       % minimum height of blocks
+\newcommand{\blocktextwidth}{22mm}    % minimum text width of blocks
+\newcommand{\filterwidth}{12mm}       % minimum width of filters
+\newcommand{\filterheight}{8mm}       % minimum height of filters
+\newcommand{\filtertextwidth}{6mm}    % minimum text width of filters
+\usetikzlibrary{signalflowoperators}
+
 \begin{document}
@@
 \tikzstyle{resample}=[draw, circle, line width=1pt, inner sep=0pt, minimum size=0.3cm]
+\tikzstyle{vdots}=[xshift=-0.75mm, inner sep=0pt, text width=0pt, minimum size=0pt]
 
@@
 \begin{tikzpicture}[
+	every label/.style={text height=4pt},
+	column sep=1mm,
+	row sep=1mm,
 	node distance=2mm,
 	every path/.style={line width=1pt},
-	every node/.style={line width=1pt}
+	every node/.style={line width=1pt, anchor=center}
 ]
-	\foreach \x in {0,1,2} {
-		\node[draw] at (0, {\x == 0 ? 0cm : (\x == 1 ? -3cm : -7.5cm)}) (u0) {};
-		\node[draw, below=of u0] (u1) {};
-		\node[below=of u1] (uellip\x) {$\vdots$};
-		\node[draw, below=of uellip\x] (u2) {};
 
-		\node[junction] at ($0.5*(u0) + 0.5*(u2) - (0.5, 0)$) (in\x tee) {};
-		\node[resample, left=1.5cm of in\x tee] (in\x) {\tiny$\downarrow$};
+	\matrix (m) [
+		matrix of nodes,
+		column 1/.style={column sep=8mm},
+		column 2/.style={column sep=15mm},
+		column 3/.style={column sep=3mm},
+		column 4/.style={column sep=8mm},
+		column 5/.style={column sep=15mm},
+		column 6/.style={column sep=8mm},
+		column 11/.style={column sep=8mm}
+	]{
+		& \node[yshift=0.5cm](in){}; & \node[yshift=0.5cm](intee){}; & \node[yshift=0.5cm](u){}; & \node[yshift=0.5cm](utee){}; & \node[yshift=0.5cm](v){}; & \node[yshift=0.5cm](add0){}; & \node[yshift=0.5cm](add1){}; & \node[yshift=0.5cm](add2){}; & & \node[yshift=0.5cm](add3){}; & \node[yshift=0.5cm](out){}; \\
+		& & & \node[draw] (u00) {}; & \node[junction] (u00tee) {}; & \node[adder] (v00) {}; & \node[adder] (add00) {}; & & & & & \node[signal, label=right:{$\rho_0^0$}] (out00) {}; \\
+		& & & \node[draw] (u10) {}; & \node[junction] (u10tee) {}; & \node[adder] (v10) {}; & & \node[adder] (add10) {}; & & & & \node[signal, label=right:{$\rho_1^0$}] (out10) {}; \\
+		\node[signal, label=left:{$x$}] (x) {}; & \node[junction] (in0) {}; & \node[junction] (in0tee) {}; & & & \node[adder] (v20) {}; & & & \node[adder] (add20) {}; & & & \node[signal, label=right:{$\rho_2^0$}] (out20) {}; \\
+		& & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
+		& & & \node[draw] (u20) {}; & \node[junction] (u20tee) {}; & \node[adder] (v30) {}; & & & & & \node[adder] (add30) {}; & \node[signal, label=right:{$\rho_{M-1}^0$}] (out30) {}; \\
+		& & & & & & & & \node[minimum height=6mm](delay0){}; & & & \\
+		& & & & & & \node[resample] (resample01) {\tiny$\uparrow$}; & \node[resample] (resample11) {\tiny$\uparrow$}; & \node[resample] (resample21) {\tiny$\uparrow$}; & \node{$\cdot \cdot$}; & \node[resample] (resample31) {\tiny$\uparrow$}; \\
+%
+		& & & \node[draw] (u01) {}; & \node[junction] (u01tee) {}; & \node[adder] (v01) {}; & \node[adder] (add01) {}; & & & & & \node[signal, label=right:{$\rho_0^1$}] (out01) {}; \\
+		& & & \node[draw] (u11) {}; & \node[junction] (u11tee) {}; & \node[adder] (v11) {}; & & \node[adder] (add11) {}; & & & & \node[signal, label=right:{$\rho_1^1$}] (out11) {}; \\
+		& \node[resample] (in1) {\tiny$\downarrow$}; & \node[junction] (in1tee) {}; & & & \node[adder] (v21) {}; & & & \node[adder] (add21) {}; & & & \node[signal, label=right:{$\rho_2^1$}] (out21) {}; \\
+		& & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
+		& & & \node[draw] (u21) {}; & \node[junction] (u21tee) {}; & \node[adder] (v31) {}; & & & & & \node[adder] (add31) {}; & \node[signal, label=right:{$\rho_{M-1}^1$}] (out31) {}; \\
+		& & & & & & & & \node[minimum height=6mm](delay1){}; & & & \\
+		& & & & & & \node[resample] (resample02) {\tiny$\uparrow$}; & \node[resample] (resample12) {\tiny$\uparrow$}; & \node[resample] (resample22) {\tiny$\uparrow$}; & \node{$\cdot \cdot$}; & \node[resample] (resample32) {\tiny$\uparrow$}; \\
+%
+		& \node[minimum size=0pt, text height=8pt, anchor=base] (invdots) {$\vdots$}; & & \node[minimum size=0pt, text height=0pt, anchor=base] {$\vdots$}; & \node[minimum size=0pt, text height=8pt, anchor=base] {$\vdots$}; & \node[minimum size=0pt, text height=8pt, anchor=base] {$\vdots$}; & \node[minimum height=8pt]{}; & & & & & \node[minimum size=0pt, text height=0pt, anchor=base] {$\vdots$}; \\
+		& & & & & & \node[minimum size=0pt, text height=8pt] (vdots0) {$\vdots$}; & \node[minimum size=0pt, text height=8pt] (vdots1) {$\vdots$}; & \node[minimum size=0pt, text height=8pt] (vdots2) {$\vdots$}; & & \node[minimum size=0pt, text height=8pt] (vdots3) {$\vdots$}; \\
+%
+		& & & \node[draw] (u02) {}; & \node[junction] (u02tee) {}; & \node[adder] (v02) {}; & \node[junction] (add02) {}; & & & & & \node[signal, label=right:{$\rho_0^{S-1}$}] (out02) {}; \\
+		& & & \node[draw] (u12) {}; & \node[junction] (u12tee) {}; & \node[adder] (v12) {}; & & \node[junction] (add12) {}; & & & & \node[signal, label=right:{$\rho_1^{S-1}$}] (out12) {}; \\
+		& \node[junction] (in2) {}; & \node[junction] (in2tee) {}; & & & \node[adder] (v22) {}; & & & \node[junction] (add22) {}; & & & \node[signal, label=right:{$\rho_2^{S-1}$}] (out22) {}; \\
+		& & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
+		& & & \node[draw] (u22) {}; & \node[junction] (u22tee) {}; & \node[adder] (v32) {}; & & & & & \node[junction] (add32) {}; & \node[signal, label=right:{$\rho_{M-1}^{S-1}$}] (out32) {}; \\
+	};
 
 	\foreach \i in {0,1,2} {
-			\draw (in\x tee) |- (u\i);
-			\node[junction, right=1cm of u\i] (tee u\i) {};
-			\draw (u\i) -- (tee u\i);
+		\draw[loosely dotted] ($0.3*(add2\i) + 0.7*(add3\i)$) -- ($0.7*(add2\i) + 0.3*(add3\i)$);
+		\node[draw,dashed,line width=0.5pt,fit=(in\i tee) (u0\i) (u2\i)] {};
+		\node[draw,dashed,line width=0.5pt,yshift=0.5mm,fit=(u0\i tee) (v3\i)] {};
+		\foreach \j in {0,1,2,3}
+			\foreach \k in {0,1,2}
+				\draw[->] (u\k \i tee) -- (v\j \i) -- (add\j \i);
+		\foreach \k in {0,1,2}
+			\draw (in\i tee) |- (u\k \i) -- (u\k \i tee);
+		\node[vdots] at ($0.5*(u1\i tee) + 0.5*(u2\i tee)$) {$\vdots$};
+		\node[vdots] at ($0.5*(u1\i) + 0.5*(u2\i)$) {$\vdots$};
 	}
 
-		\node[adder, right=2cm of tee u0] (v0\x) {};
-		\node[adder, below=of v0\x] (v1\x) {};
-		\node[adder, below=of v1\x] (v2\x) {};
-		\node[right=0.85cm of uellip\x] {$\vdots$};
-		\node[adder, right=2cm of tee u2] (v3\x) {};
-		\node[above=0cm of v3\x] {$\vdots$};
+	\foreach \i in {1,2}
+		\foreach \j in {0,1,2,3}
+			\draw[->, dashed] (add\j \i) -- (out\j \i);
 
-		\foreach \i in {0,1,2}
 	\foreach \j in {0,1,2,3}
-				\draw (tee u\i) -- (v\j\x);
-
-		\node[draw,dashed,line width=0.5pt,fit=(in\x tee) (u0) (u2)] (ortho\x) {};
-		\node[draw,dashed,line width=0.5pt,fit=(tee u0) (tee u2) (v0\x) (v3\x)] (reconstruct\x) {};
-	}
-
-	\foreach \x in {1,2} {
-		\node[resample, right=1cm of v0\x, yshift=0.5cm] (resample0\x) {\tiny$\uparrow$};
-		\node[resample, right=1.5cm of v0\x, yshift=0.5cm] (resample1\x) {\tiny$\uparrow$};
-		\node[resample, right=2cm of v0\x, yshift=0.5cm] (resample2\x) {\tiny$\uparrow$};
-		\node[right=2.5cm of v0\x, yshift=0.5cm] (dots) {$\cdots$};
-		\node[below=of dots] (dots) {$\cdots$};
-		\node[below=of dots] (dots) {$\cdots$};
-		\node[resample, right=3.5cm of v0\x, yshift=0.5cm] (resample3\x) {\tiny$\uparrow$};
-	}
-
-	\foreach \x in {0,1} {
-		\node[adder, right=1cm of v0\x] (adder0\x) {};
-		\node[adder, right=1.5cm of v1\x] (adder1\x) {};
-		\node[adder, right=2cm of v2\x] (adder2\x) {};
-		\node[adder, right=3.5cm of v3\x] (adder3\x) {};
-		\foreach \i in {0,1,2,3} {
-			\draw[->] (v\i \x) -- (adder\i \x);
-		}
-	}
-
-	\foreach \i in {0,1,2,3} {
-		\node[signal, right=4.5cm of v\i 0, label=right:$\rho_\i$] (out\i) {};
-		\draw[->] (v\i 2) -| (resample\i 2);
-		\draw[->] (resample\i 2) -- (adder\i 1) node[above=0.5cm of resample\i 2,fill=white] {$\vdots$};
-		\draw[->] (adder\i 1) -- (resample\i 1);
-		\draw[->] (resample\i 1) -- (adder\i 0);
-		\draw[->] (adder\i 0) -- (out\i);
-	}
-
-	\node[draw, left=0.5cm of in0] (whiten) {\emph{whiten}};
-	\node[signal, label=left:$x$, label=below:{\small 16384 Hz}, left=0.75cm of whiten] (x) {};
-	\node[below=1.5cm of in1] (ellip) {$\vdots$};
-	\node[right=1.6cm of ellip] {$\vdots$};
-	\node[right=4.1cm of ellip] {$\vdots$};
-	\node[above=1mm of out3] {\vdots};
-	\draw[->] (x) -- (whiten);
-	\draw[->] (whiten) -- (in0);
-	\draw[->] (in0) -- (in1);
-	\draw (in1) -- (ellip);
-	\draw[->] (ellip) -- (in2);
+		\draw[->] (add\j 0) -- (out\j 0);
 
@@
 
-	\node[above=0.5cm of ortho0,align=center] (fir label) {Orthogonal \\ FIR filters \\ $\overbrace{\hspace{1cm}}^{}$};
-	\node[right=0.1cm of fir label,align=center] {Reconstruction \\ matrices \\ $\overbrace{\hspace{2.75cm}}^{}$};
-	\node[right=3.5cm of fir label,align=center] {Upsampling and \\ SNR accumulation \\ $\overbrace{\hspace{2.75cm}}^{}$};
-	\node[left=0.1cm of fir label,align=center] {Down-\\ sampling \\ $\overbrace{\hspace{1cm}}^{}$};
-	\node[left=1.75cm of fir label,align=center] {Data acquisition \\ and whitening \\ $\overbrace{\hspace{2cm}}^{}$};
+	\foreach \i in {0,1,2,3} {
+		\draw[->] (resample\i 1) -- (add\i 0);
+		\draw[->] (resample\i 2) -- (add\i 1);
+		\draw (add\i 2) -- (vdots\i);
+		\draw[->] (vdots\i) -- (resample\i 2);
+		\draw[->] (add\i 1) -- (resample\i 1);
+	}
 
+	\node[draw,fill=white,minimum width=22mm,xshift=1mm,label=right:{\small delay}] at (delay0) {\tiny$z^{-t^1 f^0}$};
+	\node[draw,fill=white,minimum width=22mm,xshift=1mm,label=right:{\small delay}] at (delay1) {\tiny$z^{-t^2 f^1}$};
+	\draw[->] (in0) -- (in1);
+	\draw (in1) -- (invdots);
+	\draw[->] (invdots) -- (in2);
+	\draw[->] (x) -- (in0);
+
+	\draw[decorate, decoration={brace, amplitude=4pt}] ($(in)+(-4mm,0)$) -- ($(in)+(4mm,0)$) node[midway,above=2mm,align=center] {Decimation \\ of input};
+	\draw[decorate, decoration={brace, amplitude=4pt}] ($(intee)+(-2mm,0)$) -- ($(u)+(2mm,0)$) node[midway,above=2mm,align=center] {Orthogonal \\ FIR filters};
+	\draw[decorate, decoration={brace, amplitude=4pt}] ($(utee)+(-2mm,0)$) -- ($(v)+(2mm,0)$) node[midway,above=2mm,align=center] {Reconstruction \\ matrices};
+	\draw[decorate, decoration={brace, amplitude=4pt}] ($(add0)+(-2mm,0)$) -- ($(add3)+(2mm,0)$) node[midway,above=2mm,align=center] {Interpolation and \\ SNR accumulation};
 \end{tikzpicture}

### 2011-06-01 analysis.tex
adding more to section 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@799 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 online, low latency filtering stage.  The offline stage can be done before the
-analysis is started an updated asychronously, whereas the online stage must keep
+analysis is started and updated asychronously, whereas the online stage must keep
 up with the detector output and produce search results as rapidly as possible.
@@
 Dividing the mass space into smaller sub-banks also aids in the computational
-cost of the singular value decomposition.  Using our understanding of the
-time-frequency evolution of the templates, we choose time slice boundaries as
-in \eqref{FIXME} such that all of the templates within a sub-bank are
-sub-critically sampled at progressively lower sample rates.  Next, the
-templates within the sub-bank are realized as \textsc{fir} filter coefficients.
-For each time slice, the templates are downsampled to the appropriate sample
-rate.  Finally, the \textsc{svd} is applied to each time slice in order to
-produce a set of orthogonal \textsc{fir} filters and a reconstruction matrix
-that maps them back to the original templates.  The downsampled orthogonal
-\textsc{fir} filter coefficients, the reconstruction matrix, and the time slice
-boundaries are all saved to disk.
+cost of the singular value decomposition and is the approach considered
+in~\cite{Cannon:2010p10398}.  Using our understanding of the time-frequency
+evolution of the templates, we choose time slice boundaries as in \eqref{FIXME}
+such that all of the templates within a sub-bank are sub-critically sampled at
+progressively lower sample rates.  Next, the templates within the sub-bank are
+realized as \textsc{fir} filter coefficients.  For each time slice, the
+templates are downsampled to the appropriate sample rate.  Finally, the
+\textsc{svd} is applied to each time slice in the sub-bank in order to produce
+a set of orthogonal \textsc{fir} filters and a reconstruction matrix that maps
+them back to the original templates as described in \eqref{FIXME}.  The
+downsampled orthogonal \textsc{fir} filter coefficients, the reconstruction
+matrix, and the time slice boundaries are all saved to disk.
 
@@
 
+The filtering algorithm described in the previous section could be used in a
+true sample-in-sample-out realtime system.  This would likely require a
+nontrivial integration directly into the data acquisition and storage system of
+the gravitational wave observatories.  A slightly more modest goal is to leverage
+existing low latency, but not realtime, signal processing infrastructure in
+order to implement the algorithm described in the previous section.
+
 We have implemented a prototype of the low latency filtering stage using an
-open source signal processing environment called GStreamer \cite{gstreamer}.
-GStreamer is a vital component of many Linux systems, providing media playback,
-authoring, and streaming on devices from cell phones to desktop computers to
-streaming media servers.  It turns out that it is also uniquely suited to
-gravitational wave data analysis.  In our application, GStreamer excels at
-queueing, synchronizing, adding, and bookkeeping many different signals at
-different sample rates.  It also provides some useful signal processing
-primitives such as decimators, \textsc{fir} filters, and interpolators.  Most
-importantly, it permits us to recruit all of the host system's \textsc{cpu}s
-without having to write complicated and error-prone multithreaded code.
+open source signal processing environment called \gstreamer\ \cite{gstreamer}.
+\gstreamer\ is a vital component of many Linux systems, providing media
+playback, authoring, and streaming on devices from cell phones to desktop
+computers to streaming media servers.  Given the similarities of gravitational
+wave detector data to audio data it is not surprising that \gstreamer\ is
+useful.  In our application, GStreamer excels at queueing, synchronizing,
+adding, and bookkeeping many different signals at different sample rates.  It
+also provides some useful signal processing elements such as decimators,
+\textsc{fir} filters, and interpolators.
 
@@
 filter to the data, and then downsampling by deleting samples.  We use a
-192-tap \textsc{fir} decimator provided by GStreamer.
+192-tap \textsc{fir} decimator provided by {\tt Gstreamer's audioresample} element.
+The detector data is provided at every power-of-two sample rate required by the 
+template time sliced described in \ref{FIXME}.
 
 \subsubsection{Time delays}
+
+Rather than implement the time sliced templates as zero padded \fir\ filters as
+described in \eqref{FIXME} we instead implement them as shorter, but
+appropriately time delayed \fir\ filters that contain only the nonzero samples.
+In order to assure that we add the \SNR from the time slices adds appropriate
 
@@
 \subsubsection{SNR accumulation}
-
-%Our new method consists of six distinct stages:
-%\begin{enumerate}
-%\item\label{item:decimation} Decimation of the detector data to successively lower sample rates
-%\item\label{item:delay} Delay lines to appropriately synchronize of all of the time slices
-%\item\label{item:ortho} Application of orthogonal \textsc{fir} filters to the decimated detector data
-%\item\label{item:reconstruct} Mixing of orthogonal filter output using reconstruction matrices
-%\item\label{item:interpolation} Interpolation of reconstructed output to a common sample rate
-%\item\label{item:accumulation} Accumulation of \textsc{snr} from each time slice
-%\end{enumerate}

### 2011-06-01 analysis.tex
more stuff
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@801 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In this section we describe an implementation of the method described in
-section \ref{label{sec:method}} suitable for rapid gravitational-wave searches
-for compact binary coalescence.  The previous method requires several
+In this section we describe an implementation of the \lloid\ method described
+in section \ref{label{sec:method}} suitable for rapid gravitational-wave
+searches for compact binary coalescence.  The \lloid\ method requires several
 computations that can be done before the analysis is actually underway.  Thus
 we divide the procedure into two stages 1) an offline planning stage and 2) an
-online, low latency filtering stage.  The offline stage can be done before the
-analysis is started and updated asychronously, whereas the online stage must keep
-up with the detector output and produce search results as rapidly as possible.
-In the next two subsections we describe what these stages entail.
+online, low-latency filtering stage.  The offline stage can be done before the
+analysis is started and updated asychronously, whereas the online stage must
+keep up with the detector output and produce search results as rapidly as
+possible.  In the next two subsections we describe what these stages entail.
 
@@
 
-The filtering algorithm described in the previous section could be used in a
-true sample-in-sample-out realtime system.  This would likely require a
-nontrivial integration directly into the data acquisition and storage system of
-the gravitational wave observatories.  A slightly more modest goal is to leverage
-existing low latency, but not realtime, signal processing infrastructure in
-order to implement the algorithm described in the previous section.
+The \lloid\ algorithm could be used in a true sample-in-sample-out realtime
+system.  However, such a system would likely require integration directly into
+the data acquisition and storage system of the gravitational wave
+observatories.  A slightly more modest goal is to leverage existing low
+latency, but not realtime, signal processing infrastructure in order to
+implement the \lloid\ algorithm.  For the near-term this should be a viable 
+solution for ultra-low latency searches with $\sim 1s$ intrinsic latency.
 
@@
 wave detector data to audio data it is not surprising that \gstreamer\ is
-useful.  In our application, GStreamer excels at queueing, synchronizing,
-adding, and bookkeeping many different signals at different sample rates.  It
-also provides some useful signal processing elements such as decimators,
-\textsc{fir} filters, and interpolators.
-
+useful for our purpose. \gstreamer\ also provides some useful stock signal
+processing elements such as decimators, filters, and interpolators.  
+We have extended the \gstreamer\ framework by developing custom elements
+that are available in the \gstlal\ library~\cite{gstlal}.
+Figure \ref{fig:pipeline} describes schematically how we implement the \lloid\
+algorithm using \gstlal\ and \gstreamer\ components.
+%
+%
 \begin{figure}[htbp]
 	\includegraphics{figures/lloid-diagram.pdf}
-	\caption{Schematic of LLOID pipeline illustrating signal flow.
-Circles with arrows represent upsampling \protect\includegraphics{figures/upsample-symbol.pdf} or downsampling \protect\includegraphics{figures/downsample-symbol.pdf}.
-Circles with plus signs represent summing junctions \protect\includegraphics{figures/adder-symbol.pdf}.
-Squares \protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.
-Sample rate decreases from the top of the diagram to the bottom.}
+	\caption{\label{fig:pipeline} Schematic of LLOID pipeline illustrating
+signal flow.  Circles with arrows represent upsampling
+\protect\includegraphics{figures/upsample-symbol.pdf} or downsampling
+\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
+signs represent summing junctions
+\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
+\protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.  Sample
+rate decreases from the top of the diagram to the bottom.}
 \end{figure}
 
-The filter pipeline consists of six distinct stages.
+The filter pipeline drawn in figure \ref{fig:pipeline} consists of several
+distinct stages and parallel branches.  We outline the stages below.
 
@@
 filter to the data, and then downsampling by deleting samples.  We use a
-192-tap \textsc{fir} decimator provided by {\tt Gstreamer's audioresample} element.
-The detector data is provided at every power-of-two sample rate required by the 
-template time sliced described in \ref{FIXME}.
-
-\subsubsection{Time delays}
-
-Rather than implement the time sliced templates as zero padded \fir\ filters as
-described in \eqref{FIXME} we instead implement them as shorter, but
-appropriately time delayed \fir\ filters that contain only the nonzero samples.
-In order to assure that we add the \SNR from the time slices adds appropriate
-
-Each decimated detector data stream becomes the input for one time slice, but
-it must be appropriately delayed.
+192-tap \textsc{fir} decimator provided by {\tt Gstreamer's audioresample}
+element.  The detector data is provided at every power-of-two sample rate
+required by the template time sliced described in \eqref{FIXME}.  These
+parallel decimated data streams are fed into parallel fir filter banks in the
+next stage.
 
@@
 
+The \fir\ filter banks are implemented using a \gstlal\ element called {\tt
+lal\_firbank}, which produces N channels of filter output from an NxM matrix of
+\fir\ filter coefficients.  This element is used in parallel branches in the
+pipeline to implement the filtering of the SVD basis filters in each time
+slice.  Rather than implement the time sliced templates as zero padded \fir\
+filters as described in \eqref{FIXME} we instead implement them as shorter, but
+appropriately time advanced \fir\ filters that contain only the nonzero
+samples.  Adding the appropriate time offset to the filter output makes up for
+the lack of explicit zero padding.  The orthogonal filter outputs must be
+reconstructed into the filter output for the underlying physical filter ouputs.
+
 \subsubsection{Reconstruction}
+
+The orthogonal filter output provided by the SVD must be used in linear
+combinations in order to obtain the physical filter output for the template
+waveforms.  This involves a matrix multiplication.  We use a \gstlal\ element
+called {\tt lal\_matrix\_mixer} on the output of each orthogonal filter bank in
+order to produce the filter output for the original waveforms in each time
+slice.  At this point the time slice filter outputs are not yet at the same
+sample rate.
 
@@
 
-\subsubsection{SNR accumulation}
+Before we can construct the final \SNR\, the individual time slice filter
+outputs must be brought to a common sample rate, or upsampled.  This is done
+using the same \gstreamer\ element as was used for decimation, {\tt
+audioresample}.
+
+\subsubsection{\SNR accumulation}
+
+In order to accumulate \SNR\ for an early warning, the time slice filter
+output is upsampled to the next highest frequency time slice and summed. This
+process is iterated until the final \SNR is computed.  However at each stage
+of upsampling the summed filter outputs are related to the early warning \SNR
+by a simple normalization factor that can be computed in advance. Thus the
+\lloid\ algorithm and this implementation leads to a simple early warning
+pipeline with little additional work.

### 2011-06-01 results.tex
added figure
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@806 ccdd05d5-56c1-461a-bf98-85f6876523b4
some section headings
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@809 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+In this section we test our implementation of the \lloid\ method using the
+pipeline described in the previous section.  We answer the following questions
+\begin{itemize}
+\item What is the measured loss of SNR due to the approximations used in the \lloid\ algorithm in a realistic analysis pipeline?
+\item What is the measured latency of the implementation described in section 
+\ref{sec:implementation}?
+\item What is the measured computational cost of the implementation described in section
+\ref{sec:implementation}?
+\end{itemize}
+
+\subsection{Measured \SNR\ loss}
+
+\begin{figure}
+\includegraphics{resamp_mm.pdf}
+\caption{\label{fig:resamp_mm}}
+\end{figure}
+
+\subsection{Measured latency}
+
+\subsection{Measured computational cost}
+
 \FIXME{
-I would like to see this section changed to emphasize the low-latency time domain filtering
-1. Use a live simulated white noise source (this ignores the latency of whitening, but that goes
-beyond the scope of this paper, and we mention this and perhaps suggest some exporation
-2. Use TD filtering of N templates
-3. Present the performance and latency and provide estimates for number of cores required for realtime ALIGO search based on current infrastructure
-4. Compute the impulse responses of the templates and histogram the SNR loss for the SVD and time/slice/resampling
-5. Put  a tee after some of the low frequency stages and perhaps test the possibility of
-predictive filtering (also with latency measurments)
-This will remove the complications of running a full pipeline.  We need to sort things out before
-we do that, and this paper should not be delayed any further.  The above tests will make the point we need to make
+I would like to see this section changed to emphasize the low-latency time
+domain filtering 1. Use a live simulated white noise source (this ignores the
+latency of whitening, but that goes beyond the scope of this paper, and we
+mention this and perhaps suggest some exporation 2. Use TD filtering of N
+templates 3. Present the performance and latency and provide estimates for
+number of cores required for realtime ALIGO search based on current
+infrastructure 4. Compute the impulse responses of the templates and histogram
+the SNR loss for the SVD and time/slice/resampling 5. Put  a tee after some of
+the low frequency stages and perhaps test the possibility of predictive
+filtering (also with latency measurments) This will remove the complications of
+running a full pipeline.  We need to sort things out before we do that, and
+this paper should not be delayed any further.  The above tests will make the
+point we need to make
 }

### 2011-06-01 analysis.tex
label and typos
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@810 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \section{Implementation}
+\label{sec:implementation}
 
@@
 
-\subsubsection{\SNR accumulation}
+\subsubsection{\SNR\ accumulation}
 
@@
 output is upsampled to the next highest frequency time slice and summed. This
-process is iterated until the final \SNR is computed.  However at each stage
-of upsampling the summed filter outputs are related to the early warning \SNR
+process is iterated until the final \SNR\ is computed.  However at each stage
+of upsampling the summed filter outputs are related to the early warning \SNR\
 by a simple normalization factor that can be computed in advance. Thus the

### 2011-06-01 results.tex
filled in some text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@811 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+We expect two contributions to the \SNR\ loss to arise in our implementation of
+the \lloid\ algorithm.  The first is the \SNR\ loss due to the truncation of
+the SVD basis, which is fundamental, and estimates for it
+exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
+implementations of resampling that cause signal loss.  We have measured the
+effect of the resampling in the pipeline alone on a single waveform to gauge
+the magnitude of the loss.  It is shown in figure \ref{fig:resamp_mm} as a
+function of the quality of the {\tt audioresample} element (which is
+proportional to the filter length used).  In both cases we have to compare the
+\SNR\ loss to the expected \SNR\ loss that arises from the discreteness of the
+template bank which is typically $\sim 3\%$.  We will consider a comparable
+\SNR\ loss to be tolerable, but ideally would like to have a factor of 10 lower
+loss from the \lloid\ implementation (i.e. no more than $\sim 0.3 \%$).  
+%
 \begin{figure}
 \includegraphics{resamp_mm.pdf}
-\caption{\label{fig:resamp_mm}}
+\caption{\label{fig:resamp_mm}Mis-match as a function of the {\tt
+audioresample} element's quality parameter.  Increasing the quality changes the
+length of the sinc function filter that is used to decimate and interpolate the
+data and filter output in our test pipeline.  Choosing a quality of 4
+corresponds to a filter length of 64 samples and results in an acceptable loss
+of .2 \% in \SNR\ for the chosen waveform.}
 \end{figure}

### 2011-06-01 results.tex
removed old text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@814 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 pipeline described in the previous section.  We answer the following questions
-\begin{itemize}
-\item What is the measured loss of SNR due to the approximations used in the \lloid\ algorithm in a realistic analysis pipeline?
-\item What is the measured latency of the implementation described in section 
+%
+I) What is the measured loss of SNR due to the approximations used in the
+\lloid\ algorithm in a realistic analysis pipeline?
+%
+II) What is the measured latency of the implementation described in section
 \ref{sec:implementation}?
-\item What is the measured computational cost of the implementation described in section
-\ref{sec:implementation}?
-\end{itemize}
+%
+III) What is the measured computational cost of the implementation described in
+section \ref{sec:implementation}?
 
@@
 
-\FIXME{
-I would like to see this section changed to emphasize the low-latency time
-domain filtering 1. Use a live simulated white noise source (this ignores the
-latency of whitening, but that goes beyond the scope of this paper, and we
-mention this and perhaps suggest some exporation 2. Use TD filtering of N
-templates 3. Present the performance and latency and provide estimates for
-number of cores required for realtime ALIGO search based on current
-infrastructure 4. Compute the impulse responses of the templates and histogram
-the SNR loss for the SVD and time/slice/resampling 5. Put  a tee after some of
-the low frequency stages and perhaps test the possibility of predictive
-filtering (also with latency measurments) This will remove the complications of
-running a full pipeline.  We need to sort things out before we do that, and
-this paper should not be delayed any further.  The above tests will make the
-point we need to make
-}
+%\FIXME{
+%I would like to see this section changed to emphasize the low-latency time
+%domain filtering 1. Use a live simulated white noise source (this ignores the
+%latency of whitening, but that goes beyond the scope of this paper, and we
+%mention this and perhaps suggest some exporation 2. Use TD filtering of N
+%templates 3. Present the performance and latency and provide estimates for
+%number of cores required for realtime ALIGO search based on current
+%infrastructure 4. Compute the impulse responses of the templates and histogram
+%the SNR loss for the SVD and time/slice/resampling 5. Put  a tee after some of
+%the low frequency stages and perhaps test the possibility of predictive
+%filtering (also with latency measurments) This will remove the complications of
+%running a full pipeline.  We need to sort things out before we do that, and
+%this paper should not be delayed any further.  The above tests will make the
+%point we need to make
+%}
 
-\subsection{Detector noise characteristics}
+%\subsection{Detector noise characteristics}
 
-We tested the new detection method with mock Advanced \textsc{ligo} data having a power spectrum prescribed by the ``zero detuning, high power'' noise model in \cite{Shoemaker:2009p9770}.
+%We tested the new detection method with mock Advanced \textsc{ligo} data having a power spectrum prescribed by the ``zero detuning, high power'' noise model in \cite{Shoemaker:2009p9770}.

### 2011-06-01 analysis.tex
fixed some wording about delays
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@815 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 slice.  Rather than implement the time sliced templates as zero padded \fir\
-filters as described in \eqref{FIXME} we instead implement them as shorter, but
-appropriately time advanced \fir\ filters that contain only the nonzero
-samples.  Adding the appropriate time offset to the filter output makes up for
-the lack of explicit zero padding.  The orthogonal filter outputs must be
-reconstructed into the filter output for the underlying physical filter ouputs.
+filters as described in \eqref{FIXME} we instead implement them as shorter
+filters that contain only the nonzero samples.  Adding the appropriate time
+offset to the filter output later in the pipeline makes up for the lack of
+explicit zero padding.  The orthogonal filter outputs must be reconstructed
+into the filter output for the underlying physical filter ouputs.
 

### 2011-06-02 results.tex
added hist
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@816 ccdd05d5-56c1-461a-bf98-85f6876523b4
tweaked plot
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@817 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added box and whisker plot
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@822 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed macro
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@825 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
-I) What is the measured loss of SNR due to the approximations used in the
+I) What is the measured loss of \SNR\ due to the approximations used in the
 \lloid\ algorithm in a realistic analysis pipeline?
@@
 
+\begin{figure}
+	\label{fig:hist}
+	\begin{center}
+		\includegraphics{bw.pdf}
+		\caption{Box-and-whisker plot of mismatch between nominal template bank and \lloid\ measured impulse respones.
+The \textsc{svd} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
+The whiskers denote the minimum and maximum mismatch over all templates.}
+	\end{center}
+\end{figure}
+
 \subsection{Measured latency}

### 2011-06-02 analysis.tex
Fixed name of element
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@826 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 waveforms.  This involves a matrix multiplication.  We use a \gstlal\ element
-called {\tt lal\_matrix\_mixer} on the output of each orthogonal filter bank in
+called {\tt lal\_matrixmixer} on the output of each orthogonal filter bank in
 order to produce the filter output for the original waveforms in each time

### 2011-06-02 results.tex
some more text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@828 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \caption{\label{fig:resamp_mm}Mis-match as a function of the {\tt
-audioresample} element's quality parameter.  Increasing the quality changes the
-length of the sinc function filter that is used to decimate and interpolate the
-data and filter output in our test pipeline.  Choosing a quality of 4
-corresponds to a filter length of 64 samples and results in an acceptable loss
-of .2 \% in \SNR\ for the chosen waveform.}
+audioresample} element's filter length.  Increasing the filter length changes
+the length of the sinc function filter that is used to decimate and interpolate
+the data and filtered time series in our test pipeline.  Choosing a filter
+length of 64 samples results in an acceptable loss of .2 \% in \SNR\ for the
+chosen waveform.}
 \end{figure}
@@
 		\includegraphics{bw.pdf}
-		\caption{Box-and-whisker plot of mismatch between nominal template bank and \lloid\ measured impulse respones.
-The \textsc{svd} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
-The whiskers denote the minimum and maximum mismatch over all templates.}
+		\caption{Box-and-whisker plot of mismatch between nominal
+template bank and \lloid\ measured impulse respones.  The \textsc{svd}
+tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
+The whiskers denote the minimum and maximum mismatch over all templates.  The
+diminishing improvement is caused by hitting the resampling match limit.}
 	\end{center}
@@
 
+In order to check the latency of our implementation of the \lloid\ algorithm we
+used the pipeline described in \ref{fig:pipeline} with a live source connected
+to the input.  We then measured the timestamp of the live source and the output
+of the final \SNR\ time series.  Since \gstreamer\ is not a
+sample-in-sample-out realtime infrastructure, it does work by passing small
+buffers of data and filtered data through the pipeline.  We set the buffer size
+on the input to be 1/4 s.  In an ideal case the latency should be nearly 1/4 s.
+However, it is necessary to queue things in the pipeline to assure that there
+are no issues with threading. Sometimes buffers get delayed.  We measured the
+average filtering latency to be \FIXME{1 s}.
+
 \subsection{Measured computational cost}
 
-%\FIXME{
-%I would like to see this section changed to emphasize the low-latency time
-%domain filtering 1. Use a live simulated white noise source (this ignores the
-%latency of whitening, but that goes beyond the scope of this paper, and we
-%mention this and perhaps suggest some exporation 2. Use TD filtering of N
-%templates 3. Present the performance and latency and provide estimates for
-%number of cores required for realtime ALIGO search based on current
-%infrastructure 4. Compute the impulse responses of the templates and histogram
-%the SNR loss for the SVD and time/slice/resampling 5. Put  a tee after some of
-%the low frequency stages and perhaps test the possibility of predictive
-%filtering (also with latency measurments) This will remove the complications of
-%running a full pipeline.  We need to sort things out before we do that, and
-%this paper should not be delayed any further.  The above tests will make the
-%point we need to make
-%}
-
-%\subsection{Detector noise characteristics}
-
-%We tested the new detection method with mock Advanced \textsc{ligo} data having a power spectrum prescribed by the ``zero detuning, high power'' noise model in \cite{Shoemaker:2009p9770}.
+Using the pipeline described previously, we also measured the computational
+cost.  We used an 8 socket, quad core AMD\texttrademark\ machine operating at
+2.7 GHz per core.  We found that filtering \FIXME{657} templates in real time
+required about \FIXME{6} of the available cores.  We thus conclude that we
+require about 1 core for every \FIXME{100} templates.  We note that this is
+well within the computational limits of a modern, computing cluster assuming
+that the scaling holds for the entire mass range of interest, we would need
+about \FIXME{300} cores per detector.  This requirment should be even less of a
+burden in the advanced detecor era, when presumably clusters built in that era
+will have a factor of 4 or 8 more cores than the ones today.

### 2011-06-02 notation.tex
more words
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@794 ccdd05d5-56c1-461a-bf98-85f6876523b4
thoughts
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@830 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-05-31 notation.tex
@@
 \usepackage{stmaryrd}  % gives \shortuparrow and \shortdownarrow
+
+\newcommand{\downsample}[1]{\ensuremath{\, \downarrow^{s-1} \star \,\, #1}}
+\newcommand{\upsample}[1]{\ensuremath{\, \uparrow^{s+1} \star \,\, #1}}
 
@@
 
-Definition of matched filter:
+We define a general discrete time series $p[k]$ with k indicating the time
+samples at the maximum sample rate of the analysis.  We define a time series
+$p^{s}[k^{s}]$ as being the time series $p[k]$ at a different (and lower)
+sample rate $s$.  It is also necessary to thus define the operation that
+changes the sample rate.
+
+Decimation:
+\begin{equation}
+p^{s-1}[k^{s-1}] = \downsample{p^{s}[k^{s}]}
+\end{equation}
+
+Interpolation:
+\begin{equation}
+p^{s+1}[k^{s+1}] = \upsample{p^{s}[k^{s}]}
+\end{equation}
+
+In general downsampling and upsampling are a convolution of a antialiasing
+filter or an interpolation function, respectively.  In the special case of a
+band limited function, decimation may simply reduce to taking every nth sample
+point, depending on the relative sample rates.
+
+It is also necessary to introduce notation representing the time slices
+of a template $\tau$
+
+Definition of time sliced templates:
+\begin{align}
+h_{i}[k] &= \sum_{\tau} h_{i}^{\tau}[k] \\
+h_{i}^{\tau}[k] &= \left\{ \begin{aligned}
+    h_{i}[k] & \; \text{if } t^{\tau} < \frac{k}{f} \leq t^{\tau+1} \\
+    0 & \; \text{, otherwise }
+    \end{aligned} \right.
+\end{align}
+
+Since the templates are band limited an antialiasing filter is not required to
+downsample, we need only to remove certain sample points as long as the desired
+sample rate is at least twice the frequency content of the template in a time
+
+The index $\tau$ when present will be used to denote a quantity that pertains
+to a given time slice.
+
+Definition of matched filter at the full sample rate:
 \begin{equation}
@@
 
-Definition of time slices:
-\begin{align}
-h_{i}[k] &= \sum_s h_{i}^{s}[k] \\
-h_{i}^{s}[k] &= \left\{ \begin{aligned}
-    h_{i}[k] & \; \text{if } t^{s+1} < \frac{k}{f} \leq t^{s} \\
-    0 & \; \text{, otherwise }
-    \end{aligned} \right. \\
-h_{i}^{s}[k^{s}] &\equiv h_{i}^{s}\!\left[k\frac{f}{f^{s}}\right]
-\end{align}
-
-Throughout, we will use the convention that $k$ indexes time-series sampled at
-the full sample rate $f$ and $k^s$ indexes time-series sampled at a reduced
-sample rate $f^s$.  For the case of band limited functions it is sufficient to
-simply remove sample points to define the downsampled version as is shown in
-the previous equation.
-
-Definition of decimation for a function $f$:
-\begin{equation}
-f[k^{s+1}] = \sum_{n=-N^\shortdownarrow/2}^{N^\shortdownarrow/2-1} b^\shortdownarrow[n] f\!\left[k^{s} - n\right]
-\end{equation}
-
-Definition of interpolation for a function $f$:
-\begin{equation}
-f[k^{s}] = \sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] f\!\left[k^{s+1} - n\right]
-\end{equation}
+Quantities involving the SVD are at a given sample rate $s$ and in a time slice
+$\tau$
 
@@
 \begin{equation}
-h^{s}_{i}\!\left[k^{s}\right] = \sum_{m=0}^{M-1} v^{s}_{im} \sigma^{s}_m u^{s}_{m}[k^{s}] \approx \sum_{l=0}^{L^s-1} v^{s}_{il} \sigma^{s}_l u^{s}_{l}[k^{s}]
+h^{\tau s}_{i}\!\left[k^{s}\right] = \sum_{m=0}^{M-1} v^{\tau s}_{im} \sigma^{\tau s}_m u^{\tau s}_{m}[k^{s}] \approx \sum_{l=0}^{L^s-1} v^{\tau s}_{il} \sigma^{\tau s}_l u^{\tau s}_{l}[k^{s}]
 \end{equation}
@@
 \begin{equation}
-    \rho^{s}_i[k^{s}] = \underbrace{\mystrut\sum_{l=0}^{L^{s}-1} v^{s}_{il} \sigma^{s}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
-    \underbrace{\mystrut\sum_{n=0}^{N^{s} - 1} u^{s}_{l}[n] x[k^{s}-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + 
+    \rho^{\tau s}_i[k^{s}] = \underbrace{\mystrut\sum_{l=0}^{L^{s}-1} v^{\tau s}_{il} \sigma^{\tau s}_l}_{\substack{\text{Reconstruction}\\ \text{with {\sc SVD}}}}
+    \underbrace{\mystrut\sum_{n=0}^{N^{s} - 1} u^{\tau s}_{l}[n] x[k^{s}-n]}_{\substack{\text{orthogonal}\\ \text{{\sc FIR} filter}}} + 
     % + 
     % \underbrace{\mystrut\sum_{n=-N^\shortuparrow/2}^{N^\shortuparrow/2-1} b^\shortuparrow[n] \sum_{q \in \mathbb{Z}} \delta\!\left[k-n, q \frac{f^{s}}{f^{s+1}}\right]}_\text{interpolation filter} 
-    \underbrace{\mystrut\rho^{s+1}_i[k^{s}]}_{\substack{\text{accum.}\\ \text{prior {\sc snr}}}}
+    \underbrace{\mystrut\rho^{\tau - 1 s}_i[k^{s}]}_{\substack{\text{accum.}\\ \text{prior {\sc snr}}}}
 \end{equation}

### 2011-06-02 results.tex
Added second box and whisker plot
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@831 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{figure}
-\includegraphics{resamp_mm.pdf}
-\caption{\label{fig:resamp_mm}Mis-match as a function of the {\tt
-audioresample} element's filter length.  Increasing the filter length changes
-the length of the sinc function filter that is used to decimate and interpolate
-the data and filtered time series in our test pipeline.  Choosing a filter
-length of 64 samples results in an acceptable loss of .2 \% in \SNR\ for the
-chosen waveform.}
+	\label{fig:hist-interpolate}
+	\begin{center}
+		\includegraphics{bw_resample.pdf}
+		\caption{Box-and-whisker plot of mismatch between nominal template bank and \lloid\ measured impulse respones.
+The length $N^\shortuparrow$ of the interpolation filter is varied from 8 to 192.
+The \textsc{svd} tolerance is kept fixed at $(1-10^{-6})$.
+The lines in the center of the boxes denote the medians.
+The upper and lower boundaries of the boxes show the upper and lower quartiles.
+The whiskers denote the minimum and maximum mismatch over all templates.}
+	\end{center}
 \end{figure}
@@
 \begin{figure}
-	\label{fig:hist}
+	\label{fig:hist-svd-tolerance}
 	\begin{center}
@@
 tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
-The whiskers denote the minimum and maximum mismatch over all templates.  The
-diminishing improvement is caused by hitting the resampling match limit.}
+The diminishing improvement is caused by hitting the resampling match limit.}
 	\end{center}

### 2011-06-02 analysis.tex
some more text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@827 ccdd05d5-56c1-461a-bf98-85f6876523b4
more revisions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@833 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In this section we describe an implementation of the \lloid\ method described
-in section \ref{label{sec:method}} suitable for rapid gravitational-wave
+in section \ref{sec:method} suitable for rapid gravitational-wave
 searches for compact binary coalescence.  The \lloid\ method requires several
-computations that can be done before the analysis is actually underway.  Thus
+computations that can be completed before the analysis is underway.  Thus
 we divide the procedure into two stages 1) an offline planning stage and 2) an
@@
 The planning stage begins with choosing templates that cover the space of mass
-parameters with a hexagonal grid \cite{PhysRevD.76.102004} in order to satisfy
-the minimum match criterion, which assures a user specified maximum loss in SNR
-for signals that fall in-between the chosen templates.  Next, the templates are
-subdivided into groups of neighbors called ``sub-banks'' that are appropriately
-sized so that each bank can be efficiently handled by a single computer.
-Dividing the mass space into smaller sub-banks also aids in the computational
-cost of the singular value decomposition and is the approach considered
-in~\cite{Cannon:2010p10398}.  Using our understanding of the time-frequency
-evolution of the templates, we choose time slice boundaries as in \eqref{FIXME}
-such that all of the templates within a sub-bank are sub-critically sampled at
-progressively lower sample rates.  Next, the templates within the sub-bank are
-realized as \textsc{fir} filter coefficients.  For each time slice, the
-templates are downsampled to the appropriate sample rate.  Finally, the
-\textsc{svd} is applied to each time slice in the sub-bank in order to produce
-a set of orthogonal \textsc{fir} filters and a reconstruction matrix that maps
-them back to the original templates as described in \eqref{FIXME}.  The
-downsampled orthogonal \textsc{fir} filter coefficients, the reconstruction
-matrix, and the time slice boundaries are all saved to disk.
+parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to satisfy
+a minimum match criterion.  This assures a user-specified, maximum loss in SNR
+for signals that fall in-between the chosen templates.  Typically the minimum
+match is 97\% corresponding to a maximum mismatch of 3\%.  Next, the templates
+are subdivided into groups of neighbors called ``sub-banks'' that are
+appropriately sized so that each bank can be efficiently handled by a single
+computer.  The neighbors are chosen to have similar chirp mass, which produces
+subbanks with very similar waveforms.  Dividing the mass space into smaller
+sub-banks aids in the computational cost of the singular value decomposition
+and is the approach considered in~\cite{Cannon:2010p10398}.  Using our
+understanding of the time-frequency evolution of the templates, we choose time
+slice boundaries as in \eqref{FIXME} such that all of the templates within a
+sub-bank are sub-critically sampled at progressively lower sample rates.  Next,
+the templates within the sub-bank are realized as \fir\ filter
+coefficients.  For each time slice, the templates are downsampled to the
+appropriate sample rate.  Finally, the \SVD\ is applied to each time
+slice in the sub-bank in order to produce a set of orthogonal \fir\
+filters and a reconstruction matrix that maps them back to the original
+templates as described in \eqref{FIXME}.  The downsampled orthogonal
+\fir\ filter coefficients, the reconstruction matrix, and the time slice
+boundaries are all saved to disk.
 
@@
 useful for our purpose. \gstreamer\ also provides some useful stock signal
-processing elements such as decimators, filters, and interpolators.  
-We have extended the \gstreamer\ framework by developing custom elements
-that are available in the \gstlal\ library~\cite{gstlal}.
-Figure \ref{fig:pipeline} describes schematically how we implement the \lloid\
-algorithm using \gstlal\ and \gstreamer\ components.
+processing elements such as resamplers and filters.  We have extended the
+\gstreamer\ framework by developing custom elements that are available in the
+\gstlal\ library~\cite{gstlal}.  Figure \ref{fig:pipeline} describes
+schematically how we implement the \lloid\ algorithm using \gstlal\ and
+\gstreamer\ components.
 %
@@
 \protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.  Sample
-rate decreases from the top of the diagram to the bottom.}
+rate decreases from the top of the diagram to the bottom.  In this diagram each
+time slice contains three \fir\ filters that are linearly combined to produce
+four output channels.  In a typical pipeline the number of \fir\ filters is
+much less than the number of output channels.}
 \end{figure}
@@
 The filter pipeline drawn in figure \ref{fig:pipeline} consists of several
-distinct stages and parallel branches.  We outline the stages below.
+distinct stages and parallel branches.  The stages are decimation, filtering
+the SVD basis, reconstruction of the original template basis, interpolation and
+accumulation of the early warning and final \SNR\.  We outline the stages
+below.
 
@@
 filter to the data, and then downsampling by deleting samples.  We use a
-192-tap \textsc{fir} decimator provided by {\tt Gstreamer's audioresample}
+192-tap \fir\ decimator provided by {\tt Gstreamer's audioresample}
 element.  The detector data is provided at every power-of-two sample rate
-required by the template time sliced described in \eqref{FIXME}.  These
+required by the template time slices described in \eqref{FIXME}.  These
 parallel decimated data streams are fed into parallel fir filter banks in the
@@
 
-\subsubsection{Orthogonal \textsc{fir} filters}
+\subsubsection{Orthogonal \fir\ filters}
 
@@
 slice.  At this point the time slice filter outputs are not yet at the same
-sample rate.
+sample rate, however, before they are upsampled we sum the reconstructed output
+of any time slices that have a common sample rate.
 
@@
 pipeline with little additional work.
+
+In the next section we present an end-to-end test of the \lloid\ algorithm
+implemented in \gstlal\.

### 2011-06-02 inspiral_svd.tex
better title
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@802 ccdd05d5-56c1-461a-bf98-85f6876523b4
more honest title
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@807 ccdd05d5-56c1-461a-bf98-85f6876523b4
another tweak to the title
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@808 ccdd05d5-56c1-461a-bf98-85f6876523b4
more revisions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@833 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\title[LLOID]{Improving the prospects for multimessenger astronomy with \earlywarning\ detection of compact binary coalescence.}
+\title[LLOID]{Technique for \earlywarning\ detection of compact binary coalescence and its implication for multimessenger astronomy}
 
@@
 
-\input{analysis}
+%\input{analysis}
 
@@
 \textsc{phy}-\oldstylenums{0107417}. CH would like to thank Ilya Mandel for
-many discussions about rate estimates and the prospects of early detection.
-NF would like to thank Alessandra Corsi for illuminating discussions on
-astronomical motivations.
+many discussions about rate estimates and the prospects of early detection.  CH
+would also like to thank Patrick Brady for countless fruitful conversations
+about low latency analysis methods.  NF would like to thank Alessandra Corsi
+for illuminating discussions on astronomical motivations.
 

### 2011-06-02 results.tex
Wrapped text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@832 ccdd05d5-56c1-461a-bf98-85f6876523b4
more revisions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@833 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In this section we test our implementation of the \lloid\ method using the
-pipeline described in the previous section.  We answer the following questions
+pipeline described in the previous section.  We answer the following questions:
 %
@@
 the \lloid\ algorithm.  The first is the \SNR\ loss due to the truncation of
-the SVD basis, which is fundamental, and estimates for it
+the \SVD\ basis, which is fundamental, and estimates for it
 exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
@@
 loss from the \lloid\ implementation (i.e. no more than $\sim 0.3 \%$).
+
+It is also within reason that some \SNR loss could arise from other suboptimal
+implementations of the \lloid\ algorithm in our test pipeline.  To gauge
+accurately how well the pipeline is performing we tested the response to a
+delta function.  This should produce time reversed versions of the original
+template waveforms. By taking the innerproduct of the impulse response for each
+channel with the time reversed template, we can gauge very accurately the loss
+in expected \SNR\ due to the approximations we have made and any inadequacies
+in our implementation.
+
+\subsubsection{Checking the effect of the \SVD\ tolerance}
+
+In this section we check the effect of the \SVD\ tolerance parameter defined in
+~\cite{Cannon:2010p10398}.  By changing the tolerance of the reconstruction of
+the physical waveforms one effects directly the mismatch that results from the
+truncation of the orthogonal filter matrix.  The conditions presented here are
+more complicated than in the original work~\cite{Cannon:2010p10398} due to the
+inclusion of the time sliced templates and resampling.  \FIXME{However, we find
+that the results are quite as expected.}  Figure \ref{fig:hist-svd-tolerance}
+demonstrates that it is possible to acheive typical \SNR\ losses of $\ll1\%$.
+However, we do notice that the mismatch saturates with respect to the tolerance.
+This could be the effect of the resampling, or another \SNR\ loss that we did
+not model or expect.  However the saturation is still an order of magnitude below our target mismatch of $\sim 0.3 \%$.  We find that an \SVD\ tolerance of
+$1-1\times10^{-4}$ is adequate. 
 %
 \begin{figure}
-	\label{fig:hist-interpolate}
 	\begin{center}
-		\includegraphics{bw_resample.pdf}
-		\caption{Box-and-whisker plot of mismatch between nominal template bank and \lloid\ measured impulse respones.
-The length $N^\shortuparrow$ of the interpolation filter is varied from 8 to 192.
-The \textsc{svd} tolerance is kept fixed at $(1-10^{-6})$.
-The lines in the center of the boxes denote the medians.
-The upper and lower boundaries of the boxes show the upper and lower quartiles.
-The whiskers denote the minimum and maximum mismatch over all templates.}
+		\includegraphics{bw.pdf}
+		\caption{\label{fig:hist-svd-tolerance}
+Box-and-whisker plot of mismatch between nominal
+template bank and \lloid\ measured impulse respones.  The \textsc{svd}
+tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
+The diminishing improvement is caused by some other limitation in the pipeline.}
 	\end{center}
@@
 
+\subsubsection{Checking the effect of the interpolation filter length}
+
+Next, keeping the \SVD\ tolerance fixed at $1-1\times10^{-6}$, we checked the
+impact of changing the filter length used to interpolate the decimated data
+back up to the full sample rate.  The results are in figure
+\ref{fig:hist-interpolate}.  We find that a filter length of 16 is sufficient
+to produce mismatches at our target of $\sim 0.3 \%$.
+%
 \begin{figure}
-	\label{fig:hist-svd-tolerance}
 	\begin{center}
-		\includegraphics{bw.pdf}
-		\caption{Box-and-whisker plot of mismatch between nominal
-template bank and \lloid\ measured impulse respones.  The \textsc{svd}
-tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
-The diminishing improvement is caused by hitting the resampling match limit.}
+		\includegraphics{bw_resample.pdf}
+		\caption{\label{fig:hist-interpolate} Box-and-whisker plot of
+mismatch between nominal template bank and \lloid\ measured impulse respones.
+The length $N^\uparrow$ of the interpolation filter is varied from 8 to 192.
+The \textsc{svd} tolerance is kept fixed at $(1-10^{-6})$.  The lines in the
+center of the boxes denote the medians.  The upper and lower boundaries of the
+boxes show the upper and lower quartiles.  The whiskers denote the minimum and
+maximum mismatch over all templates.}
 	\end{center}

### 2011-06-02 analysis.tex
fixed broken tex
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@834 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 coefficients.  For each time slice, the templates are downsampled to the
-appropriate sample rate.  Finally, the \SVD\ is applied to each time
+appropriate sample rate.  Finally, the SVD is applied to each time
 slice in the sub-bank in order to produce a set of orthogonal \fir\
@@
 pipeline with little additional work.
-
-In the next section we present an end-to-end test of the \lloid\ algorithm
-implemented in \gstlal\.

### 2011-06-02 inspiral_svd.tex
fixed broken tex
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@834 ccdd05d5-56c1-461a-bf98-85f6876523b4
spell check
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@835 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\title[LLOID]{Technique for \earlywarning\ detection of compact binary coalescence and its implication for multimessenger astronomy}
+\title[LLOID]{Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy}
 
@@
 The rapid detection of compact binary coalescence with a network of advanced
-gravitational-wave detectors will offer a unique opportunity for multimessenger astronomy.
+gravitational-wave detectors will offer a unique opportunity for multi-messenger astronomy.
 Prompt detection alerts to the astronomical community may make it possible to observe the onset of electromagnetic emission from compact binary coalescence.
 We demonstrate a computationally practical analysis strategy that produces early warning triggers even before gravitational radiation from the final merger has arrived at the detectors.
-With current rate estimates for the Advanced \LIGO{} design configuration, we should detect $\sim$10 sources earlier than 10 seconds before merger in 1 year of livetime.
+With current rate estimates for the Advanced \LIGO{} design configuration, we should detect $\sim$10 sources earlier than 10 seconds before merger in 1 year of live time.
 \end{abstract}
@@
 
-%\input{analysis}
+\input{analysis}
 

### 2011-06-02 results.tex
commented out measured cost and latency
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@836 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In this section we test our implementation of the \lloid\ method using the
-pipeline described in the previous section.  We answer the following questions:
+pipeline described in the previous section.  We calculate the measured \SNR\
+loss due to the approximations of the \lloid\ method and our implementation of
+it.
+%We answer the following questions:
 %
-I) What is the measured loss of \SNR\ due to the approximations used in the
-\lloid\ algorithm in a realistic analysis pipeline?
+%I) What is the measured loss of \SNR\ due to the approximations used in the
+%\lloid\ algorithm in a realistic analysis pipeline?
 %
-II) What is the measured latency of the implementation described in section
-\ref{sec:implementation}?
+%II) What is the measured latency of the implementation described in section
+%\ref{sec:implementation}?
 %
-III) What is the measured computational cost of the implementation described in
-section \ref{sec:implementation}?
+%III) What is the measured computational cost of the implementation described in
+%section \ref{sec:implementation}?
 
@@
 
-We expect two contributions to the \SNR\ loss to arise in our implementation of
-the \lloid\ algorithm.  The first is the \SNR\ loss due to the truncation of
-the \SVD\ basis, which is fundamental, and estimates for it
+We expect two known contributions to the \SNR\ loss to arise in our
+implementation of the \lloid\ algorithm.  The first is the \SNR\ loss due to
+the truncation of the \SVD\ basis, which is fundamental, and estimates for it
 exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
@@
 accurately how well the pipeline is performing we tested the response to a
-delta function.  This should produce time reversed versions of the original
-template waveforms. By taking the innerproduct of the impulse response for each
-channel with the time reversed template, we can gauge very accurately the loss
-in expected \SNR\ due to the approximations we have made and any inadequacies
-in our implementation.
+delta function.  This produces time reversed versions of the original template
+waveforms. By taking the inner product of the impulse response for each channel
+with the time reversed template, we can gauge very accurately the loss in
+expected \SNR\ due to the approximations we have made and any inadequacies in
+our implementation.
 
@@
 
-\subsection{Measured latency}
+%\subsection{Measured latency}
 
-In order to check the latency of our implementation of the \lloid\ algorithm we
-used the pipeline described in \ref{fig:pipeline} with a live source connected
-to the input.  We then measured the timestamp of the live source and the output
-of the final \SNR\ time series.  Since \gstreamer\ is not a
-sample-in-sample-out realtime infrastructure, it does work by passing small
-buffers of data and filtered data through the pipeline.  We set the buffer size
-on the input to be 1/4 s.  In an ideal case the latency should be nearly 1/4 s.
-However, it is necessary to queue things in the pipeline to assure that there
-are no issues with threading. Sometimes buffers get delayed.  We measured the
-average filtering latency to be \FIXME{1 s}.
+%In order to check the latency of our implementation of the \lloid\ algorithm we
+%used the pipeline described in \ref{fig:pipeline} with a live source connected
+%to the input.  We then measured the timestamp of the live source and the output
+%of the final \SNR\ time series.  Since \gstreamer\ is not a
+%sample-in-sample-out realtime infrastructure, it does work by passing small
+%buffers of data and filtered data through the pipeline.  We set the buffer size
+%on the input to be 1/4 s.  In an ideal case the latency should be nearly 1/4 s.
+%However, it is necessary to queue things in the pipeline to assure that there
+%are no issues with threading. Sometimes buffers get delayed.  We measured the
+%average filtering latency to be \FIXME{1 s}.
 
-\subsection{Measured computational cost}
+%\subsection{Measured computational cost}
 
-Using the pipeline described previously, we also measured the computational
-cost.  We used an 8 socket, quad core AMD\texttrademark\ machine operating at
-2.7 GHz per core.  We found that filtering \FIXME{657} templates in real time
-required about \FIXME{6} of the available cores.  We thus conclude that we
-require about 1 core for every \FIXME{100} templates.  We note that this is
-well within the computational limits of a modern, computing cluster assuming
-that the scaling holds for the entire mass range of interest, we would need
-about \FIXME{300} cores per detector.  This requirment should be even less of a
-burden in the advanced detecor era, when presumably clusters built in that era
-will have a factor of 4 or 8 more cores than the ones today.
+%Using the pipeline described previously, we also measured the computational
+%cost.  We used an 8 socket, quad core AMD\texttrademark\ machine operating at
+%2.7 GHz per core.  We found that filtering \FIXME{657} templates in real time
+%required about \FIXME{6} of the available cores.  We thus conclude that we
+%require about 1 core for every \FIXME{100} templates.  We note that this is
+%well within the computational limits of a modern, computing cluster assuming
+%that the scaling holds for the entire mass range of interest, we would need
+%about \FIXME{300} cores per detector.  This requirment should be even less of a
+%burden in the advanced detecor era, when presumably clusters built in that era
+%will have a factor of 4 or 8 more cores than the ones today.

### 2011-06-02 introduction.tex
spell check
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@837 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 % - Motivation for low latency
-%   - Observational applications of near-realtime detection
+%   - Observational applications of near-real-time detection
 %   - Mention S6 and VSR2, online running, EM followup
@@
 Prompt electromagnetic emission can arise as shells of relativistically
-outflowing matter collide in the inner shock. Such an inner shock from a
+out-flowing matter collide in the inner shock. Such an inner shock from a
 compact binary coalescence is believed to be a mechanism for short gamma-ray
@@
 evolve slowly in frequency, so that they can be broken into contiguous
-bandlimited time intervals and processed at possibly lower sample rates.
+band-limited time intervals and processed at possibly lower sample rates.
 Second, inspiral filter banks consist of highly similar templates, admitting

### 2011-06-02 analysis.tex
macros
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@838 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-02 analysis.tex
@@
 computations that can be completed before the analysis is underway.  Thus
-we divide the procedure into two stages 1) an offline planning stage and 2) an
-online, low-latency filtering stage.  The offline stage can be done before the
-analysis is started and updated asychronously, whereas the online stage must
+we divide the procedure into two stages 1) an off-line planning stage and 2) an
+online, low-latency filtering stage.  The off-line stage can be done before the
+analysis is started and updated asynchronously, whereas the online stage must
 keep up with the detector output and produce search results as rapidly as
@@
 parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to satisfy
-a minimum match criterion.  This assures a user-specified, maximum loss in SNR
+a minimum match criterion.  This assures a user-specified, maximum loss in \SNR\
 for signals that fall in-between the chosen templates.  Typically the minimum
@@
 computer.  The neighbors are chosen to have similar chirp mass, which produces
-subbanks with very similar waveforms.  Dividing the mass space into smaller
+sub-banks with very similar waveforms.  Dividing the mass space into smaller
 sub-banks aids in the computational cost of the singular value decomposition
@@
 the templates within the sub-bank are realized as \fir\ filter
-coefficients.  For each time slice, the templates are downsampled to the
+coefficients.  For each time slice, the templates are down-sampled to the
 appropriate sample rate.  Finally, the \SVD\ is applied to each time
@@
 filters and a reconstruction matrix that maps them back to the original
-templates as described in \eqref{FIXME}.  The downsampled orthogonal
+templates as described in \eqref{FIXME}.  The down-sampled orthogonal
 \fir\ filter coefficients, the reconstruction matrix, and the time slice
@@
 
-The \lloid\ algorithm could be used in a true sample-in-sample-out realtime
+The \lloid\ algorithm could be used in a true sample-in-sample-out real-time
 system.  However, such a system would likely require integration directly into
@@
 observatories.  A slightly more modest goal is to leverage existing low
-latency, but not realtime, signal processing infrastructure in order to
+latency, but not real-time, signal processing infrastructure in order to
 implement the \lloid\ algorithm.  For the near-term this should be a viable 
@@
 	\caption{\label{fig:pipeline} Schematic of LLOID pipeline illustrating
-signal flow.  Circles with arrows represent upsampling
-\protect\includegraphics{figures/upsample-symbol.pdf} or downsampling
+signal flow.  Circles with arrows represent up-sampling
+\protect\includegraphics{figures/upsample-symbol.pdf} or down-sampling
 \protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
@@
 distinct stages and parallel branches.  The stages are decimation, filtering
-the SVD basis, reconstruction of the original template basis, interpolation and
-accumulation of the early warning and final \SNR\.  We outline the stages
+the \SVD\ basis, reconstruction of the original template basis, interpolation
+and accumulation of the early warning and final \SNR\.  We outline the stages
 below.
@@
 First, the sample rate of the whitened detector data is reduced to successively
-lower sample rates by decimation.  Decimation involves applying an antialiasing
-filter to the data, and then downsampling by deleting samples.  We use a
+lower sample rates by decimation.  Decimation involves applying an anti-aliasing
+filter to the data, and then down-sampling by deleting samples.  We use a
 192-tap \fir\ decimator provided by {\tt Gstreamer's audioresample}
@@
 \fir\ filter coefficients.  This element is used in parallel branches in the
-pipeline to implement the filtering of the SVD basis filters in each time
+pipeline to implement the filtering of the \SVD\ basis filters in each time
 slice.  Rather than implement the time sliced templates as zero padded \fir\
@@
 explicit zero padding.  The orthogonal filter outputs must be reconstructed
-into the filter output for the underlying physical filter ouputs.
+into the filter output for the underlying physical filter outputs.
 
@@
 
-The orthogonal filter output provided by the SVD must be used in linear
+The orthogonal filter output provided by the \SVD\ must be used in linear
 combinations in order to obtain the physical filter output for the template
@@
 slice.  At this point the time slice filter outputs are not yet at the same
-sample rate, however, before they are upsampled we sum the reconstructed output
+sample rate, however, before they are up-sampled we sum the reconstructed output
 of any time slices that have a common sample rate.
@@
 Before we can construct the final \SNR\, the individual time slice filter
-outputs must be brought to a common sample rate, or upsampled.  This is done
+outputs must be brought to a common sample rate, or up-sampled.  This is done
 using the same \gstreamer\ element as was used for decimation, {\tt
@@
 In order to accumulate \SNR\ for an early warning, the time slice filter
-output is upsampled to the next highest frequency time slice and summed. This
+output is up-sampled to the next highest frequency time slice and summed. This
 process is iterated until the final \SNR\ is computed.  However at each stage
-of upsampling the summed filter outputs are related to the early warning \SNR\
+of up-sampling the summed filter outputs are related to the early warning \SNR\
 by a simple normalization factor that can be computed in advance. Thus the

### 2011-06-02 conclusions.tex
commented some text in the conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@812 ccdd05d5-56c1-461a-bf98-85f6876523b4
removed missing reference
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@839 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 computer cores to analyze a four-detector network of gravitational-wave data
-for such systems assuming spin effects can be ignored in parameterizing the
-expected signals~\cite{the-papers-that-say-that-is-okay}.  This is within the
+for binary neutron stars and stellar mass black holes.  This is within the
 current computing capabilities of the \textsc{lsc} Data Grid~\cite{LDG}.
@@
 
-Latency budget, including `before' and `after' quotes for:
+%Latency budget, including `before' and `after' quotes for:
 
-\begin{itemize}
-\item Data acquisition
-\item Calibration
-\item Data aggregation
-\item Analysis
-\item Localization
-\item Alert
-\item Telescope actuation
-\item Total
-\end{itemize}
+%\begin{itemize}
+%\item Data acquisition
+%\item Calibration
+%\item Data aggregation
+%\item Analysis
+%\item Localization
+%\item Alert
+%\item Telescope actuation
+%\item Total
+%\end{itemize}
 
-Future work:
+%Future work:
 
-\begin{itemize}
-\item Sub-solar mass search
-\item Hierarchical detection
-\end{itemize}
+%\begin{itemize}
+%\item Sub-solar mass search
+%\item Hierarchical detection
+%\end{itemize}

### 2011-06-02 analysis.tex
Added template placement plot
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@841 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 updated spectrum estimates as they are available.
+
+\begin{figure}[htbp]
+	\label{fig:tmpltbank}
+	\begin{center}
+		\includegraphics[scale=0.75]{figures/tmpltbank.png}
+		\caption{Placement of template parameters used in this paper.
+The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.
+We shall design a filter bank to search for a small subset of these template parameters with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
+	\end{center}
+\end{figure}
 

### 2011-06-02 results.tex
Corrected figure reference
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@842 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 effect of the resampling in the pipeline alone on a single waveform to gauge
-the magnitude of the loss.  It is shown in figure \ref{fig:resamp_mm} as a
+the magnitude of the loss.  It is shown in figure \ref{fig:hist-interpolate} as a
 function of the quality of the {\tt audioresample} element (which is

### 2011-06-02 method.tex
Notation update work in progress
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@846 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 expand on the ideas of~\cite{Marion2004, Buskulic2010} that describe a
-multiband decomposition of the compact binary parameter space that resulted in
-search with $\sim$minutes latency in \LIGO{}'s S6 and Virgo's VSR2 science runs.
+multiband decomposition of the compact binary parameter space that resulted in a
+search with minutes latency in \LIGO{}'s S6 and Virgo's VSR2 science runs.
 We combine this with the orthogonal decomposition described
@@
 
-Inspiral signals are parameterized by a set of intrinsic parameters
-$\bar{\theta}$%
-%
-\editorial{Change this to a harpoon or vector symbol.}
-%
-that determine the amplitude and phase evolution of the gravitational wave strain. For systems where the effects of spin can be ignored, the
-intrinsic parameters are the component masses of the binary, $\bar{\theta} =
-(m_1, m_2)$. Searches for gravitational waves from compact binary coalescence
+Inspiral signals are continuously parameterized by a set of intrinsic source parameters
+$\vec\theta$ that determine the amplitude and phase evolution of the gravitational wave strain. For systems where the effects of spin can be ignored, the
+intrinsic source parameters are the component masses of the binary, $\vec\theta =
+(m_1, m_2)$. Searches for inspiral signals
 typically employ matched filter banks that discretely
-sample the possible intrinsic parameters~\cite{findchirppaper}.  The filters
-for the waveforms $h(t,\bar{\theta})$, known as templates, are the waveforms
-weighted by the inverse detector noise amplitude spectral density.
-To construct a template bank, templates are chosen with discrete signal parameters $\theta_1,\, \theta_2,\, \cdots$, $\theta_N$ to assure a
+sample the possible intrinsic parameters~\cite{findchirppaper}.
+For a given source, the strain observed by the detector is a linear combination of two waveforms corresponding to the `$+$' and `$\times$' gravitational wave polarizations, so for any value of $\vec\theta$ we must implement 2 filters.
+The coefficients for the $M$ filters
+are known as templates, and are formed by discretizing and time reversing the waveforms and weighting them by the inverse amplitude spectral density of the detector's noise.
+To construct a template bank, templates are chosen with discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\, \vec\theta_{M/2-1}$ to assure a
 bounded loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}. That is, any possible
-signal within the search space will have a cross-correlation of $\geq0.97$ with
+signal within the a given range of intrinsic source parameters will have an inner product that is $\geqslant 0.97$ with
 at least one template. Such a template bank is said to have a {\em
-minimum match} of 0.97. Data are filtered against each template to produce an \SNR{}
-time-series. Local peak-finding across time and templates determines detection
+minimum match} of 0.97. The data from the detector are whitened and convolved with each template to produce $2M$ \SNR\
+time series. Local peak-finding across time and templates determines detection
 candidates.
-%
-\footnote{There are two gravitational-wave polarizations, $+$ and $\times$. A
-given detector will observe a combination of these polarizations that will
-largely be degenerate with an overall unknown constant phase.  This can be
-maximized over by filtering for quadrature phases and taking the magnitude of
-the result. For simplicity we will ignore that aspect in this work, as it is
-straightforward to generalize but not necessary for understanding any of the
-points that will be made.}
 
-We will denote the $i^\mathrm{th}$ filter with parameters $\bar{\theta_i}$
-as a function of time $x_i(\tau)$.
-In this work we will discuss transformations to a set of filters
-$\{x_i(\tau)\}$.  Some of these transformations are not useful or practical
-over the entire parameter space.  For that reason we assume from here onward
-that the set of filters $\{x_i(\tau)\}$ refer to a set of near-neighbor filters
-that can be chosen as a subset of the full parameter space.  Several such
-subsets can be chosen until all of the filters are a member of one local set.
-\editorial{nvf: I don't think that the splitting of the bank is a crucial point here. Perhaps later if we discuss the up-front cost of the SVD.}
-
-Filtering the detector data $h(t)$ involves a convolution of the data with the
-filter.  For a unit-normalized filter, and whitened detector data, the result
-can be interpreted as the signal-to-noise ratio, $\rho_i(t)$ and is defined as
+Filtering the detector data involves a convolution of the data with the
+template.  For a unit-normalized template $h_i[k]$ and whitened detector data $x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the signal-to-noise ratio, $\rho_i[k]$ defined as
 %
@@
 %
-\begin{eqnarray}
-\rho_i(t) &=& \int x_i(\tau) \, h(t-\tau) \, d \tau \label{eq:SNRTD} \\
-          &=& \int \tilde{x}_i(f) \, \tilde{h}(f) \, \e^{-2\pi\i ft} \label{eq:SNRFD} df,
-\end{eqnarray}
-%
-%
-where the second line is a result of the convolution theorem and
-$\tilde{x}_i(f)$ is the Fourier transform of $x_i(\tau)$ as is $\tilde{h}(f)$ the
-Fourier transform of $h(t)$.
+\begin{equation}
+	\label{eq:SNRTD}
+	\rho_i [k] = \sum_{n=0}^{N-1} h_{i}[n] x [k-n].
+\end{equation}
 
-The evaluation of the integrals in \eqref{eq:SNRTD} and \eqref{eq:SNRFD} are
-implemented as sums over sample points for the digitized gravitational-wave
-detector output.  Discrete Fourier transforms can be computed efficiently
-numerically.  For that reason \eqref{eq:SNRFD} is typically far faster,
-computationally.  To evaluate \eqref{eq:SNRTD} requires \order{\tmpsamps
-\hoftsamps} floating point operations per filter $x_i$, where \tmpsamps\ is the
-number of sample points in the filter $x_i$ and \hoftsamps\ is the number of
-sample points in the data $h$.  Assuming \numtmps\ filters are required, the
-total cost is \order{\numtmps \tmpsamps \hoftsamps} However, \eqref{eq:SNRFD}
-requires only \order{\hoftsamps \log \hoftsamps} operations per filter assuming
-transform lengths that are longer than the filter (i.e. \hoftsamps\ $>$
-\tmpsamps), resulting in a total cost of \order{\numtmps \hoftsamps \log
-\hoftsamps}.  In most cases \tmpsamps\ $\gg$ $\log$ \hoftsamps\ and the
-computational savings by choosing the frequency-domain integral form
-\eqref{eq:SNRFD} is clear.  However, to take full advantage of the
-computational efficiency of \eqref{eq:SNRFD} requires an acausal knowledge of
-the detector data $h(t)$, which implies an inherent latency.  In contrast,
-\eqref{eq:SNRTD} can be updated every time a new sample point of detector data
-is taken.
+Equation~(\ref{eq:SNRTD}) can be implemented in the time domain as an \textsc{fir} filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point operations per sample.
+However, it is typically much more computationally efficient to use the convolution theorem and the \textsc{fft} to implement fast convolution, requiring only $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample.
 
@@
 
-In order to minimize latency we propose using the time-domain convolution
+\editorial{This paragraph is hereby condemned!
+It's too early to introduce this equation, and this should be more of a summary of our method.}In order to minimize latency we propose using the time-domain convolution
 presented in \eqref{eq:SNRTD}.  However, because the brute-force evaluation of
@@
 
-The first step of the orthogonal decomposition described in \eqref{eq:decomp}
-is to divide the templates into \emph{time slices}.  This is a time-domain
+Our first innovation and the first step of our proposed method
+is to divide the templates into \emph{time slices} in a time-domain
 analogue to the frequency-domain decomposition described in ~\cite{Marion2004,
-Buskulic2010, beauville2006, beauville2008}.  A matched filter is constructed
+Buskulic2010, beauville2006, beauville2008}\editorial{Four citations is way too many!
+We should pick just \emph{one} citation for \textsc{mbta}, and it should be the most recent and complete description of it.}.
+A matched filter is constructed
 for each time slice.  The outputs form an ensemble of partial \SNR{}
 streams.  By linearity, these partial \SNR\ streams can be suitably time
-delayed and summed to reproduce the \SNR\ of the full template.  We will show
+delayed and summed to reproduce the \SNR\ of the full template.  To wit,
+%
+\begin{equation}
+\label{eq:time-slices}
+h_{i}[k] = \sum_{s=0}^{S-1}
+	\cases{%
+		h_i^s[k] & for $t^s \leqslant k / f^0 < t^{s+1}$ \\
+		0 & otherwise
+	}
+\end{equation}
+%
+for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0 t^S = N$.
+We will show
 in the next section that this, combined with the singular value docomposition,
@@
 coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
-template is truncated at some prescribed time $t_0$, or equivalently frequency \fmax.
-This is often chosen to correspond to the \ISCO\ defined previously. An inspiral signal
-will enter the detection band at a low frequency, $f = f_\mathrm{low}$
+template is truncated at some prescribed time $t^0$, or equivalently frequency $f_\textrm{hi}$.
+This is often chosen to correspond to the \ISCO. An inspiral signal
+will enter the detection band at a low frequency, $f = f_\mathrm{low}$,
 corresponding to a time $t_\mathrm{low}$.  The template is assumed to be zero
-outside the interval $[t_\mathrm{low}, t_\mathrm{0})$ and is said to have  a
-duration of $t_\mathrm{0} - t_\mathrm{low}$. It is critically sampled at a
-rate of $2 \fmax$.
+outside the interval $[t_\mathrm{low}, t^0)$ and is said to have  a
+duration of $t^0 - t_\mathrm{low}$. It is critically sampled at a
+rate of $2 f_\mathrm{hi}$.
 
@@
 highest power-of-two sample rate that critically samples the time sliced
-filter.  The time slices for this template consist of the $k$ intervals $(t_k,
-t_{k-1}], \dots, (t_2, t_1], (t_1, t_0]$ sampled at frequencies
-$f_\mathrm{k-1}, \dots, f_1, f_0$ where $f_0 \geq 2 f_\mathrm{ISCO}$, $t_0 =
-t_\mathrm{ISCO}$, $f_{k-1} \geqslant 2 f_\mathrm{low}$, and $t_k \leqslant
-t_\mathrm{low}$. An example time-slice design satisfying these constraints for
-a $1.4 - 1.4 \, M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
+template.  The time slices consist of the $S$ intervals $\left(t^S,
+t^{S-1}\right],\, \dots,\, \left(t^2, t^1\right],\, \left(t^1, t^0\right]$ sampled at frequencies
+$f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0 \geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.
+The time sliced templates may be downsampled without aliasing, so we define them as
+%
+\begin{equation}
+\label{eq:time-sliced-templates}
+h_{i}^{s}[k] \equiv
+	\cases{
+		h_{i}\!\left[k\frac{f}{f^s}\right] & if $t^s \leqslant k/f^s < t^{s+1}$ \\
+		0 & otherwise.
+	}
+\end{equation}
+%
+An example time-slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
 %
@@
 $f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
-with a time frequency structure given by ($\ref{eq:fgw})$. $f_k$ is the sample
-rate of the time slice, $(t_{k+1}, t_k]$ are the boundaries in seconds
+with a time frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
+rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
 preceeding coalescence and \slicessamps\ are the number of sample points in the
-$k^{\mathrm{th}}$ filter.}
+$s^{\mathrm{th}}$ filter.}
 \end{table}
@@
 
-Rather than applying a unique time-slice decomposition to each template
+\editorial{Need to convert some continuous time notation to discrete time here.}Rather than applying a unique time-slice decomposition to each template
 waveform, we find a decomposition that is adequate for the entire set
@@
 \end{equation}
-where \numslices\ is the number of time slices required.  Note that we write
-the relationship as approximate since the resampling implementation may have
-some inherent loss in quality.  We note that by construction these filters are
-orthogonal over the index $k$ since they are disjoint in time.  In the next
-section we examine how to reduce the number of filters $y_{ik}(\tau)$ via
-singular value decomposition to construct a set of filters that is also
-orthogonal within a given time slice.
+where \numslices\ is the number of time slices required.  We note that
+the time slice decomposition in equation~(\ref{eq:time-slices}) is manifestly
+orthogonal since the time slices are disjoint in time.  In the next
+section we examine how to reduce the number of filters within each time slice via
+singular value decomposition of the time sliced templates.
 
@@
 \hline
-\hoftsamps	& The number of sample points in the data \\
 \tmpsamps	& The number of sample points in template $x_i$ \\
@@
 \fmax		& The maximum frequency of a filter \\
-$f_k$		& The sample frequency of the $k^{\mathrm{th}}$ filter \\
+$f^s$		& The sample frequency of the $k^{\mathrm{th}}$ filter \\
 \resampsamps	& The number of sample points in the resample filter
@@
 
+\begin{comment}
 The filter bank can be implemented using finite impulse response (\fir{}) filters,
@@
 
-\editorial{It would be good to illustrate the layout of this particular
-template bank: masses spanned, time-slice layout \dots}
-\begin{table}[!h]
-\label{table:comp_cost}
-\begin{center}
-\input{flop_budget_example}
-\end{center}
-\caption{\hl{FIXME: Operation counts per sample for four different detection methods.}}
-\end{table}
+\end{comment}
 

### 2011-06-02 time_slices.py

@@ -71,13 +71,11 @@ pylab.savefig('time_slices.pdf')
 
 print r"\begin{tabular}{lrr}"
 #FIXME change the symbol for the number of sample points per slice if the macro changes
-print r"$f_k$ (Hz) & $(t_{k+1}$,$t_k]$ (s) & $\slicessamps$ \\"
+print r"$f^s$ (Hz) & $\left(t^{s+1}, t^s\right]$ (s) & $\slicessamps$ \\"
 print r"\hline"
 for slice in slices:
 	begin = slice['begin']
-	if begin != 0.: # arrgh! floats have both negative and positive 0!
-		begin = -begin
-	print r"%(rate)d & $(%(end)g$,$%(begin)g]$ & %(samples)d \\" % {'begin': begin, 'end': -slice['end'], 'rate': slice['rate'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
+	print r"%(rate)d & $(%(end)g, %(begin)g]$ & %(samples)d \\" % {'begin': begin, 'end': slice['end'], 'rate': slice['rate'], 'samples': int(round((slice['end'] - slice['begin']) * slice['rate']))}
 print r"\end{tabular}"
 
 # Write latency

### 2011-06-02 method.tex
Made number of signal parameters explicit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@848 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 are known as templates, and are formed by discretizing and time reversing the waveforms and weighting them by the inverse amplitude spectral density of the detector's noise.
-To construct a template bank, templates are chosen with discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\, \vec\theta_{M/2-1}$ to assure a
+To construct a template bank, templates are chosen with the $M/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\, \vec\theta_{M/2-1}$ to assure a
 bounded loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}. That is, any possible

### 2011-06-02 method.tex
Removed some more continuous time notation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@849 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \end{table}
-%
 
-\editorial{Need to convert some continuous time notation to discrete time here.}Rather than applying a unique time-slice decomposition to each template
-waveform, we find a decomposition that is adequate for the entire set
-$\{x_i\}$.  This is facilated by choosing templates in the set $\{x_i\}$ with
-similar chirp masses, $\mathcal{M}$.  The time-slice decompositions of the
-filters $x_i(\tau)$ lead to new filters $y_{ik}(\tau)$ satisfying
-%
-\begin{equation}
-\label{eq:time_slices}
-x_i(\tau) \approx \sum_k^{\numslices} y_{ik}(\tau)
-\end{equation}
-where \numslices\ is the number of time slices required.  We note that
+Since waveforms with neighboring intrinsic source parameters $\vec\theta$ have similar
+time-frequency evolution, it is possible to design computationally efficient time slices
+for an extended region of parameter space rather than to design different time slices for
+each template.
+
+We note that
 the time slice decomposition in equation~(\ref{eq:time-slices}) is manifestly

### 2011-06-03 method.tex
Demoted some subsubsections to subsections
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@850 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsubsection{Selectively reducing the sample rate of the data and template waveforms}
+\subsection{Selectively reducing the sample rate of the data and template waveforms}
 
@@
 
-\subsubsection{Reducing the number of filters with the singular value
+\subsection{Reducing the number of filters with the singular value
 decomposition}

### 2011-06-03 method.tex
Some tweaks to symbol table
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@851 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \hline
-\tmpsamps	& The number of sample points in template $x_i$ \\
-\numtmps	& The number of templates in the set $\{x_i\}$ \\
+\tmpsamps	& The number of sample points per template \\
+\numtmps		& The number of templates \\
 \numslices	& The number of time slices \\
-\numsvdtmps	& The number of orthogonal filters in time slice $k$ \\
-\slicessamps	& The number of samples in time slice $k$ \\
-\fmax		& The maximum frequency of a filter \\
-$f^s$		& The sample frequency of the $k^{\mathrm{th}}$ filter \\
-\resampsamps	& The number of sample points in the resample filter
+\numsvdtmps	& The number of orthogonal templates in time slice $s$ \\
+\slicessamps	& The number of samples in time slice $s$ \\
+$f^s$		& The sample rate in time slice $s$ \\
+\resampsamps	& The number of coefficients in the decimation and interpolation filters
 \end{tabular}
-\caption{\label{tab:recap} Notation used to describe filtering.  This table
+\caption{\label{tab:recap} Notation used to describe filters.  This table
 provides a quick reference for symbols used.}

### 2011-06-03 method.tex
Moved proposed method subsection below svd
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@852 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed broken equation reference
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@853 ccdd05d5-56c1-461a-bf98-85f6876523b4
some words
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@854 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsection{Proposed method}
+\subsection{The \lloid\ method}
 
-\editorial{This paragraph is hereby condemned!
-It's too early to introduce this equation, and this should be more of a summary of our method.}In order to minimize latency we propose using the time-domain convolution
-presented in \eqref{eq:SNRTD}.  However, because the brute-force evaluation of
-\eqref{eq:SNRTD} is far too costly to be useful, we will consider an
-approximation to \eqref{eq:SNRTD} that can reduce substantially the cost of
-\realtime\ filtering. This approximation has the form
-%
-% orthogonal decomposition filtering
-%
-\begin{equation}
-\rho_i(t) \approx \sum_k^{\numslices} \sum_j^{\numsvdtmps} 
-	\int_{\tau_{k}}^{\tau_{k+1}} v_{ijk} \sigma_{jk} u_{jk}\,(\tau) \, h(t-\tau) \, d \tau \label{eq:decomp}
-\end{equation} 
-%
-%
-where $u_{jk}(\tau)$ is an orthogonal basis set of filters spanning the space
-of $\{x_i(\tau)\}$ and $\sigma_{jk} v_{ijk}$ is a tensor relating the filters
-$u_{jk}(\tau)$ to the original filter set $\{x_i(\tau)\}$.  We claim that with
-a suitable choice of filters $u_{jk}(\tau)$ one can reduce the computational
-cost of \eqref{eq:SNRTD} sufficiently to feasibly search for
-gravitational waves from compact binary coalescence in \realtime.  This
-requires 1) exploiting
-the time-frequency characteristics of the binary waveforms and 2) exploiting the redundancy of the template bank. We describe our
-procedure for producing the decompostion in \eqref{eq:decomp} in the remainder
-of this section.
+In the remainder of this section we explore a method for reducing the
+computational cost of a time domain search for compact binary coalescence.  We
+will show that it is possible to have a theoretically zero latency algorithm
+that competes with the computational cost of the standard frequency domain
+matched filtering, which requires a signficant latency in order to be
+computationally cheap. Our method, \lloid\ involves two transformations to the
+template waveforms that produce a set of orthogonal filters with far fewer
+sample points than the original template waveforms.  The reduction in the
+number of sample points in all of the filters required to search the entire
+parameter space is dramatic.
+
+The first transformation is to chop the time domain templates up into time
+slices.  Since each template piece is disjoint in time, the resulting set for a
+single template is orthogonal.  Given the chirp like structure of the
+templates, the early time slices have significantly lower bandwidth and can be
+safely downsampled.  The downsampling removes a factor of $\sim 10$ samples
+overall and also allows the filters to be evaluated at about a $\sim 10$ times
+lower rate.  This amounts to more than a factor of 100 in the floating point
+operations per second required.  However, the resulting filters are still not
+orthogonal across the mass parameter space, and are in fact highly redundant.
+We use the singular value decomposition to produce an orthogonal filter set
+from the time sliced templates~\cite{Cannon:2010p10398}.  We find that this
+reduces the number of sample points that need to be filtered by another factor
+of $\sim 100$.  The combined methods reduce the number of floating point
+operations to the level where they are competitive with the conventional
+matched filter approach.  In the remainder of this section we describe the
+\lloid\ algorithm in detail and provide some basic computational cost scaling.
 
@@
 than the original number of filters \numtmps.  We combine \eqref{eq:svddecomp}
-with \eqref{eq:time_slices} to arrive at \eqref{eq:decomp}.  In the next
+with \eqref{eq:time-slices} to arrive at \eqref{eq:decomp}.  In the next
 section we compute the expected computational cost scaling of this
@@
 \eqref{eq:SNRTD} and higher latency \fft\ methods.
+
+\subsection{Proposed method}
+
+\editorial{This paragraph is hereby condemned!
+It's too early to introduce this equation, and this should be more of a summary of our method.}In order to minimize latency we propose using the time-domain convolution
+presented in \eqref{eq:SNRTD}.  However, because the brute-force evaluation of
+\eqref{eq:SNRTD} is far too costly to be useful, we will consider an
+approximation to \eqref{eq:SNRTD} that can reduce substantially the cost of
+\realtime\ filtering. This approximation has the form
+%
+% orthogonal decomposition filtering
+%
+\begin{equation}
+\rho_i(t) \approx \sum_k^{\numslices} \sum_j^{\numsvdtmps} 
+	\int_{\tau_{k}}^{\tau_{k+1}} v_{ijk} \sigma_{jk} u_{jk}\,(\tau) \, h(t-\tau) \, d \tau \label{eq:decomp}
+\end{equation} 
+%
+%
+where $u_{jk}(\tau)$ is an orthogonal basis set of filters spanning the space
+of $\{x_i(\tau)\}$ and $\sigma_{jk} v_{ijk}$ is a tensor relating the filters
+$u_{jk}(\tau)$ to the original filter set $\{x_i(\tau)\}$.  We claim that with
+a suitable choice of filters $u_{jk}(\tau)$ one can reduce the computational
+cost of \eqref{eq:SNRTD} sufficiently to feasibly search for
+gravitational waves from compact binary coalescence in \realtime.  This
+requires 1) exploiting
+the time-frequency characteristics of the binary waveforms and 2) exploiting the redundancy of the template bank. We describe our
+procedure for producing the decompostion in \eqref{eq:decomp} in the remainder
+of this section.
 

### 2011-06-03 method.tex
more words
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@855 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 templates, the early time slices have significantly lower bandwidth and can be
-safely downsampled.  The downsampling removes a factor of $\sim 10$ samples
-overall and also allows the filters to be evaluated at about a $\sim 10$ times
-lower rate.  This amounts to more than a factor of 100 in the floating point
-operations per second required.  However, the resulting filters are still not
-orthogonal across the mass parameter space, and are in fact highly redundant.
-We use the singular value decomposition to produce an orthogonal filter set
-from the time sliced templates~\cite{Cannon:2010p10398}.  We find that this
-reduces the number of sample points that need to be filtered by another factor
-of $\sim 100$.  The combined methods reduce the number of floating point
-operations to the level where they are competitive with the conventional
-matched filter approach.  In the remainder of this section we describe the
-\lloid\ algorithm in detail and provide some basic computational cost scaling.
+safely downsampled.  The downsampling removes a factor of $\sim 10$ samples on
+average and also allows the filters to be evaluated at about a $\sim 10$ times
+lower rate on average.  This amounts to more than a factor of 100 in the
+floating point operations per second required.  However, the resulting filters
+are still not orthogonal across the mass parameter space, and are in fact
+highly redundant.  We use the singular value decomposition to produce an
+orthogonal filter set from the time sliced templates~\cite{Cannon:2010p10398}.
+We find that this reduces the number of sample points that need to be filtered
+by another factor of $\sim 100$.  The combined methods reduce the number of
+floating point operations to the level where they are competitive with the
+conventional matched filter approach.  In the remainder of this section we
+describe the \lloid\ algorithm in detail and provide some basic computational
+cost scaling.
 
@@
 
-Our first innovation and the first step of our proposed method
-is to divide the templates into \emph{time slices} in a time-domain
-analogue to the frequency-domain decomposition described in ~\cite{Marion2004,
-Buskulic2010, beauville2006, beauville2008}\editorial{Four citations is way too many!
-We should pick just \emph{one} citation for \textsc{mbta}, and it should be the most recent and complete description of it.}.
-A matched filter is constructed
-for each time slice.  The outputs form an ensemble of partial \SNR{}
-streams.  By linearity, these partial \SNR\ streams can be suitably time
-delayed and summed to reproduce the \SNR\ of the full template.  To wit,
+The first step of our proposed method is to divide the templates into
+\emph{time slices} in a time-domain analogue to the frequency-domain
+decomposition described in ~\cite{Marion2004, Buskulic2010, beauville2006,
+beauville2008}\editorial{Four citations is way too many!  We should pick just
+\emph{one} citation for \textsc{mbta}, and it should be the most recent and
+complete description of it.}.  A matched filter is constructed for each time
+slice.  The outputs form an ensemble of partial \SNR{} streams.  By linearity,
+these partial \SNR\ streams can be suitably time delayed and summed to
+reproduce the \SNR\ of the full template.  To wit,
 %

### 2011-06-03 analysis.tex
fixed equation references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@856 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 appropriately sized so that each bank can be efficiently handled by a single
-computer.  The neighbors are chosen to have similar chirp mass, which produces
+computer.  The neighbors are chosen to have comparable chirp mass, which produces
 sub-banks with very similar waveforms.  Dividing the mass space into smaller
-sub-banks aids in the computational cost of the singular value decomposition
+sub-banks reduces the computational cost of the singular value decomposition
 and is the approach considered in~\cite{Cannon:2010p10398}.  Using our
@@
 the \SVD\ basis, reconstruction of the original template basis, interpolation
-and accumulation of the early warning and final \SNR\.  We outline the stages
+and accumulation of the early warning and final {\SNR}.  We outline the stages
 below.
@@
 element.  The detector data is provided at every power-of-two sample rate
-required by the template time slices described in \eqref{FIXME}.  These
+required by the template time slices described in \eqref{eq:time-slices}.  These
 parallel decimated data streams are fed into parallel fir filter banks in the
@@
 slice.  Rather than implement the time sliced templates as zero padded \fir\
-filters as described in \eqref{FIXME} we instead implement them as shorter
+filters as described in \eqref{eq:time-slices} we instead implement them as shorter
 filters that contain only the nonzero samples.  Adding the appropriate time

### 2011-06-03 method.tex
Added notation for detection statistic
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@857 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\editorial{This paragraph is hereby condemned!
-It's too early to introduce this equation, and this should be more of a summary of our method.}In order to minimize latency we propose using the time-domain convolution
+\editorial{I just dropped our new notation in here; I haven't edited the text yet.}%
+Definition of decimated detector data:
+\begin{equation}
+	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k]
+\end{equation}
+
+In order to minimize latency we propose using the time-domain convolution
 presented in \eqref{eq:SNRTD}.  However, because the brute-force evaluation of
@@
 \begin{equation}
-\rho_i(t) \approx \sum_k^{\numslices} \sum_j^{\numsvdtmps} 
-	\int_{\tau_{k}}^{\tau_{k+1}} v_{ijk} \sigma_{jk} u_{jk}\,(\tau) \, h(t-\tau) \, d \tau \label{eq:decomp}
+	\rho_i^s [k] =%
+		% Reconstruction
+		\underbrace{
+			\sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s
+		}_\textrm{\clap{reconstruction}}
+		% Orthogonal FIR filter
+		\overbrace{
+			\sum_{\mathclap{n=0}}^{\mathclap{N^s-1}} u_l^s[n] x^s[k-n]
+		}^\textrm{\clap{orthogonal {\sc fir} filters}}
+		% Plus ...
+		+
+		% Interpolation SNR
+		\underbrace{
+			\left(H^\uparrow \rho_i^{s+1}\right)[k]
+		}_\textrm{\clap{{\sc snr} from previous time slices}}
 \end{equation}

### 2011-06-03 method.tex
Fixed some notation in svd section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@858 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 in~\cite{Cannon:2010p10398}.  Using the results of ~\cite{Cannon:2010p10398} we
-establish that the filters of the previous section can be approximated to high
-accuracy by the expansion in the singular value basis
+establish that the time sliced templates of the previous section can be
+approximated to high accuracy by the expansion
 %
 \begin{equation}
-y_{ik}(\tau) \approx \sum_j^{\numsvdtmps} \sigma_{jk} v_{ijk} u_{j}(\tau)
+h_i^s[k] = \sum_{\mathclap{l=0}}^{\mathclap{M-1}} v_{il}^s \sigma_l^s u_l^s[k] \approx \sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s u_l^s[k]
 \label{eq:svddecomp}
@@
 %
-where $\sigma_{jk}$ is the $j^{\mathrm{th}}$ singular value for the
-$k^{\mathrm{th}}$ time slice, $v_{ijk}$ is an orthogonal matrix for the
-$k^{\mathrm{th}}$ time slice  and $u_{jk}$ is a new orthogonal basis filter set
-for the $k^{\mathrm{th}}$ time slice.  The authors of \cite{Cannon:2010p10398}
+where $\sigma_l^s$ is the $l^{\mathrm{th}}$ singular value for the
+$s^{\mathrm{th}}$ time slice, $v_{il}^s$ is an orthogonal matrix for the
+$s^{\mathrm{th}}$ time slice, and $u_l^s[k]$ is a new set of orthogonal basis filters
+for the $s^{\mathrm{th}}$ time slice.  The authors of \cite{Cannon:2010p10398}
 showed that to high reconstruction accuracy far fewer filters are needed than

### 2011-06-03 results.tex
Fixed some math typesetting
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@859 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 not model or expect.  However the saturation is still an order of magnitude below our target mismatch of $\sim 0.3 \%$.  We find that an \SVD\ tolerance of
-$1-1\times10^{-4}$ is adequate. 
+$\left(1-10^{-4}\right)$ is adequate. 
 %
@@
 
-Next, keeping the \SVD\ tolerance fixed at $1-1\times10^{-6}$, we checked the
+Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$, we checked the
 impact of changing the filter length used to interpolate the decimated data

### 2011-06-03 method.tex
Updated flops section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@861 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We now examine the computational cost scaling of the approximate implementation
-of \eqref{eq:SNRTD} as \eqref{eq:decomp}.  An actual implementation of this
-decomposition in a working analysis pipeline is discussed in the next section
-along with measured computational requirements.  For convenience, table
-\ref{tab:recap} is a recap of the meaning of various symbols used in this
-calculation.
+of a conventional time domain or frequency domain matched filter bank as compared
+with \lloid{}.  For convenience, \ref{tab:recap} provides a review of the notation
+that we will need in this section.
 %
@@
 %
-\begin{table}
-\begin{tabular}{rl}
-\bf{Symbol}	& \bf{Definition} \\
-\hline
-\tmpsamps	& The number of sample points per template \\
+\Table{\label{tab:recap} Notation used to describe filters.}
+\br
+Symbol	& Definition \\
+\mr
 \numtmps		& The number of templates \\
+\tmpsamps	& The number of samples per template \\
 \numslices	& The number of time slices \\
@@
 $f^s$		& The sample rate in time slice $s$ \\
-\resampsamps	& The number of coefficients in the decimation and interpolation filters
+$N^\shortdownarrow$ & The number of coefficients in the decimation filter \\
+$N^\shortuparrow$ & The number of coefficients in the interpolation filter \\
+\br
 \end{tabular}
-\caption{\label{tab:recap} Notation used to describe filters.  This table
-provides a quick reference for symbols used.}
+\end{indented}
 \end{table}
 
-In table \ref{table:flops} we present the computational cost scaling in
-floating point operations per sample for common tasks in the pipeline.  
-%
-% common tasks FLOPS table
-%
-\begin{table}[htdp]
-\begin{center}
-\begin{tabular}{l l}
-\bf{Process} & \bf{ops/sample} \\
-\hline
-% FIR
-\fir\ matched filter, \numtmps\ templates of length \tmpsamps\ & $2 \numtmps \tmpsamps$ \\
-% FFT
-\fft\ matched filter, \numtmps\ templates of length \tmpsamps\, & \multirow{2}{*}{$\mbox{\fontsize{14}{17}\selectfont $\frac{4 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock}$}$} \\
-% FFT contd...
-$\,\,\,\,\,\,$ blocks of length \fftblock & \\
-% Resampling
-\fir\ resampling filter, length \resampsamps\ for each of \numtmps\ templates& \multirow{2}{*}{$2 \numtmps \resampsamps\ f_1 / f_2$} \\
-% Resampling contd...
-$\,\,\,\,\,\,$ and sample rates $f_1 < f_2$ & \\
-% Matrix multiply
-multiply $M \times L$ real matrix by $L\times1$ real vector & $2 M L$ \\
-%
+\Table{\label{table:flops}Minimum computational cost of the \textsc{fir}, \textsc{fft} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
+\br
+Method & \flops \\
+\mr
+\textsc{fir} & $2.4\times10^{13}$ \\
+\textsc{fft} convolution & $2.6\times10^8$ \\
+\lloid & $4.7\times10^8$ \\
+\br
 \end{tabular}
-\end{center}
-\caption{Number of floating point operations per sample (multiplications and
-divisions) required for a selection of signal processing operations used in
-\textsc{lloid}.}
-\label{table:flops}
+\end{indented}
 \end{table}
 
-\begin{comment}
-The filter bank can be implemented using finite impulse response (\fir{}) filters,
-which are just sliding window dot products.  If there are \numtmps\ templates
-of length \tmpsamps\, and the data stream contains \hoftsamps\ samples, then
-applying the filter bank requires $2 \numtmps \tmpsamps \hoftsamps$ operations.
+\subsubsection{Conventional time domain method}
 
-More commonly, the matched filters are implemented using the \fft\ convolution.
-This entails applying {\fft}s to blocks of \fftblock\ samples, with $\tmpsamps
-\leq \fftblock$, each block overlapping the previous one by $ \fftblock -
-\tmpsamps $ samples.  There are $\hoftsamps/(\fftblock-\tmpsamps)$ such blocks
-required to filter \hoftsamps\ samples of data.  Modern implementations of the
-Cooley-Tukey \fft, such as the ubiquitous \texttt{fftw}, require about $4 N \lg
-N$ operations to evaluate a \textsc{dft} of size $N$~\cite{Johnson:2007p9654}.
-%
-%
-\editorial{This is more commonly known as ``overlap-save''.  We should find
-someone else's operation count and cite it.}  
-%
-%
-A \fftblock\ sample cross-correlation consists of a forward \fft, a \fftblock
-sample dot product, and an inverse \fft\, totaling $8 \fftblock \lg \fftblock +
-2 D$ operations per block.  Per sample, this is $(8 \lg \fftblock + 2) / (1 -
-\tmpsamps/\fftblock)$ operations.  As this expression indicates the number
-of operations increases as the block size \fftblock\ approaches the
-filter length \tmpsamps.
-%
-%
-\editorial{Drew: Why don't we change this to an overlap of $m$ samples so we
-can see what happens as we increase the overlap to reduce latency.}
+The conventional time domain method consists of a bank of \textsc{fir} filters, or sliding window dot products.
+If there are $\numtmps$ templates, each $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications and additions per sample, or $2 \numtmps \tmpsamps f^0$ floating point operations per second (\flops) at a sample rate $f^0$.
 
-The \fir\ filter implementation has the advantage that it has no intrinsic
-latency, whereas the \fft\ convolution has latency of $\fftblock-\tmpsamps$. 
-%
-%
-%For example, for a $1.4 -
-%1.4 \, M_\odot$ template with duration $\sim 1 \, \mathrm{ks}$, the
-%\textsc{fft} convolution has a latency $\geq 2 \, \mathrm{ks}$.  
-%
-However, the \fir\ filter implementation has the disadvantage of much greater
-overhead per sample than the \fft\ convolution.  For a $1\,\mathrm{ks}$
-template sampled at $4096\,\mathrm{Hz}$, the \textsc{fir} implementation
-requires about about $\tmpsamps / 8 \lg 2 \tmpsamps = 2.2 \times 10^4$ times
-more operations per sample than the \fft{} implementation.  We will now consider
-the computational cost of the \fir\ filter implementation described in the
-previous sections.
+\subsubsection{Conventional frequency domain method}
 
-It is convenient to express the computational cost of the entire filtering
-procedure in floating point operations per second {\flops}.  The cost will be
-the sum of the cost of the \fir\ filtering for the orthogonal filter in each
-time slice plus the cost of reconstructing the original waveforms with matrix
-operations and resampling.  Using the formulas in table \ref{table:flops} we
-arrive at
-%
-\begin{equation}
-%\mathrm{Computational\, cost} 
-\mathrm{C.C.} \propto 2 \sum_k^{\numslices} (\slicessamps + \numtmps) \numsvdtmps f_k \nonumber
-	r \sum_{k, f_k \neq f_{\fmax}}^{\numslices} \numtmps \resampsamps f_k 
-\label{eq:TDcost}
-\end{equation}
-%
-where the first sum is the total cost for filtering and reconstructing the
-orthogonal filters $u_{jk}$.  The second sum is the cost of resampling the
-reconstructed time-slice outputs to the original sample rate {\fmax}.  It
-assumes a \fir\ filter with \resampsamps\ sample points.  Note that the cost
-for resampling only occurs for the downsampled time slices. The resampling cost
-largely depends on {\resampsamps}.  The computational cost of \eqref{eq:TDcost}
-is dominated by the highest frequency terms in the sum.  Comparing just the
-highest frequency term of \eqref{eq:TDcost} with \eqref{eq:SNRTD} shows \order{
-(\slicesamps{\mathrm{max}} + \numtmps) \svdtmps{\mathrm{max}} \fmax} \flops{}
-versus \order{\numtmps \tmpsamps \fmax} {\flops}.  The full calculation for a
-particular patch of parameter space is shown in table \ref{table:comp_cost}.
-\hl{FIXME we need to actually say what went into this}
+The most common frequency domain method is known as the \emph{overlap-save} algorithm, described in \cite{numerical-recipes-chapter-13}.
+It entails splitting the input into blocks of $D$ samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$ samples.
+For each block, the algorithm computes the forward \fft\ of the data and the templates, multiplies them, and then computes the reverse \fft.
 
-\editorial{From here down introduce an example, but don't go through the FFT
-methods in lloid focus just on the comparison of the TD method without lloid
-tricks, the FD method without lloid tricks, and the TD lloid method (no comp
-detection statistic though), remember that the goal is near realtime detection
-not computational cost improvements on FD methods}
+Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about $2 \fftblock \lg \fftblock$ operations to evaluate a real \textsc{dft} of size $\fftblock$~\cite{Johnson:2007p9654}.
+Including the forward transform of the data and $M$ reverse transforms for all of the templates, the \fft\ costs $2 (\numtmps + 1) \fftblock \lg \fftblock$ operations per block.
+The multiplication of the transforms adds a further $2 \numtmps \fftblock$ operations per block.
+Since each block produces $\fftblock - \tmpsamps$ usable samples of output, the overlap-save method requires
+$$
+f^0 \cdot \frac{2 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock} \; \mathrm{\flops}.
+$$
 
-\end{comment}
+\subsubsection{\lloid\ method}
+
+For time slices $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ for evaluating the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ for the linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops to accumulate the resultant partial \SNR\ stream.
+
+The computational cost of the decimation of the detector data and the interpolation of the partial \SNR\ streams is a little bit more subtle.
+Decimation is achieved by applying an \fir\ antialiasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
+Naively, the antialiasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.
+However, it is only necessary to evaluate the antialiasing filter for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.
+Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 N^\shortdownarrow f^s$ \flops\ is possible.
+
+The story is similar for the interpolation filters used to change the sample rates of the partial \textsc{snr} streams.
+Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a lowpass filter with $N^\shortdownarrow$.
+The lowpass filter requires $2 M N^\shortdownarrow f^{s-1}$ \flops.
+However, by taking advantage of the fact that by construction a fraction $f^{s-1}/f$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 M N^\shortuparrow f^s$ \flops.
+
+Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total \lloid\ requires
+$$
+\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right) f \textrm{ \flops.}
+$$
+The second sum is carried out over the set of sample rates of all of the time slices rather than over the time slices themselves because some time slices may have the same sample rates.
 

### 2011-06-03 inspiral_svd.tex
Added acknowledgement for my fellowship
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@862 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed backup title
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@863 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed some accent errors in author addresses
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@864 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added editorial note about addresses for co-authors
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@867 ccdd05d5-56c1-461a-

@@
 
-% Backup title.
-%\title[LLOID]{Method for early-warning detection of compact binary coalescence.}
-
 \date{\today}
@@
 
+\editorial{We have only Institution, City, State, Country for some co-authors.  Can we get rid of the full addresses, or is this at the discretion of the co-authors?}
 \address{$^1$ Canadian Institute for Theoretical Astrophysics, Toronto, ON,
@@
 CA, USA} 
-\address{$^4$ Facultat de Física, Universitat de València, E-46100 Burjassot,
+\address{$^4$ Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, E-46100 Burjassot,
 Spain} 
@@
 Foundation and operates under cooperative agreement
-\textsc{phy}-\oldstylenums{0107417}. CH would like to thank Ilya Mandel for
+\textsc{phy}-\oldstylenums{0107417}.  This research is supported by the National Science Foundation through a Graduate Research Fellowship to LS.  CH would like to thank Ilya Mandel for
 many discussions about rate estimates and the prospects of early detection.  CH

### 2011-06-03 introduction.tex
Added editorial comment about too many detector citations
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@871 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 and/or black holes (BH) is the most promising source of gravitational radiation
-for Advanced \LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb},
+for Advanced \LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb},%
+\editorial{Can we condense all of the detector citations into one -- perhaps find a whitepaper?}%
 and \LCGT~\cite{LCGTWeb}.  Tens of binary coalescence events are expected to be

### 2011-06-03 introduction.tex
'past' --> 'anything before'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@872 ccdd05d5-56c1-461a-bf98-85f6876523b4
Got rid of editorial comment that has been addressed
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@873 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 same field~\cite{2008Natur.455..183R}. Short GRBs, on the other hand, typically
-fade too quickly to observe past the initial burst of gamma rays and hard
+fade too quickly to observe anything before the initial burst of gamma rays and hard
 x-rays. Rapid gravitational-wave transient alerts could enable the observation
@@
 To this end, we have the ambition of reporting candidates not minutes
-\emph{after} the merger, but seconds \emph{before}.  \editorial{I think we
-should say also something about how we can also get the full \SNR{} within moments
-after the coalescence, it is obvious maybe, but we should be explicit}  By
+\emph{after} the merger, but seconds \emph{before}.  By
 looking for threshold crossings before the gravitational-wave signal leaves

### 2011-06-03 introduction.tex
made candidates more specific
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@874 ccdd05d5-56c1-461a-bf98-85f6876523b4
Changed some wording
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@875 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-To this end, we have the ambition of reporting candidates not minutes
+To this end, we have the ambition of reporting gravitational-wave candidates not minutes
 \emph{after} the merger, but seconds \emph{before}.  By
@@
 assume that in the advanced detector era the vetting process will be automated,
-so current trigger generation and telescope actuation would then dominate
-latency.
+so current gravitational wave search methodology and telescope actuation would
+dominate latency.
 
-Predictive detection of \CBC{}s will require striking a balance between latency
+Advance detection of \CBC{}s will require striking a balance between latency
 and throughput.  \CBC{} searches consist of banks of matched filters, or

### 2011-06-03 inspiral_svd.tex
Picked PACS numbers -- same as SVD paper
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@868 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed unused documentclass options
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@869 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added editorial note about author e-mails
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@870 ccdd05d5-56c1-461a-bf98-85f6876523b4
Author list conforms to guidelines: no full stops, last author separated by 'and'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floo

@@
-\documentclass[showpacs,groupedaddress,showkeys,preprintnumbers]{iopart}
+\documentclass{iopart}
 \input{packages}
@@
 
-\author{Kipp Cannon$^{1}$, Romain Cariou$^{2}$, Adrian Chapman$^3$, M.
+\author{Kipp Cannon$^{1}$, Romain Cariou$^{2}$, Adrian Chapman$^3$, M
 Crispin-Ortuzar$^4$, Nickolas Fotopoulos$^3$, Melissa Frei $^5$, Chad
 Hanna$^{6}$, Erin Kara$^{7}$, Drew Keppel$^{8,9}$, Laura Liao$^{10}$, Stephen
-Privitera $^3$, Antony C.  Searle$^3$, Leo Singer$^3$, Alan J.  Weinstein$^3$}
+Privitera $^3$, Antony C Searle$^3$, Leo Singer$^3$ and Alan J  Weinstein$^3$}
 
-\editorial{We have only Institution, City, State, Country for some co-authors.  Can we get rid of the full addresses, or is this at the discretion of the co-authors?}
+\editorial{We have only Institution, City, State, Country for some co-authors.  Can we get rid of the full addresses, or is this at the discretion of the co-authors?  How about author e-mails?}
 \address{$^1$ Canadian Institute for Theoretical Astrophysics, Toronto, ON,
@@
 
-\pacs{95.55.Ym, 84.30.Vn, 95.75.Wx}
+\pacs{04.30.-w, 07.05.Kf}
 

### 2011-06-03 time_slices.py
Adjusted some font sizes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@879 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -15,6 +15,7 @@ matplotlib.rcParams.update({
         "subplots.right": 0.75,
         "subplots.bottom": 0.25,
         "subplots.top": 0.75,
+		"legend.fontsize": 8.0
 })
 
 # Command line interface

### 2011-06-03 time_slices.py
Fixed clipped-off caption
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@880 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -63,7 +63,7 @@ pylab.xlim(-tmax, 0.)
 pylab.ylim(-.15, .15)
 pylab.xlabel('time relative to coalescence (s)')
 pylab.ylabel(r'gravitational wave strain amplitude')
-matplotlib.pyplot.subplots_adjust(left=0.06, right=0.88, top=0.95, bottom=0.12)
+matplotlib.pyplot.subplots_adjust(left=0.06, right=0.88, top=0.95, bottom=0.13)
 
 pylab.savefig('time_slices.pdf')
 

### 2011-06-03 introduction.tex
Introduced term 'CBC'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@876 ccdd05d5-56c1-461a-bf98-85f6876523b4
Centered figure
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@878 ccdd05d5-56c1-461a-bf98-85f6876523b4
Some tweaks to figure 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@881 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{figure}
-\includegraphics{N_before_Tc.pdf}
+\begin{center}
+\includegraphics{figures/snr_in_time.pdf}
 \caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
-be detectable $T_\mathrm{bc}$ seconds before coalescence.  The solid line is the most
-likely yearly rate estimate $N_{\mathrm{re}}$ for Advanced \LIGO\ and the
-shaded region is the interval $N_{\mathrm{low}}$ to $N_{\mathrm{high}}$ from
-\cite{Abadie:2010p10836}.  Note that assuming \SNR\ 8 is sufficient for
-detection and that we observe $N_{\mathrm{re}} = 40$ events per year with a
-detector having the ZERO\_DET\_high\_P noise model described
-in~\cite{ALIGONoiseZERO_DET_high_P}, $\sim10$ sources detectable within 10
-seconds of merger and $\sim1$ sources detectable within 100 seconds of
-merger.  Other noise models, ZERO\_DET\_low\_P~\cite{ALIGONoiseZERO_DET_low_P}
-and BHBH\_20deg~\cite{ALIGONoiseBHBH_20deg} provide better results for early
-detection but at the cost of fewer total events observed above \SNR{} 8.}
+be detectable $t$ seconds before coalescence.  The solid line is the most
+likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
+shaded region corresponds to the confidence interval $\dot N_{\mathrm{low}}$ to $\dot 
+N_{\mathrm{high}}$ described in
+\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in 
+\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for
+detection and that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a
+detector in the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be 
+detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be detectable 
+within 100 seconds of merger.  The `zero detuning, lower power' and `\textsc{bbh} optimized' configurations are still more conducive to advance detection
+but at the cost of fewer total detectable events above an \SNR\ of 8.}
+\end{center}
 \end{figure}
@@
 
-Advance detection of \CBC{}s will require striking a balance between latency
+Advance detection of compact binary coalescences (\CBC{}s) will require striking a balance between latency
 and throughput.  \CBC{} searches consist of banks of matched filters, or

### 2011-06-03 snr_in_time.py

@@ -52,7 +52,7 @@ snrd = {}
 
 # loop over the noise curves
 # ZERO_DET_high_P.txt must be first since it normalizes the rate
-for i, cfile in enumerate(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt', 'data/BHBH_20deg.txt']):
+for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt', 'data/BHBH_20deg.txt'], ['zero detuning, high power', 'zero detuning, low power', 'BHBH optimized'])):
 
 	# Load the data
 	A = numpy.loadtxt(cfile)
@@ -87,15 +87,15 @@ for i, cfile in enumerate(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt'
 
 	# generate the figure
 	ax1 = fig.add_subplot(1,1,1, adjustable='box', aspect=0.5)
-	pylab.loglog(t, num, markers[i], lw=2, label=cfile.replace('.txt','').replace('data/','').replace('_', ' '))
+	pylab.loglog(t, num, markers[i], lw=2, label=label)
 	#pylab.hold(1)
 	pylab.fill_between(t, numminus, numplus, color='k', alpha=0.15)
 
 pylab.grid()
 pylab.legend(loc='lower left')
-pylab.ylabel(r'$\langle N \rangle$ yr$^{-1}$')
-pylab.xlabel(r'Time before coalescence, $T_\mathrm{bc}$')
+pylab.ylabel(r'$\dot N$ yr$^{-1}$')
+pylab.xlabel(r'Time before coalescence, $t$')
 pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.08,right=0.95)
 pylab.xlim([0.01,1000])
 pylab.ylim([.1, 1000])
-pylab.savefig('N_before_Tc.pdf')
+pylab.savefig(sys.argv[1])

### 2011-06-03 introduction.tex
Misc edits to introduction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@882 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \cite{Cannon:2010p10398}.  We will use both aspects to demonstrate that a very
-low latency analysis with predictive detection of compact binary sources is
-possible with current computing resources.  Assuming other aspects of
-gravitational-wave observation latency can be reduced significantly, this
-should allow the possibility for prompt alerts to be sent to the astronomical
-community.
+low latency analysis with advance detection of compact binary sources is
+possible with current computing resources.  Assuming other technical sources of
+latency can be reduced significantly, this should allow the possibility for
+prompt alerts to be sent to the astronomical community.
 

### 2011-06-03 introduction.tex
Don't use shading; can't tell confidence intervals apart
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@883 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
-be detectable $t$ seconds before coalescence.  The solid line is the most
+be detectable $t$ seconds before coalescence.  The heavy solid line is the most
 likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
-shaded region corresponds to the confidence interval $\dot N_{\mathrm{low}}$ to $\dot 
-N_{\mathrm{high}}$ described in
+light solid lines represent the confidence interval $[\dot N_{\mathrm{low}}, \dot 
+N_{\mathrm{high}}]$ described in
 \cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in 

### 2011-06-03 snr_in_time.py

@@ -46,13 +46,13 @@ def snr_to_num(fracsnr, number=40):
 # Start a new figure
 fig = pylab.figure(figsize=(5.5,2.3))
 
-markers = ['k-', 'k--', 'k-.']
+markers = ['k-', 'k--', 'k:']
 
 snrd = {}
 
 # loop over the noise curves
 # ZERO_DET_high_P.txt must be first since it normalizes the rate
-for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt', 'data/BHBH_20deg.txt'], ['zero detuning, high power', 'zero detuning, low power', 'BHBH optimized'])):
+for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt', 'data/BHBH_20deg.txt'], ['zero det., high power', 'zero det., low power', 'BHBH optimized'])):
 
 	# Load the data
 	A = numpy.loadtxt(cfile)
@@ -88,8 +88,10 @@ for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt', 'data/ZERO_D
 	# generate the figure
 	ax1 = fig.add_subplot(1,1,1, adjustable='box', aspect=0.5)
 	pylab.loglog(t, num, markers[i], lw=2, label=label)
+	pylab.loglog(t, numminus, markers[i], lw=0.5)
+	pylab.loglog(t, numplus, markers[i], lw=0.5)
 	#pylab.hold(1)
-	pylab.fill_between(t, numminus, numplus, color='k', alpha=0.15)
+	#pylab.fill_between(t, numminus, numplus, color='k', alpha=0.15)
 
 pylab.grid()
 pylab.legend(loc='lower left')

### 2011-06-03 method.tex
Minor edits to section 2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@884 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 To construct a template bank, templates are chosen with the $M/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\, \vec\theta_{M/2-1}$ to assure a
-bounded loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}. That is, any possible
+bounded loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}\editorial{Can we cut this down to one citation?}. That is, any possible
 signal within the a given range of intrinsic source parameters will have an inner product that is $\geqslant 0.97$ with
@@
 Filtering the detector data involves a convolution of the data with the
-template.  For a unit-normalized template $h_i[k]$ and whitened detector data $x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the signal-to-noise ratio, $\rho_i[k]$ defined as
+templates.  For a unit-normalized template $h_i[k]$ and whitened detector data $x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the signal-to-noise ratio, $\rho_i[k]$ defined as
 %

### 2011-06-03 method.tex
Minor edit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@885 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 matched filtering, which requires a signficant latency in order to be
-computationally cheap. Our method, \lloid\ involves two transformations to the
+computationally cheap. Our method, \lloid\ involves two transformations of the
 template waveforms that produce a set of orthogonal filters with far fewer

### 2011-06-03 method.tex
Fixed appositive
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@886 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 matched filtering, which requires a signficant latency in order to be
-computationally cheap. Our method, \lloid\ involves two transformations of the
+computationally cheap. Our method, \lloid{}, involves two transformations of the
 template waveforms that produce a set of orthogonal filters with far fewer

### 2011-06-03 method.tex
Minor edits in section 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@887 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-03 method.tex
@@
 computational cost of a time domain search for compact binary coalescence.  We
-will show that it is possible to have a theoretically zero latency algorithm
-that competes with the computational cost of the standard frequency domain
-matched filtering, which requires a signficant latency in order to be
-computationally cheap. Our method, \lloid\ involves two transformations of the
-template waveforms that produce a set of orthogonal filters with far fewer
-sample points than the original template waveforms.  The reduction in the
-number of sample points in all of the filters required to search the entire
+will give a truly zero latency algorithm that competes in terms of floating point
+operations per second with the conventional frequency domain method, which requires a
+signficant latency in order to be computationally cheap. Our method, \lloid{},
+involves two transformations of the template waveforms that produce a set of orthogonal
+filters with far fewer coefficients than the original templates.  The reduction
+in the number of sample points in all of the filters required to search the entire
 parameter space is dramatic.
@@
 safely downsampled.  The downsampling removes a factor of $\sim 10$ samples on
+%
+\editorial{Where do these numbers come from?}
+%
 average and also allows the filters to be evaluated at about a $\sim 10$ times
@@
 %
-f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{-t}
+f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
 \right]^{3/8}.
@@
 
-The monotonic time-frequency relationship of \eqref{eq:fgw} allows us to choose
+The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us to choose
 time-slice boundaries that require substantially less bandwidth at early times
 in the inspiral.  Our goal is to reduce the filtering cost of a large fraction
-of the waveform by computing part of the filter integral at a lower sample
-rate.  Specifically we consider here time slice boundaries with the next
-highest power-of-two sample rate that critically samples the time sliced
+of the waveform by computing part of the convolution at a lower sample
+rate.  Specifically we consider here time slice boundaries with the smallest
+power-of-two sample rate that subcritically samples the time sliced
 template.  The time slices consist of the $S$ intervals $\left(t^S,
@@
 rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
-preceeding coalescence and \slicessamps\ are the number of sample points in the
+preceeding coalescence and \slicessamps\ is the number of samples in the
 $s^{\mathrm{th}}$ filter.}

### 2011-06-03 method.tex
Edited sections 2.4 and 2.5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@888 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 particular minimum match by designing an appropriate set of orthonormal {\em
-basis templates}.  A purely numerical technique based on the application of the
-singular value decomposition (\SVD) to inspiral waveforms is demonstrated
-in~\cite{Cannon:2010p10398}.  Using the results of ~\cite{Cannon:2010p10398} we
-establish that the time sliced templates of the previous section can be
-approximated to high accuracy by the expansion
+basis templates}.  A numerical technique based on the
+singular value decomposition (\SVD) of inspiral template banks is demonstrated
+in~\cite{Cannon:2010p10398}.  Similarly, the time sliced templates described above
+can be approximated to arbitrary accuracy by expansion into a set of orthonormal
+\emph{basis templates}, $u_l^s[k]$, related to the original time sliced templates
+through the \emph{reconstruction matrix}, $v_{il}^s\sigma_l^s$:
 %
 \begin{equation}
-h_i^s[k] = \sum_{\mathclap{l=0}}^{\mathclap{M-1}} v_{il}^s \sigma_l^s u_l^s[k] \approx \sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s u_l^s[k]
+h_i^s[k] = \sum_{\mathclap{l=0}}^{\mathclap{M-1}} v_{il}^s \sigma_l^s u_l^s[k] \approx \sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s u_l^s[k].
 \label{eq:svddecomp}
@@
 %
-where $\sigma_l^s$ is the $l^{\mathrm{th}}$ singular value for the
-$s^{\mathrm{th}}$ time slice, $v_{il}^s$ is an orthogonal matrix for the
-$s^{\mathrm{th}}$ time slice, and $u_l^s[k]$ is a new set of orthogonal basis filters
-for the $s^{\mathrm{th}}$ time slice.  The authors of \cite{Cannon:2010p10398}
-showed that to high reconstruction accuracy far fewer filters are needed than
-were in the original template bank. We find that when combined with the
-time-slice decomposition, the number of \SVD\ filters, \numsvdtmps\ is much smaller
-than the original number of filters \numtmps.  We combine \eqref{eq:svddecomp}
-with \eqref{eq:time-slices} to arrive at \eqref{eq:decomp}.  In the next
-section we compute the expected computational cost scaling of this
-decomposition and compare it with the brute-force implementation in
-\eqref{eq:SNRTD} and higher latency \fft\ methods.
+\editorial{Say a word or two about \SVD\ tolerance?}%
+%
+The parameter $\numsvdtmps$ sets the number of basis templates that are kept in the
+approximation.  This determines the \SVD\ tolerance, which affects the \SNR\ loss due
+to the approximation.  The authors of \cite{Cannon:2010p10398} showed that high accuracy
+could be achieved with far fewer basis templates than templates in the original template
+bank.  We find that when combined with the time slice decomposition, the number of basis 
+templates \numsvdtmps\ is much smaller than the original number of templates \numtmps.
+In the next section we describe how we form our novel detection statistic using the time slice decomposition and the \SVD.
 
@@
 
-\editorial{I just dropped our new notation in here; I haven't edited the text yet.}%
 Definition of decimated detector data:
@@
 
+\editorial{Chad: This whole subsection needs a lot of work.
+I just dropped our new notation in here.}%
 In order to minimize latency we propose using the time-domain convolution
@@
 
+In the next section we compute the expected computational cost scaling of this
+decomposition and compare it with the brute force time domain implementation of
+\eqref{eq:SNRTD} and higher latency frequency domain methods.
+
 \subsection{Comparison of computational costs}
@@
 of a conventional time domain or frequency domain matched filter bank as compared
-with \lloid{}.  For convenience, \ref{tab:recap} provides a review of the notation
+with \lloid{}.  For convenience, table~\ref{tab:recap} provides a review of the notation
 that we will need in this section.

### 2011-06-03 method.tex
Edited sections 2.4 and 2.5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@889 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \mr
-\numtmps		& The number of templates \\
-\tmpsamps	& The number of samples per template \\
-\numslices	& The number of time slices \\
-\numsvdtmps	& The number of orthogonal templates in time slice $s$ \\
-\slicessamps	& The number of samples in time slice $s$ \\
-$f^s$		& The sample rate in time slice $s$ \\
-$N^\shortdownarrow$ & The number of coefficients in the decimation filter \\
-$N^\shortuparrow$ & The number of coefficients in the interpolation filter \\
+\numtmps		& number of templates \\
+\tmpsamps	& number of samples per template \\
+\numslices	& number of time slices \\
+\numsvdtmps	& number of orthogonal templates in time slice $s$ \\
+\slicessamps	& number of samples in time slice $s$ \\
+$f^s$		& sample rate in time slice $s$ \\
+$N^\shortdownarrow$ & number of coefficients in the decimation filter \\
+$N^\shortuparrow$ & number of coefficients in the interpolation filter \\
 \br

### 2011-06-03 method.tex
Made notation table shorter
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@890 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \br
-Symbol	& Definition \\
+	& Definition \\
 \mr
@@
 $f^s$		& sample rate in time slice $s$ \\
-$N^\shortdownarrow$ & number of coefficients in the decimation filter \\
-$N^\shortuparrow$ & number of coefficients in the interpolation filter \\
+$N^\shortdownarrow$ & number of coefficients in decimation filter \\
+$N^\shortuparrow$ & number of coefficients in interpolation filter \\
 \br

### 2011-06-03 method.tex
dft -> transform
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@891 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about $2 \fftblock \lg \fftblock$ operations to evaluate a real \textsc{dft} of size $\fftblock$~\cite{Johnson:2007p9654}.
+Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about $2 \fftblock \lg \fftblock$ operations to evaluate a real transform of size $\fftblock$~\cite{Johnson:2007p9654}.
 Including the forward transform of the data and $M$ reverse transforms for all of the templates, the \fft\ costs $2 (\numtmps + 1) \fftblock \lg \fftblock$ operations per block.

### 2011-06-03 method.tex
Edited section 2.6
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@892 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\Table{\label{table:flops}Minimum computational cost of the \textsc{fir}, \textsc{fft} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
+\editorial{This section's two tables are small.  Any way to put them side by side or next to another figure?}
+
+\Table{\label{table:flops}Minimum computational cost of the \textsc{fir} filter bank, \textsc{fft} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
 \br
@@
 
-For time slices $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ for evaluating the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ for the linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops to accumulate the resultant partial \SNR\ stream.
+For time slices $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ \flops\ 
+for evaluating the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ for the 
+linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops\ to add the resultant partial \SNR\ stream.
 
-The computational cost of the decimation of the detector data and the interpolation of the partial \SNR\ streams is a little bit more subtle.
+The computational cost of the decimation of the detector data is a little bit more subtle.
 Decimation is achieved by applying an \fir\ antialiasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
-Naively, the antialiasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.
-However, it is only necessary to evaluate the antialiasing filter for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.
+Naively, an antialiasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.
+However, it is necessary to evaluate the antialiasing filter only for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.
 Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 N^\shortdownarrow f^s$ \flops\ is possible.
@@
 The story is similar for the interpolation filters used to change the sample rates of the partial \textsc{snr} streams.
-Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a lowpass filter with $N^\shortdownarrow$.
+Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a lowpass filter with $N^\shortdownarrow$ coefficients.
 The lowpass filter requires $2 M N^\shortdownarrow f^{s-1}$ \flops.

### 2011-06-03 implementation.tex
Renamed analysis.tex to implementation.tex
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@893 ccdd05d5-56c1-461a-bf98-85f6876523b4
Edited section 3.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@897 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-03 analysis.tex
@@
 computations that can be completed before the analysis is underway.  Thus
-we divide the procedure into two stages 1) an off-line planning stage and 2) an
-online, low-latency filtering stage.  The off-line stage can be done before the
+we divide the procedure into two stages 1) an offline planning stage and 2) an
+online, low-latency filtering stage.  The offline stage can be done before the
 analysis is started and updated asynchronously, whereas the online stage must
@@
 
-The choice of filter waveforms and singular value decomposition can be done in
+The choice of templates and the \SVD\ can be done in
 advance and will be valid as long as the detector noise spectrum remains
-roughly constant.  New filter waveforms can be computed asynchronously using
+roughly constant%
+\editorial{I'd rather not go into this.}%
+.  New templates and the \SVD\ can be computed asynchronously using
 updated spectrum estimates as they are available.
@@
 	\begin{center}
-		\includegraphics[scale=0.75]{figures/tmpltbank.png}
+		\includegraphics{figures/tmpltbank.png}
 		\caption{Placement of template parameters used in this paper.
@@
 
-The planning stage begins with choosing templates that cover the space of mass
+\begin{table}
+\begin{indented}
+\caption{Filter design for these 657 templates.
+From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.
+The \textsc{svd} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$}
+\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
+\br
+\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
+\ns
+\multicolumn{4}{c}{} &\crule{6} \\
+$f^s$ (Hz) & $(t^{s+1}$&$t^s]$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
+\mr
+4096 & (0.5&0] & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
+512 & (4.5&0.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
+256 & (12.5&4.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
+128 & (76.5&12.5] & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
+64 & (140.5&76.5] & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
+64 & (268.5&140.5] & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
+64 & (396.5&268.5] & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
+32 & (460.5&396.5] & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
+32 & (588.5&460.5] & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
+32 & (844.5&588.5] & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
+32 & (1100.5&844.5] & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
+\br
+\end{tabular}
+\end{indented}
+\end{table}
+
+The planning stage begins with choosing templates that cover the space of source
 parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to satisfy
-a minimum match criterion.  This assures a user-specified, maximum loss in \SNR\
-for signals that fall in-between the chosen templates.  Typically the minimum
-match is 97\% corresponding to a maximum mismatch of 3\%.  Next, the templates
-are subdivided into groups of neighbors called ``sub-banks'' that are
+a minimum match criterion.  This assures a prescribed maximum loss in \SNR\
+for signals that fall between the chosen templates.  Typically the minimum
+match is 0.97 corresponding to a maximum mismatch of 0.03.  Next, the templates
+are subdivided into groups of neighbors called \emph{sub-banks} that are
 appropriately sized so that each bank can be efficiently handled by a single
 computer.  The neighbors are chosen to have comparable chirp mass, which produces
-sub-banks with very similar waveforms.  Dividing the mass space into smaller
-sub-banks reduces the computational cost of the singular value decomposition
-and is the approach considered in~\cite{Cannon:2010p10398}.  Using our
-understanding of the time-frequency evolution of the templates, we choose time
-slice boundaries as in \eqref{FIXME} such that all of the templates within a
+sub-banks with similar time-frequency evolution.  Dividing the source parameter space
+into smaller sub-banks reduces the computational cost of the singular value decomposition
+and is the approach considered in~\cite{Cannon:2010p10398}.  We choose time
+slice boundaries as in equation~\eqref{eq:time-sliced-templates} such that all of the templates within a
 sub-bank are sub-critically sampled at progressively lower sample rates.  Next,
@@
 filters and a reconstruction matrix that maps them back to the original
-templates as described in \eqref{FIXME}.  The down-sampled orthogonal
+templates as described in equation~\eqref{eq:svddecomp}.  The down-sampled orthogonal
 \fir\ filter coefficients, the reconstruction matrix, and the time slice

### 2011-06-03 implementation.tex
Edited implementation section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@898 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 processing elements such as resamplers and filters.  We have extended the
-\gstreamer\ framework by developing custom elements that are available in the
-\gstlal\ library~\cite{gstlal}.  Figure \ref{fig:pipeline} describes
-schematically how we implement the \lloid\ algorithm using \gstlal\ and
-\gstreamer\ components.
+\gstreamer\ framework by developing a library called \gstlal{}~\cite{gstlal}
+that provides elements for gravitational wave data analysis.  Figure
+\ref{fig:pipeline} describes schematically how we implement the \lloid\ algorithm
+using \gstlal\ and \gstreamer\ components.
 %
@@
 	\caption{\label{fig:pipeline} Schematic of LLOID pipeline illustrating
-signal flow.  Circles with arrows represent up-sampling
-\protect\includegraphics{figures/upsample-symbol.pdf} or down-sampling
+signal flow.  Circles with arrows represent interpolation
+\protect\includegraphics{figures/upsample-symbol.pdf} or decimation
 \protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
@@
 
-The filter pipeline drawn in figure \ref{fig:pipeline} consists of several
+The signal flow diagram drawn in figure \ref{fig:pipeline} contains several
 distinct stages and parallel branches.  The stages are decimation, filtering
-the \SVD\ basis, reconstruction of the original template basis, interpolation
+the data against the orthogonal templates, reconstruction of the original template basis, interpolation
 and accumulation of the early warning and final {\SNR}.  We outline the stages
@@
 filter to the data, and then down-sampling by deleting samples.  We use a
-192-tap \fir\ decimator provided by {\tt Gstreamer's audioresample}
+192-tap \fir\ decimator provided by {\tt Gstreamer}'s {\tt audioresample}
 element.  The detector data is provided at every power-of-two sample rate
 required by the template time slices described in \eqref{eq:time-slices}.  These
-parallel decimated data streams are fed into parallel fir filter banks in the
+parallel decimated data streams are fed into parallel \fir\ filter banks in the
 next stage.
@@
 \fir\ filter coefficients.  This element is used in parallel branches in the
-pipeline to implement the filtering of the \SVD\ basis filters in each time
+pipeline to implement the \SVD\ basis filters in each time
 slice.  Rather than implement the time sliced templates as zero padded \fir\
@@
 offset to the filter output later in the pipeline makes up for the lack of
-explicit zero padding.  The orthogonal filter outputs must be reconstructed
-into the filter output for the underlying physical filter outputs.
+explicit zero padding.  The orthogonal filter outputs must next be reconstructed
+into the output of the underlying time sliced templates.
 
@@
 
-The orthogonal filter output provided by the \SVD\ must be used in linear
-combinations in order to obtain the physical filter output for the template
-waveforms.  This involves a matrix multiplication.  We use a \gstlal\ element
-called {\tt lal\_matrixmixer} on the output of each orthogonal filter bank in
-order to produce the filter output for the original waveforms in each time
-slice.  At this point the time slice filter outputs are not yet at the same
-sample rate, however, before they are up-sampled we sum the reconstructed output
-of any time slices that have a common sample rate.
+From the outputs of the \SVD\ basis filters, we form the partial \SNR{}s for each time 
+slice by multiplying by the reconstruction matrices.  This is accomplished by connecting 
+the \texttt{gstlal} element \texttt{lal\_matrixmixer} to the output of
+\texttt{lal\_firbank}.
 
@@
 
-Before we can construct the final \SNR\, the individual time slice filter
-outputs must be brought to a common sample rate, or up-sampled.  This is done
+In order to form the early-warning \SNR\ from each time slice, we hav eto add the partial 
+\SNR\ to the early-warning \SNR\ from the subsequent time slices.  If the next time slice 
+does not have the same sample rate, its output must first be interpolated.  This is done
 using the same \gstreamer\ element as was used for decimation, {\tt
@@
 
-In order to accumulate \SNR\ for an early warning, the time slice filter
-output is up-sampled to the next highest frequency time slice and summed. This
-process is iterated until the final \SNR\ is computed.  However at each stage
-of up-sampling the summed filter outputs are related to the early warning \SNR\
-by a simple normalization factor that can be computed in advance. Thus the
-\lloid\ algorithm and this implementation leads to a simple early warning
+The early-warning \SNR\ for each time slice is formed by accumulating interpolated early
+warning \SNR\ from the subsequent time slice.  This process continuous until we have 
+worked our way to the \SNR\ of the original templates at the full sample rate $f^0$.  In 
+this way, the \lloid\ algorithm and this implementation leads to a simple early warning
 pipeline with little additional work.

### 2011-06-03 conclusions.tex
Added comments to conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@899 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-We have demonstrated a computationally feasible procedure for the rapid
+\editorial{We should drive home the point that our method is as fast as the fft convolution but without any latency at all.}%
+We have demonstrated a computationally feasible procedure for the rapid and even advance
 detection of gravitational waves emitted during the coalescence of neutron
@@
 X-ray down to radio.  We antipate requiring no more than \numcpus\ modern
+%
+\editorial{We need to do this calculation using the \flops\ counts and the number of templates.}%
+%
 computer cores to analyze a four-detector network of gravitational-wave data
@@
 procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
-to a 1/16 Hz cadence introducing an up-front latency of 125 ms.  Data
+to a 1/(16~Hz) cadence introducing an up-front latency of 125~ms.  Data
+%
+\editorial{This is redundant: 1/(16~Hz) and 125~ms are just reciprocals of each other.}%
+%
 aggregation from the observatories will travel over various networks, each
@@
 
+\editorial{The way this is worded, it sounds like a big omission.}%
 We have omitted discussion of source localization. Localization is known to

### 2011-06-03 results.tex
Fixed run-on
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@860 ccdd05d5-56c1-461a-bf98-85f6876523b4
Edited results section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@902 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In this section we test our implementation of the \lloid\ method using the
+In this section we test our implementation of the \lloid\ method using the \gstreamer\
 pipeline described in the previous section.  We calculate the measured \SNR\
@@
 implementation of the \lloid\ algorithm.  The first is the \SNR\ loss due to
-the truncation of the \SVD\ basis, which is fundamental, and estimates for it
+the truncation of the \SVD\ basis and estimates for it
 exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
-implementations of resampling that cause signal loss.  We have measured the
-effect of the resampling in the pipeline alone on a single waveform to gauge
-the magnitude of the loss.  It is shown in figure \ref{fig:hist-interpolate} as a
-function of the quality of the {\tt audioresample} element (which is
-proportional to the filter length used).  In both cases we have to compare the
-\SNR\ loss to the expected \SNR\ loss that arises from the discreteness of the
-template bank which is typically $\sim 3\%$.  We will consider a comparable
-\SNR\ loss to be tolerable, but ideally would like to have a factor of 10 lower
-loss from the \lloid\ implementation (i.e. no more than $\sim 0.3 \%$).
+implementations of resampling that cause signal loss.  The \SNR\ loss is to be compared
+with the mismatch of 0.03 that arises from the discreteness of the
+template bank which is typically.  We will consider an acceptable
+\SNR\ loss to be a factor of 10 smaller than this, i.e., no more than $\sim 0.003$.
 
-It is also within reason that some \SNR loss could arise from other suboptimal
+It is also within reason that some \SNR\ loss could arise from other suboptimal
 implementations of the \lloid\ algorithm in our test pipeline.  To gauge
 accurately how well the pipeline is performing we tested the response to a
-delta function.  This produces time reversed versions of the original template
-waveforms. By taking the inner product of the impulse response for each channel
-with the time reversed template, we can gauge very accurately the loss in
-expected \SNR\ due to the approximations we have made and any inadequacies in
+unit impulse.  By taking the inner product of the impulse response for each channel
+with the template, we can gauge very accurately the loss in
+\SNR\ due to the approximations we have made and any inadequacies in
 our implementation.
@@
 This could be the effect of the resampling, or another \SNR\ loss that we did
-not model or expect.  However the saturation is still an order of magnitude below our target mismatch of $\sim 0.3 \%$.  We find that an \SVD\ tolerance of
+not model or expect.  However the saturation is still an order of magnitude below our target mismatch of $\sim 0.003$.  We find that an \SVD\ tolerance of
 $\left(1-10^{-4}\right)$ is adequate. 
@@
 tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
-The diminishing improvement is caused by some other limitation in the pipeline.}
+The diminishing improvement is caused by some other limitation in the pipeline.
+The lines in the
+center of the boxes denote the medians.  The upper and lower boundaries of the
+boxes show the upper and lower quartiles.  The whiskers denote the minimum and
+maximum mismatch over all templates.}
 	\end{center}
@@
 \ref{fig:hist-interpolate}.  We find that a filter length of 16 is sufficient
-to produce mismatches at our target of $\sim 0.3 \%$.
+to produce mismatches at our target of $\sim 0.003$.
 %
@@
 The length $N^\uparrow$ of the interpolation filter is varied from 8 to 192.
-The \textsc{svd} tolerance is kept fixed at $(1-10^{-6})$.  The lines in the
-center of the boxes denote the medians.  The upper and lower boundaries of the
-boxes show the upper and lower quartiles.  The whiskers denote the minimum and
-maximum mismatch over all templates.}
+The \textsc{svd} tolerance is kept fixed at $(1-10^{-6})$.}
 	\end{center}

### 2011-06-03 introduction.tex
rearrange detector references fix typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@903 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\editorial{Would the introduction be more effective without this paragraph?}
+\editorial{Would the introduction be more effective without this paragraph?, should we find whitepapers?}
+
 The coalescence of compact binary systems consisting of neutron stars (NS)
 and/or black holes (BH) is the most promising source of gravitational radiation
-for Advanced \LIGO~\cite{ALIGOWeb}, Virgo~\cite{AVirgoWeb}, \GEO~\cite{GEOWeb},%
-\editorial{Can we condense all of the detector citations into one -- perhaps find a whitepaper?}%
-and \LCGT~\cite{LCGTWeb}.  Tens of binary coalescence events are expected to be
-observed in the advanced detector era later this
-decade~\cite{Abadie:2010p10836}.
+for Advanced \LIGO, Virgo, \GEO\ and \LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb,
+LCGTWeb}.  Tens of binary coalescence events are expected to be observed in the
+advanced detector era later this decade~\cite{Abadie:2010p10836}.
 
@@
 searches.  The most obvious example of the latter is the time domain
-(\textsc{td}) convolution, which is has no latency at all.  However, its
+(\textsc{td}) convolution, which has no latency at all.  However, its
 computational complexity is quadratic in the length of the templates, so it is

### 2011-06-03 results.tex
Combined two box and whisker plots into one compound figure
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@905 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	\begin{center}
-		\includegraphics{bw.pdf}
-		\caption{\label{fig:hist-svd-tolerance}
-Box-and-whisker plot of mismatch between nominal
-template bank and \lloid\ measured impulse respones.  The \textsc{svd}
-tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
-The diminishing improvement is caused by some other limitation in the pipeline.
-The lines in the
-center of the boxes denote the medians.  The upper and lower boundaries of the
-boxes show the upper and lower quartiles.  The whiskers denote the minimum and
-maximum mismatch over all templates.}
+		\subfloat[Mismatch versus \textsc{svd} tolerance]{\label{fig:hist-svd-tolerance}\includegraphics{bw.pdf}}
+		\subfloat[Mismatch versus $N^\shortuparrow$]{\label{fig:hist-interpolate}\includegraphics{bw_resample.pdf}}
+		\caption{Box-and-whisker plot of mismatch between nominal
+template bank and \lloid\ measured impulse respones.  The upper and lower boundaries of
+the boxes show the upper and lower quartiles; the lines in the center denote the medians.
+The whiskers represent the minimum and maximum mismatch over all templates.  In 
+\subref{fig:hist-svd-tolerance} the interpolation filter length is held fixed
+at $N^\shortuparrow = 192$, while the \SVD\ tolerance is varied from
+$\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
+In \subref{fig:hist-interpolate}, the \SVD\ tolerance is fixed at $\left(1-10^{-6}\right)$ while $N^\shortuparrow$ is varied from 8 to 192 coefficients.}
 	\end{center}
@@
 
-Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$, we checked the
-impact of changing the filter length used to interpolate the decimated data
-back up to the full sample rate.  The results are in figure
-\ref{fig:hist-interpolate}.  We find that a filter length of 16 is sufficient
-to produce mismatches at our target of $\sim 0.003$.
-%
-\begin{figure}
-	\begin{center}
-		\includegraphics{bw_resample.pdf}
-		\caption{\label{fig:hist-interpolate} Box-and-whisker plot of
-mismatch between nominal template bank and \lloid\ measured impulse respones.
-The length $N^\uparrow$ of the interpolation filter is varied from 8 to 192.
-The \textsc{svd} tolerance is kept fixed at $(1-10^{-6})$.}
-	\end{center}
-\end{figure}
+Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$, we studied the
+impact of the length $N^\shortuparrow$ of the interpolation filter.  The results are in
+figure \ref{fig:hist-interpolate}.  We find that a filter length of 16 is sufficient
+to meet our target mismatch of 0.003.
 

### 2011-06-03 conclusions.tex
lloid_paper: Many small fixes and fill-ins
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@906 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 capable of high bandwith but perhaps only modest latency.  This could amount to
-a similar latency of $\sim$ 100 ms.  Lastly, unless a realtime infrastructure is
+a similar latency of $\sim$100 ms.  Lastly, unless a realtime infrastructure is
 adopted post data acquisition, it is likely that there will be an inherent
@@
 opinion, significant work would have to be done in order to improve upon this
-number. However, it should be considered for third generation detector design.
+number. However, it should be considered for third-generation detector design.
 For example, a tighter integration of analysis and data acquisition would be
@@
 \editorial{The way this is worded, it sounds like a big omission.}%
-We have omitted discussion of source localization. Localization is known to
+We have omitted discussion of source sky localization. Localization is known to
 be poor for signals of low \SNR~\cite{Fairhurst2009}.
@@
 simulations using the infrastructure and techniques described in this work.
+\editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring
+the signal recycling mirror to optimize SNR at merger. There's a lot of science there.}
 

### 2011-06-03 figures/diagram.tex
Corrected error in second delay
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@821 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: Many small fixes and fill-ins
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@906 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	\node[draw,fill=white,minimum width=22mm,xshift=1mm,label=right:{\small delay}] at (delay0) {\tiny$z^{-t^1 f^0}$};
-	\node[draw,fill=white,minimum width=22mm,xshift=1mm,label=right:{\small delay}] at (delay1) {\tiny$z^{-t^2 f^1}$};
+	\node[draw,fill=white,minimum width=22mm,xshift=1mm,label=right:{\small delay}] at (delay1) {\tiny$z^{-(t^2-t^1) f^1}$};
 	\draw[->] (in0) -- (in1);
@@
 	\draw[decorate, decoration={brace, amplitude=4pt}] ($(in)+(-4mm,0)$) -- ($(in)+(4mm,0)$) node[midway,above=2mm,align=center] {Decimation \\ of input};
-	\draw[decorate, decoration={brace, amplitude=4pt}] ($(intee)+(-2mm,0)$) -- ($(u)+(2mm,0)$) node[midway,above=2mm,align=center] {Orthogonal \\ FIR filters};
+	\draw[decorate, decoration={brace, amplitude=4pt}] ($(intee)+(-2mm,0)$) -- ($(u)+(2mm,0)$) node[midway,above=2mm,align=center] {Orthogonal \\ \textsc{fir} filters};
 	\draw[decorate, decoration={brace, amplitude=4pt}] ($(utee)+(-2mm,0)$) -- ($(v)+(2mm,0)$) node[midway,above=2mm,align=center] {Reconstruction \\ matrices};
-	\draw[decorate, decoration={brace, amplitude=4pt}] ($(add0)+(-2mm,0)$) -- ($(add3)+(2mm,0)$) node[midway,above=2mm,align=center] {Interpolation and \\ SNR accumulation};
+	\draw[decorate, decoration={brace, amplitude=4pt}] ($(add0)+(-2mm,0)$) -- ($(add3)+(2mm,0)$) node[midway,above=2mm,align=center] {Interpolation and \\ \textsc{snr} accumulation};
 \end{tikzpicture}

### 2011-06-03 implementation.tex
Added missing period
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@904 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: Many small fixes and fill-ins
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@906 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.
-The \textsc{svd} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$}
-\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
+The \SVD{} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
+\item[]\begin{tabular}{rr@{,\,}lc*{6}{r}}
 \br
-\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
+\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 $-$ \textsc{svd} tolerance)} \\
 \ns
 \multicolumn{4}{c}{} &\crule{6} \\
-$f^s$ (Hz) & $(t^{s+1}$&$t^s]$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
+$f^s$ (Hz) & $(t^{s+1}$&$t^s]$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\
 \mr
@@
 system.  However, such a system would likely require integration directly into
-the data acquisition and storage system of the gravitational wave
+the data acquisition and storage system of the gravitational-wave
 observatories.  A slightly more modest goal is to leverage existing low
@@
 
-We have implemented a prototype of the low latency filtering stage using an
-open source signal processing environment called \gstreamer\ \cite{gstreamer}.
+We have implemented a prototype of the low-latency filtering stage using an
+open-source signal processing environment called \gstreamer\ \cite{gstreamer}.
 \gstreamer\ is a vital component of many Linux systems, providing media
 playback, authoring, and streaming on devices from cell phones to desktop
-computers to streaming media servers.  Given the similarities of gravitational
-wave detector data to audio data it is not surprising that \gstreamer\ is
+computers to streaming media servers.  Given the similarities of
+gravitational- wave detector data to audio data it is not surprising that
+\gstreamer\ is
 useful for our purpose. \gstreamer\ also provides some useful stock signal
@@
 \gstreamer\ framework by developing a library called \gstlal{}~\cite{gstlal}
-that provides elements for gravitational wave data analysis.  Figure
+that provides elements for gravitational-wave data-analysis.  Figure
 \ref{fig:pipeline} describes schematically how we implement the \lloid\ algorithm
@@
 	\includegraphics{figures/lloid-diagram.pdf}
-	\caption{\label{fig:pipeline} Schematic of LLOID pipeline illustrating
+	\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
 signal flow.  Circles with arrows represent interpolation
@@
 \protect\includegraphics{figures/adder-symbol.pdf}.  Squares
-\protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.  Sample
+\protect\includegraphics{figures/fir-symbol.pdf} stand for \fir{} filters.  Sample
 rate decreases from the top of the diagram to the bottom.  In this diagram each
@@
 the data against the orthogonal templates, reconstruction of the original template basis, interpolation
-and accumulation of the early warning and final {\SNR}.  We outline the stages
+and accumulation of the early-warning and final {\SNR}.  We outline the stages
 below.
@@
 filter to the data, and then down-sampling by deleting samples.  We use a
-192-tap \fir\ decimator provided by {\tt Gstreamer}'s {\tt audioresample}
+192-tap \fir\ decimator provided by \gstreamer{}'s {\tt audioresample}
 element.  The detector data is provided at every power-of-two sample rate
@@
 The \fir\ filter banks are implemented using a \gstlal\ element called {\tt
-lal\_firbank}, which produces N channels of filter output from an NxM matrix of
+lal\_firbank}, which produces $N$ channels of filter output from an $N\times M$ matrix of
 \fir\ filter coefficients.  This element is used in parallel branches in the
 pipeline to implement the \SVD\ basis filters in each time
-slice.  Rather than implement the time sliced templates as zero padded \fir\
+slice.  Rather than implement the time-sliced templates as zero-padded \fir\
 filters as described in \eqref{eq:time-slices} we instead implement them as shorter
@@
 offset to the filter output later in the pipeline makes up for the lack of
-explicit zero padding.  The orthogonal filter outputs must next be reconstructed
-into the output of the underlying time sliced templates.
+explicit zero-padding.  The orthogonal filter outputs must next be reconstructed
+into the output of the underlying time-sliced templates.
 
@@
 
-In order to form the early-warning \SNR\ from each time slice, we hav eto add the partial 
-\SNR\ to the early-warning \SNR\ from the subsequent time slices.  If the next time slice 
-does not have the same sample rate, its output must first be interpolated.  This is done
-using the same \gstreamer\ element as was used for decimation, {\tt
-audioresample}.
+In order to form the early-warning \SNR\ from each time slice, we have to add
+the partial \SNR\ to the early-warning \SNR\ from the subsequent time slices.
+If the next time slice does not have the same sample rate, its output must
+first be interpolated.  This is done using the same \gstreamer\ element as was
+used for decimation, {\tt audioresample}.
 

### 2011-06-03 introduction.tex
lloid_paper: Many small fixes and fill-ins
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@906 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 (S6) and Virgo completed its third science run (VSR3).  While both
-\textsc{ligo} detectors and Virgo were operating, several all-sky detection
+\LIGO{} detectors and Virgo were operating, several all-sky detection
 pipelines operated in a low-latency configuration, namely \textsc{mbta}, ihope,
@@
 low throughput.  The former are epitomized by the overlap-save algorithm for
-\textsc{fft} convolution, currently the preferred method in gravitational wave
+\fft{} convolution, currently the preferred method in gravitational wave
 searches.  The most obvious example of the latter is the time domain

### 2011-06-03 method.tex
subcritically -> sub-critically
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@895 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: Many small fixes and fill-ins
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@906 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 multiband decomposition of the compact binary parameter space that resulted in a
-search with minutes latency in \LIGO{}'s S6 and Virgo's VSR2 science runs.
+search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 science runs.
 We combine this with the orthogonal decomposition described
@@
 
-Equation~(\ref{eq:SNRTD}) can be implemented in the time domain as an \textsc{fir} filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point operations per sample.
-However, it is typically much more computationally efficient to use the convolution theorem and the \textsc{fft} to implement fast convolution, requiring only $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample.
+Equation~(\ref{eq:SNRTD}) can be implemented in the time domain as an \fir{} filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point operations per sample.
+However, it is typically much more computationally efficient to use the convolution theorem and the \fft{} to implement fast convolution, requiring only $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample.
 
@@
 In the remainder of this section we explore a method for reducing the
-computational cost of a time domain search for compact binary coalescence.  We
-will give a truly zero latency algorithm that competes in terms of floating point
-operations per second with the conventional frequency domain method, which requires a
+computational cost of a time-domain search for compact binary coalescence.  We
+will give a truly zero-latency algorithm that competes in terms of floating point
+operations per second with the conventional frequency-domain method, which requires a
 signficant latency in order to be computationally cheap. Our method, \lloid{},
@@
 
-The first transformation is to chop the time domain templates up into time
+The first transformation is to chop the time-domain templates up into time
 slices.  Since each template piece is disjoint in time, the resulting set for a
@@
 coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
-template is truncated at some prescribed time $t^0$, or equivalently frequency $f_\textrm{hi}$.
+template is truncated at some prescribed time $t^0$, or equivalently frequency $f_\mathrm{hi}$.
 This is often chosen to correspond to the \ISCO. An inspiral signal
@@
 rate.  Specifically we consider here time slice boundaries with the smallest
-power-of-two sample rate that subcritically samples the time sliced
+power-of-two sample rate that sub-critically samples the time sliced
 template.  The time slices consist of the $S$ intervals $\left(t^S,
@@
 $f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0 \geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.
-The time sliced templates may be downsampled without aliasing, so we define them as
+The time-sliced templates may be downsampled without aliasing, so we define them as
 %
@@
 power-of-two time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from
-$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
-with a time frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
+$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{hi} = f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
+with a time-frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
 rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
@@
 \begin{equation}
+\label{eq:decomp}
 	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k]
@@
 the time-frequency characteristics of the binary waveforms and 2) exploiting the redundancy of the template bank. We describe our
-procedure for producing the decompostion in \eqref{eq:decomp} in the remainder
+procedure for producing the decomposition in \eqref{eq:decomp} in the remainder
 of this section.
@@
 In the next section we compute the expected computational cost scaling of this
-decomposition and compare it with the brute force time domain implementation of
-\eqref{eq:SNRTD} and higher latency frequency domain methods.
+decomposition and compare it with the brute-force time-domain implementation of
+\eqref{eq:SNRTD} and higher latency frequency-domain methods.
 
@@
 We now examine the computational cost scaling of the approximate implementation
-of a conventional time domain or frequency domain matched filter bank as compared
+of a conventional time-domain or frequency-domain matched filter bank as compared
 with \lloid{}.  For convenience, table~\ref{tab:recap} provides a review of the notation
@@
 
-\Table{\label{table:flops}Minimum computational cost of the \textsc{fir} filter bank, \textsc{fft} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
+\Table{\label{table:flops}Minimum computational cost of the \fir{} filter bank, \fft{} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
 \br
@@
 \mr
-\textsc{fir} & $2.4\times10^{13}$ \\
-\textsc{fft} convolution & $2.6\times10^8$ \\
+\fir{} & $2.4\times10^{13}$ \\
+\fft{} convolution & $2.6\times10^8$ \\
 \lloid & $4.7\times10^8$ \\
@@
 
-\subsubsection{Conventional time domain method}
+\subsubsection{Conventional time-domain method}
 
-The conventional time domain method consists of a bank of \textsc{fir} filters, or sliding window dot products.
+The conventional time-domain method consists of a bank of \fir{} filters, or sliding-window dot products.
 If there are $\numtmps$ templates, each $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications and additions per sample, or $2 \numtmps \tmpsamps f^0$ floating point operations per second (\flops) at a sample rate $f^0$.
 
-\subsubsection{Conventional frequency domain method}
+\subsubsection{Conventional frequency-domain method}
 
-The most common frequency domain method is known as the \emph{overlap-save} algorithm, described in \cite{numerical-recipes-chapter-13}.
+The most common frequency-domain method is known as the \emph{overlap-save} algorithm, described in \cite{numerical-recipes-chapter-13}.
 It entails splitting the input into blocks of $D$ samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$ samples.
@@
 
-The story is similar for the interpolation filters used to change the sample rates of the partial \textsc{snr} streams.
-Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a lowpass filter with $N^\shortdownarrow$ coefficients.
-The lowpass filter requires $2 M N^\shortdownarrow f^{s-1}$ \flops.
+The story is similar for the interpolation filters used to change the sample rates of the partial \SNR{} streams.
+Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a low-pass filter with $N^\shortdownarrow$ coefficients.
+The low-pass filter requires $2 M N^\shortdownarrow f^{s-1}$ \flops.
 However, by taking advantage of the fact that by construction a fraction $f^{s-1}/f$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 M N^\shortuparrow f^s$ \flops.

### 2011-06-03 method.tex
merged
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@908 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 expand on the ideas of~\cite{Marion2004, Buskulic2010} that describe a
-multiband decomposition of the compact binary parameter space that resulted in a
-search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 science runs.
-We combine this with the orthogonal decomposition described
+multiband decomposition of the compact binary parameter space that resulted in
+a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR2 science runs.
+We combine this with the \SVD\ rank reduction method described
 in~\cite{Cannon:2010p10398} that exploits the redundancy of the template banks.
@@
 
-Inspiral signals are continuously parameterized by a set of intrinsic source parameters
-$\vec\theta$ that determine the amplitude and phase evolution of the gravitational wave strain. For systems where the effects of spin can be ignored, the
-intrinsic source parameters are the component masses of the binary, $\vec\theta =
-(m_1, m_2)$. Searches for inspiral signals
-typically employ matched filter banks that discretely
-sample the possible intrinsic parameters~\cite{findchirppaper}.
-For a given source, the strain observed by the detector is a linear combination of two waveforms corresponding to the `$+$' and `$\times$' gravitational wave polarizations, so for any value of $\vec\theta$ we must implement 2 filters.
-The coefficients for the $M$ filters
-are known as templates, and are formed by discretizing and time reversing the waveforms and weighting them by the inverse amplitude spectral density of the detector's noise.
-To construct a template bank, templates are chosen with the $M/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\, \vec\theta_{M/2-1}$ to assure a
-bounded loss of \SNR~\cite{Owen:1995tm,Owen:1998dk}\editorial{Can we cut this down to one citation?}. That is, any possible
-signal within the a given range of intrinsic source parameters will have an inner product that is $\geqslant 0.97$ with
-at least one template. Such a template bank is said to have a {\em
-minimum match} of 0.97. The data from the detector are whitened and convolved with each template to produce $2M$ \SNR\
-time series. Local peak-finding across time and templates determines detection
-candidates.
+Inspiral signals are continuously parameterized by a set of intrinsic source
+parameters $\vec\theta$ that determine the amplitude and phase evolution of the
+gravitational wave strain. For systems where the effects of spin can be
+ignored, the intrinsic source parameters are the component masses of the
+binary, $\vec\theta = (m_1, m_2)$. Searches for inspiral signals typically
+employ matched filter banks that discretely sample the possible intrinsic
+parameters~\cite{findchirppaper}.  For a given source, the strain observed by
+the detector is a linear combination of two waveforms corresponding to the
+`$+$' and `$\times$' gravitational wave polarizations, so for any value of
+$\vec\theta$ we must implement 2 filters.  The coefficients for the $M$ filters
+are known as templates, and are formed by discretizing and time reversing the
+waveforms and weighting them by the inverse amplitude spectral density of the
+detector's noise. To construct a template bank, templates are chosen with the
+$M/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\,
+\vec\theta_{M/2-1}$ to assure a bounded loss of
+\SNR~\cite{Owen:1995tm,Owen:1998dk}\editorial{Can we cut this down to one
+citation?}. That is, any possible signal within a given range of intrinsic
+source parameters will have an inner product that is $\geqslant 0.97$ with at
+least one template. Such a template bank is said to have a {\em minimum match}
+of 0.97. The data from the detector are whitened and convolved with each
+template to produce $2M$ \SNR\ time series. Local peak-finding across time and
+templates determines detection candidates.
 
 Filtering the detector data involves a convolution of the data with the
-templates.  For a unit-normalized template $h_i[k]$ and whitened detector data $x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the signal-to-noise ratio, $\rho_i[k]$ defined as
+templates.  For a unit-normalized template $h_i[k]$ and whitened detector data
+$x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the
+signal-to-noise ratio, $\rho_i[k]$ defined as
 %
@@
 
-Equation~(\ref{eq:SNRTD}) can be implemented in the time domain as an \fir{} filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point operations per sample.
-However, it is typically much more computationally efficient to use the convolution theorem and the \fft{} to implement fast convolution, requiring only $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample.
+Equation~(\ref{eq:SNRTD}) can be implemented in the time domain (\TD) as an
+\fir\ filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point
+operations per sample.  However, it is typically much more computationally
+efficient to use the convolution theorem and the \fft\ to implement fast
+convolution in the frequency domain (\FD), requiring only $\mathcal O(\numtmps
+\lg \tmpsamps)$ operations per sample.
 
@@
 In the remainder of this section we explore a method for reducing the
-computational cost of a time-domain search for compact binary coalescence.  We
-will give a truly zero-latency algorithm that competes in terms of floating point
-operations per second with the conventional frequency-domain method, which requires a
+computational cost of a \TD\ search for compact binary coalescence.  We will
+give a truly zero latency algorithm that competes in terms of floating point
+operations per second with the conventional \FD\ method, which requires a
 signficant latency in order to be computationally cheap. Our method, \lloid{},
-involves two transformations of the template waveforms that produce a set of orthogonal
-filters with far fewer coefficients than the original templates.  The reduction
-in the number of sample points in all of the filters required to search the entire
-parameter space is dramatic.
+involves two transformations of the template waveforms that produce a set of
+orthogonal filters with far fewer coefficients than the original templates.
+The reduction in sample count for all of the filters required to search the
+entire parameter space is dramatic.
 
 The first transformation is to chop the time-domain templates up into time
-slices.  Since each template piece is disjoint in time, the resulting set for a
+slices.  Since each template slice is disjoint in time, the resulting set for a
 single template is orthogonal.  Given the chirp like structure of the
 templates, the early time slices have significantly lower bandwidth and can be
-safely downsampled.  The downsampling removes a factor of $\sim 10$ samples on
+safely downsampled.  The downsampling removes a factor of $\sim 100$ samples
+from
 %
-\editorial{Where do these numbers come from?}
+\editorial{Where do these numbers come from? CHAD: I have actually changed the statement to 100 for the early time slices.}
 %
-average and also allows the filters to be evaluated at about a $\sim 10$ times
-lower rate on average.  This amounts to more than a factor of 100 in the
-floating point operations per second required.  However, the resulting filters
-are still not orthogonal across the mass parameter space, and are in fact
-highly redundant.  We use the singular value decomposition to produce an
-orthogonal filter set from the time sliced templates~\cite{Cannon:2010p10398}.
-We find that this reduces the number of sample points that need to be filtered
-by another factor of $\sim 100$.  The combined methods reduce the number of
-floating point operations to the level where they are competitive with the
-conventional matched filter approach.  In the remainder of this section we
-describe the \lloid\ algorithm in detail and provide some basic computational
-cost scaling.
+the early part of the waveform and allows the filters to be evaluated at about
+$\sim 100$ times lower rate.  This amounts to more than a factor of 10000
+reduction in the floating point operations per second required to filter the
+early part of the waveform.  However, the resulting filters are still not
+orthogonal across the mass parameter space, and are in fact highly redundant.
+We use the singular value decomposition to produce an orthogonal filter set
+from the time sliced templates~\cite{Cannon:2010p10398}.  We find that this
+reduces the number of sample points of the filters by another factor of $\sim
+100$.  The combined methods reduce the number of floating point operations to
+the level where they are competitive with the conventional high latency \FD\
+matched filter approach.  In the remainder of this section we describe the
+\lloid\ algorithm in detail and provide some basic computational cost scaling.
 
@@
 coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
-template is truncated at some prescribed time $t^0$, or equivalently frequency $f_\mathrm{hi}$.
-This is often chosen to correspond to the \ISCO. An inspiral signal
-will enter the detection band at a low frequency, $f = f_\mathrm{low}$,
+template is truncated at some prescribed time $t^0$, or equivalently frequency
+$f_\textrm{hi}$.  This is often chosen to correspond to the \ISCO. An inspiral
+signal will enter the detection band at a low frequency, $f = f_\mathrm{low}$,
 corresponding to a time $t_\mathrm{low}$.  The template is assumed to be zero
-outside the interval $[t_\mathrm{low}, t^0)$ and is said to have  a
-duration of $t^0 - t_\mathrm{low}$. It is critically sampled at a
-rate of $2 f_\mathrm{hi}$.
+outside the interval $[t_\mathrm{low}, t^0)$ and is said to have  a duration of
+$t^0 - t_\mathrm{low}$. It is critically sampled at a rate of $2
+f_\mathrm{hi}$.
 
-The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us to choose
-time-slice boundaries that require substantially less bandwidth at early times
-in the inspiral.  Our goal is to reduce the filtering cost of a large fraction
-of the waveform by computing part of the convolution at a lower sample
-rate.  Specifically we consider here time slice boundaries with the smallest
-power-of-two sample rate that sub-critically samples the time sliced
+The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us
+to choose time-slice boundaries that require substantially less bandwidth at
+early times in the inspiral.  Our goal is to reduce the filtering cost of a
+large fraction of the waveform by computing part of the convolution at a lower
+sample rate.  Specifically we consider here time slice boundaries with the
+smallest power-of-two sample rate that sub-critically samples the time sliced
 template.  The time slices consist of the $S$ intervals $\left(t^S,
-t^{S-1}\right],\, \dots,\, \left(t^2, t^1\right],\, \left(t^1, t^0\right]$ sampled at frequencies
-$f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0 \geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.
-The time-sliced templates may be downsampled without aliasing, so we define them as
+t^{S-1}\right],\, \dots,\, \left(t^2, t^1\right],\, \left(t^1, t^0\right]$
+sampled at frequencies $f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0
+\geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time
+sliced templates may be downsampled without aliasing, so we define them as
 %
@@
 %
-An example time-slice design satisfying these constraints for a $1.4 - 1.4 \, M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
+An example time-slice design satisfying these constraints for a $1.4 - 1.4 \,
+M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
 %
@@
 
-Since waveforms with neighboring intrinsic source parameters $\vec\theta$ have similar
-time-frequency evolution, it is possible to design computationally efficient time slices
-for an extended region of parameter space rather than to design different time slices for
-each template.
+Since waveforms with neighboring intrinsic source parameters $\vec\theta$ have
+similar time-frequency evolution, it is possible to design computationally
+efficient time slices for an extended region of parameter space rather than to
+design different time slices for each template.
 
-We note that
-the time slice decomposition in equation~(\ref{eq:time-slices}) is manifestly
-orthogonal since the time slices are disjoint in time.  In the next
-section we examine how to reduce the number of filters within each time slice via
-singular value decomposition of the time sliced templates.
+We note that the time slice decomposition in equation~(\ref{eq:time-slices}) is
+manifestly orthogonal since the time slices are disjoint in time.  In the next
+section we examine how to reduce the number of filters within each time slice
+via singular value decomposition of the time sliced templates.
 
@@
 It is possible to greatly reduce the number of filters required to achieve a
-particular minimum match by designing an appropriate set of orthonormal {\em
-basis templates}.  A numerical technique based on the
-singular value decomposition (\SVD) of inspiral template banks is demonstrated
-in~\cite{Cannon:2010p10398}.  Similarly, the time sliced templates described above
-can be approximated to arbitrary accuracy by expansion into a set of orthonormal
-\emph{basis templates}, $u_l^s[k]$, related to the original time sliced templates
-through the \emph{reconstruction matrix}, $v_{il}^s\sigma_l^s$:
+particular minimum match by designing an appropriate set of \SVD\ {\em
+basis templates}.  A numerical technique based on the singular value
+decomposition (\SVD) of inspiral template banks is demonstrated
+in~\cite{Cannon:2010p10398}.  Similarly, the time sliced templates described
+above can be approximated to arbitrary accuracy by expansion into a set of
+\SVD\ \emph{basis templates}, $u_l^s[k]$, related to the original time
+sliced templates through the \emph{reconstruction matrix},
+$v_{il}^s\sigma_l^s$:
 %
@@
 %
-The parameter $\numsvdtmps$ sets the number of basis templates that are kept in the
-approximation.  This determines the \SVD\ tolerance, which affects the \SNR\ loss due
-to the approximation.  The authors of \cite{Cannon:2010p10398} showed that high accuracy
-could be achieved with far fewer basis templates than templates in the original template
-bank.  We find that when combined with the time slice decomposition, the number of basis 
-templates \numsvdtmps\ is much smaller than the original number of templates \numtmps.
-In the next section we describe how we form our novel detection statistic using the time slice decomposition and the \SVD.
+The parameter $\numsvdtmps$ sets the number of basis templates that are kept in
+the approximation.  This determines the \SVD\ tolerance, which affects the
+\SNR\ loss due to the approximation.  The authors of \cite{Cannon:2010p10398}
+showed that high accuracy could be achieved with far fewer basis templates than
+templates in the original template bank.  We find that when combined with the
+time slice decomposition, the number of basis templates \numsvdtmps\ is much
+smaller than the original number of templates \numtmps.  In the next section we
+describe how we form our novel detection statistic using the time slice
+decomposition and the \SVD.
 
-\subsection{Proposed method}
+\subsection{Early warning \SNR }
 
-Definition of decimated detector data:
+In the previous two sections we have described two transformation that greatly
+reduce the burden of filtering the compact binary parameter space and make \TD\
+filtering of the data possible.  We are now prepared to define the early
+warning \SNR\ and to comment on the computational cost of evaluating it.  But
+first, we will introduce some notation refering to the decimation of data,
+which means reducing the sample rate after low pass filtering,
 \begin{equation}
@@
 \end{equation}
+and notation refering to the interpolation of data to increase the sample rate,
+\begin{equation}
+	x^{s}[k] = \left( H^\shortuparrow x^{s+1}\right)[k]
+\end{equation}
 
-\editorial{Chad: This whole subsection needs a lot of work.
-I just dropped our new notation in here.}%
-In order to minimize latency we propose using the time-domain convolution
-presented in \eqref{eq:SNRTD}.  However, because the brute-force evaluation of
-\eqref{eq:SNRTD} is far too costly to be useful, we will consider an
-approximation to \eqref{eq:SNRTD} that can reduce substantially the cost of
-\realtime\ filtering. This approximation has the form
+From the combination of transformations to the compact binary templates defined
+in equation \eqref{eq:time-sliced-templates} and \eqref{eq:svddecomp} we define
+the early warning filter output accumulated up to time slice $s$ as,
 %
@@
 %
-where $u_{jk}(\tau)$ is an orthogonal basis set of filters spanning the space
-of $\{x_i(\tau)\}$ and $\sigma_{jk} v_{ijk}$ is a tensor relating the filters
-$u_{jk}(\tau)$ to the original filter set $\{x_i(\tau)\}$.  We claim that with
-a suitable choice of filters $u_{jk}(\tau)$ one can reduce the computational
-cost of \eqref{eq:SNRTD} sufficiently to feasibly search for
-gravitational waves from compact binary coalescence in \realtime.  This
-requires 1) exploiting
-the time-frequency characteristics of the binary waveforms and 2) exploiting the redundancy of the template bank. We describe our
-procedure for producing the decomposition in \eqref{eq:decomp} in the remainder
-of this section.
+This quantity is related to the early warning \SNR\ by the normalization of
+the partial filter defined by the present time slice.  This normalization
+may be computed in the planning stage and can be easily multiplied here.
+Figure \ref{fig:pipeline} describes the computation of the \SNR\ and
+early warning \SNR\ schematically.
+%
+%
+\begin{figure}[htbp]
+	\includegraphics{figures/lloid-diagram.pdf}
+	\caption{\label{fig:pipeline} Schematic of LLOID pipeline illustrating
+signal flow.  Circles with arrows represent interpolation
+\protect\includegraphics{figures/upsample-symbol.pdf} or decimation
+\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
+signs represent summing junctions
+\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
+\protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.  Sample
+rate decreases from the top of the diagram to the bottom.  In this diagram each
+time slice contains three \fir\ filters that are linearly combined to produce
+four output channels.  In a typical pipeline the number of \fir\ filters is
+much less than the number of output channels.}
+\end{figure}
+%
+%
+The signal flow diagram drawn in figure \ref{fig:pipeline} contains several
+distinct stages and parallel branches.  The stages are decimation, filtering
+the data against the orthogonal templates, reconstruction of the original
+template basis, interpolation and accumulation of the early warning and final
+{\SNR}.
 
@@
 
+\subsubsection{Extrapolation of the computational cost of the \lloid\ method to an Advanced LIGO search}
+
+Table \ref{table:flops} shows that it requires about 0.5 GFLOPS to conduct the
+filtering, reconstruction and resampling stages of the \lloid\ method for 657
+templates chosen around typical NS--NS mass parameters.  Given the capability
+of modern CPUs it should be possible to filter that sub-bank on one CPU.
+Assuming that the cost holds for the other sub-banks we expect that the full
+parameter space should require $\sim 150$ CPUs for realtime filtering with the
+\lloid\ method per detector.  Even if real implementations fall short of these
+estimates by factors of several, there will still be ample computing resources
+for an early warning search in the advanced detector era.
+
 \begin{comment}

### 2011-06-03 introduction.tex
merged
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@909 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 low throughput.  The former are epitomized by the overlap-save algorithm for
-\fft{} convolution, currently the preferred method in gravitational wave
+\fft\ convolution, currently the preferred method in gravitational wave
 searches.  The most obvious example of the latter is the time domain
-(\textsc{td}) convolution, which has no latency at all.  However, its
+(\TD) convolution, which has no latency at all.  However, its
 computational complexity is quadratic in the length of the templates, so it is

### 2011-06-03 conclusions.tex
merge
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@912 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 improve the chances of detecting an electromagnetic counterpart in bands from
-X-ray down to radio.  We antipate requiring no more than \numcpus\ modern
+X-ray down to radio.  We anticipate requiring $\sim$\numcpus\ modern
 %
-\editorial{We need to do this calculation using the \flops\ counts and the number of templates.}%
+\editorial{We need to do this calculation using the \flops\ counts and the
+number of templates. CHAD: Agreed, but if you do it by the books it seems
+misleading.  I have adjusted it to 600.  This number is taken from assuming
+that 1 657 template sub-bank can be filtered on one core and dividing the total
+number of templates (10$^5$) by 657 which gives you 150 cores per detector. I
+put it in section 2. }%
 %
@@
 procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
-to a 1/(16~Hz) cadence introducing an up-front latency of 125~ms.  Data
-%
-\editorial{This is redundant: 1/(16~Hz) and 125~ms are just reciprocals of each other.}%
-%
+to a 1/(16~Hz) cadence introducing an up-front latency.  Data
 aggregation from the observatories will travel over various networks, each
@@
 \editorial{The way this is worded, it sounds like a big omission.}%
-We have omitted discussion of source sky localization. Localization is known to
-be poor for signals of low \SNR~\cite{Fairhurst2009}.
-However, one should not immediately dismiss the practical usefulness of a
-poorly localized source. Even with poor localization, it should be possible to
-begin downselecting what observatories could view a potential signal and for
-such observatories to begin any necessary prerequisite activities. In future
-works we wil explore more rigorously the pointing prospects with realistic
-simulations using the infrastructure and techniques described in this work.
+We have omitted discussion of source localization though point
+the reader to some theoretical estimates \SNR~\cite{Fairhurst2009}.
+%However, one should not immediately dismiss the practical usefulness of a
+%poorly localized source. Even with poor localization, it should be possible to
+%begin downselecting what observatories could view a potential signal and for
+%such observatories to begin any necessary prerequisite activities. 
+In future works we will explore more rigorously the pointing prospects with
+realistic simulations using the infrastructure and techniques described here.
 \editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring

### 2011-06-03 implementation.tex
merge
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@913 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The planning stage begins with choosing templates that cover the space of source
-parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to satisfy
-a minimum match criterion.  This assures a prescribed maximum loss in \SNR\
-for signals that fall between the chosen templates.  Typically the minimum
-match is 0.97 corresponding to a maximum mismatch of 0.03.  Next, the templates
-are subdivided into groups of neighbors called \emph{sub-banks} that are
-appropriately sized so that each bank can be efficiently handled by a single
-computer.  The neighbors are chosen to have comparable chirp mass, which produces
-sub-banks with similar time-frequency evolution.  Dividing the source parameter space
-into smaller sub-banks reduces the computational cost of the singular value decomposition
-and is the approach considered in~\cite{Cannon:2010p10398}.  We choose time
-slice boundaries as in equation~\eqref{eq:time-sliced-templates} such that all of the templates within a
-sub-bank are sub-critically sampled at progressively lower sample rates.  Next,
-the templates within the sub-bank are realized as \fir\ filter
+The planning stage begins with choosing templates that cover the space of
+source parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to
+satisfy a minimum match criterion.  This assures a prescribed maximum loss in
+\SNR\ for signals that fall between the chosen templates.  Typically the
+minimum match is 0.97 corresponding to a maximum mismatch of 0.03.  Next, the
+templates are subdivided into groups of neighbors called \emph{sub-banks} that
+are appropriately sized so that each bank can be efficiently handled by a
+single computer.  The neighbors are chosen to have comparable chirp mass, which
+produces sub-banks with similar time-frequency evolution.  Dividing the source
+parameter space into smaller sub-banks reduces the computational cost of the
+singular value decomposition and is the approach considered
+in~\cite{Cannon:2010p10398}.  We choose time slice boundaries as in
+equation~\eqref{eq:time-sliced-templates} such that all of the templates within
+a sub-bank are sub-critically sampled at progressively lower sample rates.
+Next, the templates within the sub-bank are realized as \fir\ filter
 coefficients.  For each time slice, the templates are down-sampled to the
-appropriate sample rate.  Finally, the \SVD\ is applied to each time
-slice in the sub-bank in order to produce a set of orthogonal \fir\
-filters and a reconstruction matrix that maps them back to the original
-templates as described in equation~\eqref{eq:svddecomp}.  The down-sampled orthogonal
-\fir\ filter coefficients, the reconstruction matrix, and the time slice
-boundaries are all saved to disk.
+appropriate sample rate.  Finally, the \SVD\ is applied to each time slice in
+the sub-bank in order to produce a set of orthogonal \fir\ filters and a
+reconstruction matrix that maps them back to the original templates as
+described in equation~\eqref{eq:svddecomp}.  The down-sampled orthogonal \fir\
+filter coefficients, the reconstruction matrix, and the time slice boundaries
+are all saved to disk.
 
@@
 \end{figure}
-
 The signal flow diagram drawn in figure \ref{fig:pipeline} contains several
@@
 
-\subsubsection{Orthogonal \fir\ filters}
+\subsubsection{\fir\ filters}
 
@@
 this way, the \lloid\ algorithm and this implementation leads to a simple early warning
-pipeline with little additional work.
+pipeline.

### 2011-06-03 implementation.tex
Removed duplicate figure
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@914 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 using \gstlal\ and \gstreamer\ components.
-%
-%
-\begin{figure}[htbp]
-	\includegraphics{figures/lloid-diagram.pdf}
-	\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
-signal flow.  Circles with arrows represent interpolation
-\protect\includegraphics{figures/upsample-symbol.pdf} or decimation
-\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
-signs represent summing junctions
-\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
-\protect\includegraphics{figures/fir-symbol.pdf} stand for \fir{} filters.  Sample
-rate decreases from the top of the diagram to the bottom.  In this diagram each
-time slice contains three \fir\ filters that are linearly combined to produce
-four output channels.  In a typical pipeline the number of \fir\ filters is
-much less than the number of output channels.}
-\end{figure}
-The signal flow diagram drawn in figure \ref{fig:pipeline} contains several
-distinct stages and parallel branches.  The stages are decimation, filtering
-the data against the orthogonal templates, reconstruction of the original template basis, interpolation
-and accumulation of the early-warning and final {\SNR}.  We outline the stages
-below.
 

### 2011-06-03 method.tex
Commented out figure 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@916 ccdd05d5-56c1-461a-bf98-85f6876523b4
remove sent
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@917 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
+\begin{comment}
 An example time-slice design satisfying these constraints for a $1.4 - 1.4 \,
@@
 \end{table}
+\end{comment}
 
@@
 %
-This quantity is related to the early warning \SNR\ by the normalization of
-the partial filter defined by the present time slice.  This normalization
-may be computed in the planning stage and can be easily multiplied here.
-Figure \ref{fig:pipeline} describes the computation of the \SNR\ and
-early warning \SNR\ schematically.
+\FIXME{This quantity is related to the early warning SNR by the normalization
+of the partial filter defined by the present time slice.}  Figure
+\ref{fig:pipeline} describes the computation of the \SNR\ and early warning
+\SNR\ schematically.
 %

### 2011-06-03 inspiral_svd.tex
Renamed analysis.tex to implementation.tex
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@893 ccdd05d5-56c1-461a-bf98-85f6876523b4
Keywords not needed
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@894 ccdd05d5-56c1-461a-bf98-85f6876523b4
Moved NSF fellowship to end of acknowledgements
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@900 ccdd05d5-56c1-461a-bf98-85f6876523b4
merged
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@911 ccdd05d5-56c1-461a-bf98-85f6876523b4
fixed we
git-svn-id: h

@@
 
-\title[LLOID]{Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy}
+\title[LLOID]{Technique for \earlywarning\ detection of compact binary
+coalescence and its implications for multi-messenger astronomy}
 
@@
 The rapid detection of compact binary coalescence with a network of advanced
-gravitational-wave detectors will offer a unique opportunity for multi-messenger astronomy.
-Prompt detection alerts to the astronomical community may make it possible to observe the onset of electromagnetic emission from compact binary coalescence.
-We demonstrate a computationally practical analysis strategy that produces early warning triggers even before gravitational radiation from the final merger has arrived at the detectors.
-With current rate estimates for the Advanced \LIGO{} design configuration, we should detect $\sim$10 sources earlier than 10 seconds before merger in 1 year of live time.
+gravitational-wave detectors will offer a unique opportunity for
+multi-messenger astronomy.  Prompt detection alerts to the astronomical
+community may make it possible to observe the onset of electromagnetic emission
+from compact binary coalescence.  We demonstrate a computationally practical
+analysis strategy that produces early warning triggers even before
+gravitational radiation from the final merger has arrived at the detectors.
+With current rate estimates for the Advanced \LIGO{} design configuration, it
+should be possible to detect $\sim$10 sources earlier than 10 seconds before
+merger in 1 year of live time.
 \end{abstract}
@@
 \pacs{04.30.-w, 07.05.Kf}
-
-%\keywords{Gravitational Waves, Laser Interferometry}
 
@@
 
-\input{analysis}
+\input{implementation}
 
@@
 Foundation and operates under cooperative agreement
-\textsc{phy}-\oldstylenums{0107417}.  This research is supported by the National Science Foundation through a Graduate Research Fellowship to LS.  CH would like to thank Ilya Mandel for
+\textsc{phy}-\oldstylenums{0107417}.  CH would like to thank Ilya Mandel for
 many discussions about rate estimates and the prospects of early detection.  CH
@@
 about low latency analysis methods.  NF would like to thank Alessandra Corsi
-for illuminating discussions on astronomical motivations.
+for illuminating discussions on astronomical motivations.  This research is supported by the National Science Foundation through a Graduate Research Fellowship to LS.
 

### 2011-06-03 method.tex
editorial about lloid
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@920 ccdd05d5-56c1-461a-bf98-85f6876523b4
Reworked last part of section 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@922 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\editorial{if we are gonna stick with lloid we should define it}
 \subsection{The \lloid\ method}
@@
 \FIXME{This quantity is related to the early warning SNR by the normalization
-of the partial filter defined by the present time slice.}  Figure
-\ref{fig:pipeline} describes the computation of the \SNR\ and early warning
-\SNR\ schematically.
+of the partial filter defined by the present time slice.}
 %
@@
 %
-The signal flow diagram drawn in figure \ref{fig:pipeline} contains several
-distinct stages and parallel branches.  The stages are decimation, filtering
-the data against the orthogonal templates, reconstruction of the original
-template basis, interpolation and accumulation of the early warning and final
-{\SNR}.
+The signal flow diagram in figure~\ref{fig:pipeline} illustrates how this
+recursion relation can be realized as a multirate filter network with outputs for
+each of the early warning \SNR{}s.
 

### 2011-06-03 implementation.tex
lloid_paper: cleanups including trimming references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@923 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 implement the \lloid\ algorithm.  For the near-term this should be a viable 
-solution for ultra-low latency searches with $\sim 1s$ intrinsic latency.
+solution for ultra-low latency searches with $\sim 1$s intrinsic latency.
 
@@
 computers to streaming media servers.  Given the similarities of
-gravitational- wave detector data to audio data it is not surprising that
+gravitational-wave detector data to audio data it is not surprising that
 \gstreamer\ is
@@
 \gstreamer\ framework by developing a library called \gstlal{}~\cite{gstlal}
-that provides elements for gravitational-wave data-analysis.  Figure
-\ref{fig:pipeline} describes schematically how we implement the \lloid\ algorithm
-using \gstlal\ and \gstreamer\ components.
+that provides elements for gravitational-wave data analysis.
+Figure~\ref{fig:pipeline} describes schematically how we implement the \lloid\
+algorithm using \gstlal\ and \gstreamer\ components.
 

### 2011-06-03 method.tex
lloid_paper: cleanups including trimming references
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@923 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR2 science runs.
-We combine this with the \SVD\ rank reduction method described
+We combine this with the \SVD\ rank-reduction method described
 in~\cite{Cannon:2010p10398} that exploits the redundancy of the template banks.
@@
 $M/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\,
-\vec\theta_{M/2-1}$ to assure a bounded loss of
-\SNR~\cite{Owen:1995tm,Owen:1998dk}\editorial{Can we cut this down to one
-citation?}. That is, any possible signal within a given range of intrinsic
+\vec\theta_{M/2-1}$ to assure a bounded loss of \SNR~\cite{Owen:1998dk}. That
+is, any possible signal within a given range of intrinsic
 source parameters will have an inner product that is $\geqslant 0.97$ with at
@@
 computational cost of a \TD\ search for compact binary coalescence.  We will
-give a truly zero latency algorithm that competes in terms of floating point
+give a truly zero-latency algorithm that competes in terms of floating point
 operations per second with the conventional \FD\ method, which requires a
@@
 slices.  Since each template slice is disjoint in time, the resulting set for a
-single template is orthogonal.  Given the chirp like structure of the
+single template is orthogonal.  Given the chirp-like structure of the
 templates, the early time slices have significantly lower bandwidth and can be
@@
 $\sim 100$ times lower rate.  This amounts to more than a factor of 10000
-reduction in the floating point operations per second required to filter the
+reduction in the floating-point operations per second required to filter the
 early part of the waveform.  However, the resulting filters are still not
@@
 We use the singular value decomposition to produce an orthogonal filter set
-from the time sliced templates~\cite{Cannon:2010p10398}.  We find that this
+from the time-sliced templates~\cite{Cannon:2010p10398}.  We find that this
 reduces the number of sample points of the filters by another factor of $\sim
 100$.  The combined methods reduce the number of floating point operations to
-the level where they are competitive with the conventional high latency \FD\
+the level where they are competitive with the conventional high-latency \FD\
 matched filter approach.  In the remainder of this section we describe the
@@
 \emph{time slices} in a time-domain analogue to the frequency-domain
-decomposition described in ~\cite{Marion2004, Buskulic2010, beauville2006,
-beauville2008}\editorial{Four citations is way too many!  We should pick just
-\emph{one} citation for \textsc{mbta}, and it should be the most recent and
-complete description of it.}.  A matched filter is constructed for each time
+decomposition described in ~\cite{Marion2004, Buskulic2010}.
+\editorial{nvf: I've picked the two most relevant; I think they are reasonably complementary.}%
+A matched filter is constructed for each time
 slice.  The outputs form an ensemble of partial \SNR{} streams.  By linearity,
-these partial \SNR\ streams can be suitably time delayed and summed to
+these partial \SNR\ streams can be suitably time-delayed and summed to
 reproduce the \SNR\ of the full template.  To wit,
@@
 templates in the original template bank.  We find that when combined with the
-time slice decomposition, the number of basis templates \numsvdtmps\ is much
+time-slice decomposition, the number of basis templates \numsvdtmps\ is much
 smaller than the original number of templates \numtmps.  In the next section we
-describe how we form our novel detection statistic using the time slice
+describe how we form our novel detection statistic using the time-slice
 decomposition and the \SVD.
@@
 
-In the previous two sections we have described two transformation that greatly
+In the previous two sections we have described two transformations that greatly
 reduce the burden of filtering the compact binary parameter space and make \TD\
@@
 first, we will introduce some notation refering to the decimation of data,
-which means reducing the sample rate after low pass filtering,
+which means reducing the sample rate after low-pass filtering,
 \begin{equation}
 \label{eq:decomp}
-	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k]
+	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k] \,,
 \end{equation}
@@
 \begin{equation}
-	x^{s}[k] = \left( H^\shortuparrow x^{s+1}\right)[k]
+	x^{s}[k] = \left( H^\shortuparrow x^{s+1}\right)[k] \,.
 \end{equation}
@@
 in equation \eqref{eq:time-sliced-templates} and \eqref{eq:svddecomp} we define
-the early warning filter output accumulated up to time slice $s$ as,
+the early-warning filter output accumulated up to time slice $s$ as,
 %
@@
 %
-\FIXME{This quantity is related to the early warning SNR by the normalization
+\FIXME{This quantity is related to the early warning \SNR{} by the normalization
 of the partial filter defined by the present time slice.}
@@
 
-Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total \lloid\ requires
+Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time-sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total \lloid\ requires
 $$

### 2011-06-03 conclusions.tex
lloid_paper: reworked astrophysics a bit in intro and conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@924 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \editorial{We should drive home the point that our method is as fast as the fft convolution but without any latency at all.}%
-We have demonstrated a computationally feasible procedure for the rapid and even advance
+We have demonstrated a computationally feasible procedure for the rapid and even preemptive
 detection of gravitational waves emitted during the coalescence of neutron
-stars and stellar-mass black holes.  These sources are expected to produce
-prompt electromagnetic signals and may be the progenitors of some short hard
-gamma-ray bursts.  Rapid alerts to the broader astronomical community may
+stars and stellar-mass black holes. Our method is as fast as standard \fft{}
+convolutions but without any latency at all. Our search targets are expected to
+produce prompt electromagnetic signals and may be the progenitors of some short
+hard gamma-ray bursts.  Rapid alerts to the broader astronomical community will
 improve the chances of detecting an electromagnetic counterpart in bands from
-X-ray down to radio.  We anticipate requiring $\sim$\numcpus\ modern
+gamma-rays down to radio.  We anticipate requiring $\sim$\numcpus\ modern
 %
@@
 computer cores to analyze a four-detector network of gravitational-wave data
-for binary neutron stars and stellar mass black holes.  This is within the
+for binary neutron stars and stellar-mass black holes.  This is well within the
 current computing capabilities of the \textsc{lsc} Data Grid~\cite{LDG}.
@@
 capable of high bandwith but perhaps only modest latency.  This could amount to
-a similar latency of $\sim$100 ms.  Lastly, unless a realtime infrastructure is
+a similar latency of $\sim$100 ms.  Lastly, unless a real-time infrastructure is
 adopted post data acquisition, it is likely that there will be an inherent
@@
 opinion, significant work would have to be done in order to improve upon this
-number. However, it should be considered for third-generation detector design.
-For example, a tighter integration of analysis and data acquisition would be
-beneficial.
+number. For example, a tighter integration of analysis and data acquisition
+would be beneficial. This should be considered for third-generation detector
+design.
 

### 2011-06-03 introduction.tex
lloid_paper: reworked astrophysics a bit in intro and conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@924 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 compact binary coalescence is believed to be a mechanism for short gamma-ray
-bursts (short GRBs)~\cite{Lee:2005, nakar07}. The same inner shocks, or
+bursts (short \GRB{}s)~\cite{Lee:2005, nakar07}. The same inner shocks, or
 potentially reverse shocks, can produce a bright accompanying optical flash~\cite{Sari99}. Prompt emission is a probe into the extreme
@@
 have only been observed for a handful of long
-GRBs~\cite{1999Natur.398..400A,2003Natur.422..284F,2006A&A...454L.119J,0004-637X-660-1-489,2008Natur.455..183R,2011A&A...528A..15G}
-by telescopes with extremely rapid response or, in the case of GRB 080319B, by
-pure serendipity, where several telescopes were observing a previous GRB in the
-same field~\cite{2008Natur.455..183R}. Short GRBs, on the other hand, typically
-fade too quickly to observe anything before the initial burst of gamma rays and hard
+\GRB{}s~\cite{2011CRPhy..12..255A}
+by telescopes with extremely rapid response or, in the case of \textsc{grb 080319b}, by
+pure serendipity, where several telescopes were observing a previous \GRB{} in the
+same field~\cite{2008Natur.455..183R}. Short \GRB{}s, on the other hand, typically
+fade too quickly to observe anything before the initial burst of gamma-rays and hard
 x-rays. Rapid gravitational-wave transient alerts could enable the observation
-of optical flashes from short GRBs. An optical counterpart would vastly boost
+of optical flashes from short \GRB{}s and perhaps even
+precursors~\cite{0004-637X-723-2-1711}. An electromagnetic counterpart would
+vastly boost
 confidence in the gravitational-wave detection and provide the tight sky
@@
 
-To this end, we have the ambition of reporting gravitational-wave candidates not minutes
-\emph{after} the merger, but seconds \emph{before}.  By
+To this end, we have the ambition of reporting gravitational-wave candidates
+not minutes \emph{after} the merger, as has already been accomplished, but
+seconds \emph{before}.  By
 looking for threshold crossings before the gravitational-wave signal leaves
@@
 Fortunately, the morphology of inspiral signals can be exploited to offset some
-of the computational complexity of low latency algorithms.  First, the signals
+of the computational complexity of low-latency algorithms.  First, the signals
 evolve slowly in frequency, so that they can be broken into contiguous

### 2011-06-03 method.tex
Added cpu extrapolation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@925 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsubsection{Extrapolation of the computational cost of the \lloid\ method to an Advanced LIGO search}
+\subsubsection{Extrapolation of the computational cost to an Advanced LIGO search}
 
-Table \ref{table:flops} shows that it requires about 0.5 GFLOPS to conduct the
-filtering, reconstruction and resampling stages of the \lloid\ method for 657
-templates chosen around typical NS--NS mass parameters.  Given the capability
-of modern CPUs it should be possible to filter that sub-bank on one CPU.
-Assuming that the cost holds for the other sub-banks we expect that the full
-parameter space should require $\sim 150$ CPUs for realtime filtering with the
-\lloid\ method per detector.  Even if real implementations fall short of these
-estimates by factors of several, there will still be ample computing resources
-for an early warning search in the advanced detector era.
+Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$ \flops\
+to cover a sub-bank of comprising $\sim 10^2$ out of $\sim 10^5$ sets of intrinsic
+source parameters.  Given that modern (ca. 2011) workstations can sustain
+computation rates up to $\sim 10^{10}$ \flops{}, and assuming that other regions of the
+parameter space have similar computational scaling, an entire search could be implemented
+with $\gtrsim 10$ machines.
+
+By comparison, using the \TD\ method to achieve the same latency costs $\sim 10^{13}$ \flops\ for this particular sub-bank, and so it would require $\sim 10^6$ present-day machines.
+Presently, the \LIGO\ data grid consists of only $\sim 10^4$ machines.
 

### 2011-06-03 inspiral_svd.tex
Removed author addresses
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@926 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\editorial{We have only Institution, City, State, Country for some co-authors.  Can we get rid of the full addresses, or is this at the discretion of the co-authors?  How about author e-mails?}
 \address{$^1$ Canadian Institute for Theoretical Astrophysics, Toronto, ON,
 Canada}
-\address{$^2$  D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, 61
-Avenue du Pr\'{e}sident Wilson, 94235 Cachan Cedex, France}
+\address{$^2$  D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
 \address{$^3$ LIGO Laboratory - California Institute of Technology, Pasadena,
 CA, USA} 
-\address{$^4$ Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, E-46100 Burjassot,
+\address{$^4$ Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot,
 Spain} 
@@
 \address{$^7$ Department of Physics and Astronomy, Barnard College, Columbia
-University, New York, NY 10027, USA} 
+University, New York, NY, USA} 
 \address{$^8$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r

### 2011-06-03 method.tex
Deleted funky indent
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@927 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed an error in section 2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@928 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 `$+$' and `$\times$' gravitational wave polarizations, so for any value of
-$\vec\theta$ we must implement 2 filters.  The coefficients for the $M$ filters
+$\vec\theta$ we must implement 2 filters.  The coefficients for the $\numtmps$ filters
 are known as templates, and are formed by discretizing and time reversing the
@@
 detector's noise. To construct a template bank, templates are chosen with the
-$M/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\,
-\vec\theta_{M/2-1}$ to assure a bounded loss of \SNR~\cite{Owen:1998dk}. That
+$\numtmps/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\,
+\vec\theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~\cite{Owen:1998dk}. That
 is, any possible signal within a given range of intrinsic
@@
 of 0.97. The data from the detector are whitened and convolved with each
-template to produce $2M$ \SNR\ time series. Local peak-finding across time and
+template to produce $\numtmps$ \SNR\ time series. Local peak-finding across time and
 templates determines detection candidates.

### 2011-06-03 conclusions.tex
changed future plan discussion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@929 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 improve the chances of detecting an electromagnetic counterpart in bands from
-gamma-rays down to radio.  We anticipate requiring $\sim$\numcpus\ modern
-%
-\editorial{We need to do this calculation using the \flops\ counts and the
-number of templates. CHAD: Agreed, but if you do it by the books it seems
-misleading.  I have adjusted it to 600.  This number is taken from assuming
-that 1 657 template sub-bank can be filtered on one core and dividing the total
-number of templates (10$^5$) by 657 which gives you 150 cores per detector. I
-put it in section 2. }%
-%
-computer cores to analyze a four-detector network of gravitational-wave data
+gamma-rays down to radio.  We anticipate requiring $\sim10\times4$ modern
+multi-core computers to analyze a four-detector network of gravitational-wave data
 for binary neutron stars and stellar-mass black holes.  This is well within the
@@
 
-\editorial{The way this is worded, it sounds like a big omission.}%
-We have omitted discussion of source localization though point
-the reader to some theoretical estimates \SNR~\cite{Fairhurst2009}.
-%However, one should not immediately dismiss the practical usefulness of a
-%poorly localized source. Even with poor localization, it should be possible to
-%begin downselecting what observatories could view a potential signal and for
-%such observatories to begin any necessary prerequisite activities. 
-In future works we will explore more rigorously the pointing prospects with
-realistic simulations using the infrastructure and techniques described here.
+We plan to explore more rigorously the prospects of early warning detection
+with realistic simulations of binary mergers using the infrastructure and
+techniques described here.  We will compare the performance of the early
+warning pipeline sky localization capabilities with the theoretical
+estimates~\cite{Fairhurst2009}.
+
 \editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring

### 2011-06-03 method.tex
Fleshed out caption
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@930 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\Table{\label{table:flops}Minimum computational cost of the \fir{} filter bank, \fft{} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
+\Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
+A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.}
 \br

### 2011-06-03 conclusions.tex
reworked end
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@931 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-We plan to explore more rigorously the prospects of early warning detection
-with realistic simulations of binary mergers using the infrastructure and
-techniques described here.  We will compare the performance of the early
-warning pipeline sky localization capabilities with the theoretical
-estimates~\cite{Fairhurst2009}.
+Although we have demonstrated a feasible method for advance detection we have
+not explored the accuracy of sky localization that is possible before merger.
+~\cite{Fairhurst2009} discusses some of the theoretical prospects for early sky
+localization.  Our future work will explore the prospects of early warning
+detection with realistic simulations of binary mergers using the infrastructure
+and techniques described here.
 
+\editorial{SHould we mention composite detection?}
 \editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring

### 2011-06-03 inspiral_svd.tex
fix dcc and abs
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@932 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 should be possible to detect $\sim$10 sources earlier than 10 seconds before
-merger in 1 year of live time.
+merger in 1 year of live time using the method we describe.
 \end{abstract}
@@
 about low latency analysis methods.  NF would like to thank Alessandra Corsi
-for illuminating discussions on astronomical motivations.  This research is supported by the National Science Foundation through a Graduate Research Fellowship to LS.
+for illuminating discussions on astronomical motivations.  This research is
+supported by the National Science Foundation through a Graduate Research
+Fellowship to LS.
 
 This paper has \LIGO\ Document
-Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{3}.}
+Number \LIGO-\textsc{p}\oldstylenums{0900004}.}
 

### 2011-06-04 gstlal_flop_budget.py
Removed some old scripts
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@933 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -1,310 +1,0 @@
-#!/usr/bin/env python
-"""Floating point operation count budgets for LLOID."""
-__author__ = "Leo Singer <leo.singer@ligo.org>"
-# FIXME: are we counting a multiply-accumulate as one operation or two?
-
-
-import math
-import sys
-
-
-#
-# Miscellaneous utility functions.
-#
-
-
-def lg(x):
-	"""Logarithm base 2."""
-	return math.log(x) / math.log(2.)
-
-
-#
-# Operation counts for convolution (or cross correlation) filters.
-#
-
-
-def td_conv_ops(kernel_len, count_channels=1):
-	"""Calculate the number of floating point operations per sample required to
-	apply a convolution filter with a kernel of kernel_len samples and
-	count_channels channels, using the FIR (time domain) implementation."""
-
-	# One multiply and one add per sample.
-	return 2 * kernel_len * count_channels
-
-
-def fd_conv_ops(kernel_len, block_len_factor=2, count_channels=1):
-	"""Calculate the number of floating point operations per sample required to
-	apply a convolution filter with a kernel of kernel_len samples and
-	count_channels channels, using the FFT (frequency domain) implementation and
-	a block size of (block_len) or (block_len_factor * kernel_len)."""
-
-	# 1 forward FFT and count_channels inverse FFTs
-	ops = 4 * lg(kernel_len * block_len_factor) * (count_channels + 1)
-
-	# count_channels dot products
-	ops += 2 * count_channels
-
-	# (1/block_len_factor) of the data gets reused from one block to the next
-	ops /= (1 - 1. / block_len_factor)
-
-	return ops
-
-
-#
-# Operation counts for miscellaneious signal processing operations.
-#
-
-
-def resample_filter_ops(kernel_len, in_rate, out_rate, count_channels=1):
-	"""Calculate the number of floating point operations per sample require to
-	apply an FIR resampling filter."""
-
-	# no-op if in_rate == out_rate
-	if in_rate == out_rate:
-		return 0
-
-	# FIR filter...
-	ops = td_conv_ops(kernel_len, count_channels)
-
-	# ... but (min(rates) / max(rates)) of the samples are zero.
-	ops *= float(min(in_rate, out_rate)) / max(in_rate, out_rate)
-
-	return ops
-
-
-#
[…]

### 2011-06-05 method.tex
Switched order of time slices in [ , ) intervals
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@937 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 template is truncated at some prescribed time $t^0$, or equivalently frequency
-$f_\textrm{hi}$.  This is often chosen to correspond to the \ISCO. An inspiral
+$f_\mathrm{hi}$.  This is often chosen to correspond to the \ISCO. An inspiral
 signal will enter the detection band at a low frequency, $f = f_\mathrm{low}$,
 corresponding to a time $t_\mathrm{low}$.  The template is assumed to be zero
-outside the interval $[t_\mathrm{low}, t^0)$ and is said to have  a duration of
-$t^0 - t_\mathrm{low}$. It is critically sampled at a rate of $2
+outside the interval $[t^0, t_\mathrm{low})$ and is said to have  a duration of
+$t_\mathrm{low} - t^0$. It is critically sampled at a rate of $2
 f_\mathrm{hi}$.
@@
 smallest power-of-two sample rate that sub-critically samples the time sliced
-template.  The time slices consist of the $S$ intervals $\left(t^S,
-t^{S-1}\right],\, \dots,\, \left(t^2, t^1\right],\, \left(t^1, t^0\right]$
-sampled at frequencies $f^\mathrm{S-1},\, \dots,\, f^1,\, f^0$ where $f^0
+template.  The time slices consist of the $S$ intervals
+$\left[t^0, t^1\right),\, \left[t^1, t^2\right),\, \dots,\, \left[t^{S-1}, t^S\right)$,
+sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^0
 \geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time

### 2011-06-05 method.tex
Added latency column to table 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@938 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
-A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.}
+A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.
+Both the \FD\ method and \lloid\ are five orders of magnitude faster than the conventional \TD\ method.
+However, the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no latency at all.}
 \br
-Method & \flops \\
+Method & \flops\ & latency (s) \\
 \mr
-\fir{} & $2.4\times10^{13}$ \\
-\fft{} convolution & $2.6\times10^8$ \\
-\lloid & $4.7\times10^8$ \\
+Time domain (\TD) convolution & $2.4\times10^{13}$ & 0 \\
+Frequency domain (\FD) convolution & $2.6\times10^8$ & 2201 \\
+\lloid\ (this work) & $4.7\times10^8$ & 0 \\
 \br

### 2011-06-05 introduction.tex
Cited Numerical Recipes earlier on for overlap-save
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@939 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 have high throughput at the cost of high latency, or low latency at the cost of
-low throughput.  The former are epitomized by the overlap-save algorithm for
-\fft\ convolution, currently the preferred method in gravitational wave
+low throughput.  The former are epitomized by the overlap-save algorithm
+\cite{numerical-recipes-chapter-13} for
+frequency domain (\FD) convolution, currently the preferred method in gravitational wave
 searches.  The most obvious example of the latter is the time domain

### 2011-06-05 inspiral_svd.tex
Improved short title
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@940 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\title[LLOID]{Technique for \earlywarning\ detection of compact binary
+\title[Early-warning detection of compact binary coalescence]{Technique for \earlywarning\ detection of compact binary
 coalescence and its implications for multi-messenger astronomy}

### 2011-06-05 introduction.tex
NS and BH now small caps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@941 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The coalescence of compact binary systems consisting of neutron stars (NS)
-and/or black holes (BH) is the most promising source of gravitational radiation
+The coalescence of compact binary systems consisting of neutron stars (\textsc{ns})
+and/or black holes (\textsc{bh}) is the most promising source of gravitational radiation
 for Advanced \LIGO, Virgo, \GEO\ and \LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb,
@@
 for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
-for NS--NS binaries in Advanced \LIGO\ assuming the event rate predictions
+for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming the event rate predictions
 in~\cite{Abadie:2010p10836}.
@@
 \includegraphics{figures/snr_in_time.pdf}
-\caption{\label{fig:earlywarning} Expected number of NS--NS sources that will
+\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns} sources that will
 be detectable $t$ seconds before coalescence.  The heavy solid line is the most

### 2011-06-06 inspiral_svd.tex
Fixing Mireia's name
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@942 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\author{Kipp Cannon$^{1}$, Romain Cariou$^{2}$, Adrian Chapman$^3$, M
-Crispin-Ortuzar$^4$, Nickolas Fotopoulos$^3$, Melissa Frei $^5$, Chad
+\author{Kipp Cannon$^{1}$, Romain Cariou$^{2}$, Adrian Chapman$^3$, Mireia
+Crisp\'{i}n-Ortuzar$^4$, Nickolas Fotopoulos$^3$, Melissa Frei $^5$, Chad
 Hanna$^{6}$, Erin Kara$^{7}$, Drew Keppel$^{8,9}$, Laura Liao$^{10}$, Stephen

### 2011-06-06 conclusions.tex
Minor fixes and change suggestions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@943 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 stars and stellar-mass black holes. Our method is as fast as standard \fft{}
-convolutions but without any latency at all. Our search targets are expected to
+convolutions but without any latency at all.
+%
+\editorial{Drew: ``but without any latency at all" is awkward and should be reworded.}
+%
+Our search targets are expected to
 produce prompt electromagnetic signals and may be the progenitors of some short
@@
 improve the chances of detecting an electromagnetic counterpart in bands from
-gamma-rays down to radio.  We anticipate requiring $\sim10\times4$ modern
+gamma-rays down to radio.  We anticipate requiring $\sim$$10\times4$ modern
 multi-core computers to analyze a four-detector network of gravitational-wave data
@@
 not explored the accuracy of sky localization that is possible before merger.
-~\cite{Fairhurst2009} discusses some of the theoretical prospects for early sky
+Ref.~\cite{Fairhurst2009} discusses some of the theoretical prospects for early sky
 localization.  Our future work will explore the prospects of early warning
@@
 
-\editorial{SHould we mention composite detection?}
+\editorial{Should we mention composite detection?}
+\editorial{Yes, although only as something else that we could employ to further reduce compuational cost.}
+
 \editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring

### 2011-06-06 implementation.tex
Switched order of time slices in [ , ) intervals
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@937 ccdd05d5-56c1-461a-bf98-85f6876523b4
Minor fixes and change suggestions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@943 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 in section \ref{sec:method} suitable for rapid gravitational-wave
-searches for compact binary coalescence.  The \lloid\ method requires several
+searches for \CBC{}s.  The \lloid\ method requires several
 computations that can be completed before the analysis is underway.  Thus
@@
 The \SVD{} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
-\item[]\begin{tabular}{rr@{,\,}lc*{6}{r}}
+\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
 \br
-\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 $-$ \textsc{svd} tolerance)} \\
+\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
 \ns
 \multicolumn{4}{c}{} &\crule{6} \\
-$f^s$ (Hz) & $(t^{s+1}$&$t^s]$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\
+$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
 \mr
-4096 & (0.5&0] & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
-512 & (4.5&0.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
-256 & (12.5&4.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
-128 & (76.5&12.5] & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
-64 & (140.5&76.5] & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
-64 & (268.5&140.5] & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
-64 & (396.5&268.5] & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
-32 & (460.5&396.5] & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
-32 & (588.5&460.5] & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
-32 & (844.5&588.5] & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
-32 & (1100.5&844.5] & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
+4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
+512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
+256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
+128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
+64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
+64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
+64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
+32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
+32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
+32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
+32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
 \br
@@
 offset to the filter output later in the pipeline makes up for the lack of
-explicit zero-padding.  The orthogonal filter outputs must next be reconstructed
+explicit zero-padding.
+%
+\editorial{Drew: This final sentence should be removed. It belongs to the next section.}
+%
+The orthogonal filter outputs must next be reconstructed
 into the output of the underlying time-sliced templates.
@@
 The early-warning \SNR\ for each time slice is formed by accumulating interpolated early
-warning \SNR\ from the subsequent time slice.  This process continuous until we have 
+warning \SNR\ from the subsequent time slice.  This process continues until we have 
 worked our way to the \SNR\ of the original templates at the full sample rate $f^0$.  In 

### 2011-06-06 introduction.tex
Minor fixes and change suggestions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@943 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 not minutes \emph{after} the merger, as has already been accomplished, but
-seconds \emph{before}.  By
-looking for threshold crossings before the gravitational-wave signal leaves
-the detection band, it is possible to trade some signal to noise ratio (\SNR{})
-for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
-for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming the event rate predictions
-in~\cite{Abadie:2010p10836}.
+seconds \emph{before}.  By looking for threshold crossings before the
+gravitational-wave signal leaves the detection band, it is possible to trade
+some signal to noise ratio (\SNR{}) and source localization
+accuracy~\cite{Fairhurst2009} for latency.  Figure \ref{fig:earlywarning} shows
+projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
+\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.
 %
@@
 the best gravitational-wave trigger generation latencies of 2--5 minutes.  We
-assume that in the advanced detector era the vetting process will be automated,
+assume that in the Advanced detector era the vetting process will be automated,
 so current gravitational wave search methodology and telescope actuation would
@@
 \cite{numerical-recipes-chapter-13} for
-frequency domain (\FD) convolution, currently the preferred method in gravitational wave
+frequency domain (\FD) convolution, currently the preferred method in gravitational-wave
 searches.  The most obvious example of the latter is the time domain

### 2011-06-06 method.tex
Minor fixes and change suggestions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@943 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In this section we describe a decomposition of the compact binary parameter
+In this section we describe a decomposition of the \CBC{} waveform
 space that reduces low-latency filtering cost sufficiently to allow for the
@@
 the detector is a linear combination of two waveforms corresponding to the
-`$+$' and `$\times$' gravitational wave polarizations, so for any value of
+`$+$' and `$\times$' gravitational wave polarizations.  Thus, for any value of
 $\vec\theta$ we must implement 2 filters.  The coefficients for the $\numtmps$ filters
@@
 convolution in the frequency domain (\FD), requiring only $\mathcal O(\numtmps
-\lg \tmpsamps)$ operations per sample.
+\lg \tmpsamps)$ operations per sample with increased latency.
 
@@
 entire parameter space is dramatic.
+%
+\editorial{Drew: is ``sample count" the correct wording?}
 
@@
 templates, the early time slices have significantly lower bandwidth and can be
-safely downsampled.  The downsampling removes a factor of $\sim 100$ samples
+safely downsampled.  The downsampling removes a factor of $\sim$100 samples
 from
@@
 the early part of the waveform and allows the filters to be evaluated at about
-$\sim 100$ times lower rate.  This amounts to more than a factor of 10000
+$\sim$100 times lower rate.  This amounts to more than a factor of 10000
 reduction in the floating-point operations per second required to filter the
@@
 from the time-sliced templates~\cite{Cannon:2010p10398}.  We find that this
-reduces the number of sample points of the filters by another factor of $\sim
-100$.  The combined methods reduce the number of floating point operations to
+reduces the number of sample points of the filters by another factor of $\sim$100.
+The combined methods reduce the number of floating point operations to
 the level where they are competitive with the conventional high-latency \FD\
-matched filter approach.  In the remainder of this section we describe the
+matched filter approach.  In the remainder of this section 
+%
+\editorial{Drew: We used this wording at the beginning of this paragraph.}
+%
+we describe the
 \lloid\ algorithm in detail and provide some basic computational cost scaling.
@@
 The first step of our proposed method is to divide the templates into
-\emph{time slices} in a time-domain analogue to the frequency-domain
-decomposition described in ~\cite{Marion2004, Buskulic2010}.
+\emph{time slices} 
+%
+\editorial{Drew: This was introduced in the previous section. It should be italicized there.}
+%
+in a time-domain analogue to the frequency-domain
+decomposition described in~\cite{Marion2004, Buskulic2010}.
 \editorial{nvf: I've picked the two most relevant; I think they are reasonably complementary.}%
@@
 \end{equation}
+\editorial{Drew: Should this be $h_{i}[k]$ or $\rho_{i}[k]$?}
 %
@@
 %
-The parameter $\numsvdtmps$ sets the number of basis templates that are kept in
+The parameter $\numsvdtmps$ is the number of basis templates that are kept in
 the approximation.  This determines the \SVD\ tolerance, which affects the
@@
 In the previous two sections we have described two transformations that greatly
-reduce the burden of filtering the compact binary parameter space and make \TD\
+reduce the burden of filtering waveforms from the the compact binary parameter space and make \TD\
 filtering of the data possible.  We are now prepared to define the early
@@
 \end{table}
+\editorial{Drew: Is \slicessamps before or after decimation? We should clarify that here.}
 
@@
 \end{table}
+\editorial{Drew:``$1.955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$" can't be true. Should the signs be greater than signs?}
 
@@
 \subsubsection{Extrapolation of the computational cost to an Advanced LIGO search}
+\editorial{Drew: I think this section be the last section of the results section. Should we move it to be Sec. 4.2?}
 

### 2011-06-06 results.tex
Minor fixes and change suggestions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@943 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 template bank which is typically.  We will consider an acceptable
-\SNR\ loss to be a factor of 10 smaller than this, i.e., no more than $\sim 0.003$.
+\SNR\ loss to be a factor of 10 smaller than this, i.e., no more than $\sim$0.003.
 
@@
 
-In this section we check the effect of the \SVD\ tolerance parameter defined in
-~\cite{Cannon:2010p10398}.  By changing the tolerance of the reconstruction of
+In this section we check the effect of the \SVD\ tolerance parameter defined
+in~\cite{Cannon:2010p10398}.  By changing the tolerance of the reconstruction of
 the physical waveforms one effects directly the mismatch that results from the
@@
 This could be the effect of the resampling, or another \SNR\ loss that we did
-not model or expect.  However the saturation is still an order of magnitude below our target mismatch of $\sim 0.003$.  We find that an \SVD\ tolerance of
+not model or expect.  However the saturation is still an order of magnitude below our target mismatch of $\sim$0.003.  We find that an \SVD\ tolerance of
 $\left(1-10^{-4}\right)$ is adequate. 

### 2011-06-06 conclusions.tex
lloid_paper: addressed comments from Steve and Missy; significant introduction reordering
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@944 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\editorial{We should drive home the point that our method is as fast as the fft convolution but without any latency at all.}%
 We have demonstrated a computationally feasible procedure for the rapid and even preemptive
@@
 stars and stellar-mass black holes. Our method is as fast as standard \fft{}
-convolutions but without any latency at all.
-%
-\editorial{Drew: ``but without any latency at all" is awkward and should be reworded.}
-%
+convolutions but allows for sample-in-sample-out processing.
 Our search targets are expected to
@@
 for binary neutron stars and stellar-mass black holes.  This is well within the
-current computing capabilities of the \textsc{lsc} Data Grid~\cite{LDG}.
+current computing capabilities of the \LIGO{} Data Grid~\cite{LDG}.
 

### 2011-06-06 inspiral_svd.tex
lloid_paper: addressed comments from Steve and Missy; significant introduction reordering
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@944 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 from compact binary coalescence.  We demonstrate a computationally practical
-analysis strategy that produces early warning triggers even before
+analysis strategy that produces early-warning triggers even before
 gravitational radiation from the final merger has arrived at the detectors.
 With current rate estimates for the Advanced \LIGO{} design configuration, it
-should be possible to detect $\sim$10 sources earlier than 10 seconds before
-merger in 1 year of live time using the method we describe.
+should be possible to detect $\sim$10 such sources earlier than 10 seconds
+before merger in 1 year of live time using the method we describe.
 \end{abstract}

### 2011-06-06 introduction.tex
lloid_paper: addressed comments from Steve and Missy; significant introduction reordering
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@944 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-05 introduction.tex
@@
 
-\editorial{Would the introduction be more effective without this paragraph?, should we find whitepapers?}
-
-The coalescence of compact binary systems consisting of neutron stars (\textsc{ns})
-and/or black holes (\textsc{bh}) is the most promising source of gravitational radiation
-for Advanced \LIGO, Virgo, \GEO\ and \LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb,
-LCGTWeb}.  Tens of binary coalescence events are expected to be observed in the
-advanced detector era later this decade~\cite{Abadie:2010p10836}.
-
 As a compact binary system loses energy to gravitational waves, its orbital
-separation decreases. This causes a run-away inspiral with the
+separation decays, leading to a run-away inspiral with the
 gravitational-wave amplitude and frequency increasing until the system
-eventually merges near the innermost stable circular orbit (\ISCO).  If a
+eventually merges soon after the innermost stable circular orbit (\ISCO). If a
 neutron star is involved it may become tidally disrupted near the merger.  This
-disrupted matter can fuel a bright electromagnetic counterpart in the system's
-final moments as a binary~\cite{shibata:2007}.
+disrupted matter can fuel a bright electromagnetic counterpart in the merger's aftermath~\cite{shibata:2007}.
+The coalescence of compact binary systems consisting of neutron stars
+(\textsc{ns}) and/or black holes (\textsc{bh}) is the most promising source of
+gravitational radiation for Advanced \LIGO, Virgo, \GEO\, and
+\LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb, LCGTWeb}.  Tens of binary coalescence
+events are expected to be observed in the advanced detector era later this
+decade~\cite{Abadie:2010p10836}. In the era of first-generation
+detectors the gravitational-wave community
+initiated a project to send alerts when potential gravitational-wave transients
+were observed in order to trigger followup observations by optical telescopes.
+The best demonstrated latencies were 30--60 minutes~\cite{LOOCUP}, which was
+an important achievement, but far too late to catch prompt electromagnetic emission.
 
@@
 bursts (short \GRB{}s)~\cite{Lee:2005, nakar07}. The same inner shocks, or
-potentially reverse shocks, can produce a bright accompanying optical flash~\cite{Sari99}. Prompt emission is a probe into the extreme
-initial conditions of the outflow, in contrast with afterglows, which are
-relatively insensitive to initial conditions. Optical flashes
-have only been observed for a handful of long
-\GRB{}s~\cite{2011CRPhy..12..255A}
-by telescopes with extremely rapid response or, in the case of \textsc{grb 080319b}, by
-pure serendipity, where several telescopes were observing a previous \GRB{} in the
-same field~\cite{2008Natur.455..183R}. Short \GRB{}s, on the other hand, typically
-fade too quickly to observe anything before the initial burst of gamma-rays and hard
-x-rays. Rapid gravitational-wave transient alerts could enable the observation
-of optical flashes from short \GRB{}s and perhaps even
-precursors~\cite{0004-637X-723-2-1711}. An electromagnetic counterpart would
-vastly boost
-confidence in the gravitational-wave detection and provide the tight sky
-localization necessary to allow determination of the source's host galaxy,
-which leads to a redshift measurement. With both redshift and a coincident
-gravitational-wave observation, we can produce precision measurements of the
-Hubble constant~\cite{2010ApJ...725..496N}.
+potentially reverse shocks, can produce a bright accompanying optical
+flash~\cite{Sari99}. Prompt emission is a probe into the extreme initial
+conditions of the outflow, in contrast with afterglows, which arise in the
+external shock with the local medium and are relatively
+insensitive to initial conditions. Optical flashes have only been observed to
+peak within tens of seconds for a handful of long
+\GRB{}s~\cite{2011CRPhy..12..255A} by telescopes with extremely rapid response
+or, in the case of \textsc{grb 080319b}, by pure serendipity, where several
+telescopes were observing a previous \GRB{} in the same
+field~\cite{2008Natur.455..183R}. Short \GRB{}s, on the other hand, typically
+fade too quickly to even catch the tails of the afterglows. Rapid
+gravitational-wave transient alerts could enable the first observation of
+prompt optical flashes and the rise of afterglows from short \GRB{}s, and, for
+an exceptional event, perhaps even a glimpse of a
+precursor~\cite{0004-637X-723-2-1711}. An electromagnetic counterpart would
+vastly boost confidence in the gravitational-wave detection and provide the
+tight sky localization necessary to allow determination of the source's host
+galaxy, which leads to a redshift measurement. A coincident gravitational-wave
+observation will provide a luminosity distance measurement. With these
+together, we can produce precision measurements of the Hubble
+constant~\cite{2010ApJ...725..496N}.
 
@@
 not minutes \emph{after} the merger, as has already been accomplished, but
-seconds \emph{before}.  By
-looking for threshold crossings before the gravitational-wave signal leaves
-the detection band, it is possible to trade some signal to noise ratio (\SNR{})
+seconds \emph{before}.  By examining the signal to noise ratio (\SNR{}) stream
+for threshold crossings before the gravitational-wave signal leaves the
+detection band, it is possible to trade some \SNR{}
 for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
@@
 
-\editorial{Citation needed for LOOC-UP} The gravitational-wave community
-initiated a project to send alerts when potential gravitational-wave transients
-are observed.  In October 2010, \LIGO{} completed its sixth science run
+In October 2010, \LIGO{} completed its sixth science run
 (S6) and Virgo completed its third science run (VSR3).  While both
 \LIGO{} detectors and Virgo were operating, several all-sky detection
-pipelines operated in a low-latency configuration, namely \textsc{mbta}, ihope,
+pipelines operated in a low-latency configuration to send astronomical alerts,
+namely \textsc{mbta}, ihope,
 Coherent WaveBurst, and Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
@@
 HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved
-the best gravitational-wave trigger generation latencies of 2--5 minutes.  We
+the best gravitational-wave trigger-generation latencies of 2--5 minutes.  We
 assume that in the advanced detector era the vetting process will be automated,
-so current gravitational wave search methodology and telescope actuation would
+so current gravitational-wave search methodology and telescope actuation would
 dominate latency.
@@
 low throughput.  The former are epitomized by the overlap-save algorithm
-\cite{numerical-recipes-chapter-13} for
-frequency domain (\FD) convolution, currently the preferred method in gravitational wave
+\cite{numerical-recipes-chapter-13} for frequency domain (\FD) convolution,
+currently the preferred method in gravitational-wave
 searches.  The most obvious example of the latter is the time domain
-(\TD) convolution, which has no latency at all.  However, its
+(\TD) convolution, which has no algorithmic latency.  However, its
 computational complexity is quadratic in the length of the templates, so it is
@@
 To validate the approach we present results of simulations and conclude with
-some remarks on what remains to prepare for the Advanced detector era.
+some remarks on what remains to prepare for the advanced detector era.

### 2011-06-06 method.tex
lloid_paper: addressed comments from Steve and Missy; significant introduction reordering
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@944 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 multiband decomposition of the compact binary parameter space that resulted in
-a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR2 science runs.
-We combine this with the \SVD\ rank-reduction method described
+a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 joint science
+runs.  We combine this with the \SVD\ rank-reduction method described
 in~\cite{Cannon:2010p10398} that exploits the redundancy of the template banks.
@@
 
-\subsubsection{Extrapolation of the computational cost to an Advanced LIGO search}
-\editorial{Drew: I think this section be the last section of the results section. Should we move it to be Sec. 4.2?}
+\subsubsection{Extrapolation of the computational cost to an Advanced \LIGO{} search}
+\editorial{Drew: I think this section be the last section of the results section.
+Should we move it to be Sec.
+4.2?
+Nick: I'm inclined to agree with Drew, though I see that we want to make explicit that it's an analysis of the algorithm and not of the implementation and the current placement aids that.
+Also, the given FLOPS numbers correspond to what choice of \SVD{} tolerance and $N^\shortuparrow$?}
 
@@
 By comparison, using the \TD\ method to achieve the same latency costs $\sim 10^{13}$ \flops\ for this particular sub-bank, and so it would require $\sim 10^6$ present-day machines.
-Presently, the \LIGO\ data grid consists of only $\sim 10^4$ machines.
+Presently, the \LIGO{} Data Grid consists of only $\sim 10^4$ machines.
 

### 2011-06-06 implementation.tex
Addressed a bunch of Drew's margin comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@945 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The \fir\ filter banks are implemented using a \gstlal\ element called {\tt
+The \fir\ filtering is implemented using a \gstlal\ element called {\tt
 lal\_firbank}, which produces $N$ channels of filter output from an $N\times M$ matrix of
-\fir\ filter coefficients.  This element is used in parallel branches in the
+\fir\ filter coefficients, representing the \SVD{} basis filters in our application.  This element is used in parallel branches in the
 pipeline to implement the \SVD\ basis filters in each time
@@
 explicit zero-padding.
-%
-\editorial{Drew: This final sentence should be removed. It belongs to the next section.}
-%
-The orthogonal filter outputs must next be reconstructed
-into the output of the underlying time-sliced templates.
 

### 2011-06-06 method.tex
Addressed a bunch of Drew's margin comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@945 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In the remainder of this section we explore a method for reducing the
+Here we explore a method for reducing the
 computational cost of a \TD\ search for compact binary coalescence.  We will
@@
 orthogonal filters with far fewer coefficients than the original templates.
-The reduction in sample count for all of the filters required to search the
-entire parameter space is dramatic.
-%
-\editorial{Drew: is ``sample count" the correct wording?}
 
-The first transformation is to chop the time-domain templates up into time
-slices.  Since each template slice is disjoint in time, the resulting set for a
+The first transformation is to chop the time-domain templates up into \emph{time
+slices}.  Since each template slice is disjoint in time, the resulting set for a
 single template is orthogonal.  Given the chirp-like structure of the
@@
 the level where they are competitive with the conventional high-latency \FD\
-matched filter approach.  In the remainder of this section 
-%
-\editorial{Drew: We used this wording at the beginning of this paragraph.}
-%
-we describe the
+matched filter approach.  In the remainder of this section we describe the
 \lloid\ algorithm in detail and provide some basic computational cost scaling.
@@
 The first step of our proposed method is to divide the templates into
-\emph{time slices} 
-%
-\editorial{Drew: This was introduced in the previous section. It should be italicized there.}
-%
-in a time-domain analogue to the frequency-domain
+time slices in a time-domain analogue to the frequency-domain
 decomposition described in~\cite{Marion2004, Buskulic2010}.
-\editorial{nvf: I've picked the two most relevant; I think they are reasonably complementary.}%
 A matched filter is constructed for each time
@@
 \numsvdtmps	& number of orthogonal templates in time slice $s$ \\
-\slicessamps	& number of samples in time slice $s$ \\
+\slicessamps	& number of samples in decimated time slice $s$\\
 $f^s$		& sample rate in time slice $s$ \\
@@
 \end{table}
-\editorial{Drew: Is \slicessamps before or after decimation? We should clarify that here.}
 
@@
 
-\Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
+\Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
 A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.
@@
 \end{table}
-\editorial{Drew:``$1.955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$" can't be true. Should the signs be greater than signs?}
 

### 2011-06-06 implementation.tex
Replace long words with macro acronyms. We're back to 15 pages.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@946 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 parameter space into smaller sub-banks reduces the computational cost of the
-singular value decomposition and is the approach considered
-in~\cite{Cannon:2010p10398}.  We choose time slice boundaries as in
+\SVD{} and is the approach considered
+in~\cite{Cannon:2010p10398}.  We choose time-slice boundaries as in
 equation~\eqref{eq:time-sliced-templates} such that all of the templates within
@@
 described in equation~\eqref{eq:svddecomp}.  The down-sampled orthogonal \fir\
-filter coefficients, the reconstruction matrix, and the time slice boundaries
+filter coefficients, the reconstruction matrix, and the time-slice boundaries
 are all saved to disk.

### 2011-06-06 method.tex
Replace long words with macro acronyms. We're back to 15 pages.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@946 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the
-signal-to-noise ratio, $\rho_i[k]$ defined as
+\SNR{}, $\rho_i[k]$ defined as
 %
@@
 
-Equation~(\ref{eq:SNRTD}) can be implemented in the time domain (\TD) as an
+Equation~(\ref{eq:SNRTD}) can be implemented in \TD as an
 \fir\ filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point
@@
 efficient to use the convolution theorem and the \fft\ to implement fast
-convolution in the frequency domain (\FD), requiring only $\mathcal O(\numtmps
+convolution in the \FD, requiring only $\mathcal O(\numtmps
 \lg \tmpsamps)$ operations per sample with increased latency.
@@
 
-The first transformation is to chop the time-domain templates up into \emph{time
+The first transformation is to chop the \TD{} templates up into \emph{time
 slices}.  Since each template slice is disjoint in time, the resulting set for a
@@
 The first step of our proposed method is to divide the templates into
-time slices in a time-domain analogue to the frequency-domain
+time slices in a \TD{} analogue to the \FD{}
 decomposition described in~\cite{Marion2004, Buskulic2010}.
@@
 in the next section that this, combined with the singular value docomposition,
-is sufficient to enable a computationally efficient time-domain search and
+is sufficient to enable a computationally efficient \TD{} search and
 furthermore is an essential part of an \earlywarning\ detection scheme.
@@
 In the next section we compute the expected computational cost scaling of this
-decomposition and compare it with the brute-force time-domain implementation of
-\eqref{eq:SNRTD} and higher latency frequency-domain methods.
+decomposition and compare it with the brute-force \TD{} implementation of
+\eqref{eq:SNRTD} and higher latency \FD{} methods.
 
@@
 We now examine the computational cost scaling of the approximate implementation
-of a conventional time-domain or frequency-domain matched filter bank as compared
+of a conventional \TD{} or \FD{} matched filter bank as compared
 with \lloid{}.  For convenience, table~\ref{tab:recap} provides a review of the notation
@@
 \mr
-Time domain (\TD) convolution & $2.4\times10^{13}$ & 0 \\
-Frequency domain (\FD) convolution & $2.6\times10^8$ & 2201 \\
+Time-domain (\TD) convolution & $2.4\times10^{13}$ & 0 \\
+Frequency-domain (\FD) convolution & $2.6\times10^8$ & 2201 \\
 \lloid\ (this work) & $4.7\times10^8$ & 0 \\
@@
 
-\subsubsection{Conventional time-domain method}
+\subsubsection{Conventional \TD{} method}
 
-The conventional time-domain method consists of a bank of \fir{} filters, or sliding-window dot products.
+The conventional \TD{} method consists of a bank of \fir{} filters, or sliding-window dot products.
 If there are $\numtmps$ templates, each $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications and additions per sample, or $2 \numtmps \tmpsamps f^0$ floating point operations per second (\flops) at a sample rate $f^0$.
 
-\subsubsection{Conventional frequency-domain method}
+\subsubsection{Conventional \FD{} method}
 
-The most common frequency-domain method is known as the \emph{overlap-save} algorithm, described in \cite{numerical-recipes-chapter-13}.
+The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in \cite{numerical-recipes-chapter-13}.
 It entails splitting the input into blocks of $D$ samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$ samples.
@@
 $$
-f^0 \cdot \frac{2 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock} \; \mathrm{\flops}.
+f^0 \cdot \frac{2 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock} \; \mathrm{\flops} \,.
 $$
@@
 The computational cost of the decimation of the detector data is a little bit more subtle.
-Decimation is achieved by applying an \fir\ antialiasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
-Naively, an antialiasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.
-However, it is necessary to evaluate the antialiasing filter only for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.
+Decimation is achieved by applying an \fir\ anti-aliasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
+Naively, an anti-aliasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.
+However, it is necessary to evaluate the anti-aliasing filter only for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.
 Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 N^\shortdownarrow f^s$ \flops\ is possible.
@@
 $$
-\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right) f \textrm{ \flops.}
+\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right) f \textrm{ \flops}\,.
 $$

### 2011-06-07 method.tex
some rearrangment
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@948 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In this section we describe a decomposition of the \CBC{} waveform
-space that reduces low-latency filtering cost sufficiently to allow for the
+space that reduces time-domain (\TD) filtering cost sufficiently to allow for the
 possibility of \earlywarning\ detection with modest computing requirements.  We
@@
 
-Equation~(\ref{eq:SNRTD}) can be implemented in \TD as an
+Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as an
 \fir\ filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point
@@
 efficient to use the convolution theorem and the \fft\ to implement fast
-convolution in the \FD, requiring only $\mathcal O(\numtmps
+convolution in the frequency-domain, requiring only $\mathcal O(\numtmps
 \lg \tmpsamps)$ operations per sample with increased latency.
 
-\editorial{if we are gonna stick with lloid we should define it}
 \subsection{The \lloid\ method}
 
-Here we explore a method for reducing the
-computational cost of a \TD\ search for compact binary coalescence.  We will
-give a truly zero-latency algorithm that competes in terms of floating point
-operations per second with the conventional \FD\ method, which requires a
-signficant latency in order to be computationally cheap. Our method, \lloid{},
+Here we explore a method for reducing the computational cost of a \TD\ search
+for compact binary coalescence.  We will give a truly zero-latency algorithm
+that competes in terms of floating point operations per second with the
+conventional \FD\ method, which requires a signficant latency in order to be
+computationally cheap. Our method, \lloid\footnote{\lloid, The term strictly
+means \emph{Low Latency Online Inspiral Detection} and was a code name for
+this procedure as it was being developed. It is now known in some GW circles.
+We feel that although the name doesn't carry much specific information about
+the details of the algorithm it is far too late to do anything about it now.}
 involves two transformations of the template waveforms that produce a set of
@@
 The first transformation is to chop the \TD{} templates up into \emph{time
-slices}.  Since each template slice is disjoint in time, the resulting set for a
-single template is orthogonal.  Given the chirp-like structure of the
+slices}.  Since each template slice is disjoint in time, the resulting set for
+a single template is orthogonal.  Given the chirp-like structure of the
 templates, the early time slices have significantly lower bandwidth and can be
 safely downsampled.  The downsampling removes a factor of $\sim$100 samples
-from
-%
-\editorial{Where do these numbers come from? CHAD: I have actually changed the statement to 100 for the early time slices.}
-%
-the early part of the waveform and allows the filters to be evaluated at about
-$\sim$100 times lower rate.  This amounts to more than a factor of 10000
+from the early part of the waveform and allows the filters to be evaluated at
+about $\sim$100 times lower rate.  This amounts to more than a factor of 10000
 reduction in the floating-point operations per second required to filter the
@@
 from the time-sliced templates~\cite{Cannon:2010p10398}.  We find that this
-reduces the number of sample points of the filters by another factor of $\sim$100.
-The combined methods reduce the number of floating point operations to
-the level where they are competitive with the conventional high-latency \FD\
+reduces the number of sample points of the filters by another factor of
+$\sim$100.  The combined methods reduce the number of floating point operations
+to the level where they are competitive with the conventional high-latency \FD\
 matched filter approach.  In the remainder of this section we describe the
@@
 time-slice decomposition, the number of basis templates \numsvdtmps\ is much
-smaller than the original number of templates \numtmps.  In the next section we
-describe how we form our novel detection statistic using the time-slice
+smaller than the original number of templates \numtmps\ and improves on
+\cite{Cannon:2010p10398} by nearly an order of magnitude.  In the next section
+we describe how we form our novel detection statistic using the time-slice
 decomposition and the \SVD.
@@
 
-\Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
-A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.
-Both the \FD\ method and \lloid\ are five orders of magnitude faster than the conventional \TD\ method.
-However, the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no latency at all.}
-\br
-Method & \flops\ & latency (s) \\
-\mr
-Time-domain (\TD) convolution & $2.4\times10^{13}$ & 0 \\
-Frequency-domain (\FD) convolution & $2.6\times10^8$ & 2201 \\
-\lloid\ (this work) & $4.7\times10^8$ & 0 \\
-\br
-\end{tabular}
-\end{indented}
-\end{table}
-
 \subsubsection{Conventional \TD{} method}
@@
 
-Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time-sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total \lloid\ requires
+Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time-sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total the \lloid\ algorithm requires
 $$
@@
 The second sum is carried out over the set of sample rates of all of the time slices rather than over the time slices themselves because some time slices may have the same sample rates.
-
-\subsubsection{Extrapolation of the computational cost to an Advanced \LIGO{} search}
-\editorial{Drew: I think this section be the last section of the results section.
-Should we move it to be Sec.
-4.2?
-Nick: I'm inclined to agree with Drew, though I see that we want to make explicit that it's an analysis of the algorithm and not of the implementation and the current placement aids that.
-Also, the given FLOPS numbers correspond to what choice of \SVD{} tolerance and $N^\shortuparrow$?}
-
-Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$ \flops\
-to cover a sub-bank of comprising $\sim 10^2$ out of $\sim 10^5$ sets of intrinsic
-source parameters.  Given that modern (ca. 2011) workstations can sustain
-computation rates up to $\sim 10^{10}$ \flops{}, and assuming that other regions of the
-parameter space have similar computational scaling, an entire search could be implemented
-with $\gtrsim 10$ machines.
-
-By comparison, using the \TD\ method to achieve the same latency costs $\sim 10^{13}$ \flops\ for this particular sub-bank, and so it would require $\sim 10^6$ present-day machines.
-Presently, the \LIGO{} Data Grid consists of only $\sim 10^4$ machines.
-
-\begin{comment}
-
-% Do we want to spend this much copy on the whitener?
-% What matters is that we have picked a low-latency whitening procedure.
-
-\subsection{Data Whitening}
-Matched filtering the SVD basis vectors has motivated a decoupling of the
-whitening routine from the matched filtering engine. This was necessary in
-order weight templates appropriately by whitening them before the SVD
-calculation, and, as such, we have also moved the data whitening outside the
-match filter engine. We have chosen to whiten the data using a running
-geometric average of the PSD computed by Hann windowing 8 second buffers of
-data and using 50\% overlapping buffers. The running geometric average is
-updated once per buffer from a median of the recent PSDs. We find that this
-algorithm is extremely fast to converge to an accurate PSD and also very robust
-against glitches.
-
-An inspiral, to leading order, has gravitational-wave frequency evolution given
-by
-%
-\begin{equation}
-%
-f(t) = \frac{1}{8 \pi \mathcal{M}} \left( \frac{t_c - t}{5 \mathcal{M}}
-\right)^{-3/8}\;,
-%
-\end{equation}
-%
-whose time derivative is given by
-%
-\begin{equation}
-%
-\frac{df}{dt} = \frac{3}{320 \pi \mathcal{M}^2} \left( \frac{t_c - t}{5
-\mathcal{M}} \right)^{-11/8}\;.
-%
-\end{equation}
-%
-Combining these, we find the frequency evolution as a function of frequency to
-be
-%
-\begin{equation}
-%
-\frac{df}{dt}(f_0) = \frac{3}{320 \pi \mathcal{M}^2} \left( 8 \pi \mathcal{M}
-f_0 \right)^{3/11}\;,
-%
-\end{equation}
-%
-which can be inverted to get the minimum frequency which has a given frequency
-derivative,
-%
-\begin{equation}
-%
-f_0 = \frac{1}{8 \pi \mathcal{M}} \left( \frac{320 \pi \mathcal{M}^2}{3}
-\frac{df}{dt} \right)^{11/3}\;.
-%
-\end{equation}
-
-If we allow frequency bins to be affected by an injection only once within the
-median history, we need to calculate the corresponding $df/dt$ for our PSD
-calculation. An FFT of a buffer with length $T$ will have a frequency
-resolution $df = 1/(2T)$. Hann windowing the data and overlapping buffers by
-50\% introduces correlations in neighboring frequency bins of the FFT. This
-means that we actually want the frequency to change by $df = 3/(2T)$ before the
-next PSD calculation, which happens a $dt = T/2$ later, resulting in a minimum
-$df/dt = 3/T^2$.
-
-Combining these results we find the lower frequency bound for our injections,
-assuming a chirp mass for a binary with $m_1 = m_2 = 1 M_{\odot}$ and $T = 8$
-s, is 23 Hz.
-
-We want to compute the equivalent injection density which would be biased as
-much as lalapps\_inspiral's PSD estimator, which allows $\sim3$ injections per
-2048 seconds. It computes the PSD by breaking the segment up into 16 256 second
-chunks. These chunks are then combined into to sets of 8 chunks each. The
-median of each set is then averaged across the sets for each frequency bin to
-produce the PSD estimate for that 2048 second segment. This procedure results
-in an average injection density of 1.5 per 8 PSDs in the median history, which
-is comparable to LLOID's 1 per 7 PSDs in the median history. Since LLOID has an
-injection history of 7 PSDs which span 32 seconds, this means LLOID can perform
-one injection every 32 seconds, or 64 injections per 2048 second segment, more
-than an order of magnitude increase in density over lalapps\_inspiral.
-\end{comment}

### 2011-06-07 implementation.tex
some rearrangment
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@949 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The choice of templates and the \SVD\ can be done in
-advance and will be valid as long as the detector noise spectrum remains
-roughly constant%
-\editorial{I'd rather not go into this.}%
-.  New templates and the \SVD\ can be computed asynchronously using
-updated spectrum estimates as they are available.
-
-\begin{figure}[htbp]
-	\label{fig:tmpltbank}
-	\begin{center}
-		\includegraphics{figures/tmpltbank.png}
-		\caption{Placement of template parameters used in this paper.
-The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.
-We shall design a filter bank to search for a small subset of these template parameters with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
-	\end{center}
-\end{figure}
-
-\begin{table}
-\begin{indented}
-\caption{Filter design for these 657 templates.
-From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.
-The \SVD{} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
-\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
-\br
-\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
-\ns
-\multicolumn{4}{c}{} &\crule{6} \\
-$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
-\mr
-4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
-512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
-256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
-128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
-64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
-64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
-64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
-32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
-32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
-32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
-32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
-\br
-\end{tabular}
-\end{indented}
-\end{table}
+The choice of templates and the \SVD\ can be done in advance.  New templates
+and the \SVD\ can be computed asynchronously using updated spectrum estimates
+as they are available.
 

### 2011-06-07 results.tex
some rearrangment
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@950 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 it.
-%We answer the following questions:
-%
-%I) What is the measured loss of \SNR\ due to the approximations used in the
-%\lloid\ algorithm in a realistic analysis pipeline?
-%
-%II) What is the measured latency of the implementation described in section
-%\ref{sec:implementation}?
-%
-%III) What is the measured computational cost of the implementation described in
-%section \ref{sec:implementation}?
+
+\subsection{Setup}
+
+We examine the performance of the \lloid\ algorithm on a small region of
+compact binary parameter space centered on typical NS-NS masses.  We begin by
+constructing a template bank that spans component masses between
+1~and~3~$M_\odot$ using a simulated advanced LIGO noise curve~\cite{lal}.  Then
+we create sub banks based on chirp mass slices in that parameter space.
+Figure{fig:tmpltbank} illustrates this procedure.
+\begin{figure}[htbp]
+	\label{fig:tmpltbank}
+	\begin{center}
+		\includegraphics{figures/tmpltbank.png}
+		\caption{Placement of template parameters used in this paper.
+The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.
+We shall design a filter bank to search for a small subset of these template parameters with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
+	\end{center}
+\end{figure}
+The result is that we obtain 657 templates with chirpmasses between 1.1955~and~1.2045~$M_\odot$.  With this
+sub-bank we were able to construct an efficient time slice decomposition that consisted of 11 time slices
+with sample rates between 32 and 4096 Hz shown in table \ref{tab:time_slices}.
+\begin{table}
+\begin{indented}
+\caption{\label{tab:time_slices} Filter design for these 657 templates.
+From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.
+The \SVD{} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
+\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
+\br
+\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
+\ns
+\multicolumn{4}{c}{} &\crule{6} \\
+$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
+\mr
+4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
+512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
+256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
+128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
+64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
+64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
+64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
+32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
+32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
+32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
+32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
+\br
+\end{tabular}
+\end{indented}
+\end{table}
+We use this template bank and decomposition for the remainder of this section.
 
@@
 
-%\subsection{Measured latency}
+\subsection{Lower bounds on computational cost compared to other methods}
 
-%In order to check the latency of our implementation of the \lloid\ algorithm we
-%used the pipeline described in \ref{fig:pipeline} with a live source connected
-%to the input.  We then measured the timestamp of the live source and the output
-%of the final \SNR\ time series.  Since \gstreamer\ is not a
-%sample-in-sample-out realtime infrastructure, it does work by passing small
-%buffers of data and filtered data through the pipeline.  We set the buffer size
-%on the input to be 1/4 s.  In an ideal case the latency should be nearly 1/4 s.
-%However, it is necessary to queue things in the pipeline to assure that there
-%are no issues with threading. Sometimes buffers get delayed.  We measured the
-%average filtering latency to be \FIXME{1 s}.
+\Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
+A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.
+Both the \FD\ method and \lloid\ are five orders of magnitude faster than the conventional \TD\ method.
+However, the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no latency at all.}
+\br
+Method & \flops\ & latency (s) \\
+\mr
+Time-domain (\TD) convolution & $2.4\times10^{13}$ & 0 \\
+Frequency-domain (\FD) convolution & $2.6\times10^8$ & 2201 \\
+\lloid\ (this work) & $4.7\times10^8$ & 0 \\
+\br
+\end{tabular}
+\end{indented}
+\end{table}
 
-%\subsection{Measured computational cost}
+\subsubsection{Extrapolation of the computational cost to an Advanced \LIGO{} search}
+\editorial{Drew: I think this section be the last section of the results section.
+Should we move it to be Sec.
+4.2?
+Nick: I'm inclined to agree with Drew, though I see that we want to make explicit that it's an analysis of the algorithm and not of the implementation and the current placement aids that.
+Also, the given FLOPS numbers correspond to what choice of \SVD{} tolerance and $N^\shortuparrow$?}
 
-%Using the pipeline described previously, we also measured the computational
-%cost.  We used an 8 socket, quad core AMD\texttrademark\ machine operating at
-%2.7 GHz per core.  We found that filtering \FIXME{657} templates in real time
-%required about \FIXME{6} of the available cores.  We thus conclude that we
-%require about 1 core for every \FIXME{100} templates.  We note that this is
-%well within the computational limits of a modern, computing cluster assuming
-%that the scaling holds for the entire mass range of interest, we would need
-%about \FIXME{300} cores per detector.  This requirment should be even less of a
-%burden in the advanced detecor era, when presumably clusters built in that era
-%will have a factor of 4 or 8 more cores than the ones today.
+Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$ \flops\
+to cover a sub-bank of comprising $\sim 10^2$ out of $\sim 10^5$ sets of intrinsic
+source parameters.  Given that modern (ca. 2011) workstations can sustain
+computation rates up to $\sim 10^{10}$ \flops{}, and assuming that other regions of the
+parameter space have similar computational scaling, an entire search could be implemented
+with $\gtrsim 10$ machines.
+
+By comparison, using the \TD\ method to achieve the same latency costs $\sim 10^{13}$ \flops\ for this particular sub-bank, and so it would require $\sim 10^6$ present-day machines.
+Presently, the \LIGO{} Data Grid consists of only $\sim 10^4$ machines.

### 2011-06-07 introduction.tex
minor rewording, added citation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@951 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 were observed in order to trigger followup observations by optical telescopes.
-The best demonstrated latencies were 30--60 minutes~\cite{LOOCUP}, which was
+The best demonstrated latencies were 30--60 minutes~\cite{HugheyGWPAW2011}, which was
 an important achievement, but far too late to catch prompt electromagnetic emission.
@@
 Second, inspiral filter banks consist of highly similar templates, admitting
-principal component analysis to reduce the number of templates.  We described a
-rank-reduction scheme based on singular value decomposition in
-\cite{Cannon:2010p10398}.  We will use both aspects to demonstrate that a very
+methods such as singular value decomposition to reduce the number of templates\cite{Cannon:2010p10398}
+We will use both aspects to demonstrate that a very
 low latency analysis with advance detection of compact binary sources is

### 2011-06-07 method.tex
pass after rearrangement
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@952 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In this section we describe a decomposition of the \CBC{} waveform
-space that reduces time-domain (\TD) filtering cost sufficiently to allow for the
+In this section we describe a decomposition of the \CBC{} waveform space that
+reduces time-domain (\TD) filtering cost sufficiently to allow for the
 possibility of \earlywarning\ detection with modest computing requirements.  We
@@
 a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 joint science
-runs.  We combine this with the \SVD\ rank-reduction method described
-in~\cite{Cannon:2010p10398} that exploits the redundancy of the template banks.
+runs~\cite{HugheyGWPAW2011}.  We combine this with the \SVD\ rank-reduction
+method described in~\cite{Cannon:2010p10398} that exploits the redundancy of
+the template banks.
 
@@
 \cite{Cannon:2010p10398} by nearly an order of magnitude.  In the next section
-we describe how we form our novel detection statistic using the time-slice
+we describe how we form our early warning detection statistic using the time-slice
 decomposition and the \SVD.
@@
 %
-\FIXME{This quantity is related to the early warning \SNR{} by the normalization
-of the partial filter defined by the present time slice.}
+This quantity is related to the early warning \SNR{} by the normalization
+of the partial filter defined by the present time slice, which may be computed
+in advance. \editorial{Chad: are we happy with that statement?}
 %
@@
 
-We now examine the computational cost scaling of the approximate implementation
-of a conventional \TD{} or \FD{} matched filter bank as compared
-with \lloid{}.  For convenience, table~\ref{tab:recap} provides a review of the notation
-that we will need in this section.
+We now examine the computational cost scaling of the conventional \TD{} or
+\FD{} matched filter procedure as compared with \lloid{}.  For convenience,
+table~\ref{tab:recap} provides a review of the notation that we will need in
+this section.
 %

### 2011-06-07 implementation.tex
pass after rearrangement
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@953 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 implement the \lloid\ algorithm.  For the near-term this should be a viable 
-solution for ultra-low latency searches with $\sim 1$s intrinsic latency.
+solution for ultra-low latency searches with order seconds of intrinsic latency.
 
@@
 gravitational-wave detector data to audio data it is not surprising that
-\gstreamer\ is
-useful for our purpose. \gstreamer\ also provides some useful stock signal
-processing elements such as resamplers and filters.  We have extended the
-\gstreamer\ framework by developing a library called \gstlal{}~\cite{gstlal}
-that provides elements for gravitational-wave data analysis.
-Figure~\ref{fig:pipeline} describes schematically how we implement the \lloid\
-algorithm using \gstlal\ and \gstreamer\ components.
+\gstreamer\ is useful for our purpose. \gstreamer\ also provides some useful
+stock signal processing elements such as resamplers and filters.  We have
+extended the \gstreamer\ framework by developing a library called
+\gstlal{}~\cite{gstlal} that provides elements for gravitational-wave data
+analysis.
 

### 2011-06-07 conclusions.tex
pass after rearrangement
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@954 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 improve the chances of detecting an electromagnetic counterpart in bands from
-gamma-rays down to radio.  We anticipate requiring $\sim$$10\times4$ modern
+gamma-rays down to radio.  We anticipate requiring $\sim100$ modern
 multi-core computers to analyze a four-detector network of gravitational-wave data
@@
 procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
-to a 1/(16~Hz) cadence introducing an up-front latency.  Data
-aggregation from the observatories will travel over various networks, each
-capable of high bandwith but perhaps only modest latency.  This could amount to
-a similar latency of $\sim$100 ms.  Lastly, unless a real-time infrastructure is
-adopted post data acquisition, it is likely that there will be an inherent
-latency introduced by such infrastructure.  We have shown a prototype
-implementation using \gstlal\ that is capable of $\sim$1 s latency. In our
-opinion, significant work would have to be done in order to improve upon this
-number. For example, a tighter integration of analysis and data acquisition
-would be beneficial. This should be considered for third-generation detector
-design.
+to a 1/(16~Hz) cadence introducing an up-front latency~\cite{CITE_CDS}.  Data aggregation from
+the observatories will travel over various networks, each capable of high
+bandwith but perhaps only modest latency.  This could amount to a similar
+latency of $\sim$100 ms.  Lastly, unless a real-time infrastructure is adopted
+post data acquisition, it is likely that there will be an inherent latency
+introduced by such infrastructure.  We have shown a prototype implementation
+using \gstlal\ that is capable of order seconds latency. In our opinion,
+significant work would have to be done in order to improve upon this number.
+For example, a tighter integration of analysis and data acquisition would be
+beneficial. This should be considered for third-generation detector design.
 

### 2011-06-07 results.tex
pass after rearrangement
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@955 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 we create sub banks based on chirp mass slices in that parameter space.
-Figure{fig:tmpltbank} illustrates this procedure.
+Figure \ref{fig:tmpltbank} illustrates this procedure.
 \begin{figure}[htbp]
-	\label{fig:tmpltbank}
 	\begin{center}
 		\includegraphics{figures/tmpltbank.png}
-		\caption{Placement of template parameters used in this paper.
+		\caption{\label{fig:tmpltbank}Placement of template parameters used in this paper.
 The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.
@@
 exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
-implementations of resampling that cause signal loss.  The \SNR\ loss is to be compared
-with the mismatch of 0.03 that arises from the discreteness of the
-template bank which is typically.  We will consider an acceptable
-\SNR\ loss to be a factor of 10 smaller than this, i.e., no more than $\sim$0.003.
+implementations of resampling that cause signal loss.  The \SNR\ loss is to be
+compared with the mismatch of 0.03 that arises from the discreteness of the
+template bank.  We will consider a target \SNR\ loss to be a factor of 10
+smaller than this, i.e., no more than $\sim$0.003.
 
@@
 In this section we check the effect of the \SVD\ tolerance parameter defined
-in~\cite{Cannon:2010p10398}.  By changing the tolerance of the reconstruction of
-the physical waveforms one effects directly the mismatch that results from the
-truncation of the orthogonal filter matrix.  The conditions presented here are
-more complicated than in the original work~\cite{Cannon:2010p10398} due to the
-inclusion of the time sliced templates and resampling.  \FIXME{However, we find
-that the results are quite as expected.}  Figure \ref{fig:hist-svd-tolerance}
-demonstrates that it is possible to acheive typical \SNR\ losses of $\ll1\%$.
-However, we do notice that the mismatch saturates with respect to the tolerance.
-This could be the effect of the resampling, or another \SNR\ loss that we did
-not model or expect.  However the saturation is still an order of magnitude below our target mismatch of $\sim$0.003.  We find that an \SVD\ tolerance of
-$\left(1-10^{-4}\right)$ is adequate. 
+in~\cite{Cannon:2010p10398}.  By changing the tolerance of the reconstruction
+of the physical waveforms one effects directly the mismatch that results from
+the truncation of the orthogonal filter matrix.  The conditions presented here
+are more complicated than in the original work~\cite{Cannon:2010p10398} due to
+the inclusion of the time sliced templates and resampling.  Figure
+\ref{fig:hist-svd-tolerance} demonstrates that it is possible to acheive
+typical \SNR\ losses of $\ll1\%$.  However, we do notice that the mismatch
+saturates with respect to the tolerance.  This could be the effect of the
+resampling, or another \SNR\ loss that we did not model or expect.  However the
+saturation is still an order of magnitude below our target mismatch of
+$\sim$0.003.  We find that an \SVD\ tolerance of $\left(1-10^{-4}\right)$ is
+adequate to acheive our target SNR loss.
 %
@@
 
+We are now prepared to offer the estimated computational cost of filtering this
+sub-bank of templates compared to other methods.  We used the results of the
+previous sub sections to set the \SVD\ tolerance to $1-10^{-4}$  and the
+interpolation filter length to 16.  We find that the \lloid\ method provides
+comparable performance to the standard \FD\ approach but with no intrinsic
+latency.  Table \ref{table:flops} lists the costs for filtering this sub-bank.
+%
 \Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
@@
 
-\subsubsection{Extrapolation of the computational cost to an Advanced \LIGO{} search}
-\editorial{Drew: I think this section be the last section of the results section.
-Should we move it to be Sec.
-4.2?
-Nick: I'm inclined to agree with Drew, though I see that we want to make explicit that it's an analysis of the algorithm and not of the implementation and the current placement aids that.
-Also, the given FLOPS numbers correspond to what choice of \SVD{} tolerance and $N^\shortuparrow$?}
+\subsection{Extrapolation of the computational cost to an Advanced \LIGO{} search}
 
-Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$ \flops\
-to cover a sub-bank of comprising $\sim 10^2$ out of $\sim 10^5$ sets of intrinsic
-source parameters.  Given that modern (ca. 2011) workstations can sustain
-computation rates up to $\sim 10^{10}$ \flops{}, and assuming that other regions of the
-parameter space have similar computational scaling, an entire search could be implemented
-with $\gtrsim 10$ machines.
+Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$
+\flops\ to cover a sub-bank comprising of $\sim 10^2$ out of the total $\sim
+10^5$ mass pairs.  Given that modern (ca. 2011) workstations can sustain
+computation rates up to $\sim 10^{10}$ \flops{}, and assuming that other
+regions of the parameter space have similar computational scaling\footnote{The
+length of templates varies over the parameter space.  Lower mass templates are
+longer and would require more computational cost, higher mass templates are
+shorter and would require less.  We consider the estimates based on this
+sub-bank to be a reasonable representation, however}, an entire search could be
+implemented with $\gtrsim 10$ machines.
 

### 2011-06-07 conclusions.tex
lloid_paper: Polishing pass post-Chad
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@958 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
-to a 1/(16~Hz) cadence introducing an up-front latency~\cite{CITE_CDS}.  Data aggregation from
+to a 1/(16~Hz) cadence, introducing an up-front latency~\cite{CITE_CDS}. Data
+aggregation from
 the observatories will travel over various networks, each capable of high
@@
 introduced by such infrastructure.  We have shown a prototype implementation
-using \gstlal\ that is capable of order seconds latency. In our opinion,
-significant work would have to be done in order to improve upon this number.
-For example, a tighter integration of analysis and data acquisition would be
-beneficial. This should be considered for third-generation detector design.
+of \lloid{} using \gstlal\ that is capable of order seconds latency. Based
+on our experience, significant work would have to be done in order to improve
+upon this number, for example, more tightly integrating analysis and data
+acquisition. This should be considered for third-generation detector design.
 
@@
 Ref.~\cite{Fairhurst2009} discusses some of the theoretical prospects for early sky
-localization.  Our future work will explore the prospects of early warning
+localization.  Our future work will explore the prospects of early-warning
 detection with realistic simulations of binary mergers using the infrastructure

### 2011-06-07 implementation.tex
lloid_paper: Polishing pass post-Chad
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@958 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \subsection{Planning stage}
-
-The choice of templates and the \SVD\ can be done in advance.  New templates
-and the \SVD\ can be computed asynchronously using updated spectrum estimates
-as they are available.
 
@@
 the data acquisition and storage system of the gravitational-wave
-observatories.  A slightly more modest goal is to leverage existing low
+observatories.  A slightly more modest goal is to leverage existing low-%
 latency, but not real-time, signal processing infrastructure in order to
 implement the \lloid\ algorithm.  For the near-term this should be a viable 
-solution for ultra-low latency searches with order seconds of intrinsic latency.
+solution for searches with order seconds of intrinsic latency.
 
@@
 
-The early-warning \SNR\ for each time slice is formed by accumulating interpolated early
-warning \SNR\ from the subsequent time slice.  This process continues until we have 
-worked our way to the \SNR\ of the original templates at the full sample rate $f^0$.  In 
-this way, the \lloid\ algorithm and this implementation leads to a simple early warning
-pipeline.
+The early-warning \SNR\ for each time slice is formed by accumulating
+interpolated early-warning \SNR\ from the subsequent time slice.  This process
+continues until we have worked our way to the \SNR\ of the original templates
+at the full sample rate $f^0$.  In this way, the \lloid\ algorithm and this
+implementation naturally leads to a simple early-warning pipeline.

### 2011-06-07 introduction.tex
lloid_paper: Polishing pass post-Chad
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@958 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Second, inspiral filter banks consist of highly similar templates, admitting
-methods such as singular value decomposition to reduce the number of templates\cite{Cannon:2010p10398}
+methods such as singular value decomposition (\SVD{}) to reduce the number of templates\cite{Cannon:2010p10398}
 We will use both aspects to demonstrate that a very

### 2011-06-07 method.tex
lloid_paper: Polishing pass post-Chad
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@958 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 runs~\cite{HugheyGWPAW2011}.  We combine this with the \SVD\ rank-reduction
-method described in~\cite{Cannon:2010p10398} that exploits the redundancy of
+method of~\cite{Cannon:2010p10398} that exploits the redundancy of
 the template banks.
@@
 conventional \FD\ method, which requires a signficant latency in order to be
-computationally cheap. Our method, \lloid\footnote{\lloid, The term strictly
-means \emph{Low Latency Online Inspiral Detection} and was a code name for
-this procedure as it was being developed. It is now known in some GW circles.
-We feel that although the name doesn't carry much specific information about
-the details of the algorithm it is far too late to do anything about it now.}
+computationally cheap. Our method, \lloid\footnote{\lloid stands for \emph{Low
+Latency Online Inspiral Detection} and was a code name for this procedure
+throughout development. Although the name doesn't carry any specific
+information about the algorithm, it became known within the gravitational-wave
+community and has stuck.}%
 involves two transformations of the template waveforms that produce a set of
@@
 orthogonal across the mass parameter space, and are in fact highly redundant.
-We use the singular value decomposition to produce an orthogonal filter set
+We use the \SVD{} to produce an orthogonal filter set
 from the time-sliced templates~\cite{Cannon:2010p10398}.  We find that this
@@
 We will show
-in the next section that this, combined with the singular value docomposition,
+in the next section that this, combined with the \SVD{},
 is sufficient to enable a computationally efficient \TD{} search and
@@
 sample rate.  Specifically we consider here time slice boundaries with the
-smallest power-of-two sample rate that sub-critically samples the time sliced
+smallest power-of-two sample rate that sub-critically samples the time-sliced
 template.  The time slices consist of the $S$ intervals
@@
 
-We note that the time slice decomposition in equation~(\ref{eq:time-slices}) is
+We note that the time-slice decomposition in equation~(\ref{eq:time-slices}) is
 manifestly orthogonal since the time slices are disjoint in time.  In the next
 section we examine how to reduce the number of filters within each time slice
-via singular value decomposition of the time sliced templates.
+via \SVD{} of the time-sliced templates.
 
-\subsection{Reducing the number of filters with the singular value
-decomposition}
+\subsection{Reducing the number of filters with the \SVD{}}
 
@@
 particular minimum match by designing an appropriate set of \SVD\ {\em
-basis templates}.  A numerical technique based on the singular value
-decomposition (\SVD) of inspiral template banks is demonstrated
-in~\cite{Cannon:2010p10398}.  Similarly, the time sliced templates described
+basis templates}, previously demonstrated
+in~\cite{Cannon:2010p10398}.  Similarly, the time-sliced templates described
 above can be approximated to arbitrary accuracy by expansion into a set of
@@
 \cite{Cannon:2010p10398} by nearly an order of magnitude.  In the next section
-we describe how we form our early warning detection statistic using the time-slice
+we describe how we form our early-warning detection statistic using the time-slice
 decomposition and the \SVD.
@@
 %
-This quantity is related to the early warning \SNR{} by the normalization
+This quantity is related to the early-warning \SNR{} by the normalization
 of the partial filter defined by the present time slice, which may be computed
@@
 recursion relation can be realized as a multirate filter network with outputs for
-each of the early warning \SNR{}s.
+each of the early-warning \SNR{}s.
 

### 2011-06-07 introduction.tex
Remove reference to the secret shadow ihope pipeline.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@960 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 pipelines operated in a low-latency configuration to send astronomical alerts,
-namely \textsc{mbta}, ihope,
-Coherent WaveBurst, and Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
+namely \textsc{mbta}, Coherent WaveBurst, and
+Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
 \editorial{Get references for these low-latency pipelines.} The S6 analyses

### 2011-06-07 method.tex
lloid_paper: Forgot {}, so a macro ate a space
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@959 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: Changes from Mireia and misc polishing and clipping
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@961 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 conventional \FD\ method, which requires a signficant latency in order to be
-computationally cheap. Our method, \lloid\footnote{\lloid stands for \emph{Low
+computationally cheap. Our method, \lloid{}\footnote{\lloid{} stands for
+\emph{Low
 Latency Online Inspiral Detection} and was a code name for this procedure
@@
 templates, the early time slices have significantly lower bandwidth and can be
-safely downsampled.  The downsampling removes a factor of $\sim$100 samples
-from the early part of the waveform and allows the filters to be evaluated at
-about $\sim$100 times lower rate.  This amounts to more than a factor of 10000
+safely downsampled.  The downsampling reduces the number of data samples by a
+factor of $\sim$100 from the early part of the waveform and allows the filters
+to be evaluated at about $\sim$$1/100^\mathrm{th}$ the sample rate. This
+amounts to more than a factor of 10000
 reduction in the floating-point operations per second required to filter the
@@
 $$
-The second sum is carried out over the set of sample rates of all of the time slices rather than over the time slices themselves because some time slices may have the same sample rates.
+The second sum is carried out over the set of distinct sample rates rather than over the time slices themselves.

### 2011-06-07 results.tex
lloid_paper: Changes from Mireia and misc polishing and clipping
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@961 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 constructing a template bank that spans component masses between
-1~and~3~$M_\odot$ using a simulated advanced LIGO noise curve~\cite{lal}.  Then
+1~and~3~$M_\odot$ using a simulated Advanced LIGO noise curve~\cite{lal}.  Then
 we create sub banks based on chirp mass slices in that parameter space.
@@
 
-It is also within reason that some \SNR\ loss could arise from other suboptimal
-implementations of the \lloid\ algorithm in our test pipeline.  To gauge
+It is also within reason that some \SNR\ loss could arise from other
+suboptimalities of implementation in our test pipeline.  To gauge
 accurately how well the pipeline is performing we tested the response to a
@@
 are more complicated than in the original work~\cite{Cannon:2010p10398} due to
-the inclusion of the time sliced templates and resampling.  Figure
+the inclusion of the time sliced templates and resampling.  Figure~%
 \ref{fig:hist-svd-tolerance} demonstrates that it is possible to acheive
@@
 $\sim$0.003.  We find that an \SVD\ tolerance of $\left(1-10^{-4}\right)$ is
-adequate to acheive our target SNR loss.
+adequate to acheive our target \SNR{} loss.
 %
@@
 sub-bank of templates compared to other methods.  We used the results of the
-previous sub sections to set the \SVD\ tolerance to $1-10^{-4}$  and the
-interpolation filter length to 16.  We find that the \lloid\ method provides
-comparable performance to the standard \FD\ approach but with no intrinsic
-latency.  Table \ref{table:flops} lists the costs for filtering this sub-bank.
+previous subsections to set the \SVD\ tolerance to $1-10^{-4}$  and the
+interpolation filter length to 16.
+Table \ref{table:flops} lists the costs for filtering this sub-bank.
+Both the \FD\ method and \lloid\ are five orders of magnitude faster than the conventional \TD\ method.
+However, the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no latency at all.
 %
 \Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
-A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.
-Both the \FD\ method and \lloid\ are five orders of magnitude faster than the conventional \TD\ method.
-However, the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no latency at all.}
+A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.}
 \br
-Method & \flops\ & latency (s) \\
+Method & \flops\ & Latency (s) \\
 \mr

### 2011-06-07 implementation.tex
lloid_paper: more cleanups
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@962 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-First, the sample rate of the whitened detector data is reduced to successively
+The whitened detector data is reduced to successively
 lower sample rates by decimation. Decimation involves applying an anti-aliasing
@@
 element.  The detector data is provided at every power-of-two sample rate
-required by the template time slices described in \eqref{eq:time-slices}. These
-parallel decimated data streams are fed into parallel \fir\ filter banks in the
-next stage.
+required by the template time slices described in \eqref{eq:time-slices}. Next,
+these decimated data streams are fed into parallel \fir\ filter banks.
 

### 2011-06-07 results.tex
lloid_paper: more cleanups
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@962 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In this section we test our implementation of the \lloid\ method using the \gstreamer\
-pipeline described in the previous section.  We calculate the measured \SNR\
-loss due to the approximations of the \lloid\ method and our implementation of
-it.
+In this section we evaluate the \lloid{} algorithm using our \gstreamer{}-based implementation described in the previous section. We calculate the
+measured \SNR\ loss due to the approximations of the \lloid\ method and our
+implementation of it. Using settings that give acceptable \SNR{} loss for
+our chosen parameter space, we then compute the operation counts.
 
@@
 We examine the performance of the \lloid\ algorithm on a small region of
-compact binary parameter space centered on typical NS-NS masses.  We begin by
-constructing a template bank that spans component masses between
-1~and~3~$M_\odot$ using a simulated Advanced LIGO noise curve~\cite{lal}.  Then
-we create sub banks based on chirp mass slices in that parameter space.
-Figure \ref{fig:tmpltbank} illustrates this procedure.
+compact binary parameter space centered on typical \textsc{ns}--\textsc{ns}
+masses.  We begin by constructing a template bank that spans component masses
+between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
+curve~\cite{lal}.  Then we create sub-banks based on chirp-mass slices in that
+parameter space.  Figure \ref{fig:tmpltbank} illustrates this procedure.
 \begin{figure}[htbp]
@@
 The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.
-We shall design a filter bank to search for a small subset of these template parameters with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
+We design a filter bank to search for a small subset of these templates with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
 	\end{center}
 \end{figure}
-The result is that we obtain 657 templates with chirpmasses between 1.1955~and~1.2045~$M_\odot$.  With this
-sub-bank we were able to construct an efficient time slice decomposition that consisted of 11 time slices
+The result is that we obtain 657 templates with chirp masses between 1.1955~and~1.2045~$M_\odot$.  With this
+sub-bank we were able to construct an efficient time-slice decomposition that consisted of 11 time slices
 with sample rates between 32 and 4096 Hz shown in table \ref{tab:time_slices}.
@@
 From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.
-The \SVD{} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
-\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
+We vary \SVD{} tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
+\item[]\begin{tabular}{rr@{,\,}lc*{6}{r}}
 \br
-\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
+\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\
 \ns
 \multicolumn{4}{c}{} &\crule{6} \\
-$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
+$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\
 \mr
@@
 in~\cite{Cannon:2010p10398}.  By changing the tolerance of the reconstruction
-of the physical waveforms one effects directly the mismatch that results from
+of the physical waveforms one directly affects the mismatch that results from
 the truncation of the orthogonal filter matrix.  The conditions presented here
 are more complicated than in the original work~\cite{Cannon:2010p10398} due to
-the inclusion of the time sliced templates and resampling.  Figure~%
-\ref{fig:hist-svd-tolerance} demonstrates that it is possible to acheive
+the inclusion of the time-sliced templates and resampling.  Figure~%
+\ref{fig:hist-svd-tolerance} demonstrates that it is possible to achieve
 typical \SNR\ losses of $\ll1\%$.  However, we do notice that the mismatch
@@
 previous subsections to set the \SVD\ tolerance to $1-10^{-4}$  and the
-interpolation filter length to 16.
-Table \ref{table:flops} lists the costs for filtering this sub-bank.
-Both the \FD\ method and \lloid\ are five orders of magnitude faster than the conventional \TD\ method.
-However, the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no latency at all.
+interpolation filter length to 16. Table \ref{table:flops} lists the
+theoretical costs for filtering this sub-bank. Both the \FD\ method and \lloid\
+are five orders of magnitude faster than the conventional \TD\ method.
+However, the \FD\ method has a latency of over half of an hour, whereas \lloid\
+has no latency at all.
 %
@@
 regions of the parameter space have similar computational scaling\footnote{The
-length of templates varies over the parameter space.  Lower mass templates are
-longer and would require more computational cost, higher mass templates are
-shorter and would require less.  We consider the estimates based on this
-sub-bank to be a reasonable representation, however}, an entire search could be
-implemented with $\gtrsim 10$ machines.
+lengths of templates varies over the parameter space; lower-mass templates are
+longer and would require more computation to analyze while higher-mass
+templates are shorter and would require less. However, we consider the
+estimates based on this sub-bank to be a reasonable representation.}, an entire
+search could be implemented with $\gtrsim 10$ machines.
 

### 2011-06-08 introduction.tex
Fixed punctuation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@964 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Second, inspiral filter banks consist of highly similar templates, admitting
-methods such as singular value decomposition (\SVD{}) to reduce the number of templates\cite{Cannon:2010p10398}
-We will use both aspects to demonstrate that a very
+methods such as singular value decomposition (\SVD{}) to reduce the number of
+templates \cite{Cannon:2010p10398}. We will use both aspects to demonstrate that a very
 low latency analysis with advance detection of compact binary sources is

### 2011-06-08 method.tex
Removed a footnote
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@965 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 conventional \FD\ method, which requires a signficant latency in order to be
-computationally cheap. Our method, \lloid{}\footnote{\lloid{} stands for
-\emph{Low
-Latency Online Inspiral Detection} and was a code name for this procedure
-throughout development. Although the name doesn't carry any specific
-information about the algorithm, it became known within the gravitational-wave
-community and has stuck.}%
+computationally cheap. Our method, called \lloid{} (a backronym for
+Low Latency Online Inspiral Detection),
 involves two transformations of the template waveforms that produce a set of

### 2011-06-08 results.tex
Removed some fluff  ('Checking the effect of' --> 'Effect of'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@963 ccdd05d5-56c1-461a-bf98-85f6876523b4
reworded a sentence in section 4.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@966 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
-curve~\cite{lal}.  Then we create sub-banks based on chirp-mass slices in that
-parameter space.  Figure \ref{fig:tmpltbank} illustrates this procedure.
+curve~\cite{lal}.  Then we create sub-banks by partitioning the parameter space by
+chirp mass.  Figure \ref{fig:tmpltbank} illustrates this procedure.
 \begin{figure}[htbp]
@@
 
-\subsubsection{Checking the effect of the \SVD\ tolerance}
+\subsubsection{Effect of the \SVD\ tolerance}
 
@@
 
-\subsubsection{Checking the effect of the interpolation filter length}
+\subsubsection{Effect of the interpolation filter length}
 

### 2011-06-08 results.tex
Fixed typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@967 ccdd05d5-56c1-461a-bf98-85f6876523b4
Small tweaks to figure 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@970 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.
-We design a filter bank to search for a small subset of these templates with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
+We design a filter bank to search for a subset of 657 of these templates with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
 	\end{center}
@@
 Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$
-\flops\ to cover a sub-bank comprising of $\sim 10^2$ out of the total $\sim
+\flops\ to cover a sub-bank comprising $\sim 10^2$ out of the total $\sim
 10^5$ mass pairs.  Given that modern (ca. 2011) workstations can sustain

### 2011-06-08 introduction.tex
Updates
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@971 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 As a compact binary system loses energy to gravitational waves, its orbital
-separation decays, leading to a run-away inspiral with the
-gravitational-wave amplitude and frequency increasing until the system
-eventually merges soon after the innermost stable circular orbit (\ISCO). If a
-neutron star is involved it may become tidally disrupted near the merger.  This
-disrupted matter can fuel a bright electromagnetic counterpart in the merger's aftermath~\cite{shibata:2007}.
-The coalescence of compact binary systems consisting of neutron stars
-(\textsc{ns}) and/or black holes (\textsc{bh}) is the most promising source of
-gravitational radiation for Advanced \LIGO, Virgo, \GEO\, and
-\LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb, LCGTWeb}.  Tens of binary coalescence
-events are expected to be observed in the advanced detector era later this
-decade~\cite{Abadie:2010p10836}. In the era of first-generation
-detectors the gravitational-wave community
-initiated a project to send alerts when potential gravitational-wave transients
-were observed in order to trigger followup observations by optical telescopes.
-The best demonstrated latencies were 30--60 minutes~\cite{HugheyGWPAW2011}, which was
-an important achievement, but far too late to catch prompt electromagnetic emission.
+separation decays, leading to a run-away inspiral with the gravitational-wave
+amplitude and frequency increasing until the system eventually merges soon
+after the innermost stable circular orbit (\ISCO). If a neutron star is
+involved it may become tidally disrupted near the merger.  This disrupted
+matter can fuel a bright electromagnetic counterpart in the merger's
+aftermath~\cite{shibata:2007}.
+
+Gravitational-wave detectors are sensitive the to these gravitational-wave
+signals.  Indeed, the coalescence of compact binary systems consisting of
+neutron stars (\textsc{ns}) and/or black holes (\textsc{bh}) is the most
+promising source of gravitational radiation for Advanced \LIGO, Virgo, \GEO\,
+and \LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb, LCGTWeb}.  Tens of binary
+coalescence events per year are expected to be observed in the advanced
+detector era later this decade~\cite{Abadie:2010p10836}. In the era of
+first-generation detectors the gravitational-wave community initiated a project
+to send alerts when potential gravitational-wave transients were observed in
+order to trigger followup observations by optical telescopes.  The best
+demonstrated latencies were 30--60 minutes~\cite{HugheyGWPAW2011}, which was an
+important achievement, but far too late to catch prompt electromagnetic
+emission.
 
@@
 for threshold crossings before the gravitational-wave signal leaves the
-detection band, it is possible to trade some \SNR{}
-for latency.  Figure \ref{fig:earlywarning} shows projected early trigger rates
-for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming the event rate predictions
-in~\cite{Abadie:2010p10836}.
+detection band, it is possible to trade some \SNR{} and sky localization
+accuracy~\cite{Fairhurst2009} for latency.  Figure \ref{fig:earlywarning} shows
+projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
+\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.
 %

### 2011-06-08 method.tex
Updates
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@971 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The first step of our proposed method is to divide the templates into
-time slices in a \TD{} analogue to the \FD{}
-decomposition described in~\cite{Marion2004, Buskulic2010}.
-A matched filter is constructed for each time
-slice.  The outputs form an ensemble of partial \SNR{} streams.  By linearity,
-these partial \SNR\ streams can be suitably time-delayed and summed to
-reproduce the \SNR\ of the full template.  To wit,
+The first step of our proposed method is to divide the templates into time
+slices in a \TD{} analogue to the \FD{} decomposition described
+in~\cite{Marion2004, Buskulic2010}.  To wit,
 %
@@
 \end{equation}
-\editorial{Drew: Should this be $h_{i}[k]$ or $\rho_{i}[k]$?}
 %
-for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0 t^S = N$.
-We will show
-in the next section that this, combined with the \SVD{},
-is sufficient to enable a computationally efficient \TD{} search and
+for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
+t^S = N$.  A matched filter is constructed for each time slice.  The outputs
+form an ensemble of partial \SNR{} streams.  By linearity, these partial \SNR\
+streams can be suitably time-delayed and summed to reproduce the \SNR\ of the
+full template.  We will show in the next section that this, combined with the
+\SVD{}, is sufficient to enable a computationally efficient \TD{} search and
 furthermore is an essential part of an \earlywarning\ detection scheme.
@@
 \end{equation}
-%
-\editorial{Say a word or two about \SVD\ tolerance?}%
 %

### 2011-06-08 conclusions.tex
Minor rewording in conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@968 ccdd05d5-56c1-461a-bf98-85f6876523b4
Adding line on comp det stat
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@972 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-We have demonstrated a computationally feasible procedure for the rapid and even preemptive
-detection of gravitational waves emitted during the coalescence of neutron
-stars and stellar-mass black holes. Our method is as fast as standard \fft{}
-convolutions but allows for sample-in-sample-out processing.
-Our search targets are expected to
-produce prompt electromagnetic signals and may be the progenitors of some short
-hard gamma-ray bursts.  Rapid alerts to the broader astronomical community will
-improve the chances of detecting an electromagnetic counterpart in bands from
-gamma-rays down to radio.  We anticipate requiring $\sim100$ modern
-multi-core computers to analyze a four-detector network of gravitational-wave data
-for binary neutron stars and stellar-mass black holes.  This is well within the
-current computing capabilities of the \LIGO{} Data Grid~\cite{LDG}.
+We have demonstrated a computationally feasible procedure for the rapid and
+even preemptive detection of gravitational waves emitted during the coalescence
+of neutron stars and stellar-mass black holes. Our method is as fast as
+standard \fft{} convolutions but allows for zero latency, sample-in-sample-out
+operation.  Our search targets are expected to produce prompt electromagnetic
+signals and may be the progenitors of some short hard gamma-ray bursts.  Rapid
+alerts to the broader astronomical community will improve the chances of
+detecting an electromagnetic counterpart in bands from gamma-rays down to
+radio.  We anticipate requiring $\sim100$ modern multi-core computers to
+analyze a four-detector network of gravitational-wave data for binary neutron
+stars and stellar-mass black holes.  This is well within the current computing
+capabilities of the \LIGO{} Data Grid~\cite{LDG}. In the future, this could be
+improved upon further by conditionally reconstructing the \SNR{} time-series
+only during times when a composite dection statistic crosses a
+threshold~\cite{svd-compdetstat}.
 
@@
 
-\editorial{Should we mention composite detection?}
-\editorial{Yes, although only as something else that we could employ to further reduce compuational cost.}
-
 \editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring

### 2011-06-08 conclusions.tex
Adding GW macro
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@974 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We have demonstrated a computationally feasible procedure for the rapid and
-even preemptive detection of gravitational waves emitted during the coalescence
+even preemptive detection of \GW{}s emitted during the coalescence
 of neutron stars and stellar-mass black holes. Our method is as fast as
@@
 radio.  We anticipate requiring $\sim100$ modern multi-core computers to
-analyze a four-detector network of gravitational-wave data for binary neutron
+analyze a four-detector network of \GW{} data for binary neutron
 stars and stellar-mass black holes.  This is well within the current computing

### 2011-06-08 implementation.tex
Adding GW macro
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@974 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In this section we describe an implementation of the \lloid\ method described
-in section \ref{sec:method} suitable for rapid gravitational-wave
+in section \ref{sec:method} suitable for rapid \GW{}
 searches for \CBC{}s.  The \lloid\ method requires several
@@
 system.  However, such a system would likely require integration directly into
-the data acquisition and storage system of the gravitational-wave
+the data acquisition and storage system of the \GW{}
 observatories.  A slightly more modest goal is to leverage existing low-%
@@
 computers to streaming media servers.  Given the similarities of
-gravitational-wave detector data to audio data it is not surprising that
+\GW{} detector data to audio data it is not surprising that
 \gstreamer\ is useful for our purpose. \gstreamer\ also provides some useful
@@
 extended the \gstreamer\ framework by developing a library called
-\gstlal{}~\cite{gstlal} that provides elements for gravitational-wave data
+\gstlal{}~\cite{gstlal} that provides elements for \GW{} data
 analysis.

### 2011-06-08 introduction.tex
Adding GW macro
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@974 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-As a compact binary system loses energy to gravitational waves, its orbital
-separation decays, leading to a run-away inspiral with the gravitational-wave
+As a compact binary system loses energy to gravitational waves (\GW{}s), its
+orbital separation decays, leading to a run-away inspiral with the \GW{}
 amplitude and frequency increasing until the system eventually merges soon
@@
 
-Gravitational-wave detectors are sensitive the to these gravitational-wave
+\GW{} detectors are sensitive the to these \GW{}
 signals.  Indeed, the coalescence of compact binary systems consisting of
@@
 detector era later this decade~\cite{Abadie:2010p10836}. In the era of
-first-generation detectors the gravitational-wave community initiated a project
-to send alerts when potential gravitational-wave transients were observed in
+first-generation detectors the \GW{} community initiated a project
+to send alerts when potential \GW{} transients were observed in
 order to trigger followup observations by optical telescopes.  The best
@@
 fade too quickly to even catch the tails of the afterglows. Rapid
-gravitational-wave transient alerts could enable the first observation of
+\GW{} transient alerts could enable the first observation of
 prompt optical flashes and the rise of afterglows from short \GRB{}s, and, for
@@
 precursor~\cite{0004-637X-723-2-1711}. An electromagnetic counterpart would
-vastly boost confidence in the gravitational-wave detection and provide the
+vastly boost confidence in the \GW{} detection and provide the
 tight sky localization necessary to allow determination of the source's host
-galaxy, which leads to a redshift measurement. A coincident gravitational-wave
+galaxy, which leads to a redshift measurement. A coincident \GW{}
 observation will provide a luminosity distance measurement. With these
@@
 
-To this end, we have the ambition of reporting gravitational-wave candidates
+To this end, we have the ambition of reporting \GW{} candidates
 not minutes \emph{after} the merger, as has already been accomplished, but
 seconds \emph{before}.  By examining the signal to noise ratio (\SNR{}) stream
-for threshold crossings before the gravitational-wave signal leaves the
+for threshold crossings before the \GW{} signal leaves the
 detection band, it is possible to trade some \SNR{} and sky localization
@@
 HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved
-the best gravitational-wave trigger-generation latencies of 2--5 minutes.  We
+the best \GW{} trigger-generation latencies of 2--5 minutes.  We
 assume that in the advanced detector era the vetting process will be automated,
-so current gravitational-wave search methodology and telescope actuation would
+so current \GW{} search methodology and telescope actuation would
 dominate latency.
@@
 \cite{numerical-recipes-chapter-13} for frequency domain (\FD) convolution,
-currently the preferred method in gravitational-wave
+currently the preferred method in \GW{}
 searches.  The most obvious example of the latter is the time domain

### 2011-06-08 method.tex
Adding GW macro
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@974 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 parameters $\vec\theta$ that determine the amplitude and phase evolution of the
-gravitational wave strain. For systems where the effects of spin can be
+\GW{} strain. For systems where the effects of spin can be
 ignored, the intrinsic source parameters are the component masses of the
@@
 the detector is a linear combination of two waveforms corresponding to the
-`$+$' and `$\times$' gravitational wave polarizations.  Thus, for any value of
+`$+$' and `$\times$' \GW{} polarizations.  Thus, for any value of
 $\vec\theta$ we must implement 2 filters.  The coefficients for the $\numtmps$ filters

### 2011-06-08 introduction.tex
next round of comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@975 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-To this end, we have the ambition of reporting \GW{} candidates
-not minutes \emph{after} the merger, as has already been accomplished, but
-seconds \emph{before}.  By examining the signal to noise ratio (\SNR{}) stream
-for threshold crossings before the \GW{} signal leaves the
-detection band, it is possible to trade some \SNR{} and sky localization
+To this end, we have the ambition of reporting \GW{} candidates not minutes
+\emph{after} the merger, as has already been accomplished, but seconds
+\emph{before}.  By examining the signal to noise ratio (\SNR{}) stream for
+threshold crossings before the \GW{} signal leaves the detection band, it is
+possible to trade some \SNR{} and sky localization
 accuracy~\cite{Fairhurst2009} for latency.  Figure \ref{fig:earlywarning} shows
 projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
-\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.
+\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.  The
+most likely estimates indicate that 10 sources could be audible (meaning
+producing an \SNR\ above 8) 10 seconds or more prior to the merger during one
+year of observation.  This number is highly uncertain.  In fact as many as 100
+events detected 10 seconds early is plausible. It is also plausible that no
+sources will be detected at all.  The gray bands give the pessimistic to
+optimistic interval defined in ~\cite{Abadie:2010p10836}. 
 %
@@
 \includegraphics{figures/snr_in_time.pdf}
-\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns} sources that will
-be detectable $t$ seconds before coalescence.  The heavy solid line is the most
-likely yearly rate estimate $\dot N_{\mathrm{re}}$ during Advanced \LIGO.  The
-light solid lines represent the confidence interval $[\dot N_{\mathrm{low}}, \dot 
-N_{\mathrm{high}}]$ described in
+\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns}
+sources that will be detectable $t$ seconds before coalescence.  The heavy
+solid line is the most likely yearly rate estimate $\dot N_{\mathrm{re}}$
+during Advanced \LIGO.  The shaded region represents the confidence interval
+$[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
 \cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in
-\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for
-detection and that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a
-detector in the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be 
-detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be detectable 
-within 100 seconds of merger.  The `zero detuning, lower power' and `\textsc{bbh} optimized' configurations are still more conducive to advance detection
-but at the cost of fewer total detectable events above an \SNR\ of 8.}
+\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for detection and
+that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a detector in
+the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be
+detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be
+detectable within 100 seconds of merger.} 
 \end{center}

### 2011-06-08 snr_in_time.py
next round of comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@976 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -52,7 +52,8 @@ snrd = {}
 
 # loop over the noise curves
 # ZERO_DET_high_P.txt must be first since it normalizes the rate
-for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt', 'data/BHBH_20deg.txt'], ['zero det., high power', 'zero det., low power', 'BHBH optimized'])):
+#for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt', 'data/ZERO_DET_low_P.txt', 'data/BHBH_20deg.txt'], ['zero det., high power', 'zero det., low power', 'BHBH optimized'])):
+for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt'], ['zero det'])):
 
 	# Load the data
 	A = numpy.loadtxt(cfile)
@@ -91,12 +92,12 @@ for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt', 'data/ZERO_D
 	pylab.loglog(t, numminus, markers[i], lw=0.5)
 	pylab.loglog(t, numplus, markers[i], lw=0.5)
 	#pylab.hold(1)
-	#pylab.fill_between(t, numminus, numplus, color='k', alpha=0.15)
+	pylab.fill_between(t, numminus, numplus, color='k', alpha=0.15)
 
 pylab.grid()
-pylab.legend(loc='lower left')
-pylab.ylabel(r'$\dot N$ yr$^{-1}$')
-pylab.xlabel(r'Time before coalescence, $t$')
+#pylab.legend(loc='lower left')
+pylab.ylabel(r'Observable NS-NS systems / yr')
+pylab.xlabel(r'Time before coalescence, $t$ (s)')
 pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.08,right=0.95)
 pylab.xlim([0.01,1000])
 pylab.ylim([.1, 1000])

### 2011-06-08 conclusions.tex
next round of comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@977 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
-to a 1/(16~Hz) cadence, introducing an up-front latency~\cite{CITE_CDS}. Data
-aggregation from
-the observatories will travel over various networks, each capable of high
-bandwith but perhaps only modest latency.  This could amount to a similar
-latency of $\sim$100 ms.  Lastly, unless a real-time infrastructure is adopted
-post data acquisition, it is likely that there will be an inherent latency
-introduced by such infrastructure.  We have shown a prototype implementation
-of \lloid{} using \gstlal\ that is capable of order seconds latency. Based
-on our experience, significant work would have to be done in order to improve
-upon this number, for example, more tightly integrating analysis and data
-acquisition. This should be considered for third-generation detector design.
+to a 1/(16~Hz) cadence, introducing an up-front latency~\cite{Bork2001}. Data
+aggregation from the observatories will travel over various networks, each
+capable of high bandwith but perhaps only modest latency.  This could amount to
+a similar latency of $\sim$100 ms.  Lastly, unless a real-time infrastructure
+is adopted post data acquisition, it is likely that there will be an inherent
+latency introduced by such infrastructure.  We have shown a prototype
+implementation of \lloid{} using the open source signal processing software
+\gstreamer\ and \gstlal. Significant work would have to be done in order to
+improve upon the latency capability of our implementation, for example, more
+tightly integrating analysis and data acquisition. This should be considered
+for third-generation detector design.
 
@@
 not explored the accuracy of sky localization that is possible before merger.
-Ref.~\cite{Fairhurst2009} discusses some of the theoretical prospects for early sky
-localization.  Our future work will explore the prospects of early-warning
+Ref.~\cite{Fairhurst2009} discusses some of the theoretical prospects for early
+sky localization.  Our future work will explore the prospects of early-warning
 detection with realistic simulations of binary mergers using the infrastructure
@@
 
-\editorial{nvf: It's a bit pie-in-the-sky, but we could also mention reconfiguring
-the signal recycling mirror to optimize SNR at merger. There's a lot of science there.}
+\editorial{nvf: It's a bit pie-in-the-sky, but we could also mention
+reconfiguring the signal recycling mirror to optimize SNR at merger. There's a
+lot of science there.}
 

### 2011-06-08 results.tex
next round of comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@978 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$
-\flops\ to cover a sub-bank comprising $\sim 10^2$ out of the total $\sim
-10^5$ mass pairs.  Given that modern (ca. 2011) workstations can sustain
-computation rates up to $\sim 10^{10}$ \flops{}, and assuming that other
-regions of the parameter space have similar computational scaling\footnote{The
-lengths of templates varies over the parameter space; lower-mass templates are
-longer and would require more computation to analyze while higher-mass
-templates are shorter and would require less. However, we consider the
-estimates based on this sub-bank to be a reasonable representation.}, an entire
-search could be implemented with $\gtrsim 10$ machines.
+\flops\ to cover a sub-bank comprising $\sim 10^2$ out of the total $\sim 10^5$
+mass pairs.  Given that modern (ca. 2011) workstations can sustain computation
+rates up to $\sim 10^{10}$ \flops{}, and assuming that other regions of the
+parameter space have similar computational scaling, an entire search could be
+implemented with $\gtrsim 10$ machines.  The lengths of templates does vary
+over the parameter space; lower-mass templates are longer and would require
+more computation to analyze while higher-mass templates are shorter and would
+require less. However, we consider the estimates based on this sub-bank to be a
+reasonable representation.
 
-By comparison, using the \TD\ method to achieve the same latency costs $\sim 10^{13}$ \flops\ for this particular sub-bank, and so it would require $\sim 10^6$ present-day machines.
-Presently, the \LIGO{} Data Grid consists of only $\sim 10^4$ machines.
+By comparison, using the \TD\ method to achieve the same latency costs $\sim
+10^{13}$ \flops\ for this particular sub-bank, and so it would require $\sim
+10^6$ present-day machines to search the full parameter space.  Presently, the
+\LIGO{} Data Grid consists of only $\sim 10^4$ machines.

### 2011-06-08 method.tex
spell check
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@980 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 expand on the ideas of~\cite{Marion2004, Buskulic2010} that describe a
-multiband decomposition of the compact binary parameter space that resulted in
+multi-band decomposition of the compact binary parameter space that resulted in
 a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 joint science
@@
 that competes in terms of floating point operations per second with the
-conventional \FD\ method, which requires a signficant latency in order to be
+conventional \FD\ method, which requires a significant latency in order to be
 computationally cheap. Our method, called \lloid{} (a backronym for
@@
 templates, the early time slices have significantly lower bandwidth and can be
-safely downsampled.  The downsampling reduces the number of data samples by a
+safely down-sampled.  The down-sampling reduces the number of data samples by a
 factor of $\sim$100 from the early part of the waveform and allows the filters
@@
 For concreteness and simplicity, we will consider an inspiral waveform in the
-quadrupole approximation, for which the time-frequency relation is
+quadrapole approximation, for which the time-frequency relation is
 %
@@
 \geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time
-sliced templates may be downsampled without aliasing, so we define them as
+sliced templates may be down-sampled without aliasing, so we define them as
 %
@@
 rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
-preceeding coalescence and \slicessamps\ is the number of samples in the
+preceding coalescence and \slicessamps\ is the number of samples in the
 $s^{\mathrm{th}}$ filter.}
@@
 warning \SNR\ and to comment on the computational cost of evaluating it.  But
-first, we will introduce some notation refering to the decimation of data,
+first, we will introduce some notation referring to the decimation of data,
 which means reducing the sample rate after low-pass filtering,
@@
 \end{equation}
-and notation refering to the interpolation of data to increase the sample rate,
+and notation referring to the interpolation of data to increase the sample rate,
 \begin{equation}
@@
 The signal flow diagram in figure~\ref{fig:pipeline} illustrates how this
-recursion relation can be realized as a multirate filter network with outputs for
+recursion relation can be realized as a multi-rate filter network with outputs for
 each of the early-warning \SNR{}s.
@@
 The computational cost of the decimation of the detector data is a little bit more subtle.
-Decimation is achieved by applying an \fir\ anti-aliasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
+Decimation is achieved by applying an \fir\ anti-aliasing filter and then down-sampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
 Naively, an anti-aliasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.

### 2011-06-08 implementation.tex
spell check
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@981 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 computations that can be completed before the analysis is underway.  Thus
-we divide the procedure into two stages 1) an offline planning stage and 2) an
-online, low-latency filtering stage.  The offline stage can be done before the
+we divide the procedure into two stages 1) an off-line planning stage and 2) an
+online, low-latency filtering stage.  The off-line stage can be done before the
 analysis is started and updated asynchronously, whereas the online stage must
@@
 \gstreamer\ is useful for our purpose. \gstreamer\ also provides some useful
-stock signal processing elements such as resamplers and filters.  We have
+stock signal processing elements such as re-samplers and filters.  We have
 extended the \gstreamer\ framework by developing a library called

### 2011-06-08 conclusions.tex
spell check
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@982 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 improved upon further by conditionally reconstructing the \SNR{} time-series
-only during times when a composite dection statistic crosses a
+only during times when a composite detection statistic crosses a
 threshold~\cite{svd-compdetstat}.
@@
 aggregation from the observatories will travel over various networks, each
-capable of high bandwith but perhaps only modest latency.  This could amount to
+capable of high bandwidth but perhaps only modest latency.  This could amount to
 a similar latency of $\sim$100 ms.  Lastly, unless a real-time infrastructure

### 2011-06-08 results.tex
spell check
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@983 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
-implementations of resampling that cause signal loss.  The \SNR\ loss is to be
+implementations of re-sampling that cause signal loss.  The \SNR\ loss is to be
 compared with the mismatch of 0.03 that arises from the discreteness of the
@@
 are more complicated than in the original work~\cite{Cannon:2010p10398} due to
-the inclusion of the time-sliced templates and resampling.  Figure~%
+the inclusion of the time-sliced templates and re-sampling.  Figure~%
 \ref{fig:hist-svd-tolerance} demonstrates that it is possible to achieve
@@
 saturates with respect to the tolerance.  This could be the effect of the
-resampling, or another \SNR\ loss that we did not model or expect.  However the
+re-sampling, or another \SNR\ loss that we did not model or expect.  However the
 saturation is still an order of magnitude below our target mismatch of
 $\sim$0.003.  We find that an \SVD\ tolerance of $\left(1-10^{-4}\right)$ is
-adequate to acheive our target \SNR{} loss.
+adequate to achieve our target \SNR{} loss.
 %
@@
 		\caption{Box-and-whisker plot of mismatch between nominal
-template bank and \lloid\ measured impulse respones.  The upper and lower boundaries of
+template bank and \lloid\ measured impulse responses.  The upper and lower boundaries of
 the boxes show the upper and lower quartiles; the lines in the center denote the medians.

### 2011-06-08 method.tex
fir
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@984 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as an
-\fir\ filter, requiring $\mathcal O(\numtmps \tmpsamps)$ floating point
-operations per sample.  However, it is typically much more computationally
-efficient to use the convolution theorem and the \fft\ to implement fast
-convolution in the frequency-domain, requiring only $\mathcal O(\numtmps
-\lg \tmpsamps)$ operations per sample with increased latency.
+finite impulse response (\fir) filter, requiring $\mathcal O(\numtmps
+\tmpsamps)$ floating point operations per sample.  However, it is typically
+much more computationally efficient to use the convolution theorem and the
+\fft\ to implement fast convolution in the frequency-domain, requiring only
+$\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample with increased
+latency.
 

### 2011-06-08 introduction.tex
Made the intro rock your socks.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@985 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-As a compact binary system loses energy to gravitational waves (\GW{}s), its
+The coalescence of compact binary systems consisting of neutron stars
+ (\textsc{ns}) and/or black holes (\textsc{bh}) is presently the best
+understood and most actively pursued target source of gravitational
+radiation for ground-based gravitational wave (\GW{}) detectors.
+As a compact binary system loses energy to \GW{}s, its
 orbital separation decays, leading to a run-away inspiral with the \GW{}
-amplitude and frequency increasing until the system eventually merges soon
-after the innermost stable circular orbit (\ISCO). If a neutron star is
-involved it may become tidally disrupted near the merger.  This disrupted
-matter can fuel a bright electromagnetic counterpart in the merger's
-aftermath~\cite{shibata:2007}.
+amplitude and frequency increasing until the system eventually merges. At the
+merger, our understanding of the system's behavior is incomplete. It is thought
+that if a neutron star is involved, it may become tidally disrupted near
+the merger and fuel a bright electromagnetic counterpart ~\cite{shibata:2007}.
+This counterpart is expected to be rich in information about the source, but 
+extracting that information requires rapid identification of candidate
+gravitational wave events.
 
-\GW{} detectors are sensitive the to these \GW{}
-signals.  Indeed, the coalescence of compact binary systems consisting of
-neutron stars (\textsc{ns}) and/or black holes (\textsc{bh}) is the most
-promising source of gravitational radiation for Advanced \LIGO, Virgo, \GEO\,
-and \LCGT~\cite{ALIGOWeb, AVirgoWeb, GEOWeb, LCGTWeb}.  Tens of binary
-coalescence events per year are expected to be observed in the advanced
-detector era later this decade~\cite{Abadie:2010p10836}. In the era of
-first-generation detectors the \GW{} community initiated a project
-to send alerts when potential \GW{} transients were observed in
-order to trigger followup observations by optical telescopes.  The best
-demonstrated latencies were 30--60 minutes~\cite{HugheyGWPAW2011}, which was an
-important achievement, but far too late to catch prompt electromagnetic
-emission.
+Though compact binary coalescence (\CBC{}) events may be responsible for several
+ classes of electromagnetic transients, \CBC{}s are in particular believed to be a
+mechanism for short gamma-ray bursts (short \GRB{}s)~\cite{Lee:2005, nakar07}.
+In this scenario, prompt electromagnetic emission arises as shells of
+ relativistically out-flowing matter collide in the inner shock.
+The same inner shocks, or potentially reverse shocks, can produce a bright
+ accompanying optical flash~\cite{Sari99}. Prompt emission is a probe into
+ the extreme initial conditions of the outflow, in contrast with afterglows,
+which arise in the external shock with the local medium and are relatively
+insensitive to initial conditions.
 
-Prompt electromagnetic emission can arise as shells of relativistically
-out-flowing matter collide in the inner shock. Such an inner shock from a
-compact binary coalescence is believed to be a mechanism for short gamma-ray
-bursts (short \GRB{}s)~\cite{Lee:2005, nakar07}. The same inner shocks, or
-potentially reverse shocks, can produce a bright accompanying optical
-flash~\cite{Sari99}. Prompt emission is a probe into the extreme initial
-conditions of the outflow, in contrast with afterglows, which arise in the
-external shock with the local medium and are relatively
-insensitive to initial conditions. Optical flashes have only been observed to
-peak within tens of seconds for a handful of long
+Optical flashes typically peak in brightness within tens of seconds and
+consequently have only been observed for a handful of long
 \GRB{}s~\cite{2011CRPhy..12..255A} by telescopes with extremely rapid response
@@
 an exceptional event, perhaps even a glimpse of a
-precursor~\cite{0004-637X-723-2-1711}. An electromagnetic counterpart would
-vastly boost confidence in the \GW{} detection and provide the
-tight sky localization necessary to allow determination of the source's host
-galaxy, which leads to a redshift measurement. A coincident \GW{}
-observation will provide a luminosity distance measurement. With these
-together, we can produce precision measurements of the Hubble
-constant~\cite{2010ApJ...725..496N}.
+precursor~\cite{0004-637X-723-2-1711}.
 
-To this end, we have the ambition of reporting \GW{} candidates not minutes
-\emph{after} the merger, as has already been accomplished, but seconds
-\emph{before}.  By examining the signal to noise ratio (\SNR{}) stream for
+Observation of an electromagnetic counterpart to a candidate \GW{} event would
+vastly boost confidence in the \GW{} detection and provide the best possible
+estimation of the event's positon in the sky localization. With such detailed
+information, it will be possible to determine the source's host
+galaxy and therefore measure the source's redshift. A coincident \GW{}
+observation of this event will also provide a luminosity distance measurement.
+Together, these two quantities immediately lead to measurements of the Hubble
+constant of unprecedented accuracy ~\cite{2010ApJ...725..496N}.
+
+In the era of first-generation detectors, the \GW{} community initiated a project
+to send alerts when potential \GW{} transients were observed in
+order to trigger followup observations by optical telescopes.  The best
+demonstrated latencies were 30--60 minutes~\cite{HugheyGWPAW2011}, which was an
+important achievement, but far too late to catch prompt electromagnetic
+emission. In order to extract the maximal science from these events, we have 
+the ambition of reporting \GW{} candidates not minutes \emph{after} the merger, 
+as has already been accomplished, but seconds \emph{before}.
+
+We show below that by examining the signal to noise ratio (\SNR{}) stream for
 threshold crossings before the \GW{} signal leaves the detection band, it is

### 2011-06-09 conclusions.tex
lloid_paper: polishing pass post-Steve
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@986 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 detecting an electromagnetic counterpart in bands from gamma-rays down to
-radio.  We anticipate requiring $\sim100$ modern multi-core computers to
+radio.  We anticipate requiring $\sim$100 modern multi-core computers to
 analyze a four-detector network of \GW{} data for binary neutron

### 2011-06-09 introduction.tex
lloid_paper: polishing pass post-Steve
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@986 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 understood and most actively pursued target source of gravitational
-radiation for ground-based gravitational wave (\GW{}) detectors.
+radiation for ground-based gravitational-wave (\GW{}) detectors.
 As a compact binary system loses energy to \GW{}s, its
 orbital separation decays, leading to a run-away inspiral with the \GW{}
-amplitude and frequency increasing until the system eventually merges. At the
-merger, our understanding of the system's behavior is incomplete. It is thought
-that if a neutron star is involved, it may become tidally disrupted near
-the merger and fuel a bright electromagnetic counterpart ~\cite{shibata:2007}.
-This counterpart is expected to be rich in information about the source, but 
-extracting that information requires rapid identification of candidate
-gravitational wave events.
+amplitude and frequency increasing until the system eventually merges.  It is
+thought that if a \textsc{ns} is involved, it may become tidally disrupted near
+the merger and fuel a bright electromagnetic (\EM{})
+counterpart~\cite{shibata:2007}. Observation of an \EM{} counterpart to a
+candidate \GW{} event will vastly boost confidence in the \GW{} detection and
+will often provide sufficient sky localization to determine
+the source's host galaxy and therefore measure the source's redshift. The
+\GW{} observation of this event will also provide a luminosity
+distance measurement. Together, these two quantities immediately lead to
+measurements of the Hubble constant of unprecedented
+accuracy~\cite{2010ApJ...725..496N}. The most efficient means of obtaining
+such joint observations is through the rapid identification of candidate \GW{}
+events.
 
 Though compact binary coalescence (\CBC{}) events may be responsible for several
- classes of electromagnetic transients, \CBC{}s are in particular believed to be a
+classes of \EM{} transients, \CBC{}s are in particular believed to be a
 mechanism for short gamma-ray bursts (short \GRB{}s)~\cite{Lee:2005, nakar07}.
-In this scenario, prompt electromagnetic emission arises as shells of
+In this scenario, prompt \EM{} emission arises as shells of
 relativistically out-flowing matter collide in the inner shock.
@@
 
-Optical flashes typically peak in brightness within tens of seconds and
-consequently have only been observed for a handful of long
+Optical flashes have only been observed for a handful of long
 \GRB{}s~\cite{2011CRPhy..12..255A} by telescopes with extremely rapid response
@@
 telescopes were observing a previous \GRB{} in the same
-field~\cite{2008Natur.455..183R}. Short \GRB{}s, on the other hand, typically
-fade too quickly to even catch the tails of the afterglows. Rapid
+field~\cite{2008Natur.455..183R}. The observed optical flashes peaked
+within tens of seconds and decayed quickly.
+Short \GRB{}s, on the other hand, typically
+fade too quickly to even catch the tails of the afterglows in any band. Rapid
 \GW{} transient alerts could enable the first observation of
@@
 
-Observation of an electromagnetic counterpart to a candidate \GW{} event would
-vastly boost confidence in the \GW{} detection and provide the best possible
-estimation of the event's positon in the sky localization. With such detailed
-information, it will be possible to determine the source's host
-galaxy and therefore measure the source's redshift. A coincident \GW{}
-observation of this event will also provide a luminosity distance measurement.
-Together, these two quantities immediately lead to measurements of the Hubble
-constant of unprecedented accuracy ~\cite{2010ApJ...725..496N}.
+In the era of first-generation detectors, the \GW{} community initiated a
+project to send alerts when potential \GW{} transients were observed in
+order to trigger followup observations by optical telescopes.  The best
+demonstrated latencies were
+$\mathcal{O}(1\textrm{ hour})$~\cite{HugheyGWPAW2011}, which was an important
+achievement, but far too late to catch prompt \EM{}
+emission. In order to extract the maximum science from Advanced \LIGO{}
+\CBC{} events, we have the ambition of reporting \GW{} candidates not minutes
+\emph{after} the merger, as has already been accomplished, but seconds
+\emph{before}.
 
-In the era of first-generation detectors, the \GW{} community initiated a project
-to send alerts when potential \GW{} transients were observed in
-order to trigger followup observations by optical telescopes.  The best
-demonstrated latencies were 30--60 minutes~\cite{HugheyGWPAW2011}, which was an
-important achievement, but far too late to catch prompt electromagnetic
-emission. In order to extract the maximal science from these events, we have 
-the ambition of reporting \GW{} candidates not minutes \emph{after} the merger, 
-as has already been accomplished, but seconds \emph{before}.
-
-We show below that by examining the signal to noise ratio (\SNR{}) stream for
+We show that by examining the signal to noise ratio (\SNR{}) stream for
 threshold crossings before the \GW{} signal leaves the detection band, it is
 possible to trade some \SNR{} and sky localization
-accuracy~\cite{Fairhurst2009} for latency.  Figure \ref{fig:earlywarning} shows
+accuracy for negative latency.  Figure~\ref{fig:earlywarning} shows
 projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
 \LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.  The
-most likely estimates indicate that 10 sources could be audible (meaning
-producing an \SNR\ above 8) 10 seconds or more prior to the merger during one
-year of observation.  This number is highly uncertain.  In fact as many as 100
-events detected 10 seconds early is plausible. It is also plausible that no
-sources will be detected at all.  The gray bands give the pessimistic to
-optimistic interval defined in ~\cite{Abadie:2010p10836}. 
+most likely estimates indicate that $\sim$10 sources will produce an \SNR\
+above 8 in the detector 10 seconds or more prior to the merger during one
+year of observation, though the uncertainty spans orders of magnitude.
+The gray bands give the pessimistic to optimistic interval defined
+in~\cite{Abadie:2010p10836}.
 %
@@
 that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a detector in
-the `zero detuning, high power' configuration, about 10~yr$^{-1}$ will be
-detectable within 10 seconds before merger and $\sim1$~yr$^{-1}$ will be
+the `zero detuning, high power' configuration, $\sim$10~yr$^{-1}$ will be
+detectable within 10 seconds before merger and $\sim$1~yr$^{-1}$ will be
 detectable within 100 seconds of merger.} 
@@
 achieved latencies of 30--60 minutes, which were dominated by a human vetting
-process. Candidates were sent for electromagnetic followup to several
+process. Candidates were sent for \EM{} followup to several
 telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008,
@@
 
-Advance detection of compact binary coalescences (\CBC{}s) will require striking a balance between latency
-and throughput.  \CBC{} searches consist of banks of matched filters, or
+Realizing advance detection of compact binary coalescences (\CBC{}s) will
+require striking a balance between latency and throughput. \CBC{} searches
+consist of banks of matched filters, or
 cross-correlations between the data stream and a bank of nominal ``template''

### 2011-06-09 results.tex
lloid_paper: polishing pass post-Steve
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@986 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Table~\ref{table:flops} shows that the \lloid\ method requires $\sim 10^8$
-\flops\ to cover a sub-bank comprising $\sim 10^2$ out of the total $\sim 10^5$
+Table~\ref{table:flops} shows that the \lloid\ method requires $\sim$$10^8$
+\flops\ to cover a sub-bank comprising $\sim$$10^2$ out of the total $\sim$$10^5$
 mass pairs.  Given that modern (ca. 2011) workstations can sustain computation
-rates up to $\sim 10^{10}$ \flops{}, and assuming that other regions of the
+rates up to $\sim$$10^{10}$ \flops{}, and assuming that other regions of the
 parameter space have similar computational scaling, an entire search could be
-implemented with $\gtrsim 10$ machines.  The lengths of templates does vary
+implemented with $\gtrsim$$10$ machines.  The lengths of templates does vary
 over the parameter space; lower-mass templates are longer and would require
@@
 
-By comparison, using the \TD\ method to achieve the same latency costs $\sim
-10^{13}$ \flops\ for this particular sub-bank, and so it would require $\sim
-10^6$ present-day machines to search the full parameter space.  Presently, the
-\LIGO{} Data Grid consists of only $\sim 10^4$ machines.
+By comparison, using the \TD\ method to achieve the same latency costs
+$\sim$$10^{13}$ \flops\ for this particular sub-bank, and so it would require
+$\sim$$10^6$ present-day machines to search the full parameter space.
+Presently, the \LIGO{} Data Grid consists of only $\sim$$10^4$ machines.

### 2011-06-09 introduction.tex
lloid_paper: one more reordering
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@987 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed improper use of big-O notation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@988 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+In the era of first-generation detectors, the \GW{} community initiated a
+project to send alerts when potential \GW{} transients were observed in
+order to trigger followup observations by optical telescopes.  The best
+demonstrated latencies were on the order of 1 hour%
+~\cite{HugheyGWPAW2011}, which was an important
+achievement, but far too late to catch prompt \EM{}
+emission. In order to extract the maximum science from Advanced \LIGO{}
+\CBC{} events, we have the ambition of reporting \GW{} candidates not minutes
+\emph{after} the merger, as has already been accomplished, but seconds
+\emph{before}.
+
 Though compact binary coalescence (\CBC{}) events may be responsible for several
@@
 precursor~\cite{0004-637X-723-2-1711}.
-
-In the era of first-generation detectors, the \GW{} community initiated a
-project to send alerts when potential \GW{} transients were observed in
-order to trigger followup observations by optical telescopes.  The best
-demonstrated latencies were
-$\mathcal{O}(1\textrm{ hour})$~\cite{HugheyGWPAW2011}, which was an important
-achievement, but far too late to catch prompt \EM{}
-emission. In order to extract the maximum science from Advanced \LIGO{}
-\CBC{} events, we have the ambition of reporting \GW{} candidates not minutes
-\emph{after} the merger, as has already been accomplished, but seconds
-\emph{before}.
 

### 2011-06-09 introduction.tex
Added 'dissemination'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@989 ccdd05d5-56c1-461a-bf98-85f6876523b4
Reworded funny-sounding sentence.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@991 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \GW{} observation of this event will also provide a luminosity
-distance measurement. Together, these two quantities immediately lead to
-measurements of the Hubble constant of unprecedented
+distance measurement. With both quantities, we can
+immediately measure the Hubble constant with unprecedented
 accuracy~\cite{2010ApJ...725..496N}. The most efficient means of obtaining
-such joint observations is through the rapid identification of candidate \GW{}
-events.
+such joint observations is through the rapid identification and dissemination
+of candidate \GW{} events.
 

### 2011-06-09 introduction.tex
Made description of confidence intervals more quantitative
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@992 ccdd05d5-56c1-461a-bf98-85f6876523b4
Polished caption for Figure 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@993 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
-\LIGO\ assuming the event rate predictions in~\cite{Abadie:2010p10836}.  The
+\LIGO\ assuming event rates predicted in~\cite{Abadie:2010p10836} and anticipated
+detector sensitivity described in \cite{ALIGONoise}.  The
 most likely estimates indicate that $\sim$10 sources will produce an \SNR\
@@
 year of observation, though the uncertainty spans orders of magnitude.
-The gray bands give the pessimistic to optimistic interval defined
+The gray bands give the 5 to 95\% confidence interval as described
 in~\cite{Abadie:2010p10836}.
@@
 $[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
-\cite{Abadie:2010p10836}.  Advanced \LIGO\ noise projections are described in
-\cite{ALIGONoise}.  Assuming that an \SNR\ of 8 is sufficient for detection and
-that we observe $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$ with a detector in
-the `zero detuning, high power' configuration, $\sim$10~yr$^{-1}$ will be
-detectable within 10 seconds before merger and $\sim$1~yr$^{-1}$ will be
-detectable within 100 seconds of merger.} 
+\cite{Abadie:2010p10836}.  With an advanced detector in the `zero detuning,
+high power' configuration \cite{ALIGONoise} and an \SNR\ threshold of 8,
+we will observe a total of $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$, but
+$\sim$10~yr$^{-1}$ will be detectable within 10~s of merger and
+$\sim$1~yr$^{-1}$ within 100~s.} 
 \end{center}

### 2011-06-09 implementation.tex
Converted some software citations to links
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@995 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We have implemented a prototype of the low-latency filtering stage using an
-open-source signal processing environment called \gstreamer\ \cite{gstreamer}.
+open-source signal processing environment called
+\gstreamer\footnote{\url{http://gstreamer.net/}}.
 \gstreamer\ is a vital component of many Linux systems, providing media
@@
 extended the \gstreamer\ framework by developing a library called
-\gstlal{}~\cite{gstlal} that provides elements for \GW{} data
-analysis.
+\gstlal\footnote{\url{https://www.lsc-group.phys.uwm.edu/daswg/projects/gstlal.html}}
+that provides elements for \GW{} data analysis.
 

### 2011-06-09 results.tex
Converted some software citations to links
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@995 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
-curve~\cite{lal}.  Then we create sub-banks by partitioning the parameter space by
-chirp mass.  Figure \ref{fig:tmpltbank} illustrates this procedure.
+curve~\cite{ALIGONoise}.  Then we create sub-banks by partitioning the parameter
+space by chirp mass.  Figure \ref{fig:tmpltbank} illustrates this procedure.
 \begin{figure}[htbp]

### 2011-06-09 packages/astronat/nat2jour.pl
Added astronat
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@996 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -1,0 +1,374 @@
+#!/usr/local/bin/perl -- # -*-Perl-*-
+
+######################################################################
+#
+# nat2jour.pl
+#
+# Program for converting tex/bbl files generated using the natbib
+# (v5.3) BibTeX style into files which may be submitted electronically
+# to astronomy journals.
+#
+# Written by Jonathan Baker <jbaker@astro.berkeley.edu>
+# inspired by 2apj.pl code by Eddie Baron
+#
+# $Log: nat2jour.pl,v $
+# Revision 1.6  2000/04/21 17:48:02  jbaker
+# Changed addition of space before \cite's so that it only happens when
+# \cite is preceeded by a \... command (so that it doesn't happen with
+# ~\cite, for example, as suggested by Norman Gray).
+#
+# Revision 1.5  2000/04/20 22:16:26  jbaker
+# If using -references, insert a space before \cite in case it is
+# terminating a command (like \protect, for example).
+#
+# Revision 1.4  1999/09/01 02:45:19  jbaker
+# Added flag to generate `references' environment instead of `thebibliography'.
+#
+# Revision 1.3  1998/08/30 19:48:32  jbaker
+# Added RCS keywords.  Non-MNRAS lettering style changed from "1998a,b" to
+# "1998a, 1998b".
+#
+#
+######################################################################
+
+$MaxAuths = 8;
+$inline = 0;
+$refenv = 0; $bibenv = "thebibliography";
+$etal = "{et al.}";
+
+$usage = <<EOT;
+
+Usage: $0 [options] INPUT [OUTPUT]
+
+$Id: nat2jour.pl,v 1.6 2000/04/21 17:48:02 jbaker Exp $
+
+Convert BibTeX/Natbib tex/bbl files into files which may be submitted
+to astronomy journals.  Inlines citations adding \\markcite or
+\\nocite and formats \\bibitems and removes natbib package dependencies,
+stripping "%nat" comments.  Also implements astronomy 3-author citation
+style and truncates long author lists.
+
+Arguments:
+  INPUT        If this is "foo", input files are "foo.tex" and "foo.bbl"
+  OUTPUT       (optional) Same for output files; default appends "-aas".
+
+Options:
+  -inline      Inline the bibliography into a single LaTeX file
[…]

### 2011-06-09 conclusions.tex
Converted to AASTeX
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@997 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \section{Conclusions}
-\label{SECV}\label{sec:conclusions}
+\label{sec:conclusions}
 
@@
 stars and stellar-mass black holes.  This is well within the current computing
-capabilities of the \LIGO{} Data Grid~\cite{LDG}. In the future, this could be
+capabilities of the \LIGO{} Data Grid~\citep{LDG}. In the future, this could be
 improved upon further by conditionally reconstructing the \SNR{} time-series
 only during times when a composite detection statistic crosses a
-threshold~\cite{svd-compdetstat}.
+threshold~\citep{svd-compdetstat}.
 
@@
 procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
-to a 1/(16~Hz) cadence, introducing an up-front latency~\cite{Bork2001}. Data
+to a 1/(16~Hz) cadence, introducing an up-front latency~\citep{Bork2001}. Data
 aggregation from the observatories will travel over various networks, each
@@
 not explored the accuracy of sky localization that is possible before merger.
-Ref.~\cite{Fairhurst2009} discusses some of the theoretical prospects for early
+Ref.~\citep{Fairhurst2009} discusses some of the theoretical prospects for early
 sky localization.  Our future work will explore the prospects of early-warning

### 2011-06-09 implementation.tex
Converted to AASTeX
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@997 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The planning stage begins with choosing templates that cover the space of
-source parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to
+source parameters with a hexagonal grid~\citep{PhysRevD.76.102004} in order to
 satisfy a minimum match criterion.  This assures a prescribed maximum loss in
@@
 \SVD{} and is the approach considered
-in~\cite{Cannon:2010p10398}.  We choose time-slice boundaries as in
+in~\citep{Cannon:2010p10398}.  We choose time-slice boundaries as in
 equation~\eqref{eq:time-sliced-templates} such that all of the templates within

### 2011-06-09 inspiral_svd.tex
added funding ack
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@957 ccdd05d5-56c1-461a-bf98-85f6876523b4
Updates
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@971 ccdd05d5-56c1-461a-bf98-85f6876523b4
Converted to AASTeX
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@997 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\documentclass{iopart}
+\documentclass[preprint2]{aastex}
 \input{packages}
 \input{macros}
-\renewcommand{\arraystretch}{1.2}
 
@@
 
-\title[Early-warning detection of compact binary coalescence]{Technique for \earlywarning\ detection of compact binary
-coalescence and its implications for multi-messenger astronomy}
+\title{Technique for \earlywarning\ detection of compact binary coalescence
+and its implications for multi-messenger astronomy}
 
-\date{\today}
+\shorttitle{Early-warning of compact binary coalescence}
 
-\author{Kipp Cannon$^{1}$, Romain Cariou$^{2}$, Adrian Chapman$^3$, Mireia
-Crisp\'{i}n-Ortuzar$^4$, Nickolas Fotopoulos$^3$, Melissa Frei $^5$, Chad
-Hanna$^{6}$, Erin Kara$^{7}$, Drew Keppel$^{8,9}$, Laura Liao$^{10}$, Stephen
-Privitera $^3$, Antony C Searle$^3$, Leo Singer$^3$ and Alan J  Weinstein$^3$}
+\author{
+	Kipp Cannon\altaffilmark{1},
+	Romain Cariou\altaffilmark{2},
+	Adrian Chapman\altaffilmark{3},
+	Mireia Crisp\'{i}n-Ortuzar\altaffilmark{4},
+	Nickolas Fotopoulos\altaffilmark{3},
+	Melissa Frei\altaffilmark{5},
+	Chad Hanna\altaffilmark{6},
+	Erin Kara\altaffilmark{7},
+	Drew Keppel\altaffilmark{8,9},
+	Laura Liao\altaffilmark{10},
+	Stephen Privitera\altaffilmark{3},
+	Antony Searle\altaffilmark{3},
+	Leo Singer\altaffilmark{3}, and
+	Alan Weinstein\altaffilmark{3}
+}
 
-\address{$^1$ Canadian Institute for Theoretical Astrophysics, Toronto, ON,
-Canada}
-\address{$^2$  D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
-\address{$^3$ LIGO Laboratory - California Institute of Technology, Pasadena,
-CA, USA} 
-\address{$^4$ Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot,
-Spain} 
-\address{$^5$ The University of Texas at Austin, Austin, TX, USA}
-\address{$^6$ Perimeter Institute for Theoretical Physics, Waterloo, ON,
-Canada} 
-\address{$^7$ Department of Physics and Astronomy, Barnard College, Columbia
-University, New York, NY, USA} 
-\address{$^8$ Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r
-Gravitationphysik, Hannover, Germany} 
-\address{$^9$ Leibniz Universit\"{a}t Hannover, Hannover, Germany}
-\address{$^{10}$ Ryerson University, Toronto, ON, Canada}
+\altaffiltext{1}{Canadian Institute for Theoretical Astrophysics, Toronto, ON, Canada}
+\altaffiltext{2}{D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
+\altaffiltext{3}{LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA} 
+\altaffiltext{4}{Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot, Spain} 
+\altaffiltext{5}{University of Texas at Austin, Austin, TX, USA}
+\altaffiltext{6}{Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
+\altaffiltext{7}{Department of Physics and Astronomy, Barnard College, Columbia University, New York, NY, USA}
+\altaffiltext{8}{Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
+\altaffiltext{9}{Leibniz Universit\"{a}t Hannover, Hannover, Germany}
+\altaffiltext{10}{Ryerson University, Toronto, ON, Canada}
+
+\keywords{gamma ray burst: general --- gravitational waves --- methods: data analysis --- methods: numerical}
 
@@
 
-\pacs{04.30.-w, 07.05.Kf}
-
-%\preprint{}
 \input{introduction}
-
 \input{method}
-
 \input{implementation}
-
 \input{results}
-
 \input{conclusions}
 
-\ack{\LIGO\ was constructed by the California Institute of Technology and
+\acknowledgements
+
+\LIGO\ was constructed by the California Institute of Technology and
 Massachusetts Institute of Technology with funding from the National Science
@@
 about low latency analysis methods.  NF would like to thank Alessandra Corsi
-for illuminating discussions on astronomical motivations.  This research is
-supported by the National Science Foundation through a Graduate Research
-Fellowship to LS.
+for illuminating discussions on astronomical motivations.  DK was supported
+from the Max Planck Gesellschaft. This research is supported by the National
+Science Foundation through a Graduate Research Fellowship to LS and by a
+fellowship from The Perimeter Institute for Theoretical Physics to CH.
 
-This paper has \LIGO\ Document
-Number \LIGO-\textsc{p}\oldstylenums{0900004}.}
+This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}.
 
-%\appendix
-%\input{appendix}
-
-\section*{References}
-
-\bibliographystyle{iopart-num}
+\bibliographystyle{apj}
 \bibliography{references}

### 2011-06-09 introduction.tex
Converted to AASTeX
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@997 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the merger and fuel a bright electromagnetic (\EM{})
-counterpart~\cite{shibata:2007}. Observation of an \EM{} counterpart to a
+counterpart~\citep{shibata:2007}. Observation of an \EM{} counterpart to a
 candidate \GW{} event will vastly boost confidence in the \GW{} detection and
@@
 immediately measure the Hubble constant with unprecedented
-accuracy~\cite{2010ApJ...725..496N}. The most efficient means of obtaining
+accuracy~\citep{2010ApJ...725..496N}. The most efficient means of obtaining
 such joint observations is through the rapid identification and dissemination
@@
 demonstrated latencies were on the order of 1 hour%
-~\cite{HugheyGWPAW2011}, which was an important
+~\citep{HugheyGWPAW2011}, which was an important
 achievement, but far too late to catch prompt \EM{}
@@
 classes of \EM{} transients, \CBC{}s are in particular believed to be a
-mechanism for short gamma-ray bursts (short \GRB{}s)~\cite{Lee:2005, nakar07}.
+mechanism for short gamma-ray bursts (short \GRB{}s)~\citep{Lee:2005, nakar07}.
 In this scenario, prompt \EM{} emission arises as shells of
@@
 The same inner shocks, or potentially reverse shocks, can produce a bright
-accompanying optical flash~\cite{Sari99}. Prompt emission is a probe into
+accompanying optical flash~\citep{Sari99}. Prompt emission is a probe into
 the extreme initial conditions of the outflow, in contrast with afterglows,
@@
 Optical flashes have only been observed for a handful of long
-\GRB{}s~\cite{2011CRPhy..12..255A} by telescopes with extremely rapid response
+\GRB{}s~\citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
 or, in the case of \textsc{grb 080319b}, by pure serendipity, where several
 telescopes were observing a previous \GRB{} in the same
-field~\cite{2008Natur.455..183R}. The observed optical flashes peaked
+field~\citep{2008Natur.455..183R}. The observed optical flashes peaked
 within tens of seconds and decayed quickly.
@@
 an exceptional event, perhaps even a glimpse of a
-precursor~\cite{0004-637X-723-2-1711}.
+precursor~\citep{0004-637X-723-2-1711}.
 
@@
 projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
-\LIGO\ assuming event rates predicted in~\cite{Abadie:2010p10836} and anticipated
-detector sensitivity described in \cite{ALIGONoise}.  The
+\LIGO\ assuming event rates predicted in~\citep{Abadie:2010p10836} and anticipated
+detector sensitivity described in \citep{ALIGONoise}.  The
 most likely estimates indicate that $\sim$10 sources will produce an \SNR\
@@
 The gray bands give the 5 to 95\% confidence interval as described
-in~\cite{Abadie:2010p10836}.
+in~\citep{Abadie:2010p10836}.
 %
-\begin{figure}
-\begin{center}
-\includegraphics{figures/snr_in_time.pdf}
+\begin{figure}[h]
+\plotone{figures/snr_in_time.pdf}
 \caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns}
@@
 $[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
-\cite{Abadie:2010p10836}.  With an advanced detector in the `zero detuning,
-high power' configuration \cite{ALIGONoise} and an \SNR\ threshold of 8,
+\citep{Abadie:2010p10836}.  With an advanced detector in the `zero detuning,
+high power' configuration \citep{ALIGONoise} and an \SNR\ threshold of 8,
 we will observe a total of $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$, but
@@
 $\sim$1~yr$^{-1}$ within 100~s.}
-\end{center}
 \end{figure}
@@
 namely \textsc{mbta}, Coherent WaveBurst, and
-Omega~\cite{HugheyGWPAW2011, S6lowlatency}.
-\editorial{Get references for these low-latency pipelines.} The S6 analyses
+Omega~\citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
 achieved latencies of 30--60 minutes, which were dominated by a human vetting
 process. Candidates were sent for \EM{} followup to several
-telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko~\cite{kanner2008,
+telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko~\citep{kanner2008,
 HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved
@@
 low throughput.  The former are epitomized by the overlap-save algorithm
-\cite{numerical-recipes-chapter-13} for frequency domain (\FD) convolution,
+\citep{numerical-recipes-chapter-13} for frequency domain (\FD) convolution,
 currently the preferred method in \GW{}
@@
 methods such as singular value decomposition (\SVD{}) to reduce the number of
-templates \cite{Cannon:2010p10398}. We will use both aspects to demonstrate that a very
+templates \citep{Cannon:2010p10398}. We will use both aspects to demonstrate that a very
 low latency analysis with advance detection of compact binary sources is

### 2011-06-09 method.tex
Converted to AASTeX
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@997 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \section{Early warning searches for compact binary coalescence}
-\label{SECII}\label{sec:method}
+\label{sec:method}
 
@@
 possibility of \earlywarning\ detection with modest computing requirements.  We
-expand on the ideas of~\cite{Marion2004, Buskulic2010} that describe a
+expand on the ideas of~\citep{Marion2004, Buskulic2010} that describe a
 multi-band decomposition of the compact binary parameter space that resulted in
 a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 joint science
-runs~\cite{HugheyGWPAW2011}.  We combine this with the \SVD\ rank-reduction
-method of~\cite{Cannon:2010p10398} that exploits the redundancy of
+runs~\citep{HugheyGWPAW2011}.  We combine this with the \SVD\ rank-reduction
+method of~\citep{Cannon:2010p10398} that exploits the redundancy of
 the template banks.
@@
 Inspiral signals are continuously parameterized by a set of intrinsic source
-parameters $\vec\theta$ that determine the amplitude and phase evolution of the
+parameters $\mathbf\Theta$ that determine the amplitude and phase evolution of the
 \GW{} strain. For systems where the effects of spin can be
 ignored, the intrinsic source parameters are the component masses of the
-binary, $\vec\theta = (m_1, m_2)$. Searches for inspiral signals typically
+binary, $\mathbf\Theta = (m_1, m_2)$. Searches for inspiral signals typically
 employ matched filter banks that discretely sample the possible intrinsic
-parameters~\cite{findchirppaper}.  For a given source, the strain observed by
+parameters~\citep{findchirppaper}.  For a given source, the strain observed by
 the detector is a linear combination of two waveforms corresponding to the
 `$+$' and `$\times$' \GW{} polarizations.  Thus, for any value of
-$\vec\theta$ we must implement 2 filters.  The coefficients for the $\numtmps$ filters
+$\mathbf\Theta$ we must implement 2 filters.  The coefficients for the $\numtmps$ filters
 are known as templates, and are formed by discretizing and time reversing the
@@
 detector's noise. To construct a template bank, templates are chosen with the
-$\numtmps/2$ discrete signal parameters $\vec\theta_0,\, \vec\theta_1,\, \dots,\,
-\vec\theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~\cite{Owen:1998dk}. That
+$\numtmps/2$ discrete signal parameters $\mathbf\Theta_0,\, \mathbf\Theta_1,\, \dots,\,
+\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~\citep{Owen:1998dk}. That
 is, any possible signal within a given range of intrinsic
@@
 We use the \SVD{} to produce an orthogonal filter set
-from the time-sliced templates~\cite{Cannon:2010p10398}.  We find that this
+from the time-sliced templates~\citep{Cannon:2010p10398}.  We find that this
 reduces the number of sample points of the filters by another factor of
@@
 slices in a \TD{} analogue to the \FD{} decomposition described
-in~\cite{Marion2004, Buskulic2010}.  To wit,
+in~\citep{Marion2004, Buskulic2010}.  To wit,
 %
@@
 h_{i}[k] = \sum_{s=0}^{S-1}
-	\cases{%
-		h_i^s[k] & for $t^s \leqslant k / f^0 < t^{s+1}$ \\
-		0 & otherwise
-	}
+	\begin{cases}
+		h_i^s[k] & \textrm{if } t^s \leqslant k / f^0 < t^{s+1} \\
+		0 & \textrm{otherwise}
+	\end{cases}
 \end{equation}
@@
 Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
-M_\odot / c^3 \approx 5 \umu\mathrm{s}$) and $t$ is the time relative to the
-coalescence of the binary~\cite{findchirppaper, kidder1992}.  Usually the
+M_\odot / c^3 \approx 5 \upmu\mathrm{s}$) and $t$ is the time relative to the
+coalescence of the binary~\citep{findchirppaper, kidder1992}.  Usually the
 template is truncated at some prescribed time $t^0$, or equivalently frequency
@@
 h_{i}^{s}[k] \equiv
-	\cases{
-		h_{i}\!\left[k\frac{f}{f^s}\right] & if $t^s \leqslant k/f^s < t^{s+1}$ \\
-		0 & otherwise.
-	}
+	\begin{cases}
+		h_{i}\!\left[k\frac{f}{f^s}\right] & \textrm{if } t^s \leqslant k/f^s < t^{s+1} \\
+		0 & \textrm{otherwise.}
+	\end{cases}
 \end{equation}
@@
 
-Since waveforms with neighboring intrinsic source parameters $\vec\theta$ have
+Since waveforms with neighboring intrinsic source parameters $\mathbf\Theta$ have
 similar time-frequency evolution, it is possible to design computationally
@@
 basis templates}, previously demonstrated
-in~\cite{Cannon:2010p10398}.  Similarly, the time-sliced templates described
+in~\citep{Cannon:2010p10398}.  Similarly, the time-sliced templates described
 above can be approximated to arbitrary accuracy by expansion into a set of
@@
 the approximation.  This determines the \SVD\ tolerance, which affects the
-\SNR\ loss due to the approximation.  The authors of \cite{Cannon:2010p10398}
+\SNR\ loss due to the approximation.  The authors of \citep{Cannon:2010p10398}
 showed that high accuracy could be achieved with far fewer basis templates than
@@
 smaller than the original number of templates \numtmps\ and improves on
-\cite{Cannon:2010p10398} by nearly an order of magnitude.  In the next section
+\citep{Cannon:2010p10398} by nearly an order of magnitude.  In the next section
 we describe how we form our early-warning detection statistic using the time-slice
@@
 which means reducing the sample rate after low-pass filtering,
-\begin{equation}
+\begin{equation*}
 \label{eq:decomp}
 	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k] \,,
-\end{equation}
+\end{equation*}
 and notation referring to the interpolation of data to increase the sample rate,
-\begin{equation}
+\begin{equation*}
 	x^{s}[k] = \left( H^\shortuparrow x^{s+1}\right)[k] \,.
-\end{equation}
+\end{equation*}
 
@@
 %
-\begin{figure}[htbp]
-	\includegraphics{figures/lloid-diagram.pdf}
+\begin{figure}[h]
+	\plotone{figures/lloid-diagram.pdf}
 	\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
@@
 %
-\Table{\label{tab:recap} Notation used to describe filters.}
-\br
+\begin{table}
+\caption{\label{tab:recap}Notation used to describe filters.}
+\begin{tabular}{ll}
+\tableline\tableline
 & Definition \\
-\mr
+\tableline
 \numtmps		& number of templates \\
@@
 $N^\shortuparrow$ & number of coefficients in interpolation filter \\
-\br
+\tableline
 \end{tabular}
-\end{indented}
 \end{table}
-
-\editorial{This section's two tables are small.  Any way to put them side by side or next to another figure?}
 
@@
 
-The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in \cite{numerical-recipes-chapter-13}.
+The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in \citep{numerical-recipes-chapter-13}.
 It entails splitting the input into blocks of $D$ samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$ samples.
@@
 
-Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about $2 \fftblock \lg \fftblock$ operations to evaluate a real transform of size $\fftblock$~\cite{Johnson:2007p9654}.
+Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about $2 \fftblock \lg \fftblock$ operations to evaluate a real transform of size $\fftblock$~\citep{Johnson:2007p9654}.
 Including the forward transform of the data and $M$ reverse transforms for all of the templates, the \fft\ costs $2 (\numtmps + 1) \fftblock \lg \fftblock$ operations per block.

### 2011-06-09 results.tex
Converted to AASTeX
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@997 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-09 results.tex
@@
 \section{Results}
-\label{SECIV}\label{sec:results}
+\label{sec:results}
 
@@
 between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
-curve~\cite{lal}.  Then we create sub-banks by partitioning the parameter space by
-chirp mass.  Figure \ref{fig:tmpltbank} illustrates this procedure.
-\begin{figure}[htbp]
-	\begin{center}
-		\includegraphics{figures/tmpltbank.png}
+curve~\citep{ALIGONoise}.  Then we create sub-banks by partitioning the parameter
+space by chirp mass.  Figure \ref{fig:tmpltbank} illustrates this procedure.
+\begin{figure}[h]
+	\plotone{figures/tmpltbank.png}
 	\caption{\label{fig:tmpltbank}Placement of template parameters used in this paper.
@@
 We design a filter bank to search for a subset of 657 of these templates with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
-	\end{center}
 \end{figure}
@@
 \begin{table}
-\begin{indented}
 \caption{\label{tab:time_slices} Filter design for these 657 templates.
@@
 We vary \SVD{} tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
-\item[]\begin{tabular}{rr@{,\,}lc*{6}{r}}
-\br
-\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\
-\ns
-\multicolumn{4}{c}{} &\crule{6} \\
+\begin{tabular}{rr@{,\,}lc*{6}{r}}
+\tableline\tableline
+\multicolumn{4}{c}{} &\multicolumn{6}{c}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\
+%\multicolumn{4}{c}{} &\tableline \\
 $f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\
-\mr
+\tableline
 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
@@
 32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
-\br
+\tableline
 \end{tabular}
-\end{indented}
 \end{table}
@@
 the truncation of the \SVD\ basis and estimates for it
-exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
+exist~\citep{Cannon:2010p10398}.  The second comes from non ideal
 implementations of re-sampling that cause signal loss.  The \SNR\ loss is to be
@@
 In this section we check the effect of the \SVD\ tolerance parameter defined
-in~\cite{Cannon:2010p10398}.  By changing the tolerance of the reconstruction
+in~\citep{Cannon:2010p10398}.  By changing the tolerance of the reconstruction
 of the physical waveforms one directly affects the mismatch that results from
 the truncation of the orthogonal filter matrix.  The conditions presented here
-are more complicated than in the original work~\cite{Cannon:2010p10398} due to
+are more complicated than in the original work~\citep{Cannon:2010p10398} due to
 the inclusion of the time-sliced templates and re-sampling.  Figure~%
@@
 \begin{figure}
-	\begin{center}
-		\subfloat[Mismatch versus \textsc{svd} tolerance]{\label{fig:hist-svd-tolerance}\includegraphics{bw.pdf}}
-		\subfloat[Mismatch versus $N^\shortuparrow$]{\label{fig:hist-interpolate}\includegraphics{bw_resample.pdf}}
+		\subfloat[Mismatch versus \textsc{svd} tolerance]{\label{fig:hist-svd-tolerance}\includegraphics{figures/bw.pdf}}
+		\subfloat[Mismatch versus $N^\shortuparrow$]{\label{fig:hist-interpolate}\includegraphics{figures/bw_resample.pdf}}
 		\caption{Box-and-whisker plot of mismatch between nominal
@@
 In \subref{fig:hist-interpolate}, the \SVD\ tolerance is fixed at $\left(1-10^{-6}\right)$ while $N^\shortuparrow$ is varied from 8 to 192 coefficients.}
-	\end{center}
 \end{figure}
@@
 %
-\Table{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
+\begin{table}
+\caption{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
 A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.}
-\br
+\begin{tabular}{lll}
+\tableline\tableline
 Method & \flops\ & Latency (s) \\
-\mr
-Time-domain (\TD) convolution & $2.4\times10^{13}$ & 0 \\
-Frequency-domain (\FD) convolution & $2.6\times10^8$ & 2201 \\
+\tableline
+Time domain (\TD) & $2.4\times10^{13}$ & 0 \\
+Frequency domain (\FD) & $2.6\times10^8$ & 2201 \\
 \lloid\ (this work) & $4.7\times10^8$ & 0 \\
-\br
+\tableline
 \end{tabular}
-\end{indented}
 \end{table}

### 2011-06-09 snr_in_time.py

@@ -44,7 +44,7 @@ def snr_to_num(fracsnr, number=40):
 	return number * (fracsnr)**3
 
 # Start a new figure
-fig = pylab.figure(figsize=(5.5,2.3))
+fig = pylab.figure(figsize=(3,2.5))
 
 markers = ['k-', 'k--', 'k:']
 
@@ -98,7 +98,7 @@ pylab.grid()
 #pylab.legend(loc='lower left')
 pylab.ylabel(r'Observable NS-NS systems / yr')
 pylab.xlabel(r'Time before coalescence, $t$ (s)')
-pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.08,right=0.95)
+pylab.subplots_adjust(bottom=0.05,top=0.95,left=0.1,right=0.95)
 pylab.xlim([0.01,1000])
 pylab.ylim([.1, 1000])
 pylab.savefig(sys.argv[1])

### 2011-06-09 introduction.tex
Changing optical to EM
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1000 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 project to send alerts when potential \GW{} transients were observed in
-order to trigger followup observations by optical telescopes.  The best
+order to trigger followup observations by \EM{} telescopes.  The best
 demonstrated latencies were on the order of 1 hour%

### 2011-06-10 method.tex
Moved units out of expression to fit in column
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1001 ccdd05d5-56c1-461a-bf98-85f6876523b4
Diagram is now in figure* environment instead of figure
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1003 ccdd05d5-56c1-461a-bf98-85f6876523b4
Shortened rows in table to make it fit better
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1004 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
-\begin{figure}[h]
-	\plotone{figures/lloid-diagram.pdf}
+\begin{figure*}[h!]
+	\begin{center}
+		\includegraphics{figures/lloid-diagram.pdf}
 		\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
@@
 much less than the number of output channels.}
-\end{figure}
+	\end{center}
+\end{figure*}
 %
@@
 \tableline
-\numtmps		& number of templates \\
-\tmpsamps	& number of samples per template \\
-\numslices	& number of time slices \\
-\numsvdtmps	& number of orthogonal templates in time slice $s$ \\
-\slicessamps	& number of samples in decimated time slice $s$\\
+\numtmps		& num. of templates \\
+\tmpsamps	& num. of samples per template \\
+\numslices	& num. of time slices \\
+\numsvdtmps	& num. of orthogonal templates in time slice $s$ \\
+\slicessamps	& num. of samples in decimated time slice $s$\\
 $f^s$		& sample rate in time slice $s$ \\
-$N^\shortdownarrow$ & number of coefficients in decimation filter \\
-$N^\shortuparrow$ & number of coefficients in interpolation filter \\
+$N^\shortdownarrow$ & num. of coefficients in decimation filter \\
+$N^\shortuparrow$ & num. of coefficients in interpolation filter \\
 \tableline
@@
 $$
-\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right) f \textrm{ \flops}\,.
+\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right)
 $$
-The second sum is carried out over the set of distinct sample rates rather than over the time slices themselves.
+\flops.  The second sum is carried out over the set of distinct sample rates rather than over the time slices themselves.

### 2011-06-10 method.tex
Fixing column overflow for long equation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1007 ccdd05d5-56c1-461a-bf98-85f6876523b4
Reorganize section 2.1 to make grumpy grad students happy.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1008 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 parameters $\mathbf\Theta$ that determine the amplitude and phase evolution of the
-\GW{} strain. For systems where the effects of spin can be
-ignored, the intrinsic source parameters are the component masses of the
-binary, $\mathbf\Theta = (m_1, m_2)$. Searches for inspiral signals typically
-employ matched filter banks that discretely sample the possible intrinsic
-parameters~\citep{findchirppaper}.  For a given source, the strain observed by
-the detector is a linear combination of two waveforms corresponding to the
+\GW{} strain. For systems in which the effects of spin can be ignored, the intrinsic
+source parameters are the just component masses of the binary,
+ $\mathbf\Theta = (m_1, m_2)$. For a given source, the strain observed by the
+ detector is a linear combination of two waveforms corresponding to the
 `$+$' and `$\times$' \GW{} polarizations.  Thus, for any value of
-$\mathbf\Theta$ we must implement 2 filters.  The coefficients for the $\numtmps$ filters
-are known as templates, and are formed by discretizing and time reversing the
-waveforms and weighting them by the inverse amplitude spectral density of the
-detector's noise. To construct a template bank, templates are chosen with the
+$\mathbf\Theta$ we must implement two filters (real-valued) filters.
+
+Searches for inspiral signals typically employ matched filter
+banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.
+To construct a template bank, templates are chosen with
 $\numtmps/2$ discrete signal parameters $\mathbf\Theta_0,\, \mathbf\Theta_1,\, \dots,\,
-\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~\citep{Owen:1998dk}. That
+\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~ for signals given by
+intermediate values of the intrinsic parameters~\citep{Owen:1998dk}. That
 is, any possible signal within a given range of intrinsic
@@
 least one template. Such a template bank is said to have a {\em minimum match}
-of 0.97. The data from the detector are whitened and convolved with each
-template to produce $\numtmps$ \SNR\ time series. Local peak-finding across time and
-templates determines detection candidates.
+of 0.97.
 
@@
 \end{equation}
+The coefficients for the $\numtmps$ filters are known as templates, 
+and are formed by discretizing and time reversing the
+waveforms and weighting them by the inverse amplitude spectral density of the
+detector's noise. The data from the detector are whitened and convolved with each
+template to produce $\numtmps$ \SNR\ time series. Local peak-finding across time and
+templates determines detection candidates.
 
@@
 %
-\begin{equation}
+\begin{multline}
 	\rho_i^s [k] =%
+		% Interpolation SNR
+		\overbrace{
+			\left(H^\uparrow \rho_i^{s+1}\right)[k]
+		}^\textrm{\clap{{\sc snr} from previous time slices}} \\
+		% Plus ...
+		+
 		% Reconstruction
@@
 			\sum_{\mathclap{n=0}}^{\mathclap{N^s-1}} u_l^s[n] x^s[k-n]
-		}^\textrm{\clap{orthogonal {\sc fir} filters}}
-		% Plus ...
-		+
-		% Interpolation SNR
-		\underbrace{
-			\left(H^\uparrow \rho_i^{s+1}\right)[k]
-		}_\textrm{\clap{{\sc snr} from previous time slices}}
-\end{equation}
+		}^\textrm{\clap{orthogonal {\sc fir} filters}} .
+\end{multline}
 %

### 2011-06-10 method.tex
Reorganize section 2.2 to make a less grumpy grad students even happier.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1010 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Here we explore a method for reducing the computational cost of a \TD\ search
-for compact binary coalescence.  We will give a truly zero-latency algorithm
+Here we describe a method for reducing the computational cost of a \TD\ search
+for compact binary coalescence.  We give a truly zero-latency algorithm
 that competes in terms of floating point operations per second with the
@@
 Low Latency Online Inspiral Detection),
-involves two transformations of the template waveforms that produce a set of
-orthogonal filters with far fewer coefficients than the original templates.
+involves two transformations of the template waveforms that produce a
+smaller set of \emph{orthogonal} filters with far fewer coefficients than
+ the original templates.
 
@@
 factor of $\sim$100 from the early part of the waveform and allows the filters
-to be evaluated at about $\sim$$1/100^\mathrm{th}$ the sample rate. This
-amounts to more than a factor of 10000
-reduction in the floating-point operations per second required to filter the
-early part of the waveform.  However, the resulting filters are still not
+to be evaluated at about $\sim$$1/100^\mathrm{th}$ the sample rate. Since the
+computational cost of \TD\ filtering is quadratic in the filter length, this
+mounts to more than a factor of $10^4$ reduction in the floating-point operations
+ per second required to filter the early part of the waveform.
+
+However, the resulting filters are still not
 orthogonal across the mass parameter space, and are in fact highly redundant.
-We use the \SVD{} to produce an orthogonal filter set
-from the time-sliced templates~\citep{Cannon:2010p10398}.  We find that this
-reduces the number of sample points of the filters by another factor of
+We use the \SVD{} to approximate the template bank by a set of orthogonal filters
+~\citep{Cannon:2010p10398}.  We find that this approximation
+reduces the number of filters needed by another factor of
 $\sim$100.  The combined methods reduce the number of floating point operations

### 2011-06-10 method.tex
Fixed typo.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1011 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 computational cost of \TD\ filtering is quadratic in the filter length, this
-mounts to more than a factor of $10^4$ reduction in the floating-point operations
+amounts to more than a factor of $10^4$ reduction in the floating-point operations
  per second required to filter the early part of the waveform.

### 2011-06-10 method.tex
Reorganize section 2.3 to make happy grad student happier.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1012 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 slices in a \TD{} analogue to the \FD{} decomposition described
-in~\citep{Marion2004, Buskulic2010}.  To wit,
+in~\citep{Marion2004, Buskulic2010}.  To wit, we decompose each template
+$h_{i}[k]$ into a sum of $S$ non-overlapping filters
 %
@@
 t^S = N$.  A matched filter is constructed for each time slice.  The outputs
-form an ensemble of partial \SNR{} streams.  By linearity, these partial \SNR\
-streams can be suitably time-delayed and summed to reproduce the \SNR\ of the
-full template.  We will show in the next section that this, combined with the
-\SVD{}, is sufficient to enable a computationally efficient \TD{} search and
-furthermore is an essential part of an \earlywarning\ detection scheme.
+form an ensemble of partial \SNR{} streams.  By linearity of the filtering
+process, these partial \SNR\ streams can be suitably time-delayed and summed to
+ reproduce the \SNR\ of the full template.
 
-For concreteness and simplicity, we will consider an inspiral waveform in the
+Since waveforms with neighboring intrinsic source parameters $\mathbf\Theta$
+ have similar time-frequency evolution, it is possible to design computationally
+efficient time slices for an extended region of parameter space rather than to
+design different time slices for each template.
+For concreteness and simplicity, consider an inspiral waveform in the
 quadrapole approximation, for which the time-frequency relation is
@@
 M_\odot / c^3 \approx 5 \upmu\mathrm{s}$) and $t$ is the time relative to the
-coalescence of the binary~\citep{findchirppaper, kidder1992}.  Usually the
-template is truncated at some prescribed time $t^0$, or equivalently frequency
-$f_\mathrm{hi}$.  This is often chosen to correspond to the \ISCO. An inspiral
-signal will enter the detection band at a low frequency, $f = f_\mathrm{low}$,
-corresponding to a time $t_\mathrm{low}$.  The template is assumed to be zero
-outside the interval $[t^0, t_\mathrm{low})$ and is said to have  a duration of
-$t_\mathrm{low} - t^0$. It is critically sampled at a rate of $2
-f_\mathrm{hi}$.
-
+coalescence of the binary~\citep{findchirppaper, kidder1992}.
 The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us
 to choose time-slice boundaries that require substantially less bandwidth at
-early times in the inspiral.  Our goal is to reduce the filtering cost of a
+early times in the inspiral.
+
+In any given interval of time, the filters are band-limited across the enitre
+template bank. Usually the template is truncated at some prescribed time $t^0$,
+or equivalently frequency $f_\mathrm{hi}$, often chosen to correspond to the
+\ISCO. An inspiral signal will enter the detection band at some low frequency,
+$f = f_\mathrm{low}$, corresponding to a time $t_\mathrm{low}$.  The end of the
+template is then critically sampled at a rate of $2 f_\mathrm{hi}$ but the 
+beginning of the template is critically sampled at $2 f_\mathrm{lo}$. In any
+time interval smaller than the duration of the template, the bandwidth of the
+filters across the entire template bank may be significantly less and
+a lower sampling rate may be used for all filters without aliasing.
+
+Our goal is to reduce the filtering cost of a
 large fraction of the waveform by computing part of the convolution at a lower
@@
 $\left[t^0, t^1\right),\, \left[t^1, t^2\right),\, \dots,\, \left[t^{S-1}, t^S\right)$,
-sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^0
-\geqslant 2 f_\mathrm{hi}$ and $f^{S-1} \geqslant 2 f_\mathrm{low}$.  The time
-sliced templates may be down-sampled without aliasing, so we define them as
+sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^j$ is at
+least twice the highest frequency component of any filter in the bank for the $j$th
+time slice.
+
+The time sliced templates may then be down-sampled in each interval without
+aliasing, so we define them as
 %
@@
 %
+We note that the time-slice decomposition in equation~(\ref{eq:time-slices}) is
+manifestly orthogonal since the time slices are disjoint in time.  In the next
+section we examine how to reduce the number of filters within each time slice
+via \SVD{} of the time-sliced templates.
+
 \begin{comment}
@@
 \end{comment}
-
-Since waveforms with neighboring intrinsic source parameters $\mathbf\Theta$ have
-similar time-frequency evolution, it is possible to design computationally
-efficient time slices for an extended region of parameter space rather than to
-design different time slices for each template.
-
-We note that the time-slice decomposition in equation~(\ref{eq:time-slices}) is
-manifestly orthogonal since the time slices are disjoint in time.  In the next
-section we examine how to reduce the number of filters within each time slice
-via \SVD{} of the time-sliced templates.
 

### 2011-06-10 introduction.tex
lloid_paper: citep -> citet where appropriate
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1013 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 immediately measure the Hubble constant with unprecedented
-accuracy~\citep{2010ApJ...725..496N}. The most efficient means of obtaining
+accuracy and systematics independent of the standard cosmic distance
+ladder~\citep{2010ApJ...725..496N}. The most efficient means of obtaining
 such joint observations is through the rapid identification and dissemination
@@
 projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
-\LIGO\ assuming event rates predicted in~\citep{Abadie:2010p10836} and anticipated
-detector sensitivity described in \citep{ALIGONoise}.  The
+\LIGO\ assuming event rates predicted in \citet{Abadie:2010p10836} and anticipated
+detector sensitivity described in \citet{ALIGONoise}.  The
 most likely estimates indicate that $\sim$10 sources will produce an \SNR\
@@
 The gray bands give the 5 to 95\% confidence interval as described
-in~\citep{Abadie:2010p10836}.
+in \citet{Abadie:2010p10836}.
 %
@@
 have high throughput at the cost of high latency, or low latency at the cost of
-low throughput.  The former are epitomized by the overlap-save algorithm
+low throughput.  The former are epitomized by the overlap-save algorithm~%
 \citep{numerical-recipes-chapter-13} for frequency domain (\FD) convolution,

### 2011-06-10 method.tex
lloid_paper: citep -> citet where appropriate
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1013 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 possibility of \earlywarning\ detection with modest computing requirements.  We
-expand on the ideas of~\citep{Marion2004, Buskulic2010} that describe a
+expand on the ideas of \citet{Marion2004, Buskulic2010} that describe a
 multi-band decomposition of the compact binary parameter space that resulted in
@@
 runs~\citep{HugheyGWPAW2011}.  We combine this with the \SVD\ rank-reduction
-method of~\citep{Cannon:2010p10398} that exploits the redundancy of
+method of \citet{Cannon:2010p10398} that exploits the redundancy of
 the template banks.
@@
 slices in a \TD{} analogue to the \FD{} decomposition described
-in~\citep{Marion2004, Buskulic2010}.  To wit, we decompose each template
+in \citet{Marion2004, Buskulic2010}.  To wit, we decompose each template
 $h_{i}[k]$ into a sum of $S$ non-overlapping filters
@@
 basis templates}, previously demonstrated
-in~\citep{Cannon:2010p10398}.  Similarly, the time-sliced templates described
+in \citet{Cannon:2010p10398}.  Similarly, the time-sliced templates described
 above can be approximated to arbitrary accuracy by expansion into a set of
@@
 the approximation.  This determines the \SVD\ tolerance, which affects the
-\SNR\ loss due to the approximation.  The authors of \citep{Cannon:2010p10398}
+\SNR\ loss due to the approximation.  The authors of \citet{Cannon:2010p10398}
 showed that high accuracy could be achieved with far fewer basis templates than
@@
 
-The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in \citep{numerical-recipes-chapter-13}.
+The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in \citet{numerical-recipes-chapter-13}.
 It entails splitting the input into blocks of $D$ samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$ samples.

### 2011-06-10 results.tex
Table now spans two columns
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1002 ccdd05d5-56c1-461a-bf98-85f6876523b4
Tweaks to figure 4
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1005 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: citep -> citet where appropriate
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1013 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 with sample rates between 32 and 4096 Hz shown in table \ref{tab:time_slices}.
-\begin{table}
+\begin{table*}
+\begin{center}
 \caption{\label{tab:time_slices} Filter design for these 657 templates.
@@
 \tableline\tableline
-\multicolumn{4}{c}{} &\multicolumn{6}{c}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\
-%\multicolumn{4}{c}{} &\tableline \\
-$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\
-\tableline
+\\ [-1.5ex]
+\multicolumn{4}{c}{} &\multicolumn{6}{c}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\ [1ex]
+\cline{5-10}
+\\ [-1.5ex]
+$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\ \tableline
 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
@@
 \end{tabular}
-\end{table}
+\end{center}
+\end{table*}
 We use this template bank and decomposition for the remainder of this section.
@@
 In this section we check the effect of the \SVD\ tolerance parameter defined
-in~\citep{Cannon:2010p10398}.  By changing the tolerance of the reconstruction
+in \citet{Cannon:2010p10398}.  By changing the tolerance of the reconstruction
 of the physical waveforms one directly affects the mismatch that results from
@@
 adequate to achieve our target \SNR{} loss.
-%
-\begin{figure}
+
+\subsubsection{Effect of the interpolation filter length}
+
+Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$, we studied the
+impact of the length $N^\shortuparrow$ of the interpolation filter.  The results are in
+figure \ref{fig:hist-interpolate}.  We find that a filter length of 16 is sufficient
+to meet our target mismatch of 0.003.
+
+\begin{figure*}[b]
 	\subfloat[Mismatch versus \textsc{svd} tolerance]{\label{fig:hist-svd-tolerance}\includegraphics{figures/bw.pdf}}
@@
 In \subref{fig:hist-interpolate}, the \SVD\ tolerance is fixed at $\left(1-10^{-6}\right)$ while $N^\shortuparrow$ is varied from 8 to 192 coefficients.}
-\end{figure}
-
-\subsubsection{Effect of the interpolation filter length}
-
-Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$, we studied the
-impact of the length $N^\shortuparrow$ of the interpolation filter.  The results are in
-figure \ref{fig:hist-interpolate}.  We find that a filter length of 16 is sufficient
-to meet our target mismatch of 0.003.
+\end{figure*}
 

### 2011-06-11 introduction.tex
spelling correction: run-away --> runaway
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1014 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 As a compact binary system loses energy to \GW{}s, its
-orbital separation decays, leading to a run-away inspiral with the \GW{}
+orbital separation decays, leading to a runaway inspiral with the \GW{}
 amplitude and frequency increasing until the system eventually merges.  It is

### 2011-06-11 introduction.tex
Minor edit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1015 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 immediately measure the Hubble constant with unprecedented
-accuracy and systematics independent of the standard cosmic distance
+accuracy and without systematics dependent on the standard cosmic distance
 ladder~\citep{2010ApJ...725..496N}. The most efficient means of obtaining

### 2011-06-11 introduction.tex
Moved discussion of Figure 1 to main text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1017 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 accuracy for negative latency.  Figure~\ref{fig:earlywarning} shows
-projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
+projected early detection rates for \textsc{ns}--\textsc{ns} binaries in Advanced
 \LIGO\ assuming event rates predicted in \citet{Abadie:2010p10836} and anticipated
-detector sensitivity described in \citet{ALIGONoise}.  The
-most likely estimates indicate that $\sim$10 sources will produce an \SNR\
-above 8 in the detector 10 seconds or more prior to the merger during one
-year of observation, though the uncertainty spans orders of magnitude.
-The gray bands give the 5 to 95\% confidence interval as described
-in \citet{Abadie:2010p10836}.
+detector sensitivity for the `zero detuning, high power' configuration described in 
+\citet{ALIGONoise}.  The most realistic estimates indicate that at an \SNR\ threshold
+of 8, we will observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
+detectable within 10~s of merger and $\sim$1~yr$^{-1}$ will be detectable within
+100~s of merger.  These rates are uncertain by an order of magnitude.
 %
@@
 \plotone{figures/snr_in_time.pdf}
-\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns}
-sources that will be detectable $t$ seconds before coalescence.  The heavy
-solid line is the most likely yearly rate estimate $\dot N_{\mathrm{re}}$
-during Advanced \LIGO.  The shaded region represents the confidence interval
-$[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
-\citep{Abadie:2010p10836}.  With an advanced detector in the `zero detuning,
-high power' configuration \citep{ALIGONoise} and an \SNR\ threshold of 8,
-we will observe a total of $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$, but
-$\sim$10~yr$^{-1}$ will be detectable within 10~s of merger and
-$\sim$1~yr$^{-1}$ within 100~s.}
+\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
+sources that will be detectable by Advanced \LIGO\ $t$ seconds before coalescence.
+The heavy solid line is the most realistic yearly rate estimate.  The shaded
+region represents the 5 to 95\% confidence interval arising from uncertainty
+in predicted event rates \citep{Abadie:2010p10836}.}
 \end{figure}

### 2011-06-11 snr_in_time.py

@@ -44,7 +44,7 @@ def snr_to_num(fracsnr, number=40):
 	return number * (fracsnr)**3
 
 # Start a new figure
-fig = pylab.figure(figsize=(3,2.5))
+fig = pylab.figure(figsize=(3,2))
 
 markers = ['k-', 'k--', 'k:']
 
@@ -87,7 +87,7 @@ for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt'], ['zero det'
 	numminus = snr_to_num(snr1, number=.4)
 
 	# generate the figure
-	ax1 = fig.add_subplot(1,1,1, adjustable='box', aspect=0.5)
+	ax1 = fig.add_subplot(1,1,1, adjustable='box')
 	pylab.loglog(t, num, markers[i], lw=2, label=label)
 	pylab.loglog(t, numminus, markers[i], lw=0.5)
 	pylab.loglog(t, numplus, markers[i], lw=0.5)
@@ -96,9 +96,9 @@ for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt'], ['zero det'
 
 pylab.grid()
 #pylab.legend(loc='lower left')
-pylab.ylabel(r'Observable NS-NS systems / yr')
-pylab.xlabel(r'Time before coalescence, $t$ (s)')
-pylab.subplots_adjust(bottom=0.05,top=0.95,left=0.1,right=0.95)
+pylab.ylabel(r'detections yr$^{-1}$')
+pylab.xlabel(r'time before coalescence, $t$ (s)')
+pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.2,right=0.95)
 pylab.xlim([0.01,1000])
 pylab.ylim([.1, 1000])
 pylab.savefig(sys.argv[1])

### 2011-06-11 method.tex
waveform -> signal
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1019 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In this section we describe a decomposition of the \CBC{} waveform space that
+In this section we describe a decomposition of the \CBC{} signal space that
 reduces time-domain (\TD) filtering cost sufficiently to allow for the
@@
 expand on the ideas of \citet{Marion2004, Buskulic2010} that describe a
-multi-band decomposition of the compact binary parameter space that resulted in
+multi-band decomposition of the compact binary signal space that resulted in
 a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 joint science
@@
 $\numtmps/2$ discrete signal parameters $\mathbf\Theta_0,\, \mathbf\Theta_1,\, \dots,\,
-\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~ for signals given by
+\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR\ for signals given by
 intermediate values of the intrinsic parameters~\citep{Owen:1998dk}. That

### 2011-06-11 method.tex
Filters are implicitly real-valued
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1020 ccdd05d5-56c1-461a-bf98-85f6876523b4
Slight rewording
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1021 ccdd05d5-56c1-461a-bf98-85f6876523b4
Reordered section 2.1 in a more sensible way
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1022 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
  detector is a linear combination of two waveforms corresponding to the
-`$+$' and `$\times$' \GW{} polarizations.  Thus, for any value of
-$\mathbf\Theta$ we must implement two filters (real-valued) filters.
+`$+$' and `$\times$' \GW{} polarizations.  Thus, we must design two filters
+for each $\mathbf\Theta$.
 
@@
 banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.
+The coefficients for the $\numtmps$ filters are known as templates, 
+and are formed by discretizing and time reversing the
+waveforms and weighting them by the inverse amplitude spectral density of the
+detector's noise.
 To construct a template bank, templates are chosen with
 $\numtmps/2$ discrete signal parameters $\mathbf\Theta_0,\, \mathbf\Theta_1,\, \dots,\,
-\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR\ for signals given by
-intermediate values of the intrinsic parameters~\citep{Owen:1998dk}. That
-is, any possible signal within a given range of intrinsic
-source parameters will have an inner product that is $\geqslant 0.97$ with at
-least one template. Such a template bank is said to have a {\em minimum match}
-of 0.97.
+\mathbf\Theta_{\numtmps/2-1}$. These are chosen such that any possible signal
+will have an inner product $\geqslant 0.97$ with at least one template.
+Such a template bank is said to have a {\em minimum match} of 0.97~\citep{Owen:1998dk}.
 
@@
 \end{equation}
-The coefficients for the $\numtmps$ filters are known as templates, 
-and are formed by discretizing and time reversing the
-waveforms and weighting them by the inverse amplitude spectral density of the
-detector's noise. The data from the detector are whitened and convolved with each
-template to produce $\numtmps$ \SNR\ time series. Local peak-finding across time and
+This results in $\numtmps$ \SNR\ time series. Local peak-finding across time and
 templates determines detection candidates.

### 2011-06-11 method.tex
Minor changes to section 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1023 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 that competes in terms of floating point operations per second with the
-conventional \FD\ method, which requires a significant latency in order to be
-computationally cheap. Our method, called \lloid{} (a backronym for
-Low Latency Online Inspiral Detection),
+conventional \FD\ method, which by contrast requires a significant latency due to the
+inherent acausality of the Fourier transform.  Our method, called \lloid{}
+(a backronym for Low Latency Online Inspiral Detection),
 involves two transformations of the template waveforms that produce a
-smaller set of \emph{orthogonal} filters with far fewer coefficients than
- the original templates.
+network of orthogonal filters that is far more computationally
+efficient than the original bank of matched filters.
 

### 2011-06-11 implementation.tex
lloid_paper: citep -> citet where appropriate
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1013 ccdd05d5-56c1-461a-bf98-85f6876523b4
down-sample --> downsample
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1024 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \SVD{} and is the approach considered
-in~\citep{Cannon:2010p10398}.  We choose time-slice boundaries as in
+in \citet{Cannon:2010p10398}.  We choose time-slice boundaries as in
 equation~\eqref{eq:time-sliced-templates} such that all of the templates within
@@
 Next, the templates within the sub-bank are realized as \fir\ filter
-coefficients.  For each time slice, the templates are down-sampled to the
+coefficients.  For each time slice, the templates are downsampled to the
 appropriate sample rate.  Finally, the \SVD\ is applied to each time slice in
@@
 reconstruction matrix that maps them back to the original templates as
-described in equation~\eqref{eq:svddecomp}.  The down-sampled orthogonal \fir\
+described in equation~\eqref{eq:svddecomp}.  The downsampled orthogonal \fir\
 filter coefficients, the reconstruction matrix, and the time-slice boundaries
@@
 \gstreamer\ is useful for our purpose. \gstreamer\ also provides some useful
-stock signal processing elements such as re-samplers and filters.  We have
+stock signal processing elements such as resamplers and filters.  We have
 extended the \gstreamer\ framework by developing a library called

### 2011-06-11 method.tex
down-sample --> downsample
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1024 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 templates, the early time slices have significantly lower bandwidth and can be
-safely down-sampled.  The down-sampling reduces the number of data samples by a
+safely downsampled.  The downsampling reduces the number of data samples by a
 factor of $\sim$100 from the early part of the waveform and allows the filters
@@
 
-The time sliced templates may then be down-sampled in each interval without
+The time sliced templates may then be downsampled in each interval without
 aliasing, so we define them as
@@
 The computational cost of the decimation of the detector data is a little bit more subtle.
-Decimation is achieved by applying an \fir\ anti-aliasing filter and then down-sampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
+Decimation is achieved by applying an \fir\ anti-aliasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
 Naively, an anti-aliasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.

### 2011-06-11 method.tex
Clarified order of magnitude statements in section 2.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1025 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 templates, the early time slices have significantly lower bandwidth and can be
-safely downsampled.  The downsampling reduces the number of data samples by a
-factor of $\sim$100 from the early part of the waveform and allows the filters
-to be evaluated at about $\sim$$1/100^\mathrm{th}$ the sample rate. Since the
+safely downsampled.  Downsampling reduces the total number of filter coefficients
+by a factor of $\sim$100 by treating the earliest part of the waveform
+at $\sim$$1/100$ of the full sample rate. Since the
 computational cost of \TD\ filtering is quadratic in the filter length, this

### 2011-06-11 method.tex
Minor tweaks in section 2.3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1026 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 slices in a \TD{} analogue to the \FD{} decomposition described
-in \citet{Marion2004, Buskulic2010}.  To wit, we decompose each template
-$h_{i}[k]$ into a sum of $S$ non-overlapping filters
+in \citet{Marion2004, Buskulic2010}.  We decompose each template
+$h_{i}[k]$ into a sum of $S$ non-overlapping templates
 %
@@
 for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
-t^S = N$.  A matched filter is constructed for each time slice.  The outputs
+t^S = N$.  The outputs of these new time-slice filters
 form an ensemble of partial \SNR{} streams.  By linearity of the filtering
-process, these partial \SNR\ streams can be suitably time-delayed and summed to
- reproduce the \SNR\ of the full template.
+process, these partial \SNR\ streams can be suitably summed to reproduce the
+\SNR\ of the full template.
 
@@
 design different time slices for each template.
+
 For concreteness and simplicity, consider an inspiral waveform in the
@@
 coalescence of the binary~\citep{findchirppaper, kidder1992}.
-The monotonic time-frequency relationship of equation~\eqref{eq:fgw} allows us
+This monotonic time-frequency relationship allows us
 to choose time-slice boundaries that require substantially less bandwidth at
@@
 template is then critically sampled at a rate of $2 f_\mathrm{hi}$ but the 
-beginning of the template is critically sampled at $2 f_\mathrm{lo}$. In any
+beginning of the template is critically sampled at $2 f_\mathrm{low}$. In any
 time interval smaller than the duration of the template, the bandwidth of the

### 2011-06-11 method.tex
template waveform --> template
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1027 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 (a backronym for Low Latency Online Inspiral Detection),
-involves two transformations of the template waveforms that produce a
+involves two transformations of the templates that produce a
 network of orthogonal filters that is far more computationally
@@
 
-\subsection{Selectively reducing the sample rate of the data and template waveforms}
+\subsection{Selectively reducing the sample rate of the data and templates}
 

### 2011-06-12 method.tex
Minor corrections in section 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1031 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as an
+Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as bank of
 finite impulse response (\fir) filter, requiring $\mathcal O(\numtmps
@@
 much more computationally efficient to use the convolution theorem and the
-\fft\ to implement fast convolution in the frequency-domain, requiring only
-$\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample with increased
-latency.
+\fft\ to implement fast convolution in the frequency domain, requiring only
+$\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample but incurring
+a latency comparable to the length of the templates.
 

### 2011-06-12 method.tex
Reworded last sentence of section 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1032 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample but incurring
-a latency comparable to the length of the templates.
+a latency of $\gtrsim \tmpsamps$ samples.
 

### 2011-06-12 method.tex
Corrections to sectino 2.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1033 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The first transformation is to chop the \TD{} templates up into \emph{time
-slices}.  Since each template slice is disjoint in time, the resulting set for
-a single template is orthogonal.  Given the chirp-like structure of the
-templates, the early time slices have significantly lower bandwidth and can be
-safely downsampled.  Downsampling reduces the total number of filter coefficients
-by a factor of $\sim$100 by treating the earliest part of the waveform
-at $\sim$$1/100$ of the full sample rate. Since the
-computational cost of \TD\ filtering is quadratic in the filter length, this
-amounts to more than a factor of $10^4$ reduction in the floating-point operations
- per second required to filter the early part of the waveform.
+The first transformation is to chop the templates into disjointly supported
+intervals, or \emph{time slices}.  Since the time slices of a given template
+are disjoint in time, they are orthogonal with respect to time.  Given the
+chirp-like structure of the templates, the ``early'' (lowest frequency) time
+slices have significantly lower bandwidth and can be safely downsampled.
+Downsampling reduces the total number of filter coefficients by a factor of
+$\sim$100 by treating the earliest part of the waveform at $\sim$$1/100$ of
+the full sample rate.  Together, the factor of 100 reduction in the number of
+filter coefficients and the factor of 100 reduction in the sample rate save a
+factor of $\sim 10^4$ floating point operations per second (\flops) over the
+original (full sample rate) templates.
 
 However, the resulting filters are still not
-orthogonal across the mass parameter space, and are in fact highly redundant.
-We use the \SVD{} to approximate the template bank by a set of orthogonal filters
-~\citep{Cannon:2010p10398}.  We find that this approximation
+orthogonal across the parameter space, and are in fact highly redundant.
+We use the \SVD{} to approximate the template bank by a set of orthogonal
+\emph{basis filters}~\citep{Cannon:2010p10398}.  We find that this approximation
 reduces the number of filters needed by another factor of

### 2011-06-12 method.tex
Edited section 2.4
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1034 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
-t^S = N$.  The outputs of these new time-slice filters
+t^S = N$.  The outputs of these new time-sliced filters
 form an ensemble of partial \SNR{} streams.  By linearity of the filtering
@@
 This monotonic time-frequency relationship allows us
-to choose time-slice boundaries that require substantially less bandwidth at
+to choose time slice boundaries that require substantially less bandwidth at
 early times in the inspiral.
@@
 %
-We note that the time-slice decomposition in equation~(\ref{eq:time-slices}) is
+We note that the time slice decomposition in equation~(\ref{eq:time-slices}) is
 manifestly orthogonal since the time slices are disjoint in time.  In the next
@@
 \begin{comment}
-An example time-slice design satisfying these constraints for a $1.4 - 1.4 \,
+An example time slice design satisfying these constraints for a $1.4 - 1.4 \,
 M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
@@
 
-As described previously, the template banks are, by design, highly correlated.
-It is possible to greatly reduce the number of filters required to achieve a
-particular minimum match by designing an appropriate set of \SVD\ {\em
-basis templates}, previously demonstrated
-in \citet{Cannon:2010p10398}.  Similarly, the time-sliced templates described
-above can be approximated to arbitrary accuracy by expansion into a set of
-\SVD\ \emph{basis templates}, $u_l^s[k]$, related to the original time
-sliced templates through the \emph{reconstruction matrix},
-$v_{il}^s\sigma_l^s$:
+As noted previously, the template banks used in inspiral searches are by design
+highly correlated.  \citet{Cannon:2010p10398} showed that applying the \SVD\
+to inspiral template banks greatly reduces the number of filters required to achieve a
+particular minimum match.  A similar technique can be applied to the time-sliced
+templates as defined in equation~\ref{eq:time-sliced-templates} above.  The \SVD\
+is a matrix factorization that takes the form
 %
@@
 \end{equation}
+where $u_l^s[k]$ are orthonormal \emph{basis templates} related to the original
+time-sliced templates through the \emph{reconstruction matrix}, $v_{il}^s\sigma_l^s$.
+The expectation value of the fractional loss in \SNR\ is the \SVD\ tolerance, given by
 %
-The parameter $\numsvdtmps$ is the number of basis templates that are kept in
-the approximation.  This determines the \SVD\ tolerance, which affects the
-\SNR\ loss due to the approximation.  The authors of \citet{Cannon:2010p10398}
-showed that high accuracy could be achieved with far fewer basis templates than
-templates in the original template bank.  We find that when combined with the
-time-slice decomposition, the number of basis templates \numsvdtmps\ is much
-smaller than the original number of templates \numtmps\ and improves on
-\citep{Cannon:2010p10398} by nearly an order of magnitude.  In the next section
-we describe how we form our early-warning detection statistic using the time-slice
-decomposition and the \SVD.
+\begin{equation*}
+\left[ \sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} \left( \sigma_l^s \right)^2 \right]\left[ \sum_{\mathclap{l=0}}^{\mathclap{M-1}} \left( \sigma_l^s \right)^2 \right]^{-1},
+\end{equation*}
+%
+determined by the number $\numsvdtmps$ of basis templates that are kept in
+the approximation.  The authors of \citet{Cannon:2010p10398}
+showed that highly accurate approximations of inspiral template banks could be
+achieved with $L^s \ll M$.  We find that when combined with the
+time slice decomposition, the number of basis templates \numsvdtmps\ is much
+smaller than the original number of templates \numtmps\ and improves on the
+rank reduction demonstrated in \citep{Cannon:2010p10398} by nearly an order
+of magnitude.  In the next section we describe how we form our early-warning
+detection statistic using the time slice decomposition and the \SVD.
 

### 2011-06-12 implementation.tex
Edited section 2.5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1035 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In order to form the early-warning \SNR\ from each time slice, we have to add
-the partial \SNR\ to the early-warning \SNR\ from the subsequent time slices.
+In order to form the early-warning output from each time slice, we have to add
+the partial \SNR\ to the early-warning output from the subsequent time slices.
 If the next time slice does not have the same sample rate, its output must
@@
 
-The early-warning \SNR\ for each time slice is formed by accumulating
-interpolated early-warning \SNR\ from the subsequent time slice.  This process
+The early-warning output for each time slice is formed by accumulating
+interpolated early-warning output from the subsequent time slice.  This process
 continues until we have worked our way to the \SNR\ of the original templates

### 2011-06-12 method.tex
Edited section 2.5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1035 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsection{Early warning \SNR }
+\subsection{Early-warning output}
 
 In the previous two sections we have described two transformations that greatly
-reduce the burden of filtering waveforms from the the compact binary parameter space and make \TD\
-filtering of the data possible.  We are now prepared to define the early
-warning \SNR\ and to comment on the computational cost of evaluating it.  But
-first, we will introduce some notation referring to the decimation of data,
-which means reducing the sample rate after low-pass filtering,
+reduce the computational burden of \TD\ filtering.  We are now prepared to define
+our detection statistic, the early-warning output, and to comment on the computational
+cost of evaluating it.
+
+First, the sample rate of the detector data must be decimated to match sample rates
+with each of the time slices.  We will denote the decimated detector data streams using
+a superscript $^s$ to indicate the time slices to which they correspond, and the
+operator $H^\shortdownarrow$ will represent the appropriate decimation filter:
 \begin{equation*}
 \label{eq:decomp}
-	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k] \,,
+	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k].
 \end{equation*}
-and notation referring to the interpolation of data to increase the sample rate,
+We shall use the symbol $H^\shortuparrow$ to represent an interpolation filter,
 \begin{equation*}
-	x^{s}[k] = \left( H^\shortuparrow x^{s+1}\right)[k] \,.
+	x^{s}[k] = \left( H^\shortuparrow x^{s+1}\right)[k].
 \end{equation*}
 
-From the combination of transformations to the compact binary templates defined
-in equation \eqref{eq:time-sliced-templates} and \eqref{eq:svddecomp} we define
-the early-warning filter output accumulated up to time slice $s$ as,
+From the combination of the time slice decomposition in
+equation~(\ref{eq:time-sliced-templates}) and the \SVD\ defined in
+equation~(\ref{eq:svddecomp}), we define the early-warning output accumulated
+up to time slice $s$ using the recursion relation,
 %
@@
 %
-This quantity is related to the early-warning \SNR{} by the normalization
-of the partial filter defined by the present time slice, which may be computed
-in advance. \editorial{Chad: are we happy with that statement?}
+Observe that the early-warning output for time slice 0, $\rho_i^0[k]$,
+approximates the \SNR\ of the original templates.
 %
@@
 %
-The signal flow diagram in figure~\ref{fig:pipeline} illustrates how this
-recursion relation can be realized as a multi-rate filter network with outputs for
-each of the early-warning \SNR{}s.
+The signal flow diagram in figure~\ref{fig:pipeline} illustrates this
+recursion relation as a multirate filter network with a number of early-warning outputs.
 

### 2011-06-12 method.tex
Demoted some subsections to subsubsections
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1036 ccdd05d5-56c1-461a-bf98-85f6876523b4
Abbreviation already defined
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1038 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsection{Selectively reducing the sample rate of the data and templates}
+\subsubsection{Selectively reducing the sample rate of the data and templates}
 
@@
 
-\subsection{Reducing the number of filters with the \SVD{}}
+\subsubsection{Reducing the number of filters with the \SVD{}}
 
@@
 
-\subsection{Early-warning output}
+\subsubsection{Early-warning output}
 
@@
 The conventional \TD{} method consists of a bank of \fir{} filters, or sliding-window dot products.
-If there are $\numtmps$ templates, each $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications and additions per sample, or $2 \numtmps \tmpsamps f^0$ floating point operations per second (\flops) at a sample rate $f^0$.
+If there are $\numtmps$ templates, each $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications and additions per sample, or $2 \numtmps \tmpsamps f^0$ \flops\ at a sample rate $f^0$.
 

### 2011-06-12 method.tex
Edited section 2.3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1039 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \tableline
-\numtmps		& num. of templates \\
-\tmpsamps	& num. of samples per template \\
-\numslices	& num. of time slices \\
-\numsvdtmps	& num. of orthogonal templates in time slice $s$ \\
-\slicessamps	& num. of samples in decimated time slice $s$\\
 $f^s$		& sample rate in time slice $s$ \\
-$N^\shortdownarrow$ & num. of coefficients in decimation filter \\
-$N^\shortuparrow$ & num. of coefficients in interpolation filter \\
+\numtmps		& num. templates \\
+\tmpsamps	& num. samples per template \\
+\numslices	& num. time slices \\
+\numsvdtmps	& num. basis templates in time slice $s$ \\
+\slicessamps	& num. samples in decimated time slice $s$\\
+$N^\shortdownarrow$ & num. coefficients in decimation filter \\
+$N^\shortuparrow$ & num. coefficients in interpolation filter \\
 \tableline
@@
 
-The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in \citet{numerical-recipes-chapter-13}.
-It entails splitting the input into blocks of $D$ samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$ samples.
-For each block, the algorithm computes the forward \fft\ of the data and the templates, multiplies them, and then computes the reverse \fft.
+The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in
+\citet{numerical-recipes-chapter-13}.  It entails splitting the input into blocks of $D$
+samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$
+samples.  For each block, the algorithm computes the forward \fft\ of the data and
+each of the templates, multiplies them, and then computes the reverse \fft.
 
-Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about $2 \fftblock \lg \fftblock$ operations to evaluate a real transform of size $\fftblock$~\citep{Johnson:2007p9654}.
-Including the forward transform of the data and $M$ reverse transforms for all of the templates, the \fft\ costs $2 (\numtmps + 1) \fftblock \lg \fftblock$ operations per block.
-The multiplication of the transforms adds a further $2 \numtmps \fftblock$ operations per block.
-Since each block produces $\fftblock - \tmpsamps$ usable samples of output, the overlap-save method requires
+Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about
+$2 \fftblock \lg \fftblock$ operations to evaluate a real transform of size
+$\fftblock$~\citep{Johnson:2007p9654}.  Including the forward transform of the data and
+$M$ reverse transforms for each of the templates, the \fft\ costs $2 (\numtmps + 1)
+\fftblock \lg \fftblock$ operations per block.  The multiplication of the transforms adds
+a further $2 \numtmps \fftblock$ operations per block.  Since each block produces
+$\fftblock - \tmpsamps$ usable samples of output, the overlap-save method requires
 $$
@@
 
-The story is similar for the interpolation filters used to change the sample rates of the partial \SNR{} streams.
+The story is similar for the interpolation filters used to match the sample rates of the partial \SNR{} streams.
 Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a low-pass filter with $N^\shortdownarrow$ coefficients.
@@
 Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time-sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total the \lloid\ algorithm requires
-$$
-\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right)
-$$
+\begin{multline*}
+\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s \\ + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right) f^s
+\end{multline*}
 \flops.  The second sum is carried out over the set of distinct sample rates rather than over the time slices themselves.

### 2011-06-12 implementation.tex
Edited section 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1040 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 computations that can be completed before the analysis is underway.  Thus
-we divide the procedure into two stages 1) an off-line planning stage and 2) an
-online, low-latency filtering stage.  The off-line stage can be done before the
+we divide the procedure into an offline planning stage and an
+online, low-latency filtering stage.  The offline stage can be done before the
 analysis is started and updated asynchronously, whereas the online stage must

### 2011-06-12 implementation.tex
Edited section 3.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1041 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 satisfy a minimum match criterion.  This assures a prescribed maximum loss in
-\SNR\ for signals that fall between the chosen templates.  Typically the
-minimum match is 0.97 corresponding to a maximum mismatch of 0.03.  Next, the
-templates are subdivided into groups of neighbors called \emph{sub-banks} that
-are appropriately sized so that each bank can be efficiently handled by a
-single computer.  The neighbors are chosen to have comparable chirp mass, which
-produces sub-banks with similar time-frequency evolution.  Dividing the source
-parameter space into smaller sub-banks reduces the computational cost of the
-\SVD{} and is the approach considered
-in \citet{Cannon:2010p10398}.  We choose time-slice boundaries as in
-equation~\eqref{eq:time-sliced-templates} such that all of the templates within
-a sub-bank are sub-critically sampled at progressively lower sample rates.
-Next, the templates within the sub-bank are realized as \fir\ filter
-coefficients.  For each time slice, the templates are downsampled to the
+\SNR\ for signals whose parameters do not lie on the hexagonal grid.  Next, the
+grid is partitioned into groups of neighbors called \emph{sub-banks} that
+are appropriately sized so that each sub-bank can be efficiently handled by a
+single computer.  The sub-banks are chosen to have comparable chirp mass, which
+groups together templates with similar time-frequency evolution.  Dividing the source
+parameter space into smaller sub-banks also reduces the offline cost of the
+\SVD{} and is the approach considered in \citet{Cannon:2010p10398}.  Next, we choose
+time-slice boundaries as in equation~\eqref{eq:time-sliced-templates} such that all
+of the templates within a sub-bank are sub-critically sampled at progressively lower
+sample rates.  For each time slice, the templates are downsampled to the
 appropriate sample rate.  Finally, the \SVD\ is applied to each time slice in
-the sub-bank in order to produce a set of orthogonal \fir\ filters and a
+the sub-bank in order to produce a set of orthonormal basis templates and a
 reconstruction matrix that maps them back to the original templates as
-described in equation~\eqref{eq:svddecomp}.  The downsampled orthogonal \fir\
-filter coefficients, the reconstruction matrix, and the time-slice boundaries
-are all saved to disk.
+described in equation~\eqref{eq:svddecomp}.  The downsampled basis templates,
+the reconstruction matrix, and the time-slice boundaries are all saved to disk.
 

### 2011-06-12 implementation.tex
Revised section 3.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1042 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The \lloid\ algorithm could be used in a true sample-in-sample-out real-time
-system.  However, such a system would likely require integration directly into
-the data acquisition and storage system of the \GW{}
-observatories.  A slightly more modest goal is to leverage existing low-%
-latency, but not real-time, signal processing infrastructure in order to
-implement the \lloid\ algorithm.  For the near-term this should be a viable 
-solution for searches with order seconds of intrinsic latency.
+The \lloid\ algorithm is amenable to truly latency free, realtime implementation.
+However, such a realtime search pipeline would likely require integration directly into
+the data acquisition and storage system of the \LIGO\ observatories.  A slightly more 
+modest goal is to leverage existing low-latency, but not real-time, signal processing
+software in order to implement the \lloid\ algorithm.
 

### 2011-06-12 implementation.tex
Revised section 3.2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1043 ccdd05d5-56c1-461a-bf98-85f6876523b4
Revised section 3.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1044 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsubsection{Decimation}
+\paragraph{Decimation}
 
@@
 element.  The detector data is provided at every power-of-two sample rate
-required by the template time slices described in \eqref{eq:time-slices}. Next,
-these decimated data streams are fed into parallel \fir\ filter banks.
+required by the time slices described in \eqref{eq:time-slices}.  At this point,
+the \gstreamer\ pipeline splits into several parallel branches to handle each
+time slice.  Next, these decimated data streams are fed into parallel \fir\ filter banks.
 
-\subsubsection{\fir\ filters}
+\paragraph{Basis filters}
 
-The \fir\ filtering is implemented using a \gstlal\ element called {\tt
-lal\_firbank}, which produces $N$ channels of filter output from an $N\times M$ matrix of
-\fir\ filter coefficients, representing the \SVD{} basis filters in our application.  This element is used in parallel branches in the
-pipeline to implement the \SVD\ basis filters in each time
-slice.  Rather than implement the time-sliced templates as zero-padded \fir\
-filters as described in \eqref{eq:time-slices} we instead implement them as shorter
-filters that contain only the nonzero samples.  Adding the appropriate time
-offset to the filter output later in the pipeline makes up for the lack of
-explicit zero-padding.
+The basis filters are implemented using a \gstlal\ element called {\tt
+lal\_firbank}, which provides a single input, multiple output bank of \fir\ filters
+from a matrix of filter coefficients.  Several instances of {\tt lal\_firbank}
+execute in parallel to handle each of the time slices.  Filter coefficients are taken
+from the nonzero samples of the time slices in equation~\eqref{eq:time-sliced-templates},
+with the zero-padding effectively accomplished by appropriate delays.
 
-\subsubsection{Reconstruction}
+\paragraph{Reconstruction}
 
-From the outputs of the \SVD\ basis filters, we form the partial \SNR{}s for each time 
-slice by multiplying by the reconstruction matrices.  This is accomplished by connecting 
-the \texttt{gstlal} element \texttt{lal\_matrixmixer} to the output of
-\texttt{lal\_firbank}.
+From the outputs of the \SVD\ basis filters, we form the partial \SNR streams for
+each time  slice by multiplying by the reconstruction matrices.  This is accomplished
+with the \texttt{gstlal} element \texttt{lal\_matrixmixer}.
 
-\subsubsection{Interpolation}
+\paragraph{Interpolation}
 
@@
 If the next time slice does not have the same sample rate, its output must
-first be interpolated.  This is done using the same \gstreamer\ element as was
-used for decimation, {\tt audioresample}.
+first be interpolated.  The \gstreamer\ element {\tt audioresample} is used here.
 
-\subsubsection{\SNR\ accumulation}
+\paragraph{\SNR\ accumulation}
 

### 2011-06-12 results.tex
Revised section 4
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1045 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In this section we evaluate the \lloid{} algorithm using our \gstreamer{}-based implementation described in the previous section. We calculate the
+In this section we evaluate the accuracy of the \lloid{} algorithm using our 
+\gstreamer{}-based implementation described in the previous section. We calculate the
 measured \SNR\ loss due to the approximations of the \lloid\ method and our
-implementation of it. Using settings that give acceptable \SNR{} loss for
-our chosen parameter space, we then compute the operation counts.
+implementation of it. Using a configuration that gives acceptable \SNR{} loss for
+our chosen set of source parameters, we then compare the computational cost in \flops\
+for \TD\ method, the \FD\ method, and \lloid.
 
@@
 between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
-curve~\citep{ALIGONoise}.  Then we create sub-banks by partitioning the parameter
-space by chirp mass.  Figure \ref{fig:tmpltbank} illustrates this procedure.
+curve~\citep{ALIGONoise}.  This results in a grid of 98544 points, or
+$2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
+the parameter space by chirp mass.  We shall concentrate on a sub-bank with 657
+chirp masses between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
+Figure \ref{fig:tmpltbank} illustrates this procedure.
 \begin{figure}[h]
 	\plotone{figures/tmpltbank.png}
-	\caption{\label{fig:tmpltbank}Placement of template parameters used in this paper.
-The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.
-We design a filter bank to search for a subset of 657 of these templates with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
+	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
+case study, consisting of component masses $m_1$, $m_2$, between 1 and 3~$M_\odot$, and
+chirp masses $\mathcal{M}$ between 1.1955 and 1.2045~$M_\odot$.}
 \end{figure}
-The result is that we obtain 657 templates with chirp masses between 1.1955~and~1.2045~$M_\odot$.  With this
-sub-bank we were able to construct an efficient time-slice decomposition that consisted of 11 time slices
-with sample rates between 32 and 4096 Hz shown in table \ref{tab:time_slices}.
+With this sub-bank we were able to construct an efficient time-slice decomposition
+that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
+in table \ref{tab:time_slices}.
 \begin{table*}

### 2011-06-12 results.tex
Edited section 4.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1046 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-We expect two known contributions to the \SNR\ loss to arise in our
+The \SNR\ loss is to be compared with the mismatch of 0.03 that arises from the
+discreteness of template bank designed for a minimum match of 0.97.  We will consider
+an acceptable \SNR\ loss to be a factor of 10 smaller than this, that is, no more
+than 0.003.
+
+We expect two main contributions to the \SNR\ loss to arise in our
 implementation of the \lloid\ algorithm.  The first is the \SNR\ loss due to
-the truncation of the \SVD\ basis and estimates for it
-exist~\citep{Cannon:2010p10398}.  The second comes from non ideal
-implementations of re-sampling that cause signal loss.  The \SNR\ loss is to be
-compared with the mismatch of 0.03 that arises from the discreteness of the
-template bank.  We will consider a target \SNR\ loss to be a factor of 10
-smaller than this, i.e., no more than $\sim$0.003.
+the truncation of the \SVD\ at $L^s < M$ basis templates.  As remarked upon in
+section~\ref{sec:svd} and \citet{Cannon:2010p10398}, this effect is measured by
+the \SVD\ tolerance.  The second comes from the limited bandwidth of the
+interpolation filters used to match the sample rates of the partial \SNR\ streams.
+\SNR\ loss could also arise if the series rather than parallel connection of the
+decimation and interpolation filters reduces their bandwidth measurably,
+if the decimation and interpolation filters do not have perfectly uniform phase
+response, or if there is an unintended subsample time delay at any stage.
 
-It is also within reason that some \SNR\ loss could arise from other
-suboptimalities of implementation in our test pipeline.  To gauge
-accurately how well the pipeline is performing we tested the response to a
-unit impulse.  By taking the inner product of the impulse response for each channel
-with the template, we can gauge very accurately the loss in
-\SNR\ due to the approximations we have made and any inadequacies in
-our implementation.
+To measure the accuracy of our \gstreamer\ implemention of \lloid\ including all of
+the above potential sources of \SNR\ loss, we conducted impulse response tests.  The
+\gstreamer\ pipeline was presented with an input consisting of a unit impulse.  By
+recording the outputs, we can effectively ``play back'' the templates.  These impulse
+responses will be similar, but not identical, to the original, nominal templates.  By
+taking the inner product between the impulses responses for each output 
+channel with the corresponding nominal template, we can gauge exactly how much \SNR\
+is lost due to the approximations in the \lloid\ algorithm and any of the technical
+imperfections mentioned above.
 

### 2011-06-12 results.tex
Removed some articles from subsection names
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1047 ccdd05d5-56c1-461a-bf98-85f6876523b4
Revised section 4.2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1048 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 discreteness of template bank designed for a minimum match of 0.97.  We will consider
-an acceptable \SNR\ loss to be a factor of 10 smaller than this, that is, no more
+an acceptable target \SNR\ loss to be a factor of 10 smaller than this, that is, no more
 than 0.003.
@@
 the truncation of the \SVD\ at $L^s < M$ basis templates.  As remarked upon in
-section~\ref{sec:svd} and \citet{Cannon:2010p10398}, this effect is measured by
+\citet{Cannon:2010p10398} and section~\ref{sec:svd}, this effect is measured by
 the \SVD\ tolerance.  The second comes from the limited bandwidth of the
 interpolation filters used to match the sample rates of the partial \SNR\ streams.
-\SNR\ loss could also arise if the series rather than parallel connection of the
-decimation and interpolation filters reduces their bandwidth measurably,
-if the decimation and interpolation filters do not have perfectly uniform phase
-response, or if there is an unintended subsample time delay at any stage.
+The maximum possible bandwidth is determined by the length of the filter,
+$N^\shortuparrow$.  \SNR\ loss could also arise if the series rather than parallel
+connection of the decimation and interpolation filters reduces their bandwidth
+measurably, if the decimation and interpolation filters do not have perfectly uniform
+phase response, or if there is an unintended subsample time delay at any stage.
 
@@
 is lost due to the approximations in the \lloid\ algorithm and any of the technical
-imperfections mentioned above.
+imperfections mentioned above.  We call one minus this dot product the \emph{mismatch}
+relative to the nominal template.
 
-\subsubsection{Effect of the \SVD\ tolerance}
+\subsubsection{Effect of \SVD\ tolerance}
 
-In this section we check the effect of the \SVD\ tolerance parameter defined
-in \citet{Cannon:2010p10398}.  By changing the tolerance of the reconstruction
-of the physical waveforms one directly affects the mismatch that results from
-the truncation of the orthogonal filter matrix.  The conditions presented here
-are more complicated than in the original work~\citep{Cannon:2010p10398} due to
-the inclusion of the time-sliced templates and re-sampling.  Figure~%
-\ref{fig:hist-svd-tolerance} demonstrates that it is possible to achieve
-typical \SNR\ losses of $\ll1\%$.  However, we do notice that the mismatch
-saturates with respect to the tolerance.  This could be the effect of the
-re-sampling, or another \SNR\ loss that we did not model or expect.  However the
-saturation is still an order of magnitude below our target mismatch of
-$\sim$0.003.  We find that an \SVD\ tolerance of $\left(1-10^{-4}\right)$ is
-adequate to achieve our target \SNR{} loss.
+We studied how the \SVD\ tolerance affected \SNR\ loss by holding
+$N^\shortuparrow = 192$ fixed as we vaired the \SVD\ tolerance from
+$\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The minimum, maximum, and median
+mismatch is shown as a function of \SVD\ tolerance in
+figure~\ref{fig:hist-svd-tolerance}.  As the \SVD\ tolerance increases toward 1, the
+\SVD\ becomes an exact matrix factorization, but the computational cost increases as
+the number of basis filters increases.  The conditions presented here are more
+complicated than in the original work~\citep{Cannon:2010p10398} due to the inclusion
+of the time-sliced templates and interpolation, though we still see that the average
+mismatch is approximately proportional to the \SVD\ tolerance down to
+$\left(1-10^{-4}\right)$.  However, as the \SVD\ tolerance becomes even higher, the
+mismatch seems to saturate around $2 \times 10^{-4}$.  This could be the effect of the
+interpolation, or an unintended technical imperfection that we did not model or expect.
+However, this is still an order of magnitude below our target mismatch of 0.003.  We
+find that an \SVD\ tolerance of $\left(1-10^{-4}\right)$ is adequate to achieve our
+target \SNR{} loss.
 
-\subsubsection{Effect of the interpolation filter length}
+\subsubsection{Effect of interpolation filter length}
 

### 2011-06-12 results.tex
Edited section 4.2.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1049 ccdd05d5-56c1-461a-bf98-85f6876523b4
Changed 2 subsubsections to paragraphs
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1050 ccdd05d5-56c1-461a-bf98-85f6876523b4
Revised section 4.3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1051 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \subsection{Setup}
+\label{sec:bank-setup}
 
@@
 
-\subsubsection{Effect of \SVD\ tolerance}
+\paragraph{Effect of \SVD\ tolerance}
 
@@
 
-\subsubsection{Effect of interpolation filter length}
+\paragraph{Effect of interpolation filter length}
 
 Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$, we studied the
-impact of the length $N^\shortuparrow$ of the interpolation filter.  The results are in
-figure \ref{fig:hist-interpolate}.  We find that a filter length of 16 is sufficient
-to meet our target mismatch of 0.003.
+impact of the length $N^\shortuparrow$ of the interpolation filter.  The mismatch as
+a function of $N^\shortuparrow$ is shown in figure~\ref{fig:hist-interpolate}.  The
+mismatch saturates at $\sim 2 \times 10^{-4}$ with $N^\shortuparrow = 64$.  We find
+that a filter length of 16 is sufficient to meet our target mismatch of 0.003.
 
@@
 sub-bank of templates compared to other methods.  We used the results of the
-previous subsections to set the \SVD\ tolerance to $1-10^{-4}$  and the
-interpolation filter length to 16. Table \ref{table:flops} lists the
-theoretical costs for filtering this sub-bank. Both the \FD\ method and \lloid\
-are five orders of magnitude faster than the conventional \TD\ method.
-However, the \FD\ method has a latency of over half of an hour, whereas \lloid\
-has no latency at all.
+previous subsections to set the \SVD\ tolerance at $\left(1-10^{-4}\right)$ and the
+interpolation filter length to 16. Table \ref{table:flops} shows the computational cost
+in \flops\ for this sub-bank.  Fort the \FD\ method, an \fft\ block size of $\fftblock = 
+2 \tmpsamps$ is assumed, resulting in a latency of $\left(\tmpsamps f^0\right)$
+seconds.  Both the \FD\ method and \lloid\ are five orders of magnitude faster than
+the conventional \TD\ method.  However, the \FD\ method has a latency of over half of an
+hour, whereas \lloid\ has no latency at all.
 %
 \begin{table}
-\caption{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the bank of $2 \times 657$ templates sampled at 4096 Hz for chirp masses $1.1955 \leqslant \mathcal{M} / M_\odot \leqslant 1.2045$.
-A block size of $D = 2 N$ is used for the \fft{}s in the \FD\ method.}
+\caption{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the sub-bank described in section~\ref{sec:bank-setup}.}
 \begin{tabular}{lll}
@@
 \tableline
-Time domain (\TD) & $2.4\times10^{13}$ & 0 \\
-Frequency domain (\FD) & $2.6\times10^8$ & 2201 \\
+Time domain & $2.4\times10^{13}$ & 0 \\
+Frequency domain & $2.6\times10^8$ & 2201 \\
 \lloid\ (this work) & $4.7\times10^8$ & 0 \\

### 2011-06-12 conclusions.tex
lloid_paper: citep -> citet where appropriate
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1013 ccdd05d5-56c1-461a-bf98-85f6876523b4
Edited section 5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1052 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We have demonstrated a computationally feasible procedure for the rapid and
-even preemptive detection of \GW{}s emitted during the coalescence
+even early-warning detection of \GW{}s emitted during the coalescence
 of neutron stars and stellar-mass black holes. Our method is as fast as
-standard \fft{} convolutions but allows for zero latency, sample-in-sample-out
+standard \fft{} convolutions but allows for latency-free, realtime
 operation.  Our search targets are expected to produce prompt electromagnetic
@@
 detecting an electromagnetic counterpart in bands from gamma-rays down to
-radio.  We anticipate requiring $\sim$100 modern multi-core computers to
-analyze a four-detector network of \GW{} data for binary neutron
-stars and stellar-mass black holes.  This is well within the current computing
-capabilities of the \LIGO{} Data Grid~\citep{LDG}. In the future, this could be
-improved upon further by conditionally reconstructing the \SNR{} time-series
+radio.  We anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
+search for binary neutron stars using coincident \GW\ data from a four-detector network.
+This is well within the current computing capabilities of the \LIGO{} Data Grid%
+\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}. In the future, this
+could be improved upon further by conditionally reconstructing the \SNR{} time-series
 only during times when a composite detection statistic crosses a
@@
 fundamental and practical latencies associated with the analysis and detection
-procedure. For example, the \LIGO{} detectors, data acquisition is synchronized
-to a 1/(16~Hz) cadence, introducing an up-front latency~\citep{Bork2001}. Data
-aggregation from the observatories will travel over various networks, each
-capable of high bandwidth but perhaps only modest latency.  This could amount to
-a similar latency of $\sim$100 ms.  Lastly, unless a real-time infrastructure
-is adopted post data acquisition, it is likely that there will be an inherent
-latency introduced by such infrastructure.  We have shown a prototype
-implementation of \lloid{} using the open source signal processing software
+procedure. For example, the \LIGO{} data acquisition records science data in 1/16~s
+blocks~\citep{Bork2001}. Data must also be aggregated from all of the \GW\
+observatories using global computer networks capable of high bandwidth but perhaps only
+modest latency.  This could introduce a latency of $\sim$100~ms.  Lastly, unless a
+realtime infrastructure is adopted post data acquisition, it is likely that there will be 
+an inherent latency introduced by the software architecture itself.  We have shown 
+prototype implementation of \lloid{} using the open source signal processing software
 \gstreamer\ and \gstlal. Significant work would have to be done in order to
-improve upon the latency capability of our implementation, for example, more
-tightly integrating analysis and data acquisition. This should be considered
-for third-generation detector design.
+improve upon the latency of our implementation, including tighter integration between
+data acquisition and analysis. This should be considered for third-generation detector
+design.
 
@@
 not explored the accuracy of sky localization that is possible before merger.
-Ref.~\citep{Fairhurst2009} discusses some of the theoretical prospects for early
+\citet{Fairhurst2009} discusses some of the theoretical prospects for early
 sky localization.  Our future work will explore the prospects of early-warning

### 2011-06-12 method.tex
Edited section 4.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1046 ccdd05d5-56c1-461a-bf98-85f6876523b4
Abbreviation already defined
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1054 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In this section we describe a decomposition of the \CBC{} signal space that
-reduces time-domain (\TD) filtering cost sufficiently to allow for the
+reduces \TD\ filtering cost sufficiently to allow for the
 possibility of \earlywarning\ detection with modest computing requirements.  We
@@
 \subsubsection{Selectively reducing the sample rate of the data and templates}
+\label{sec:time-slices}
 
@@
 \subsubsection{Reducing the number of filters with the \SVD{}}
+\label{sec:svd}
 

### 2011-06-12 method.tex
Wrong index
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1055 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $\left[t^0, t^1\right),\, \left[t^1, t^2\right),\, \dots,\, \left[t^{S-1}, t^S\right)$,
-sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^j$ is at
-least twice the highest frequency component of any filter in the bank for the $j$th
+sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^s$ is at
+least twice the highest frequency component of any filter in the bank for the $s$th
 time slice.

### 2011-06-12 method.tex
Minor edit in section 2.2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1056 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^s$ is at
-least twice the highest frequency component of any filter in the bank for the $s$th
-time slice.
+least twice the highest nonzero frequency component of any filter in the bank for the
+$s$th time slice.
 

### 2011-06-13 inspiral_svd.tex
Changing optical to EM
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1000 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added DCC version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1028 ccdd05d5-56c1-461a-bf98-85f6876523b4
Updated DCC version
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1053 ccdd05d5-56c1-461a-bf98-85f6876523b4
Incremented DCC version to 17
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1060 ccdd05d5-56c1-461a-bf98-85f6876523b4
Reworded acknowledgements
git-svn-id:

@@
 about low latency analysis methods.  NF would like to thank Alessandra Corsi
-for illuminating discussions on astronomical motivations.  DK was supported
-from the Max Planck Gesellschaft. This research is supported by the National
-Science Foundation through a Graduate Research Fellowship to LS and by a
-fellowship from The Perimeter Institute for Theoretical Physics to CH.
+for illuminating discussions on astronomical motivations.  This research is
+supported by the National Science Foundation through a Graduate Research
+Fellowship to LS and by the Perimeter Institute for Theoretical Physics
+through a fellowship to CH. DK is supported from the Max Planck Gesellschaft.
 
-This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}.
+This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{17}.
 

### 2011-06-13 method.tex
Removed mathclaps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1057 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixing grammar
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1063 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \GW{} strain. For systems in which the effects of spin can be ignored, the intrinsic
-source parameters are the just component masses of the binary,
+source parameters are just the component masses of the binary,
  $\mathbf\Theta = (m_1, m_2)$. For a given source, the strain observed by the
@@
 will have an inner product $\geqslant 0.97$ with at least one template.
-Such a template bank is said to have a {\em minimum match} of 0.97~\citep{Owen:1998dk}.
+Such a template bank is said to have a {\em minimal match} of 0.97~\citep{Owen:1998dk}.
 
@@
 \begin{equation*}
-\left[ \sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} \left( \sigma_l^s \right)^2 \right]\left[ \sum_{\mathclap{l=0}}^{\mathclap{M-1}} \left( \sigma_l^s \right)^2 \right]^{-1},
+\left[ \sum_{l=0}^{L^s-1} \left( \sigma_l^s \right)^2 \right]\left[ \sum_{l=0}^{M-1} \left( \sigma_l^s \right)^2 \right]^{-1},
 \end{equation*}

### 2011-06-13 conclusions.tex
Minor edits
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1064 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 of neutron stars and stellar-mass black holes. Our method is as fast as
-standard \fft{} convolutions but allows for latency-free, realtime
+standard \fft{} convolutions but allows for latency-free, real-time
 operation.  Our search targets are expected to produce prompt electromagnetic
@@
 modest latency.  This could introduce a latency of $\sim$100~ms.  Lastly, unless a
-realtime infrastructure is adopted post data acquisition, it is likely that there will be 
+real-time infrastructure is adopted post data acquisition, it is likely that there will be 
 an inherent latency introduced by the software architecture itself.  We have shown 

### 2011-06-13 implementation.tex
Minor edits
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1064 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The \lloid\ algorithm is amenable to truly latency free, realtime implementation.
-However, such a realtime search pipeline would likely require integration directly into
+The \lloid\ algorithm is amenable to truly latency free, real-time implementation.
+However, such a real-time search pipeline would likely require integration directly into
 the data acquisition and storage system of the \LIGO\ observatories.  A slightly more 

### 2011-06-13 method.tex
Minor edits
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1064 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 possibility of \earlywarning\ detection with modest computing requirements.  We
-expand on the ideas of \citet{Marion2004, Buskulic2010} that describe a
+expand on the ideas of \citet{Marion2004} and \citet{Buskulic2010} that describe a
 multi-band decomposition of the compact binary signal space that resulted in
@@
 \mathbf\Theta_{\numtmps/2-1}$. These are chosen such that any possible signal
-will have an inner product $\geqslant 0.97$ with at least one template.
+will have an inner product $\geqslant$0.97 with at least one template.
 Such a template bank is said to have a {\em minimal match} of 0.97~\citep{Owen:1998dk}.

### 2011-06-13 conclusions.tex
Moved two sentences in conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1066 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 standard \fft{} convolutions but allows for latency-free, real-time
-operation.  Our search targets are expected to produce prompt electromagnetic
-signals and may be the progenitors of some short hard gamma-ray bursts.  Rapid
-alerts to the broader astronomical community will improve the chances of
-detecting an electromagnetic counterpart in bands from gamma-rays down to
-radio.  We anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
+operation.  We anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
 search for binary neutron stars using coincident \GW\ data from a four-detector network.
@@
 
-The algorithm we described has no intrinsic latency.  However, there are
+\CBC\ events may be the progenitors of some short hard gamma-ray bursts and are
+expected to be accompanied prompt electromagnetic signals.  Rapid
+alerts to the broader astronomical community will improve the chances of
+detecting an electromagnetic counterpart in bands from gamma-rays down to
+radio.  The detection algorithm we described has no intrinsic latency.  However, there are
 fundamental and practical latencies associated with the analysis and detection

### 2011-06-13 conclusions.tex
Fleshed out future work section a little bit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1068 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Although we have demonstrated a feasible method for advance detection we have
-not explored the accuracy of sky localization that is possible before merger.
-\citet{Fairhurst2009} discusses some of the theoretical prospects for early
-sky localization.  Our future work will explore the prospects of early-warning
-detection with realistic simulations of binary mergers using the infrastructure
-and techniques described here.
+not yet explored triggering, coincidence, and ranking of gravitational wave
+candidates in a framework that supports early \EM\ followup.  If adjacent time slices
+have non-negligible cross-correlation with each other, then there may be some
+subtleties in the design of these postprocessing stages that we must address in order
+to translate the early-warning outputs into candidate events.
+
+Our future work must also address sky localization accuracy prior to merger as well as
+observing strategies.  \citep{Fairhurst2009} derives the area of 90\% localization
+confidence in terms of timing uncertainties, but it would be advantageous to use
+a galaxy catalog to inform the telescope tiling \citep{galaxy-catalog}.  Because
+early detection is likely to be especially effective for nearby sources, the galaxy
+catalog technique might be an important ingredient in reducing the fraction of
+sky that must be imaged.  Extensive injection campaigns and simulations
+incorporating realistic binary merger rates will be necessary in order to fully
+understand the prospects for early-warning detection, localization, and \EM\ folluwup
+using the techniques we have described.
 
@@
 %\end{itemize}
-
-%Future work:
-
-%\begin{itemize}
-%\item Sub-solar mass search
-%\item Hierarchical detection
-%\end{itemize}

### 2011-06-13 conclusions.tex
Added mention of signal based vetoes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1069 ccdd05d5-56c1-461a-bf98-85f6876523b4
Slight rewording
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1070 ccdd05d5-56c1-461a-bf98-85f6876523b4
Moved composite detection statistic into future work section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1071 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 This is well within the current computing capabilities of the \LIGO{} Data Grid%
-\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}. In the future, this
-could be improved upon further by conditionally reconstructing the \SNR{} time-series
-only during times when a composite detection statistic crosses a
-threshold~\citep{svd-compdetstat}.
+\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.
 
@@
 
+% This passage goes in the 'future work' part, but I am having trouble working it in.
+\begin{comment}
+In the future, this
+could be improved upon further by conditionally reconstructing the \SNR{} time-series
+only during times when a composite detection statistic crosses a
+threshold~\citep{svd-compdetstat}.
+\end{comment}
+
 Although we have demonstrated a feasible method for advance detection we have
@@
 subtleties in the design of these postprocessing stages that we must address in order
-to translate the early-warning outputs into candidate events.
+to translate the early-warning outputs into candidate events.  The time-slice
+decomposition and the \SVD\ will help us form low-latency signal based vetoes
+(for example, a $\chi^2$ statistic) that will help us here.
 

### 2011-06-14 method.tex
minor edits
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1072 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample but incurring
-a latency of $\gtrsim \tmpsamps$ samples.
+a latency of $\mathcal O(\tmpsamps)$ samples.
 
@@
 inherent acausality of the Fourier transform.  Our method, called \lloid{}
-(a backronym for Low Latency Online Inspiral Detection),
+(Low Latency Online Inspiral Detection),
 involves two transformations of the templates that produce a

### 2011-06-14 introduction.tex
removed algorithmic latency
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1018 ccdd05d5-56c1-461a-bf98-85f6876523b4
minor edits
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1073 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In the era of first-generation detectors, the \GW{} community initiated a
-project to send alerts when potential \GW{} transients were observed in
-order to trigger followup observations by \EM{} telescopes.  The best
-demonstrated latencies were on the order of 1 hour%
-~\citep{HugheyGWPAW2011}, which was an important
-achievement, but far too late to catch prompt \EM{}
-emission. In order to extract the maximum science from Advanced \LIGO{}
-\CBC{} events, we have the ambition of reporting \GW{} candidates not minutes
-\emph{after} the merger, as has already been accomplished, but seconds
-\emph{before}.
+project to send alerts when potential \GW{} transients were observed in order
+to trigger followup observations by \EM{} telescopes.  The typical latencies
+were 30--60 minutes ~\citep{HugheyGWPAW2011}, which was an important
+achievement, but far too late to catch prompt \EM{} emission. We have the
+ambition of reporting \GW{} candidates not minutes \emph{after} the merger, as
+has already been accomplished, but seconds \emph{before}.
 
@@
 threshold crossings before the \GW{} signal leaves the detection band, it is
-possible to trade some \SNR{} and sky localization
-accuracy for negative latency.  Figure~\ref{fig:earlywarning} shows
-projected early detection rates for \textsc{ns}--\textsc{ns} binaries in Advanced
-\LIGO\ assuming event rates predicted in \citet{Abadie:2010p10836} and anticipated
-detector sensitivity for the `zero detuning, high power' configuration described in 
-\citet{ALIGONoise}.  The most realistic estimates indicate that at an \SNR\ threshold
-of 8, we will observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
-detectable within 10~s of merger and $\sim$1~yr$^{-1}$ will be detectable within
-100~s of merger.  These rates are uncertain by an order of magnitude.
+possible to trade some \SNR{} and sky localization accuracy for pre-merger
+detection.  Figure~\ref{fig:earlywarning} shows projected early detection rates
+for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
+predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
+the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
+The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
+observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
+detectable within 10~s of merger and $\sim$1~yr$^{-1}$ will be detectable
+within 100~s of merger.  These rates are uncertain and the 5 to 95\% confidence
+interval is shown in figure \ref{fig:earlywarning}.
 %
@@
 searches.  The most obvious example of the latter is the time domain
-(\TD) convolution, which has no algorithmic latency.  However, its
+(\TD) convolution, which is latency-free.  However, its
 computational complexity is quadratic in the length of the templates, so it is

### 2011-06-14 conclusions.tex
add new paragraph on sky localization and reorganized intro a bit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1074 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Our future work must also address sky localization accuracy prior to merger as well as
-observing strategies.  \citep{Fairhurst2009} derives the area of 90\% localization
+observing strategies.  \citet{Fairhurst2009} derives the area of 90\% localization
 confidence in terms of timing uncertainties, but it would be advantageous to use

### 2011-06-14 introduction.tex
add new paragraph on sky localization and reorganized intro a bit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1074 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\subsection{\EM{} counterparts}
+
 Though compact binary coalescence (\CBC{}) events may be responsible for several
@@
 
-We show that by examining the signal to noise ratio (\SNR{}) stream for
-threshold crossings before the \GW{} signal leaves the detection band, it is
-possible to trade some \SNR{} and sky localization accuracy for pre-merger
-detection.  Figure~\ref{fig:earlywarning} shows projected early detection rates
-for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
-predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
-the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
-The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
-observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
-detectable within 10~s of merger and $\sim$1~yr$^{-1}$ will be detectable
-within 100~s of merger.  These rates are uncertain and the 5 to 95\% confidence
-interval is shown in figure \ref{fig:earlywarning}.
-%
-\begin{figure}[h]
-\plotone{figures/snr_in_time.pdf}
-\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
-sources that will be detectable by Advanced \LIGO\ $t$ seconds before coalescence.
-The heavy solid line is the most realistic yearly rate estimate.  The shaded
-region represents the 5 to 95\% confidence interval arising from uncertainty
-in predicted event rates \citep{Abadie:2010p10836}.}
-\end{figure}
+\subsection{Rapid \EM{} followup}
 
@@
 
-Realizing advance detection of compact binary coalescences (\CBC{}s) will
+It has been observed that by examining the signal to noise ratio (\SNR{}) stream for
+threshold crossings before the \GW{} signal leaves the detection band, it is
+possible to trade some \SNR{} and sky localization accuracy for pre-merger
+detection. Figure~\ref{fig:earlywarning} shows projected early detection rates
+for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
+predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
+the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
+The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
+observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
+detectable within 10~s of merger and $\sim$1~yr$^{-1}$ will be detectable
+within 100~s of merger. These rates are shown with their substantial uncertainties in figure~\ref{fig:earlywarning}.
+%
+\begin{figure}[h]
+\plotone{figures/snr_in_time.pdf}
+\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
+sources that will be detectable by Advanced \LIGO\ $t$ seconds before coalescence.
+The heavy solid line is the most realistic yearly rate estimate.  The shaded
+region represents the 5 to 95\% confidence interval arising from uncertainty
+in predicted event rates \citep{Abadie:2010p10836}.}
+\end{figure}
+
+Telescopic followup requires estimating the location of the \GW{} source. The
+localization uncertainty is dominated by uncertainty in the time of arrival of
+the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
+\citep{Fairhurst2009}. Unfortunately, for a \GW{} signal with an \SNR{} of
+blah, the 90\% confidence area is at least blah deg$^2$ (blah\% of the sky) at
+10~s before merger and blah deg$^2$ (blah\% of the sky) at 1~s before merger.
+If we can take galactic luminosities to accurately trace mass and thus proxy
+for the \CBC{} event rate, it is possible to reduce the localization
+uncertainty using a galaxy catalog as a prior probability distribution
+\citep{galaxy-catalog}, along with the sky location and luminosity distance
+estimate from the \GW{} signal, as was done in S6/VSR3. At Advanced \LIGO{}
+\textsc{ns}--\textsc{ns} horizon distances of $\sim$400~Mpc, the galaxy catalog
+is fairly isotropic and thus not as useful for downselecting pointings, but
+sources that are detectable before merger are likely to be exceptionally
+nearby. Thus there will be relatively few galaxies to consider for a given
+candidate and the galaxy catalog is likely to be sufficiently complete. This
+may allow us to shrink the 90\% confidence area substantially. Even so, it will
+be desirable to mobilize a large number of heterogeneous telescopes in order to
+guarantee imaging the source at the instant of merger. Because the sky
+localization will improve rapidly down to blah deg$^2$ at merger, by the time
+that the entire \CBC{} event is over it will be possible to downselect the
+images that have already been taken as well as repoint telescopes to more
+likely locations.
+
+\subsection{Low-latency \CBC{} search algorithms}
+
+Realizing advance detection of \CBC{}s will
 require striking a balance between latency and throughput. \CBC{} searches

### 2011-06-14 method.tex
add new paragraph on sky localization and reorganized intro a bit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1074 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 slices in a \TD{} analogue to the \FD{} decomposition described
-in \citet{Marion2004, Buskulic2010}.  We decompose each template
+in \citet{Marion2004} and \citet{Buskulic2010}.  We decompose each template
 $h_{i}[k]$ into a sum of $S$ non-overlapping templates

### 2011-06-14 introduction.tex
Added LSST comparison
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1075 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsection{Rapid \EM{} followup}
+\subsection{Early-warning \EM{} followup}
 
@@
 10~s before merger and blah deg$^2$ (blah\% of the sky) at 1~s before merger.
+It is infeasible to search this large an area for a prompt counterpart. For
+reference, \textsc{lsst} should see first light around the same time as
+Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
+\citep{2008arXiv0805.2366I}.
+
 If we can take galactic luminosities to accurately trace mass and thus proxy
@@
 nearby. Thus there will be relatively few galaxies to consider for a given
-candidate and the galaxy catalog is likely to be sufficiently complete. This
-may allow us to shrink the 90\% confidence area substantially. Even so, it will
-be desirable to mobilize a large number of heterogeneous telescopes in order to
-guarantee imaging the source at the instant of merger. Because the sky
-localization will improve rapidly down to blah deg$^2$ at merger, by the time
-that the entire \CBC{} event is over it will be possible to downselect the
-images that have already been taken as well as repoint telescopes to more
-likely locations.
+early-warning candidate and the galaxy catalog is likely to be sufficiently
+complete. This should allow us to shrink the 90\% confidence area substantially.
+Even so, it will be desirable to mobilize a large number of heterogeneous
+telescopes in order to guarantee imaging this exceptional event at the instant
+of merger. Because the sky localization will improve rapidly down to blah
+deg$^2$ at merger, by the time that the entire \CBC{} event is over it will be
+possible to send a second alert to downselect the images that have already been
+taken as well as repoint telescopes to more likely locations.
 
-\subsection{Low-latency \CBC{} search algorithms}
+\subsection{Latency in \CBC{} search algorithms}
 

### 2011-06-14 introduction.tex
Added figures for sky localization paragraph
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1076 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \citep{Fairhurst2009}. Unfortunately, for a \GW{} signal with an \SNR{} of
-blah, the 90\% confidence area is at least blah deg$^2$ (blah\% of the sky) at
-10~s before merger and blah deg$^2$ (blah\% of the sky) at 1~s before merger.
+10, the 90\% confidence area is at least $2 \times 10^3$~deg$^2$ (4.8\% of the sky) at
+10~s before merger and $10^2$~deg$^2$ (0.3\% of the sky) at 1~s before merger.
 It is infeasible to search this large an area for a prompt counterpart. For

### 2011-06-15 data/localization_uncertainty.py
Added script to compute localization uncertainty
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1078 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -1,0 +1,78 @@
+#!/usr/bin/env python
+
+import numpy
+from numpy import pi
+
+f_ligo, a_ligo = numpy.loadtxt('ZERO_DET_high_P.txt').T
+f_virgo, a_virgo = numpy.loadtxt('AdV_baseline_sensitivity_12May09.txt').T
+
+# Component masses (in M_sun)
+m1 = 1.4
+m2 = 1.4
+
+# Total mass
+M = m1 + m2
+
+# Chirp mass
+mchirp = (m1 * m2)**.6 / (m1+m2)**.2
+
+# Low frequency cutoff
+fLOW = 10
+
+# ISCO frequency (equation 3.6, FINDCHIRP paper)
+fISCO = 4400 / M
+
+def freq_to_time(Mc,f):
+	LAL_MTSUN_SI = 4.9254909500000001e-06
+	Mc = Mc * LAL_MTSUN_SI
+	return 5. * Mc / 256. * (pi * Mc * f) ** (-8./3.)
+
+# Frequency step in Hz
+df = 0.1
+
+# Frequency grid
+f = numpy.arange(fLOW, fISCO + df, df)
+f0_weights = 4 * df * f ** (-7./3)
+f1_weights = 4 * df * f ** (-4./3)
+f2_weights = 4 * df * f ** (-1./3)
+
+# Inverse power spectrum
+invS_ligo = numpy.interp(f, f_ligo, a_ligo) ** -2
+invS_virgo = numpy.interp(f, f_virgo, a_virgo) ** -2
+
+# SNR
+rho_ligo = numpy.cumsum(f0_weights * invS_ligo)
+rho_virgo = numpy.cumsum(f0_weights * invS_virgo)
+
+# Scale everything for an SNR of 10 in LIGO detectors
+factor = 10 / rho_ligo[-1]
+rho_ligo *= factor
+rho_virgo *= factor
+f2_weights *= factor
+f1_weights *= factor
+
+# Effective bandwidth.
+# Note an omission in Fairhurst (2009): the frequency moments have to be
+# defined with a normalization of 1 / \int |h(f)|^2 / S(f) df, or 1 / SNR.
+sigmaf_ligo = numpy.sqrt(numpy.cumsum(f2_weights * invS_ligo) / rho_ligo - (numpy.cumsum(f1_weights * invS_ligo) / rho_ligo)**2)
[…]

### 2011-06-15 data/localization_uncertainty.py
Localization uncertainty is now calculated for snr of 8 at early-warning time
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1079 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -28,7 +28,7 @@ def freq_to_time(Mc,f):
 	return 5. * Mc / 256. * (pi * Mc * f) ** (-8./3.)
 
 # Frequency step in Hz
-df = 0.1
+df = 0.01
 
 # Frequency grid
 f = numpy.arange(fLOW, fISCO + df, df)
@@ -41,11 +41,12 @@ invS_ligo = numpy.interp(f, f_ligo, a_ligo) ** -2
 invS_virgo = numpy.interp(f, f_virgo, a_virgo) ** -2
 
 # SNR
+rho_threshold = 8
 rho_ligo = numpy.cumsum(f0_weights * invS_ligo)
 rho_virgo = numpy.cumsum(f0_weights * invS_virgo)
 
 # Scale everything for an SNR of 10 in LIGO detectors
-factor = 10 / rho_ligo[-1]
+factor = 1 / rho_ligo[-1]
 rho_ligo *= factor
 rho_virgo *= factor
 f2_weights *= factor
@@ -58,8 +59,8 @@ sigmaf_ligo = numpy.sqrt(numpy.cumsum(f2_weights * invS_ligo) / rho_ligo - (nump
 sigmaf_virgo = numpy.sqrt(numpy.cumsum(f2_weights * invS_virgo) / rho_virgo - (numpy.cumsum(f1_weights * invS_virgo) / rho_virgo)**2)
 
 # Timing uncertainty
-sigmat_ligo = 1. / (2 * pi * rho_ligo * sigmaf_ligo)
-sigmat_virgo = 1. / (2 * pi * rho_virgo * sigmaf_virgo)
+sigmat_ligo = 1. / (2 * pi * rho_threshold * sigmaf_ligo)
+sigmat_virgo = 1. / (2 * pi * rho_virgo * (rho_threshold/rho_ligo) * sigmaf_virgo)
 
 # Uncertainty in direction cosines in plane of detector
 sigmax = sigmat_ligo / 7e-3
@@ -73,6 +74,7 @@ t = freq_to_time(mchirp, f)
 
 print '  t   deg^2   %'
 print '-----------------'
-for t_before_merger in [10, 1, 0.1]:
-	a = a90best[t <= t_before_merger][0]
+for t_before_merger in [10, 1, 0.1, 0]:
+	rho = rho_ligo[t >= t_before_merger][-1]
+	a = a90best[t >= t_before_merger][-1]
[…]

### 2011-06-16 introduction.tex
Added sky localization uncertainty table
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1081 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-14 introduction.tex
@@
 observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
-detectable within 10~s of merger and $\sim$1~yr$^{-1}$ will be detectable
-within 100~s of merger. These rates are shown with their substantial uncertainties in figure~\ref{fig:earlywarning}.
+detectable within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable
+within 25~s of merger. These rates are shown with their substantial uncertainties in figure~\ref{fig:earlywarning}.
+%
+\begin{table}
+\caption{\label{table:sky-localization-accuracy}Horizon distance, most likely yearly detection rate, and \SNR\ at merger for a source that produces an \SNR\ of 8 in all three detectors at $t$~s before merger.
+The fifth and sixth columns show the minimum area of the 90\% confidence region on the sky before merger and at merger.}
+\begin{tabular}{rrrrrr}
+\tableline\tableline
+& horiz. & rate & final & \multicolumn{2}{c}{$A$(90\%) (deg$^2$)} \\
+\cline{5-6}
+$t$ (s) & (Mpc) & (yr$^{-1}$) & \SNR\ & early & final \\
+\tableline
+25.0 & 203 & 5 & 15.8 & 2893.2 & 1.8 \\
+10.0 & 269 & 12 & 11.9 & 1150.6 & 3.6 \\
+5.0 & 307 & 18 & 10.4 & 594.6 & 5.1 \\
+1.0 & 360 & 29 & 8.9 & 153.9 & 7.5 \\
+0.1 & 391 & 37 & 8.2 & 32.5 & 9.1 \\
+0.0 & 400 & 40 & 8.0 & 9.6 & 9.6 \\
+\tableline
+\end{tabular}
+\end{table}
 %
@@
 the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}. Unfortunately, for a \GW{} signal with an \SNR{} of
-blah, the 90\% confidence area is at least blah deg$^2$ (blah\% of the sky) at
-10~s before merger and blah deg$^2$ (blah\% of the sky) at 1~s before merger.
+\citep{Fairhurst2009}.  Table~\ref{table:sky-localization-accuracy} shows the
+minimum 90\% confidence area for a source that produces an \SNR\ of 8 in the
+\LIGO\ detectors in Hanford, Washington, and Livingston, Louisiana, and, the
+Virgo detector in Cascina, Italy, at a time $t$ before merger.  Also shown are
+the final \SNR\ and 90\% confidence area at merger for the same source.
+At 10~s before merger, the pre-merger 90\% confidence area is about $1150.6$~deg$^2$
+(2.8\% of the sky) and at 1~s before merger is 153.9~deg$^2$ (0.37\% of the sky).
+The confidence region shrinks quadratically in area as the overall amplitude of the
+signal increases.
+
 It is infeasible to search this large an area for a prompt counterpart. For
@@
 Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
-\citep{2008arXiv0805.2366I}.
-
-If we can take galactic luminosities to accurately trace mass and thus proxy
+\citep{2008arXiv0805.2366I}.  However, if we can take galactic luminosities
+to accurately trace mass and thus proxy
 for the \CBC{} event rate, it is possible to reduce the localization
@@
 telescopes in order to guarantee imaging this exceptional event at the instant
-of merger. Because the sky localization will improve rapidly down to blah
-deg$^2$ at merger, by the time that the entire \CBC{} event is over it will be
+of merger. Because the sky localization will improve rapidly until merger,
+by the time that the entire \CBC{} event is over it will be
 possible to send a second alert to downselect the images that have already been

### 2011-06-16 data/localization_uncertainty.py

@@ -3,6 +3,23 @@
 import numpy
 from numpy import pi
 
+def float_as_string(num, sigfigs = 2):
+	"""Convert a floating point number to a string in scientific notation,
+	with sigfigs significant figures, suitable for a LaTeX document."""
+	import decimal
+	with decimal.localcontext() as ctx:
+		ctx.prec = 2
+		ctx.rounding = decimal.ROUND_HALF_UP
+		sci_str = str(decimal.Decimal(str(num)) + decimal.Decimal('0'))
+
+		sci_str_parts = sci_str.split('E')
+		if len(sci_str_parts) == 1:
+			return sci_str
+		elif len(sci_str_parts) == 2:
+			return r"\ensuremath{%s \times 10^{%s}}" % (sci_str_parts[0],
+				sci_str_parts[1].lstrip('+'))
+
+
 f_ligo, a_ligo = numpy.loadtxt('ZERO_DET_high_P.txt').T
 f_virgo, a_virgo = numpy.loadtxt('AdV_baseline_sensitivity_12May09.txt').T
 
@@ -42,25 +59,25 @@ invS_virgo = numpy.interp(f, f_virgo, a_virgo) ** -2
 
 # SNR
 rho_threshold = 8
-rho_ligo = numpy.cumsum(f0_weights * invS_ligo)
-rho_virgo = numpy.cumsum(f0_weights * invS_virgo)
+rho2_ligo = numpy.cumsum(f0_weights * invS_ligo)
+rho2_virgo = numpy.cumsum(f0_weights * invS_virgo)
 
 # Scale everything for an SNR of 10 in LIGO detectors
-factor = 1 / rho_ligo[-1]
-rho_ligo *= factor
-rho_virgo *= factor
-f2_weights *= factor
-f1_weights *= factor
+#factor = 1 / rho_ligo[-1]
+invS_ligo *= 1 / rho2_ligo[-1]
+invS_virgo *= 1 / rho2_virgo[-1]
+rho2_ligo *= 1 / rho2_ligo[-1]
+rho2_virgo *= 1 / rho2_virgo[-1]
 
 # Effective bandwidth.
 # Note an omission in Fairhurst (2009): the frequency moments have to be
 # defined with a normalization of 1 / \int |h(f)|^2 / S(f) df, or 1 / SNR.
[…]

### 2011-06-16 introduction.tex
Tweaked figures and tables
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1082 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{table}
-\caption{\label{table:sky-localization-accuracy}Horizon distance, most likely yearly detection rate, and \SNR\ at merger for a source that produces an \SNR\ of 8 in all three detectors at $t$~s before merger.
-The fifth and sixth columns show the minimum area of the 90\% confidence region on the sky before merger and at merger.}
+\caption{\label{table:sky-localization-accuracy}Horizon distance, expected detection rate, and \SNR\ at merger for a source that produces an \SNR\ of 8 in all three detectors at time $t$ before merger.
+The fifth and sixth columns show the minimum area of the 90\% confidence region before merger and at merger.}
+\begin{center}
 \begin{tabular}{rrrrrr}
@@
 \end{tabular}
+\end{center}
 \end{table}

### 2011-06-16 results.tex
Beautified figure 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1058 ccdd05d5-56c1-461a-bf98-85f6876523b4
Table entries are lowercase
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1080 ccdd05d5-56c1-461a-bf98-85f6876523b4
Tweaked figures and tables
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1082 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{figure}[h]
-	\plotone{figures/tmpltbank.png}
+	\plotone{figures/tmpltbank.pdf}
 	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
@@
 \begin{table*}
-\begin{center}
 \caption{\label{tab:time_slices} Filter design for these 657 templates.
@@
 We vary \SVD{} tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
+\begin{center}
 \begin{tabular}{rr@{,\,}lc*{6}{r}}
@@
 mismatch is shown as a function of \SVD\ tolerance in
-figure~\ref{fig:hist-svd-tolerance}.  As the \SVD\ tolerance increases toward 1, the
+figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the
 \SVD\ becomes an exact matrix factorization, but the computational cost increases as
@@
 impact of the length $N^\shortuparrow$ of the interpolation filter.  The mismatch as
-a function of $N^\shortuparrow$ is shown in figure~\ref{fig:hist-interpolate}.  The
+a function of $N^\shortuparrow$ is shown in figure~\ref{fig:bw}b.  The
 mismatch saturates at $\sim 2 \times 10^{-4}$ with $N^\shortuparrow = 64$.  We find
@@
 \begin{figure*}[b]
-	\subfloat[Mismatch versus \textsc{svd} tolerance]{\label{fig:hist-svd-tolerance}\includegraphics{figures/bw.pdf}}
-	\subfloat[Mismatch versus $N^\shortuparrow$]{\label{fig:hist-interpolate}\includegraphics{figures/bw_resample.pdf}}
-	\caption{Box-and-whisker plot of mismatch between nominal
+	\begin{minipage}[t]{0.5\textwidth}
+		\begin{center}
+			\includegraphics{figures/bw.pdf}
+			(a) Mismatch versus \SVD\ tolerance
+		\end{center}
+	\end{minipage}
+	\begin{minipage}[t]{0.5\textwidth}
+		\begin{center}
+			\includegraphics{figures/bw_resample.pdf}
+			(b) Mismatch versus $N^\shortuparrow$
+		\end{center}
+	\end{minipage}
+	\caption{\label{fig:bw}Box-and-whisker plot of mismatch between nominal
 template bank and \lloid\ measured impulse responses.  The upper and lower boundaries of
@@
 The whiskers represent the minimum and maximum mismatch over all templates.  In 
-\subref{fig:hist-svd-tolerance} the interpolation filter length is held fixed
-at $N^\shortuparrow = 192$, while the \SVD\ tolerance is varied from
-$\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
-In \subref{fig:hist-interpolate}, the \SVD\ tolerance is fixed at $\left(1-10^{-6}\right)$ while $N^\shortuparrow$ is varied from 8 to 192 coefficients.}
+(a) the interpolation filter length is held fixed at $N^\shortuparrow = 192$, while
+the \SVD\ tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
+In (b), the \SVD\ tolerance is fixed at $\left(1-10^{-6}\right)$ while $N^\shortuparrow$
+is varied from 8 to 192 coefficients.}
 \end{figure*}
@@
 \begin{table}
-\caption{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the sub-bank described in section~\ref{sec:bank-setup}.}
+\caption{Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the sub-bank described in section~\ref{sec:bank-setup}.}
+\begin{center}
 \begin{tabular}{lll}
 \tableline\tableline
-Method & \flops\ & Latency (s) \\
+method & \flops\ & latency (s) \\
 \tableline
-Time domain & $2.4\times10^{13}$ & 0 \\
-Frequency domain & $2.6\times10^8$ & 2201 \\
+time domain & $2.4\times10^{13}$ & 0 \\
+frequency domain & $2.6\times10^8$ & 2201 \\
 \lloid\ (this work) & $4.7\times10^8$ & 0 \\

### 2011-06-16 introduction.tex
Weakened language about galaxy catalog completeness
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1085 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 sources that are detectable before merger are likely to be exceptionally
-nearby. Thus there will be relatively few galaxies to consider for a given
+nearby. Thus there will be fewer galaxies to consider for a given
 early-warning candidate and the galaxy catalog is likely to be sufficiently
-complete. This should allow us to shrink the 90\% confidence area substantially.
+complete. This should allow us to shrink the 90\% confidence area.
 Even so, it will be desirable to mobilize a large number of heterogeneous

### 2011-06-16 introduction.tex
Another attempt at tightening the galaxy catalog discussion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1087 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-16 introduction.tex
@@
 \citep{2008arXiv0805.2366I}.  However, if we can take galactic luminosities
-to accurately trace mass and thus proxy
-for the \CBC{} event rate, it is possible to reduce the localization
-uncertainty using a galaxy catalog as a prior probability distribution
-\citep{galaxy-catalog}, along with the sky location and luminosity distance
-estimate from the \GW{} signal, as was done in S6/VSR3. At Advanced \LIGO{}
-\textsc{ns}--\textsc{ns} horizon distances of $\sim$400~Mpc, the galaxy catalog
-is fairly isotropic and thus not as useful for downselecting pointings, but
+to accurately trace mass and thus proxy for the \CBC{} event rate, it is
+possible to reduce the localization uncertainty using a galaxy catalog, along
+with the sky location and luminosity distance estimate from the \GW{} signal,
+as a prior probability distribution \citep{galaxy-catalog} as was done in
+S6/VSR3. Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon
+distance, the number of galaxies that can produce a given signal
+amplitude is much larger than the number for initial \LIGO{} and thus the
+catalog will not be as useful for downselecting pointings generally. However,
 sources that are detectable before merger are likely to be exceptionally
-nearby. Thus there will be relatively few galaxies to consider for a given
-early-warning candidate and the galaxy catalog is likely to be sufficiently
-complete. This should allow us to shrink the 90\% confidence area substantially.
+nearby. Thus there will be fewer galaxies to consider for a given
+early-warning candidate and catalog completeness will be less of a concern.
+This should allow us to shrink the 90\% confidence area substantially.
 Even so, it will be desirable to mobilize a large number of heterogeneous

### 2011-06-17 introduction.tex
target -> potential
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1089 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
  (\textsc{ns}) and/or black holes (\textsc{bh}) is presently the best
-understood and most actively pursued target source of gravitational
+understood and most actively pursued potential source of gravitational
 radiation for ground-based gravitational-wave (\GW{}) detectors.

### 2011-06-17 introduction.tex
Word choice
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1090 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \citep{2008arXiv0805.2366I}.  However, if we can take galactic luminosities
-to accurately trace mass and thus proxy for the \CBC{} event rate, it is
+to accurately reflect mass and thus be a proxy for the \CBC{} event rate, it is
 possible to reduce the localization uncertainty using a galaxy catalog, along

### 2011-06-17 method.tex
Tweaked figures and tables
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1082 ccdd05d5-56c1-461a-bf98-85f6876523b4
Mentioned triggers and coincidences
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1091 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed plural
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1092 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 This results in $\numtmps$ \SNR\ time series. Local peak-finding across time and
-templates determines detection candidates.
+template indices results in single-detector triggers.  Coincidences are sought
+between triggers in different \GW\ detectors in order to form detection candidates.
 
 Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as bank of
-finite impulse response (\fir) filter, requiring $\mathcal O(\numtmps
+finite impulse response (\fir) filters, requiring $\mathcal O(\numtmps
 \tmpsamps)$ floating point operations per sample.  However, it is typically
@@
 \caption{\label{tab:recap}Notation used to describe filters.}
+\begin{center}
 \begin{tabular}{ll}

### 2011-06-17 method.tex
Reworded one sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1093 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \emph{basis filters}~\citep{Cannon:2010p10398}.  We find that this approximation
-reduces the number of filters needed by another factor of
-$\sim$100.  The combined methods reduce the number of floating point operations
-to the level where they are competitive with the conventional high-latency \FD\
+reduces the number of filters needed by another factor of $\sim$100.  These two
+transformations combined reduce the number of floating point operations
+to the level that is competitive with the conventional high-latency \FD\
 matched filter approach.  In the remainder of this section we describe the

### 2011-06-17 method.tex
Moved citation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1094 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 For concreteness and simplicity, consider an inspiral waveform in the
-quadrapole approximation, for which the time-frequency relation is
+quadrapole approximation, for which the time-frequency relation~\citep{kidder1992, findchirppaper} is
 %
@@
 M_\odot / c^3 \approx 5 \upmu\mathrm{s}$) and $t$ is the time relative to the
-coalescence of the binary~\citep{findchirppaper, kidder1992}.
+coalescence of the binary.
 This monotonic time-frequency relationship allows us

### 2011-06-17 method.tex
Removed 'suitably'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1095 ccdd05d5-56c1-461a-bf98-85f6876523b4
Spelling correction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1096 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 form an ensemble of partial \SNR{} streams.  By linearity of the filtering
-process, these partial \SNR\ streams can be suitably summed to reproduce the
+process, these partial \SNR\ streams can be summed to reproduce the
 \SNR\ of the full template.
@@
 
-In any given interval of time, the filters are band-limited across the enitre
+In any given interval of time, the filters are band-limited across the entire
 template bank. Usually the template is truncated at some prescribed time $t^0$,

### 2011-06-17 method.tex
Removed inappropriate inequality
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1097 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 showed that highly accurate approximations of inspiral template banks could be
-achieved with $L^s \ll M$.  We find that when combined with the
+achieved with relatively few basis templates.  We find that when combined with the
 time slice decomposition, the number of basis templates \numsvdtmps\ is much

### 2011-06-17 method.tex
Tense agreement
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1098 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 For time slices $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ \flops\ 
-for evaluating the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ for the 
+to evaluate the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ to apply the 
 linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops\ to add the resultant partial \SNR\ stream.

### 2011-06-17 implementation.tex
fixed acronym spacing
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1065 ccdd05d5-56c1-461a-bf98-85f6876523b4
Moved sentence from decimation section to basis filters section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1099 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added 'queues'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1100 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the \gstreamer\ pipeline splits into several parallel branches to handle each
-time slice.  Next, these decimated data streams are fed into parallel \fir\ filter banks.
+time slice.
 
@@
 
+Next, these decimated data streams are fed into parallel \fir\ filter banks.
 The basis filters are implemented using a \gstlal\ element called {\tt
@@
 from the nonzero samples of the time slices in equation~\eqref{eq:time-sliced-templates},
-with the zero-padding effectively accomplished by appropriate delays.
+with the zero-padding effectively accomplished by appropriate queues, or delay lines.
 
@@
 
-From the outputs of the \SVD\ basis filters, we form the partial \SNR streams for
-each time  slice by multiplying by the reconstruction matrices.  This is accomplished
-with the \texttt{gstlal} element \texttt{lal\_matrixmixer}.
+From the outputs of the \SVD\ basis filters, we form the partial \SNR\ streams
+for each time  slice by multiplying by the reconstruction matrices.  This is
+accomplished with the \texttt{gstlal} element \texttt{lal\_matrixmixer}.
 

### 2011-06-17 results.tex
subject-verb agreement
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1102 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed broken table reference
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1103 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The minimum, maximum, and median
-mismatch is shown as a function of \SVD\ tolerance in
+mismatch are shown as functions of \SVD\ tolerance in
 figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the
@@
 \begin{table}
-\caption{Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the sub-bank described in section~\ref{sec:bank-setup}.}
+\caption{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the sub-bank described in section~\ref{sec:bank-setup}.}
 \begin{center}

### 2011-06-17 results.tex
Spelling correction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1104 ccdd05d5-56c1-461a-bf98-85f6876523b4
Reword
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1105 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 interpolation filter length to 16. Table \ref{table:flops} shows the computational cost
-in \flops\ for this sub-bank.  Fort the \FD\ method, an \fft\ block size of $\fftblock = 
+in \flops\ for this sub-bank.  For the \FD\ method, an \fft\ block size of $\fftblock = 
 2 \tmpsamps$ is assumed, resulting in a latency of $\left(\tmpsamps f^0\right)$
@@
 parameter space have similar computational scaling, an entire search could be
-implemented with $\gtrsim$$10$ machines.  The lengths of templates does vary
+implemented with $\gtrsim$$10$ machines.  Template length does vary
 over the parameter space; lower-mass templates are longer and would require

### 2011-06-17 conclusions.tex
Don't mention gstlal in conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1106 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed syntax error
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1107 ccdd05d5-56c1-461a-bf98-85f6876523b4
software -> platform
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1108 ccdd05d5-56c1-461a-bf98-85f6876523b4
should -> could
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1109 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 an inherent latency introduced by the software architecture itself.  We have shown 
-prototype implementation of \lloid{} using the open source signal processing software
-\gstreamer\ and \gstlal. Significant work would have to be done in order to
+prototype implementation of \lloid{} using the open source signal processing platform
+\gstreamer. Significant work would have to be done in order to
 improve upon the latency of our implementation, including tighter integration between
-data acquisition and analysis. This should be considered for third-generation detector
+data acquisition and analysis. This could be considered for third-generation detector
 design.

### 2011-06-17 data/localization_uncertainty.py
Adding Horizon computation.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1110 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -62,6 +62,19 @@ rho_threshold = 8
 rho2_ligo = numpy.cumsum(f0_weights * invS_ligo)
 rho2_virgo = numpy.cumsum(f0_weights * invS_virgo)
 
+def horizon(Mc, rho2):
+	LAL_C = 299792458.
+	LAL_PI = 3.1415926535897932384626433832795029
+	LAL_MTSUN_SI = 4.9254909500000001e-06
+	LAL_PC_SI = 3.0856775807e16
+	Mc = Mc * LAL_MTSUN_SI	
+	D = 2. * LAL_C * (5./96.)**.5 * Mc**(5./6.) * LAL_PI**(-2./3.) * rho2**.5 / 8.
+	return D / 1e6 / LAL_PC_SI
+
+# Horizon
+H_ligo = horizon(mchirp, rho2_ligo[-1])
+H_virgo = horizon(mchirp, rho2_virgo[-1])
+
 # Scale everything for an SNR of 10 in LIGO detectors
 #factor = 1 / rho_ligo[-1]
 invS_ligo *= 1 / rho2_ligo[-1]
@@ -108,8 +121,8 @@ for t_before_merger in [25., 10, 5., 1., 0.1, 0]:
 
 	a = a90best[t >= t_before_merger][-1] * (180 / pi) ** 2
 
-	horizon = 400. * (8. / rho_final_ligo)
-	rate = 40. * (horizon / 400.) ** 3
+	horizon = H_ligo * (8. / rho_final_ligo)
+	rate = 40. * (horizon / H_ligo) ** 3
 	print r"%.1f & %.0f & %d & %.1f & %.1f & %.1f \\" % (t_before_merger, horizon, round(rate), rho_final_ligo, a, a_final)
 print r"\tableline"
 print r"\end{tabular}"

### 2011-06-17 implementation.tex
again
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1101 ccdd05d5-56c1-461a-bf98-85f6876523b4
minimum match -> minimal match
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1112 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 source parameters with a hexagonal grid~\citep{PhysRevD.76.102004} in order to
-satisfy a minimum match criterion.  This assures a prescribed maximum loss in
+satisfy a minimal match criterion.  This assures a prescribed maximum loss in
 \SNR\ for signals whose parameters do not lie on the hexagonal grid.  Next, the
@@
 If the next time slice does not have the same sample rate, its output must
-first be interpolated.  The \gstreamer\ element {\tt audioresample} is used here.
+first be interpolated.  The \gstreamer\ element {\tt audioresample} is used again here.
 

### 2011-06-17 method.tex
minimum match -> minimal match
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1112 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 to inspiral template banks greatly reduces the number of filters required to achieve a
-particular minimum match.  A similar technique can be applied to the time-sliced
+particular minimal match.  A similar technique can be applied to the time-sliced
 templates as defined in equation~\ref{eq:time-sliced-templates} above.  The \SVD\

### 2011-06-17 method.tex
Merged two paragraphs in section 2.2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1113 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In any given interval of time, the filters are band-limited across the entire
-template bank. Usually the template is truncated at some prescribed time $t^0$,
-or equivalently frequency $f_\mathrm{hi}$, often chosen to correspond to the
-\ISCO. An inspiral signal will enter the detection band at some low frequency,
-$f = f_\mathrm{low}$, corresponding to a time $t_\mathrm{low}$.  The end of the
-template is then critically sampled at a rate of $2 f_\mathrm{hi}$ but the 
-beginning of the template is critically sampled at $2 f_\mathrm{low}$. In any
-time interval smaller than the duration of the template, the bandwidth of the
-filters across the entire template bank may be significantly less and
-a lower sampling rate may be used for all filters without aliasing.
+An inspiral signal will enter the detection band at some low frequency
+$f_\mathrm{low}$, at a time $t_\mathrm{low}$ before merger.  Usually the template
+is truncated at some prescribed time $t^0$, or equivalent frequency $f_\mathrm{hi}$,
+often chosen to correspond to the \ISCO. The beginning of the template is critically
+sampled at $2 f_\mathrm{low}$, but the end of the template is critically sampled at a
+rate of $2 f_\mathrm{hi}$. In any time interval smaller than the duration of the template,
+the bandwidth of the filters across the entire template bank may be significantly less
+than the full sample rate at which data is acquired.
 

### 2011-06-17 method.tex
Added asymptotic expression for fd flop/s
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1115 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+In the limit of many templates, $M \gg 1$, we may neglect the cost of the forward
+transform of the data and of the multiplication of the transforms.  By making the common
+choice of an \fft\ block size of twice the length of the filter,
+$\fftblock = 2 \tmpsamps$, this simplifies to $\approx 4 f^0 \numtmps \lg \fftblock$
+\flops.
+
 \subsubsection{\lloid\ method}

### 2011-06-17 method.tex
Revised asymptotic quantity
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1116 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-17 method.tex
@@
 
+In the limit of many templates, $M \gg 1$, we may neglect the cost of the forward
+transform of the data and of the multiplication of the transforms.  The computational
+cost will reach an optimum at some large but finite \fft\ block size
+$\fftblock \gg \tmpsamps$.  In this limit, the \FD\ method costs
+$\approx 2 f^0 \numtmps \lg \fftblock$ \flops.
+
 \subsubsection{\lloid\ method}

### 2011-06-17 method.tex
Asymptotic expression for LLOID
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1117 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \flops.  The second sum is carried out over the set of distinct sample rates rather than over the time slices themselves.
+
+We can simplify this expression quite a bit by taking some limits that arise from sensible filter design.
+In the limit of many templates, the cost of the decimation filters is negligible as compared to the cost of the interpolation filters.
+Typically, we will design the interpolation filters with $N^\shortuparrow \lesssim \numsvdtmps$ so that the interpolation cost itself is negligible compared with the reconstruction cost.
+Finally, if the number of basis templates per time slices $\numsvdtmps$ is not too small, the reconstruction cost dominates over the cost of accumulating the partial \SNR.
+In these limits, the cost of \lloid\ is dominated by the basis filters themselves and the reconstruction, totaling $2 \sum_{s=0}^{S-1} f^s \numsvdtmps \left( \slicessamps + \numtmps \right)$ \flops.

### 2011-06-18 introduction.tex
Adding Horizon computation.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1110 ccdd05d5-56c1-461a-bf98-85f6876523b4
Adjusting Table 1 location.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1111 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: Telescopic -> EM
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1114 ccdd05d5-56c1-461a-bf98-85f6876523b4
Adding localization figure
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1118 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\begin{figure}[h]
+\plotone{figures/snr_in_time.pdf}
+\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
+sources that will be detectable by Advanced \LIGO\ $t$ seconds before coalescence.
+The heavy solid line is the most realistic yearly rate estimate.  The shaded
+region represents the 5 to 95\% confidence interval arising from uncertainty
+in predicted event rates \citep{Abadie:2010p10836}.}
+\end{figure}
+%
 It has been observed that by examining the signal to noise ratio (\SNR{}) stream for
@@
 within 25~s of merger. These rates are shown with their substantial uncertainties in figure~\ref{fig:earlywarning}.
-%
-\begin{table}
-\caption{\label{table:sky-localization-accuracy}Horizon distance, expected detection rate, and \SNR\ at merger for a source that produces an \SNR\ of 8 in all three detectors at time $t$ before merger.
+
+\begin{table}[h]
+\caption{\label{table:sky-localization-accuracy}Horizon distance, expected detection rate, and \SNR\ at merger for sources that produce an \SNR\ of 8 in all three detectors at time $t$ before merger.
 The fifth and sixth columns show the minimum area of the 90\% confidence region before merger and at merger.}
@@
 \tableline
-25.0 & 203 & 5 & 15.8 & 2893.2 & 1.8 \\
-10.0 & 269 & 12 & 11.9 & 1150.6 & 3.6 \\
-5.0 & 307 & 18 & 10.4 & 594.6 & 5.1 \\
-1.0 & 360 & 29 & 8.9 & 153.9 & 7.5 \\
-0.1 & 391 & 37 & 8.2 & 32.5 & 9.1 \\
-0.0 & 400 & 40 & 8.0 & 9.6 & 9.6 \\
+25.0 & 226 & 5 & 15.8 & 2893.2 & 1.8 \\
+10.0 & 299 & 12 & 11.9 & 1150.6 & 3.6 \\
+5.0 & 341 & 18 & 10.4 & 594.6 & 5.1 \\
+1.0 & 401 & 29 & 8.9 & 153.9 & 7.5 \\
+0.1 & 435 & 37 & 8.2 & 32.5 & 9.1 \\
+0.0 & 445 & 40 & 8.0 & 9.6 & 9.6 \\
 \tableline
@@
 \begin{figure}[h]
-\plotone{figures/snr_in_time.pdf}
-\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
-sources that will be detectable by Advanced \LIGO\ $t$ seconds before coalescence.
-The heavy solid line is the most realistic yearly rate estimate.  The shaded
-region represents the 5 to 95\% confidence interval arising from uncertainty
-in predicted event rates \citep{Abadie:2010p10836}.}
+\plotone{figures/loc_in_time.pdf}
+\caption{}
 \end{figure}
-
-Telescopic followup requires estimating the location of the \GW{} source. The
+%
+\EM{} followup requires estimating the location of the \GW{} source. The
 localization uncertainty is dominated by uncertainty in the time of arrival of

### 2011-06-18 localization_uncertainty.py

@@ -2,6 +2,8 @@
 
 import numpy
 from numpy import pi
+import pylab
+import sys
 
 def float_as_string(num, sigfigs = 2):
 	"""Convert a floating point number to a string in scientific notation,
@@ -20,8 +22,8 @@ def float_as_string(num, sigfigs = 2):
 				sci_str_parts[1].lstrip('+'))
 
 
-f_ligo, a_ligo = numpy.loadtxt('ZERO_DET_high_P.txt').T
-f_virgo, a_virgo = numpy.loadtxt('AdV_baseline_sensitivity_12May09.txt').T
+f_ligo, a_ligo = numpy.loadtxt('data/ZERO_DET_high_P.txt').T
+f_virgo, a_virgo = numpy.loadtxt('data/AdV_baseline_sensitivity_12May09.txt').T
 
 # Component masses (in M_sun)
 m1 = 1.4
@@ -126,3 +128,37 @@ for t_before_merger in [25., 10, 5., 1., 0.1, 0]:
 	print r"%.1f & %.0f & %d & %.1f & %.1f & %.1f \\" % (t_before_merger, horizon, round(rate), rho_final_ligo, a, a_final)
 print r"\tableline"
 print r"\end{tabular}"
+
+
+# exceptional event is 1 per year
+horizon = H_ligo / 40. ** (1./3.)
+rho_threshold = H_ligo * (8. / horizon)
+
+# SNR
+rho_ligo = rho2_ligo ** .5
+rho_virgo = rho2_virgo ** .5
+
+# Timing uncertainty
+sigmat_ligo = 1. / (2 * pi * rho_threshold * rho_ligo * sigmaf_ligo)
+sigmat_virgo = 1. / (2 * pi * rho_threshold * rho_virgo * sigmaf_virgo)
+
+# Uncertainty in direction cosines in plane of detector
+sigmax = sigmat_ligo / 7e-3
+sigmay = numpy.sqrt((2 * sigmat_virgo ** 2 + sigmat_ligo ** 2) / 3.) / 22e-3
+
+# Minimum area of 90% confidence region
+a90best = 2 * pi * numpy.log(10) * sigmax * sigmay * (180 / pi) ** 2
+
+fig = pylab.figure(figsize=(3,2))
+ax = fig.add_subplot(1,1,1, adjustable='box')
[…]

### 2011-06-18 method.tex
Added speedup
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1119 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
-f = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
+f(t) = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
 \right]^{3/8}.
@@
 In these limits, the cost of \lloid\ is dominated by the basis filters themselves and the reconstruction, totaling $2 \sum_{s=0}^{S-1} f^s \numsvdtmps \left( \slicessamps + \numtmps \right)$ \flops.
+
+If the cost of the basis filters dominates, and the frequency of the templates evolves slowly enough in time, then we can use the time-frequency relationship of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional \TD\ method.
+The reduction in \flops\ is approximately
+%
+\begin{multline}
+\label{eq:speedup}
+\frac{2 \sum_{s=0}^{S-1} f^s \numsvdtmps \slicessamps}{2 \numtmps \tmpsamps f^0} \\
+\approx \frac{\alpha}{\left(t_\mathrm{low} - t_\mathrm{hi}\right) \left(f^0\right)^2} \int_{t_\mathrm{low}}^{t_\mathrm{hi}} \left(2 f(t) \right)^2 \, \mathrm{d} t \\
+= \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}^2) - t_\mathrm{hi} f^2 (t_\mathrm{hi}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{hi}\right)}
+\end{multline}
+%
+where $\alpha \approx \numsvdtmps / \numtmps$ is the rank reduction factor, or ratio between the number of basis templates and the number of templates.
+This approximation assumes that the frequency of the signal is evolving very slowly so that we can approximate the time slice sample rate as twice the instantaneous gravitational wave frequency, $f^s \approx 2 f(t)$, and the number of samples in the decimated time slice as the sample rate times an infinitesimally short time interval, $\slicessamps \approx 2 f(t) \, \mathrm{d}t$.
+Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 1000$~s, $f_\mathrm{low} = 10$~Hz, $t_\mathrm{hi} = 1/1570$~s $f_\mathrm{hi} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires a factor $\sim 10^{-6}$ fewer \flops\ than the conventional \TD\ method.

### 2011-06-18 method.tex
Got rid of sigfigs
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1120 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 This approximation assumes that the frequency of the signal is evolving very slowly so that we can approximate the time slice sample rate as twice the instantaneous gravitational wave frequency, $f^s \approx 2 f(t)$, and the number of samples in the decimated time slice as the sample rate times an infinitesimally short time interval, $\slicessamps \approx 2 f(t) \, \mathrm{d}t$.
-Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 1000$~s, $f_\mathrm{low} = 10$~Hz, $t_\mathrm{hi} = 1/1570$~s $f_\mathrm{hi} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires a factor $\sim 10^{-6}$ fewer \flops\ than the conventional \TD\ method.
+Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{hi} = 1/1570$~s $f_\mathrm{hi} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires a factor $\sim 10^{-6}$ fewer \flops\ than the conventional \TD\ method.

### 2011-06-18 localization_uncertainty.py
Made figure 2 semilogy
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1121 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -151,12 +151,13 @@ a90best = 2 * pi * numpy.log(10) * sigmax * sigmay * (180 / pi) ** 2
 
 fig = pylab.figure(figsize=(3,2))
 ax = fig.add_subplot(1,1,1, adjustable='box')
-ax.loglog(t[rho_ligo >= 8. / rho_threshold], a90best[rho_ligo >= 8. / rho_threshold], 'k')
+ax.semilogy(t[rho_ligo >= 8. / rho_threshold], a90best[rho_ligo >= 8. / rho_threshold], 'k')
 for rho in [8., 12., 16., 20.]:
-	ax.loglog(t[rho_ligo >= rho / rho_threshold][0], a90best[rho_ligo >= rho / rho_threshold][0], 'k+')
-ax.set_xlim(1e-1, 1e2)
+	ax.semilogy(t[rho_ligo >= rho / rho_threshold][0], a90best[rho_ligo >= rho / rho_threshold][0], 'k+')
 ax.axhline(a90best[-1], color='k', linestyle='--')
+pylab.xlim(0, t[rho_ligo >= 8 / rho_threshold][0])
 ax.invert_xaxis()
+pylab.grid()
 pylab.ylabel(r'$A$(90%) (deg$^2$)')
 pylab.xlabel(r'time before coalescence, $t$ (s)')
 pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.2,right=0.95)

### 2011-06-20 introduction.tex
Adding caption and discussion of Figure 2.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1122 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \plotone{figures/loc_in_time.pdf}
-\caption{}
+\caption{\label{fig:sky-localization-accuracy} Area of the 90\% confidence
+region as a function of time for an exceptional source, where such a source is
+defined to produce a signal with an \SNR\ of at least 27.4. These signals are
+expected to be observed in the detectors once per year.  80~s, 35~s, 16~s, and
+6~s before coalescence, such a signal will have accumulated an \SNR\ of 8, 12,
+16, and 20, respectively. After merger, the minimum 90\% confidence area
+shrinks to $0.82$~deg$^2$, shown by the dashed line.}
 \end{figure}
-%
+
 \EM{} followup requires estimating the location of the \GW{} source. The
@@
 \citep{Fairhurst2009}.  Table~\ref{table:sky-localization-accuracy} shows the
-minimum 90\% confidence area for a source that produces an \SNR\ of 8 in the
-\LIGO\ detectors in Hanford, Washington, and Livingston, Louisiana, and, the
-Virgo detector in Cascina, Italy, at a time $t$ before merger.  Also shown are
-the final \SNR\ and 90\% confidence area at merger for the same source.
-At 10~s before merger, the pre-merger 90\% confidence area is about $1150.6$~deg$^2$
-(2.8\% of the sky) and at 1~s before merger is 153.9~deg$^2$ (0.37\% of the sky).
-The confidence region shrinks quadratically in area as the overall amplitude of the
-signal increases.
+minimum 90\% confidence area for sources that produces signals with an \SNR\ of
+8 in the \LIGO\ detectors in Hanford, Washington, and Livingston, Louisiana,
+and, the Virgo detector in Cascina, Italy, at a time $t$ before merger.  Also
+shown are the final \SNR\ and 90\% confidence area at merger for the same
+source.  For a source that accumulates an \SNR\ of 8 at a time 10~s before
+merger, the pre-merger 90\% confidence area is about $1150.6$~deg$^2$ (2.8\% of
+the sky) whereas a source that accumulates an \SNR\ of 8 at a time 1~s before
+merger is 153.9~deg$^2$ (0.37\% of the sky).  The confidence region shrinks
+quadratically in area as the overall amplitude of the signal increases.
+Figure~\ref{fig:sky-localization-accuracy} tracks the minimum 90\% confidence
+area as a function of time before coalescence for an \emph{exceptional source},
+which is defined to be a source that produces a signal with an \SNR\ of at
+least 27.4. Such sources are expected to be seen by the detectors on average
+once per year. These sources would have minimum 90\% confidence areas of
+$2250$~deg$^2$ and $270$~deg$^2$ at a time 30~s and 10~s before coalescence,
+respectively.  After merger, the minimum 90\% confidence area shrinks to
+$0.82$~deg$^2$.
 

### 2011-06-20 localization_uncertainty.py
Changing thickness of line in figure 2.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1123 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -151,7 +151,7 @@ a90best = 2 * pi * numpy.log(10) * sigmax * sigmay * (180 / pi) ** 2
 
 fig = pylab.figure(figsize=(3,2))
 ax = fig.add_subplot(1,1,1, adjustable='box')
-ax.semilogy(t[rho_ligo >= 8. / rho_threshold], a90best[rho_ligo >= 8. / rho_threshold], 'k')
+ax.semilogy(t[rho_ligo >= 8. / rho_threshold], a90best[rho_ligo >= 8. / rho_threshold], 'k', linewidth=2)
 for rho in [8., 12., 16., 20.]:
 	ax.semilogy(t[rho_ligo >= rho / rho_threshold][0], a90best[rho_ligo >= rho / rho_threshold][0], 'k+')
 ax.axhline(a90best[-1], color='k', linestyle='--')

### 2011-06-20 introduction.tex
lloid_paper: trim and rework the EM followup section anticipating that table 1 will merge into Figure 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1124 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-18 introduction.tex
@@
 the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}.  Table~\ref{table:sky-localization-accuracy} shows the
-minimum 90\% confidence area for a source that produces an \SNR\ of 8 in the
-\LIGO\ detectors in Hanford, Washington, and Livingston, Louisiana, and, the
-Virgo detector in Cascina, Italy, at a time $t$ before merger.  Also shown are
-the final \SNR\ and 90\% confidence area at merger for the same source.
-At 10~s before merger, the pre-merger 90\% confidence area is about $1150.6$~deg$^2$
-(2.8\% of the sky) and at 1~s before merger is 153.9~deg$^2$ (0.37\% of the sky).
-The confidence region shrinks quadratically in area as the overall amplitude of the
-signal increases.
+\citep{Fairhurst2009}. Figure~\ref{fig:sky-localization-accuracy} estimates the
+minimum 90\% confidence area versus time for a few exceptionally loud
+coalescence events detectable by Advanced \LIGO{} and Advanced Virgo. A
+once-per-year event would have minimum 90\% confidence areas of $2250$~deg$^2$
+($X$\% of the sky) and $270$~deg$^2$ ($Y$\% of the sky) at times 30~s and 10~s
+before coalescence, respectively. The confidence region shrinks quadratically in
+area as the overall amplitude of the signal increases. By the time of merger,
+the minimum 90\% confidence area shrinks to $0.82$~deg$^2$.
 
-It is infeasible to search this large an area for a prompt counterpart. For
+It is infeasible to search this hundreds of square degrees for a prompt
+counterpart. For
 reference, \textsc{lsst} should see first light around the same time as
@@
 catalog will not be as useful for downselecting pointings generally. However,
-sources that are detectable before merger are likely to be exceptionally
-nearby. Thus there will be fewer galaxies to consider for a given
-early-warning candidate and catalog completeness will be less of a concern.
+our exceptional sources will necessarily be extremely nearby. Within this
+reduced volume there will be fewer galaxies to consider for a given
+candidate and catalog completeness will be less of a concern.
 This should allow us to shrink the 90\% confidence area substantially.
-Even so, it will be desirable to mobilize a large number of heterogeneous
-telescopes in order to guarantee imaging this exceptional event at the instant
-of merger. Because the sky localization will improve rapidly until merger,
-by the time that the entire \CBC{} event is over it will be
-possible to send a second alert to downselect the images that have already been
-taken as well as repoint telescopes to more likely locations.
 

### 2011-06-21 localization_uncertainty.py
Changing variable name in horizon function
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1126 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -64,13 +64,13 @@ rho_threshold = 8
 rho2_ligo = numpy.cumsum(f0_weights * invS_ligo)
 rho2_virgo = numpy.cumsum(f0_weights * invS_virgo)
 
-def horizon(Mc, rho2):
+def horizon(Mc, hdoth):
 	LAL_C = 299792458.
 	LAL_PI = 3.1415926535897932384626433832795029
 	LAL_MTSUN_SI = 4.9254909500000001e-06
 	LAL_PC_SI = 3.0856775807e16
 	Mc = Mc * LAL_MTSUN_SI	
-	D = 2. * LAL_C * (5./96.)**.5 * Mc**(5./6.) * LAL_PI**(-2./3.) * rho2**.5 / 8.
+	D = 2. * LAL_C * (5./96.)**.5 * Mc**(5./6.) * LAL_PI**(-2./3.) * hdoth**.5 / 8.
 	return D / 1e6 / LAL_PC_SI
 
 # Horizon

### 2011-06-21 introduction.tex
lloid_paper: one more typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1125 ccdd05d5-56c1-461a-bf98-85f6876523b4
Updated figure 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1127 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \plotone{figures/loc_in_time.pdf}
-\caption{\label{fig:sky-localization-accuracy} Area of the 90\% confidence
-region as a function of time for an exceptional source, where such a source is
-defined to produce a signal with an \SNR\ of at least 27.4. These signals are
-expected to be observed in the detectors once per year.  80~s, 35~s, 16~s, and
-6~s before coalescence, such a signal will have accumulated an \SNR\ of 8, 12,
-16, and 20, respectively. After merger, the minimum 90\% confidence area
-shrinks to $0.82$~deg$^2$, shown by the dashed line.}
+\caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
+region as a function of time for a sources with anticipated detection rates
+of 10~yr$^{-1}$, 1~yr$^{-1}$, and 0.1~yr$^{-1}$. The heavy dot indicates the
+time at which the accumulated \SNR\ exceeds a threshold of 8.}
 \end{figure}
@@
 
-It is infeasible to search this hundreds of square degrees for a prompt
+It is infeasible to search hundreds of square degrees for a prompt
 counterpart. For

### 2011-06-21 localization_uncertainty.py

@@ -25,6 +25,13 @@ def float_as_string(num, sigfigs = 2):
 f_ligo, a_ligo = numpy.loadtxt('data/ZERO_DET_high_P.txt').T
 f_virgo, a_virgo = numpy.loadtxt('data/AdV_baseline_sensitivity_12May09.txt').T
 
+# Some constants
+LAL_C = 299792458.
+LAL_PI = 3.1415926535897932384626433832795029
+LAL_MTSUN_SI = 4.9254909500000001e-06
+LAL_PC_SI = 3.0856775807e16
+
+
 # Component masses (in M_sun)
 m1 = 1.4
 m2 = 1.4
@@ -42,7 +49,6 @@ fLOW = 10
 fISCO = 4400 / M
 
 def freq_to_time(Mc,f):
-	LAL_MTSUN_SI = 4.9254909500000001e-06
 	Mc = Mc * LAL_MTSUN_SI
 	return 5. * Mc / 256. * (pi * Mc * f) ** (-8./3.)
 
@@ -65,24 +71,23 @@ rho2_ligo = numpy.cumsum(f0_weights * invS_ligo)
 rho2_virgo = numpy.cumsum(f0_weights * invS_virgo)
 
 def horizon(Mc, hdoth):
-	LAL_C = 299792458.
-	LAL_PI = 3.1415926535897932384626433832795029
-	LAL_MTSUN_SI = 4.9254909500000001e-06
-	LAL_PC_SI = 3.0856775807e16
 	Mc = Mc * LAL_MTSUN_SI	
 	D = 2. * LAL_C * (5./96.)**.5 * Mc**(5./6.) * LAL_PI**(-2./3.) * hdoth**.5 / 8.
 	return D / 1e6 / LAL_PC_SI
 
-# Horizon
-H_ligo = horizon(mchirp, rho2_ligo[-1])
-H_virgo = horizon(mchirp, rho2_virgo[-1])
-
-# Scale everything for an SNR of 10 in LIGO detectors
-#factor = 1 / rho_ligo[-1]
+# Scale everything for an SNR of 1 in each detector (just to prevent floating point overflow)
 invS_ligo *= 1 / rho2_ligo[-1]
 invS_virgo *= 1 / rho2_virgo[-1]
 rho2_ligo *= 1 / rho2_ligo[-1]
 rho2_virgo *= 1 / rho2_virgo[-1]
+
+# SNR
[…]

### 2011-06-21 localization_uncertainty.py
Upper limit on plot should be total square degrees of the sky.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1128 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -24,6 +24,7 @@ def float_as_string(num, sigfigs = 2):
 
 f_ligo, a_ligo = numpy.loadtxt('data/ZERO_DET_high_P.txt').T
 f_virgo, a_virgo = numpy.loadtxt('data/AdV_baseline_sensitivity_12May09.txt').T
+f_ET, a_ET = numpy.loadtxt('data/ET_D_data.txt', usecols=(0,3)).T
 
 # Some constants
 LAL_C = 299792458.
@@ -43,7 +44,7 @@ M = m1 + m2
 mchirp = (m1 * m2)**.6 / (m1+m2)**.2
 
 # Low frequency cutoff
-fLOW = 10
+fLOW = 2
 
 # ISCO frequency (equation 3.6, FINDCHIRP paper)
 fISCO = 4400 / M
@@ -148,7 +149,7 @@ for rate in (10., 1., 0.1):
 	pylab.loglog(t[pred], a90[pred], 'k')
 	pylab.text(.8*t[-1], a90[-1], r"%g yr$^{-1}$" % rate, {"size": 8.}, horizontalalignment='left', verticalalignment='center')
 pylab.xlim(0., 1000.)
-pylab.ylim(1e-1, 1e6)
+pylab.ylim(1e-1, 41253)
 ax.invert_xaxis()
 pylab.grid()
 pylab.ylabel(r'$A$(90%) (deg$^2$)')

### 2011-06-21 localization_uncertainty.py
File not checked in
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1129 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -24,7 +24,7 @@ def float_as_string(num, sigfigs = 2):
 
 f_ligo, a_ligo = numpy.loadtxt('data/ZERO_DET_high_P.txt').T
 f_virgo, a_virgo = numpy.loadtxt('data/AdV_baseline_sensitivity_12May09.txt').T
-f_ET, a_ET = numpy.loadtxt('data/ET_D_data.txt', usecols=(0,3)).T
+# f_ET, a_ET = numpy.loadtxt('data/ET_D_data.txt', usecols=(0,3)).T # not checked in
 
 # Some constants
 LAL_C = 299792458.

### 2011-06-22 introduction.tex
Updated table 1, figure 2, and their captions
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1130 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{table}[h]
-\caption{\label{table:sky-localization-accuracy}Horizon distance, expected detection rate, and \SNR\ at merger for sources that produce an \SNR\ of 8 in all three detectors at time $t$ before merger.
-The fifth and sixth columns show the minimum area of the 90\% confidence region before merger and at merger.}
+\caption{\label{table:sky-localization-accuracy}Horizon distance, \SNR\ at merger, and area of 90\% confidence at selected times before merger for sources with expected detection rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
 \begin{center}
-\begin{tabular}{rrrrrr}
+\begin{tabular}{rrrrrrr}
 \tableline\tableline
-& horiz. & rate & final & \multicolumn{2}{c}{$A$(90\%) (deg$^2$)} \\
-\cline{5-6}
-$t$ (s) & (Mpc) & (yr$^{-1}$) & \SNR\ & early & final \\
+rate & horiz. & final & \multicolumn{4}{c}{$A$(90\%) (deg$^2$)} \\
+\cline{4-7}
+yr$^{-1}$ & (Mpc) & \SNR\ & 25 s & 10 s & 1 s & 0 s \\
 \tableline
-25.0 & 226 & 5 & 15.8 & 2893.2 & 1.8 \\
-10.0 & 299 & 12 & 11.9 & 1150.6 & 3.6 \\
-5.0 & 341 & 18 & 10.4 & 594.6 & 5.1 \\
-1.0 & 401 & 29 & 8.9 & 153.9 & 7.5 \\
-0.1 & 435 & 37 & 8.2 & 32.5 & 9.1 \\
-0.0 & 445 & 40 & 8.0 & 9.6 & 9.6 \\
+40\phd\phn & 445 & 8.0 & 15500 & 3100 & 200 & 9.6 \\
+10\phd\phn & 280 & 12.7 & 6100 & 1200 & 78 & 3.8 \\
+1\phd\phn & 130 & 27.4 & 1300 & 260 & 17 & 0.8 \\
+0.1 & 60 & 58.9 & 280 & 56 & 3.6 & 0.2 \\
 \tableline
@@
 region as a function of time for a sources with anticipated detection rates
-of 10~yr$^{-1}$, 1~yr$^{-1}$, and 0.1~yr$^{-1}$. The heavy dot indicates the
+of 40, 10, 1, and 0.1~yr$^{-1}$. The heavy dot indicates the
 time at which the accumulated \SNR\ exceeds a threshold of 8.}

### 2011-06-22 localization_uncertainty.py

@@ -73,8 +73,12 @@ rho2_virgo = numpy.cumsum(f0_weights * invS_virgo)
 
 def horizon(Mc, hdoth):
 	Mc = Mc * LAL_MTSUN_SI	
-	D = 2. * LAL_C * (5./96.)**.5 * Mc**(5./6.) * LAL_PI**(-2./3.) * hdoth**.5 / 8.
+	D = 2. * LAL_C * (5./96.)**.5 * Mc**(5./6.) * LAL_PI**(-2./3.) * hdoth**.5 / 8
 	return D / 1e6 / LAL_PC_SI
+
+# Horizon
+H_ligo = horizon(mchirp, rho2_ligo[-1])
+# H_virgo = horizon(mchirp, rho2_virgo[-1]) # not used
 
 # Scale everything for an SNR of 1 in each detector (just to prevent floating point overflow)
 invS_ligo *= 1 / rho2_ligo[-1]
@@ -85,10 +89,6 @@ rho2_virgo *= 1 / rho2_virgo[-1]
 # SNR
 rho_ligo = numpy.sqrt(rho2_ligo)
 rho_virgo = numpy.sqrt(rho2_virgo)
-
-# Horizon
-H_ligo = horizon(mchirp, 1.) # not used
-# H_virgo = horizon(mchirp, rho2_virgo[-1]) # not used
 
 # Effective bandwidth.
 # Note an omission in Fairhurst (2009): the frequency moments have to be
@@ -110,37 +110,41 @@ a90best = 2 * pi * numpy.log(10) * sigmax * sigmay
 # time
 t = freq_to_time(mchirp, f)
 
-if 0:
-	print r"\begin{tabular}{rrrrrr}"
-	print r"\tableline\tableline"
-	print r"& horiz. & rate & final & \multicolumn{2}{c}{$A$(90\%) (deg$^2$)} \\"
-	print r"\cline{5-6}"
-	print r"$t$ (s) & (Mpc) & (yr$^{-1}$) & \SNR\ & early & final \\"
-	print r"\tableline"
-	for t_before_merger in [25., 10, 5., 1., 0.1, 0]:
-		rho_final_ligo = rho_threshold / numpy.sqrt(rho2_ligo[t >= t_before_merger][-1])
-		rho_final_virgo = rho_threshold / numpy.sqrt(rho2_virgo[t >= t_before_merger][-1])
+def localization_uncertainty_as_str(a90):
+	if a90 >= 4 * 180**2 / numpy.pi:
+		return "-"
[…]

### 2011-06-22 method.tex
Broke speedup into separate subsubsection
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1132 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1133 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\subsubsection{Speedup of \lloid\ relative to \TD\ method}
+
 If the cost of the basis filters dominates, and the frequency of the templates evolves slowly enough in time, then we can use the time-frequency relationship of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional \TD\ method.
@@
 \approx \frac{\alpha}{\left(t_\mathrm{low} - t_\mathrm{hi}\right) \left(f^0\right)^2} \int_{t_\mathrm{low}}^{t_\mathrm{hi}} \left(2 f(t) \right)^2 \, \mathrm{d} t \\
-= \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}^2) - t_\mathrm{hi} f^2 (t_\mathrm{hi}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{hi}\right)}
+= \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}) - t_\mathrm{hi} f^2 (t_\mathrm{hi}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{hi}\right)}
 \end{multline}

### 2011-06-22 conclusions.tex
lloid_paper: update intro text to match figures; tightened intro and conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1134 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\CBC\ events may be the progenitors of some short hard gamma-ray bursts and are
-expected to be accompanied prompt electromagnetic signals.  Rapid
+\CBC\ events may be the progenitors of some short hard \GRB{}s and are
+expected to be accompanied prompt \EM{} signals.  Rapid
 alerts to the broader astronomical community will improve the chances of
-detecting an electromagnetic counterpart in bands from gamma-rays down to
-radio.  The detection algorithm we described has no intrinsic latency.  However, there are
+detecting an \EM{} counterpart in bands from gamma-rays down to
+radio.
+
+The detection algorithm we described has no intrinsic latency.  However, there are
 fundamental and practical latencies associated with the analysis and detection
@@
 design.
+Also possible for third-generation instruments, the \lloid{} method could
+provide the input for a dynamic tuning of detector response via the signal
+recycling mirror to match the frequency of maximum sensitivity to the
+instantaneous frequency of the \GW{} waveform; this challenging technique has
+the potential for vast gains in \SNR{} and timing uncertainty
+\citep{PhysRevD.47.2184}.
 
@@
 to translate the early-warning outputs into candidate events.  The time-slice
-decomposition and the \SVD\ will help us form low-latency signal based vetoes
+decomposition and the \SVD\ will help us form low-latency signal-based vetoes
 (for example, a $\chi^2$ statistic) that will help us here.
 
-Our future work must also address sky localization accuracy prior to merger as well as
-observing strategies.  \citet{Fairhurst2009} derives the area of 90\% localization
-confidence in terms of timing uncertainties, but it would be advantageous to use
-a galaxy catalog to inform the telescope tiling \citep{galaxy-catalog}.  Because
-early detection is likely to be especially effective for nearby sources, the galaxy
-catalog technique might be an important ingredient in reducing the fraction of
-sky that must be imaged.  Extensive injection campaigns and simulations
-incorporating realistic binary merger rates will be necessary in order to fully
-understand the prospects for early-warning detection, localization, and \EM\ folluwup
-using the techniques we have described.
-
-\editorial{nvf: It's a bit pie-in-the-sky, but we could also mention
-reconfiguring the signal recycling mirror to optimize SNR at merger. There's a
-lot of science there.}
+Our future work must more deeply address sky localization accuracy prior to
+merger as well as observing strategies. Here, we have followed
+\citet{Fairhurst2009} in estimating the area of 90\% localization confidence in
+terms of timing uncertainties alone, but it would be advantageous to use a
+galaxy catalog to inform the telescope tiling \citep{galaxy-catalog}. Because
+early detections will arise from nearby sources, the galaxy catalog technique
+might be an important ingredient in reducing the fraction of sky that must be
+imaged.  Extensive simulation campaigns incorporating realistic binary merger
+rates and detector networks will be necessary in order to fully understand the
+prospects for early-warning detection, localization, and \EM\ folluwup using
+the techniques we have described.
 

### 2011-06-22 introduction.tex
lloid_paper: update intro text to match figures; tightened intro and conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1134 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-% Outline:
-%  - Low-mass inspirals are the most promising source of gravitational radiation for advanced \textsc{ligo}
-% - Motivation for high throughput
-%   - Point out challenges of advanced detectors: more templates, longer templates
-% - Motivation for low latency
-%   - Observational applications of near-real-time detection
-%   - Mention S6 and VSR2, online running, EM followup
-% - Layout of paper: description of method, procedure followed
-
-The coalescence of compact binary systems consisting of neutron stars
- (\textsc{ns}) and/or black holes (\textsc{bh}) is presently the best
-understood and most actively pursued potential source of gravitational
-radiation for ground-based gravitational-wave (\GW{}) detectors.
 As a compact binary system loses energy to \GW{}s, its
@@
 the merger and fuel a bright electromagnetic (\EM{})
-counterpart~\citep{shibata:2007}. Observation of an \EM{} counterpart to a
-candidate \GW{} event will vastly boost confidence in the \GW{} detection and
-will often provide sufficient sky localization to determine
-the source's host galaxy and therefore measure the source's redshift. The
-\GW{} observation of this event will also provide a luminosity
-distance measurement. With both quantities, we can
-immediately measure the Hubble constant with unprecedented
-accuracy and without systematics dependent on the standard cosmic distance
-ladder~\citep{2010ApJ...725..496N}. The most efficient means of obtaining
-such joint observations is through the rapid identification and dissemination
-of candidate \GW{} events.
+counterpart \citep{shibata:2007}. One way of obtaining such joint observations
+is through the rapid identification and dissemination of candidate \GW{} events
+for \EM{} followup.
 
@@
 to trigger followup observations by \EM{} telescopes.  The typical latencies
-were 30--60 minutes ~\citep{HugheyGWPAW2011}, which was an important
+were 30--60 minutes \citep{HugheyGWPAW2011}, which was an important
 achievement, but far too late to catch prompt \EM{} emission. We have the
@@
 classes of \EM{} transients, \CBC{}s are in particular believed to be a
-mechanism for short gamma-ray bursts (short \GRB{}s)~\citep{Lee:2005, nakar07}.
+mechanism for short gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}.
 In this scenario, prompt \EM{} emission arises as shells of
@@
 The same inner shocks, or potentially reverse shocks, can produce a bright
-accompanying optical flash~\citep{Sari99}. Prompt emission is a probe into
+accompanying optical flash \citep{Sari99}. Prompt emission is a probe into
 the extreme initial conditions of the outflow, in contrast with afterglows,
@@
 Optical flashes have only been observed for a handful of long
-\GRB{}s~\citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
+\GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
 or, in the case of \textsc{grb 080319b}, by pure serendipity, where several
 telescopes were observing a previous \GRB{} in the same
-field~\citep{2008Natur.455..183R}. The observed optical flashes peaked
+field \citep{2008Natur.455..183R}. The observed optical flashes peaked
 within tens of seconds and decayed quickly.
@@
 prompt optical flashes and the rise of afterglows from short \GRB{}s, and, for
-an exceptional event, perhaps even a glimpse of a
-precursor~\citep{0004-637X-723-2-1711}.
+a truly exceptional event, perhaps even a glimpse of a
+precursor \citep{0004-637X-723-2-1711}.
 
@@
 namely \textsc{mbta}, Coherent WaveBurst, and
-Omega~\citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
+Omega \citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
 achieved latencies of 30--60 minutes, which were dominated by a human vetting
 process. Candidates were sent for \EM{} followup to several
-telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko~\citep{kanner2008,
+telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008,
 HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved
@@
 the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}. Figure~\ref{fig:sky-localization-accuracy} estimates the
-minimum 90\% confidence area versus time for a few exceptionally loud
+\citep{Fairhurst2009}. Table~\ref{fig:earlywarning} and Fig.~\ref{fig:sky-localization-accuracy} estimate the
+minimum 90\% confidence area versus time for a few rare, loud
 coalescence events detectable by Advanced \LIGO{} and Advanced Virgo. A
-once-per-year event would have minimum 90\% confidence areas of $2250$~deg$^2$
-($X$\% of the sky) and $270$~deg$^2$ ($Y$\% of the sky) at times 30~s and 10~s
-before coalescence, respectively. The confidence region shrinks quadratically in
-area as the overall amplitude of the signal increases. By the time of merger,
-the minimum 90\% confidence area shrinks to $0.82$~deg$^2$.
+once-per-year event would have minimum 90\% confidence areas of $1300$~deg$^2$
+($3.1$\% of the sky) and $260$~deg$^2$ ($0.63$\% of the sky) at times 25~s and
+10~s before coalescence, respectively. The confidence region shrinks
+quadratically in area as the overall amplitude of the signal increases. By the
+time of merger, the minimum 90\% confidence area shrinks to $0.82$~deg$^2$.
 

### 2011-06-22 implementation.tex
Added text about buffering, benchmark
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1135 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\gstreamer\ pipelines typically operate with very low (in some consumer applications, imperceptibly low) latency rather than in true real-time because signals are partitioned into blocks of samples, or \emph{buffers}.
+This affords a number of advantages, including amortizing the overhead of passing signals between elements and grouping together sequences of similar operations.
+However, buffering a signal incurs a latency of up to one buffer length.
+This latency can be made small at the cost of some additional overhead by making the buffers sufficiently small.
+
 \paragraph{Decimation}

### 2011-06-22 results.tex
minimum match -> minimal match
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1112 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added text about buffering, benchmark
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1135 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The \SNR\ loss is to be compared with the mismatch of 0.03 that arises from the
-discreteness of template bank designed for a minimum match of 0.97.  We will consider
+discreteness of template bank designed for a minimal match of 0.97.  We will consider
 an acceptable target \SNR\ loss to be a factor of 10 smaller than this, that is, no more
@@
 Presently, the \LIGO{} Data Grid consists of only $\sim$$10^4$ machines.
+
+\subsection{Measured latency and overhead}
+
+Our \gstreamer\ pipeline for measuring impulse responses contained instrumentation that would not be necessary for an actual search, including additional interpolation filters to bring the early-warning outputs back to the full sample rate and additional outputs for recording signals to disk.
+
+We wrote a second, stripped pipeline to evaluate the actual latency and computational overhead.
+We executed this pipeline on one of the submit machines of the \LIGO-Caltech cluster, a Sun Microsystems Sun Fire\texttrademark\ X4600~M2 server with eight quad-core 2.7~GHz AMD Opteron\texttrademark\ 8384 processors.
+This test consumed 78\% of the capacity of just one out of the 32 cores, maintaining a constant latency of 5~s.
+
+The measured overhead is consistent to within an order of magnitude with the lower bound from the \flops\ budget.
+The additional overhead is probably dominated by thread synchronization.
+A carefully optimized \gstreamer\ pipeline or a hand-tuned C implementation of the pipeline might reduce overhead further.
+
+The 5~s latency is probably due to buffering and synchronization.  The latency could probably be reduced by carefully tuning buffer lengths at every stage in the pipeline.

### 2011-06-22 conclusions.tex
timing uncertainty -> timing accuracy
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1137 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 instantaneous frequency of the \GW{} waveform; this challenging technique has
-the potential for vast gains in \SNR{} and timing uncertainty
+the potential for vast gains in \SNR{} and timing accuracy
 \citep{PhysRevD.47.2184}.

### 2011-06-22 conclusions.tex
Split compound sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1138 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 recycling mirror to match the frequency of maximum sensitivity to the
-instantaneous frequency of the \GW{} waveform; this challenging technique has
-the potential for vast gains in \SNR{} and timing accuracy
+instantaneous frequency of the \GW{} waveform.  This is a challenging technique, but
+it has the potential for vast gains in \SNR{} and timing accuracy
 \citep{PhysRevD.47.2184}.

### 2011-06-22 conclusions.tex
Reworded an awkward sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1139 ccdd05d5-56c1-461a-bf98-85f6876523b4
Addressing some issues in conclusion.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1140 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-22 conclusions.tex
@@
 We have demonstrated a computationally feasible procedure for the rapid and
-even early-warning detection of \GW{}s emitted during the coalescence
-of neutron stars and stellar-mass black holes. Our method is as fast as
-standard \fft{} convolutions but allows for latency-free, real-time
-operation.  We anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
-search for binary neutron stars using coincident \GW\ data from a four-detector network.
-This is well within the current computing capabilities of the \LIGO{} Data Grid%
-\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.
+even early-warning detection of \GW{}s emitted during the coalescence of
+neutron stars and stellar-mass black holes. Our method is as fast as standard
+\fft{} convolutions but allows for latency-free, real-time operation.  We
+anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
+search for binary neutron stars using coincident \GW\ data from a four-detector
+network.  In the future, additional computational savings could be acheived by
+conditionally reconstructing the \SNR{} time-series only during times when a
+composite detection statistic crosses a threshold~\citep{svd-compdetstat}.
+However, the anticipated required number of computers is well within the
+current computing capabilities of the \LIGO{} Data
+Grid\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.
 
-\CBC\ events may be the progenitors of some short hard \GRB{}s and are
-expected to be accompanied prompt \EM{} signals.  Rapid
-alerts to the broader astronomical community will improve the chances of
-detecting an \EM{} counterpart in bands from gamma-rays down to
-radio.
+\CBC\ events may be the progenitors of some short hard \GRB{}s and are expected
+to be accompanied prompt \EM{} signals.  Rapid alerts to the broader
+astronomical community will improve the chances of detecting an \EM{}
+counterpart in bands from gamma-rays down to radio.
 
-The detection algorithm we described has no intrinsic latency.  However, there are
-fundamental and practical latencies associated with the analysis and detection
-procedure. For example, the \LIGO{} data acquisition records science data in 1/16~s
-blocks~\citep{Bork2001}. Data must also be aggregated from all of the \GW\
-observatories using global computer networks capable of high bandwidth but perhaps only
-modest latency.  This could introduce a latency of $\sim$100~ms.  Lastly, unless a
-real-time infrastructure is adopted post data acquisition, it is likely that there will be 
-an inherent latency introduced by the software architecture itself.  We have shown 
-prototype implementation of \lloid{} using the open source signal processing platform
-\gstreamer. Significant work would have to be done in order to
-improve upon the latency of our implementation, including tighter integration between
-data acquisition and analysis. This could be considered for third-generation detector
-design.
-Also possible for third-generation instruments, the \lloid{} method could
-provide the input for a dynamic tuning of detector response via the signal
-recycling mirror to match the frequency of maximum sensitivity to the
-instantaneous frequency of the \GW{} waveform; this challenging technique has
-the potential for vast gains in \SNR{} and timing accuracy
-\citep{PhysRevD.47.2184}.
+The detection algorithm we described has no intrinsic latency.  However, there
+are fundamental and practical latencies associated with the analysis and
+detection procedure. For example, the \LIGO{} data acquisition records science
+data in 1/16~s blocks~\citep{Bork2001}. Data must also be aggregated from all
+of the \GW\ observatories using global computer networks capable of high
+bandwidth but perhaps only modest latency.  This could introduce a latency of
+$\sim$100~ms.  Lastly, the software implementation of the algorithm itself may
+introduced latency.  We have shown prototype implementation of \lloid{} using
+the open source signal processing platform \gstreamer. Significant work would
+have to be done in order to improve upon the latency of our implementation,
+including tighter integration between data acquisition and analysis. This could
+be considered for third-generation detector design.  Also possible for
+third-generation instruments, the \lloid{} method could provide the input for a
+dynamic tuning of detector response via the signal recycling mirror to match
+the frequency of maximum sensitivity to the instantaneous frequency of the
+\GW{} waveform.  This is a challenging technique, but it has the potential for
+vast gains in \SNR{} and timing accuracy \citep{PhysRevD.47.2184}.
 
-% This passage goes in the 'future work' part, but I am having trouble working it in.
-\begin{comment}
-In the future, this
-could be improved upon further by conditionally reconstructing the \SNR{} time-series
-only during times when a composite detection statistic crosses a
-threshold~\citep{svd-compdetstat}.
-\end{comment}
-
-Although we have demonstrated a feasible method for advance detection we have
-not yet explored triggering, coincidence, and ranking of gravitational wave
-candidates in a framework that supports early \EM\ followup.  If adjacent time slices
-have non-negligible cross-correlation with each other, then there may be some
-subtleties in the design of these postprocessing stages that we must address in order
-to translate the early-warning outputs into candidate events.  The time-slice
-decomposition and the \SVD\ will help us form low-latency signal-based vetoes
-(for example, a $\chi^2$ statistic) that will help us here.
+Although we have demonstrated a feasible method for advance detection
+\emph{filtering}, we have not yet explored data whitening, triggering,
+coincidence, and ranking of gravitational wave candidates in a framework that
+supports early \EM\ followup.
+% I don't understand what is trying to be said here.  Time-slices have no
+% cross-correlation with eachother since they are orthogonal.
+\begin{comment}If adjacent time slices have non-negligible cross-correlation
+with each other, then there may be some subtleties in the design of these
+postprocessing stages that we must address in order to translate the
+early-warning outputs into candidate events.\end{comment}
+%
+As one possible method of addressing the latter, the time-slice decomposition
+and the \SVD\ will help us form low-latency signal-based vetoes (for example, a
+$\chi^2$ statistic) that have been essential for glitch rejection used in
+previous \GW{} \CBC{} searches.
 

### 2011-06-22 conclusions.tex
Modifying conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1142 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 supports early \EM\ followup.
-% I don't understand what is trying to be said here.  Time-slices have no
+% Drew: I don't understand what is trying to be said here.  Time-slices have no
 % cross-correlation with eachother since they are orthogonal.
@@
 %
-As one possible method of addressing the latter, the time-slice decomposition
-and the \SVD\ will help us form low-latency signal-based vetoes (for example, a
-$\chi^2$ statistic) that have been essential for glitch rejection used in
-previous \GW{} \CBC{} searches.
+However, we believe these obstacles are surmountable.
+% Drew: Do we really want to say this. It doesn't really address any of the
+% issues above.
+\begin{comment}For an example of how the last obstacle could be approached, the
+time-slice decomposition and the \SVD\ will help us form low-latency
+signal-based vetoes (for example, a $\chi^2$ statistic) that have been
+essential for glitch rejection used in previous \GW{} \CBC{}
+searches.\end{comment}
 

### 2011-06-23 conclusions.tex
Restructured second to last paragraph in conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1143 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-06-22 conclusions.tex
@@
 
-Although we have demonstrated a feasible method for advance detection
-\emph{filtering}, we have not yet explored data whitening, triggering,
+Although we have demonstrated a computationally feasible detection statistic for
+advance detection, we have not yet explored data whitening, triggering,
 coincidence, and ranking of gravitational wave candidates in a framework that
-supports early \EM\ followup.
-% I don't understand what is trying to be said here.  Time-slices have no
-% cross-correlation with eachother since they are orthogonal.
-\begin{comment}If adjacent time slices have non-negligible cross-correlation
-with each other, then there may be some subtleties in the design of these
-postprocessing stages that we must address in order to translate the
-early-warning outputs into candidate events.\end{comment}
-%
-As one possible method of addressing the latter, the time-slice decomposition
-and the \SVD\ will help us form low-latency signal-based vetoes (for example, a
-$\chi^2$ statistic) that have been essential for glitch rejection used in
-previous \GW{} \CBC{} searches.
+supports early \EM\ followup.  One complication in deploying a low-latency or
+early-warning search is that whitening the detector data involves estimating the
+noise power spectrum, which is often done in acausally.  We will have to look into
+causal or low latency alternatives to conventional spectral estimation techniques.
+Another potential complication is that if adjacent time slices have non-negligible
+cross-correlation with each other, then there may be some subtleties in the design
+of the triggering and coincidence stages that we must address in order to translate
+the early-warning outputs into candidate events.  The time-slice decomposition and
+the \SVD\ may also be beneficial for the postprocessing stage because they will
+help us form low-latency signal-based vetoes (e.g. $\chi^2$ statistics)
+that have been essential for glitch rejection used in previous \GW{} \CBC{}
+searches.
 

### 2011-06-23 conclusions.tex
Minor corrections to conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1144 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $\sim$100~ms.  Lastly, the software implementation of the algorithm itself may
-introduced latency.  We have shown prototype implementation of \lloid{} using
-the open source signal processing platform \gstreamer. Significant work would
-have to be done in order to improve upon the latency of our implementation,
+introduce latency.  We have shown a prototype implementation of \lloid{} using
+\gstreamer, an open source signal processing platform. Some work could
+be done in order to improve upon the latency of our implementation,
 including tighter integration between data acquisition and analysis. This could

### 2011-06-23 introduction.tex
Reworded sentence in introduction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1145 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In the era of first-generation detectors, the \GW{} community initiated a
+In the first generation of ground-based laser interferometers, the \GW{}
+community initiated a
 project to send alerts when potential \GW{} transients were observed in order

### 2011-06-23 introduction.tex
Improved wording of sentence in section 2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1146 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 or, in the case of \textsc{grb 080319b}, by pure serendipity, where several
-telescopes were observing a previous \GRB{} in the same
-field \citep{2008Natur.455..183R}. The observed optical flashes peaked
+telescopes were already observing the afterglow of another \GRB{} in the same
+field of view~\citep{2008Natur.455..183R}. The observed optical flashes peaked
 within tens of seconds and decayed quickly.

### 2011-06-23 introduction.tex
Added URL footnotes for detectors
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1147 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In October 2010, \LIGO{} completed its sixth science run
-(S6) and Virgo completed its third science run (VSR3).  While both
-\LIGO{} detectors and Virgo were operating, several all-sky detection
+In October 2010, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its sixth science run
+(S6) and Virgo\footnote{\url{http://www.ego-gw.it/}} completed its third science run (VSR3).
+While both \LIGO{} detectors and Virgo were operating, several all-sky detection
 pipelines operated in a low-latency configuration to send astronomical alerts,

### 2011-06-23 introduction.tex
Minor rewording in section 1.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1148 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008,
-HugheyGWPAW2011} took images of likely sky locations.  \textsc{mbta} achieved
+HugheyGWPAW2011} imaged some of the most likely sky locations.  \textsc{mbta} achieved
 the best \GW{} trigger-generation latencies of 2--5 minutes.  We

### 2011-06-23 introduction.tex
Removed unnecessary variable (t)
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1149 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
-sources that will be detectable by Advanced \LIGO\ $t$ seconds before coalescence.
+sources that will be detectable by Advanced \LIGO\ a given number of seconds before coalescence.
 The heavy solid line is the most realistic yearly rate estimate.  The shaded

### 2011-06-23 snr_in_time.py

@@ -97,7 +97,7 @@ for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt'], ['zero det'
 pylab.grid()
 #pylab.legend(loc='lower left')
 pylab.ylabel(r'detections yr$^{-1}$')
-pylab.xlabel(r'time before coalescence, $t$ (s)')
+pylab.xlabel(r'time before coalescence (s)')
 pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.2,right=0.95)
 pylab.xlim([0.01,1000])
 pylab.ylim([.1, 1000])

### 2011-06-23 introduction.tex
Reworded a paragraph in section 1.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1150 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}. Table~\ref{fig:earlywarning} and Fig.~\ref{fig:sky-localization-accuracy} estimate the
-minimum 90\% confidence area versus time for a few rare, loud
-coalescence events detectable by Advanced \LIGO{} and Advanced Virgo. A
-once-per-year event would have minimum 90\% confidence areas of $1300$~deg$^2$
-($3.1$\% of the sky) and $260$~deg$^2$ ($0.63$\% of the sky) at times 25~s and
-10~s before coalescence, respectively. The confidence region shrinks
-quadratically in area as the overall amplitude of the signal increases. By the
-time of merger, the minimum 90\% confidence area shrinks to $0.82$~deg$^2$.
+\citep{Fairhurst2009}. Table~\ref{fig:earlywarning} and figure~\ref{fig:sky-localization-accuracy} show the estimated
+minimum 90\% confidence area versus time of the loudest
+coalescence events detectable by Advanced \LIGO{} and Advanced Virgo.  Once per year,
+we expect to observe an event with a final \SNR\ of 27.4 whose location can be
+constrained to about 1300~deg$^2$ (3.1\% of the sky) 25~s before merger, 260~deg$^2$
+(0.63\% of the sky) 10~s before merger, and 0.82~deg$^2$ (0.0020\% of the sky) at merger.
 

### 2011-06-23 introduction.tex
Reworded galaxy catalog sentene
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1151 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 to accurately reflect mass and thus be a proxy for the \CBC{} event rate, it is
-possible to reduce the localization uncertainty using a galaxy catalog, along
-with the sky location and luminosity distance estimate from the \GW{} signal,
-as a prior probability distribution \citep{galaxy-catalog} as was done in
-S6/VSR3. Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon
+possible to reduce the localization uncertainty by only looking at galaxies from a
+catalog that lie near the sky location and luminosity distance estimate from the \GW{}
+signal~\citep{galaxy-catalog} as was done in S6/VSR3.
+Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon
 distance, the number of galaxies that can produce a given signal

### 2011-06-24 method.tex
lloid_paper: update intro text to match figures; tightened intro and conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1134 ccdd05d5-56c1-461a-bf98-85f6876523b4
Eq. -> equation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1136 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed spacing
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1152 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 filter coefficients and the factor of 100 reduction in the sample rate save a
-factor of $\sim 10^4$ floating point operations per second (\flops) over the
+factor of $\sim$$10^4$ floating point operations per second (\flops) over the
 original (full sample rate) templates.
@@
 This approximation assumes that the frequency of the signal is evolving very slowly so that we can approximate the time slice sample rate as twice the instantaneous gravitational wave frequency, $f^s \approx 2 f(t)$, and the number of samples in the decimated time slice as the sample rate times an infinitesimally short time interval, $\slicessamps \approx 2 f(t) \, \mathrm{d}t$.
+The integral is evaluated using the power-law form of $f(t)$ from equation~(\ref{eq:fgw}).
 Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{hi} = 1/1570$~s $f_\mathrm{hi} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires a factor $\sim 10^{-6}$ fewer \flops\ than the conventional \TD\ method.

### 2011-06-24 method.tex
Removed unnecessary word
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1154 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 showed that highly accurate approximations of inspiral template banks could be
-achieved with relatively few basis templates.  We find that when combined with the
+achieved with few basis templates.  We find that when combined with the
 time slice decomposition, the number of basis templates \numsvdtmps\ is much

### 2011-06-24 method.tex
Fixed typo
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1155 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-For time slices $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ \flops\ 
+For time slice $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ \flops\ 
 to evaluate the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ to apply the 

### 2011-06-24 conclusions.tex
Minor rewording in conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1156 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 of the triggering and coincidence stages that we must address in order to translate
-the early-warning outputs into candidate events.  The time-slice decomposition and
-the \SVD\ may also be beneficial for the postprocessing stage because they will
-help us form low-latency signal-based vetoes (e.g. $\chi^2$ statistics)
+the early-warning outputs into candidate events.  On the other hand, the time-slice
+decomposition and the \SVD\ may be beneficial for the postprocessing stage
+because they will help us form low-latency signal-based vetoes (e.g. $\chi^2$ statistics)
 that have been essential for glitch rejection used in previous \GW{} \CBC{}

### 2011-06-25 introduction.tex
Checked that all figure and table references  meet style guide
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1160 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 detectable within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable
-within 25~s of merger. These rates are shown with their substantial uncertainties in figure~\ref{fig:earlywarning}.
+within 25~s of merger. These rates are shown with their substantial uncertainties in Figure~\ref{fig:earlywarning}.
 
@@
 the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}. Table~\ref{fig:earlywarning} and figure~\ref{fig:sky-localization-accuracy} show the estimated
+\citep{Fairhurst2009}. Table~\ref{fig:earlywarning} and Figure~\ref{fig:sky-localization-accuracy} show the estimated
 minimum 90\% confidence area versus time of the loudest

### 2011-06-25 method.tex
Checked that all figure and table references  meet style guide
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1160 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 An example time slice design satisfying these constraints for a $1.4 - 1.4 \,
-M_{\odot}$ binary is shown in table~\ref{table:time_slices}.
+M_{\odot}$ binary is shown in Table~\ref{table:time_slices}.
 %
@@
 %
-The signal flow diagram in figure~\ref{fig:pipeline} illustrates this
+The signal flow diagram in Figure~\ref{fig:pipeline} illustrates this
 recursion relation as a multirate filter network with a number of early-warning outputs.
@@
 \FD{} matched filter procedure as compared with \lloid{}.  For convenience,
-table~\ref{tab:recap} provides a review of the notation that we will need in
+Table~\ref{tab:recap} provides a review of the notation that we will need in
 this section.

### 2011-06-25 results.tex
Checked that all figure and table references  meet style guide
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1160 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
-in table \ref{tab:time_slices}.
+in Table \ref{tab:time_slices}.
 \begin{table*}
@@
 mismatch are shown as functions of \SVD\ tolerance in
-figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the
+Figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the
 \SVD\ becomes an exact matrix factorization, but the computational cost increases as
@@
 impact of the length $N^\shortuparrow$ of the interpolation filter.  The mismatch as
-a function of $N^\shortuparrow$ is shown in figure~\ref{fig:bw}b.  The
+a function of $N^\shortuparrow$ is shown in Figure~\ref{fig:bw}b.  The
 mismatch saturates at $\sim 2 \times 10^{-4}$ with $N^\shortuparrow = 64$.  We find
@@
 previous subsections to set the \SVD\ tolerance at $\left(1-10^{-4}\right)$ and the
-interpolation filter length to 16. Table \ref{table:flops} shows the computational cost
+interpolation filter length to 16. Table~\ref{table:flops} shows the computational cost
 in \flops\ for this sub-bank.  For the \FD\ method, an \fft\ block size of $\fftblock = 

### 2011-06-26 results.tex
Added template envelope plot
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1161 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
-in Table \ref{tab:time_slices}.
+in Table~\ref{tab:time_slices} and Figure~\ref{fig:envelope}.
 \begin{table*}
-\caption{\label{tab:time_slices} Filter design for these 657 templates.
+\caption{\label{tab:time_slices} Filter design sub-bank of 657 templates.
 From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.
@@
 \end{table*}
-We use this template bank and decomposition for the remainder of this section.
+%
+\begin{figure}[h]
+	\plotone{figures/envelope.pdf}
+	\caption{\label{fig:envelope}Amplitude envelope, time slices boundaries, and sample rates for sub-bank used for impulse response test.}
+\end{figure}
+%
+We use this sub-bank and decomposition for the remainder of this section.
 

### 2011-06-26 envelope.py

@@ -1,0 +1,42 @@
+#!/usr/bin/env python
+
+from optparse import OptionParser, Option
+from pylal import spawaveform
+import numpy as np
+import matplotlib
+from operator import attrgetter
+import pylab
+from itertools import groupby, izip
+import sys
+
+# Command line interface
+
+# Place time slices
+
+# Generate plot
+from gstlal.svd_bank import read_bank
+bank = read_bank('data/svd_0_9.xml')
+pylab.figure(figsize=(3,2.5))
+ax = pylab.subplot(111)
+legend_artists = []
+legend_labels = []
+num_rates = len(set(x.rate for x in bank.bank_fragments))
+tmax = max(x.end for x in bank.bank_fragments)
+mc = spawaveform.chirpmass(1.4, 1.4)
+for color, (rate, fragments) in izip(pylab.linspace(0.25, 1., num_rates), groupby(bank.bank_fragments, attrgetter('rate'))):
+	legend_artists += [pylab.Rectangle((0, 0), 1, 1, facecolor = str(color))]
+	legend_labels += ['%d Hz' % rate]
+	for fragment in fragments:
+		t = pylab.linspace(-fragment.end, -max(fragment.start, 1./4096))
+		a = (-0.2 * t / mc) ** (-1./4)
+		pylab.fill_between(t, -a, a, facecolor = str(color))
+pylab.legend(reversed(legend_artists), reversed(legend_labels), loc='lower left', ncol=2)
+pylab.xlim(-tmax, 0.)
+pylab.ylim(-.1, .1)
+pylab.yticks([], [])
+pylab.xticks([-x.end for x in bank.bank_fragments if x.end > 10], [str(x.end) for x in bank.bank_fragments if x.end > 10], rotation=45, size=8.)
+pylab.grid()
+pylab.xlabel('time relative to coalescence (s)')
+pylab.ylabel(r'strain amplitude')
+pylab.subplots_adjust(left=0.075, right=0.97, top=0.98, bottom=0.225)
+pylab.savefig(sys.argv[1])

### 2011-06-26 snr_in_time.py
Fixed alpha
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1166 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -92,7 +92,7 @@ for i, (cfile, label) in enumerate(zip(['data/ZERO_DET_high_P.txt'], ['zero det'
 	pylab.loglog(t, numminus, markers[i], lw=0.5)
 	pylab.loglog(t, numplus, markers[i], lw=0.5)
 	#pylab.hold(1)
-	pylab.fill_between(t, numminus, numplus, color='k', alpha=0.15)
+	pylab.fill_between(t, numminus, numplus, color='0.85')
 
 pylab.grid()
 #pylab.legend(loc='lower left')

### 2011-06-28 introduction.tex
Converted some figures to eps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1162 ccdd05d5-56c1-461a-bf98-85f6876523b4
tone down
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1167 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-As a compact binary system loses energy to \GW{}s, its
+As a compact binary system loses energy to gravitational waves (\GW{}s), its
 orbital separation decays, leading to a runaway inspiral with the \GW{}
@@
 thought that if a \textsc{ns} is involved, it may become tidally disrupted near
-the merger and fuel a bright electromagnetic (\EM{})
-counterpart \citep{shibata:2007}. One way of obtaining such joint observations
-is through the rapid identification and dissemination of candidate \GW{} events
-for \EM{} followup.
+the merger and fuel a bright electromagnetic (\EM{}) counterpart
+\citep{shibata:2007}.  The gravitational wave signal is in principle detectable
+before merger.  Effort from both the \GW{} and astronomy communities may make
+it possible to use \GW{} observations as an early warning trigger for \EM{}
+followup.  We have the ambition of reporting GW candidates not minutes after
+the merger, as has already been accomplished, but seconds before.  We discuss a
+real-time filtering algorithm for compact binary coalescence waveforms, which
+is one aspect of the work required to make this possible.  We also present the
+prospects for advanced \GW\ detectors and discuss other areas of work that
+would be required for rapid analysis.
 
 In the first generation of ground-based laser interferometers, the \GW{}
-community initiated a
-project to send alerts when potential \GW{} transients were observed in order
-to trigger followup observations by \EM{} telescopes.  The typical latencies
-were 30--60 minutes \citep{HugheyGWPAW2011}, which was an important
-achievement, but far too late to catch prompt \EM{} emission. We have the
-ambition of reporting \GW{} candidates not minutes \emph{after} the merger, as
-has already been accomplished, but seconds \emph{before}.
+community initiated a project to send alerts when potential \GW{} transients
+were observed in order to trigger followup observations by \EM{} telescopes.
+The typical latencies were 30 minutes \citep{HugheyGWPAW2011}, which was an
+important achievement, but probably too late to catch the potential prompt
+\EM{} emission.
 
@@
 
-Though compact binary coalescence (\CBC{}) events may be responsible for several
-classes of \EM{} transients, \CBC{}s are in particular believed to be a
+Though compact binary coalescence (\CBC{}) events may be responsible for
+several classes of \EM{} transients, \CBC{}s are in particular believed to be a
 mechanism for short gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}.
-In this scenario, prompt \EM{} emission arises as shells of
-relativistically out-flowing matter collide in the inner shock.
-The same inner shocks, or potentially reverse shocks, can produce a bright
-accompanying optical flash \citep{Sari99}. Prompt emission is a probe into
-the extreme initial conditions of the outflow, in contrast with afterglows,
-which arise in the external shock with the local medium and are relatively
-insensitive to initial conditions.
+In this scenario, prompt gamma-ray emission arises on the accretion timescale
+of a central compact object formed after the merger, 0.1--1.0\,s
+\citep{Janka1999}.  Interestingly, roughly 1/4 -- 1/2 of observed short \GRB{}s
+also exhibit extended emission $\sim100$\,s.  Potential explanations for the
+emission are delayed fall-back accretion \citep{Metzger2010} caused by
+$r$-process heating or the formation of a proto-magnetar that interacts with
+ejecta \citep{Bucciantini2011}.  In the case of the proto-magnetar, it might be
+formed from accretion induced collapse, or the merger of two neutron stars.  An
+accompanying \GW{} observation would confirm the latter.
 
-Optical flashes have only been observed for a handful of long
-\GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
-or, in the case of \textsc{grb 080319b}, by pure serendipity, where several
-telescopes were already observing the afterglow of another \GRB{} in the same
-field of view~\citep{2008Natur.455..183R}. The observed optical flashes peaked
-within tens of seconds and decayed quickly.
-Short \GRB{}s, on the other hand, typically
-fade too quickly to even catch the tails of the afterglows in any band. Rapid
-\GW{} transient alerts could enable the first observation of
-prompt optical flashes and the rise of afterglows from short \GRB{}s, and, for
-a truly exceptional event, perhaps even a glimpse of a
-precursor \citep{0004-637X-723-2-1711}.
+Optical flashes have only been observed for a handful of long \GRB{}s
+\citep{2011CRPhy..12..255A} by telescopes with extremely rapid response or, in
+the case of \textsc{grb 080319b}, by pure serendipity, where several telescopes
+were already observing the afterglow of another \GRB{} in the same field of
+view~\citep{2008Natur.455..183R}. The observed optical flashes peaked within
+tens of seconds and decayed quickly.  Short \GRB{}s, on the other hand,
+typically fade too quickly to even catch the tails of the afterglows in any
+band. Rapid \GW{} transient alerts could enable the first observation of prompt
+optical flashes and the rise of afterglows from short \GRB{}s, and, for a truly
+exceptional event, perhaps even a glimpse of a precursor
+\citep{0004-637X-723-2-1711}.
 
@@
 
-In October 2010, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its sixth science run
-(S6) and Virgo\footnote{\url{http://www.ego-gw.it/}} completed its third science run (VSR3).
-While both \LIGO{} detectors and Virgo were operating, several all-sky detection
-pipelines operated in a low-latency configuration to send astronomical alerts,
-namely \textsc{mbta}, Coherent WaveBurst, and
-Omega \citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
+In October 2010, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its
+sixth science run (S6) and Virgo\footnote{\url{http://www.ego-gw.it/}}
+completed its third science run (VSR3).  While both \LIGO{} detectors and Virgo
+were operating, several all-sky detection pipelines operated in a low-latency
+configuration to send astronomical alerts, namely \textsc{mbta}, Coherent
+WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
 achieved latencies of 30--60 minutes, which were dominated by a human vetting
-process. Candidates were sent for \EM{} followup to several
-telescopes; Swift, \textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008,
-HugheyGWPAW2011} imaged some of the most likely sky locations.  \textsc{mbta} achieved
-the best \GW{} trigger-generation latencies of 2--5 minutes.  We
-assume that in the advanced detector era the vetting process will be automated,
-so current \GW{} search methodology and telescope actuation would
-dominate latency.
+process. Candidates were sent for \EM{} followup to several telescopes; Swift,
+\textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008, HugheyGWPAW2011}
+imaged some of the most likely sky locations.  \textsc{mbta} achieved the best
+\GW{} trigger-generation latencies of 2--5 minutes.  
+%We assume that in the
+%advanced detector era the vetting process will be automated, so current \GW{}
+%search methodology and telescope actuation would dominate latency.
 
 \begin{figure}[h]
-\plotone{figures/snr_in_time.pdf}
+\plotone{figures/snr_in_time}
 \caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
-sources that will be detectable by Advanced \LIGO\ a given number of seconds before coalescence.
-The heavy solid line is the most realistic yearly rate estimate.  The shaded
-region represents the 5 to 95\% confidence interval arising from uncertainty
-in predicted event rates \citep{Abadie:2010p10836}.}
+sources that could be detectable by Advanced \LIGO\ a given number of seconds
+before coalescence.  The time before coalescence that a source could be
+detectable is a lower bound that does not include any latency inducing aspects
+of a real analysis.  The heavy solid line is the most realistic yearly rate
+estimate.  The shaded region represents the 5 to 95\% confidence interval
+arising from uncertainty in predicted event rates \citep{Abadie:2010p10836}.}
 \end{figure}
 %
-It has been observed that by examining the signal to noise ratio (\SNR{}) stream for
-threshold crossings before the \GW{} signal leaves the detection band, it is
-possible to trade some \SNR{} and sky localization accuracy for pre-merger
-detection. Figure~\ref{fig:earlywarning} shows projected early detection rates
-for \textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
+Examining the signal to noise ratio (\SNR{}) stream for threshold crossings
+before the \GW{} signal leaves the detection band makes it possible to trade
+some \SNR{} and sky localization accuracy for pre-merger detection.
+Figure~\ref{fig:earlywarning} shows projected early detectability rates for
+\textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
 predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
@@
 The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
-observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
-detectable within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable
-within 25~s of merger. These rates are shown with their substantial uncertainties in Figure~\ref{fig:earlywarning}.
+observe a total of 40~events~yr$^{-1}$. $\sim$10~yr$^{-1}$ will be detectable
+within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable within 25~s of
+merger if analysis can proceed with near zero latency. These rates are shown
+with their substantial uncertainties in Figure~\ref{fig:earlywarning}.
+
+We emphasize that Figure~\ref{fig:earlywarning} provides only the lower bound
+for early detection and should serve as a baseline to understand the potential
+of early warning \GW{} searches in a ideal world.  Real analysis will have
+sources of latency that are unavoidable meaning that the actual detection will
+occur with some delay.
 
 \begin{table}[h]
-\caption{\label{table:sky-localization-accuracy}Horizon distance, \SNR\ at merger, and area of 90\% confidence at selected times before merger for sources with expected detection rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
+\caption{\label{table:sky-localization-accuracy}Horizon distance, \SNR\ at
+merger, and area of 90\% confidence at selected times before merger for sources
+with expected detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
 \begin{center}
@@
 \begin{figure}[h]
-\plotone{figures/loc_in_time.pdf}
+\plotone{figures/loc_in_time}
 \caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
-region as a function of time for a sources with anticipated detection rates
-of 40, 10, 1, and 0.1~yr$^{-1}$. The heavy dot indicates the
-time at which the accumulated \SNR\ exceeds a threshold of 8.}
+region as a function of time before coalescence for sources with anticipated
+detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$. The heavy dot indicates
+the time at which the accumulated \SNR\ exceeds a threshold of 8.}
 \end{figure}
 
-\EM{} followup requires estimating the location of the \GW{} source. The
-localization uncertainty is dominated by uncertainty in the time of arrival of
-the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}. Table~\ref{fig:earlywarning} and Figure~\ref{fig:sky-localization-accuracy} show the estimated
-minimum 90\% confidence area versus time of the loudest
-coalescence events detectable by Advanced \LIGO{} and Advanced Virgo.  Once per year,
-we expect to observe an event with a final \SNR\ of 27.4 whose location can be
-constrained to about 1300~deg$^2$ (3.1\% of the sky) 25~s before merger, 260~deg$^2$
-(0.63\% of the sky) 10~s before merger, and 0.82~deg$^2$ (0.0020\% of the sky) at merger.
+Before discussing realistic latencies, it is important to note that \EM{}
+followup requires estimating the location of the \GW{} source. The localization
+uncertainty is dominated by uncertainty in the time of arrival of the \GW{}s,
+which is determined by the signal's effective bandwidth and \SNR{}
+\citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
+Figure~\ref{fig:sky-localization-accuracy} show the estimated minimum 90\%
+confidence area versus time of the loudest coalescence events detectable by
+Advanced \LIGO{} and Advanced Virgo.  Once per year, we expect to observe an
+event with a final \SNR\ of $\sim 27$ whose location can be constrained to about
+1300~deg$^2$ (3.1\% of the sky) using only the waveform 25~s before merger,
+260~deg$^2$ (0.63\% of the sky) using the waveform 10~s before merger, and
+0.82~deg$^2$ (0.0020\% of the sky) using the full \GW\ waveform.
 
-It is infeasible to search hundreds of square degrees for a prompt
-counterpart. For
-reference, \textsc{lsst} should see first light around the same time as
+It is infeasible to search hundreds of square degrees for a prompt counterpart.
+For reference, \textsc{lsst} should see first light around the same time as
 Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
-\citep{2008arXiv0805.2366I}.  However, if we can take galactic luminosities
-to accurately reflect mass and thus be a proxy for the \CBC{} event rate, it is
-possible to reduce the localization uncertainty by only looking at galaxies from a
-catalog that lie near the sky location and luminosity distance estimate from the \GW{}
-signal~\citep{galaxy-catalog} as was done in S6/VSR3.
-Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon
-distance, the number of galaxies that can produce a given signal
-amplitude is much larger than the number for initial \LIGO{} and thus the
-catalog will not be as useful for downselecting pointings generally. However,
-our exceptional sources will necessarily be extremely nearby. Within this
-reduced volume there will be fewer galaxies to consider for a given
-candidate and catalog completeness will be less of a concern.
-This should allow us to shrink the 90\% confidence area substantially.
+\citep{2008arXiv0805.2366I}.  However, 
+%if we can take galactic luminosities to
+%accurately reflect mass and thus be a proxy for the \CBC{} event rate, 
+it is possible to reduce the localization uncertainty by only looking at
+galaxies from a catalog that lie near the sky location and luminosity distance
+estimate from the \GW{} signal~\citep{galaxy-catalog} as was done in S6/VSR3.
+Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon distance,
+the number of galaxies that can produce a given signal amplitude is much larger
+than the number for initial \LIGO{} and thus the catalog will not be as useful
+for downselecting pointings generally. However, exceptional \GW{} sources will
+necessarily be extremely nearby. Within this reduced volume there will be fewer
+galaxies to consider for a given candidate and catalog completeness will be
+less of a concern.  This should reduce the 90\% confidence area substantially.
+Also future \GW{} detectors, should they be built, will improve the \SNR\ and
+likewise the error box.
 
-\subsection{Latency in \CBC{} search algorithms}
+\subsection{Latency in \CBC{} searches}
 
-Realizing advance detection of \CBC{}s will
-require striking a balance between latency and throughput. \CBC{} searches
-consist of banks of matched filters, or
-cross-correlations between the data stream and a bank of nominal ``template''
-signals.  There are many different implementations of matched filters, but most
-have high throughput at the cost of high latency, or low latency at the cost of
-low throughput.  The former are epitomized by the overlap-save algorithm~%
-\citep{numerical-recipes-chapter-13} for frequency domain (\FD) convolution,
-currently the preferred method in \GW{}
-searches.  The most obvious example of the latter is the time domain
-(\TD) convolution, which is latency-free.  However, its
-computational complexity is quadratic in the length of the templates, so it is
-prohibitively expensive for long templates.
+Here we provide a brief list of technical latencies associated with \CBC\
+analysis in S6, VSR3 in chronological order \citep{HugheyGWPAW2011}.
+%
+\begin{itemize}
+%
+\item {\it Data calibration, quality control and distribution ($\sim
+2$\,minutes):}  This is the time it took to calibrate the detector data, assess
+its quality and distribute it to a central location for analysis.  Improvements
+in other areas could make this the dominating source of latency. If possible
+work should be done to improve this.
+%
+\item {\it Trigger production ($2-5$\,minutes):} This was an impressive latency
+and was certainly sub-dominant in the S6, VSR3 analysis.  However, even if no
+other sources of latency exist it is too long to catch prompt or extended
+emission.  It is also more challenging for searches with advanced detectors
+because the templates needed to filter the data are about 10 times longer.  We
+will focus on reducing the latency of the trigger production step in this work.
+%
+\item {\it Event gathering, downselection and pointing calculations ($2-3$\,
+minutes):} This was the time required to gather the event parameters and sky
+localization from the various online analyses, downselect the events and
+calculate pointings.  If other sources of latency were improved this could
+dominate.  If possible work should be done to improve this.
+%
+\item {\it Human vetting of the trigger quality ($10-20$\,minutes):}  This was
+by far the largest latency during the S6, VSR3 low latency analysis and will
+hopefully be automated by the advanced detector era.  We therefore give it no
+more consideration here.
+%
+\end{itemize}
+%
+
+In addition to the latencies described above, which may be overcome, there are
+fundamental latencies associated with data acquisition and transfer that cannot
+be overcome.  For example, the \LIGO{} data acquisition records science data in
+1/16~s blocks~\citep{Bork2001}. Data must also be aggregated from all of the
+\GW\ observatories using global computer networks capable of high bandwidth but
+perhaps only modest latency.  This could introduce an additional latency of
+$\sim$100~ms for a total of $\sim 1/4$\,s of unavoidable latency.
+
+This work will focus on reducing the latency of trigger production.  Realizing
+advance detection of \CBC{}s will require striking a balance between latency
+and throughput of filtering \GW\ wave data. \CBC{} searches consist of banks of
+matched filters, or cross-correlations between the data stream and a bank of
+nominal ``template'' signals.  There are many different implementations of
+matched filters, but most have high throughput at the cost of high latency, or
+low latency at the cost of low throughput.  The former are epitomized by the
+overlap-save algorithm~% \citep{numerical-recipes-chapter-13}
+for frequency domain (\FD) convolution, currently the preferred method in \GW{}
+searches.  The most obvious example of the latter is the time domain (\TD)
+convolution, which is latency-free.  However, its computational complexity is
+quadratic in the length of the templates, so it is prohibitively expensive for
+long templates.
 
@@
 methods such as singular value decomposition (\SVD{}) to reduce the number of
-templates \citep{Cannon:2010p10398}. We will use both aspects to demonstrate that a very
-low latency analysis with advance detection of compact binary sources is
-possible with current computing resources.  Assuming other technical sources of
-latency can be reduced significantly, this should allow the possibility for
-prompt alerts to be sent to the astronomical community.
+templates \citep{Cannon:2010p10398}. We will use both aspects to demonstrate
+that a very low latency detection statistic is possible with current computing
+resources.  Assuming the other technical sources of latency can be reduced
+significantly, this should allow the possibility for prompt alerts to be sent
+to the astronomical community.
 

### 2011-06-28 inspiral_svd.tex
Incremented DCC version number.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1084 ccdd05d5-56c1-461a-bf98-85f6876523b4
Incremented DCC version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1086 ccdd05d5-56c1-461a-bf98-85f6876523b4
Upped DCC version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1088 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added my e-mail address as corresponding author
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1131 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\title{Technique for \earlywarning\ detection of compact binary coalescence
+\title{Technique for \earlywarning\ filtering for compact binary coalescence
 and its implications for multi-messenger astronomy}
@@
 }
+% we either need to include all of our emails, or none.
+%\email{leo.singer@ligo.org}
 
@@
 from compact binary coalescence.  We demonstrate a computationally practical
-analysis strategy that produces early-warning triggers even before
+filtering strategy that could produce early-warning triggers before
 gravitational radiation from the final merger has arrived at the detectors.
-With current rate estimates for the Advanced \LIGO{} design configuration, it
-should be possible to detect $\sim$10 such sources earlier than 10 seconds
-before merger in 1 year of live time using the method we describe.
 \end{abstract}
@@
 many discussions about rate estimates and the prospects of early detection.  CH
-would also like to thank Patrick Brady for countless fruitful conversations
-about low latency analysis methods.  NF would like to thank Alessandra Corsi
-for illuminating discussions on astronomical motivations.  This research is
-supported by the National Science Foundation through a Graduate Research
-Fellowship to LS and by the Perimeter Institute for Theoretical Physics
-through a fellowship to CH. DK is supported from the Max Planck Gesellschaft.
+would like to thank Patrick Brady for countless fruitful conversations about
+low latency analysis methods.  CH would like to thank John Zweizig for
+disscussions about \LIGO\ data acquisition.  NF would like to thank Alessandra
+Corsi for illuminating discussions on astronomical motivations.  This research
+is supported by the National Science Foundation through a Graduate Research
+Fellowship to LS and by the Perimeter Institute for Theoretical Physics through
+a fellowship to CH. DK is supported from the Max Planck Gesellschaft.
 
-This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{17}.
+This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{21}.
 

### 2011-06-28 conclusions.tex
tone down
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1169 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We have demonstrated a computationally feasible procedure for the rapid and
-even early-warning detection of \GW{}s emitted during the coalescence of
-neutron stars and stellar-mass black holes. Our method is as fast as standard
-\fft{} convolutions but allows for latency-free, real-time operation.  We
-anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
-search for binary neutron stars using coincident \GW\ data from a four-detector
-network.  In the future, additional computational savings could be acheived by
-conditionally reconstructing the \SNR{} time-series only during times when a
-composite detection statistic crosses a threshold~\citep{svd-compdetstat}.
-However, the anticipated required number of computers is well within the
-current computing capabilities of the \LIGO{} Data
+even early-warning filtering of data for \GW{}s emitted during the coalescence
+of neutron stars and stellar-mass black holes.  It is one part of a complicated
+analysis and observation strategy that will unfortunately have other sources of
+latency.  However, we hope that it will motivate further work to reduce such
+technical sources of \GW{} observation latency and encourage the possibility of
+even more rapid \EM\ followup observations to catch prompt emission in the
+advanced detector era.
+
+The \lloid{} method is as fast as standard \fft{} convolutions but allows for a
+technically latency-free, real-time operation.  We anticipate requiring
+$\gtrsim 4 \times 10 = 40$ modern multi-core computers to search for binary
+neutron stars using coincident \GW\ data from a four-detector network.  In the
+future, additional computational savings could be acheived by conditionally
+reconstructing the \SNR{} time-series only during times when a composite
+detection statistic crosses a threshold~\citep{svd-compdetstat}.  However, the
+anticipated required number of computers is well within the current computing
+capabilities of the \LIGO{} Data
 Grid\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.
@@
 \CBC\ events may be the progenitors of some short hard \GRB{}s and are expected
-to be accompanied prompt \EM{} signals.  Rapid alerts to the broader
+to be accompanied by prompt \EM{} signals.  Rapid alerts to the broader
 astronomical community will improve the chances of detecting an \EM{}
-counterpart in bands from gamma-rays down to radio.
+counterpart in bands from gamma-rays down to radio.  It will be extremely
+challenging for all but the rarest events to localize the source well enough in
+advance to detect the onset of prompt emission, however localization prior to
+the onset of extended emission on the 10--100\,s time-scale is more easily
+achieved and will benefit from the full bandwidth and \SNR\ of the \GW{} event
+at merger.  Probing even the extended emission timescale requires very low
+latency filtering schemes such as the one described here.
 
+\begin{comment}
 The detection algorithm we described has no intrinsic latency.  However, there
@@
 $\sim$100~ms.  Lastly, the software implementation of the algorithm itself may
-introduce latency.  We have shown a prototype implementation of \lloid{} using
-\gstreamer, an open source signal processing platform. Some work could
-be done in order to improve upon the latency of our implementation,
-including tighter integration between data acquisition and analysis. This could
-be considered for third-generation detector design.  Also possible for
+introduce latency.  
+\end{comment}
+
+We have shown a prototype implementation of the \lloid{} algorithm using
+\gstreamer, an open source signal processing platform.  Work will be done in
+order to improve upon the latency of our implementation.  Ultimately the best
+possible latency would be achieved by tighter integration between data
+acquisition and analysis with dedicated hardware and software. This could be
+considered for third-generation detector design.  Also possible for
 third-generation instruments, the \lloid{} method could provide the input for a
@@
 
-Although we have demonstrated a computationally feasible detection statistic for
-advance detection, we have not yet explored data whitening, triggering,
-coincidence, and ranking of gravitational wave candidates in a framework that
-supports early \EM\ followup.  One complication in deploying a low-latency or
-early-warning search is that whitening the detector data involves estimating the
-noise power spectrum, which is often done in acausally.  We will have to look into
-causal or low latency alternatives to conventional spectral estimation techniques.
+Although we have demonstrated a computationally feasible detection statistic
+for advance detection, we have not yet explored data calibration and whitening,
+triggering, coincidence, and ranking of gravitational wave candidates in a
+framework that supports early \EM\ followup.
+\begin{comment}
+One complication in deploying a
+low-latency or early-warning search is that whitening the detector data
+involves estimating the noise power spectrum, which is often done acausally.
+We will have to look into causal or low latency alternatives to conventional
+spectral estimation techniques.  
+\end{comment}
+We will explore these and also using the time-slice decomposition and the \SVD\
+to form low-latency signal-based vetoes (e.g.  $\chi^2$ statistics) that have
+been essential for glitch rejection used in previous \GW{} \CBC{} searches.
+
+\begin{comment}
 Another potential complication is that if adjacent time slices have non-negligible
@@
 searches.
+\end{comment}
 
-Our future work must more deeply address sky localization accuracy prior to
-merger as well as observing strategies. Here, we have followed
+Our future work must more deeply address sky localization accuracy in a
+realistic setting as well as observing strategies. Here, we have followed
 \citet{Fairhurst2009} in estimating the area of 90\% localization confidence in
@@
 rates and detector networks will be necessary in order to fully understand the
-prospects for early-warning detection, localization, and \EM\ folluwup using
+prospects for early-warning detection, localization, and \EM\ followup using
 the techniques we have described.

### 2011-06-28 results.tex
Converted some figures to eps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1162 ccdd05d5-56c1-461a-bf98-85f6876523b4
tone down
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1170 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{figure}[h]
-	\plotone{figures/envelope.pdf}
+	\plotone{figures/envelope}
 	\caption{\label{fig:envelope}Amplitude envelope, time slices boundaries, and sample rates for sub-bank used for impulse response test.}
@@
 
-\subsection{Lower bounds on computational cost compared to other methods}
+\subsection{Lower bounds on computational cost and latency compared to other
+methods}
 
@@
 sub-bank of templates compared to other methods.  We used the results of the
-previous subsections to set the \SVD\ tolerance at $\left(1-10^{-4}\right)$ and the
-interpolation filter length to 16. Table~\ref{table:flops} shows the computational cost
-in \flops\ for this sub-bank.  For the \FD\ method, an \fft\ block size of $\fftblock = 
-2 \tmpsamps$ is assumed, resulting in a latency of $\left(\tmpsamps f^0\right)$
-seconds.  Both the \FD\ method and \lloid\ are five orders of magnitude faster than
-the conventional \TD\ method.  However, the \FD\ method has a latency of over half of an
-hour, whereas \lloid\ has no latency at all.
+previous subsections to set the \SVD\ tolerance at $\left(1-10^{-4}\right)$ and
+the interpolation filter length to 16. Table~\ref{table:flops} shows the
+computational cost in \flops\ for this sub-bank.  For the \FD\ method, an \fft\
+block size of $\fftblock = 2 \tmpsamps$ is assumed, resulting in a latency of
+$\left(\tmpsamps f^0\right)$ seconds.  Both the \FD\ method and \lloid\ are
+five orders of magnitude faster than the conventional \TD\ method.  However,
+the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no
+latency at all.
 %
@@
 
-Our \gstreamer\ pipeline for measuring impulse responses contained instrumentation that would not be necessary for an actual search, including additional interpolation filters to bring the early-warning outputs back to the full sample rate and additional outputs for recording signals to disk.
+Our \gstreamer\ pipeline for measuring impulse responses contained
+instrumentation that would not be necessary for an actual search, including
+additional interpolation filters to bring the early-warning outputs back to the
+full sample rate and additional outputs for recording signals to disk.
 
-We wrote a second, stripped pipeline to evaluate the actual latency and computational overhead.
-We executed this pipeline on one of the submit machines of the \LIGO-Caltech cluster, a Sun Microsystems Sun Fire\texttrademark\ X4600~M2 server with eight quad-core 2.7~GHz AMD Opteron\texttrademark\ 8384 processors.
-This test consumed 78\% of the capacity of just one out of the 32 cores, maintaining a constant latency of 5~s.
+We wrote a second, stripped pipeline to evaluate the actual latency and
+computational overhead.  We executed this pipeline on one of the submit
+machines of the \LIGO-Caltech cluster, a Sun Microsystems Sun
+Fire\texttrademark\ X4600~M2 server with eight quad-core 2.7~GHz AMD
+Opteron\texttrademark\ 8384 processors.  This test consumed 78\% of the
+capacity of just one out of the 32 cores, maintaining a constant latency of
+5~s.
 
-The measured overhead is consistent to within an order of magnitude with the lower bound from the \flops\ budget.
-The additional overhead is probably dominated by thread synchronization.
-A carefully optimized \gstreamer\ pipeline or a hand-tuned C implementation of the pipeline might reduce overhead further.
+The measured overhead is consistent to within an order of magnitude with the
+lower bound from the \flops\ budget.  The additional overhead is probably
+dominated by thread synchronization.  A carefully optimized \gstreamer\
+pipeline or a hand-tuned C implementation of the pipeline might reduce overhead
+further.
 
-The 5~s latency is probably due to buffering and synchronization.  The latency could probably be reduced by carefully tuning buffer lengths at every stage in the pipeline.
+The 5~s latency is probably due to buffering and synchronization.  The latency
+might be reduced by carefully tuning buffer lengths at every stage in
+the pipeline.
+
+Even without further refinements, our implementation of the \lloid\ algorithm
+achieves latencies that would make other aspects of the analysis the limiting
+steps assuming that the human vetting process is removed.

### 2011-06-28 results.tex
addressed leos comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1172 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 achieves latencies that would make other aspects of the analysis the limiting
-steps assuming that the human vetting process is removed.
+steps assuming that the human vetting process is automated.

### 2011-06-28 introduction.tex
addressed leos comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1173 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 followup.  We have the ambition of reporting GW candidates not minutes after
-the merger, as has already been accomplished, but seconds before.  We discuss a
-real-time filtering algorithm for compact binary coalescence waveforms, which
-is one aspect of the work required to make this possible.  We also present the
-prospects for advanced \GW\ detectors and discuss other areas of work that
-would be required for rapid analysis.
+the merger, as has already been accomplished, but seconds before.  We discuss
+one essential ingredient in this effort, a computationally inexpensive
+real-time filtering algorithm for detecting inspiral signals in gravitational
+wave data.  We also present the prospects for advanced \GW\ detectors and
+discuss other areas of work that would be required for rapid analysis.
 
@@
 sources that could be detectable by Advanced \LIGO\ a given number of seconds
-before coalescence.  The time before coalescence that a source could be
-detectable is a lower bound that does not include any latency inducing aspects
-of a real analysis.  The heavy solid line is the most realistic yearly rate
+before coalescence.  The heavy solid line is the most realistic yearly rate
 estimate.  The shaded region represents the 5 to 95\% confidence interval
@@
 %
+\item {\it raw data aggregation and broadcasting ($sim 0.25$\,s):}  There are
+fundamental latencies associated with data acquisition and broadcasting that
+cannot be overcome.  For example, the \LIGO{} data acquisition records science
+data in 1/16~s blocks~\citep{Bork2001}. Data must also be aggregated from all
+of the \GW\ observatories using global computer networks capable of high
+bandwidth but perhaps only modest latency.  This could introduce an additional
+latency of $\sim$100~ms.  These latencies are microscopic compared with the
+other technical sources of latency below. 
+%
 \item {\it Data calibration, quality control and distribution ($\sim
@@
 %
-
-In addition to the latencies described above, which may be overcome, there are
-fundamental latencies associated with data acquisition and transfer that cannot
-be overcome.  For example, the \LIGO{} data acquisition records science data in
-1/16~s blocks~\citep{Bork2001}. Data must also be aggregated from all of the
-\GW\ observatories using global computer networks capable of high bandwidth but
-perhaps only modest latency.  This could introduce an additional latency of
-$\sim$100~ms for a total of $\sim 1/4$\,s of unavoidable latency.
 

### 2011-06-28 introduction.tex
lloid_paper: polished introduction; not much change in material.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1174 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 amplitude and frequency increasing until the system eventually merges.  It is
-thought that if a \textsc{ns} is involved, it may become tidally disrupted near
+thought that if a neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
 the merger and fuel a bright electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  The gravitational wave signal is in principle detectable
+\citep{shibata:2007}.  The gravitational-wave signal is in principle detectable
 before merger.  Effort from both the \GW{} and astronomy communities may make
 it possible to use \GW{} observations as an early warning trigger for \EM{}
-followup.  We have the ambition of reporting GW candidates not minutes after
+followup. In the first generation of ground-based laser interferometers, the \GW{}
+community initiated a project to send alerts when potential \GW{} transients
+were observed in order to trigger followup observations by \EM{} telescopes.
+The typical latencies were 30 minutes \citep{HugheyGWPAW2011}, which was an
+important achievement, but too late to catch the potential prompt \EM{}
+emission. We have the ambition of reporting GW candidates not minutes after
 the merger, as has already been accomplished, but seconds before.  We discuss
@@
 discuss other areas of work that would be required for rapid analysis.
-
-In the first generation of ground-based laser interferometers, the \GW{}
-community initiated a project to send alerts when potential \GW{} transients
-were observed in order to trigger followup observations by \EM{} telescopes.
-The typical latencies were 30 minutes \citep{HugheyGWPAW2011}, which was an
-important achievement, but probably too late to catch the potential prompt
-\EM{} emission.
 
@@
 of a central compact object formed after the merger, 0.1--1.0\,s
-\citep{Janka1999}.  Interestingly, roughly 1/4 -- 1/2 of observed short \GRB{}s
-also exhibit extended emission $\sim100$\,s.  Potential explanations for the
+\citep{Janka1999}.  Interestingly, roughly a quarter to half of observed short
+\GRB{}s
+also exhibit extended emission $\sim$$100$\,s.  Potential explanations for the
 emission are delayed fall-back accretion \citep{Metzger2010} caused by
@@
 ejecta \citep{Bucciantini2011}.  In the case of the proto-magnetar, it might be
-formed from accretion induced collapse, or the merger of two neutron stars.  An
+formed from accretion induced collapse, or the merger of two \textsc{ns}s.  An
 accompanying \GW{} observation would confirm the latter.
@@
 were already observing the afterglow of another \GRB{} in the same field of
-view~\citep{2008Natur.455..183R}. The observed optical flashes peaked within
+view \citep{2008Natur.455..183R}. The observed optical flashes peaked within
 tens of seconds and decayed quickly.  Short \GRB{}s, on the other hand,
@@
 band. Rapid \GW{} transient alerts could enable the first observation of prompt
-optical flashes and the rise of afterglows from short \GRB{}s, and, for a truly
-exceptional event, perhaps even a glimpse of a precursor
-\citep{0004-637X-723-2-1711}.
+optical flashes and the rise of afterglows from short \GRB{}s.
 
@@
 The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
-observe a total of 40~events~yr$^{-1}$. $\sim$10~yr$^{-1}$ will be detectable
-within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable within 25~s of
+observe a total of 40\,events\,yr$^{-1}$. $\sim$10\,yr$^{-1}$ will be detectable
+within 10\,s of merger and $\sim$5\,yr$^{-1}$ will be detectable within 25\,s of
 merger if analysis can proceed with near zero latency. These rates are shown
@@
 merger, and area of 90\% confidence at selected times before merger for sources
-with expected detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
+with expected detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$.}
 \begin{center}
@@
 region as a function of time before coalescence for sources with anticipated
-detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$. The heavy dot indicates
+detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$. The heavy dot indicates
 the time at which the accumulated \SNR\ exceeds a threshold of 8.}
@@
 event with a final \SNR\ of $\sim 27$ whose location can be constrained to about
-1300~deg$^2$ (3.1\% of the sky) using only the waveform 25~s before merger,
-260~deg$^2$ (0.63\% of the sky) using the waveform 10~s before merger, and
-0.82~deg$^2$ (0.0020\% of the sky) using the full \GW\ waveform.
+1300\,deg$^2$ (3.1\% of the sky) using only the waveform 25\,s before merger,
+260\,deg$^2$ (0.63\% of the sky) using the waveform 10\,s before merger, and
+0.82\,deg$^2$ (0.0020\% of the sky) using the full \GW\ waveform.
 
@@
 
-Here we provide a brief list of technical latencies associated with \CBC\
-analysis in S6, VSR3 in chronological order \citep{HugheyGWPAW2011}.
+There were a number of sources of latency associated with \CBC{}
+analysis in S6/VSR3 \citep{HugheyGWPAW2011}, listed here in chronological order.
 %
@@
 %
-\item {\it raw data aggregation and broadcasting ($sim 0.25$\,s):}  There are
-fundamental latencies associated with data acquisition and broadcasting that
-cannot be overcome.  For example, the \LIGO{} data acquisition records science
-data in 1/16~s blocks~\citep{Bork2001}. Data must also be aggregated from all
+\item {\it Data aggregation and broadcasting ($\sim$$0.25$\,s):}
+The \LIGO{} data acquisition system records science
+data in 1/16\,s blocks \citep{Bork2001}. Data must also be aggregated from all
 of the \GW\ observatories using global computer networks capable of high
 bandwidth but perhaps only modest latency.  This could introduce an additional
-latency of $\sim$100~ms.  These latencies are microscopic compared with the
-other technical sources of latency below. 
+latency of $\sim$100\,ms.  These latencies are much smaller than the
+other sources of latency below and would require significant reengineering and
+capital investments to overcome.
 %
-\item {\it Data calibration, quality control and distribution ($\sim
-2$\,minutes):}  This is the time it took to calibrate the detector data, assess
-its quality and distribute it to a central location for analysis.  Improvements
-in other areas could make this the dominating source of latency. If possible
-work should be done to improve this.
+\item {\it Data calibration, quality control, and distribution
+($\sim$$2$\,min):} There is technical latency in calibrating the detector data,
+assessing its quality, and distributing it to a central location for analysis.
+None of this latency is fundamental and improvements in other areas could make
+this the dominating source of latency, so work should be done to improve this.
 %
-\item {\it Trigger production ($2-5$\,minutes):} This was an impressive latency
-and was certainly sub-dominant in the S6, VSR3 analysis.  However, even if no
-other sources of latency exist it is too long to catch prompt or extended
+\item {\it Trigger production ($2$--$5$\,min):} This was an impressive latency
+and was certainly sub-dominant in the S6/VSR3 analysis.  However, even if no
+other sources of latency exist, triggering latency is too long to catch prompt
+or extended
 emission. It is also more challenging for searches with advanced detectors
-because the templates needed to filter the data are about 10 times longer.  We
-will focus on reducing the latency of the trigger production step in this work.
+because the templates needed to filter the data are about ten times longer.  We
+will focus on reducing the latency of trigger production in this work.
 %
-\item {\it Event gathering, downselection and pointing calculations ($2-3$\,
-minutes):} This was the time required to gather the event parameters and sky
-localization from the various online analyses, downselect the events and
-calculate pointings.  If other sources of latency were improved this could
-dominate.  If possible work should be done to improve this.
+\item {\it Event gathering, downselection, and pointing calculations
+($2$--$3$\,min):} S6/VSR3 saw the introduction of low-latency astronomical
+alerts, which required gathering event parameters and sky localization from the
+various online analyses, downselecting the events, and calculating pointings.
+If other sources of latency improve, this could dominate, so work should be
+done to improve this.
 %
-\item {\it Human vetting of the trigger quality ($10-20$\,minutes):}  This was
-by far the largest latency during the S6, VSR3 low latency analysis and will
-hopefully be automated by the advanced detector era.  We therefore give it no
-more consideration here.
+\item {\it Human vetting of the trigger quality ($10$--$20$\,min):} In
+commissioning the new alert system in S6/VSR3, humans manually examined each
+candidate for obvious flaws before publishing an alert. This was by far the
+largest latency during S6/VSR3 and hopefully confidence will grow to the
+point where human operators are satisfied with checks after alerts have been
+transmitted. We give it no further consideration here.
 %
 \end{itemize}
-%
 
@@
 advance detection of \CBC{}s will require striking a balance between latency
-and throughput of filtering \GW\ wave data. \CBC{} searches consist of banks of
+and throughput of filtering \GW\ data. \CBC{} searches consist of banks of
 matched filters, or cross-correlations between the data stream and a bank of
@@
 low latency at the cost of low throughput.  The former are epitomized by the
-overlap-save algorithm~% \citep{numerical-recipes-chapter-13}
+overlap-save algorithm % \citep{numerical-recipes-chapter-13}
 for frequency domain (\FD) convolution, currently the preferred method in \GW{}

### 2011-06-28 introduction.tex
lloid_paper: polished EM counterpart section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1175 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the merger and fuel a bright electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  The gravitational-wave signal is in principle detectable
+\citep{shibata:2007}.  The \GW{} signal is in principle detectable
 before merger.  Effort from both the \GW{} and astronomy communities may make
@@
 The typical latencies were 30 minutes \citep{HugheyGWPAW2011}, which was an
-important achievement, but too late to catch the potential prompt \EM{}
-emission. We have the ambition of reporting GW candidates not minutes after
+important achievement, but too late to catch any prompt \EM{}
+emission. We have the ambition of reporting \GW{} candidates not minutes after
 the merger, as has already been accomplished, but seconds before.  We discuss
 one essential ingredient in this effort, a computationally inexpensive
-real-time filtering algorithm for detecting inspiral signals in gravitational
-wave data.  We also present the prospects for advanced \GW\ detectors and
+low-latency filtering algorithm for detecting inspiral signals in \GW{}
+data.  We also present the prospects for advanced \GW\ detectors and
 discuss other areas of work that would be required for rapid analysis.
@@
 In this scenario, prompt gamma-ray emission arises on the accretion timescale
-of a central compact object formed after the merger, 0.1--1.0\,s
+of a central compact object formed after the merger, $0.1$--$1.0$\,s
 \citep{Janka1999}.  Interestingly, roughly a quarter to half of observed short
-\GRB{}s
-also exhibit extended emission $\sim$$100$\,s.  Potential explanations for the
-emission are delayed fall-back accretion \citep{Metzger2010} caused by
-$r$-process heating or the formation of a proto-magnetar that interacts with
+\GRB{}s also exhibit extended emission of $30$--$100$\,s in duration beginning
+$\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the initial
+outburst.  Potential explanations for the emission are delayed fall-back
+accretion caused by $r$-process heating \citep{Metzger2010} or the formation of
+a proto-magnetar that interacts with
 ejecta \citep{Bucciantini2011}.  In the case of the proto-magnetar, it might be
-formed from accretion induced collapse, or the merger of two \textsc{ns}s.  An
+formed from accretion-induced collapse, or the merger of two \textsc{ns}s.  An
 accompanying \GW{} observation would confirm the latter.

### 2011-06-29 results.tex
Moved time slice table and figure together
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1176 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
-in Table~\ref{tab:time_slices} and Figure~\ref{fig:envelope}.
+in Table~\ref{tab:time_slices}.
 \begin{table*}
@@
 \begin{center}
+\begin{minipage}[c]{0.4\textwidth}
+\includegraphics{figures/envelope}
+\end{minipage}
+\begin{minipage}[c]{0.55\textwidth}
 \begin{tabular}{rr@{,\,}lc*{6}{r}}
 \tableline\tableline
-\\ [-1.5ex]
-\multicolumn{4}{c}{} &\multicolumn{6}{c}{$\log_{10}$ (1$-$\SVD{} tolerance)} \\ [1ex]
+\\ [-2ex]
+$f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD{} tolerance)} \\% [1ex]
 \cline{5-10}
-\\ [-1.5ex]
-$f^s$ (Hz) & $[t^s$&$t^{s+1})$ (s) & $N^s$ & $-1$ & $-2$ & $-3$ & $-4$ & $-5$ & $-6$ \\ \tableline
+\\[-2.5ex]
+(Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline
 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
@@
 \end{tabular}
+\end{minipage}
 \end{center}
 \end{table*}
-%
-\begin{figure}[h]
-	\plotone{figures/envelope}
-	\caption{\label{fig:envelope}Amplitude envelope, time slices boundaries, and sample rates for sub-bank used for impulse response test.}
-\end{figure}
-%
 We use this sub-bank and decomposition for the remainder of this section.

### 2011-06-29 envelope.py

@@ -16,7 +16,7 @@ import sys
 # Generate plot
 from gstlal.svd_bank import read_bank
 bank = read_bank('data/svd_0_9.xml')
-pylab.figure(figsize=(3,2.5))
+pylab.figure(figsize=(2.6,2.5))
 ax = pylab.subplot(111)
 legend_artists = []
 legend_labels = []
@@ -38,5 +38,5 @@ pylab.xticks([-x.end for x in bank.bank_fragments if x.end > 10], [str(x.end) fo
 pylab.grid()
 pylab.xlabel('time relative to coalescence (s)')
 pylab.ylabel(r'strain amplitude')
-pylab.subplots_adjust(left=0.075, right=0.97, top=0.98, bottom=0.225)
+pylab.subplots_adjust(left=0.075, right=0.97, top=0.95, bottom=0.225)
 pylab.savefig(sys.argv[1])

### 2011-06-29 inspiral_svd.tex
Removed 'The' from first sentence of abstract: 'Rapid detection of compact binary coalescence...'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1178 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{abstract}
-The rapid detection of compact binary coalescence with a network of advanced
+Rapid detection of compact binary coalescence with a network of advanced
 gravitational-wave detectors will offer a unique opportunity for

### 2011-06-29 introduction.tex
Removed fluff in section 1, paragraph 1: 'It is thought that...'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1180 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 orbital separation decays, leading to a runaway inspiral with the \GW{}
-amplitude and frequency increasing until the system eventually merges.  It is
-thought that if a neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
+amplitude and frequency increasing until the system eventually merges.
+If a neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
 the merger and fuel a bright electromagnetic (\EM{}) counterpart

### 2011-06-29 introduction.tex
In section 1, paragraph 1, switched 'discuss' with 'present'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1181 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 emission. We have the ambition of reporting \GW{} candidates not minutes after
-the merger, as has already been accomplished, but seconds before.  We discuss
+the merger, as has already been accomplished, but seconds before.  We present
 one essential ingredient in this effort, a computationally inexpensive
 low-latency filtering algorithm for detecting inspiral signals in \GW{}
-data.  We also present the prospects for advanced \GW\ detectors and
+data.  We also discuss the prospects for advanced \GW\ detectors and
 discuss other areas of work that would be required for rapid analysis.

### 2011-06-29 introduction.tex
Removed fluff in paragraph 1 of section 1.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1182 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Though compact binary coalescence (\CBC{}) events may be responsible for
-several classes of \EM{} transients, \CBC{}s are in particular believed to be a
+Compact binary coalescence (\CBC) is thought to be a
 mechanism for short gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}.

### 2011-06-29 introduction.tex
Removed some text that was commented out
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1183 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \citep{2008arXiv0805.2366I}.  However, 
-%if we can take galactic luminosities to
-%accurately reflect mass and thus be a proxy for the \CBC{} event rate, 
 it is possible to reduce the localization uncertainty by only looking at

### 2011-06-29 introduction.tex
Removed more commented out text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1184 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed citation to Numerical Recipes, since it is cited again later in the text (and overlap-save is a standard algorithm)
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1185 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \GW{} trigger-generation latencies of 2--5 minutes.
-%We assume that in the
-%advanced detector era the vetting process will be automated, so current \GW{}
-%search methodology and telescope actuation would dominate latency.
 
@@
 low latency at the cost of low throughput.  The former are epitomized by the
-overlap-save algorithm % \citep{numerical-recipes-chapter-13}
+overlap-save algorithm
 for frequency domain (\FD) convolution, currently the preferred method in \GW{}

### 2011-06-29 introduction.tex
In section 1.1, removed some fluff in sentence about formation channels for proto-magnetars
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1186 ccdd05d5-56c1-461a-bf98-85f6876523b4
Cleared up a nonsensical statement about short GRBs in section 1.1 (if the light curve fades too quickly to catch the afterglow, then how do we know about short GRBs in the first place?
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1187 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 a proto-magnetar that interacts with
-ejecta \citep{Bucciantini2011}.  In the case of the proto-magnetar, it might be
-formed from accretion-induced collapse, or the merger of two \textsc{ns}s.  An
+ejecta \citep{Bucciantini2011}.  A proto-magnetar may
+form from accretion-induced collapse or the merger of two \textsc{ns}s.  An
 accompanying \GW{} observation would confirm the latter.
@@
 tens of seconds and decayed quickly.  Short \GRB{}s, on the other hand,
-typically fade too quickly to even catch the tails of the afterglows in any
+typically fade so quickly that it is difficult to catch even the tails of the afterglows in any
 band. Rapid \GW{} transient alerts could enable the first observation of prompt

### 2011-06-29 introduction.tex
Removed some fluff in sentence listing sky localization uncertainties from Table 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1189 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 event with a final \SNR\ of $\sim 27$ whose location can be constrained to about
-1300\,deg$^2$ (3.1\% of the sky) using only the waveform 25\,s before merger,
-260\,deg$^2$ (0.63\% of the sky) using the waveform 10\,s before merger, and
-0.82\,deg$^2$ (0.0020\% of the sky) using the full \GW\ waveform.
+1300\,deg$^2$ (3.1\% of the sky) within 25\,s of merger,
+260\,deg$^2$ (0.63\% of the sky) within 10\,s of merger, and
+0.82\,deg$^2$ (0.0020\% of the sky) at merger.
 

### 2011-06-29 introduction.tex
In section 1.2, used an approximation symbol instead of an order of magnitude symbol
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1190 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Advanced \LIGO{} and Advanced Virgo.  Once per year, we expect to observe an
-event with a final \SNR\ of $\sim 27$ whose location can be constrained to about
+event with a final \SNR\ of $\approx 27$ whose location can be constrained to about
 1300\,deg$^2$ (3.1\% of the sky) within 25\,s of merger,

### 2011-06-29 introduction.tex
In section 1.2, mentioned that the 90% confidence area from Fairhurst (2009) is an overestimate
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1191 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 followup requires estimating the location of the \GW{} source. The localization
-uncertainty is dominated by uncertainty in the time of arrival of the \GW{}s,
+uncertainty can be estimated from the uncertainty in the time of arrival of the \GW{}s,
 which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
+\citep{Fairhurst2009}, though this overestimates the area on the sky for large
+timing errors.  Table~\ref{fig:earlywarning} and
 Figure~\ref{fig:sky-localization-accuracy} show the estimated minimum 90\%

### 2011-06-29 introduction.tex
In section 1.2, minor stylistic improvements to sentence about galaxy catalog completeness in Advanced LIGO versus Initial LIGO
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1192 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the number of galaxies that can produce a given signal amplitude is much larger
-than the number for initial \LIGO{} and thus the catalog will not be as useful
+than in Initial \LIGO{} and thus the catalog will not be as useful
 for downselecting pointings generally. However, exceptional \GW{} sources will

### 2011-06-29 introduction.tex
Section 1.3: this is still not in strict chronological order, and I think that readers can infer that it is in loosely chronological order, so we no longer explicitly state that it is
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1193 ccdd05d5-56c1-461a-bf98-85f6876523b4
Tuned latency budget in section 1.3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1194 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 There were a number of sources of latency associated with \CBC{}
-analysis in S6/VSR3 \citep{HugheyGWPAW2011}, listed here in chronological order.
+analysis in S6/VSR3 \citep{HugheyGWPAW2011}, listed here.
 %
@@
 %
-\item {\it Data aggregation and broadcasting ($\sim$$0.25$\,s):}
-The \LIGO{} data acquisition system records science
-data in 1/16\,s blocks \citep{Bork2001}. Data must also be aggregated from all
-of the \GW\ observatories using global computer networks capable of high
-bandwidth but perhaps only modest latency.  This could introduce an additional
-latency of $\sim$100\,ms.  These latencies are much smaller than the
-other sources of latency below and would require significant reengineering and
-capital investments to overcome.
+\item {\it Data acquisition and aggregation ($\gtrsim$100~ms):}
+The \LIGO\ data acquisition system collects data from detector subsystems 16
+times a second~\citep{Bork2001}. Data are also copied from all of the \GW\
+observatories to the analysis clusters over the Internet, which is capable of
+high bandwidth but perhaps only modest latency.  Together, these introduce a
+latency of $\gtrsim 100$~ms.  These technical sources of latency could be reduced
+with significant engineering and capital investments, but they are minor compared
+to any of the other sources of latency.
 %
-\item {\it Data calibration, quality control, and distribution
-($\sim$$2$\,min):} There is technical latency in calibrating the detector data,
-assessing its quality, and distributing it to a central location for analysis.
-None of this latency is fundamental and improvements in other areas could make
-this the dominating source of latency, so work should be done to improve this.
+\item {\it Data conditioning
+($\sim$1~min):} Science data must be calibrated using the detector's frequency
+response to gravitational radiation.  Currently, data are calibrated in blocks of
+16~s.  Within $\sim$1~min, data quality is assessed in order to create veto flags.
+These are both technical sources of latency that might be addressed with improved
+calibration and data quality software for advanced detectors.
 %
-\item {\it Trigger production ($2$--$5$\,min):} This was an impressive latency
-and was certainly sub-dominant in the S6/VSR3 analysis.  However, even if no
-other sources of latency exist, triggering latency is too long to catch prompt
-or extended
-emission. It is also more challenging for searches with advanced detectors
-because the templates needed to filter the data are about ten times longer.  We
-will focus on reducing the latency of trigger production in this work.
+\item {\it Trigger generation (2--5~min):} Low-latency data analysis pipelines
+deployed in S6/VSR3 achieved an impressive latency of minutes.  However, second
+to the human vetting process, this dominated the latency of the entire \EM\
+followup process.  Even if no other sources of latency existed, this trigger
+generation latency is too long to catch prompt or even extended emission.
+Low-latency trigger generation will become more challenging with advanced detectors
+because inspiral signals will stay in band up to ten times longer.  In this work,
+we will focus on reducing this source of latency.
 %
-\item {\it Event gathering, downselection, and pointing calculations
-($2$--$3$\,min):} S6/VSR3 saw the introduction of low-latency astronomical
+\item {\it Alert generation
+(2--3~min):} S6/VSR3 saw the introduction of low-latency astronomical
 alerts, which required gathering event parameters and sky localization from the
 various online analyses, downselecting the events, and calculating pointings.
-If other sources of latency improve, this could dominate, so work should be
-done to improve this.
+If other sources of latency improve, the technical latency associated with this
+infrastructure could dominate, so work should be done to improve it.
 %
-\item {\it Human vetting of the trigger quality ($10$--$20$\,min):} In
-commissioning the new alert system in S6/VSR3, humans manually examined each
-candidate for obvious flaws before publishing an alert. This was by far the
-largest latency during S6/VSR3 and hopefully confidence will grow to the
-point where human operators are satisfied with checks after alerts have been
-transmitted. We give it no further consideration here.
+\item {\it Human validation (10--20~min):} Because the new alert system was
+commissioned during S6/VSR3, humans examined each candidate for obvious flaws
+before publishing an alert. This was by far the largest latency during S6/VSR3.
+Hopefully, confidence in the system will grow to the point where no human intervention
+is necessary before alerts are sent, so we give it no further consideration here.
 %

### 2011-06-29 introduction.tex
Worked around bug in hyperref package
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1195 ccdd05d5-56c1-461a-bf98-85f6876523b4
added references and blurb on IIR
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1196 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsection{\EM{} counterparts}
+\subsection{\EM\ counterparts}
 
-Compact binary coalescence (\CBC) is thought to be a
+Compact binary coalescence (\CBC) events are thought to be a
 mechanism for short gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}.
@@
 configuration to send astronomical alerts, namely \textsc{mbta}, Coherent
-WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency}. The S6 analyses
+WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency, S6lowlatency2, S6lowlatency3, S6lowlatency4}. The S6 analyses
 achieved latencies of 30--60 minutes, which were dominated by a human vetting
@@
 low latency at the cost of low throughput.  The former are epitomized by the
-overlap-save algorithm
-for frequency domain (\FD) convolution, currently the preferred method in \GW{}
-searches.  The most obvious example of the latter is the time domain (\TD)
-convolution, which is latency-free.  However, its computational complexity is
-quadratic in the length of the templates, so it is prohibitively expensive for
-long templates.
+overlap-save algorithm for frequency domain (\FD) convolution, currently the
+preferred method in \GW{} searches.  The most obvious example of the latter is
+the time domain (\TD) convolution, which is latency-free.  However, its
+computational complexity is quadratic in the length of the templates, so it is
+prohibitively expensive for long templates.  We note that \citep{shaunIIR} have
+explored an infinite-impulse-response (IIR) decomposition of \CBC\ waveforms as
+a means to reduce the computational complexity of \TD\ filtering.  They
+demonstrate a factor of 50 improvement in floating point operations per second
+over the standard \TD\ method.  We will significantly improve upon that number.
 

### 2011-06-29 inspiral_svd.tex
Shortened running title to 'Early-warning detection of GW inspirals'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1179 ccdd05d5-56c1-461a-bf98-85f6876523b4
mireia asked to remove accnt
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1197 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\shorttitle{Early-warning of compact binary coalescence}
+\shorttitle{Early-warning detection of \GW\ inspirals}
 
@@
 	Adrian Chapman\altaffilmark{3},
-	Mireia Crisp\'{i}n-Ortuzar\altaffilmark{4},
+	Mireia Crispin-Ortuzar\altaffilmark{4},
 	Nickolas Fotopoulos\altaffilmark{3},

### 2011-06-29 introduction.tex
red flag pass
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1199 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 orbital separation decays, leading to a runaway inspiral with the \GW{}
-amplitude and frequency increasing until the system eventually merges.
-If a neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
+amplitude and frequency increasing until the system eventually merges.  If a
+neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
 the merger and fuel a bright electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  The \GW{} signal is in principle detectable
-before merger.  Effort from both the \GW{} and astronomy communities may make
-it possible to use \GW{} observations as an early warning trigger for \EM{}
-followup. In the first generation of ground-based laser interferometers, the \GW{}
-community initiated a project to send alerts when potential \GW{} transients
-were observed in order to trigger followup observations by \EM{} telescopes.
-The typical latencies were 30 minutes \citep{HugheyGWPAW2011}, which was an
-important achievement, but too late to catch any prompt \EM{}
-emission. We have the ambition of reporting \GW{} candidates not minutes after
-the merger, as has already been accomplished, but seconds before.  We present
-one essential ingredient in this effort, a computationally inexpensive
-low-latency filtering algorithm for detecting inspiral signals in \GW{}
-data.  We also discuss the prospects for advanced \GW\ detectors and
-discuss other areas of work that would be required for rapid analysis.
+\citep{shibata:2007}.  The \GW{} signal is in principle detectable before
+merger.  Effort from both the \GW{} and astronomy communities may make it
+possible to use \GW{} observations as an early warning trigger for \EM{}
+followup. In the first generation of ground-based laser interferometers, the
+\GW{} community initiated a project to send alerts when potential \GW{}
+transients were observed in order to trigger followup observations by \EM{}
+telescopes.  The typical latencies were 30 minutes \citep{HugheyGWPAW2011},
+which was an important achievement, but too late to catch any prompt \EM{}
+emission. We present one essential ingredient needed to report \GW{} candidates
+not minutes after the merger, as has already been accomplished, but seconds
+before.  This ingrediant is a computationally inexpensive low-latency filtering
+algorithm for detecting inspiral signals in \GW{} data.  We also discuss the
+prospects for advanced \GW\ detectors and discuss other areas of work that
+would be required for rapid analysis.
 
@@
 
-Compact binary coalescence (\CBC) events are thought to be a
-mechanism for short gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}.
-In this scenario, prompt gamma-ray emission arises on the accretion timescale
-of a central compact object formed after the merger, $0.1$--$1.0$\,s
+Compact binary coalescence (\CBC) events are thought to be a mechanism for
+short gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}.  In this
+scenario, prompt gamma-ray emission arises on the accretion timescale of a
+central compact object formed after the merger, $0.1$--$1.0$\,s
 \citep{Janka1999}.  Interestingly, roughly a quarter to half of observed short
@@
 accretion caused by $r$-process heating \citep{Metzger2010} or the formation of
-a proto-magnetar that interacts with
-ejecta \citep{Bucciantini2011}.  A proto-magnetar may
-form from accretion-induced collapse or the merger of two \textsc{ns}s.  An
-accompanying \GW{} observation would confirm the latter.
+a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}.  A
+proto-magnetar may form from accretion-induced collapse or the merger of two
+\textsc{ns}s.  An accompanying \GW{} observation would confirm the latter.
 
@@
 tens of seconds and decayed quickly.  Short \GRB{}s, on the other hand,
-typically fade so quickly that it is difficult to catch even the tails of the afterglows in any
-band. Rapid \GW{} transient alerts could enable the first observation of prompt
-optical flashes and the rise of afterglows from short \GRB{}s.
+typically fade so quickly that it is difficult to catch even the tails of the
+afterglows in any band. Rapid \GW{} transient alerts could enable the
+observation of prompt optical flashes and the rise of afterglows from short
+\GRB{}s.
 

### 2011-06-29 results.tex
red flag pass
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1200 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 rates up to $\sim$$10^{10}$ \flops{}, and assuming that other regions of the
-parameter space have similar computational scaling, an entire search could be
-implemented with $\gtrsim$$10$ machines.  Template length does vary
-over the parameter space; lower-mass templates are longer and would require
-more computation to analyze while higher-mass templates are shorter and would
-require less. However, we consider the estimates based on this sub-bank to be a
-reasonable representation.
+parameter space have similar computational scaling, an entire single detector
+search could be implemented with $\gtrsim$$10$ machines.  Template length does
+vary over the parameter space; lower-mass templates are longer and would
+require more computation to analyze while higher-mass templates are shorter and
+would require less. However, we consider the estimates based on this sub-bank
+to be a reasonable representation.
 

### 2011-06-29 conclusions.tex
softened comp reqs a bit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1201 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The \lloid{} method is as fast as standard \fft{} convolutions but allows for a
-technically latency-free, real-time operation.  We anticipate requiring
-$\gtrsim 4 \times 10 = 40$ modern multi-core computers to search for binary
-neutron stars using coincident \GW\ data from a four-detector network.  In the
-future, additional computational savings could be acheived by conditionally
-reconstructing the \SNR{} time-series only during times when a composite
-detection statistic crosses a threshold~\citep{svd-compdetstat}.  However, the
-anticipated required number of computers is well within the current computing
-capabilities of the \LIGO{} Data
+technically latency-free, real-time operation.  We anticipate requiring $\sim
+100$ modern multi-core computers to search for binary neutron stars using
+coincident \GW\ data from a multi-detector network.  In the future, additional
+computational savings could be acheived by conditionally reconstructing the
+\SNR{} time-series only during times when a composite detection statistic
+crosses a threshold~\citep{svd-compdetstat}.  However, the anticipated required
+number of computers is well within the current computing capabilities of the
+\LIGO{} Data
 Grid\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.

### 2011-06-29 method.tex
Converted diagram to eps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1165 ccdd05d5-56c1-461a-bf98-85f6876523b4
Minor rewording
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1204 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	\begin{center}
-		\includegraphics{figures/lloid-diagram.pdf}
+		\includegraphics{figures/lloid-diagram}
 		\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
 signal flow.  Circles with arrows represent interpolation
-\protect\includegraphics{figures/upsample-symbol.pdf} or decimation
-\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
+\protect\includegraphics{figures/upsample-symbol} or decimation
+\protect\includegraphics{figures/downsample-symbol}.  Circles with plus
 signs represent summing junctions
-\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
-\protect\includegraphics{figures/fir-symbol.pdf} stand for \fir{} filters.  Sample
+\protect\includegraphics{figures/adder-symbol}.  Squares
+\protect\includegraphics{figures/fir-symbol} stand for \fir{} filters.  Sample
 rate decreases from the top of the diagram to the bottom.  In this diagram each
@@
 The integral is evaluated using the power-law form of $f(t)$ from equation~(\ref{eq:fgw}).
-Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{hi} = 1/1570$~s $f_\mathrm{hi} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires a factor $\sim 10^{-6}$ fewer \flops\ than the conventional \TD\ method.
+Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{hi} = 1/1570$~s $f_\mathrm{hi} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the conventional \TD\ method.

### 2011-06-29 introduction.tex
Removed comment about overestimate of A(90%)
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1203 ccdd05d5-56c1-461a-bf98-85f6876523b4
Rewording of first paragraph in introduction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1205 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the merger and fuel a bright electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  The \GW{} signal is in principle detectable before
-merger.  Effort from both the \GW{} and astronomy communities may make it
+\citep{shibata:2007}.  Effort from both the \GW{} and astronomy communities may make it
 possible to use \GW{} observations as an early warning trigger for \EM{}
@@
 which was an important achievement, but too late to catch any prompt \EM{}
-emission. We present one essential ingredient needed to report \GW{} candidates
-not minutes after the merger, as has already been accomplished, but seconds
-before.  This ingrediant is a computationally inexpensive low-latency filtering
-algorithm for detecting inspiral signals in \GW{} data.  We also discuss the
-prospects for advanced \GW\ detectors and discuss other areas of work that
-would be required for rapid analysis.
+emission.  Since the \GW\ signal is in principle detectable even before the tidal
+disruption, one may have the ambition of reporting \GW\ candidates not minutes
+after the merger, but seconds before.  We explore one essential ingredient of this
+problem, a computationally inexpensive low-latency filtering algorithm for detecting
+inspiral signals in \GW{} data.  We also consider the prospects for advanced \GW{}
+detectors and discuss other areas of work that would be required for rapid analysis.
 
@@
 which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}, though this overestimates the area on the sky for large
-timing errors.  Table~\ref{fig:earlywarning} and
+\citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
 Figure~\ref{fig:sky-localization-accuracy} show the estimated minimum 90\%

### 2011-06-29 introduction.tex
Reworded caveat in section 1.2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1206 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-We emphasize that Figure~\ref{fig:earlywarning} provides only the lower bound
-for early detection and should serve as a baseline to understand the potential
-of early warning \GW{} searches in a ideal world.  Real analysis will have
-sources of latency that are unavoidable meaning that the actual detection will
-occur with some delay.
+We emphasize that any practical \GW\ search will include technical delays due
+to light travel time between the detectors, science infrastructure, and the
+selected data analysis strategy.  Figure~\ref{fig:earlywarning} must be understood
+in the context of all of the potential sources of latency, some of which are avoidable
+and some of which are not.
 

### 2011-06-29 introduction.tex
Reworded mention of IIR filters
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1207 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 computational complexity is quadratic in the length of the templates, so it is
-prohibitively expensive for long templates.  We note that \citep{shaunIIR} have
-explored an infinite-impulse-response (IIR) decomposition of \CBC\ waveforms as
-a means to reduce the computational complexity of \TD\ filtering.  They
+prohibitively expensive for long templates.  \citet{shaunIIR} explored decomposing
+\CBC\ matched filters into banks of infinite impulse response (\textsc{iir})
+filters as a means to reduce the computational cost of latency-free data analysis.  They
 demonstrate a factor of 50 improvement in floating point operations per second

### 2011-06-29 introduction.tex
Polished some sentence structure in section 1.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1209 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 scenario, prompt gamma-ray emission arises on the accretion timescale of a
-central compact object formed after the merger, $0.1$--$1.0$\,s
-\citep{Janka1999}.  Interestingly, roughly a quarter to half of observed short
+central compact object formed $0.1$--$1.0$~s after the merger~\citep{Janka1999}.
+Interestingly, roughly a quarter to half of observed short
 \GRB{}s also exhibit extended emission of $30$--$100$\,s in duration beginning

### 2011-06-29 introduction.tex
Moved sky localization discussion into its own section; merged subsections in introduction; juggled paragraphs discussing EM counterparts
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1210 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\subsection{\EM\ counterparts}
-
 Compact binary coalescence (\CBC) events are thought to be a mechanism for
-short gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}.  In this
+short gamma-ray bursts (\GRB{}s)~\citep{Lee:2005, nakar07}.  In this
 scenario, prompt gamma-ray emission arises on the accretion timescale of a
 central compact object formed $0.1$--$1.0$~s after the merger~\citep{Janka1999}.
-Interestingly, roughly a quarter to half of observed short
-\GRB{}s also exhibit extended emission of $30$--$100$\,s in duration beginning
-$\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the initial
-outburst.  Potential explanations for the emission are delayed fall-back
-accretion caused by $r$-process heating \citep{Metzger2010} or the formation of
-a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}.  A
-proto-magnetar may form from accretion-induced collapse or the merger of two
-\textsc{ns}s.  An accompanying \GW{} observation would confirm the latter.
-
 Optical flashes have only been observed for a handful of long \GRB{}s
@@
 
-\subsection{Early-warning \EM{} followup}
+Interestingly, roughly a quarter to half of observed short
+\GRB{}s also exhibit extended emission of $30$--$100$\,s in duration beginning
+$\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the initial
+outburst.  Potential explanations for the emission are delayed fall-back
+accretion caused by $r$-process heating \citep{Metzger2010} or the formation of
+a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}.  A
+proto-magnetar may form from accretion-induced collapse or the merger of two
+\textsc{ns}s.  An accompanying \GW{} observation would confirm the latter.
 
@@
 
-\begin{figure}[h]
-\plotone{figures/snr_in_time}
-\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
-sources that could be detectable by Advanced \LIGO\ a given number of seconds
-before coalescence.  The heavy solid line is the most realistic yearly rate
-estimate.  The shaded region represents the 5 to 95\% confidence interval
-arising from uncertainty in predicted event rates \citep{Abadie:2010p10836}.}
-\end{figure}
-%
-Examining the signal to noise ratio (\SNR{}) stream for threshold crossings
-before the \GW{} signal leaves the detection band makes it possible to trade
-some \SNR{} and sky localization accuracy for pre-merger detection.
-Figure~\ref{fig:earlywarning} shows projected early detectability rates for
-\textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
-predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
-the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
-The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
-observe a total of 40\,events\,yr$^{-1}$. $\sim$10\,yr$^{-1}$ will be detectable
-within 10\,s of merger and $\sim$5\,yr$^{-1}$ will be detectable within 25\,s of
-merger if analysis can proceed with near zero latency. These rates are shown
-with their substantial uncertainties in Figure~\ref{fig:earlywarning}.
-
-We emphasize that any practical \GW\ search will include technical delays due
-to light travel time between the detectors, science infrastructure, and the
-selected data analysis strategy.  Figure~\ref{fig:earlywarning} must be understood
-in the context of all of the potential sources of latency, some of which are avoidable
-and some of which are not.
-
-\begin{table}[h]
-\caption{\label{table:sky-localization-accuracy}Horizon distance, \SNR\ at
-merger, and area of 90\% confidence at selected times before merger for sources
-with expected detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$.}
-\begin{center}
-\begin{tabular}{rrrrrrr}
-\tableline\tableline
-rate & horiz. & final & \multicolumn{4}{c}{$A$(90\%) (deg$^2$)} \\
-\cline{4-7}
-yr$^{-1}$ & (Mpc) & \SNR\ & 25 s & 10 s & 1 s & 0 s \\
-\tableline
-40\phd\phn & 445 & 8.0 & 15500 & 3100 & 200 & 9.6 \\
-10\phd\phn & 280 & 12.7 & 6100 & 1200 & 78 & 3.8 \\
-1\phd\phn & 130 & 27.4 & 1300 & 260 & 17 & 0.8 \\
-0.1 & 60 & 58.9 & 280 & 56 & 3.6 & 0.2 \\
-\tableline
-\end{tabular}
-\end{center}
-\end{table}
-%
-\begin{figure}[h]
-\plotone{figures/loc_in_time}
-\caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
-region as a function of time before coalescence for sources with anticipated
-detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$. The heavy dot indicates
-the time at which the accumulated \SNR\ exceeds a threshold of 8.}
-\end{figure}
-
-Before discussing realistic latencies, it is important to note that \EM{}
-followup requires estimating the location of the \GW{} source. The localization
-uncertainty can be estimated from the uncertainty in the time of arrival of the \GW{}s,
-which is determined by the signal's effective bandwidth and \SNR{}
-\citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
-Figure~\ref{fig:sky-localization-accuracy} show the estimated minimum 90\%
-confidence area versus time of the loudest coalescence events detectable by
-Advanced \LIGO{} and Advanced Virgo.  Once per year, we expect to observe an
-event with a final \SNR\ of $\approx 27$ whose location can be constrained to about
-1300\,deg$^2$ (3.1\% of the sky) within 25\,s of merger,
-260\,deg$^2$ (0.63\% of the sky) within 10\,s of merger, and
-0.82\,deg$^2$ (0.0020\% of the sky) at merger.
-
-It is infeasible to search hundreds of square degrees for a prompt counterpart.
-For reference, \textsc{lsst} should see first light around the same time as
-Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
-\citep{2008arXiv0805.2366I}.  However, 
-it is possible to reduce the localization uncertainty by only looking at
-galaxies from a catalog that lie near the sky location and luminosity distance
-estimate from the \GW{} signal~\citep{galaxy-catalog} as was done in S6/VSR3.
-Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon distance,
-the number of galaxies that can produce a given signal amplitude is much larger
-than in Initial \LIGO{} and thus the catalog will not be as useful
-for downselecting pointings generally. However, exceptional \GW{} sources will
-necessarily be extremely nearby. Within this reduced volume there will be fewer
-galaxies to consider for a given candidate and catalog completeness will be
-less of a concern.  This should reduce the 90\% confidence area substantially.
-Also future \GW{} detectors, should they be built, will improve the \SNR\ and
-likewise the error box.
-
-\subsection{Latency in \CBC{} searches}
-
 There were a number of sources of latency associated with \CBC{}
 analysis in S6/VSR3 \citep{HugheyGWPAW2011}, listed here.
-%
-\begin{itemize}
-%
-\item {\it Data acquisition and aggregation ($\gtrsim$100~ms):}
+
+\paragraph{Data acquisition and aggregation ($\gtrsim$100~ms)}%
 The \LIGO\ data acquisition system collects data from detector subsystems 16
@@
 to any of the other sources of latency.
-%
-\item {\it Data conditioning
-($\sim$1~min):} Science data must be calibrated using the detector's frequency
+
+\paragraph{Data conditioning ($\sim$1~min)}%
+Science data must be calibrated using the detector's frequency
 response to gravitational radiation.  Currently, data are calibrated in blocks of
@@
 calibration and data quality software for advanced detectors.
-%
-\item {\it Trigger generation (2--5~min):} Low-latency data analysis pipelines
+
+\paragraph{Trigger generation (2--5~min)}%
+Low-latency data analysis pipelines
 deployed in S6/VSR3 achieved an impressive latency of minutes.  However, second
@@
 we will focus on reducing this source of latency.
-%
-\item {\it Alert generation
-(2--3~min):} S6/VSR3 saw the introduction of low-latency astronomical
+
+\paragraph{Alert generation (2--3~min)}%
+S6/VSR3 saw the introduction of low-latency astronomical
 alerts, which required gathering event parameters and sky localization from the
@@
 infrastructure could dominate, so work should be done to improve it.
-%
-\item {\it Human validation (10--20~min):} Because the new alert system was
+
+\paragraph{Human validation (10--20~min)}%
+Because the new alert system was
 commissioned during S6/VSR3, humans examined each candidate for obvious flaws
@@
 is necessary before alerts are sent, so we give it no further consideration here.
-%
-\end{itemize}
+
+\paragraph{}
 
@@
 
-The paper is organized as follows. First we provide an overview of our method
-for detecting compact binary coalescence signals in an \earlywarning\ analysis.
-We then describe the pipeline we have constructed that implements our method.
-To validate the approach we present results of simulations and conclude with
-some remarks on what remains to prepare for the advanced detector era.
+The paper is organized as follows.  First, we discuss prospects for early-warning
+detection.  Then, we provide an overview of our novel method for detecting compact
+binary coalescence signals with extremely low latency. We then describe a prototype
+implementation using open source signal processing software.  To validate our approach
+we present results of simulations and conclude with some remarks on what remains to
+prepare for the advanced detector era.

### 2011-06-29 method.tex
Moved sky localization discussion into its own section; merged subsections in introduction; juggled paragraphs discussing EM counterparts
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1210 ccdd05d5-56c1-461a-bf98-85f6876523b4
Deleted commented-out text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1211 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\section{Early warning searches for compact binary coalescence}
+\section{Novel data analysis strategy}
 \label{sec:method}
@@
 
-\subsection{Conventional \CBC{} matched filter searches}
+\subsection{Conventional \CBC{} searches}
 
@@
 via \SVD{} of the time-sliced templates.
-
-\begin{comment}
-An example time slice design satisfying these constraints for a $1.4 - 1.4 \,
-M_{\odot}$ binary is shown in Table~\ref{table:time_slices}.
-%
-\begin{table}[h!]
-\begin{minipage}[c]{0.52\textwidth}
-\centering
-\vspace{0.8cm}
-\includegraphics{time_slices.pdf}
-\end{minipage}
-\begin{minipage}[c]{0.3\textwidth}
-\centering
-\input{time_slices.tex}
-\end{minipage}
-\caption{\label{table:time_slices} Example of nearly critically sampled,
-power-of-two time slices for a $1.4 - 1.4 \, M_{\odot}$ template extending from
-$f_\mathrm{low} = 10 \, \mathrm{Hz}$ to $f_\mathrm{hi} = f_\mathrm{ISCO} = 1571\, \mathrm{Hz}$
-with a time-frequency structure given by ($\ref{eq:fgw})$. $f^s$ is the sample
-rate of the time slice, $(t^{s+1}, t^s]$ are the boundaries in seconds
-preceding coalescence and \slicessamps\ is the number of samples in the
-$s^{\mathrm{th}}$ filter.}
-\end{table}
-\end{comment}
 

### 2011-06-30 prospects.tex
Added missing file
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1214 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{Prospects for \earlywarning\ detection and \EM\ followup}

\begin{figure}[h]
\plotone{figures/snr_in_time}
\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
sources that could be detectable by Advanced \LIGO\ a given number of seconds
before coalescence.  The heavy solid line is the most realistic yearly rate
estimate.  The shaded region represents the 5 to 95\% confidence interval
arising from uncertainty in predicted event rates \citep{Abadie:2010p10836}.}
\end{figure}
%
Examining the signal to noise ratio (\SNR{}) stream for threshold crossings
before the \GW{} signal leaves the detection band makes it possible to trade
some \SNR{} and sky localization accuracy for pre-merger detection.
Figure~\ref{fig:earlywarning} shows projected early detectability rates for
\textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
observe a total of 40\,events\,yr$^{-1}$. $\sim$10\,yr$^{-1}$ will be detectable
within 10\,s of merger and $\sim$5\,yr$^{-1}$ will be detectable within 25\,s of
merger if analysis can proceed with near zero latency. These rates are shown
with their substantial uncertainties in Figure~\ref{fig:earlywarning}.

We emphasize that any practical \GW\ search will include technical delays due
to light travel time between the detectors, science infrastructure, and the
selected data analysis strategy.  Figure~\ref{fig:earlywarning} must be understood
in the context of all of the potential sources of latency, some of which are avoidable
and some of which are not.

\begin{table}[h]
\caption{\label{table:sky-localization-accuracy}Horizon distance, \SNR\ at
merger, and area of 90\% confidence at selected times before merger for sources
with expected detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$.}
\begin{center}
\begin{tabular}{rrrrrrr}
\tableline\tableline
rate & horiz. & final & \multicolumn{4}{c}{$A$(90\%) (deg$^2$)} \\
\cline{4-7}
yr$^{-1}$ & (Mpc) & \SNR\ & 25 s & 10 s & 1 s & 0 s \\
\tableline
40\phd\phn & 445 & 8.0 & 15500 & 3100 & 200 & 9.6 \\
10\phd\phn & 280 & 12.7 & 6100 & 1200 & 78 & 3.8 \\
1\phd\phn & 130 & 27.4 & 1300 & 260 & 17 & 0.8 \\
0.1 & 60 & 58.9 & 280 & 56 & 3.6 & 0.2 \\
\tableline
\end{tabular}
\end{center}
\end{table}
%
\begin{figure}[h]
\plotone{figures/loc_in_time}
\caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
region as a function of time before coalescence for sources with anticipated
detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$. The heavy dot indicates
the time at which the accumulated \SNR\ exceeds a threshold of 8.}
\end{figure}

\EM\ followup requires estimating the location of the \GW{} source. The localization
uncertainty can be estimated from the uncertainty in the time of arrival of the \GW{}s,
which is determined by the signal's effective bandwidth and \SNR{}
\citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
Figure~\ref{fig:sky-localization-accuracy} show the estimated minimum 90\%
confidence area versus time of the loudest coalescence events detectable by
Advanced \LIGO{} and Advanced Virgo.  Once per year, we expect to observe an
event with a final \SNR\ of $\approx 27$ whose location can be constrained to about
1300\,deg$^2$ (3.1\% of the sky) within 25\,s of merger,
260\,deg$^2$ (0.63\% of the sky) within 10\,s of merger, and
0.82\,deg$^2$ (0.0020\% of the sky) at merger.

It is infeasible to search hundreds of square degrees for a prompt counterpart.
For reference, \textsc{lsst} should see first light around the same time as
Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
\citep{2008arXiv0805.2366I}.  However, 
it is possible to reduce the localization uncertainty by only looking at
galaxies from a catalog that lie near the sky location and luminosity distance
estimate from the \GW{} signal~\citep{galaxy-catalog} as was done in S6/VSR3.
Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon distance,
the number of galaxies that can produce a given signal amplitude is much larger
than in Initial \LIGO{} and thus the catalog will not be as useful
for downselecting pointings generally. However, exceptional \GW{} sources will
necessarily be extremely nearby. Within this reduced volume there will be fewer
galaxies to consider for a given candidate and catalog completeness will be
less of a concern.  This should reduce the 90\% confidence area substantially.
Also future \GW{} detectors, should they be built, will improve the \SNR\ and
likewise the error box.

### 2011-06-30 conclusions.tex
Removed 'truly latency free'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1216 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The \lloid{} method is as fast as standard \fft{} convolutions but allows for a
-technically latency-free, real-time operation.  We anticipate requiring $\sim
-100$ modern multi-core computers to search for binary neutron stars using
+The \lloid{} method is as fast as standard \fft{} convolutions but allows for
+nearly latency-free, real-time operation.  We anticipate requiring $\sim$~100
+modern multi-core computers to search for binary neutron stars using
 coincident \GW\ data from a multi-detector network.  In the future, additional

### 2011-06-30 implementation.tex
Removed 'truly latency free'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1216 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The \lloid\ algorithm is amenable to truly latency free, real-time implementation.
-However, such a real-time search pipeline would likely require integration directly into
+The \lloid\ algorithm is amenable to nearly latency free, real-time implementation.
+However, such a near real-time search pipeline would likely require integration directly into
 the data acquisition and storage system of the \LIGO\ observatories.  A slightly more 

### 2011-06-30 method.tex
Tuned title of section 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1212 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed 'truly latency free'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1216 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\section{Novel data analysis strategy}
+\section{A novel near real-time algorithm}
 \label{sec:method}
@@
 Here we describe a method for reducing the computational cost of a \TD\ search
-for compact binary coalescence.  We give a truly zero-latency algorithm
+for compact binary coalescence.  We give a nearly zero-latency algorithm
 that competes in terms of floating point operations per second with the
@@
 \end{equation*}
+Ultimately, the latency of the entire \lloid\ algorithm is set by the decimation
+and interpolation filters because they are generally time-symmetric.  As a result,
+\lloid\ is very slightly acausal with an intrinsic latency of
+$\left(\max \left( N^\shortdownarrow, N^\shortuparrow \right) / 2 f^0 \right)$~s.
 

### 2011-06-30 conclusions.tex
Adjusted spacing
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1217 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We will explore these and also using the time-slice decomposition and the \SVD\
-to form low-latency signal-based vetoes (e.g.  $\chi^2$ statistics) that have
+to form low-latency signal-based vetoes (e.g.~$\chi^2$~statistics) that have
 been essential for glitch rejection used in previous \GW{} \CBC{} searches.

### 2011-06-30 inspiral_svd.tex
New title
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1202 ccdd05d5-56c1-461a-bf98-85f6876523b4
Bumped DCC version
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1208 ccdd05d5-56c1-461a-bf98-85f6876523b4
Moved sky localization discussion into its own section; merged subsections in introduction; juggled paragraphs discussing EM counterparts
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1210 ccdd05d5-56c1-461a-bf98-85f6876523b4
New title
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1

@@
 
-\title{Technique for \earlywarning\ filtering for compact binary coalescence
-and its implications for multi-messenger astronomy}
+\title{Toward \earlywarning\ detection of gravitational waves from compact binary coalescence}
 
@@
 }
-% we either need to include all of our emails, or none.
-%\email{leo.singer@ligo.org}
 
@@
 \altaffiltext{2}{D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
-\altaffiltext{3}{LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA} 
+\altaffiltext{3}{LIGO Laboratory, California Institute of Technology, MC 18-34, 1200 E. California Blvd., Pasadena, CA, USA; \texttt{leo.singer@ligo.org}}
 \altaffiltext{4}{Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot, Spain} 
@@
 \input{introduction}
+\input{prospects}
 \input{method}
@@
 
-This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{21}.
+This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{23}.
 

### 2011-06-30 introduction.tex
Removed duplicate citation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1220 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 configuration to send astronomical alerts, namely \textsc{mbta}, Coherent
-WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency, S6lowlatency2, S6lowlatency3, S6lowlatency4}. The S6 analyses
+WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}. The S6 analyses
 achieved latencies of 30--60 minutes, which were dominated by a human vetting

### 2011-06-30 inspiral_svd.tex
Corrected mail code
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1221 ccdd05d5-56c1-461a-bf98-85f6876523b4
abstract: Prompt detection alerts to -> Prompt detection alerts for
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1222 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \altaffiltext{2}{D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
-\altaffiltext{3}{LIGO Laboratory, California Institute of Technology, MC 18-34, 1200 E. California Blvd., Pasadena, CA, USA; \texttt{leo.singer@ligo.org}}
+\altaffiltext{3}{LIGO Laboratory, California Institute of Technology, MC 100-36, 1200 E. California Blvd., Pasadena, CA, USA; \texttt{leo.singer@ligo.org}}
 \altaffiltext{4}{Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot, Spain} 
@@
 gravitational-wave detectors will offer a unique opportunity for
-multi-messenger astronomy.  Prompt detection alerts to the astronomical
+multi-messenger astronomy.  Prompt detection alerts for the astronomical
 community may make it possible to observe the onset of electromagnetic emission

### 2011-06-30 introduction.tex
In introduction, clarified that AIC and compact binary coalescence are two possible channels, but not mutually exclusive
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1223 ccdd05d5-56c1-461a-bf98-85f6876523b4
Tuned paragraph in introduction describing S6/VSR3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1224 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 proto-magnetar may form from accretion-induced collapse or the merger of two
-\textsc{ns}s.  An accompanying \GW{} observation would confirm the latter.
+\textsc{ns}s.  An accompanying \GW{} observation would confirm the existence of
+the latter channel.
 
@@
 configuration to send astronomical alerts, namely \textsc{mbta}, Coherent
-WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}. The S6 analyses
-achieved latencies of 30--60 minutes, which were dominated by a human vetting
-process. Candidates were sent for \EM{} followup to several telescopes; Swift,
+WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
+\textsc{mbta} achieved the best \GW{} trigger-generation latencies of 2--5 minutes.
+Alerts were sent with latencies of 30--60 minutes, dominated by human vetting.
+Candidates were sent for \EM{} followup to several telescopes; Swift,
 \textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008, HugheyGWPAW2011}
-imaged some of the most likely sky locations.  \textsc{mbta} achieved the best
-\GW{} trigger-generation latencies of 2--5 minutes.
+imaged some of the most likely sky locations.
 

### 2011-06-30 introduction.tex
In latency list, changed 'calculating pointings' to 'calculating telescope pointings'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1225 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 alerts, which required gathering event parameters and sky localization from the
-various online analyses, downselecting the events, and calculating pointings.
+various online analyses, downselecting the events, and calculating telescope pointings.
 If other sources of latency improve, the technical latency associated with this

### 2011-06-30 introduction.tex
Tuned description of human validation in latency list
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1226 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \paragraph{Human validation (10--20~min)}%
-Because the new alert system was
-commissioned during S6/VSR3, humans examined each candidate for obvious flaws
-before publishing an alert. This was by far the largest latency during S6/VSR3.
+Because the new alert system was commissioned during S6/VSR3, all alerts were subjected
+to quality control checks by human operators before they were disseminated.
+This was by far the largest source of latency during S6/VSR3.
 Hopefully, confidence in the system will grow to the point where no human intervention

### 2011-06-30 introduction.tex
In description of Internet, changed 'perhaps only modest latency' to 'only modest latency'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1227 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 observatories to the analysis clusters over the Internet, which is capable of
-high bandwidth but perhaps only modest latency.  Together, these introduce a
+high bandwidth but only modest latency.  Together, these introduce a
 latency of $\gtrsim 100$~ms.  These technical sources of latency could be reduced

### 2011-06-30 introduction.tex
Tuned sentence about striking balance between latency and throughput
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1228 ccdd05d5-56c1-461a-bf98-85f6876523b4
Tuned sentence about sending prompt alerts to the astronomical community
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1229 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-This work will focus on reducing the latency of trigger production.  Realizing
-advance detection of \CBC{}s will require striking a balance between latency
-and throughput of filtering \GW\ data. \CBC{} searches consist of banks of
+This work will focus on reducing the latency of trigger production.  Data analysis
+strategies for advance detection of \CBC{}s will have to strike a balance between latency
+and throughput. \CBC{} searches consist of banks of
 matched filters, or cross-correlations between the data stream and a bank of
@@
 resources.  Assuming the other technical sources of latency can be reduced
-significantly, this should allow the possibility for prompt alerts to be sent
-to the astronomical community.
+significantly, this could make it possible to send prompt alerts to the
+astronomical community.
 

### 2011-06-30 conclusions.tex
Adjusted spacing between some figures and their units
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1231 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The \lloid{} method is as fast as standard \fft{} convolutions but allows for
-nearly latency-free, real-time operation.  We anticipate requiring $\sim$~100
+nearly latency-free, real-time operation.  We anticipate requiring $\sim$100
 modern multi-core computers to search for binary neutron stars using
@@
 advance to detect the onset of prompt emission, however localization prior to
-the onset of extended emission on the 10--100\,s time-scale is more easily
+the onset of extended emission on the 10--100~s time-scale is more easily
 achieved and will benefit from the full bandwidth and \SNR\ of the \GW{} event

### 2011-06-30 method.tex
Adjusted spacing between some figures and their units
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1231 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
-f(t) = \frac{1}{\mathcal{\pi M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
+f(t) = \frac{1}{\pi \mathcal{M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
 \right]^{3/8}.
@@
 Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
-M_\odot / c^3 \approx 5 \upmu\mathrm{s}$) and $t$ is the time relative to the
+M_\odot / c^3 \approx 5$~$\upmu$s) and $t$ is the time relative to the
 coalescence of the binary.

### 2011-06-30 prospects.tex
Adjusted spacing between some figures and their units
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1231 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
-observe a total of 40\,events\,yr$^{-1}$. $\sim$10\,yr$^{-1}$ will be detectable
-within 10\,s of merger and $\sim$5\,yr$^{-1}$ will be detectable within 25\,s of
+observe a total of 40~events~yr$^{-1}$. $\sim$10~yr$^{-1}$ will be detectable
+within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable within 25~s of
 merger if analysis can proceed with near zero latency. These rates are shown
@@
 merger, and area of 90\% confidence at selected times before merger for sources
-with expected detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$.}
+with expected detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
 \begin{center}
@@
 region as a function of time before coalescence for sources with anticipated
-detectability rates of 40, 10, 1, and 0.1\,yr$^{-1}$. The heavy dot indicates
-the time at which the accumulated \SNR\ exceeds a threshold of 8.}
+detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$. The heavy dot indicates
+the time at which the accumulated \SNR\ exceeds a threshold of~8.}
 \end{figure}
@@
 Advanced \LIGO{} and Advanced Virgo.  Once per year, we expect to observe an
-event with a final \SNR\ of $\approx 27$ whose location can be constrained to about
-1300\,deg$^2$ (3.1\% of the sky) within 25\,s of merger,
-260\,deg$^2$ (0.63\% of the sky) within 10\,s of merger, and
-0.82\,deg$^2$ (0.0020\% of the sky) at merger.
+event with a final \SNR\ of $\approx$27 whose location can be constrained to about
+1300~deg$^2$ (3.1\% of the sky) within 25~s of merger,
+260~deg$^2$ (0.63\% of the sky) within 10~s of merger, and
+0.82~deg$^2$ (0.0020\% of the sky) at merger.
 
@@
 For reference, \textsc{lsst} should see first light around the same time as
-Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
+Advanced \LIGO{} and will have an unparalleled field of view of 9.6~deg$^2$
 \citep{2008arXiv0805.2366I}.  However, 

### 2011-06-30 prospects.tex
'generally' -> 'for most events'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1232 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 than in Initial \LIGO{} and thus the catalog will not be as useful
-for downselecting pointings generally. However, exceptional \GW{} sources will
+for downselecting pointings for most events. However, exceptional \GW{} sources will
 necessarily be extremely nearby. Within this reduced volume there will be fewer

### 2011-06-30 method.tex
No need to reintroduce the abbreviation 'S6/VSR3' in section 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1234 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 multi-band decomposition of the compact binary signal space that resulted in
-a search with minutes latency in \LIGO{}'s S6 and Virgo's VSR3 joint science
-runs~\citep{HugheyGWPAW2011}.  We combine this with the \SVD\ rank-reduction
-method of \citet{Cannon:2010p10398} that exploits the redundancy of
-the template banks.
+a search with minutes latency during S6/VSR3~\citep{HugheyGWPAW2011}.  We combine this
+with the \SVD\ rank-reduction method of \citet{Cannon:2010p10398} that exploits
+the redundancy of the template banks.
 

### 2011-06-30 method.tex
Removed some fluffy commas in section 3.2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1235 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed some plural agreement issues in section 3.2.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1236 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-An inspiral signal will enter the detection band at some low frequency
-$f_\mathrm{low}$, at a time $t_\mathrm{low}$ before merger.  Usually the template
+An inspiral signal will enter the detection band with some low frequency
+$f_\mathrm{low}$ at time $t_\mathrm{low}$ before merger.  Usually the template
 is truncated at some prescribed time $t^0$, or equivalent frequency $f_\mathrm{hi}$,
@@
 sample rate.  Specifically we consider here time slice boundaries with the
-smallest power-of-two sample rate that sub-critically samples the time-sliced
-template.  The time slices consist of the $S$ intervals
+smallest power-of-two sample rates that sub-critically sample the time-sliced
+templates.  The time slices consist of the $S$ intervals
 $\left[t^0, t^1\right),\, \left[t^1, t^2\right),\, \dots,\, \left[t^{S-1}, t^S\right)$,

### 2011-06-30 results.tex
Added measured latency to flop/s table
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1215 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed 'truly latency free'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1216 ccdd05d5-56c1-461a-bf98-85f6876523b4
Adjusted spacing between some figures and their units
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1231 ccdd05d5-56c1-461a-bf98-85f6876523b4
In section 5.1, 'from 1 to 3 M_sun' sounds better to me than 'between 1 and 3 M_sun'
git-svn-id: https://svn.ligo.caltech.edu/svn/

@@
 masses.  We begin by constructing a template bank that spans component masses
-between 1~and~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
+from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
 curve~\citep{ALIGONoise}.  This results in a grid of 98544 points, or
@@
 a function of $N^\shortuparrow$ is shown in Figure~\ref{fig:bw}b.  The
-mismatch saturates at $\sim 2 \times 10^{-4}$ with $N^\shortuparrow = 64$.  We find
+mismatch saturates at $\sim$$2 \times 10^{-4}$ with $N^\shortuparrow = 64$.  We find
 that a filter length of 16 is sufficient to meet our target mismatch of 0.003.
@@
 time domain & $2.4\times10^{13}$ & 0 \\
-frequency domain & $2.6\times10^8$ & 2201 \\
-\lloid\ (this work) & $4.7\times10^8$ & 0 \\
+frequency domain & $2.6\times10^8$ & $2\times10^3$ \\
+\lloid\ (theory) & $4.7\times10^8$ & $2\times10^{-3}$ \\
+\lloid\ (prototype) & ----------- & 5 \\
 \tableline
@@
 might be reduced by carefully tuning buffer lengths at every stage in
-the pipeline.
+the pipeline.  The minimum theoretically possible latency is set by
+$N^\shortdownarrow$ and $N^\shortuparrow$.
 

### 2011-06-30 results.tex
Fixed spelling error in SVD tolerance paragraph
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1240 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We studied how the \SVD\ tolerance affected \SNR\ loss by holding
-$N^\shortuparrow = 192$ fixed as we vaired the \SVD\ tolerance from
+$N^\shortuparrow = 192$ fixed as we varied the \SVD\ tolerance from
 $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The minimum, maximum, and median

### 2011-06-30 results.tex
In section 5.3, corrected statement about LLOID being zero-latency (it has a latency of ~ms)
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1241 ccdd05d5-56c1-461a-bf98-85f6876523b4
Moved URL footnote for LIGO Data Grid to its first mention in the Results section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1242 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 five orders of magnitude faster than the conventional \TD\ method.  However,
-the \FD\ method has a latency of over half of an hour, whereas \lloid\ has no
-latency at all.
+the \FD\ method has a latency of over half of an hour, whereas \lloid\ has a
+latency of just $\sim$ms.
 %
@@
 $\sim$$10^6$ present-day machines to search the full parameter space.
-Presently, the \LIGO{} Data Grid consists of only $\sim$$10^4$ machines.
+Presently, the \LIGO{} Data Grid%
+\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}} consists of
+only $\sim$$10^4$ machines.
 

### 2011-06-30 conclusions.tex
Moved URL footnote for LIGO Data Grid to its first mention in the Results section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1242 ccdd05d5-56c1-461a-bf98-85f6876523b4
Deleted commented-out text in conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1243 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 number of computers is well within the current computing capabilities of the
-\LIGO{} Data
-Grid\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.
+\LIGO{} Data Grid.
 
@@
 latency filtering schemes such as the one described here.
-
-\begin{comment}
-The detection algorithm we described has no intrinsic latency.  However, there
-are fundamental and practical latencies associated with the analysis and
-detection procedure. For example, the \LIGO{} data acquisition records science
-data in 1/16~s blocks~\citep{Bork2001}. Data must also be aggregated from all
-of the \GW\ observatories using global computer networks capable of high
-bandwidth but perhaps only modest latency.  This could introduce a latency of
-$\sim$100~ms.  Lastly, the software implementation of the algorithm itself may
-introduce latency.  
-\end{comment}
 
@@
 
-\begin{comment}
-Another potential complication is that if adjacent time slices have non-negligible
-cross-correlation with each other, then there may be some subtleties in the design
-of the triggering and coincidence stages that we must address in order to translate
-the early-warning outputs into candidate events.  On the other hand, the time-slice
-decomposition and the \SVD\ may be beneficial for the postprocessing stage
-because they will help us form low-latency signal-based vetoes (e.g. $\chi^2$ statistics)
-that have been essential for glitch rejection used in previous \GW{} \CBC{}
-searches.
-\end{comment}
-
 Our future work must more deeply address sky localization accuracy in a
@@
 the techniques we have described.
-
-%Latency budget, including `before' and `after' quotes for:
-
-%\begin{itemize}
-%\item Data acquisition
-%\item Calibration
-%\item Data aggregation
-%\item Analysis
-%\item Localization
-%\item Alert
-%\item Telescope actuation
-%\item Total
-%\end{itemize}

### 2011-06-30 conclusions.tex
minor rewording of first sentence of conclusion
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1244 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-We have demonstrated a computationally feasible procedure for the rapid and
-even early-warning filtering of data for \GW{}s emitted during the coalescence
+We have demonstrated a computationally feasible filtering algorithm for the rapid and
+even early-warning detection of \GW{}s emitted during the coalescence
 of neutron stars and stellar-mass black holes.  It is one part of a complicated

### 2011-06-30 inspiral_svd.tex
Made CH's acknowledgements one compound sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1245 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \textsc{phy}-\oldstylenums{0107417}.  CH would like to thank Ilya Mandel for
-many discussions about rate estimates and the prospects of early detection.  CH
-would like to thank Patrick Brady for countless fruitful conversations about
-low latency analysis methods.  CH would like to thank John Zweizig for
+many discussions about rate estimates and the prospects of early detection,
+Patrick Brady for countless fruitful conversations about
+low latency analysis methods, and John Zweizig for
 disscussions about \LIGO\ data acquisition.  NF would like to thank Alessandra

### 2011-07-04 method.tex
Spelled out 'seconds' where it was not clear whether 's' was a symbol in an expression or a unit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1237 ccdd05d5-56c1-461a-bf98-85f6876523b4
More descriptive title for section 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1250 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
-\section{A novel near real-time algorithm}
+\section{Novel near real-time algorithm for \CBC\ detection}
 \label{sec:method}
@@
 \lloid\ is very slightly acausal with an intrinsic latency of
-$\left(\max \left( N^\shortdownarrow, N^\shortuparrow \right) / 2 f^0 \right)$~s.
+$\left(\max \left( N^\shortdownarrow, N^\shortuparrow \right) / 2 f^0 \right)$~seconds.
 

### 2011-07-04 prospects.tex
minor rewording of sentence baout sky localization and third generation detectors
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1233 ccdd05d5-56c1-461a-bf98-85f6876523b4
science infrastructure -> detector infrastructure
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1251 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We emphasize that any practical \GW\ search will include technical delays due
-to light travel time between the detectors, science infrastructure, and the
+to light travel time between the detectors, detector infrastructure, and the
 selected data analysis strategy.  Figure~\ref{fig:earlywarning} must be understood
@@
 less of a concern.  This should reduce the 90\% confidence area substantially.
-Also future \GW{} detectors, should they be built, will improve the \SNR\ and
-likewise the error box.
+Proposed third generation \GW{} detectors would improve the \SNR\ and likewise
+sky localization.

### 2011-07-05 snr_in_time.py
reversed x-axis in figure 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1263 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -99,6 +99,6 @@ pylab.grid()
 pylab.ylabel(r'detections yr$^{-1}$')
 pylab.xlabel(r'time before coalescence (s)')
 pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.2,right=0.95)
-pylab.xlim([0.01,1000])
+pylab.xlim([1000, 0.])
 pylab.ylim([.1, 1000])
 pylab.savefig(sys.argv[1])

### 2011-07-05 results.tex
Added 'publish' target
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1253 ccdd05d5-56c1-461a-bf98-85f6876523b4
Merged two paragraphs in section 5.5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1256 ccdd05d5-56c1-461a-bf98-85f6876523b4
657 chirp masses -> 657 points with chirp masses
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1264 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
-the parameter space by chirp mass.  We shall concentrate on a sub-bank with 657
+the parameter space by chirp mass.  We shall concentrate on a sub-bank with 657 points with
 chirp masses between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
@@
 \begin{figure}[h]
-	\plotone{figures/tmpltbank.pdf}
+	\plotone{f4}
 	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
@@
 \begin{minipage}[c]{0.4\textwidth}
-\includegraphics{figures/envelope}
+\includegraphics{t3}
 \end{minipage}
@@
 		\begin{center}
-			\includegraphics{figures/bw.pdf}
+			\includegraphics{f5a}
 			(a) Mismatch versus \SVD\ tolerance
@@
 		\begin{center}
-			\includegraphics{figures/bw_resample.pdf}
+			\includegraphics{f5b}
 			(b) Mismatch versus $N^\shortuparrow$
@@
 $N^\shortdownarrow$ and $N^\shortuparrow$.
-
 Even without further refinements, our implementation of the \lloid\ algorithm

### 2011-07-05 introduction.tex
In introduction, split apart a long compound sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1230 ccdd05d5-56c1-461a-bf98-85f6876523b4
cost of TD convolution stated in terms of flop/s instead of operations
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1265 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 preferred method in \GW{} searches.  The most obvious example of the latter is
-the time domain (\TD) convolution, which is latency-free.  However, its
-computational complexity is quadratic in the length of the templates, so it is
+the time domain (\TD) convolution, which is latency-free.  However, its cost in floating
+point operations per second is linear in the length of the templates, so it is
 prohibitively expensive for long templates.  \citet{shaunIIR} explored decomposing
@@
 implementation using open source signal processing software.  To validate our approach
-we present results of simulations and conclude with some remarks on what remains to
+we present results of simulations.  We conclude with some remarks on what remains to
 prepare for the advanced detector era.

### 2011-07-05 introduction.tex
lloid_paper: reworked EM sources section to be more complete and specific
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1266 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
-the merger and fuel a bright electromagnetic (\EM{}) counterpart
+the merger and fuel an electromagnetic (\EM{}) counterpart
 \citep{shibata:2007}.  Effort from both the \GW{} and astronomy communities may make it
@@
 
-Compact binary coalescence (\CBC) events are thought to be a mechanism for
-short gamma-ray bursts (\GRB{}s)~\citep{Lee:2005, nakar07}.  In this
-scenario, prompt gamma-ray emission arises on the accretion timescale of a
-central compact object formed $0.1$--$1.0$~s after the merger~\citep{Janka1999}.
-Optical flashes have only been observed for a handful of long \GRB{}s
-\citep{2011CRPhy..12..255A} by telescopes with extremely rapid response or, in
-the case of \textsc{grb 080319b}, by pure serendipity, where several telescopes
-were already observing the afterglow of another \GRB{} in the same field of
-view \citep{2008Natur.455..183R}. The observed optical flashes peaked within
-tens of seconds and decayed quickly.  Short \GRB{}s, on the other hand,
-typically fade so quickly that it is difficult to catch even the tails of the
-afterglows in any band. Rapid \GW{} transient alerts could enable the
-observation of prompt optical flashes and the rise of afterglows from short
-\GRB{}s.
-
-Interestingly, roughly a quarter to half of observed short
-\GRB{}s also exhibit extended emission of $30$--$100$\,s in duration beginning
-$\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the initial
-outburst.  Potential explanations for the emission are delayed fall-back
-accretion caused by $r$-process heating \citep{Metzger2010} or the formation of
-a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}.  A
-proto-magnetar may form from accretion-induced collapse or the merger of two
-\textsc{ns}s.  An accompanying \GW{} observation would confirm the existence of
-the latter channel.
+Compact binary coalescence (\CBC{}) is a plausible progenitor for most short
+gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}. The tidally
+disrupted material falls onto the newly formed, rapidly spinning compact object
+and is accelerated in jets along the spin axis with a timescale of
+$0.1$--$1.0$~s after the merger \citep{Janka1999}, matching the short \GRB{}
+duration distribution well. Prompt \EM{} emission can arise as fast outflowing
+matter collides with slower matter ejected earlier in inner shocks. The same
+inner shocks, or potentially reverse shocks, can produce an accompanying
+optical flash \citep{Sari99}. Prompt emission is a probe into the extreme
+initial conditions of the outflow, in contrast with afterglows, which arise in
+the external shock with the local medium and are relatively insensitive to
+initial conditions. Optical flashes have been observed for a handful of long
+\GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
+or, in the case of \textsc{grb 080319b}, by pure serendipity, where several
+telescopes were already observing the afterglow of another \GRB{} in the same
+field of view \citep{2008Natur.455..183R}. The observed optical flashes peaked
+within tens of seconds and decayed quickly. For short \GRB{} parameters,
+however, the reverse shock model predicts a peak flux in radio, approximately
+20 minutes after the \GRB{}, but also a relatively faint optical flash
+\citep{nakar07}; for a once-per-year Advanced LIGO event at $130$\,Mpc, the
+radio flux will peak around $9$\,GHz with $\sim$$5$\,mJy and the optical at
+$R\sim19$\,mag.  Interestingly, roughly a quarter to half of observed short
+\GRB{}s also exhibit extended X-ray emission of $30$--$100$\,s in duration
+beginning $\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the
+initial outburst. This may be caused if the merger results in the formation of
+a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}. Rapid
+\GW{} alerts would enable joint \EM{} and \GW{} observations to confirm the
+existence of \CBC{} progenitors to short \GRB{}s and allow the early \EM{}
+observation of exceptionally nearby and thus bright events.
 
@@
 
-There were a number of sources of latency associated with \CBC{}
-analysis in S6/VSR3 \citep{HugheyGWPAW2011}, listed here.
+There were a number of sources of latency associated with the search for
+\CBC{} signals in S6/VSR3 \citep{HugheyGWPAW2011}, listed here.
 

### 2011-07-05 introduction.tex
lloid_paper: give complete list of LOOCUP telescopes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1267 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Candidates were sent for \EM{} followup to several telescopes; Swift,
-\textsc{rotse}, \textsc{tarot}, and Zadko \citep{kanner2008, HugheyGWPAW2011}
-imaged some of the most likely sky locations.
+\textsc{lofar}, \textsc{rotse}, \textsc{tarot}, \textsc{quest}, SkyMapper,
+Liverpool telescope, Pi of the Sky, Zadko, and the Palomar Transient Factory
+\citep{kanner2008, HugheyGWPAW2011} imaged some of the most likely sky
+locations.
 

### 2011-07-05 prospects.tex
Added 'publish' target
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1253 ccdd05d5-56c1-461a-bf98-85f6876523b4
infeasible -> unfeasible for compatibility with British English spelling
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1255 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: try to address Larry's comment that the prospects section launched too technically without preamble.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1268 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{figure}[h]
-\plotone{figures/snr_in_time}
+\plotone{f1}
 \caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
@@
 %
-Examining the signal to noise ratio (\SNR{}) stream for threshold crossings
-before the \GW{} signal leaves the detection band makes it possible to trade
-some \SNR{} and sky localization accuracy for pre-merger detection.
+Before the \GW{} signal leaves the detection band, we can imagine examining the
+signal to noise ratio (\SNR{}) accumulated up to that point and if it is
+already significant, release an alert immediately, trading \SNR{} and sky
+localization accuracy for pre-merger detection.
 Figure~\ref{fig:earlywarning} shows projected early detectability rates for
@@
 \begin{figure}[h]
-\plotone{figures/loc_in_time}
+\plotone{f2}
 \caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
@@
 
-It is infeasible to search hundreds of square degrees for a prompt counterpart.
+It is unfeasible to search hundreds of square degrees for a prompt counterpart.
 For reference, \textsc{lsst} should see first light around the same time as

### 2011-07-05 method.tex
Added 'publish' target
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1253 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: incorporate Larry's (minor) comments and acknowledge him
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1269 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 much more computationally efficient to use the convolution theorem and the
-\fft\ to implement fast convolution in the frequency domain, requiring only
+Fast Fourier Transform (\fft) to implement fast convolution in the frequency
+domain, requiring only
 $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample but incurring
@@
 $f_\mathrm{low}$ at time $t_\mathrm{low}$ before merger.  Usually the template
-is truncated at some prescribed time $t^0$, or equivalent frequency $f_\mathrm{hi}$,
+is truncated at some prescribed time $t^0$, or equivalent frequency $f_\mathrm{high}$,
 often chosen to correspond to the \ISCO. The beginning of the template is critically
 sampled at $2 f_\mathrm{low}$, but the end of the template is critically sampled at a
-rate of $2 f_\mathrm{hi}$. In any time interval smaller than the duration of the template,
+rate of $2 f_\mathrm{high}$. In any time interval smaller than the duration of the template,
 the bandwidth of the filters across the entire template bank may be significantly less
@@
 	\begin{center}
-		\includegraphics{figures/lloid-diagram}
+		\includegraphics{f3}
 		\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
 signal flow.  Circles with arrows represent interpolation
-\protect\includegraphics{figures/upsample-symbol} or decimation
-\protect\includegraphics{figures/downsample-symbol}.  Circles with plus
+\protect\includegraphics{f3a} or decimation
+\protect\includegraphics{f3b}.  Circles with plus
 signs represent summing junctions
-\protect\includegraphics{figures/adder-symbol}.  Squares
-\protect\includegraphics{figures/fir-symbol} stand for \fir{} filters.  Sample
+\protect\includegraphics{f3c}.  Squares
+\protect\includegraphics{f3d} stand for \fir{} filters.  Sample
 rate decreases from the top of the diagram to the bottom.  In this diagram each
@@
 \frac{2 \sum_{s=0}^{S-1} f^s \numsvdtmps \slicessamps}{2 \numtmps \tmpsamps f^0} \\
-\approx \frac{\alpha}{\left(t_\mathrm{low} - t_\mathrm{hi}\right) \left(f^0\right)^2} \int_{t_\mathrm{low}}^{t_\mathrm{hi}} \left(2 f(t) \right)^2 \, \mathrm{d} t \\
-= \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}) - t_\mathrm{hi} f^2 (t_\mathrm{hi}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{hi}\right)}
+\approx \frac{\alpha}{\left(t_\mathrm{low} - t_\mathrm{high}\right) \left(f^0\right)^2} \int_{t_\mathrm{low}}^{t_\mathrm{high}} \left(2 f(t) \right)^2 \, \mathrm{d} t \\
+= \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}) - t_\mathrm{high} f^2 (t_\mathrm{high}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{high}\right)}
 \end{multline}
@@
 The integral is evaluated using the power-law form of $f(t)$ from equation~(\ref{eq:fgw}).
-Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{hi} = 1/1570$~s $f_\mathrm{hi} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the conventional \TD\ method.
+Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{high} = 1/1570$~s $f_\mathrm{high} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the conventional \TD\ method.

### 2011-07-05 introduction.tex
lloid_paper: make it clear that CBCs are not the only choice for short GRB progenitors
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1271 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Compact binary coalescence (\CBC{}) is a plausible progenitor for most short
-gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}. The tidally
+gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}, but the
+association is not iron-clad \citep(2011ApJ...727..109V}. The tidally
 disrupted material falls onto the newly formed, rapidly spinning compact object

### 2011-07-05 introduction.tex
lloid_paper: oops, forgot bibtex entry
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1272 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-07-05 introduction.tex
@@
 Compact binary coalescence (\CBC{}) is a plausible progenitor for most short
-gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}. The tidally
+gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}, but the
+association is not iron-clad \citep{2011ApJ...727..109V}. The tidally
 disrupted material falls onto the newly formed, rapidly spinning compact object

### 2011-07-07 conclusions.tex
vast gains -> substantial gains
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1257 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: incorporated Leo's comments into EM sections
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1273 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\CBC\ events may be the progenitors of some short hard \GRB{}s and are expected
+to be accompanied by a broad spectrum of \EM{} signals. Rapid alerts to the
+wider astronomical community will improve the chances of detecting an \EM{}
+counterpart in bands from gamma-rays down to radio. In the Advanced LIGO
+era, it appears possible to usefully localize a few rare events prior to the
+\GRB{}, allowing multi-wavelength observations of prompt emission. More
+frequently, low-latency alerts will be released after merger but will still
+yield extended X-ray tails and early afterglows.
+
 The \lloid{} method is as fast as standard \fft{} convolutions but allows for
@@
 \LIGO{} Data Grid.
-
-\CBC\ events may be the progenitors of some short hard \GRB{}s and are expected
-to be accompanied by prompt \EM{} signals.  Rapid alerts to the broader
-astronomical community will improve the chances of detecting an \EM{}
-counterpart in bands from gamma-rays down to radio.  It will be extremely
-challenging for all but the rarest events to localize the source well enough in
-advance to detect the onset of prompt emission, however localization prior to
-the onset of extended emission on the 10--100~s time-scale is more easily
-achieved and will benefit from the full bandwidth and \SNR\ of the \GW{} event
-at merger.  Probing even the extended emission timescale requires very low
-latency filtering schemes such as the one described here.
 
@@
 \GW{} waveform.  This is a challenging technique, but it has the potential for
-vast gains in \SNR{} and timing accuracy \citep{PhysRevD.47.2184}.
+substantial gains in \SNR{} and timing accuracy \citep{PhysRevD.47.2184}.
 

### 2011-07-07 introduction.tex
lloid_paper: incorporated Leo's comments into EM sections
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1273 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 and is accelerated in jets along the spin axis with a timescale of
-$0.1$--$1.0$~s after the merger \citep{Janka1999}, matching the short \GRB{}
-duration distribution well. Prompt \EM{} emission can arise as fast outflowing
+$0.1$--$1$~s after the merger \citep{Janka1999}, matching the short \GRB{}
+duration distribution well. Prompt \EM{} emission including the \GRB{}
+can arise as fast outflowing
 matter collides with slower matter ejected earlier in inner shocks. The same
 inner shocks, or potentially reverse shocks, can produce an accompanying
-optical flash \citep{Sari99}. Prompt emission is a probe into the extreme
+optical flash \citep{Sari99}. The prompt emission is a probe into the extreme
 initial conditions of the outflow, in contrast with afterglows, which arise in
@@
 field of view \citep{2008Natur.455..183R}. The observed optical flashes peaked
-within tens of seconds and decayed quickly. For short \GRB{} parameters,
+within tens of seconds and decayed quickly. For short \GRB{} energy balance and
+plasma density,
 however, the reverse shock model predicts a peak flux in radio, approximately
@@
 \citep{nakar07}; for a once-per-year Advanced LIGO event at $130$\,Mpc, the
-radio flux will peak around $9$\,GHz with $\sim$$5$\,mJy and the optical at
-$R\sim19$\,mag. Interestingly, roughly a quarter to half of observed short
+radio flux will peak around $9$\,GHz at $\sim$$5$\,mJy, with emission in the
+$R$-band at $\sim$$19^\mathrm{th}$ magnitude. Interestingly, roughly a quarter to half of observed short
 \GRB{}s also exhibit extended X-ray emission of $30$--$100$\,s in duration
 beginning $\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the
-initial outburst. This may be caused if the merger results in the formation of
-a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}. Rapid
-\GW{} alerts would enable joint \EM{} and \GW{} observations to confirm the
-existence of \CBC{} progenitors to short \GRB{}s and allow the early \EM{}
+initial outburst. This can be explained if the merger results in the formation
+of a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}. Rapid
+\GW{} alerts would enable joint \EM{} and \GW{} observations to confirm
+the short \GRB{}-\CBC{} link and allow the early \EM{}
 observation of exceptionally nearby and thus bright events.
@@
 \textsc{lofar}, \textsc{rotse}, \textsc{tarot}, \textsc{quest}, SkyMapper,
-Liverpool telescope, Pi of the Sky, Zadko, and the Palomar Transient Factory
+Liverpool telescope, Pi of the Sky, Zadko, and Palomar Transient Factory
 \citep{kanner2008, HugheyGWPAW2011} imaged some of the most likely sky
@@
 high bandwidth but only modest latency.  Together, these introduce a
-latency of $\gtrsim 100$~ms.  These technical sources of latency could be reduced
+latency of $\gtrsim$$100$~ms.  These technical sources of latency could be reduced
 with significant engineering and capital investments, but they are minor compared

### 2011-07-07 introduction.tex
low frequency cutoff = 20 Hz for Hooper et al.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1274 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 filters as a means to reduce the computational cost of latency-free data analysis.  They
-demonstrate a factor of 50 improvement in floating point operations per second
-over the standard \TD\ method.  We will significantly improve upon that number.
+demonstrate a factor of 250 improvement in floating point operations per second
+over the standard \TD\ method while faithfully reproducing the transfer functions of the
+matched filters above 20~Hz.  We will demonstrate a significantly greater speedup.
 

### 2011-07-08 method.tex
Boldface not used on theta
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1276 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Inspiral signals are continuously parameterized by a set of intrinsic source
-parameters $\mathbf\Theta$ that determine the amplitude and phase evolution of the
+parameters $\theta$ that determine the amplitude and phase evolution of the
 \GW{} strain. For systems in which the effects of spin can be ignored, the intrinsic
 source parameters are just the component masses of the binary,
- $\mathbf\Theta = (m_1, m_2)$. For a given source, the strain observed by the
+ $\theta = (m_1, m_2)$. For a given source, the strain observed by the
  detector is a linear combination of two waveforms corresponding to the
 `$+$' and `$\times$' \GW{} polarizations.  Thus, we must design two filters
-for each $\mathbf\Theta$.
+for each $\theta$.
 
@@
 To construct a template bank, templates are chosen with
-$\numtmps/2$ discrete signal parameters $\mathbf\Theta_0,\, \mathbf\Theta_1,\, \dots,\,
-\mathbf\Theta_{\numtmps/2-1}$. These are chosen such that any possible signal
+$\numtmps/2$ discrete signal parameters $\theta_0,\, \theta_1,\, \dots,\,
+\theta_{\numtmps/2-1}$. These are chosen such that any possible signal
 will have an inner product $\geqslant$0.97 with at least one template.
@@
 
-Since waveforms with neighboring intrinsic source parameters $\mathbf\Theta$
+Since waveforms with neighboring intrinsic source parameters $\theta$
  have similar time-frequency evolution, it is possible to design computationally

### 2011-07-08 results.tex
note about seismic wall
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1275 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed error in caption for Table 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1277 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
-curve~\citep{ALIGONoise}.  This results in a grid of 98544 points, or
+curve~\citep{ALIGONoise}.  Templates are truncated at 10~Hz, where the projected
+sensitivity of Advanced \LIGO\ is interrupted by the ``seismic wall.''
+This results in a grid of 98544 points, or
 $2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
@@
 \begin{table*}
-\caption{\label{tab:time_slices} Filter design sub-bank of 657 templates.
+\caption{\label{tab:time_slices} Filter design sub-bank of 1314 templates.
 From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.

### 2011-07-11 introduction.tex
Changed 'Liverpool telescope' to 'Liverpool Telescope'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1278 ccdd05d5-56c1-461a-bf98-85f6876523b4
removed textsc
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1280 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 amplitude and frequency increasing until the system eventually merges.  If a
-neutron star (\textsc{ns}) is involved, it may become tidally disrupted near
+neutron star (\NS{}) is involved, it may become tidally disrupted near
 the merger and fuel an electromagnetic (\EM{}) counterpart
@@
 \GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
-or, in the case of \textsc{grb 080319b}, by pure serendipity, where several
+or, in the case of GRB 080319b, by pure serendipity, where several
 telescopes were already observing the afterglow of another \GRB{} in the same
@@
 were operating, several all-sky detection pipelines operated in a low-latency
-configuration to send astronomical alerts, namely \textsc{mbta}, Coherent
+configuration to send astronomical alerts, namely \mbta{}, Coherent
 WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
-\textsc{mbta} achieved the best \GW{} trigger-generation latencies of 2--5 minutes.
+\mbta{} achieved the best \GW{} trigger-generation latencies of 2--5 minutes.
 Alerts were sent with latencies of 30--60 minutes, dominated by human vetting.
 Candidates were sent for \EM{} followup to several telescopes; Swift,
-\textsc{lofar}, \textsc{rotse}, \textsc{tarot}, \textsc{quest}, SkyMapper,
-Liverpool telescope, Pi of the Sky, Zadko, and Palomar Transient Factory
+LOFAR, ROTSE, TAROT, QUEST, SkyMapper,
+Liverpool Telescope, Pi of the Sky, Zadko, and Palomar Transient Factory
 \citep{kanner2008, HugheyGWPAW2011} imaged some of the most likely sky
@@
 prohibitively expensive for long templates.  \citet{shaunIIR} explored decomposing
-\CBC\ matched filters into banks of infinite impulse response (\textsc{iir})
+\CBC\ matched filters into banks of infinite impulse response (IIR)
 filters as a means to reduce the computational cost of latency-free data analysis.  They

### 2011-07-11 prospects.tex
removed textsc
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1281 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \plotone{f1}
-\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
+\caption{\label{fig:earlywarning}Expected number of \NS--\NS{}
 sources that could be detectable by Advanced \LIGO\ a given number of seconds
@@
 Figure~\ref{fig:earlywarning} shows projected early detectability rates for
-\textsc{ns}--\textsc{ns} binaries in Advanced \LIGO\ assuming event rates
+\NS--\NS{} binaries in Advanced \LIGO\ assuming event rates
 predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
@@
 It is unfeasible to search hundreds of square degrees for a prompt counterpart.
-For reference, \textsc{lsst} should see first light around the same time as
+For reference, LSST should see first light around the same time as
 Advanced \LIGO{} and will have an unparalleled field of view of 9.6~deg$^2$
@@
 estimate from the \GW{} signal~\citep{galaxy-catalog} as was done in S6/VSR3.
-Within the expected Advanced \LIGO{} \textsc{ns}--\textsc{ns} horizon distance,
+Within the expected Advanced \LIGO{} \NS--\NS{} horizon distance,
 the number of galaxies that can produce a given signal amplitude is much larger

### 2011-07-11 introduction.tex
made spacing between numbers and units consistent
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1283 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 20 minutes after the \GRB{}, but also a relatively faint optical flash
-\citep{nakar07}; for a once-per-year Advanced LIGO event at $130$\,Mpc, the
-radio flux will peak around $9$\,GHz at $\sim$$5$\,mJy, with emission in the
+\citep{nakar07}; for a once-per-year Advanced LIGO event at $130$~Mpc, the
+radio flux will peak around 9~GHz at $\sim$$5$~mJy, with emission in the
 $R$-band at $\sim$$19^\mathrm{th}$ magnitude. Interestingly, roughly a quarter to half of observed short
-\GRB{}s also exhibit extended X-ray emission of $30$--$100$\,s in duration
-beginning $\sim$$10$\,s after the \GRB{} and carrying comparable fluence to the
+\GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in duration
+beginning $\sim$$10$~s after the \GRB{} and carrying comparable fluence to the
 initial outburst. This can be explained if the merger results in the formation

### 2011-07-11 introduction.tex
abbreviated 'magnitude' as 'mag'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1284 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 radio flux will peak around 9~GHz at $\sim$$5$~mJy, with emission in the
-$R$-band at $\sim$$19^\mathrm{th}$ magnitude. Interestingly, roughly a quarter to half of observed short
+$R$-band at $\sim$19~mag. Interestingly, roughly a quarter to half of observed short
 \GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in duration

### 2011-07-11 method.tex
"may" --> "can" x3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1285 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 rate of $2 f_\mathrm{high}$. In any time interval smaller than the duration of the template,
-the bandwidth of the filters across the entire template bank may be significantly less
+the bandwidth of the filters across the entire template bank can be significantly less
 than the full sample rate at which data is acquired.
@@
 
-The time sliced templates may then be downsampled in each interval without
+The time sliced templates can then be downsampled in each interval without
 aliasing, so we define them as
@@
 
-In the limit of many templates, $M \gg 1$, we may neglect the cost of the forward
+In the limit of many templates, $M \gg 1$, we can neglect the cost of the forward
 transform of the data and of the multiplication of the transforms.  The computational

### 2011-07-11 inspiral_svd.tex
lloid_paper: incorporate Larry's (minor) comments and acknowledge him
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1269 ccdd05d5-56c1-461a-bf98-85f6876523b4
"may" --> "might" x1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1287 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 multi-messenger astronomy.  Prompt detection alerts for the astronomical
-community may make it possible to observe the onset of electromagnetic emission
-from compact binary coalescence.  We demonstrate a computationally practical
-filtering strategy that could produce early-warning triggers before
+community might make it possible to observe the onset of electromagnetic
+emission from compact binary coalescence.  We demonstrate a computationally
+practical filtering strategy that could produce early-warning triggers before
 gravitational radiation from the final merger has arrived at the detectors.
@@
 disscussions about \LIGO\ data acquisition.  NF would like to thank Alessandra
-Corsi for illuminating discussions on astronomical motivations.  This research
+Corsi and Larry Price for illuminating discussions on astronomical motivations.
+This research
 is supported by the National Science Foundation through a Graduate Research

### 2011-07-11 introduction.tex
"may" --> "might" x3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1288 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 amplitude and frequency increasing until the system eventually merges.  If a
-neutron star (\NS{}) is involved, it may become tidally disrupted near
+neutron star (\NS{}) is involved, it might become tidally disrupted near
 the merger and fuel an electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  Effort from both the \GW{} and astronomy communities may make it
+\citep{shibata:2007}.  Effort from both the \GW{} and astronomy communities might make it
 possible to use \GW{} observations as an early warning trigger for \EM{}
@@
 emission.  Since the \GW\ signal is in principle detectable even before the tidal
-disruption, one may have the ambition of reporting \GW\ candidates not minutes
+disruption, one might have the ambition of reporting \GW\ candidates not minutes
 after the merger, but seconds before.  We explore one essential ingredient of this

### 2011-07-11 results.tex
removed textsc
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1282 ccdd05d5-56c1-461a-bf98-85f6876523b4
Made section 5.4 more specific
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1289 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 We examine the performance of the \lloid\ algorithm on a small region of
-compact binary parameter space centered on typical \textsc{ns}--\textsc{ns}
+compact binary parameter space centered on typical \NS--\NS{}
 masses.  We begin by constructing a template bank that spans component masses
@@
 search could be implemented with $\gtrsim$$10$ machines.  Template length does
-vary over the parameter space; lower-mass templates are longer and would
-require more computation to analyze while higher-mass templates are shorter and
-would require less. However, we consider the estimates based on this sub-bank
-to be a reasonable representation.
+vary over the parameter space; lower-mass templates sweep through frequency more
+slowly and would require more computations while higher-mass templates are shorter and
+would require fewer computations.  However, we consider the estimates based on this
+sub-bank to be a reasonable representation because they are not far from estimates
+based on equation~(\ref{eq:speedup}).
 

### 2011-07-11 introduction.tex
GW astronomy is also astronomy
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1290 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the merger and fuel an electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  Effort from both the \GW{} and astronomy communities might make it
+\citep{shibata:2007}.  Effort from both the \GW{} and non-\GW{} astronomical communities might make it
 possible to use \GW{} observations as an early warning trigger for \EM{}

### 2011-07-12 introduction.tex
'the GW and non-GW' -> 'the GW and the broader'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1291 ccdd05d5-56c1-461a-bf98-85f6876523b4

--- 2011-07-11 introduction.tex
@@
 the merger and fuel an electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  Effort from both the \GW{} and astronomy communities might make it
+\citep{shibata:2007}.  Effort from both the \GW{} and the broader astronomical communities might make it
 possible to use \GW{} observations as an early warning trigger for \EM{}

### 2011-07-13 inspiral_svd.tex
add accent to Mireia's name
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1293 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	Adrian Chapman\altaffilmark{3},
-	Mireia Crispin-Ortuzar\altaffilmark{4},
+	Mireia Crisp\'{i}n-Ortuzar\altaffilmark{4},
 	Nickolas Fotopoulos\altaffilmark{3},

### 2011-07-28 results.tex
Updated flop table with corrected bug in flop script
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1294 ccdd05d5-56c1-461a-bf98-85f6876523b4
Changed figures related to benchmark script (latency -> 1 s, cpu usage -> ~90%
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1296 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \tableline
-time domain & $2.4\times10^{13}$ & 0 \\
-frequency domain & $2.6\times10^8$ & $2\times10^3$ \\
-\lloid\ (theory) & $4.7\times10^8$ & $2\times10^{-3}$ \\
-\lloid\ (prototype) & ----------- & 5 \\
+time domain & $4.9\times10^{13}$ & 0 \\
+frequency domain & $5.2\times10^8$ & $2\times10^3$ \\
+\lloid\ (theory) & $4.9\times10^8$ & $2\times10^{-3}$ \\
+\lloid\ (prototype) & ----------- & 1 \\
 \tableline
@@
 Fire\texttrademark\ X4600~M2 server with eight quad-core 2.7~GHz AMD
-Opteron\texttrademark\ 8384 processors.  This test consumed 78\% of the
+Opteron\texttrademark\ 8384 processors.  This test consumed $\sim$90\% of the
 capacity of just one out of the 32 cores, maintaining a constant latency of
-5~s.
+$\sim$1~s.
 
@@
 
-The 5~s latency is probably due to buffering and synchronization.  The latency
+The 1~s latency is probably due to buffering and synchronization.  The latency
 might be reduced by carefully tuning buffer lengths at every stage in

### 2011-07-28 implementation.tex
Streamlined sentence in section 4.1 about sub-banks containing templates of comparable chirp mass
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1238 ccdd05d5-56c1-461a-bf98-85f6876523b4
Noted that we used GStreamer 0.10.33
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1297 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 are appropriately sized so that each sub-bank can be efficiently handled by a
-single computer.  The sub-banks are chosen to have comparable chirp mass, which
-groups together templates with similar time-frequency evolution.  Dividing the source
+single computer.  Each sub-bank contains templates of comparable chirp mass, and
+therefore similar time-frequency evolution.  Dividing the source
 parameter space into smaller sub-banks also reduces the offline cost of the
@@
 This latency can be made small at the cost of some additional overhead by making the buffers sufficiently small.
+
+Our prototype pipeline, tested with GStreamer version 0.10.33, consists of the following stages.
 

### 2011-08-12 conclusions.tex
Fact checking pass: updated flop/s table, corrected use of Nup and Ndown, stripped descriptions of pipeline as 'real-time'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1339 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The \lloid{} method is as fast as standard \fft{} convolutions but allows for
-nearly latency-free, real-time operation.  We anticipate requiring $\sim$100
+nearly latency-free, real-time operation.  We anticipate requiring $\sim$400
 modern multi-core computers to search for binary neutron stars using

### 2011-08-12 figures/diagram.tex
Converted diagram to eps
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1165 ccdd05d5-56c1-461a-bf98-85f6876523b4
lloid_paper: fix LLOID diagram generation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1249 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fact checking pass: updated flop/s table, corrected use of Nup and Ndown, stripped descriptions of pipeline as 'real-time'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1339 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \tikzexternalize[force remake=true]
+\tikzset{external/system call={pdflatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"; pdftops -eps "\image".pdf "\image".eps}}
 
@@
 		column 1/.style={column sep=8mm},
-		column 2/.style={column sep=15mm},
-		column 3/.style={column sep=3mm},
-		column 4/.style={column sep=8mm},
-		column 5/.style={column sep=15mm},
-		column 6/.style={column sep=8mm},
-		column 11/.style={column sep=8mm}
+		column 2/.style={column sep=2mm},
+		column 3/.style={column sep=12mm},
+		column 4/.style={column sep=3mm},
+		column 5/.style={column sep=8mm},
+		column 6/.style={column sep=15mm},
+		column 7/.style={column sep=8mm},
+		column 12/.style={column sep=8mm}
 	]{
-		& \node[yshift=0.5cm](in){}; & \node[yshift=0.5cm](intee){}; & \node[yshift=0.5cm](u){}; & \node[yshift=0.5cm](utee){}; & \node[yshift=0.5cm](v){}; & \node[yshift=0.5cm](add0){}; & \node[yshift=0.5cm](add1){}; & \node[yshift=0.5cm](add2){}; & & \node[yshift=0.5cm](add3){}; & \node[yshift=0.5cm](out){}; \\
-		& & & \node[draw] (u00) {}; & \node[junction] (u00tee) {}; & \node[adder] (v00) {}; & \node[adder] (add00) {}; & & & & & \node[signal, label=right:{$\rho_0^0$}] (out00) {}; \\
-		& & & \node[draw] (u10) {}; & \node[junction] (u10tee) {}; & \node[adder] (v10) {}; & & \node[adder] (add10) {}; & & & & \node[signal, label=right:{$\rho_1^0$}] (out10) {}; \\
-		\node[signal, label=left:{$x$}] (x) {}; & \node[junction] (in0) {}; & \node[junction] (in0tee) {}; & & & \node[adder] (v20) {}; & & & \node[adder] (add20) {}; & & & \node[signal, label=right:{$\rho_2^0$}] (out20) {}; \\
-		& & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
-		& & & \node[draw] (u20) {}; & \node[junction] (u20tee) {}; & \node[adder] (v30) {}; & & & & & \node[adder] (add30) {}; & \node[signal, label=right:{$\rho_{M-1}^0$}] (out30) {}; \\
-		& & & & & & & & \node[minimum height=6mm](delay0){}; & & & \\
-		& & & & & & \node[resample] (resample01) {\tiny$\uparrow$}; & \node[resample] (resample11) {\tiny$\uparrow$}; & \node[resample] (resample21) {\tiny$\uparrow$}; & \node{$\cdot \cdot$}; & \node[resample] (resample31) {\tiny$\uparrow$}; \\
+		& \node[yshift=0.5cm](in){}; & & \node[yshift=0.5cm](intee){}; & \node[yshift=0.5cm](u){}; & \node[yshift=0.5cm](utee){}; & \node[yshift=0.5cm](v){}; & \node[yshift=0.5cm](add0){}; & \node[yshift=0.5cm](add1){}; & \node[yshift=0.5cm](add2){}; & & \node[yshift=0.5cm](add3){}; & \node[yshift=0.5cm](out){}; \\
+		& & & & \node[draw] (u00) {}; & \node[junction] (u00tee) {}; & \node[adder] (v00) {}; & \node[adder] (add00) {}; & & & & & \node[signal, label=right:{$\rho_0^0$}] (out00) {}; \\
+		& & & & \node[draw] (u10) {}; & \node[junction] (u10tee) {}; & \node[adder] (v10) {}; & & \node[adder] (add10) {}; & & & & \node[signal, label=right:{$\rho_1^0$}] (out10) {}; \\
+		\node[signal, label=left:{$x$}] (x) {}; & \node[junction] (in0) {}; & & \node[junction] (in0tee) {}; & & & \node[adder] (v20) {}; & & & \node[adder] (add20) {}; & & & \node[signal, label=right:{$\rho_2^0$}] (out20) {}; \\
+		& & & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
+		& & & & \node[draw] (u20) {}; & \node[junction] (u20tee) {}; & \node[adder] (v30) {}; & & & & & \node[adder] (add30) {}; & \node[signal, label=right:{$\rho_{M-1}^0$}] (out30) {}; \\
+		& & & & & & & & & \node[minimum height=6mm](delay0){}; & & & \\
+		& & & & & & & \node[resample] (resample01) {\tiny$\uparrow$}; & \node[resample] (resample11) {\tiny$\uparrow$}; & \node[resample] (resample21) {\tiny$\uparrow$}; & \node{$\cdot \cdot$}; & \node[resample] (resample31) {\tiny$\uparrow$}; \\
 %
-		& & & \node[draw] (u01) {}; & \node[junction] (u01tee) {}; & \node[adder] (v01) {}; & \node[adder] (add01) {}; & & & & & \node[signal, label=right:{$\rho_0^1$}] (out01) {}; \\
-		& & & \node[draw] (u11) {}; & \node[junction] (u11tee) {}; & \node[adder] (v11) {}; & & \node[adder] (add11) {}; & & & & \node[signal, label=right:{$\rho_1^1$}] (out11) {}; \\
-		& \node[resample] (in1) {\tiny$\downarrow$}; & \node[junction] (in1tee) {}; & & & \node[adder] (v21) {}; & & & \node[adder] (add21) {}; & & & \node[signal, label=right:{$\rho_2^1$}] (out21) {}; \\
-		& & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
-		& & & \node[draw] (u21) {}; & \node[junction] (u21tee) {}; & \node[adder] (v31) {}; & & & & & \node[adder] (add31) {}; & \node[signal, label=right:{$\rho_{M-1}^1$}] (out31) {}; \\
-		& & & & & & & & \node[minimum height=6mm](delay1){}; & & & \\
-		& & & & & & \node[resample] (resample02) {\tiny$\uparrow$}; & \node[resample] (resample12) {\tiny$\uparrow$}; & \node[resample] (resample22) {\tiny$\uparrow$}; & \node{$\cdot \cdot$}; & \node[resample] (resample32) {\tiny$\uparrow$}; \\
+		& & & & \node[draw] (u01) {}; & \node[junction] (u01tee) {}; & \node[adder] (v01) {}; & \node[adder] (add01) {}; & & & & & \node[signal, label=right:{$\rho_0^1$}] (out01) {}; \\
+		& & & & \node[draw] (u11) {}; & \node[junction] (u11tee) {}; & \node[adder] (v11) {}; & & \node[adder] (add11) {}; & & & & \node[signal, label=right:{$\rho_1^1$}] (out11) {}; \\
+		& \node[junction] (in1) {}; & \node[resample] (in1downsample) {\tiny$\downarrow$}; & \node[junction] (in1tee) {}; & & & \node[adder] (v21) {}; & & & \node[adder] (add21) {}; & & & \node[signal, label=right:{$\rho_2^1$}] (out21) {}; \\
+		& & & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
+		& & & & \node[draw] (u21) {}; & \node[junction] (u21tee) {}; & \node[adder] (v31) {}; & & & & & \node[adder] (add31) {}; & \node[signal, label=right:{$\rho_{M-1}^1$}] (out31) {}; \\
+		& & & & & & & & & \node[minimum height=6mm](delay1){}; & & & \\
+		& & & & & & & \node[resample] (resample02) {\tiny$\uparrow$}; & \node[resample] (resample12) {\tiny$\uparrow$}; & \node[resample] (resample22) {\tiny$\uparrow$}; & \node{$\cdot \cdot$}; & \node[resample] (resample32) {\tiny$\uparrow$}; \\
 %
-		& \node[minimum size=0pt, text height=8pt, anchor=base] (invdots) {$\vdots$}; & & \node[minimum size=0pt, text height=0pt, anchor=base] {$\vdots$}; & \node[minimum size=0pt, text height=8pt, anchor=base] {$\vdots$}; & \node[minimum size=0pt, text height=8pt, anchor=base] {$\vdots$}; & \node[minimum height=8pt]{}; & & & & & \node[minimum size=0pt, text height=0pt, anchor=base] {$\vdots$}; \\
-		& & & & & & \node[minimum size=0pt, text height=8pt] (vdots0) {$\vdots$}; & \node[minimum size=0pt, text height=8pt] (vdots1) {$\vdots$}; & \node[minimum size=0pt, text height=8pt] (vdots2) {$\vdots$}; & & \node[minimum size=0pt, text height=8pt] (vdots3) {$\vdots$}; \\
+		& \node[minimum size=0pt, text height=8pt, anchor=base] (invdots) {$\vdots$}; & & & \node[minimum size=0pt, text height=0pt, anchor=base] {$\vdots$}; & \node[minimum size=0pt, text height=8pt, anchor=base] {$\vdots$}; & \node[minimum size=0pt, text height=8pt, anchor=base] {$\vdots$}; & \node[minimum height=8pt]{}; & & & & & \node[minimum size=0pt, text height=0pt, anchor=base] {$\vdots$}; \\
+		& & & & & & & \node[minimum size=0pt, text height=8pt] (vdots0) {$\vdots$}; & \node[minimum size=0pt, text height=8pt] (vdots1) {$\vdots$}; & \node[minimum size=0pt, text height=8pt] (vdots2) {$\vdots$}; & & \node[minimum size=0pt, text height=8pt] (vdots3) {$\vdots$}; \\
 %
-		& & & \node[draw] (u02) {}; & \node[junction] (u02tee) {}; & \node[adder] (v02) {}; & \node[junction] (add02) {}; & & & & & \node[signal, label=right:{$\rho_0^{S-1}$}] (out02) {}; \\
-		& & & \node[draw] (u12) {}; & \node[junction] (u12tee) {}; & \node[adder] (v12) {}; & & \node[junction] (add12) {}; & & & & \node[signal, label=right:{$\rho_1^{S-1}$}] (out12) {}; \\
-		& \node[junction] (in2) {}; & \node[junction] (in2tee) {}; & & & \node[adder] (v22) {}; & & & \node[junction] (add22) {}; & & & \node[signal, label=right:{$\rho_2^{S-1}$}] (out22) {}; \\
-		& & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
-		& & & \node[draw] (u22) {}; & \node[junction] (u22tee) {}; & \node[adder] (v32) {}; & & & & & \node[junction] (add32) {}; & \node[signal, label=right:{$\rho_{M-1}^{S-1}$}] (out32) {}; \\
+		& & & & \node[draw] (u02) {}; & \node[junction] (u02tee) {}; & \node[adder] (v02) {}; & \node[junction] (add02) {}; & & & & & \node[signal, label=right:{$\rho_0^{S-1}$}] (out02) {}; \\
+		& & & & \node[draw] (u12) {}; & \node[junction] (u12tee) {}; & \node[adder] (v12) {}; & & \node[junction] (add12) {}; & & & & \node[signal, label=right:{$\rho_1^{S-1}$}] (out12) {}; \\
+		& \node[] (in2) {}; & \node[resample] (in2downsample) {\tiny$\downarrow$}; & \node[junction] (in2tee) {}; & & & \node[adder] (v22) {}; & & & \node[junction] (add22) {}; & & & \node[signal, label=right:{$\rho_2^{S-1}$}] (out22) {}; \\
+		& & & & & & \node[vdots]{$\colon$}; & & & & & & \node[vdots]{$\colon$}; \\
+		& & & & \node[draw] (u22) {}; & \node[junction] (u22tee) {}; & \node[adder] (v32) {}; & & & & & \node[junction] (add32) {}; & \node[signal, label=right:{$\rho_{M-1}^{S-1}$}] (out32) {}; \\
 	};
@@
 	\draw (in0) -- (in0tee) node[midway,above] {\small 4096 Hz};
-	\draw (in1) -- (in1tee) node[midway,above] {\small 512 Hz};
-	\draw (in2) -- (in2tee) node[midway,above] {\small 32 Hz};
+	\draw[->] (in1) -- (in1downsample);
+	\draw (in1downsample) -- (in1tee) node[pos=0.4,above] {\small 512 Hz};
+	\draw (in2downsample) -- (in2tee) node[pos=0.4,above] {\small 32 Hz};
 
@@
 	\draw (in1) -- (invdots);
-	\draw[->] (invdots) -- (in2);
+	\draw[->] (invdots) |- (in2downsample);
 	\draw[->] (x) -- (in0);
 
-	\draw[decorate, decoration={brace, amplitude=4pt}] ($(in)+(-4mm,0)$) -- ($(in)+(4mm,0)$) node[midway,above=2mm,align=center] {Decimation \\ of input};
+	\draw[decorate, decoration={brace, amplitude=4pt}] ($(in)+(-4mm,0)$) -- ($(in)+(10mm,0)$) node[midway,above=2mm,align=center] {Decimation \\ of input};
 	\draw[decorate, decoration={brace, amplitude=4pt}] ($(intee)+(-2mm,0)$) -- ($(u)+(2mm,0)$) node[midway,above=2mm,align=center] {Orthogonal \\ \textsc{fir} filters};

### 2011-08-12 implementation.tex
Fact checking pass: updated flop/s table, corrected use of Nup and Ndown, stripped descriptions of pipeline as 'real-time'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1339 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-The \lloid\ algorithm is amenable to nearly latency free, real-time implementation.
-However, such a near real-time search pipeline would likely require integration directly into
+The \lloid\ algorithm is amenable to extremely low latency implementation.
+The lowest possibly latency could be achieved by integration directly into
 the data acquisition and storage system of the \LIGO\ observatories.  A slightly more 
-modest goal is to leverage existing low-latency, but not real-time, signal processing
-software in order to implement the \lloid\ algorithm.
+modest goal is to leverage existing signal processing software in order to implement
+the \lloid\ algorithm.
 
@@
 lower sample rates by decimation. Decimation involves applying an anti-aliasing
-filter to the data, and then down-sampling by deleting samples.  We use a
-192-tap \fir\ decimator provided by \gstreamer{}'s {\tt audioresample}
-element.  The detector data is provided at every power-of-two sample rate
+filter to the data, and then down-sampling by deleting samples.  We use GStreamer's
+stock {\tt audioresample} element, which provides a FIR decimation filter with
+a Kaiser-windowed sinc function kernel.
+
+The detector data is provided at every power-of-two sample rate
 required by the time slices described in \eqref{eq:time-slices}.  At this point,

### 2011-08-12 method.tex
Fact checking pass: updated flop/s table, corrected use of Nup and Ndown, stripped descriptions of pipeline as 'real-time'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1339 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as bank of
+Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as a bank of
 finite impulse response (\fir) filters, requiring $\mathcal O(\numtmps
@@
 Here we describe a method for reducing the computational cost of a \TD\ search
-for compact binary coalescence.  We give a nearly zero-latency algorithm
+for compact binary coalescence.  We give an extremely low latency algorithm
 that competes in terms of floating point operations per second with the
@@
 rank reduction demonstrated in \citep{Cannon:2010p10398} by nearly an order
-of magnitude.  In the next section we describe how we form our early-warning
+of magnitude.
+
+Because the sets of filters from each time slice form orthogonal subspaces, and the basis
+basis filters within a given time slice are mutually orthogonal, the set of all basis filters
+from all time slices forms an orthogonal basis spanning the original templates.
+
+In the next section we describe how we form our early-warning
 detection statistic using the time slice decomposition and the \SVD.
@@
 Ultimately, the latency of the entire \lloid\ algorithm is set by the decimation
-and interpolation filters because they are generally time-symmetric.  As a result,
-\lloid\ is very slightly acausal with an intrinsic latency of
-$\left(\max \left( N^\shortdownarrow, N^\shortuparrow \right) / 2 f^0 \right)$~seconds.
+and interpolation filters because they are generally time-symmetric and slightly
+acausal.
 
@@
 \slicessamps	& num. samples in decimated time slice $s$\\
-$N^\shortdownarrow$ & num. coefficients in decimation filter \\
-$N^\shortuparrow$ & num. coefficients in interpolation filter \\
+$N^\shortdownarrow$ & length of decimation filter \\
+$N^\shortuparrow$ & length of interpolation filter \\
 \tableline
@@
 
-\subsubsection{\lloid\ method}
+By adjusting the \fft\ block size, it is possible to achieve low latency with FD
+convolution, but the computational cost grows rapidly as the latency in samples
+$(D-N$) decreases.  It is easy to show that in the limit of many templates and
+long templates, $M, \lg N \gg 1$, the computational cost scales as
+%
+$$
+\left(1 + \frac{\textrm{template length}}{\textrm{latency}}\right)
+$$
+per template per $2\lg N$.
+
+\subsubsection{\label{sec:lloid-method}\lloid\ method}
 
@@
 Decimation is achieved by applying an \fir\ anti-aliasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
-Naively, an anti-aliasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.
+Naively, an anti-aliasing filter with $(f^{s-1} / f^s) N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.
 However, it is necessary to evaluate the anti-aliasing filter only for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.
-Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 N^\shortdownarrow f^s$ \flops\ is possible.
+Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1}$ \flops\ is possible.
 
 The story is similar for the interpolation filters used to match the sample rates of the partial \SNR{} streams.
-Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a low-pass filter with $N^\shortdownarrow$ coefficients.
-The low-pass filter requires $2 M N^\shortdownarrow f^{s-1}$ \flops.
-However, by taking advantage of the fact that by construction a fraction $f^{s-1}/f$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 M N^\shortuparrow f^s$ \flops.
+Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a low-pass filter with $(f^{s-1} / f^s) N^\shortdownarrow$ coefficients.
+The low-pass filter requires $2 M N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.
+However, by taking advantage of the fact that by construction a fraction $f^s / f^{s-1}$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1}$ \flops.
 
@@
 \begin{multline*}
-\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s \\ + 2\sum_{\mathclap{f \in \{f^s\}}} \left( N^\shortdownarrow + \numtmps N^\shortuparrow \right) f^s
+\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s \\ + 2\sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} \left( N^\shortdownarrow f^0 + \numtmps N^\shortuparrow f^{s-1}\right)
 \end{multline*}
-\flops.  The second sum is carried out over the set of distinct sample rates rather than over the time slices themselves.
+\flops.
+The second sum is carried out over the set of distinct sample rates (except for the base sample rate) rather than over the time slices themselves, as we have found that it is sometimes desirable to place multiple adjacent time slices at the same sample rate in order to keep the size of matrices that enter the singular value decomposition manageable.
+Here we have assumed that the decimation filters are connected in parallel, converting from the base sample rate $f^0$ to each of the time slice sample rates $f^1$, $f^2$, $\dots$, and that the interpolation filters are connected in cascade fashion with each interpolation filter stepping from the sample rate of one time slice to the next.
 
@@
 In these limits, the cost of \lloid\ is dominated by the basis filters themselves and the reconstruction, totaling $2 \sum_{s=0}^{S-1} f^s \numsvdtmps \left( \slicessamps + \numtmps \right)$ \flops.
+
+The latency of the \lloid\ method is entirely due to the acausality of the interpolation and decimation filters.  The latency is exactly
+$$
+\frac{1}{2} \left( N^\shortdownarrow / f^{S-1} + \sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} N^\shortuparrow / f^s \right) \textrm{ seconds.}
+$$
+The first term is due to the decimation filter that converts from the base sample rate $f^0$ to the lowest sample rate $f^{S-1}$ and the sum that comprises the second term is due to the latency of all of the interpolation filters connected in series.
 

### 2011-08-12 results.tex
Fact checking pass: updated flop/s table, corrected use of Nup and Ndown, stripped descriptions of pipeline as 'real-time'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1339 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The maximum possible bandwidth is determined by the length of the filter,
-$N^\shortuparrow$.  \SNR\ loss could also arise if the series rather than parallel
-connection of the decimation and interpolation filters reduces their bandwidth
+$N^\shortuparrow$.  \SNR\ loss could also arise if the combination of both the decimation
+filters and the interpolation filters reduces their bandwidth
 measurably, if the decimation and interpolation filters do not have perfectly uniform
@@
 
+The two adjustable parameters that affect performance and mismatch the most are the SVD
+tolerance and the length of the interpolation filter.  The length of the decimation filter
+affects mismatch as well, but has very little impact on performance.
+
 \paragraph{Effect of \SVD\ tolerance}
@@
 We studied how the \SVD\ tolerance affected \SNR\ loss by holding
-$N^\shortuparrow = 192$ fixed as we varied the \SVD\ tolerance from
+$N^\shortdownarrow = N^\shortuparrow = 192$ fixed as we varied the \SVD\ tolerance from
 $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The minimum, maximum, and median
@@
 
-Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$, we studied the
-impact of the length $N^\shortuparrow$ of the interpolation filter.  The mismatch as
+Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$ and the length of the
+decimation filter fixed at $N^\shortdownarrow = 192$, we studied the
+impact of the length $N^\shortuparrow$ of the interpolation filter on mismatch.  The mismatch as
 a function of $N^\shortuparrow$ is shown in Figure~\ref{fig:bw}b.  The
@@
 that a filter length of 16 is sufficient to meet our target mismatch of 0.003.
+
+Having selected an SVD tolerance of $\left(1-10^{-4}\right)$ and $N^\shortuparrow=16$, we found that
+we could reduce $N^\shortdownarrow$ to 48 without making the median mismatch higher than 0.003.
+
+During testing, we were surprised that careful design of the decimation and interpolation stages made a crucial difference in terms of computational overhead and latency.
+When we tested the pipeline with the largest interpolation filters provided by GStreamer's audioresample element, we found that rather than making a negligible contribution to the total \flops\ budget and total measured overhead they became the dominant contribution.
+The largest interpolation filters also incurred latencies of several seconds.
+We now appreciate that the efficiency and low latency of \lloid\ hinge on proper specification of $N^\shortuparrow$.
+
+There is possibly further room for optimization here.  We have considered designing our own custom decimation and interpolation filters or tuning these filters separately for each time slice.
 
@@
 sub-bank of templates compared to other methods.  We used the results of the
-previous subsections to set the \SVD\ tolerance at $\left(1-10^{-4}\right)$ and
-the interpolation filter length to 16. Table~\ref{table:flops} shows the
-computational cost in \flops\ for this sub-bank.  For the \FD\ method, an \fft\
+previous subsections to set the \SVD\ tolerance to $\left(1-10^{-4}\right)$,
+the interpolation filter length to 16, and the decimation filter length to 48.
+Table~\ref{table:flops} shows the
+computational cost in \flops\ for the sub-bank we described above.  For the \FD\ method, an \fft\
 block size of $\fftblock = 2 \tmpsamps$ is assumed, resulting in a latency of
@@
 the \FD\ method has a latency of over half of an hour, whereas \lloid\ has a
-latency of just $\sim$ms.
+latency of just over one second.
 %
@@
 frequency domain & $5.2\times10^8$ & $2\times10^3$ \\
-\lloid\ (theory) & $4.9\times10^8$ & $2\times10^{-3}$ \\
-\lloid\ (prototype) & ----------- & 1 \\
+\lloid\ (theory) & $6.6\times10^8$ & $1.2$ \\
+\lloid\ (prototype) & ($\sim$0.9 cores) & $1.5$ \\
 \tableline
@@
 
-Table~\ref{table:flops} shows that the \lloid\ method requires $\sim$$10^8$
+Table~\ref{table:flops} shows that the \lloid\ method requires $\sim$$10^9$
 \flops\ to cover a sub-bank comprising $\sim$$10^2$ out of the total $\sim$$10^5$
@@
 parameter space have similar computational scaling, an entire single detector
-search could be implemented with $\gtrsim$$10$ machines.  Template length does
+search could be implemented with $\gtrsim$$100$ machines.  Template length does
 vary over the parameter space; lower-mass templates sweep through frequency more
@@
 would require fewer computations.  However, we consider the estimates based on this
-sub-bank to be a reasonable representation because they are not far from estimates
-based on equation~(\ref{eq:speedup}).
+sub-bank to be a reasonable representation because they are within an order of
+magnitude of estimates based on the approximate speedup in
+equation~(\ref{eq:speedup}).
 
@@
 $\sim$$10^{13}$ \flops\ for this particular sub-bank, and so it would require
-$\sim$$10^6$ present-day machines to search the full parameter space.
-Presently, the \LIGO{} Data Grid%
+a total of $\gtrsim$$10^{15}$ \flops\ or $\gtrsim$$10^6$ present-day machines
+to search the full parameter space.  Presently, the \LIGO{} Data Grid%
 \footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}} consists of
-only $\sim$$10^4$ machines.
+only $\sim$$10^4$ machines, so direct \TD\ convolution is clearly impractical.
 
@@
 capacity of just one out of the 32 cores, maintaining a constant latency of
-$\sim$1~s.
+$\sim$1.5~s.
 
 The measured overhead is consistent to within an order of magnitude with the
-lower bound from the \flops\ budget.  The additional overhead is probably
+lower bound from the \flops\ budget.  Additional overhead is possibly
 dominated by thread synchronization.  A carefully optimized \gstreamer\
@@
 
-The 1~s latency is probably due to buffering and synchronization.  The latency
-might be reduced by carefully tuning buffer lengths at every stage in
-the pipeline.  The minimum theoretically possible latency is set by
-$N^\shortdownarrow$ and $N^\shortuparrow$.
-Even without further refinements, our implementation of the \lloid\ algorithm
-achieves latencies that would make other aspects of the analysis the limiting
-steps assuming that the human vetting process is automated.
+Since in this benchmark, the pipeline's input was buffered in
+0.25~s chunks, the total measured latency of 1.5~s is entirely consistent with
+1.2~s of latency being attributed to the acausality of the decimation and
+interpolation filters, as calculated in section~\ref{sec:lloid-method}.
+Our implementation of the \lloid\ algorithm has achieved latencies that would
+make other aspects of the analysis the limiting steps assuming that the human
+vetting process is automated.

### 2011-08-13 conclusions.tex
Sterilized some other uses of the phrase 'real-time'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1340 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed speedup for IIR filter paper, improved computational cost scaling section, updated some bibliography entries
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1343 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The \lloid{} method is as fast as standard \fft{} convolutions but allows for
-nearly latency-free, real-time operation.  We anticipate requiring $\sim$400
+extremely low latency operation.  We anticipate requiring $\sim$40
 modern multi-core computers to search for binary neutron stars using
-coincident \GW\ data from a multi-detector network.  In the future, additional
+coincident \GW\ data from a 4-detector network.  In the future, additional
 computational savings could be acheived by conditionally reconstructing the

### 2011-08-13 inspiral_svd.tex
fix DCC format
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1329 ccdd05d5-56c1-461a-bf98-85f6876523b4
Updated DCC version
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1341 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed speedup for IIR filter paper, improved computational cost scaling section, updated some bibliography entries
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1343 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Foundation and operates under cooperative agreement
-\textsc{phy}-\oldstylenums{0107417}.  CH would like to thank Ilya Mandel for
+PHY-0107417.  CH would like to thank Ilya Mandel for
 many discussions about rate estimates and the prospects of early detection,
@@
 
-This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{23}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v24}.
 

### 2011-08-13 introduction.tex
Fact checking pass: updated flop/s table, corrected use of Nup and Ndown, stripped descriptions of pipeline as 'real-time'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1339 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed speedup for IIR filter paper, improved computational cost scaling section, updated some bibliography entries
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1343 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 preferred method in \GW{} searches.  The most obvious example of the latter is
-the time domain (\TD) convolution, which is latency-free.  However, its cost in floating
+time domain (\TD) convolution, which is latency-free.  However, its cost in floating
 point operations per second is linear in the length of the templates, so it is
-prohibitively expensive for long templates.  \citet{shaunIIR} explored decomposing
-\CBC\ matched filters into banks of infinite impulse response (IIR)
-filters as a means to reduce the computational cost of latency-free data analysis.  They
-demonstrate a factor of 250 improvement in floating point operations per second
-over the standard \TD\ method while faithfully reproducing the transfer functions of the
-matched filters above 20~Hz.  We will demonstrate a significantly greater speedup.
+prohibitively expensive for long templates.  Several efforts are under way to
+develop low-latency \CBC\ search pipelines with tractable computing requirements.
+Examples include \mbta{}~\citep{Marion2004, Buskulic2010}, which was deployed in
+S6/VSR3, and a novel approach using networks of IIR filters that is being explored
+by \citet{shaunIIR, linqingIIR}.
 

### 2011-08-13 results.tex
N f0 -> N / f0
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1342 ccdd05d5-56c1-461a-bf98-85f6876523b4
Removed speedup for IIR filter paper, improved computational cost scaling section, updated some bibliography entries
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1343 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 block size of $\fftblock = 2 \tmpsamps$ is assumed, resulting in a latency of
-$\left(\tmpsamps f^0\right)$ seconds.  Both the \FD\ method and \lloid\ are
+$\left(\tmpsamps / f^0\right)$ seconds.  Both the \FD\ method and \lloid\ are
 five orders of magnitude faster than the conventional \TD\ method.  However,
@@
 %
-\begin{table}
-\caption{\label{table:flops}Computational cost in \flops\ of the \TD\ method, the \FD\ method, and \lloid\ for the sub-bank described in section~\ref{sec:bank-setup}.}
+\begin{table*}
+\caption{\label{table:flops}Computational cost in \flops\ and latency in seconds
+of the \TD\ method, the \FD\ method, and \lloid.  Cost is given
+for both the sub-bank described in section~\ref{sec:bank-setup} and a full
+\NS--\NS\ search.  The last column gives the approximate number of machines per
+detector required for a full advanced LIGO \NS--\NS\ search.}
 \begin{center}
-\begin{tabular}{lll}
+\begin{tabular}{lllll}
 \tableline\tableline
-method & \flops\ & latency (s) \\
+& \flops\ & & \flops\ & number of \\
+method & (sub-bank) & latency (s) & (\NS--\NS) & machines \\[0.1em]
 \tableline
-time domain & $4.9\times10^{13}$ & 0 \\
-frequency domain & $5.2\times10^8$ & $2\times10^3$ \\
-\lloid\ (theory) & $6.6\times10^8$ & $1.2$ \\
-\lloid\ (prototype) & ($\sim$0.9 cores) & $1.5$ \\
+time domain & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$10^5$ \\
+frequency domain & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$10$ \\
+\lloid\ (theory) & $6.6\times10^8$ & $1.2$ & $\phantom{5.7}\sim 10^{10}$ & $\sim$$10$ \\
+\lloid\ (prototype) & (0.9 cores) & $1.5$ & ------------ & $\sim$$10$ \\
 \tableline
@@
 \end{center}
-\end{table}
+\end{table*}
 
-\subsection{Extrapolation of the computational cost to an Advanced \LIGO{} search}
+\subsection{Extrapolation of computational cost to an Advanced \LIGO{} search}
 
-Table~\ref{table:flops} shows that the \lloid\ method requires $\sim$$10^9$
-\flops\ to cover a sub-bank comprising $\sim$$10^2$ out of the total $\sim$$10^5$
+Table~\ref{table:flops} shows that the \lloid\ method requires $6.6 \times 10^8$
+\flops\ to cover a sub-bank comprising 657 out of the total 98544
 mass pairs.  Given that modern (ca. 2011) workstations can sustain computation
-rates up to $\sim$$10^{10}$ \flops{}, and assuming that other regions of the
+rates up to $\sim$$10^{10}$~\flops{}, and assuming that other regions of the
 parameter space have similar computational scaling, an entire single detector
-search could be implemented with $\gtrsim$$100$ machines.  Template length does
-vary over the parameter space; lower-mass templates sweep through frequency more
-slowly and would require more computations while higher-mass templates are shorter and
-would require fewer computations.  However, we consider the estimates based on this
-sub-bank to be a reasonable representation because they are within an order of
-magnitude of estimates based on the approximate speedup in
-equation~(\ref{eq:speedup}).
+search for \NS--\NS\ signals in the 1--3~$M_\odot$ component mass range could be
+implemented with $\sim$$10^{11}$ \flops\, or $\sim$$10$ machines.
+
+We computed the computational cost of a full advanced \LIGO\ \NS--\NS\ search a
+second way by computing the cost per template for the conventional \TD\ method,
+and then applying the \lloid\ speedup from equation~(\ref{eq:speedup}).  This
+should be a somewhat more accurate measure because template length varies over
+the parameter space.  Lower chirp mass templates sweep through frequency more
+slowly and require more computations while higher chirp mass templates are
+shorter and require fewer computations.  Despite these subtleties, this estimate
+gave us $4.6 \times 10^9$~\flops, differing by only a factor of $\sim$2 from the
+simple scaling argument above.  For this reason, we believe that $\sim$$10$
+machines is a realistic estimate of our computing requirements.
 
@@
 $\sim$$10^{13}$ \flops\ for this particular sub-bank, and so it would require
-a total of $\gtrsim$$10^{15}$ \flops\ or $\gtrsim$$10^6$ present-day machines
+a total of $\gtrsim$$10^{15}$ \flops\ or $\gtrsim$$10^5$ present-day machines
 to search the full parameter space.  Presently, the \LIGO{} Data Grid%

### 2011-08-14 article.tex
Merged together all tex documents
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1345 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass[preprint2]{aastex}

\usepackage[colorlinks]{hyperref}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{stmaryrd} % For \shortuparrow and \shortdownarrow
\usepackage{upgreek} % For \upmu, upright mu

\newcommand{\earlywarning}{early-warning}
\newcommand{\NS}{NS}
\newcommand{\mbta}{MBTA}
\newcommand{\GW}{GW}%{{\sc gw}}
\newcommand{\EM}{EM}%{{\sc em}}
\newcommand{\GRB}{GRB}%{{\sc grb}}
\newcommand{\CBC}{CBC}%{{\sc cbc}}
\newcommand{\LIGO}{LIGO}%{{\sc ligo}}
\newcommand{\LCGT}{LCGT}%{{\sc lcgt}}
\newcommand{\GEO}{GEO600}%{{\sc geo600}}
\newcommand{\ISCO}{ISCO}%{{\sc isco}}
\newcommand{\SNR}{SNR}%{{\sc snr}}
\newcommand{\realtime}{real-time}
\newcommand{\Msun}{\ensuremath{M_{\odot}}}
\newcommand{\order}[1]{\ensuremath{\mathcal{O}[#1]}}
\newcommand{\tmpsamps}{\ensuremath{N}}
\newcommand{\numtmps}{\ensuremath{M}}
\newcommand{\numslices}{\ensuremath{S}}
% macros for number of svd basis functions
\newcommand{\SVD}{SVD}%{{\sc svd}}
\newcommand{\svdtmps}[1]{\ensuremath{L^#1}}
\newcommand{\numsvdtmps}{\svdtmps{s}}
% macros for sample points in slices
\newcommand{\slicesamps}[1]{\ensuremath{N^#1}}
\newcommand{\slicessamps}{\slicesamps{s}}
\newcommand{\fftblock}{\ensuremath{D}}
\newcommand{\resampsamps}{\ensuremath{N^\shortdownarrow,\, N^\shortuparrow}}
\newcommand{\fir}{FIR}%{{\sc fir}}
\newcommand{\fft}{FFT}%{{\sc fft}}
\newcommand{\fmax}{\ensuremath{f^0}}
\newcommand{\flops}{flop/s}
\newcommand{\gstlal}{{\tt gstlal}}
\newcommand{\gstreamer}{GStreamer}
\newcommand{\numcpus}{{600}}
\newcommand{\lloid}{LLOID}%{{\sc lloid}}
\newcommand{\TD}{TD}%{{\sc td}}
\newcommand{\FD}{FD}%{{\sc fd}}

% Macros for collapsing sizes of things
% From TUGboat, Volume 22 (2001), No. 4
% http://www.tug.org/TUGboat/tb22-4/tb72perlS.pdf
\def\clap#1{\hbox to 0pt{\hss#1\hss}}
\def\mathllap{\mathpalette\mathllapinternal}
\def\mathrlap{\mathpalette\mathrlapinternal}
\def\mathclap{\mathpalette\mathclapinternal}
\def\mathllapinternal#1#2{\llap{$\mathsurround=0pt#1{#2}$}} 
\def\mathrlapinternal#1#2{\rlap{$\mathsurround=0pt#1{#2}$}} 
\def\mathclapinternal#1#2{\clap{$\mathsurround=0pt#1{#2}$}}

\begin{document}

\title{Toward \earlywarning\ detection of gravitational waves from compact binary coalescence}

\shorttitle{Early-warning detection of \GW\ inspirals}

\author{
	Kipp Cannon\altaffilmark{1},
	Romain Cariou\altaffilmark{2},
	Adrian Chapman\altaffilmark{3},
	Mireia Crisp\'{i}n-Ortuzar\altaffilmark{4},
	Nickolas Fotopoulos\altaffilmark{3},
	Melissa Frei\altaffilmark{5},
	Chad Hanna\altaffilmark{6},
	Erin Kara\altaffilmark{7},
	Drew Keppel\altaffilmark{8,9},
	Laura Liao\altaffilmark{10},
	Stephen Privitera\altaffilmark{3},
	Antony Searle\altaffilmark{3},
	Leo Singer\altaffilmark{3}, and
	Alan Weinstein\altaffilmark{3}
}

\altaffiltext{1}{Canadian Institute for Theoretical Astrophysics, Toronto, ON, Canada}
\altaffiltext{2}{D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
\altaffiltext{3}{LIGO Laboratory, California Institute of Technology, MC 100-36, 1200 E. California Blvd., Pasadena, CA, USA; \texttt{leo.singer@ligo.org}}
\altaffiltext{4}{Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot, Spain} 
\altaffiltext{5}{University of Texas at Austin, Austin, TX, USA}
\altaffiltext{6}{Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
\altaffiltext{7}{Department of Physics and Astronomy, Barnard College, Columbia University, New York, NY, USA}
\altaffiltext{8}{Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
\altaffiltext{9}{Leibniz Universit\"{a}t Hannover, Hannover, Germany}
\altaffiltext{10}{Ryerson University, Toronto, ON, Canada}

\keywords{gamma ray burst: general --- gravitational waves --- methods: data analysis --- methods: numerical}

\begin{abstract}
Rapid detection of compact binary coalescence with a network of advanced
gravitational-wave detectors will offer a unique opportunity for
multi-messenger astronomy.  Prompt detection alerts for the astronomical
community might make it possible to observe the onset of electromagnetic
emission from compact binary coalescence.  We demonstrate a computationally
practical filtering strategy that could produce early-warning triggers before
gravitational radiation from the final merger has arrived at the detectors.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

As a compact binary system loses energy to gravitational waves (\GW{}s), its
orbital separation decays, leading to a runaway inspiral with the \GW{}
amplitude and frequency increasing until the system eventually merges.  If a
neutron star (\NS{}) is involved, it might become tidally disrupted near
the merger and fuel an electromagnetic (\EM{}) counterpart
\citep{shibata:2007}.  Effort from both the \GW{} and the broader astronomical communities might make it
possible to use \GW{} observations as an early warning trigger for \EM{}
followup. In the first generation of ground-based laser interferometers, the
\GW{} community initiated a project to send alerts when potential \GW{}
transients were observed in order to trigger followup observations by \EM{}
telescopes.  The typical latencies were 30 minutes \citep{HugheyGWPAW2011},
which was an important achievement, but too late to catch any prompt \EM{}
emission.  Since the \GW\ signal is in principle detectable even before the tidal
disruption, one might have the ambition of reporting \GW\ candidates not minutes
after the merger, but seconds before.  We explore one essential ingredient of this
problem, a computationally inexpensive low-latency filtering algorithm for detecting
inspiral signals in \GW{} data.  We also consider the prospects for advanced \GW{}
detectors and discuss other areas of work that would be required for rapid analysis.

Compact binary coalescence (\CBC{}) is a plausible progenitor for most short
gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}, but the
association is not iron-clad \citep{2011ApJ...727..109V}. The tidally
disrupted material falls onto the newly formed, rapidly spinning compact object
and is accelerated in jets along the spin axis with a timescale of
$0.1$--$1$~s after the merger \citep{Janka1999}, matching the short \GRB{}
duration distribution well. Prompt \EM{} emission including the \GRB{}
can arise as fast outflowing
matter collides with slower matter ejected earlier in inner shocks. The same
inner shocks, or potentially reverse shocks, can produce an accompanying
optical flash \citep{Sari99}. The prompt emission is a probe into the extreme
initial conditions of the outflow, in contrast with afterglows, which arise in
the external shock with the local medium and are relatively insensitive to
initial conditions. Optical flashes have been observed for a handful of long
\GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
or, in the case of GRB 080319b, by pure serendipity, where several
telescopes were already observing the afterglow of another \GRB{} in the same
field of view \citep{2008Natur.455..183R}. The observed optical flashes peaked
within tens of seconds and decayed quickly. For short \GRB{} energy balance and
plasma density,
however, the reverse shock model predicts a peak flux in radio, approximately
20 minutes after the \GRB{}, but also a relatively faint optical flash
\citep{nakar07}; for a once-per-year Advanced LIGO event at $130$~Mpc, the
radio flux will peak around 9~GHz at $\sim$$5$~mJy, with emission in the
$R$-band at $\sim$19~mag. Interestingly, roughly a quarter to half of observed short
\GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in duration
beginning $\sim$$10$~s after the \GRB{} and carrying comparable fluence to the
initial outburst. This can be explained if the merger results in the formation
of a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}. Rapid
\GW{} alerts would enable joint \EM{} and \GW{} observations to confirm
the short \GRB{}-\CBC{} link and allow the early \EM{}
observation of exceptionally nearby and thus bright events.

In October 2010, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its
sixth science run (S6) and Virgo\footnote{\url{http://www.ego-gw.it/}}
completed its third science run (VSR3).  While both \LIGO{} detectors and Virgo
were operating, several all-sky detection pipelines operated in a low-latency
configuration to send astronomical alerts, namely \mbta{}, Coherent
WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
\mbta{} achieved the best \GW{} trigger-generation latencies of 2--5 minutes.
Alerts were sent with latencies of 30--60 minutes, dominated by human vetting.
Candidates were sent for \EM{} followup to several telescopes; Swift,
LOFAR, ROTSE, TAROT, QUEST, SkyMapper,
Liverpool Telescope, Pi of the Sky, Zadko, and Palomar Transient Factory
\citep{kanner2008, HugheyGWPAW2011} imaged some of the most likely sky
locations.

There were a number of sources of latency associated with the search for
\CBC{} signals in S6/VSR3 \citep{HugheyGWPAW2011}, listed here.

\paragraph{Data acquisition and aggregation ($\gtrsim$100~ms)}%
The \LIGO\ data acquisition system collects data from detector subsystems 16
times a second~\citep{Bork2001}. Data are also copied from all of the \GW\
observatories to the analysis clusters over the Internet, which is capable of
high bandwidth but only modest latency.  Together, these introduce a
latency of $\gtrsim$$100$~ms.  These technical sources of latency could be reduced
with significant engineering and capital investments, but they are minor compared
to any of the other sources of latency.

\paragraph{Data conditioning ($\sim$1~min)}%
Science data must be calibrated using the detector's frequency
response to gravitational radiation.  Currently, data are calibrated in blocks of
16~s.  Within $\sim$1~min, data quality is assessed in order to create veto flags.
These are both technical sources of latency that might be addressed with improved
calibration and data quality software for advanced detectors.

\paragraph{Trigger generation (2--5~min)}%
Low-latency data analysis pipelines
deployed in S6/VSR3 achieved an impressive latency of minutes.  However, second
to the human vetting process, this dominated the latency of the entire \EM\
followup process.  Even if no other sources of latency existed, this trigger
generation latency is too long to catch prompt or even extended emission.
Low-latency trigger generation will become more challenging with advanced detectors
because inspiral signals will stay in band up to ten times longer.  In this work,
we will focus on reducing this source of latency.

\paragraph{Alert generation (2--3~min)}%
S6/VSR3 saw the introduction of low-latency astronomical
alerts, which required gathering event parameters and sky localization from the
various online analyses, downselecting the events, and calculating telescope pointings.
If other sources of latency improve, the technical latency associated with this
infrastructure could dominate, so work should be done to improve it.

\paragraph{Human validation (10--20~min)}%
Because the new alert system was commissioned during S6/VSR3, all alerts were subjected
to quality control checks by human operators before they were disseminated.
This was by far the largest source of latency during S6/VSR3.
Hopefully, confidence in the system will grow to the point where no human intervention
is necessary before alerts are sent, so we give it no further consideration here.

\paragraph{}

This work will focus on reducing the latency of trigger production.  Data analysis
strategies for advance detection of \CBC{}s will have to strike a balance between latency
and throughput. \CBC{} searches consist of banks of
matched filters, or cross-correlations between the data stream and a bank of
nominal ``template'' signals.  There are many different implementations of
matched filters, but most have high throughput at the cost of high latency, or
low latency at the cost of low throughput.  The former are epitomized by the
overlap-save algorithm for frequency domain (\FD) convolution, currently the
preferred method in \GW{} searches.  The most obvious example of the latter is
time domain (\TD) convolution, which is latency-free.  However, its cost in floating
point operations per second is linear in the length of the templates, so it is
prohibitively expensive for long templates.  Several efforts are under way to
develop low-latency \CBC\ search pipelines with tractable computing requirements.
Examples include \mbta{}~\citep{Marion2004, Buskulic2010}, which was deployed in
S6/VSR3, and a novel approach using networks of IIR filters that is being explored
by \citet{shaunIIR, linqingIIR}.

Fortunately, the morphology of inspiral signals can be exploited to offset some
of the computational complexity of low-latency algorithms.  First, the signals
evolve slowly in frequency, so that they can be broken into contiguous
band-limited time intervals and processed at possibly lower sample rates.
Second, inspiral filter banks consist of highly similar templates, admitting
methods such as singular value decomposition (\SVD{}) to reduce the number of
templates \citep{Cannon:2010p10398}. We will use both aspects to demonstrate
that a very low latency detection statistic is possible with current computing
resources.  Assuming the other technical sources of latency can be reduced
significantly, this could make it possible to send prompt alerts to the
astronomical community.

The paper is organized as follows.  First, we discuss prospects for early-warning
detection.  Then, we provide an overview of our novel method for detecting compact
binary coalescence signals with extremely low latency. We then describe a prototype
implementation using open source signal processing software.  To validate our approach
we present results of simulations.  We conclude with some remarks on what remains to
prepare for the advanced detector era.

\section{Prospects for \earlywarning\ detection and \EM\ followup}

\begin{figure}[h]
\plotone{f1}
\caption{\label{fig:earlywarning}Expected number of \NS--\NS{}
sources that could be detectable by Advanced \LIGO\ a given number of seconds
before coalescence.  The heavy solid line is the most realistic yearly rate
estimate.  The shaded region represents the 5\% to 95\% confidence interval
arising from uncertainty in predicted event rates \citep{Abadie:2010p10836}.}
\end{figure}
%
Before the \GW{} signal leaves the detection band, we can imagine examining the
signal to noise ratio (\SNR{}) accumulated up to that point and if it is
already significant, release an alert immediately, trading \SNR{} and sky
localization accuracy for pre-merger detection.
Figure~\ref{fig:earlywarning} shows projected early detectability rates for
\NS--\NS{} binaries in Advanced \LIGO\ assuming event rates
predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
observe a total of 40~events~yr$^{-1}$. $\sim$10~yr$^{-1}$ will be detectable
within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable within 25~s of
merger if analysis can proceed with near zero latency. These rates are shown
with their substantial uncertainties in Figure~\ref{fig:earlywarning}.

We emphasize that any practical \GW\ search will include technical delays due
to light travel time between the detectors, detector infrastructure, and the
selected data analysis strategy.  Figure~\ref{fig:earlywarning} must be understood
in the context of all of the potential sources of latency, some of which are avoidable
and some of which are not.

\begin{table}[h]
\caption{\label{table:sky-localization-accuracy}Horizon distance, \SNR\ at
merger, and area of 90\% confidence at selected times before merger for sources
with expected detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
\begin{center}
\begin{tabular}{rrrrrrr}
\tableline\tableline
rate & horiz. & final & \multicolumn{4}{c}{$A$(90\%) (deg$^2$)} \\
\cline{4-7}
yr$^{-1}$ & (Mpc) & \SNR\ & 25 s & 10 s & 1 s & 0 s \\
\tableline
40\phd\phn & 445 & 8.0 & 15500 & 3100 & 200 & 9.6 \\
10\phd\phn & 280 & 12.7 & 6100 & 1200 & 78 & 3.8 \\
1\phd\phn & 130 & 27.4 & 1300 & 260 & 17 & 0.8 \\
0.1 & 60 & 58.9 & 280 & 56 & 3.6 & 0.2 \\
\tableline
\end{tabular}
\end{center}
\end{table}
%
\begin{figure}[h]
\plotone{f2}
\caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
region as a function of time before coalescence for sources with anticipated
detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$. The heavy dot indicates
the time at which the accumulated \SNR\ exceeds a threshold of~8.}
\end{figure}

\EM\ followup requires estimating the location of the \GW{} source. The localization
uncertainty can be estimated from the uncertainty in the time of arrival of the \GW{}s,
which is determined by the signal's effective bandwidth and \SNR{}
\citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
Figure~\ref{fig:sky-localization-accuracy} show the estimated minimum 90\%
confidence area versus time of the loudest coalescence events detectable by
Advanced \LIGO{} and Advanced Virgo.  Once per year, we expect to observe an
event with a final \SNR\ of $\approx$27 whose location can be constrained to about
1300~deg$^2$ (3.1\% of the sky) within 25~s of merger,
260~deg$^2$ (0.63\% of the sky) within 10~s of merger, and
0.82~deg$^2$ (0.0020\% of the sky) at merger.

It is unfeasible to search hundreds of square degrees for a prompt counterpart.
For reference, LSST should see first light around the same time as
Advanced \LIGO{} and will have an unparalleled field of view of 9.6~deg$^2$
\citep{2008arXiv0805.2366I}.  However, 
it is possible to reduce the localization uncertainty by only looking at
galaxies from a catalog that lie near the sky location and luminosity distance
estimate from the \GW{} signal~\citep{galaxy-catalog} as was done in S6/VSR3.
Within the expected Advanced \LIGO{} \NS--\NS{} horizon distance,
the number of galaxies that can produce a given signal amplitude is much larger
than in Initial \LIGO{} and thus the catalog will not be as useful
for downselecting pointings for most events. However, exceptional \GW{} sources will
necessarily be extremely nearby. Within this reduced volume there will be fewer
galaxies to consider for a given candidate and catalog completeness will be
less of a concern.  This should reduce the 90\% confidence area substantially.
Proposed third generation \GW{} detectors would improve the \SNR\ and likewise
sky localization.

\section{Novel low latency algorithm for \CBC\ detection}
\label{sec:method}

In this section we describe a decomposition of the \CBC{} signal space that
reduces \TD\ filtering cost sufficiently to allow for the
possibility of \earlywarning\ detection with modest computing requirements.  We
expand on the ideas of \citet{Marion2004} and \citet{Buskulic2010} that describe a
multi-band decomposition of the compact binary signal space that resulted in
a search with minutes latency during S6/VSR3~\citep{HugheyGWPAW2011}.  We combine this
with the \SVD\ rank-reduction method of \citet{Cannon:2010p10398} that exploits
the redundancy of the template banks.

\subsection{Conventional \CBC{} searches}

Inspiral signals are continuously parameterized by a set of intrinsic source
parameters $\theta$ that determine the amplitude and phase evolution of the
\GW{} strain. For systems in which the effects of spin can be ignored, the intrinsic
source parameters are just the component masses of the binary,
 $\theta = (m_1, m_2)$. For a given source, the strain observed by the
 detector is a linear combination of two waveforms corresponding to the
`$+$' and `$\times$' \GW{} polarizations.  Thus, we must design two filters
for each $\theta$.

Searches for inspiral signals typically employ matched filter
banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.
The coefficients for the $\numtmps$ filters are known as templates, 
and are formed by discretizing and time reversing the
waveforms and weighting them by the inverse amplitude spectral density of the
detector's noise.
To construct a template bank, templates are chosen with
$\numtmps/2$ discrete signal parameters $\theta_0,\, \theta_1,\, \dots,\,
\theta_{\numtmps/2-1}$. These are chosen such that any possible signal
will have an inner product $\geqslant$0.97 with at least one template.
Such a template bank is said to have a {\em minimal match} of 0.97~\citep{Owen:1998dk}.

Filtering the detector data involves a convolution of the data with the
templates.  For a unit-normalized template $h_i[k]$ and whitened detector data
$x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the
\SNR{}, $\rho_i[k]$ defined as
%
% Filtering equations
%
\begin{equation}
	\label{eq:SNRTD}
	\rho_i [k] = \sum_{n=0}^{N-1} h_{i}[n] x [k-n].
\end{equation}
This results in $\numtmps$ \SNR\ time series. Local peak-finding across time and
template indices results in single-detector triggers.  Coincidences are sought
between triggers in different \GW\ detectors in order to form detection candidates.

Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as a bank of
finite impulse response (\fir) filters, requiring $\mathcal O(\numtmps
\tmpsamps)$ floating point operations per sample.  However, it is typically
much more computationally efficient to use the convolution theorem and the
Fast Fourier Transform (\fft) to implement fast convolution in the frequency
domain, requiring only
$\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample but incurring
a latency of $\mathcal O(\tmpsamps)$ samples.

\subsection{The \lloid\ method}

Here we describe a method for reducing the computational cost of a \TD\ search
for compact binary coalescence.  We give an extremely low latency algorithm
that competes in terms of floating point operations per second with the
conventional \FD\ method, which by contrast requires a significant latency due to the
inherent acausality of the Fourier transform.  Our method, called \lloid{}
(Low Latency Online Inspiral Detection),
involves two transformations of the templates that produce a
network of orthogonal filters that is far more computationally
efficient than the original bank of matched filters.

The first transformation is to chop the templates into disjointly supported
intervals, or \emph{time slices}.  Since the time slices of a given template
are disjoint in time, they are orthogonal with respect to time.  Given the
chirp-like structure of the templates, the ``early'' (lowest frequency) time
slices have significantly lower bandwidth and can be safely downsampled.
Downsampling reduces the total number of filter coefficients by a factor of
$\sim$100 by treating the earliest part of the waveform at $\sim$$1/100$ of
the full sample rate.  Together, the factor of 100 reduction in the number of
filter coefficients and the factor of 100 reduction in the sample rate save a
factor of $\sim$$10^4$ floating point operations per second (\flops) over the
original (full sample rate) templates.

However, the resulting filters are still not
orthogonal across the parameter space, and are in fact highly redundant.
We use the \SVD{} to approximate the template bank by a set of orthogonal
\emph{basis filters}~\citep{Cannon:2010p10398}.  We find that this approximation
reduces the number of filters needed by another factor of $\sim$100.  These two
transformations combined reduce the number of floating point operations
to the level that is competitive with the conventional high-latency \FD\
matched filter approach.  In the remainder of this section we describe the
\lloid\ algorithm in detail and provide some basic computational cost scaling.

\subsubsection{Selectively reducing the sample rate of the data and templates}
\label{sec:time-slices}

The first step of our proposed method is to divide the templates into time
slices in a \TD{} analogue to the \FD{} decomposition described
in \citet{Marion2004} and \citet{Buskulic2010}.  We decompose each template
$h_{i}[k]$ into a sum of $S$ non-overlapping templates
%
\begin{equation}
\label{eq:time-slices}
h_{i}[k] = \sum_{s=0}^{S-1}
	\begin{cases}
		h_i^s[k] & \textrm{if } t^s \leqslant k / f^0 < t^{s+1} \\
		0 & \textrm{otherwise}
	\end{cases}
\end{equation}
%
for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0
t^S = N$.  The outputs of these new time-sliced filters
form an ensemble of partial \SNR{} streams.  By linearity of the filtering
process, these partial \SNR\ streams can be summed to reproduce the
\SNR\ of the full template.

Since waveforms with neighboring intrinsic source parameters $\theta$
 have similar time-frequency evolution, it is possible to design computationally
efficient time slices for an extended region of parameter space rather than to
design different time slices for each template.

For concreteness and simplicity, consider an inspiral waveform in the
quadrapole approximation, for which the time-frequency relation~\citep{kidder1992, findchirppaper} is
%
\begin{equation} \label{eq:fgw}
%
f(t) = \frac{1}{\pi \mathcal{M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
\right]^{3/8}.
%
\end{equation}
%
Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
M_\odot / c^3 \approx 5$~$\upmu$s) and $t$ is the time relative to the
coalescence of the binary.
This monotonic time-frequency relationship allows us
to choose time slice boundaries that require substantially less bandwidth at
early times in the inspiral.

An inspiral signal will enter the detection band with some low frequency
$f_\mathrm{low}$ at time $t_\mathrm{low}$ before merger.  Usually the template
is truncated at some prescribed time $t^0$, or equivalent frequency $f_\mathrm{high}$,
often chosen to correspond to the \ISCO. The beginning of the template is critically
sampled at $2 f_\mathrm{low}$, but the end of the template is critically sampled at a
rate of $2 f_\mathrm{high}$. In any time interval smaller than the duration of the template,
the bandwidth of the filters across the entire template bank can be significantly less
than the full sample rate at which data is acquired.

Our goal is to reduce the filtering cost of a
large fraction of the waveform by computing part of the convolution at a lower
sample rate.  Specifically we consider here time slice boundaries with the
smallest power-of-two sample rates that sub-critically sample the time-sliced
templates.  The time slices consist of the $S$ intervals
$\left[t^0, t^1\right),\, \left[t^1, t^2\right),\, \dots,\, \left[t^{S-1}, t^S\right)$,
sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^s$ is at
least twice the highest nonzero frequency component of any filter in the bank for the
$s$th time slice.

The time sliced templates can then be downsampled in each interval without
aliasing, so we define them as
%
\begin{equation}
\label{eq:time-sliced-templates}
h_{i}^{s}[k] \equiv
	\begin{cases}
		h_{i}\!\left[k\frac{f}{f^s}\right] & \textrm{if } t^s \leqslant k/f^s < t^{s+1} \\
		0 & \textrm{otherwise.}
	\end{cases}
\end{equation}
%
We note that the time slice decomposition in equation~(\ref{eq:time-slices}) is
manifestly orthogonal since the time slices are disjoint in time.  In the next
section we examine how to reduce the number of filters within each time slice
via \SVD{} of the time-sliced templates.

\subsubsection{Reducing the number of filters with the \SVD{}}
\label{sec:svd}

As noted previously, the template banks used in inspiral searches are by design
highly correlated.  \citet{Cannon:2010p10398} showed that applying the \SVD\
to inspiral template banks greatly reduces the number of filters required to achieve a
particular minimal match.  A similar technique can be applied to the time-sliced
templates as defined in equation~\ref{eq:time-sliced-templates} above.  The \SVD\
is a matrix factorization that takes the form
%
\begin{equation}
h_i^s[k] = \sum_{\mathclap{l=0}}^{\mathclap{M-1}} v_{il}^s \sigma_l^s u_l^s[k] \approx \sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s u_l^s[k].
\label{eq:svddecomp}
\end{equation}
where $u_l^s[k]$ are orthonormal \emph{basis templates} related to the original
time-sliced templates through the \emph{reconstruction matrix}, $v_{il}^s\sigma_l^s$.
The expectation value of the fractional loss in \SNR\ is the \SVD\ tolerance, given by
%
\begin{equation*}
\left[ \sum_{l=0}^{L^s-1} \left( \sigma_l^s \right)^2 \right]\left[ \sum_{l=0}^{M-1} \left( \sigma_l^s \right)^2 \right]^{-1},
\end{equation*}
%
determined by the number $\numsvdtmps$ of basis templates that are kept in
the approximation.  The authors of \citet{Cannon:2010p10398}
showed that highly accurate approximations of inspiral template banks could be
achieved with few basis templates.  We find that when combined with the
time slice decomposition, the number of basis templates \numsvdtmps\ is much
smaller than the original number of templates \numtmps\ and improves on the
rank reduction demonstrated in \citep{Cannon:2010p10398} by nearly an order
of magnitude.

Because the sets of filters from each time slice form orthogonal subspaces, and the basis
basis filters within a given time slice are mutually orthogonal, the set of all basis filters
from all time slices forms an orthogonal basis spanning the original templates.

In the next section we describe how we form our early-warning
detection statistic using the time slice decomposition and the \SVD.

\subsubsection{Early-warning output}

In the previous two sections we have described two transformations that greatly
reduce the computational burden of \TD\ filtering.  We are now prepared to define
our detection statistic, the early-warning output, and to comment on the computational
cost of evaluating it.

First, the sample rate of the detector data must be decimated to match sample rates
with each of the time slices.  We will denote the decimated detector data streams using
a superscript $^s$ to indicate the time slices to which they correspond, and the
operator $H^\shortdownarrow$ will represent the appropriate decimation filter:
\begin{equation*}
\label{eq:decomp}
	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k].
\end{equation*}
We shall use the symbol $H^\shortuparrow$ to represent an interpolation filter,
\begin{equation*}
	x^{s}[k] = \left( H^\shortuparrow x^{s+1}\right)[k].
\end{equation*}
Ultimately, the latency of the entire \lloid\ algorithm is set by the decimation
and interpolation filters because they are generally time-symmetric and slightly
acausal.

From the combination of the time slice decomposition in
equation~(\ref{eq:time-sliced-templates}) and the \SVD\ defined in
equation~(\ref{eq:svddecomp}), we define the early-warning output accumulated
up to time slice $s$ using the recursion relation,
%
% orthogonal decomposition filtering
%
\begin{multline}
	\rho_i^s [k] =%
		% Interpolation SNR
		\overbrace{
			\left(H^\uparrow \rho_i^{s+1}\right)[k]
		}^\textrm{\clap{{\sc snr} from previous time slices}} \\
		% Plus ...
		+
		% Reconstruction
		\underbrace{
			\sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s
		}_\textrm{\clap{reconstruction}}
		% Orthogonal FIR filter
		\overbrace{
			\sum_{\mathclap{n=0}}^{\mathclap{N^s-1}} u_l^s[n] x^s[k-n]
		}^\textrm{\clap{orthogonal {\sc fir} filters}} .
\end{multline}
%
%
Observe that the early-warning output for time slice 0, $\rho_i^0[k]$,
approximates the \SNR\ of the original templates.
%
%
\begin{figure*}[h!]
	\begin{center}
		\includegraphics{f3}
		\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
signal flow.  Circles with arrows represent interpolation
\protect\includegraphics{f3a} or decimation
\protect\includegraphics{f3b}.  Circles with plus
signs represent summing junctions
\protect\includegraphics{f3c}.  Squares
\protect\includegraphics{f3d} stand for \fir{} filters.  Sample
rate decreases from the top of the diagram to the bottom.  In this diagram each
time slice contains three \fir\ filters that are linearly combined to produce
four output channels.  In a typical pipeline the number of \fir\ filters is
much less than the number of output channels.}
	\end{center}
\end{figure*}
%
%
The signal flow diagram in Figure~\ref{fig:pipeline} illustrates this
recursion relation as a multirate filter network with a number of early-warning outputs.

In the next section we compute the expected computational cost scaling of this
decomposition and compare it with the brute-force \TD{} implementation of
\eqref{eq:SNRTD} and higher latency \FD{} methods.

\subsection{Comparison of computational costs}

We now examine the computational cost scaling of the conventional \TD{} or
\FD{} matched filter procedure as compared with \lloid{}.  For convenience,
Table~\ref{tab:recap} provides a review of the notation that we will need in
this section.
%
% symbols used in FLOPs calculations table
%
\begin{table}
\caption{\label{tab:recap}Notation used to describe filters.}
\begin{center}
\begin{tabular}{ll}
\tableline\tableline
& Definition \\
\tableline
$f^s$		& sample rate in time slice $s$ \\
\numtmps		& num. templates \\
\tmpsamps	& num. samples per template \\
\numslices	& num. time slices \\
\numsvdtmps	& num. basis templates in time slice $s$ \\
\slicessamps	& num. samples in decimated time slice $s$\\
$N^\shortdownarrow$ & length of decimation filter \\
$N^\shortuparrow$ & length of interpolation filter \\
\tableline
\end{tabular}
\end{center}
\end{table}

\subsubsection{Conventional \TD{} method}

The conventional \TD{} method consists of a bank of \fir{} filters, or sliding-window dot products.  If there are $\numtmps$ templates, each $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications and additions per sample, or $2 \numtmps \tmpsamps f^0$ \flops\ at a sample rate $f^0$.

\subsubsection{Conventional \FD{} method}

The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in
\citet{numerical-recipes-chapter-13}.  It entails splitting the input into blocks of $D$
samples, $D > \tmpsamps$, each block overlapping the previous one by $D - \tmpsamps$
samples.  For each block, the algorithm computes the forward \fft\ of the data and
each of the templates, multiplies them, and then computes the reverse \fft.

Modern implementations of the \fft, such as the ubiquitous \texttt{fftw}, require about
$2 \fftblock \lg \fftblock$ operations to evaluate a real transform of size
$\fftblock$~\citep{Johnson:2007p9654}.  Including the forward transform of the data and
$M$ reverse transforms for each of the templates, the \fft\ costs $2 (\numtmps + 1)
\fftblock \lg \fftblock$ operations per block.  The multiplication of the transforms adds
a further $2 \numtmps \fftblock$ operations per block.  Since each block produces
$\fftblock - \tmpsamps$ usable samples of output, the overlap-save method requires
$$
f^0 \cdot \frac{2 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock} \; \mathrm{\flops} \,.
$$

In the limit of many templates, $M \gg 1$, we can neglect the cost of the forward
transform of the data and of the multiplication of the transforms.  The computational
cost will reach an optimum at some large but finite \fft\ block size
$\fftblock \gg \tmpsamps$.  In this limit, the \FD\ method costs
$\approx 2 f^0 \numtmps \lg \fftblock$ \flops.

By adjusting the \fft\ block size, it is possible to achieve low latency with FD
convolution, but the computational cost grows rapidly as the latency in samples
$(D-N$) decreases.  It is easy to show that in the limit of many templates and
long templates, $M, \lg N \gg 1$, the computational cost scales as
%
$$
\left(1 + \frac{\textrm{template length}}{\textrm{latency}}\right)
$$
per template per $2\lg N$.

\subsubsection{\label{sec:lloid-method}\lloid\ method}

For time slice $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ \flops\ 
to evaluate the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ to apply the 
linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops\ to add the resultant partial \SNR\ stream.

The computational cost of the decimation of the detector data is a little bit more subtle.  Decimation is achieved by applying an \fir\ anti-aliasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.  Naively, an anti-aliasing filter with $(f^{s-1} / f^s) N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.  However, it is necessary to evaluate the anti-aliasing filter only for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.  Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1}$ \flops\ is possible.

The story is similar for the interpolation filters used to match the sample rates of the partial \SNR{} streams.  Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a low-pass filter with $(f^{s-1} / f^s) N^\shortdownarrow$ coefficients.  The low-pass filter requires $2 M N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.  However, by taking advantage of the fact that by construction a fraction $f^s / f^{s-1}$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1}$ \flops.

Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time-sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total the \lloid\ algorithm requires
\begin{multline*}
\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s \\ + 2\sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} \left( N^\shortdownarrow f^0 + \numtmps N^\shortuparrow f^{s-1}\right)
\end{multline*}
\flops.  The second sum is carried out over the set of distinct sample rates (except for the base sample rate) rather than over the time slices themselves, as we have found that it is sometimes desirable to place multiple adjacent time slices at the same sample rate in order to keep the size of matrices that enter the singular value decomposition manageable.  Here we have assumed that the decimation filters are connected in parallel, converting from the base sample rate $f^0$ to each of the time slice sample rates $f^1$, $f^2$, $\dots$, and that the interpolation filters are connected in cascade fashion with each interpolation filter stepping from the sample rate of one time slice to the next.

We can simplify this expression quite a bit by taking some limits that arise from sensible filter design.  In the limit of many templates, the cost of the decimation filters is negligible as compared to the cost of the interpolation filters.  Typically, we will design the interpolation filters with $N^\shortuparrow \lesssim \numsvdtmps$ so that the interpolation cost itself is negligible compared with the reconstruction cost.  Finally, if the number of basis templates per time slices $\numsvdtmps$ is not too small, the reconstruction cost dominates over the cost of accumulating the partial \SNR.  In these limits, the cost of \lloid\ is dominated by the basis filters themselves and the reconstruction, totaling $2 \sum_{s=0}^{S-1} f^s \numsvdtmps \left( \slicessamps + \numtmps \right)$ \flops.

The latency of the \lloid\ method is entirely due to the acausality of the interpolation and decimation filters.  The latency is exactly
$$
\frac{1}{2} \left( N^\shortdownarrow / f^{S-1} + \sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} N^\shortuparrow / f^s \right) \textrm{ seconds.}
$$
The first term is due to the decimation filter that converts from the base sample rate $f^0$ to the lowest sample rate $f^{S-1}$ and the sum that comprises the second term is due to the latency of all of the interpolation filters connected in series.

\subsubsection{Speedup of \lloid\ relative to \TD\ method}

If the cost of the basis filters dominates, and the frequency of the templates evolves slowly enough in time, then we can use the time-frequency relationship of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional \TD\ method.  The reduction in \flops\ is approximately
%
\begin{multline}
\label{eq:speedup}
\frac{2 \sum_{s=0}^{S-1} f^s \numsvdtmps \slicessamps}{2 \numtmps \tmpsamps f^0} \\
\approx \frac{\alpha}{\left(t_\mathrm{low} - t_\mathrm{high}\right) \left(f^0\right)^2} \int_{t_\mathrm{low}}^{t_\mathrm{high}} \left(2 f(t) \right)^2 \, \mathrm{d} t \\
= \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}) - t_\mathrm{high} f^2 (t_\mathrm{high}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{high}\right)}
\end{multline}
%
where $\alpha \approx \numsvdtmps / \numtmps$ is the rank reduction factor, or ratio between the number of basis templates and the number of templates.  This approximation assumes that the frequency of the signal is evolving very slowly so that we can approximate the time slice sample rate as twice the instantaneous gravitational wave frequency, $f^s \approx 2 f(t)$, and the number of samples in the decimated time slice as the sample rate times an infinitesimally short time interval, $\slicessamps \approx 2 f(t) \, \mathrm{d}t$. The integral is evaluated using the power-law form of $f(t)$ from equation~(\ref{eq:fgw}). Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{high} = 1/1570$~s $f_\mathrm{high} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the conventional \TD\ method.

\section{Implementation}
\label{sec:implementation}

In this section we describe an implementation of the \lloid\ method described
in section \ref{sec:method} suitable for rapid \GW{}
searches for \CBC{}s.  The \lloid\ method requires several
computations that can be completed before the analysis is underway.  Thus
we divide the procedure into an offline planning stage and an
online, low-latency filtering stage.  The offline stage can be done before the
analysis is started and updated asynchronously, whereas the online stage must
keep up with the detector output and produce search results as rapidly as
possible.  In the next two subsections we describe what these stages entail.

\subsection{Planning stage}

The planning stage begins with choosing templates that cover the space of
source parameters with a hexagonal grid~\citep{PhysRevD.76.102004} in order to
satisfy a minimal match criterion.  This assures a prescribed maximum loss in
\SNR\ for signals whose parameters do not lie on the hexagonal grid.  Next, the
grid is partitioned into groups of neighbors called \emph{sub-banks} that
are appropriately sized so that each sub-bank can be efficiently handled by a
single computer.  Each sub-bank contains templates of comparable chirp mass, and
therefore similar time-frequency evolution.  Dividing the source
parameter space into smaller sub-banks also reduces the offline cost of the
\SVD{} and is the approach considered in \citet{Cannon:2010p10398}.  Next, we choose
time-slice boundaries as in equation~\eqref{eq:time-sliced-templates} such that all
of the templates within a sub-bank are sub-critically sampled at progressively lower
sample rates.  For each time slice, the templates are downsampled to the
appropriate sample rate.  Finally, the \SVD\ is applied to each time slice in
the sub-bank in order to produce a set of orthonormal basis templates and a
reconstruction matrix that maps them back to the original templates as
described in equation~\eqref{eq:svddecomp}.  The downsampled basis templates,
the reconstruction matrix, and the time-slice boundaries are all saved to disk.

\subsection{Filtering stage}

The \lloid\ algorithm is amenable to extremely low latency implementation.
The lowest possibly latency could be achieved by integration directly into
the data acquisition and storage system of the \LIGO\ observatories.  A slightly more 
modest goal is to leverage existing signal processing software in order to implement
the \lloid\ algorithm.

We have implemented a prototype of the low-latency filtering stage using an
open-source signal processing environment called
\gstreamer\footnote{\url{http://gstreamer.net/}}.
\gstreamer\ is a vital component of many Linux systems, providing media
playback, authoring, and streaming on devices from cell phones to desktop
computers to streaming media servers.  Given the similarities of
\GW{} detector data to audio data it is not surprising that
\gstreamer\ is useful for our purpose. \gstreamer\ also provides some useful
stock signal processing elements such as resamplers and filters.  We have
extended the \gstreamer\ framework by developing a library called
\gstlal\footnote{\url{https://www.lsc-group.phys.uwm.edu/daswg/projects/gstlal.html}}
that provides elements for \GW{} data analysis.

\gstreamer\ pipelines typically operate with very low (in some consumer applications, imperceptibly low) latency rather than in true real-time because signals are partitioned into blocks of samples, or \emph{buffers}.  This affords a number of advantages, including amortizing the overhead of passing signals between elements and grouping together sequences of similar operations.  However, buffering a signal incurs a latency of up to one buffer length.  This latency can be made small at the cost of some additional overhead by making the buffers sufficiently small.

Our prototype pipeline, tested with GStreamer version 0.10.33, consists of the following stages.

\paragraph{Decimation}

The whitened detector data is reduced to successively
lower sample rates by decimation. Decimation involves applying an anti-aliasing
filter to the data, and then down-sampling by deleting samples.  We use GStreamer's
stock {\tt audioresample} element, which provides a FIR decimation filter with
a Kaiser-windowed sinc function kernel.

The detector data is provided at every power-of-two sample rate
required by the time slices described in \eqref{eq:time-slices}.  At this point,
the \gstreamer\ pipeline splits into several parallel branches to handle each
time slice.

\paragraph{Basis filters}

Next, these decimated data streams are fed into parallel \fir\ filter banks.
The basis filters are implemented using a \gstlal\ element called {\tt
lal\_firbank}, which provides a single input, multiple output bank of \fir\ filters
from a matrix of filter coefficients.  Several instances of {\tt lal\_firbank}
execute in parallel to handle each of the time slices.  Filter coefficients are taken
from the nonzero samples of the time slices in equation~\eqref{eq:time-sliced-templates},
with the zero-padding effectively accomplished by appropriate queues, or delay lines.

\paragraph{Reconstruction}

From the outputs of the \SVD\ basis filters, we form the partial \SNR\ streams
for each time  slice by multiplying by the reconstruction matrices.  This is
accomplished with the \texttt{gstlal} element \texttt{lal\_matrixmixer}.

\paragraph{Interpolation}

In order to form the early-warning output from each time slice, we have to add
the partial \SNR\ to the early-warning output from the subsequent time slices.
If the next time slice does not have the same sample rate, its output must
first be interpolated.  The \gstreamer\ element {\tt audioresample} is used again here.

\paragraph{\SNR\ accumulation}

The early-warning output for each time slice is formed by accumulating
interpolated early-warning output from the subsequent time slice.  This process
continues until we have worked our way to the \SNR\ of the original templates
at the full sample rate $f^0$.  In this way, the \lloid\ algorithm and this
implementation naturally leads to a simple early-warning pipeline.

\section{Results}
\label{sec:results}

In this section we evaluate the accuracy of the \lloid{} algorithm using our 
\gstreamer{}-based implementation described in the previous section. We calculate the
measured \SNR\ loss due to the approximations of the \lloid\ method and our
implementation of it. Using a configuration that gives acceptable \SNR{} loss for
our chosen set of source parameters, we then compare the computational cost in \flops\
for \TD\ method, the \FD\ method, and \lloid.

\subsection{Setup}
\label{sec:bank-setup}

We examine the performance of the \lloid\ algorithm on a small region of
compact binary parameter space centered on typical \NS--\NS{}
masses.  We begin by constructing a template bank that spans component masses
from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
curve~\citep{ALIGONoise}.  Templates are truncated at 10~Hz, where the projected
sensitivity of Advanced \LIGO\ is interrupted by the ``seismic wall.''
This results in a grid of 98544 points, or
$2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
the parameter space by chirp mass.  We shall concentrate on a sub-bank with 657 points with
chirp masses between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
Figure \ref{fig:tmpltbank} illustrates this procedure.
\begin{figure}[h]
	\plotone{f4}
	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
case study, consisting of component masses $m_1$, $m_2$, between 1 and 3~$M_\odot$, and
chirp masses $\mathcal{M}$ between 1.1955 and 1.2045~$M_\odot$.}
\end{figure}
With this sub-bank we were able to construct an efficient time-slice decomposition
that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
in Table~\ref{tab:time_slices}.
\begin{table*}
\caption{\label{tab:time_slices} Filter design sub-bank of 1314 templates.  From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.  We vary \SVD{} tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
\begin{center}
\begin{minipage}[c]{0.4\textwidth}
\includegraphics{t3}
\end{minipage}
\begin{minipage}[c]{0.55\textwidth}
\begin{tabular}{rr@{,\,}lc*{6}{r}}
\tableline\tableline
\\ [-2ex]
$f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD{} tolerance)} \\% [1ex]
\cline{5-10}
\\[-2.5ex]
(Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline
4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
\tableline
\end{tabular}
\end{minipage}
\end{center}
\end{table*}
We use this sub-bank and decomposition for the remainder of this section.

\subsection{Measured \SNR\ loss}

The \SNR\ loss is to be compared with the mismatch of 0.03 that arises from the
discreteness of template bank designed for a minimal match of 0.97.  We will consider
an acceptable target \SNR\ loss to be a factor of 10 smaller than this, that is, no more
than 0.003.

We expect two main contributions to the \SNR\ loss to arise in our
implementation of the \lloid\ algorithm.  The first is the \SNR\ loss due to
the truncation of the \SVD\ at $L^s < M$ basis templates.  As remarked upon in
\citet{Cannon:2010p10398} and section~\ref{sec:svd}, this effect is measured by
the \SVD\ tolerance.  The second comes from the limited bandwidth of the
interpolation filters used to match the sample rates of the partial \SNR\ streams.
The maximum possible bandwidth is determined by the length of the filter,
$N^\shortuparrow$.  \SNR\ loss could also arise if the combination of both the decimation
filters and the interpolation filters reduces their bandwidth
measurably, if the decimation and interpolation filters do not have perfectly uniform
phase response, or if there is an unintended subsample time delay at any stage.

To measure the accuracy of our \gstreamer\ implemention of \lloid\ including all of
the above potential sources of \SNR\ loss, we conducted impulse response tests.  The
\gstreamer\ pipeline was presented with an input consisting of a unit impulse.  By
recording the outputs, we can effectively ``play back'' the templates.  These impulse
responses will be similar, but not identical, to the original, nominal templates.  By
taking the inner product between the impulses responses for each output 
channel with the corresponding nominal template, we can gauge exactly how much \SNR\
is lost due to the approximations in the \lloid\ algorithm and any of the technical
imperfections mentioned above.  We call one minus this dot product the \emph{mismatch}
relative to the nominal template.

The two adjustable parameters that affect performance and mismatch the most are the SVD
tolerance and the length of the interpolation filter.  The length of the decimation filter
affects mismatch as well, but has very little impact on performance.

\paragraph{Effect of \SVD\ tolerance}

We studied how the \SVD\ tolerance affected \SNR\ loss by holding
$N^\shortdownarrow = N^\shortuparrow = 192$ fixed as we varied the \SVD\ tolerance from
$\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The minimum, maximum, and median
mismatch are shown as functions of \SVD\ tolerance in
Figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the
\SVD\ becomes an exact matrix factorization, but the computational cost increases as
the number of basis filters increases.  The conditions presented here are more
complicated than in the original work~\citep{Cannon:2010p10398} due to the inclusion
of the time-sliced templates and interpolation, though we still see that the average
mismatch is approximately proportional to the \SVD\ tolerance down to
$\left(1-10^{-4}\right)$.  However, as the \SVD\ tolerance becomes even higher, the
mismatch seems to saturate around $2 \times 10^{-4}$.  This could be the effect of the
interpolation, or an unintended technical imperfection that we did not model or expect.
However, this is still an order of magnitude below our target mismatch of 0.003.  We
find that an \SVD\ tolerance of $\left(1-10^{-4}\right)$ is adequate to achieve our
target \SNR{} loss.

\paragraph{Effect of interpolation filter length}

Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$ and the length of the
decimation filter fixed at $N^\shortdownarrow = 192$, we studied the
impact of the length $N^\shortuparrow$ of the interpolation filter on mismatch.  The mismatch as
a function of $N^\shortuparrow$ is shown in Figure~\ref{fig:bw}b.  The
mismatch saturates at $\sim$$2 \times 10^{-4}$ with $N^\shortuparrow = 64$.  We find
that a filter length of 16 is sufficient to meet our target mismatch of 0.003.

Having selected an SVD tolerance of $\left(1-10^{-4}\right)$ and $N^\shortuparrow=16$, we found that
we could reduce $N^\shortdownarrow$ to 48 without making the median mismatch higher than 0.003.

During testing, we were surprised that careful design of the decimation and interpolation stages made a crucial difference in terms of computational overhead and latency.  When we tested the pipeline with the largest interpolation filters provided by GStreamer's audioresample element, we found that rather than making a negligible contribution to the total \flops\ budget and total measured overhead they became the dominant contribution.  The largest interpolation filters also incurred latencies of several seconds.  We now appreciate that the efficiency and low latency of \lloid\ hinge on proper specification of $N^\shortuparrow$.

There is possibly further room for optimization here.  We have considered designing our own custom decimation and interpolation filters or tuning these filters separately for each time slice.

\begin{figure*}[b]
	\begin{minipage}[t]{0.5\textwidth}
		\begin{center}
			\includegraphics{f5a}
			(a) Mismatch versus \SVD\ tolerance
		\end{center}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\begin{center}
			\includegraphics{f5b}
			(b) Mismatch versus $N^\shortuparrow$
		\end{center}
	\end{minipage}
	\caption{\label{fig:bw}Box-and-whisker plot of mismatch between nominal
template bank and \lloid\ measured impulse responses.  The upper and lower boundaries of
the boxes show the upper and lower quartiles; the lines in the center denote the medians.
The whiskers represent the minimum and maximum mismatch over all templates.  In 
(a) the interpolation filter length is held fixed at $N^\shortuparrow = 192$, while
the \SVD\ tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
In (b), the \SVD\ tolerance is fixed at $\left(1-10^{-6}\right)$ while $N^\shortuparrow$
is varied from 8 to 192 coefficients.}
\end{figure*}

\subsection{Lower bounds on computational cost and latency compared to other
methods}

We are now prepared to offer the estimated computational cost of filtering this
sub-bank of templates compared to other methods.  We used the results of the
previous subsections to set the \SVD\ tolerance to $\left(1-10^{-4}\right)$,
the interpolation filter length to 16, and the decimation filter length to 48.
Table~\ref{table:flops} shows the
computational cost in \flops\ for the sub-bank we described above.  For the \FD\ method, an \fft\
block size of $\fftblock = 2 \tmpsamps$ is assumed, resulting in a latency of
$\left(\tmpsamps / f^0\right)$ seconds.  Both the \FD\ method and \lloid\ are
five orders of magnitude faster than the conventional \TD\ method.  However,
the \FD\ method has a latency of over half of an hour, whereas \lloid\ has a
latency of just over one second.
%
\begin{table*}
\caption{\label{table:flops}Computational cost in \flops\ and latency in seconds
of the \TD\ method, the \FD\ method, and \lloid.  Cost is given
for both the sub-bank described in section~\ref{sec:bank-setup} and a full
\NS--\NS\ search.  The last column gives the approximate number of machines per
detector required for a full advanced LIGO \NS--\NS\ search.}
\begin{center}
\begin{tabular}{lllll}
\tableline\tableline
& \flops\ & & \flops\ & number of \\
method & (sub-bank) & latency (s) & (\NS--\NS) & machines \\[0.1em]
\tableline
time domain & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$10^5$ \\
frequency domain & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$10$ \\
\lloid\ (theory) & $6.6\times10^8$ & $1.2$ & $\phantom{5.7}\sim 10^{10}$ & $\sim$$10$ \\
\lloid\ (prototype) & (0.9 cores) & $1.5$ & ------------ & $\sim$$10$ \\
\tableline
\end{tabular}
\end{center}
\end{table*}

\subsection{Extrapolation of computational cost to an Advanced \LIGO{} search}

Table~\ref{table:flops} shows that the \lloid\ method requires $6.6 \times 10^8$
\flops\ to cover a sub-bank comprising 657 out of the total 98544
mass pairs.  Given that modern (ca. 2011) workstations can sustain computation
rates up to $\sim$$10^{10}$~\flops{}, and assuming that other regions of the
parameter space have similar computational scaling, an entire single detector
search for \NS--\NS\ signals in the 1--3~$M_\odot$ component mass range could be
implemented with $\sim$$10^{11}$ \flops\, or $\sim$$10$ machines.

We computed the computational cost of a full advanced \LIGO\ \NS--\NS\ search a
second way by computing the cost per template for the conventional \TD\ method,
and then applying the \lloid\ speedup from equation~(\ref{eq:speedup}).  This
should be a somewhat more accurate measure because template length varies over
the parameter space.  Lower chirp mass templates sweep through frequency more
slowly and require more computations while higher chirp mass templates are
shorter and require fewer computations.  Despite these subtleties, this estimate
gave us $4.6 \times 10^9$~\flops, differing by only a factor of $\sim$2 from the
simple scaling argument above.  For this reason, we believe that $\sim$$10$
machines is a realistic estimate of our computing requirements.

By comparison, using the \TD\ method to achieve the same latency costs
$\sim$$10^{13}$ \flops\ for this particular sub-bank, and so it would require
a total of $\gtrsim$$10^{15}$ \flops\ or $\gtrsim$$10^5$ present-day machines
to search the full parameter space.  Presently, the \LIGO{} Data Grid%
\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}} consists of
only $\sim$$10^4$ machines, so direct \TD\ convolution is clearly impractical.

\subsection{Measured latency and overhead}

Our \gstreamer\ pipeline for measuring impulse responses contained
instrumentation that would not be necessary for an actual search, including
additional interpolation filters to bring the early-warning outputs back to the
full sample rate and additional outputs for recording signals to disk.

We wrote a second, stripped pipeline to evaluate the actual latency and
computational overhead.  We executed this pipeline on one of the submit
machines of the \LIGO-Caltech cluster, a Sun Microsystems Sun
Fire\texttrademark\ X4600~M2 server with eight quad-core 2.7~GHz AMD
Opteron\texttrademark\ 8384 processors.  This test consumed $\sim$90\% of the
capacity of just one out of the 32 cores, maintaining a constant latency of
$\sim$1.5~s.

The measured overhead is consistent to within an order of magnitude with the
lower bound from the \flops\ budget.  Additional overhead is possibly
dominated by thread synchronization.  A carefully optimized \gstreamer\
pipeline or a hand-tuned C implementation of the pipeline might reduce overhead
further.

Since in this benchmark, the pipeline's input was buffered in
0.25~s chunks, the total measured latency of 1.5~s is entirely consistent with
1.2~s of latency being attributed to the acausality of the decimation and
interpolation filters, as calculated in section~\ref{sec:lloid-method}.
Our implementation of the \lloid\ algorithm has achieved latencies that would
make other aspects of the analysis the limiting steps assuming that the human
vetting process is automated.

\section{Conclusions}
\label{sec:conclusions}

We have demonstrated a computationally feasible filtering algorithm for the rapid and
even early-warning detection of \GW{}s emitted during the coalescence
of neutron stars and stellar-mass black holes.  It is one part of a complicated
analysis and observation strategy that will unfortunately have other sources of
latency.  However, we hope that it will motivate further work to reduce such
technical sources of \GW{} observation latency and encourage the possibility of
even more rapid \EM\ followup observations to catch prompt emission in the
advanced detector era.

\CBC\ events may be the progenitors of some short hard \GRB{}s and are expected
to be accompanied by a broad spectrum of \EM{} signals. Rapid alerts to the
wider astronomical community will improve the chances of detecting an \EM{}
counterpart in bands from gamma-rays down to radio. In the Advanced LIGO
era, it appears possible to usefully localize a few rare events prior to the
\GRB{}, allowing multi-wavelength observations of prompt emission. More
frequently, low-latency alerts will be released after merger but will still
yield extended X-ray tails and early afterglows.

The \lloid{} method is as fast as standard \fft{} convolutions but allows for
extremely low latency operation.  We anticipate requiring $\sim$40
modern multi-core computers to search for binary neutron stars using
coincident \GW\ data from a 4-detector network.  In the future, additional
computational savings could be acheived by conditionally reconstructing the
\SNR{} time-series only during times when a composite detection statistic
crosses a threshold~\citep{svd-compdetstat}.  However, the anticipated required
number of computers is well within the current computing capabilities of the
\LIGO{} Data Grid.

We have shown a prototype implementation of the \lloid{} algorithm using
\gstreamer, an open source signal processing platform.  Work will be done in
order to improve upon the latency of our implementation.  Ultimately the best
possible latency would be achieved by tighter integration between data
acquisition and analysis with dedicated hardware and software. This could be
considered for third-generation detector design.  Also possible for
third-generation instruments, the \lloid{} method could provide the input for a
dynamic tuning of detector response via the signal recycling mirror to match
the frequency of maximum sensitivity to the instantaneous frequency of the
\GW{} waveform.  This is a challenging technique, but it has the potential for
substantial gains in \SNR{} and timing accuracy \citep{PhysRevD.47.2184}.

Although we have demonstrated a computationally feasible detection statistic
for advance detection, we have not yet explored data calibration and whitening,
triggering, coincidence, and ranking of gravitational wave candidates in a
framework that supports early \EM\ followup.
\begin{comment}
One complication in deploying a
low-latency or early-warning search is that whitening the detector data
involves estimating the noise power spectrum, which is often done acausally.
We will have to look into causal or low latency alternatives to conventional
spectral estimation techniques.  
\end{comment}
We will explore these and also using the time-slice decomposition and the \SVD\
to form low-latency signal-based vetoes (e.g.~$\chi^2$~statistics) that have
been essential for glitch rejection used in previous \GW{} \CBC{} searches.

Our future work must more deeply address sky localization accuracy in a
realistic setting as well as observing strategies. Here, we have followed
\citet{Fairhurst2009} in estimating the area of 90\% localization confidence in
terms of timing uncertainties alone, but it would be advantageous to use a
galaxy catalog to inform the telescope tiling \citep{galaxy-catalog}. Because
early detections will arise from nearby sources, the galaxy catalog technique
might be an important ingredient in reducing the fraction of sky that must be
imaged.  Extensive simulation campaigns incorporating realistic binary merger
rates and detector networks will be necessary in order to fully understand the
prospects for early-warning detection, localization, and \EM\ followup using
the techniques we have described.

\acknowledgements

\LIGO\ was constructed by the California Institute of Technology and
Massachusetts Institute of Technology with funding from the National Science
Foundation and operates under cooperative agreement
PHY-0107417.  CH would like to thank Ilya Mandel for
many discussions about rate estimates and the prospects of early detection,
Patrick Brady for countless fruitful conversations about
low latency analysis methods, and John Zweizig for
disscussions about \LIGO\ data acquisition.  NF would like to thank Alessandra
Corsi and Larry Price for illuminating discussions on astronomical motivations.
This research
is supported by the National Science Foundation through a Graduate Research
Fellowship to LS and by the Perimeter Institute for Theoretical Physics through
a fellowship to CH. DK is supported from the Max Planck Gesellschaft.

This paper has \LIGO\ Document Number {LIGO-P0900004-v25}.

\bibliographystyle{apj}
\bibliography{references}

\end{document}

### 2011-08-14 article.tex
Put back in real time claims, re-flowed some TeX code
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1347 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \section{Introduction}
-\label{sec:introduction}
 
@@
 amplitude and frequency increasing until the system eventually merges.  If a
-neutron star (\NS{}) is involved, it might become tidally disrupted near
-the merger and fuel an electromagnetic (\EM{}) counterpart
-\citep{shibata:2007}.  Effort from both the \GW{} and the broader astronomical communities might make it
-possible to use \GW{} observations as an early warning trigger for \EM{}
+neutron star (\NS) is involved, it might become tidally disrupted near the
+merger and fuel an electromagnetic (\EM) counterpart \citep{shibata:2007}.
+Effort from both the \GW{} and the broader astronomical communities might make
+it possible to use \GW\ observations as an early warning trigger for \EM{}
 followup. In the first generation of ground-based laser interferometers, the
-\GW{} community initiated a project to send alerts when potential \GW{}
+\GW\ community initiated a project to send alerts when potential \GW{}
 transients were observed in order to trigger followup observations by \EM{}
@@
 disruption, one might have the ambition of reporting \GW\ candidates not minutes
-after the merger, but seconds before.  We explore one essential ingredient of this
-problem, a computationally inexpensive low-latency filtering algorithm for detecting
-inspiral signals in \GW{} data.  We also consider the prospects for advanced \GW{}
-detectors and discuss other areas of work that would be required for rapid analysis.
+after the merger, but seconds before.  We explore one essential ingredient of
+this problem, a computationally inexpensive low-latency filtering algorithm for
+detecting inspiral signals in \GW\ data.  We also consider the prospects for
+advanced \GW\ detectors and discuss other areas of work that would be required
+for rapid analysis.
 
@@
 disrupted material falls onto the newly formed, rapidly spinning compact object
-and is accelerated in jets along the spin axis with a timescale of
-$0.1$--$1$~s after the merger \citep{Janka1999}, matching the short \GRB{}
-duration distribution well. Prompt \EM{} emission including the \GRB{}
-can arise as fast outflowing
-matter collides with slower matter ejected earlier in inner shocks. The same
-inner shocks, or potentially reverse shocks, can produce an accompanying
-optical flash \citep{Sari99}. The prompt emission is a probe into the extreme
-initial conditions of the outflow, in contrast with afterglows, which arise in
-the external shock with the local medium and are relatively insensitive to
-initial conditions. Optical flashes have been observed for a handful of long
-\GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely rapid response
-or, in the case of GRB 080319b, by pure serendipity, where several
-telescopes were already observing the afterglow of another \GRB{} in the same
-field of view \citep{2008Natur.455..183R}. The observed optical flashes peaked
-within tens of seconds and decayed quickly. For short \GRB{} energy balance and
-plasma density,
-however, the reverse shock model predicts a peak flux in radio, approximately
-20 minutes after the \GRB{}, but also a relatively faint optical flash
-\citep{nakar07}; for a once-per-year Advanced LIGO event at $130$~Mpc, the
-radio flux will peak around 9~GHz at $\sim$$5$~mJy, with emission in the
-$R$-band at $\sim$19~mag. Interestingly, roughly a quarter to half of observed short
-\GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in duration
-beginning $\sim$$10$~s after the \GRB{} and carrying comparable fluence to the
-initial outburst. This can be explained if the merger results in the formation
-of a proto-magnetar that interacts with ejecta \citep{Bucciantini2011}. Rapid
-\GW{} alerts would enable joint \EM{} and \GW{} observations to confirm
-the short \GRB{}-\CBC{} link and allow the early \EM{}
-observation of exceptionally nearby and thus bright events.
+and is accelerated in jets along the spin axis with a timescale of $0.1$--$1$~s
+after the merger \citep{Janka1999}, matching the short \GRB\ duration
+distribution well. Prompt \EM{} emission including the \GRB\ can arise as fast
+outflowing matter collides with slower matter ejected earlier in inner shocks.
+The same inner shocks, or potentially reverse shocks, can produce an
+accompanying optical flash \citep{Sari99}. The prompt emission is a probe into
+the extreme initial conditions of the outflow, in contrast with afterglows,
+which arise in the external shock with the local medium and are relatively
+insensitive to initial conditions. Optical flashes have been observed for a
+handful of long \GRB{}s \citep{2011CRPhy..12..255A} by telescopes with extremely
+rapid response or, in the case of GRB 080319b, by pure serendipity, where
+several telescopes were already observing the afterglow of another \GRB{} in
+the same field of view \citep{2008Natur.455..183R}. The observed optical flashes
+peaked within tens of seconds and decayed quickly. For short \GRB{} energy
+balance and plasma density, however, the reverse shock model predicts a peak
+flux in radio, approximately 20 minutes after the \GRB{}, but also a relatively
+faint optical flash \citep{nakar07}; for a once-per-year Advanced LIGO event at
+$130$~Mpc, the radio flux will peak around 9~GHz at $\sim$$5$~mJy, with emission
+in the $R$-band at $\sim$19~mag. Interestingly, roughly a quarter to half of
+observed short \GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in
+duration beginning $\sim$$10$~s after the \GRB{} and carrying comparable fluence
+to the initial outburst. This can be explained if the merger results in the
+formation of a proto-magnetar that interacts with
+ejecta~\citep{Bucciantini2011}. Rapid \GW\ alerts would enable joint \EM\ and
+\GW\ observations to confirm the short \GRB{}-\CBC{} link and allow the early
+\EM\ observation of exceptionally nearby and thus bright events.
 
@@
 observatories to the analysis clusters over the Internet, which is capable of
-high bandwidth but only modest latency.  Together, these introduce a
-latency of $\gtrsim$$100$~ms.  These technical sources of latency could be reduced
-with significant engineering and capital investments, but they are minor compared
+high bandwidth but only modest latency.  Together, these introduce a latency of
+$\gtrsim$$100$~ms.  These technical sources of latency could be reduced with
+significant engineering and capital investments, but they are minor compared
 to any of the other sources of latency.
@@
 \paragraph{Data conditioning ($\sim$1~min)}%
-Science data must be calibrated using the detector's frequency
-response to gravitational radiation.  Currently, data are calibrated in blocks of
-16~s.  Within $\sim$1~min, data quality is assessed in order to create veto flags.
-These are both technical sources of latency that might be addressed with improved
-calibration and data quality software for advanced detectors.
+Science data must be calibrated using the detector's frequency response to
+gravitational radiation.  Currently, data are calibrated in blocks of 16~s.
+Within $\sim$1~min, data quality is assessed in order to create veto flags.
+These are both technical sources of latency that might be addressed with
+improved calibration and data quality software for advanced detectors.
 
@@
 generation latency is too long to catch prompt or even extended emission.
-Low-latency trigger generation will become more challenging with advanced detectors
-because inspiral signals will stay in band up to ten times longer.  In this work,
-we will focus on reducing this source of latency.
+Low-latency trigger generation will become more challenging with advanced
+detectors because inspiral signals will stay in band up to ten times longer.  In
+this work, we will focus on reducing this source of latency.
 
 \paragraph{Alert generation (2--3~min)}%
-S6/VSR3 saw the introduction of low-latency astronomical
-alerts, which required gathering event parameters and sky localization from the
-various online analyses, downselecting the events, and calculating telescope pointings.
-If other sources of latency improve, the technical latency associated with this
+S6/VSR3 saw the introduction of low-latency astronomical alerts, which required
+gathering event parameters and sky localization from the various online
+analyses, downselecting the events, and calculating telescope pointings.  If
+other sources of latency improve, the technical latency associated with this
 infrastructure could dominate, so work should be done to improve it.
@@
 \paragraph{Human validation (10--20~min)}%
-Because the new alert system was commissioned during S6/VSR3, all alerts were subjected
-to quality control checks by human operators before they were disseminated.
-This was by far the largest source of latency during S6/VSR3.
-Hopefully, confidence in the system will grow to the point where no human intervention
-is necessary before alerts are sent, so we give it no further consideration here.
+Because the new alert system was commissioned during S6/VSR3, all alerts were
+subjected to quality control checks by human operators before they were
+disseminated. This was by far the largest source of latency during S6/VSR3.
+Hopefully, confidence in the system will grow to the point where no human
+intervention is necessary before alerts are sent, so we give it no further
+consideration here.
 
@@
 
-This work will focus on reducing the latency of trigger production.  Data analysis
-strategies for advance detection of \CBC{}s will have to strike a balance between latency
-and throughput. \CBC{} searches consist of banks of
+This work will focus on reducing the latency of trigger production.  Data
+analysis strategies for advance detection of \CBC{}s will have to strike a
+balance between latency and throughput. \CBC{} searches consist of banks of
 matched filters, or cross-correlations between the data stream and a bank of
@@
 detection.  Then, we provide an overview of our novel method for detecting compact
-binary coalescence signals with extremely low latency. We then describe a prototype
+binary coalescence signals near real-time. We then describe a prototype
 implementation using open source signal processing software.  To validate our approach
@@
 
-\section{Novel low latency algorithm for \CBC\ detection}
+\section{Novel \realtime\ algorithm for \CBC\ detection}
 \label{sec:method}
@@
 Here we describe a method for reducing the computational cost of a \TD\ search
-for compact binary coalescence.  We give an extremely low latency algorithm
+for compact binary coalescence.  We give a real-time algorithm
 that competes in terms of floating point operations per second with the
@@
 $\left[t^0, t^1\right),\, \left[t^1, t^2\right),\, \dots,\, \left[t^{S-1}, t^S\right)$,
-sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$ where $f^s$ is at
+sampled at frequencies $f^0,\, f^1,\, \dots,\, f^\mathrm{S-1}$, where $f^s$ is at
 least twice the highest nonzero frequency component of any filter in the bank for the
@@
 and interpolation filters because they are generally time-symmetric and slightly
-acausal.
+acausal.  As long as the latency introduced by the decimation and interpolation
+filters for any time slice $s$ is less than that time slice's delay $t^s$, the
+total latency of the \lloid\ algorithm will be zero.
 
@@
 
-The conventional \TD{} method consists of a bank of \fir{} filters, or sliding-window dot products.
-If there are $\numtmps$ templates, each $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications and additions per sample, or $2 \numtmps \tmpsamps f^0$ \flops\ at a sample rate $f^0$.
+The conventional \TD{} method consists of a bank of \fir{} filters, or
+sliding-window dot products.  If there are $\numtmps$ templates, each
+$\tmpsamps$ samples in length, then each filter requires $M N$ multiplications
+and additions per sample, or $2 \numtmps \tmpsamps f^0$ \flops\ at a sample rate
+$f^0$.
 
@@
 For time slice $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ \flops\ 
-to evaluate the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ to apply the 
-linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops\ to add the resultant partial \SNR\ stream.
-
-The computational cost of the decimation of the detector data is a little bit more subtle.
-Decimation is achieved by applying an \fir\ anti-aliasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.
-Naively, an anti-aliasing filter with $(f^{s-1} / f^s) N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.
-However, it is necessary to evaluate the anti-aliasing filter only for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.
-Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1}$ \flops\ is possible.
-
-The story is similar for the interpolation filters used to match the sample rates of the partial \SNR{} streams.
-Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a low-pass filter with $(f^{s-1} / f^s) N^\shortdownarrow$ coefficients.
-The low-pass filter requires $2 M N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.
-However, by taking advantage of the fact that by construction a fraction $f^s / f^{s-1}$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1}$ \flops.
-
-Taking into account the decimation of the detector data, the orthogonal \fir\ filters, the reconstruction of the time-sliced templates, the interpolation of \SNR\ from previous time slices, and the accumulation of \SNR, in total the \lloid\ algorithm requires
+to evaluate the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ to
+apply the  linear transformation from the $\numsvdtmps$ basis templates to the
+$\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops\ to add the
+resultant partial \SNR\ stream.
+
+The computational cost of the decimation of the detector data is a little bit
+more subtle.  Decimation is achieved by applying an \fir\ anti-aliasing filter
+and then downsampling, or deleting samples in order to reduce the sample rate
+from $f^{s-1}$ to $f^s$.  Naively, an anti-aliasing filter with
+$(f^{s-1} / f^s) N^\shortdownarrow$ coefficients should demand
+$2 N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.  However, it is necessary to
+evaluate the anti-aliasing filter only for the fraction $f^s / f^{s-1}$ of the
+samples that will not be deleted.  Consequently, an efficient decimator that
+requires only $2 N^\shortdownarrow f^{s-1}$ \flops\ is possible.
+
+The story is similar for the interpolation filters used to match the sample
+rates of the partial \SNR{} streams.  Interpolation of a data stream from a
+sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples
+of the original stream, and then applying a low-pass filter with
+$(f^{s-1} / f^s) N^\shortdownarrow$ coefficients.  The low-pass filter requires
+$2 M N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.  However, by taking advantage
+of the fact that by construction a fraction $f^s / f^{s-1}$ of the samples are
+zero, it is possible to build an efficient interpolator that requires only
+$M N^\shortuparrow f^{s-1}$ \flops.
+
+Taking into account the decimation of the detector data, the orthogonal \fir\
+filters, the reconstruction of the time-sliced templates, the interpolation of
+\SNR\ from previous time slices, and the accumulation of \SNR, in total the
+\lloid\ algorithm requires
+%
 \begin{multline*}
@@
 \end{multline*}
-\flops.
-The second sum is carried out over the set of distinct sample rates (except for the base sample rate) rather than over the time slices themselves, as we have found that it is sometimes desirable to place multiple adjacent time slices at the same sample rate in order to keep the size of matrices that enter the singular value decomposition manageable.
-Here we have assumed that the decimation filters are connected in parallel, converting from the base sample rate $f^0$ to each of the time slice sample rates $f^1$, $f^2$, $\dots$, and that the interpolation filters are connected in cascade fashion with each interpolation filter stepping from the sample rate of one time slice to the next.
-
-We can simplify this expression quite a bit by taking some limits that arise from sensible filter design.
-In the limit of many templates, the cost of the decimation filters is negligible as compared to the cost of the interpolation filters.
-Typically, we will design the interpolation filters with $N^\shortuparrow \lesssim \numsvdtmps$ so that the interpolation cost itself is negligible compared with the reconstruction cost.
-Finally, if the number of basis templates per time slices $\numsvdtmps$ is not too small, the reconstruction cost dominates over the cost of accumulating the partial \SNR.
-In these limits, the cost of \lloid\ is dominated by the basis filters themselves and the reconstruction, totaling $2 \sum_{s=0}^{S-1} f^s \numsvdtmps \left( \slicessamps + \numtmps \right)$ \flops.
-
-The latency of the \lloid\ method is entirely due to the acausality of the interpolation and decimation filters.  The latency is exactly
-$$
-\frac{1}{2} \left( N^\shortdownarrow / f^{S-1} + \sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} N^\shortuparrow / f^s \right) \textrm{ seconds.}
-$$
-The first term is due to the decimation filter that converts from the base sample rate $f^0$ to the lowest sample rate $f^{S-1}$ and the sum that comprises the second term is due to the latency of all of the interpolation filters connected in series.
+%
+\flops.  The second sum is carried out over the set of distinct sample rates
+(except for the base sample rate) rather than over the time slices themselves,
+as we have found that it is sometimes desirable to place multiple adjacent time
+slices at the same sample rate in order to keep the size of matrices that enter
+the singular value decomposition manageable.  Here we have assumed that the
+decimation filters are connected in parallel, converting from the base sample
+rate $f^0$ to each of the time slice sample rates $f^1$, $f^2$, $\dots$, and
+that the interpolation filters are connected in cascade fashion with each
+interpolation filter stepping from the sample rate of one time slice to the
+next.
+
+We can simplify this expression quite a bit by taking some limits that arise
+from sensible filter design.  In the limit of many templates, the cost of the
+decimation filters is negligible as compared to the cost of the interpolation
+filters.  Typically, we will design the interpolation filters with
+$N^\shortuparrow \lesssim \numsvdtmps$ so that the interpolation cost itself is
+negligible compared with the reconstruction cost.  Finally, if the number of
+basis templates per time slices $\numsvdtmps$ is not too small, the
+reconstruction cost dominates over the cost of accumulating the partial \SNR.
+In these limits, the cost of \lloid\ is dominated by the basis filters
+themselves and the reconstruction, totaling
+$2 \sum_{s=0}^{S-1} f^s \numsvdtmps \left( \slicessamps + \numtmps \right)$ \flops.
 
@@
 
-If the cost of the basis filters dominates, and the frequency of the templates evolves slowly enough in time, then we can use the time-frequency relationship of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional \TD\ method.
-The reduction in \flops\ is approximately
+If the cost of the basis filters dominates, and the frequency of the templates
+evolves slowly enough in time, then we can use the time-frequency relationship
+of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional
+\TD\ method.  The reduction in \flops\ is approximately
 %
@@
 %
-where $\alpha \approx \numsvdtmps / \numtmps$ is the rank reduction factor, or ratio between the number of basis templates and the number of templates.
-This approximation assumes that the frequency of the signal is evolving very slowly so that we can approximate the time slice sample rate as twice the instantaneous gravitational wave frequency, $f^s \approx 2 f(t)$, and the number of samples in the decimated time slice as the sample rate times an infinitesimally short time interval, $\slicessamps \approx 2 f(t) \, \mathrm{d}t$.
-The integral is evaluated using the power-law form of $f(t)$ from equation~(\ref{eq:fgw}).
-Substituting approximate values for a template bank designed for component masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{high} = 1/1570$~s $f_\mathrm{high} = 1570$~Hz, and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the \lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the conventional \TD\ method.
+where $\alpha \approx \numsvdtmps / \numtmps$ is the rank reduction factor, or
+ratio between the number of basis templates and the number of templates.  This
+approximation assumes that the frequency of the signal is evolving very slowly
+so that we can approximate the time slice sample rate as twice the instantaneous
+gravitational wave frequency, $f^s \approx 2 f(t)$, and the number of samples in
+the decimated time slice as the sample rate times an infinitesimally short time
+interval, $\slicessamps \approx 2 f(t) \, \mathrm{d}t$. The integral is
+evaluated using the power-law form of $f(t)$ from equation~(\ref{eq:fgw}).
+Substituting approximate values for a template bank designed for component
+masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s,
+$f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{high} = 1/1570$~s $f_\mathrm{high} = 1570$~Hz,
+and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the
+\lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the
+conventional \TD\ method.
 
 \section{Implementation}
-\label{sec:implementation}
 
@@
 
-The \lloid\ algorithm is amenable to extremely low latency implementation.
-The lowest possibly latency could be achieved by integration directly into
-the data acquisition and storage system of the \LIGO\ observatories.  A slightly more 
-modest goal is to leverage existing signal processing software in order to implement
+The \lloid\ algorithm is amenable to latency free, real-time implementation.
+However, such a real-time search pipeline would require integration directly into the data acquisition and storage systems of the \LIGO\ observatories.
+A slightly more
+modest goal is to leverage existing low-latency, but not real-time, signal processing software in order to implement
 the \lloid\ algorithm.
@@
 
-\gstreamer\ pipelines typically operate with very low (in some consumer applications, imperceptibly low) latency rather than in true real-time because signals are partitioned into blocks of samples, or \emph{buffers}.
-This affords a number of advantages, including amortizing the overhead of passing signals between elements and grouping together sequences of similar operations.
-However, buffering a signal incurs a latency of up to one buffer length.
-This latency can be made small at the cost of some additional overhead by making the buffers sufficiently small.
-
-Our prototype pipeline, tested with GStreamer version 0.10.33, consists of the following stages.
+\gstreamer\ pipelines typically operate with very low (in some consumer
+applications, imperceptibly low) latency rather than in true real-time because
+signals are partitioned into blocks of samples, or \emph{buffers}.  This affords
+a number of advantages, including amortizing the overhead of passing signals
+between elements and grouping together sequences of similar operations.
+However, buffering a signal incurs a latency of up to one buffer length.  This
+latency can be made small at the cost of some additional overhead by making the
+buffers sufficiently small.
+
+Our prototype pipeline, tested with GStreamer version 0.10.33, consists of the
+following stages.
 
@@
 
-The whitened detector data is reduced to successively
-lower sample rates by decimation. Decimation involves applying an anti-aliasing
-filter to the data, and then down-sampling by deleting samples.  We use GStreamer's
-stock {\tt audioresample} element, which provides a FIR decimation filter with
+The whitened detector data is reduced to successively lower sample rates by
+decimation. Decimation involves applying an anti-aliasing filter to the data,
+and then down-sampling by deleting samples.  We use GStreamer's stock
+\texttt{audioresample} element, which provides a FIR decimation filter with
 a Kaiser-windowed sinc function kernel.
@@
 Next, these decimated data streams are fed into parallel \fir\ filter banks.
-The basis filters are implemented using a \gstlal\ element called {\tt
-lal\_firbank}, which provides a single input, multiple output bank of \fir\ filters
-from a matrix of filter coefficients.  Several instances of {\tt lal\_firbank}
-execute in parallel to handle each of the time slices.  Filter coefficients are taken
-from the nonzero samples of the time slices in equation~\eqref{eq:time-sliced-templates},
-with the zero-padding effectively accomplished by appropriate queues, or delay lines.
+The basis filters are implemented using a \gstlal\ element called
+\texttt{lal\_firbank}, which provides a single input, multiple output bank of
+\fir\ filters from a matrix of filter coefficients.  Several instances of
+\texttt{lal\_firbank} execute in parallel to handle each of the time slices.
+Filter coefficients are taken from the nonzero samples of the time slices in
+equation~\eqref{eq:time-sliced-templates}, with the zero-padding effectively
+accomplished by appropriate queues, or delay lines.
 
@@
 the partial \SNR\ to the early-warning output from the subsequent time slices.
-If the next time slice does not have the same sample rate, its output must
-first be interpolated.  The \gstreamer\ element {\tt audioresample} is used again here.
+If the next time slice does not have the same sample rate, its output must first
+be interpolated.  The \gstreamer\ element \texttt{audioresample} is used again here.
 
@@
 \section{Results}
-\label{sec:results}
 
 In this section we evaluate the accuracy of the \lloid{} algorithm using our 
-\gstreamer{}-based implementation described in the previous section. We calculate the
-measured \SNR\ loss due to the approximations of the \lloid\ method and our
-implementation of it. Using a configuration that gives acceptable \SNR{} loss for
-our chosen set of source parameters, we then compare the computational cost in \flops\
-for \TD\ method, the \FD\ method, and \lloid.
+\gstreamer{}-based implementation described in the previous section. We calculate
+the measured \SNR\ loss due to the approximations of the \lloid\ method and our
+implementation of it. Using a configuration that gives acceptable \SNR{} loss
+for our chosen set of source parameters, we then compare the computational cost
+in \flops\ for \TD\ method, the \FD\ method, and \lloid.
 
@@
 \begin{table*}
-\caption{\label{tab:time_slices} Filter design sub-bank of 1314 templates.
-From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.
-We vary \SVD{} tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
+\caption{\label{tab:time_slices} Filter design sub-bank of 1314 templates.  From
+left to right, this table shows the sample rate, time interval, number of
+samples, and number of orthogonal templates for each time slice.  We vary \SVD\
+tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
 \begin{center}
@@
 
-Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$ and the length of the
-decimation filter fixed at $N^\shortdownarrow = 192$, we studied the
-impact of the length $N^\shortuparrow$ of the interpolation filter on mismatch.  The mismatch as
-a function of $N^\shortuparrow$ is shown in Figure~\ref{fig:bw}b.  The
-mismatch saturates at $\sim$$2 \times 10^{-4}$ with $N^\shortuparrow = 64$.  We find
-that a filter length of 16 is sufficient to meet our target mismatch of 0.003.
-
-Having selected an SVD tolerance of $\left(1-10^{-4}\right)$ and $N^\shortuparrow=16$, we found that
-we could reduce $N^\shortdownarrow$ to 48 without making the median mismatch higher than 0.003.
-
-During testing, we were surprised that careful design of the decimation and interpolation stages made a crucial difference in terms of computational overhead and latency.
-When we tested the pipeline with the largest interpolation filters provided by GStreamer's audioresample element, we found that rather than making a negligible contribution to the total \flops\ budget and total measured overhead they became the dominant contribution.
-The largest interpolation filters also incurred latencies of several seconds.
-We now appreciate that the efficiency and low latency of \lloid\ hinge on proper specification of $N^\shortuparrow$.
-
-There is possibly further room for optimization here.  We have considered designing our own custom decimation and interpolation filters or tuning these filters separately for each time slice.
+Next, keeping the \SVD\ tolerance fixed at $\left(1-10^{-6}\right)$ and the
+length of the decimation filter fixed at $N^\shortdownarrow = 192$, we studied
+the impact of the length $N^\shortuparrow$ of the interpolation filter on
+mismatch.  The mismatch as a function of $N^\shortuparrow$ is shown in
+Figure~\ref{fig:bw}b.  The mismatch saturates at $\sim$$2 \times 10^{-4}$ with
+$N^\shortuparrow = 64$.  We find that a filter length of 16 is sufficient to
+meet our target mismatch of 0.003.
+
+Having selected an SVD tolerance of $\left(1-10^{-4}\right)$ and
+$N^\shortuparrow=16$, we found that we could reduce $N^\shortdownarrow$ to 48
+without making the median mismatch higher than 0.003.
+
+During testing, we were surprised that careful design of the decimation and
+interpolation stages made a crucial difference in terms of computational
+overhead.  When we tested the pipeline with the largest interpolation filters
+provided by GStreamer's \texttt{audioresample} element, we found that rather than
+making a negligible contribution to the total \flops\ budget and total measured
+overhead they became the dominant contribution.  There is possibly further room
+for optimization beyond minimizing $N^\shortuparrow$.  We have considered
+designing our own custom decimation and interpolation filters or tuning these
+filters separately for each time slice.  We now appreciate that the efficiency
+of \lloid\ hinges on proper specification of $N^\shortuparrow$.
 
@@
 In (b), the \SVD\ tolerance is fixed at $\left(1-10^{-6}\right)$ while $N^\shortuparrow$
-is varied from 8 to 192 coefficients.}
+is varied from 8 to 192.}
 \end{figure*}
@@
 the interpolation filter length to 16, and the decimation filter length to 48.
-Table~\ref{table:flops} shows the
-computational cost in \flops\ for the sub-bank we described above.  For the \FD\ method, an \fft\
-block size of $\fftblock = 2 \tmpsamps$ is assumed, resulting in a latency of
+Table~\ref{table:flops} shows the computational cost in \flops\ for the sub-bank
+we described above.  For the \FD\ method, an \fft\ block size of
+$\fftblock = 2 \tmpsamps$ is assumed, resulting in a latency of
 $\left(\tmpsamps / f^0\right)$ seconds.  Both the \FD\ method and \lloid\ are
-five orders of magnitude faster than the conventional \TD\ method.  However,
-the \FD\ method has a latency of over half of an hour, whereas \lloid\ has a
-latency of just over one second.
+five orders of magnitude faster than the conventional \TD\ method.  However, the
+\FD\ method has a latency of over half of an hour, whereas the \lloid\ algorithm
+has no latency at all.
 %
@@
 frequency domain & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$10$ \\
-\lloid\ (theory) & $6.6\times10^8$ & $1.2$ & $\phantom{5.7}\sim 10^{10}$ & $\sim$$10$ \\
-\lloid\ (prototype) & (0.9 cores) & $1.5$ & ------------ & $\sim$$10$ \\
+\lloid\ (theory) & $6.6\times10^8$ & 0 & $\phantom{5.7}\sim 10^{10}$ & $\sim$$10$ \\
+\lloid\ (prototype) & (0.9 cores) & $0.5$ & ------------ & $\sim$$10$ \\
 \tableline
@@
 capacity of just one out of the 32 cores, maintaining a constant latency of
-$\sim$1.5~s.
+$\sim$0.5~s.
 
@@
 
-Since in this benchmark, the pipeline's input was buffered in
-0.25~s chunks, the total measured latency of 1.5~s is entirely consistent with
-1.2~s of latency being attributed to the acausality of the decimation and
-interpolation filters, as calculated in section~\ref{sec:lloid-method}.
-Our implementation of the \lloid\ algorithm has achieved latencies that would
-make other aspects of the analysis the limiting steps assuming that the human
-vetting process is automated.
+The 0.5~s latency is probably due to buffering and synchronization.  The latency
+might be reduced by carefully tuning buffer lengths at every stage in the
+pipeline.  Even without further refinements, our implementation of the \lloid\
+algorithm has achieved latencies comparable to the \LIGO\ data acquisition
+system itself.
 
 \section{Conclusions}
-\label{sec:conclusions}
-
-We have demonstrated a computationally feasible filtering algorithm for the rapid and
-even early-warning detection of \GW{}s emitted during the coalescence
+
+We have demonstrated a computationally feasible filtering algorithm for the rapid
+and even early-warning detection of \GW{}s emitted during the coalescence
 of neutron stars and stellar-mass black holes.  It is one part of a complicated
@@
 The \lloid{} method is as fast as standard \fft{} convolutions but allows for
-extremely low latency operation.  We anticipate requiring $\sim$40
+latency free, real-time operation.  We anticipate requiring $\sim$40
 modern multi-core computers to search for binary neutron stars using
@@
 Massachusetts Institute of Technology with funding from the National Science
-Foundation and operates under cooperative agreement
-PHY-0107417.  CH would like to thank Ilya Mandel for
-many discussions about rate estimates and the prospects of early detection,
-Patrick Brady for countless fruitful conversations about
-low latency analysis methods, and John Zweizig for
-disscussions about \LIGO\ data acquisition.  NF would like to thank Alessandra
-Corsi and Larry Price for illuminating discussions on astronomical motivations.
-This research
-is supported by the National Science Foundation through a Graduate Research
+Foundation and operates under cooperative agreement PHY-0107417.  CH would like
+to thank Ilya Mandel for many discussions about rate estimates and the prospects
+of early detection, Patrick Brady for countless fruitful conversations about low
+latency analysis methods, and John Zweizig for disscussions about \LIGO\ data
+acquisition.  NF would like to thank Alessandra Corsi and Larry Price for
+illuminating discussions on astronomical motivations.  LS would like thank Shaun
+Hooper for productive conversations on signal processing.  This research is
+supported by the National Science Foundation through a Graduate Research
 Fellowship to LS and by the Perimeter Institute for Theoretical Physics through

### 2011-08-14 article.tex
v26
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1348 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 As a compact binary system loses energy to gravitational waves (\GW{}s), its
-orbital separation decays, leading to a runaway inspiral with the \GW{}
+orbital separation decays, leading to a runaway inspiral with the \GW\
 amplitude and frequency increasing until the system eventually merges.  If a
@@
 merger and fuel an electromagnetic (\EM) counterpart \citep{shibata:2007}.
-Effort from both the \GW{} and the broader astronomical communities might make
-it possible to use \GW\ observations as an early warning trigger for \EM{}
+Effort from both the \GW\ and the broader astronomical communities might make
+it possible to use \GW\ observations as early warning triggers for \EM\
 followup. In the first generation of ground-based laser interferometers, the
-\GW\ community initiated a project to send alerts when potential \GW{}
-transients were observed in order to trigger followup observations by \EM{}
-telescopes.  The typical latencies were 30 minutes \citep{HugheyGWPAW2011},
-which was an important achievement, but too late to catch any prompt \EM{}
+\GW\ community initiated a project to send alerts when potential \GW\
+transients were observed in order to trigger followup observations by \EM\
+telescopes.  The typical latencies were 30 minutes \citep{HugheyGWPAW2011}.
+This was an important achievement, but too late to catch any prompt \EM\
 emission.  Since the \GW\ signal is in principle detectable even before the tidal
@@
 after the merger, but seconds before.  We explore one essential ingredient of
-this problem, a computationally inexpensive low-latency filtering algorithm for
-detecting inspiral signals in \GW\ data.  We also consider the prospects for
-advanced \GW\ detectors and discuss other areas of work that would be required
-for rapid analysis.
-
-Compact binary coalescence (\CBC{}) is a plausible progenitor for most short
+this problem, a computationally inexpensive latency free, real-time filtering
+algorithm for detecting inspiral signals in \GW\ data.  We also consider the
+prospects for advanced \GW\ detectors and discuss other areas of work that would
+be required for rapid analysis.
+
+Compact binary coalescence (\CBC) is a plausible progenitor for most short
 gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}, but the
@@
 after the merger \citep{Janka1999}, matching the short \GRB\ duration
-distribution well. Prompt \EM{} emission including the \GRB\ can arise as fast
+distribution well. Prompt \EM\ emission including the \GRB\ can arise as fast
 outflowing matter collides with slower matter ejected earlier in inner shocks.
@@
 rapid response or, in the case of GRB 080319b, by pure serendipity, where
-several telescopes were already observing the afterglow of another \GRB{} in
+several telescopes were already observing the afterglow of another \GRB\ in
 the same field of view \citep{2008Natur.455..183R}. The observed optical flashes
-peaked within tens of seconds and decayed quickly. For short \GRB{} energy
+peaked within tens of seconds and decayed quickly. For short \GRB\ energy
 balance and plasma density, however, the reverse shock model predicts a peak
-flux in radio, approximately 20 minutes after the \GRB{}, but also a relatively
+flux in radio, approximately 20 minutes after the \GRB, but also a relatively
 faint optical flash \citep{nakar07}; for a once-per-year Advanced LIGO event at
@@
 observed short \GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in
-duration beginning $\sim$$10$~s after the \GRB{} and carrying comparable fluence
+duration beginning $\sim$$10$~s after the \GRB\ and carrying comparable fluence
 to the initial outburst. This can be explained if the merger results in the
@@
 ejecta~\citep{Bucciantini2011}. Rapid \GW\ alerts would enable joint \EM\ and
-\GW\ observations to confirm the short \GRB{}-\CBC{} link and allow the early
+\GW\ observations to confirm the short \GRB-\CBC\ link and allow the early
 \EM\ observation of exceptionally nearby and thus bright events.
@@
 sixth science run (S6) and Virgo\footnote{\url{http://www.ego-gw.it/}}
-completed its third science run (VSR3).  While both \LIGO{} detectors and Virgo
+completed its third science run (VSR3).  While both \LIGO\ detectors and Virgo
 were operating, several all-sky detection pipelines operated in a low-latency
-configuration to send astronomical alerts, namely \mbta{}, Coherent
+configuration to send astronomical alerts, namely \mbta, Coherent
 WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
-\mbta{} achieved the best \GW{} trigger-generation latencies of 2--5 minutes.
+\mbta\ achieved the best \GW\ trigger-generation latencies of 2--5 minutes.
 Alerts were sent with latencies of 30--60 minutes, dominated by human vetting.
-Candidates were sent for \EM{} followup to several telescopes; Swift,
+Candidates were sent for \EM\ followup to several telescopes; Swift,
 LOFAR, ROTSE, TAROT, QUEST, SkyMapper,
@@
 There were a number of sources of latency associated with the search for
-\CBC{} signals in S6/VSR3 \citep{HugheyGWPAW2011}, listed here.
+\CBC\ signals in S6/VSR3 \citep{HugheyGWPAW2011}, listed here.
 
@@
 analysis strategies for advance detection of \CBC{}s will have to strike a
-balance between latency and throughput. \CBC{} searches consist of banks of
+balance between latency and throughput. \CBC\ searches consist of banks of
 matched filters, or cross-correlations between the data stream and a bank of
@@
 overlap-save algorithm for frequency domain (\FD) convolution, currently the
-preferred method in \GW{} searches.  The most obvious example of the latter is
+preferred method in \GW\ searches.  The most obvious example of the latter is
 time domain (\TD) convolution, which is latency-free.  However, its cost in floating
@@
 Second, inspiral filter banks consist of highly similar templates, admitting
-methods such as singular value decomposition (\SVD{}) to reduce the number of
+methods such as singular value decomposition (\SVD) to reduce the number of
 templates \citep{Cannon:2010p10398}. We will use both aspects to demonstrate
@@
 \plotone{f1}
-\caption{\label{fig:earlywarning}Expected number of \NS--\NS{}
+\caption{\label{fig:earlywarning}Expected number of \NS--\NS\
 sources that could be detectable by Advanced \LIGO\ a given number of seconds
@@
 %
-Before the \GW{} signal leaves the detection band, we can imagine examining the
-signal to noise ratio (\SNR{}) accumulated up to that point and if it is
-already significant, release an alert immediately, trading \SNR{} and sky
+Before the \GW\ signal leaves the detection band, we can imagine examining the
+signal to noise ratio (\SNR) accumulated up to that point and if it is
+already significant, release an alert immediately, trading \SNR\ and sky
 localization accuracy for pre-merger detection.
 Figure~\ref{fig:earlywarning} shows projected early detectability rates for
-\NS--\NS{} binaries in Advanced \LIGO\ assuming event rates
+\NS--\NS\ binaries in Advanced \LIGO\ assuming event rates
 predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
@@
 
-\EM\ followup requires estimating the location of the \GW{} source. The localization
+\EM\ followup requires estimating the location of the \GW\ source. The localization
 uncertainty can be estimated from the uncertainty in the time of arrival of the \GW{}s,
-which is determined by the signal's effective bandwidth and \SNR{}
+which is determined by the signal's effective bandwidth and \SNR\
 \citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
@@
 confidence area versus time of the loudest coalescence events detectable by
-Advanced \LIGO{} and Advanced Virgo.  Once per year, we expect to observe an
+Advanced \LIGO\ and Advanced Virgo.  Once per year, we expect to observe an
 event with a final \SNR\ of $\approx$27 whose location can be constrained to about
@@
 For reference, LSST should see first light around the same time as
-Advanced \LIGO{} and will have an unparalleled field of view of 9.6~deg$^2$
+Advanced \LIGO\ and will have an unparalleled field of view of 9.6~deg$^2$
 \citep{2008arXiv0805.2366I}.  However, 
@@
 galaxies from a catalog that lie near the sky location and luminosity distance
-estimate from the \GW{} signal~\citep{galaxy-catalog} as was done in S6/VSR3.
-Within the expected Advanced \LIGO{} \NS--\NS{} horizon distance,
+estimate from the \GW\ signal~\citep{galaxy-catalog} as was done in S6/VSR3.
+Within the expected Advanced \LIGO\ \NS--\NS\ horizon distance,
 the number of galaxies that can produce a given signal amplitude is much larger
-than in Initial \LIGO{} and thus the catalog will not be as useful
-for downselecting pointings for most events. However, exceptional \GW{} sources will
+than in Initial \LIGO\ and thus the catalog will not be as useful
+for downselecting pointings for most events. However, exceptional \GW\ sources will
 necessarily be extremely nearby. Within this reduced volume there will be fewer
@@
 less of a concern.  This should reduce the 90\% confidence area substantially.
-Proposed third generation \GW{} detectors would improve the \SNR\ and likewise
+Proposed third generation \GW\ detectors would improve the \SNR\ and likewise
 sky localization.
@@
 
-In this section we describe a decomposition of the \CBC{} signal space that
+In this section we describe a decomposition of the \CBC\ signal space that
 reduces \TD\ filtering cost sufficiently to allow for the
@@
 
-\subsection{Conventional \CBC{} searches}
+\subsection{Conventional \CBC\ searches}
 
@@
 parameters $\theta$ that determine the amplitude and phase evolution of the
-\GW{} strain. For systems in which the effects of spin can be ignored, the intrinsic
+\GW\ strain. For systems in which the effects of spin can be ignored, the intrinsic
 source parameters are just the component masses of the binary,
@@
  detector is a linear combination of two waveforms corresponding to the
-`$+$' and `$\times$' \GW{} polarizations.  Thus, we must design two filters
+`$+$' and `$\times$' \GW\ polarizations.  Thus, we must design two filters
 for each $\theta$.
@@
 $x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the
-\SNR{}, $\rho_i[k]$ defined as
+\SNR, $\rho_i[k]$ defined as
 %
@@
 Here we describe a method for reducing the computational cost of a \TD\ search
-for compact binary coalescence.  We give a real-time algorithm
+for compact binary coalescence.  We give a zero latency, real-time algorithm
 that competes in terms of floating point operations per second with the
-conventional \FD\ method, which by contrast requires a significant latency due to the
-inherent acausality of the Fourier transform.  Our method, called \lloid{}
-(Low Latency Online Inspiral Detection),
-involves two transformations of the templates that produce a
-network of orthogonal filters that is far more computationally
-efficient than the original bank of matched filters.
+conventional \FD\ method, which by contrast requires a significant latency due
+to the inherent acausality of the Fourier transform.  Our method, called \lloid\
+(Low Latency Online Inspiral Detection), involves two transformations of the
+templates that produce a network of orthogonal filters that is far more
+computationally efficient than the original bank of matched filters.
 
@@
 orthogonal across the parameter space, and are in fact highly redundant.
-We use the \SVD{} to approximate the template bank by a set of orthogonal
+We use the \SVD\ to approximate the template bank by a set of orthogonal
 \emph{basis filters}~\citep{Cannon:2010p10398}.  We find that this approximation
@@
 The first step of our proposed method is to divide the templates into time
-slices in a \TD{} analogue to the \FD{} decomposition described
+slices in a \TD\ analogue to the \FD\ decomposition described
 in \citet{Marion2004} and \citet{Buskulic2010}.  We decompose each template
@@
 t^S = N$.  The outputs of these new time-sliced filters
-form an ensemble of partial \SNR{} streams.  By linearity of the filtering
+form an ensemble of partial \SNR\ streams.  By linearity of the filtering
 process, these partial \SNR\ streams can be summed to reproduce the
@@
 section we examine how to reduce the number of filters within each time slice
-via \SVD{} of the time-sliced templates.
-
-\subsubsection{Reducing the number of filters with the \SVD{}}
+via \SVD\ of the time-sliced templates.
+
+\subsubsection{Reducing the number of filters with the \SVD}
 \label{sec:svd}
@@
 
-Because the sets of filters from each time slice form orthogonal subspaces, and the basis
-basis filters within a given time slice are mutually orthogonal, the set of all basis filters
-from all time slices forms an orthogonal basis spanning the original templates.
-
-In the next section we describe how we form our early-warning
-detection statistic using the time slice decomposition and the \SVD.
+Because the sets of filters from each time slice form orthogonal subspaces, and
+the basis filters within a given time slice are mutually orthogonal, the set of
+all basis filters from all time slices forms an orthogonal basis spanning the
+original templates.
+
+In the next section we describe how we form our early-warning detection
+statistic using the time slice decomposition and the \SVD.
 
@@
 
-In the previous two sections we have described two transformations that greatly
-reduce the computational burden of \TD\ filtering.  We are now prepared to define
-our detection statistic, the early-warning output, and to comment on the computational
-cost of evaluating it.
-
-First, the sample rate of the detector data must be decimated to match sample rates
-with each of the time slices.  We will denote the decimated detector data streams using
-a superscript $^s$ to indicate the time slices to which they correspond, and the
-operator $H^\shortdownarrow$ will represent the appropriate decimation filter:
+In the previous two sections, we described two transformations that greatly
+reduce the computational burden of \TD\ filtering.  We are now prepared to
+define our detection statistic, the early-warning output, and to comment on the
+computational cost of evaluating it.
+
+First, the sample rate of the detector data must be decimated to match sample
+rates with each of the time slices.  We will denote the decimated detector data
+streams using a superscript $^s$ to indicate the time slices to which they
+correspond, and the operator $H^\shortdownarrow$ will represent the appropriate
+decimation filter:
 \begin{equation*}
@@
 		\includegraphics{f3}
-		\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
+		\caption{\label{fig:pipeline} Schematic of \lloid\ pipeline illustrating
 signal flow.  Circles with arrows represent interpolation
@@
 \protect\includegraphics{f3c}.  Squares
-\protect\includegraphics{f3d} stand for \fir{} filters.  Sample
+\protect\includegraphics{f3d} stand for \fir\ filters.  Sample
 rate decreases from the top of the diagram to the bottom.  In this diagram each
@@
 In the next section we compute the expected computational cost scaling of this
-decomposition and compare it with the brute-force \TD{} implementation of
-\eqref{eq:SNRTD} and higher latency \FD{} methods.
+decomposition and compare it with the brute-force \TD\ implementation of
+\eqref{eq:SNRTD} and higher latency \FD\ methods.
 
@@
 
-We now examine the computational cost scaling of the conventional \TD{} or
-\FD{} matched filter procedure as compared with \lloid{}.  For convenience,
+We now examine the computational cost scaling of the conventional \TD\ or
+\FD\ matched filter procedure as compared with \lloid.  For convenience,
 Table~\ref{tab:recap} provides a review of the notation that we will need in
@@
 
-\subsubsection{Conventional \TD{} method}
-
-The conventional \TD{} method consists of a bank of \fir{} filters, or
+\subsubsection{Conventional \TD\ method}
+
+The conventional \TD\ method consists of a bank of \fir\ filters, or
 sliding-window dot products.  If there are $\numtmps$ templates, each
@@
 
-\subsubsection{Conventional \FD{} method}
-
-The most common \FD{} method is known as the \emph{overlap-save} algorithm, described in
+\subsubsection{Conventional \FD\ method}
+
+The most common \FD\ method is known as the \emph{overlap-save} algorithm, described in
 \citet{numerical-recipes-chapter-13}.  It entails splitting the input into blocks of $D$
@@
 The story is similar for the interpolation filters used to match the sample
-rates of the partial \SNR{} streams.  Interpolation of a data stream from a
+rates of the partial \SNR\ streams.  Interpolation of a data stream from a
 sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples
@@
 zero, it is possible to build an efficient interpolator that requires only
-$M N^\shortuparrow f^{s-1}$ \flops.
+$2 M N^\shortuparrow f^{s-1}$ \flops.
 
@@
 In this section we describe an implementation of the \lloid\ method described
-in section \ref{sec:method} suitable for rapid \GW{}
+in section \ref{sec:method} suitable for rapid \GW\
 searches for \CBC{}s.  The \lloid\ method requires several
@@
 parameter space into smaller sub-banks also reduces the offline cost of the
-\SVD{} and is the approach considered in \citet{Cannon:2010p10398}.  Next, we choose
+\SVD\ and is the approach considered in \citet{Cannon:2010p10398}.  Next, we choose
 time-slice boundaries as in equation~\eqref{eq:time-sliced-templates} such that all
@@
 computers to streaming media servers.  Given the similarities of
-\GW{} detector data to audio data it is not surprising that
+\GW\ detector data to audio data it is not surprising that
 \gstreamer\ is useful for our purpose. \gstreamer\ also provides some useful
@@
 \gstlal\footnote{\url{https://www.lsc-group.phys.uwm.edu/daswg/projects/gstlal.html}}
-that provides elements for \GW{} data analysis.
+that provides elements for \GW\ data analysis.
 
@@
 
-In this section we evaluate the accuracy of the \lloid{} algorithm using our 
+In this section we evaluate the accuracy of the \lloid\ algorithm using our 
 \gstreamer{}-based implementation described in the previous section. We calculate
 the measured \SNR\ loss due to the approximations of the \lloid\ method and our
-implementation of it. Using a configuration that gives acceptable \SNR{} loss
+implementation of it. Using a configuration that gives acceptable \SNR\ loss
 for our chosen set of source parameters, we then compare the computational cost
@@
 We examine the performance of the \lloid\ algorithm on a small region of
-compact binary parameter space centered on typical \NS--\NS{}
+compact binary parameter space centered on typical \NS--\NS\
 masses.  We begin by constructing a template bank that spans component masses
-from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO{} noise
+from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO\ noise
 curve~\citep{ALIGONoise}.  Templates are truncated at 10~Hz, where the projected
@@
 \\ [-2ex]
-$f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD{} tolerance)} \\% [1ex]
+$f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD\ tolerance)} \\% [1ex]
 \cline{5-10}
@@
 The maximum possible bandwidth is determined by the length of the filter,
-$N^\shortuparrow$.  \SNR\ loss could also arise if the combination of both the decimation
-filters and the interpolation filters reduces their bandwidth
+$N^\shortuparrow$.  \SNR\ loss could also arise if the combination of both the
+decimation filters and the interpolation filters reduces their bandwidth
 measurably, if the decimation and interpolation filters do not have perfectly uniform
@@
 recording the outputs, we can effectively ``play back'' the templates.  These impulse
-responses will be similar, but not identical, to the original, nominal templates.  By
-taking the inner product between the impulses responses for each output 
+responses will be similar, but not identical, to the original, nominal templates.
+By taking the inner product between the impulses responses for each output 
 channel with the corresponding nominal template, we can gauge exactly how much \SNR\
@@
 
-The two adjustable parameters that affect performance and mismatch the most are the SVD
-tolerance and the length of the interpolation filter.  The length of the decimation filter
-affects mismatch as well, but has very little impact on performance.
+The two adjustable parameters that affect performance and mismatch the most are
+the SVD tolerance and the length of the interpolation filter.  The length of the
+decimation filter affects mismatch as well, but has very little impact on
+performance.
 
@@
 We studied how the \SVD\ tolerance affected \SNR\ loss by holding
-$N^\shortdownarrow = N^\shortuparrow = 192$ fixed as we varied the \SVD\ tolerance from
-$\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The minimum, maximum, and median
-mismatch are shown as functions of \SVD\ tolerance in
-Figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the
-\SVD\ becomes an exact matrix factorization, but the computational cost increases as
+$N^\shortdownarrow = N^\shortuparrow = 192$ fixed as we varied the \SVD\
+tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The
+minimum, maximum, and median mismatch are shown as functions of \SVD\ tolerance
+in Figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the \SVD\
+becomes an exact matrix factorization, but the computational cost increases as
 the number of basis filters increases.  The conditions presented here are more
-complicated than in the original work~\citep{Cannon:2010p10398} due to the inclusion
-of the time-sliced templates and interpolation, though we still see that the average
-mismatch is approximately proportional to the \SVD\ tolerance down to
-$\left(1-10^{-4}\right)$.  However, as the \SVD\ tolerance becomes even higher, the
-mismatch seems to saturate around $2 \times 10^{-4}$.  This could be the effect of the
-interpolation, or an unintended technical imperfection that we did not model or expect.
-However, this is still an order of magnitude below our target mismatch of 0.003.  We
-find that an \SVD\ tolerance of $\left(1-10^{-4}\right)$ is adequate to achieve our
-target \SNR{} loss.
+complicated than in the original work~\citep{Cannon:2010p10398} due to the
+inclusion of the time-sliced templates and interpolation, though we still see
+that the average mismatch is approximately proportional to the \SVD\ tolerance
+down to $\left(1-10^{-4}\right)$.  However, as the \SVD\ tolerance becomes even
+higher, the mismatch seems to saturate around $2 \times 10^{-4}$.  This could be
+the effect of the interpolation, or an unintended technical imperfection that we
+did not model or expect.  However, this is still an order of magnitude below our
+target mismatch of 0.003.  We find that an \SVD\ tolerance of
+$\left(1-10^{-4}\right)$ is adequate to achieve our target \SNR\ loss.
 
@@
 interpolation stages made a crucial difference in terms of computational
-overhead.  When we tested the pipeline with the largest interpolation filters
-provided by GStreamer's \texttt{audioresample} element, we found that rather than
-making a negligible contribution to the total \flops\ budget and total measured
-overhead they became the dominant contribution.  There is possibly further room
-for optimization beyond minimizing $N^\shortuparrow$.  We have considered
-designing our own custom decimation and interpolation filters or tuning these
-filters separately for each time slice.  We now appreciate that the efficiency
-of \lloid\ hinges on proper specification of $N^\shortuparrow$.
+overhead.  Connecting the interpolation filters in cascade fashion rather than
+in parallel made a significant difference.  Also, between $N^\shortuparrow = 16$
+and $N^\shortuparrow = 192$ the total cost in \flops\ varied across an order of
+magnitude.  There is possibly further room for optimization beyond minimizing
+$N^\shortuparrow$.  We have considered designing our own custom decimation and
+interpolation filters or tuning these filters separately for each time slice.
+We now appreciate that the efficiency of \lloid\ hinges on proper specification
+of $N^\shortuparrow$.
 
@@
 \NS--\NS\ search.  The last column gives the approximate number of machines per
-detector required for a full advanced LIGO \NS--\NS\ search.}
+detector required for a full Advanced LIGO \NS--\NS\ search.}
 \begin{center}
@@
 
-\subsection{Extrapolation of computational cost to an Advanced \LIGO{} search}
+\subsection{Extrapolation of computational cost to an Advanced \LIGO\ search}
 
@@
 
-We computed the computational cost of a full advanced \LIGO\ \NS--\NS\ search a
+We computed the computational cost of a full Advanced \LIGO\ \NS--\NS\ search a
 second way by computing the cost per template for the conventional \TD\ method,
@@
 a total of $\gtrsim$$10^{15}$ \flops\ or $\gtrsim$$10^5$ present-day machines
-to search the full parameter space.  Presently, the \LIGO{} Data Grid%
+to search the full parameter space.  Presently, the \LIGO\ Data Grid%
 \footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}} consists of
@@
 latency.  However, we hope that it will motivate further work to reduce such
-technical sources of \GW{} observation latency and encourage the possibility of
+technical sources of \GW\ observation latency and encourage the possibility of
 even more rapid \EM\ followup observations to catch prompt emission in the
@@
 \CBC\ events may be the progenitors of some short hard \GRB{}s and are expected
-to be accompanied by a broad spectrum of \EM{} signals. Rapid alerts to the
-wider astronomical community will improve the chances of detecting an \EM{}
+to be accompanied by a broad spectrum of \EM\ signals. Rapid alerts to the
+wider astronomical community will improve the chances of detecting an \EM\
 counterpart in bands from gamma-rays down to radio. In the Advanced LIGO
@@
 
-The \lloid{} method is as fast as standard \fft{} convolutions but allows for
+The \lloid\ method is as fast as standard \fft\ convolutions but allows for
 latency free, real-time operation.  We anticipate requiring $\sim$40
@@
 computational savings could be acheived by conditionally reconstructing the
-\SNR{} time-series only during times when a composite detection statistic
+\SNR\ time-series only during times when a composite detection statistic
 crosses a threshold~\citep{svd-compdetstat}.  However, the anticipated required
 number of computers is well within the current computing capabilities of the
-\LIGO{} Data Grid.
-
-We have shown a prototype implementation of the \lloid{} algorithm using
+\LIGO\ Data Grid.
+
+We have shown a prototype implementation of the \lloid\ algorithm using
 \gstreamer, an open source signal processing platform.  Work will be done in
@@
 considered for third-generation detector design.  Also possible for
-third-generation instruments, the \lloid{} method could provide the input for a
+third-generation instruments, the \lloid\ method could provide the input for a
 dynamic tuning of detector response via the signal recycling mirror to match
 the frequency of maximum sensitivity to the instantaneous frequency of the
-\GW{} waveform.  This is a challenging technique, but it has the potential for
-substantial gains in \SNR{} and timing accuracy \citep{PhysRevD.47.2184}.
+\GW\ waveform.  This is a challenging technique, but it has the potential for
+substantial gains in \SNR\ and timing accuracy \citep{PhysRevD.47.2184}.
 
@@
 triggering, coincidence, and ranking of gravitational wave candidates in a
-framework that supports early \EM\ followup.
-\begin{comment}
-One complication in deploying a
-low-latency or early-warning search is that whitening the detector data
-involves estimating the noise power spectrum, which is often done acausally.
-We will have to look into causal or low latency alternatives to conventional
-spectral estimation techniques.  
-\end{comment}
-We will explore these and also using the time-slice decomposition and the \SVD\
-to form low-latency signal-based vetoes (e.g.~$\chi^2$~statistics) that have
-been essential for glitch rejection used in previous \GW{} \CBC{} searches.
+framework that supports early \EM\ followup.  We will explore these and also
+using the time-slice decomposition and the \SVD\ to form low-latency
+signal-based vetoes (e.g.~$\chi^2$~statistics) that have been essential for
+glitch rejection used in previous \GW\ \CBC\ searches.
 
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v25}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v26}.
 

### 2011-08-15 article.tex
Nick's comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1349 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-During testing, we were surprised that careful design of the decimation and
-interpolation stages made a crucial difference in terms of computational
-overhead.  Connecting the interpolation filters in cascade fashion rather than
-in parallel made a significant difference.  Also, between $N^\shortuparrow = 16$
-and $N^\shortuparrow = 192$ the total cost in \flops\ varied across an order of
-magnitude.  There is possibly further room for optimization beyond minimizing
-$N^\shortuparrow$.  We have considered designing our own custom decimation and
-interpolation filters or tuning these filters separately for each time slice.
-We now appreciate that the efficiency of \lloid\ hinges on proper specification
-of $N^\shortuparrow$.
+We found that careful design of the decimation and interpolation stages made a
+crucial difference in terms of computational overhead.  Connecting the
+interpolation filters in cascade fashion rather than in parallel made a
+significant difference.  Also, between $N^\shortuparrow = 16$ and
+$N^\shortuparrow = 192$ the total cost in \flops\ varied across an order of
+magnitude.  We now appreciate that the efficiency of \lloid\ hinges on proper
+specification of $N^\shortuparrow$.
 

### 2011-08-15 article.tex
Comments from call
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1350 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $$
-\left(1 + \frac{\textrm{template length}}{\textrm{latency}}\right)
+\left(1 + \frac{\textrm{template length}}{\textrm{latency}}\right) \left( 2 f^0 M \lg N \right).
 $$
-per template per $2\lg N$.
 
@@
 $N^\shortuparrow=16$, we found that we could reduce $N^\shortdownarrow$ to 48
-without making the median mismatch higher than 0.003.
+without exceeding a maximum mismatch of~0.003.
 
@@
 crucial difference in terms of computational overhead.  Connecting the
-interpolation filters in cascade fashion rather than in parallel made a
-significant difference.  Also, between $N^\shortuparrow = 16$ and
-$N^\shortuparrow = 192$ the total cost in \flops\ varied across an order of
-magnitude.  We now appreciate that the efficiency of \lloid\ hinges on proper
-specification of $N^\shortuparrow$.
+interpolation filters in cascade fashion rather than in parallel resulted in a
+significant speedup.  Also, only the shortest interpolation filters that met our
+maximum mismatch constraint resulted in a sub-dominant contribution to the
+overall cost.  There is possibly further room for optimization beyond minimizing
+$N^\shortuparrow$.  We could design custom decimation and interpolation filters,
+or we could tune these filters separately for each time slice.
 
@@
 for both the sub-bank described in section~\ref{sec:bank-setup} and a full
-\NS--\NS\ search.  The last column gives the approximate number of machines per
+1--3~$M_\odot$ \NS--\NS\ search.  The last column gives the approximate number of machines per
 detector required for a full Advanced LIGO \NS--\NS\ search.}
@@
 
-Although we have demonstrated a computationally feasible detection statistic
+Although we have demonstrated a computationally feasible statistic
 for advance detection, we have not yet explored data calibration and whitening,

### 2011-08-16 article.tex
flops fact checking
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1351 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
-\begin{multline*}
+\begin{multline}
+\label{eq:lloid-flops}
 \sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s \\ + 2\sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} \left( N^\shortdownarrow f^0 + \numtmps N^\shortuparrow f^{s-1}\right)
-\end{multline*}
+\end{multline}
 %
@@
 %
-\begin{multline}
+\begin{multline*}
 \label{eq:speedup}
@@
 = \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}) - t_\mathrm{high} f^2 (t_\mathrm{high}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{high}\right)}
-\end{multline}
+\end{multline*}
 %
@@
 masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s,
-$f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{high} = 1/1570$~s $f_\mathrm{high} = 1570$~Hz,
+$f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{high} = 1/1570$~s, $f_\mathrm{high} = 1570$~Hz,
 and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the
@@
 \tableline
-time domain & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$10^5$ \\
-frequency domain & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$10$ \\
-\lloid\ (theory) & $6.6\times10^8$ & 0 & $\phantom{5.7}\sim 10^{10}$ & $\sim$$10$ \\
-\lloid\ (prototype) & (0.9 cores) & $0.5$ & ------------ & $\sim$$10$ \\
+time domain & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$3.8\times10^5$ \\
+frequency domain & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$5.9$ \\
+\lloid\ (theory) & $6.6\times10^8$ & 0 & $1.1 \times 10^{11}$ & $\sim$$11$ \\
+\lloid\ (prototype) & (0.9 cores) & $0.5$ & ------------ & $\gtrsim$$10$ \\
 \tableline
@@
 We computed the computational cost of a full Advanced \LIGO\ \NS--\NS\ search a
-second way by computing the cost per template for the conventional \TD\ method,
-and then applying the \lloid\ speedup from equation~(\ref{eq:speedup}).  This
-should be a somewhat more accurate measure because template length varies over
-the parameter space.  Lower chirp mass templates sweep through frequency more
-slowly and require more computations while higher chirp mass templates are
-shorter and require fewer computations.  Despite these subtleties, this estimate
-gave us $4.6 \times 10^9$~\flops, differing by only a factor of $\sim$2 from the
-simple scaling argument above.  For this reason, we believe that $\sim$$10$
-machines is a realistic estimate of our computing requirements.
+second way by dividing the entire 1--3~$M_\odot$ parameter space into sub-banks
+of 657 points apiece, performing time slices and SVDs for each sub-bank, and
+tabulating the number of floating point operations using
+expression~(\ref{eq:lloid-flops}).  This should be a much more accurate measure
+because template length varies over the parameter space.  Lower chirp mass
+templates sweep through frequency more slowly and require more computations
+while higher chirp mass templates are shorter and require fewer computations.
+Despite these subtleties, this estimate gave us $1.1 \times 10^{11}$~\flops,
+agreeing with the simple scaling argument above.  Therefore we believe that
+$\sim$$10$ machines is a realistic estimate of our computing requirements.
 
@@
 The \lloid\ method is as fast as standard \fft\ convolutions but allows for
-latency free, real-time operation.  We anticipate requiring $\sim$40
+latency free, real-time operation.  We anticipate requiring $\gtrsim$40
 modern multi-core computers to search for binary neutron stars using
@@
 signal-based vetoes (e.g.~$\chi^2$~statistics) that have been essential for
-glitch rejection used in previous \GW\ \CBC\ searches.
+glitch rejection used in previous \GW\ \CBC\ searches.  These additional stages
+may incur some extra overhead, so computing requirements will likely be somewhat
+higher than our estimates.
 

### 2011-08-16 article.tex
Better estimate of computational scaling to all of aLIGO; more realistic flop/s numbers
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1352 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Table~\ref{table:flops} shows that the \lloid\ method requires $6.6 \times 10^8$
-\flops\ to cover a sub-bank comprising 657 out of the total 98544
-mass pairs.  Given that modern (ca. 2011) workstations can sustain computation
-rates up to $\sim$$10^{10}$~\flops{}, and assuming that other regions of the
-parameter space have similar computational scaling, an entire single detector
-search for \NS--\NS\ signals in the 1--3~$M_\odot$ component mass range could be
-implemented with $\sim$$10^{11}$ \flops\, or $\sim$$10$ machines.
+\flops\ to cover a sub-bank comprising 657 out of the total 98544 mass pairs.
+Assuming that other regions of the parameter space have similar computational
+scaling, an entire single detector search for \NS--\NS\ signals in the
+1--3~$M_\odot$ component mass range could be implemented with
+$\sim$$10^{11}$~\flops.  Modern (ca. 2011) workstations can achieve peak
+computation rates up to $\sim$$10^{11}$~\flops.  In practice, we expect that a
+software implementation of \lloid\ will reach average computation rates that are
+perhaps a factor 10 less than this, $\sim$$10^{10}$~\flops\ per machine, due to
+non-floating point tasks including bookkeeping and thread synchronization.
+Given these considerations, we estimate that a full single-detector Advanced
+\LIGO\ search with \lloid\ will require about $\sim$$10$ machines.
 
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v26}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v27}.
 

### 2011-08-16 article.tex
Decimation is from base sample rate
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1353 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 correspond, and the operator $H^\shortdownarrow$ will represent the appropriate
-decimation filter:
+decimation filter that converts between the base sample rate $f^0$ and the
+reduced sample rate $f^s$:
 \begin{equation*}
 \label{eq:decomp}
-	x^{s+1}[k] = \left( H^\shortdownarrow x^s\right)[k].
+	x^{s}[k] = \left( H^\shortdownarrow x^0\right)[k].
 \end{equation*}
-We shall use the symbol $H^\shortuparrow$ to represent an interpolation filter,
+We shall use the symbol $H^\shortuparrow$ to represent an interpolation filter
+that converts between sample rates $f^{s+1}$ and $f^s$ of adjacent time slices,
 \begin{equation*}

### 2011-08-16 article.tex
Split two sentences that don't need to be together
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1354 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 streams using a superscript $^s$ to indicate the time slices to which they
-correspond, and the operator $H^\shortdownarrow$ will represent the appropriate
+correspond.  The operator $H^\shortdownarrow$ will represent the appropriate
 decimation filter that converts between the base sample rate $f^0$ and the

### 2011-08-16 article.tex
Fixed broken equation reference
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1355 ccdd05d5-56c1-461a-bf98-85f6876523b4
Tense agreement
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1356 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
-\begin{multline*}
+\begin{multline}
 \label{eq:speedup}
@@
 = \frac{16 \alpha \left(t_\mathrm{low} f^2 (t_\mathrm{low}) - t_\mathrm{high} f^2 (t_\mathrm{high}) \right)}{\left(f^0\right)^2 \left(t_\mathrm{low} - t_\mathrm{high}\right)}
-\end{multline*}
+\end{multline}
 %
@@
 $2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
-the parameter space by chirp mass.  We shall concentrate on a sub-bank with 657 points with
-chirp masses between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
-Figure \ref{fig:tmpltbank} illustrates this procedure.
+the parameter space by chirp mass.  Figure \ref{fig:tmpltbank} illustrates this
+procedure. We concentrate on a sub-bank with 657 points with chirp masses
+between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
 \begin{figure}[h]
@@
 \end{figure}
-With this sub-bank we were able to construct an efficient time-slice decomposition
-that consisted of 11 time slices with sample rates between 32 and 4096 Hz summarized
+With this sub-bank we are able to construct an efficient time-slice decomposition
+that consists of 11 time slices with sample rates between 32 and 4096 Hz summarized
 in Table~\ref{tab:time_slices}.

### 2011-08-17 article.tex
Bumped DCC number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1357 ccdd05d5-56c1-461a-bf98-85f6876523b4
Clarified that we are talking about median mismatch
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1360 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 down to $\left(1-10^{-4}\right)$.  However, as the \SVD\ tolerance becomes even
-higher, the mismatch seems to saturate around $2 \times 10^{-4}$.  This could be
-the effect of the interpolation, or an unintended technical imperfection that we
-did not model or expect.  However, this is still an order of magnitude below our
-target mismatch of 0.003.  We find that an \SVD\ tolerance of
-$\left(1-10^{-4}\right)$ is adequate to achieve our target \SNR\ loss.
+higher, the median mismatch seems to saturate around $2 \times 10^{-4}$.  This
+could be the effect of the interpolation, or an unintended technical
+imperfection that we did not model or expect.  However, this is still an order
+of magnitude below our target mismatch of 0.003.  We find that an \SVD\
+tolerance of $\left(1-10^{-4}\right)$ is adequate to achieve our target \SNR\
+loss.
 
@@
 $N^\shortuparrow=16$, we found that we could reduce $N^\shortdownarrow$ to 48
-without exceeding a maximum mismatch of~0.003.
+without exceeding a median mismatch of~0.003.
 
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v27}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v28}.
 

### 2011-08-17 article.tex
Added RIT as an affiliation for Missy
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1361 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	Nickolas Fotopoulos\altaffilmark{3},
-	Melissa Frei\altaffilmark{5},
-	Chad Hanna\altaffilmark{6},
-	Erin Kara\altaffilmark{7},
-	Drew Keppel\altaffilmark{8,9},
-	Laura Liao\altaffilmark{10},
+	Melissa Frei\altaffilmark{5,6},
+	Chad Hanna\altaffilmark{7},
+	Erin Kara\altaffilmark{8},
+	Drew Keppel\altaffilmark{9,10},
+	Laura Liao\altaffilmark{11},
 	Stephen Privitera\altaffilmark{3},
@@
 \altaffiltext{5}{University of Texas at Austin, Austin, TX, USA}
-\altaffiltext{6}{Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
-\altaffiltext{7}{Department of Physics and Astronomy, Barnard College, Columbia University, New York, NY, USA}
-\altaffiltext{8}{Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
-\altaffiltext{9}{Leibniz Universit\"{a}t Hannover, Hannover, Germany}
-\altaffiltext{10}{Ryerson University, Toronto, ON, Canada}
+\altaffiltext{6}{Center for Computational Relativity and Gravitation and School of Mathematical Sciences, Rochester Institute of Technology, NY, USA}
+\altaffiltext{7}{Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
+\altaffiltext{8}{Department of Physics and Astronomy, Barnard College, Columbia University, New York, NY, USA}
+\altaffiltext{9}{Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
+\altaffiltext{10}{Leibniz Universit\"{a}t Hannover, Hannover, Germany}
+\altaffiltext{11}{Ryerson University, Toronto, ON, Canada}
 

### 2011-08-17 article.tex
Added city for RIT
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1364 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \altaffiltext{5}{University of Texas at Austin, Austin, TX, USA}
-\altaffiltext{6}{Center for Computational Relativity and Gravitation and School of Mathematical Sciences, Rochester Institute of Technology, NY, USA}
+\altaffiltext{6}{Center for Computational Relativity and Gravitation and School of Mathematical Sciences, Rochester Institute of Technology, Rochester, NY, USA}
 \altaffiltext{7}{Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}

### 2011-09-05 article.tex
fixed some typos
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1373 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 of the original stream, and then applying a low-pass filter with
-$(f^{s-1} / f^s) N^\shortdownarrow$ coefficients.  The low-pass filter requires
-$2 M N^\shortdownarrow (f^{s-1})^2 / f^s$ \flops.  However, by taking advantage
+$(f^{s-1} / f^s) N^\shortuparrow$ coefficients.  The low-pass filter requires
+$2 M N^\shortuparrow (f^{s-1})^2 / f^s$ \flops.  However, by taking advantage
 of the fact that by construction a fraction $f^s / f^{s-1}$ of the samples are

### 2011-09-07 article.tex
added reference to reduced basis function paper
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1374 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Second, inspiral filter banks consist of highly similar templates, admitting
-methods such as singular value decomposition (\SVD) to reduce the number of
-templates \citep{Cannon:2010p10398}. We will use both aspects to demonstrate
-that a very low latency detection statistic is possible with current computing
-resources.  Assuming the other technical sources of latency can be reduced
-significantly, this could make it possible to send prompt alerts to the
-astronomical community.
+methods such as the singular value decomposition (\SVD)~\citep{Cannon:2010p10398}
+or the Gram-Schmidt process~\citep{rbf} to reduce the number of templates.  We
+will use both aspects to demonstrate that a very low latency detection statistic
+is possible with current computing resources.  Assuming the other technical
+sources of latency can be reduced significantly, this could make it possible to
+send prompt alerts to the astronomical community.
 

### 2011-10-05 article.tex
Added Missy's grant
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1375 ccdd05d5-56c1-461a-bf98-85f6876523b4
Addressed some of referee's comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1376 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 were operating, several all-sky detection pipelines operated in a low-latency
-configuration to send astronomical alerts, namely \mbta, Coherent
-WaveBurst, and Omega \citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
+configuration to send astronomical alerts, namely Coherent WaveBurst (cWB), Omega, and Multi-Band Template Analysis (\mbta)~\citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
+cWB and Omega are both unmodeled searches for bursts based on time-frequency decomposition of the GW data.
+\mbta\ is a novel kind of template-based inspiral search that was purpose-built for low latency operation.
 \mbta\ achieved the best \GW\ trigger-generation latencies of 2--5 minutes.
@@
 develop low-latency \CBC\ search pipelines with tractable computing requirements.
-Examples include \mbta{}~\citep{Marion2004, Buskulic2010}, which was deployed in
-S6/VSR3, and a novel approach using networks of IIR filters that is being explored
+One example is \mbta{}~\citep{Marion2004, Buskulic2010}, which was deployed in
+S6/VSR3.
+MBTA consists of two \FD\ template banks for two different frequency bands, one which is matched to the early inspiral and the other which is matched to the late inspiral.
+Time coincidences are sought between triggers in both template banks.
+Another novel approach using networks of parallel, second order IIR filters is being explored
 by \citet{shaunIIR, linqingIIR}.
@@
 is truncated at some prescribed time $t^0$, or equivalent frequency $f_\mathrm{high}$,
-often chosen to correspond to the \ISCO. The beginning of the template is critically
+often chosen to correspond to the innermost stable circular orbit (\ISCO). The beginning of the template is critically
 sampled at $2 f_\mathrm{low}$, but the end of the template is critically sampled at a
@@
 \end{equation*}
-Ultimately, the latency of the entire \lloid\ algorithm is set by the decimation
-and interpolation filters because they are generally time-symmetric and slightly
-acausal.  As long as the latency introduced by the decimation and interpolation
-filters for any time slice $s$ is less than that time slice's delay $t^s$, the
-total latency of the \lloid\ algorithm will be zero.
 
@@
 recursion relation as a multirate filter network with a number of early-warning outputs.
+
+Ultimately, the latency of the entire \lloid\ algorithm is set by the decimation
+and interpolation filters because they are generally time-symmetric and slightly
+acausal.  Fortunately, as long as the latency introduced by the decimation and interpolation
+filters for any time slice $s$ is less than that time slice's delay $t^s$, the
+total latency of the \lloid\ algorithm will be zero.
+To be concrete, suppose that the first time slice, sampled at a rate $f^0 = 4096$~Hz, spans times $[t^0,\textrm{ }t^1) = [0\textrm{ s},\textrm{ }0.5\textrm{ s})$, and the second time slice, sampled at $f^1 = 512$~Hz, spans $[t^1,\textrm{ }t^2) = [0.5\textrm{ s},\textrm{ }4.5\textrm{ s})$.
+Then the second time slice's output, $\rho_i^1[k]$, will \emph{lead} the first time slice's output, $\rho_i^0[k]$, by 0.5~s.
+A decimation filter will be necessary to convert the 4096~Hz input signal $x[k] \equiv x^0[k]$ to the 512~Hz input $x^1[k]$, and an interpolation filter will be necessary to match the sample rates of the two early-warning outputs.
+In this example, as long as the decimation and interpolation filters are together acausal by less than $t^1 = 0.5$~s, the total SNR $\rho_i^0[k]$ will be available with a latency of zero samples.
+When zero latency is important, we may take this as a requirement for the decimation and interpolation filter kernels.
 
@@
 from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO\ noise
-curve~\citep{ALIGONoise}.  Templates are truncated at 10~Hz, where the projected
+power spectrum~\citep{ALIGONoise}\footnote{\url{http://dcc.ligo.org/cgi-bin/DocDB/ShowDocument?docid=T0900288&version=3}}.  Templates are truncated at 10~Hz, where the projected
 sensitivity of Advanced \LIGO\ is interrupted by the ``seismic wall.''
@@
 We have shown a prototype implementation of the \lloid\ algorithm using
-\gstreamer, an open source signal processing platform.  Work will be done in
-order to improve upon the latency of our implementation.  Ultimately the best
+\gstreamer, an open source signal processing platform.  Although our prototype
+already achieves latencies of less than one second, further fine tuning may
+reduce the latency even further.  Ultimately the best
 possible latency would be achieved by tighter integration between data
@@
 triggering, coincidence, and ranking of gravitational wave candidates in a
-framework that supports early \EM\ followup.  We will explore these and also
+framework that supports early \EM\ followup.  One might explore these and also
 using the time-slice decomposition and the \SVD\ to form low-latency
@@
 
-Our future work must more deeply address sky localization accuracy in a
+Future work must more deeply address sky localization accuracy in a
 realistic setting as well as observing strategies. Here, we have followed
@@
 a fellowship to CH. DK is supported from the Max Planck Gesellschaft.
+MAF is supported by NSF grant PHY-0855494.
 

### 2011-10-06 article.tex
Removed paragraphs describing GStreamer elements
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1377 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The \lloid\ algorithm is amenable to latency free, real-time implementation.
-However, such a real-time search pipeline would require integration directly into the data acquisition and storage systems of the \LIGO\ observatories.
+However, a real-time search pipeline would require integration directly into the data acquisition and storage systems of the \LIGO\ observatories.
 A slightly more
@@
 open-source signal processing environment called
-\gstreamer\footnote{\url{http://gstreamer.net/}}.
+\gstreamer\footnote{\url{http://gstreamer.net/}} (version 0.10.33).
 \gstreamer\ is a vital component of many Linux systems, providing media
@@
 buffers sufficiently small.
-
-Our prototype pipeline, tested with GStreamer version 0.10.33, consists of the
-following stages.
-
-\paragraph{Decimation}
-
-The whitened detector data is reduced to successively lower sample rates by
-decimation. Decimation involves applying an anti-aliasing filter to the data,
-and then down-sampling by deleting samples.  We use GStreamer's stock
-\texttt{audioresample} element, which provides a FIR decimation filter with
-a Kaiser-windowed sinc function kernel.
-
-The detector data is provided at every power-of-two sample rate
-required by the time slices described in \eqref{eq:time-slices}.  At this point,
-the \gstreamer\ pipeline splits into several parallel branches to handle each
-time slice.
-
-\paragraph{Basis filters}
-
-Next, these decimated data streams are fed into parallel \fir\ filter banks.
-The basis filters are implemented using a \gstlal\ element called
-\texttt{lal\_firbank}, which provides a single input, multiple output bank of
-\fir\ filters from a matrix of filter coefficients.  Several instances of
-\texttt{lal\_firbank} execute in parallel to handle each of the time slices.
-Filter coefficients are taken from the nonzero samples of the time slices in
-equation~\eqref{eq:time-sliced-templates}, with the zero-padding effectively
-accomplished by appropriate queues, or delay lines.
-
-\paragraph{Reconstruction}
-
-From the outputs of the \SVD\ basis filters, we form the partial \SNR\ streams
-for each time  slice by multiplying by the reconstruction matrices.  This is
-accomplished with the \texttt{gstlal} element \texttt{lal\_matrixmixer}.
-
-\paragraph{Interpolation}
-
-In order to form the early-warning output from each time slice, we have to add
-the partial \SNR\ to the early-warning output from the subsequent time slices.
-If the next time slice does not have the same sample rate, its output must first
-be interpolated.  The \gstreamer\ element \texttt{audioresample} is used again here.
-
-\paragraph{\SNR\ accumulation}
-
-The early-warning output for each time slice is formed by accumulating
-interpolated early-warning output from the subsequent time slice.  This process
-continues until we have worked our way to the \SNR\ of the original templates
-at the full sample rate $f^0$.  In this way, the \lloid\ algorithm and this
-implementation naturally leads to a simple early-warning pipeline.
+In any case, buffering is a reasonable strategy for low latency \LIGO\ data analysis because, as we previously remarked, the \LIGO\ data acquisition system has a granularity of $1/16$~s.
 
@@
 the impact of the length $N^\shortuparrow$ of the interpolation filter on
-mismatch.  The mismatch as a function of $N^\shortuparrow$ is shown in
+mismatch.  We use GStreamer's stock
+\texttt{audioresample} element, which provides a FIR decimation filter with
+a Kaiser-windowed sinc function kernel.  The mismatch as a function of $N^\shortuparrow$ is shown in
 Figure~\ref{fig:bw}b.  The mismatch saturates at $\sim$$2 \times 10^{-4}$ with

### 2011-10-06 article.tex
Addressed referee's point 12
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1378 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the full sample rate.  Together, the factor of 100 reduction in the number of
-filter coefficients and the factor of 100 reduction in the sample rate save a
+filter coefficients and the factor of 100 reduction in the sample rate during the early inspiral save a
 factor of $\sim$$10^4$ floating point operations per second (\flops) over the

### 2011-10-06 article.tex
Struck pre-detection figures from Table 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1379 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 merger, and area of 90\% confidence at selected times before merger for sources
-with expected detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
+with expected detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$.
+A dash (-----) signifies that the confidence area is omitted because at the indicated time the SNR would not have crossed the detection threshold of 8.}
 \begin{center}
@@
 \tableline
-40\phd\phn & 445 & 8.0 & 15500 & 3100 & 200 & 9.6 \\
-10\phd\phn & 280 & 12.7 & 6100 & 1200 & 78 & 3.8 \\
+40\phd\phn & 445 & 8.0 & ----- & ----- & ----- & 9.6 \\
+10\phd\phn & 280 & 12.7 & ----- & 1200 & 78 & 3.8 \\
 1\phd\phn & 130 & 27.4 & 1300 & 260 & 17 & 0.8 \\

### 2011-10-07 article.tex
Finished addressing referee's comment 6
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1380 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
-The most realistic estimates indicate that at an \SNR\ threshold of 8, we will
+The most realistic estimates indicate that at a single-detector \SNR\ threshold of 8, we will
 observe a total of 40~events~yr$^{-1}$. $\sim$10~yr$^{-1}$ will be detectable
@@
 detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$. The heavy dot indicates
-the time at which the accumulated \SNR\ exceeds a threshold of~8.}
+the time at which the accumulated \SNR\ exceeds a single-detector threshold of~8.}
 \end{figure}
@@
 \citep{Fairhurst2009}.  Table~\ref{fig:earlywarning} and
-Figure~\ref{fig:sky-localization-accuracy} show the estimated minimum 90\%
+Figure~\ref{fig:sky-localization-accuracy} show the estimated 90\%
 confidence area versus time of the loudest coalescence events detectable by
-Advanced \LIGO\ and Advanced Virgo.  Once per year, we expect to observe an
-event with a final \SNR\ of $\approx$27 whose location can be constrained to about
+Advanced \LIGO\ and Advanced Virgo.  This is the \emph{minimum} area; localization
+is best at high elevation from the plane containing the detectors and worst at zero elevation.
+As \citeauthor{Fairhurst2009} cautions, his formula also slightly \emph{overestimates} the minimum localization uncertainty for low SNR events because it makes an approximation that the timing errors are Gaussian distributed.
+Nonetheless, given the large uncertainties in rates, these approximations suffice for an order of magnitude estimate.
+Once per year, we expect to observe an
+event with a final single-detector \SNR\ of $\approx$27 whose location can be constrained to about
 1300~deg$^2$ (3.1\% of the sky) within 25~s of merger,

### 2011-10-07 article.tex
Incremented DCC version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1383 ccdd05d5-56c1-461a-bf98-85f6876523b4
Struck sentence on third generation detectors
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1384 ccdd05d5-56c1-461a-bf98-85f6876523b4
Addressed referee's point 5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1385 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 sources that could be detectable by Advanced \LIGO\ a given number of seconds
-before coalescence.  The heavy solid line is the most realistic yearly rate
-estimate.  The shaded region represents the 5\% to 95\% confidence interval
-arising from uncertainty in predicted event rates \citep{Abadie:2010p10836}.}
+before coalescence.  The heavy solid line corresponds to the most probable yearly rate
+estimate from~\citet{Abadie:2010p10836}.  The shaded region represents the 5\% to 95\% confidence interval
+arising from substantial uncertainty in predicted event rates.}
 \end{figure}
@@
 Figure~\ref{fig:earlywarning} shows projected early detectability rates for
-\NS--\NS\ binaries in Advanced \LIGO\ assuming event rates
-predicted in \citet{Abadie:2010p10836} and anticipated detector sensitivity for
-the `zero detuning, high power' configuration described in \citet{ALIGONoise}.
-The most realistic estimates indicate that at a single-detector \SNR\ threshold of 8, we will
-observe a total of 40~events~yr$^{-1}$. $\sim$10~yr$^{-1}$ will be detectable
+\NS--\NS\ binaries in Advanced \LIGO\ assuming the anticipated detector sensitivity for
+the `zero detuning, high power' configuration described in \citet{ALIGONoise} and
+\NS--\NS\ merger rates estimated in \citet{Abadie:2010p10836}.  The merger rates
+have substantial measurement uncertainty due to the small sample of known double pulsar systems that will merge within a Hubble time; they also have systematic uncertainty due sensitive dependence on the pulsar luminosity distribution function~\citep{KalogeraRates}.
+The most probable estimates indicate that at a single-detector \SNR\ threshold of 8, we will
+observe a total of 40~events~yr$^{-1}$; $\sim$10~yr$^{-1}$ will be detectable
 within 10~s of merger and $\sim$5~yr$^{-1}$ will be detectable within 25~s of
-merger if analysis can proceed with near zero latency. These rates are shown
-with their substantial uncertainties in Figure~\ref{fig:earlywarning}.
+merger if analysis can proceed with near zero latency.
 
@@
 less of a concern.  This should reduce the 90\% confidence area substantially.
-Proposed third generation \GW\ detectors would improve the \SNR\ and likewise
-sky localization.
 
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v28}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v29}.
 

### 2011-10-08 article.tex
Addressed referee's comment 3
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1389 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 preferred method in \GW\ searches.  The most obvious example of the latter is
-time domain (\TD) convolution, which is latency-free.  However, its cost in floating
+direct time domain (\TD) convolution, which is latency-free.  However, its cost in floating
 point operations per second is linear in the length of the templates, so it is
@@
 Fortunately, the morphology of inspiral signals can be exploited to offset some
-of the computational complexity of low-latency algorithms.  First, the signals
+of the computational complexity of known low-latency algorithms.  First, the signals
 evolve slowly in frequency, so that they can be broken into contiguous
@@
 that competes in terms of floating point operations per second with the
-conventional \FD\ method, which by contrast requires a significant latency due
+conventional overlap-save \FD\ method, which by contrast requires a significant latency due
 to the inherent acausality of the Fourier transform.  Our method, called \lloid\
@@
 In the next section we compute the expected computational cost scaling of this
-decomposition and compare it with the brute-force \TD\ implementation of
-\eqref{eq:SNRTD} and higher latency \FD\ methods.
+decomposition and compare it with the direct \TD\ implementation of
+\eqref{eq:SNRTD} and higher latency blockwise \FD\ methods.
 
@@
 
-The conventional \TD\ method consists of a bank of \fir\ filters, or
+The conventional, direct \TD\ method consists of a bank of \fir\ filters, or
 sliding-window dot products.  If there are $\numtmps$ templates, each
@@
 
-The most common \FD\ method is known as the \emph{overlap-save} algorithm, described in
+The most common \FD\ method is known as the overlap-save algorithm, described in
 \citet{numerical-recipes-chapter-13}.  It entails splitting the input into blocks of $D$
@@
 evolves slowly enough in time, then we can use the time-frequency relationship
-of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional
+of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional, direct
 \TD\ method.  The reduction in \flops\ is approximately
@@
 for our chosen set of source parameters, we then compare the computational cost
-in \flops\ for \TD\ method, the \FD\ method, and \lloid.
+in \flops\ for the direct \TD\ method, the overlap-save \FD\ method, and \lloid.
 
@@
 Table~\ref{table:flops} shows the computational cost in \flops\ for the sub-bank
-we described above.  For the \FD\ method, an \fft\ block size of
+we described above.  For the overlap-save \FD\ method, an \fft\ block size of
 $\fftblock = 2 \tmpsamps$ is assumed, resulting in a latency of
 $\left(\tmpsamps / f^0\right)$ seconds.  Both the \FD\ method and \lloid\ are
-five orders of magnitude faster than the conventional \TD\ method.  However, the
+five orders of magnitude faster than the conventional, direct \TD\ method.  However, the
 \FD\ method has a latency of over half of an hour, whereas the \lloid\ algorithm
@@
 \caption{\label{table:flops}Computational cost in \flops\ and latency in seconds
-of the \TD\ method, the \FD\ method, and \lloid.  Cost is given
+of the direct \TD\ method, the overlap-save \FD\ method, and \lloid.  Cost is given
 for both the sub-bank described in section~\ref{sec:bank-setup} and a full
@@
 \tableline
-time domain & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$3.8\times10^5$ \\
-frequency domain & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$5.9$ \\
+direct (\TD) & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$3.8\times10^5$ \\
+overlap-save (\FD) & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$5.9$ \\
 \lloid\ (theory) & $6.6\times10^8$ & 0 & $1.1 \times 10^{11}$ & $\sim$$11$ \\
@@
 
-By comparison, using the \TD\ method to achieve the same latency costs
+By comparison, using the conventional \TD\ method to achieve the same latency costs
 $\sim$$10^{13}$ \flops\ for this particular sub-bank, and so it would require
@@
 
-The \lloid\ method is as fast as standard \fft\ convolutions but allows for
+The \lloid\ method is as fast as conventional \fft-based, \FD\ convolution but allows for
 latency free, real-time operation.  We anticipate requiring $\gtrsim$40

### 2011-10-12 article.tex
Addressed referee's comment 4
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1392 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 localization accuracy for pre-merger detection.
+
+In the quadrupole approximation, the instantaneous frequency of the GW inspiral signal is related to the time $t$ relative to coalescence ~\citep{kidder1992, findchirppaper} through
+%
+\begin{equation} \label{eq:fgw}
+	f(t) = \frac{1}{\pi \mathcal{M}}
+		\left[ \frac{5}{256}\frac{\mathcal{M}}{t} \right]^{3/8}
+\end{equation}
+%
+where $\mathcal{M} = M^{2/5} \mu^{3/5}$ is the chirp mass of the binary in units of time ($G
+M_\odot / c^3 \approx 5$~$\upmu$s), $M$ is the total mass, and $\mu$ is the reduced mass. 
+The expected value of the single detector \SNR\ for an optimally oriented (source at detector's zenith or nadir, orbital plane face-on) inspiral source is~\citep{Abadie:2010p10836}
+%
+\begin{equation}
+	\label{eq:expected-snr}
+	\rho =
+		\frac{\mathcal{M}^{5/6} c}{\pi^{2/3} D}
+		\sqrt{
+			\frac{5}{6} \int_{f_\mathrm{low}}^{f_\mathrm{high}}
+			\frac{f^{-7/3}}{S(f)} \mathrm{d}f}
+\end{equation}
+%
+where $D$ is the luminosity distance and $S(f)$ is the power spectral density of the detector noise.
+$f_\mathrm{low}$ and $f_\mathrm{high}$ are low and high frequency limits of integration which may be chosen to extend across the entire bandwidth of the detector.
+If we want to trigger at a time $t$ before merger, then we must cut off the SNR integration at $f_\mathrm{high} = f(t)$ with $f(t)$ given by equation~(\ref{eq:fgw}) above.
+
 Figure~\ref{fig:earlywarning} shows projected early detectability rates for
@@
 For concreteness and simplicity, consider an inspiral waveform in the
-quadrapole approximation, for which the time-frequency relation~\citep{kidder1992, findchirppaper} is
-%
-\begin{equation} \label{eq:fgw}
-%
-f(t) = \frac{1}{\pi \mathcal{M}} \left[ \frac{5}{256}\frac{\mathcal{M}}{t}
-\right]^{3/8}.
-%
-\end{equation}
-%
-Here, $\mathcal{M}$ is the chirp mass of the binary in units of time (where $G
-M_\odot / c^3 \approx 5$~$\upmu$s) and $t$ is the time relative to the
-coalescence of the binary.
+quadrapole approximation, for which the time-frequency relation is given by equation~(\ref{eq:fgw}).
 This monotonic time-frequency relationship allows us

### 2011-10-12 article.tex
Completely addressed point 4
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1393 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\DeclareMathOperator{\expectation}{E}
+\DeclareMathOperator{\var}{var}
+
 % Macros for collapsing sizes of things
@@
 
+Searches for inspiral signals typically employ matched filter
+banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.
+Suppose that the observed data $x[k]$ consists of a known, nominal signal $s[k]$ and additive, zero-mean noise $n[k]$,
+%
+$$
+	x[k] = s[k] + n[k].
+$$
+%
+A matched filter is a linear filter, defined as
+%
+$$
+	y[k] = \sum_{n=0}^{N-1} h[n] x[k-n] = y_s[k] + y_n[k],
+$$
+%
+where $y_s$ is the response of the filter to the signal alone and $y_n$ is the response of the signal to noise alone.
+The matched filter's coefficients maximize the ratio of the expectation of the filter's instantaneous response to the variance in the filter's output at all times:
+%
+$$
+\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]}{\var \left[ y[k] \right]} = \frac{y_s[h]}{\var \left[ y_n[k] \right]}.
+$$
+%
+It is well known~\citep[see, for example,][]{matched-filter} that if $n[k]$ is Gaussian and wide-sense stationary, then the optimum is obtained when
+%
+$$
+\tilde{h}[n] = \tilde{s}^*[n] \tilde S^{-1}[n],
+$$
+%
+up to an arbitrary multiplicative constant.
+$\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms of $h[k]$, $s[k]$, and $x[k]$, respectively, here $\tilde{S}[n]$ is the power spectral density of $n[k]$.
+The discrete Fourier transform of the output is
+%
+\begin{multline}
+\label{eq:matched-filter-fd}
+\tilde{y}[n] = \tilde{s}^*[n] \tilde{S}^{-1}[n] \tilde{x}[n] \\
+\equiv \left(\tilde{S}^{-1/2}[n] \tilde{s}[n]\right)^* \left(\tilde{S}^{-1/2}[n] \tilde{x}[n] \right).
+\end{multline}
+%
+The placement of parentheses in equation~(\ref{eq:matched-filter-fd}) emphasizes that the matched filter can be thought of as a cross-correlation of a whitened version of the data with a whitened version of the nominal signal.
+In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.
+
 Inspiral signals are continuously parameterized by a set of intrinsic source
@@
 
-Searches for inspiral signals typically employ matched filter
-banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.
 The coefficients for the $\numtmps$ filters are known as templates, 
@@
 $x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the
-\SNR, $\rho_i[k]$ defined as
+\SNR, $\rho_i[k]$, defined as
 %

### 2011-10-12 article.tex
Corrected comments related to Fisher matrix
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1395 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 confidence area versus time of the loudest coalescence events detectable by
-Advanced \LIGO\ and Advanced Virgo.  This is the \emph{minimum} area; localization
-is best at high elevation from the plane containing the detectors and worst at zero elevation.
-As \citeauthor{Fairhurst2009} cautions, his formula also slightly \emph{overestimates} the minimum localization uncertainty for low SNR events because it makes an approximation that the timing errors are Gaussian distributed.
-Nonetheless, given the large uncertainties in rates, these approximations suffice for an order of magnitude estimate.
+Advanced \LIGO\ and Advanced Virgo.  This is the \emph{minimum} area; localization is best at high elevation
+from the plane containing the detectors and worst at zero elevation. As
+\citeauthor{Fairhurst2009} cautions, his formula also slightly
+\emph{overestimates} the minimum area for low SNR events because it
+makes an approximation that the timing errors are Gaussian distributed
+and that the sky location estimates are Gaussian distributed about the
+true location.  Nonetheless, given the large uncertainties in rates,
+\citeauthor{Fairhurst2009}'s approximations suffice for an order of
+magnitude estimate.
 Once per year, we expect to observe an

### 2011-10-12 article.tex
Added definition of whitened templates
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1396 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 up to an arbitrary multiplicative constant.
-$\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms of $h[k]$, $s[k]$, and $x[k]$, respectively, here $\tilde{S}[n]$ is the power spectral density of $n[k]$.
+$\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively, here $\tilde{S}[n]$ is the power spectral density of $n[k]$.
 The discrete Fourier transform of the output is
@@
 In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.
+Correspondingly, from this point on we shall use $h[k]$ to describe the \emph{whitened} templates, being the inverse DFT of $\left(\tilde{S}^{-1/2}[n] \tilde{s}[n]\right)^*$.
 

### 2011-10-13 article.tex
Cited Gardner (2010) and reworded reference to MBTA
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1399 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The first step of our proposed method is to divide the templates into time
-slices in a \TD\ analogue to the \FD\ decomposition described
-in \citet{Marion2004} and \citet{Buskulic2010}.  We decompose each template
+slices in a \TD\ analogue to the \FD\ decomposition employed by MBTA~\citep{Marion2004,Buskulic2010}.
+The application to GW data analysis is foreshadowed by an earlier \FD\ convolution algorithm based on splitting the impulse response of a filter into smaller blocks, proposed by \citet{gardner1995efficient}.
+We decompose each template
 $h_{i}[k]$ into a sum of $S$ non-overlapping templates

### 2011-10-13 article.tex
Corrected description of MBTA
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1400 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 S6/VSR3.
-MBTA consists of two \FD\ template banks for two different frequency bands, one which is matched to the early inspiral and the other which is matched to the late inspiral.
-Time coincidences are sought between triggers in both template banks.
+MBTA consists of multiple, usually two, template banks for different frequency bands, one which is matched to the early inspiral and the other which is matched to the late inspiral.
+An excursion in the output of any filter bank triggers coherent reconstruction of the full matched filtered output.
+Final triggers are built from the reconstructed matched filter output.
 Another novel approach using networks of parallel, second order IIR filters is being explored

### 2011-10-13 article.tex
Made section 5.4's relationshipto table 4 more obvious
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1401 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $\tmpsamps$ samples in length, then each filter requires $M N$ multiplications
-and additions per sample, or $2 \numtmps \tmpsamps f^0$ \flops\ at a sample rate
-$f^0$.
+and additions per sample, or, at a sample rate $f^0$,
+%
+\begin{equation}
+	\label{eq:td-flops}
+	2 \numtmps \tmpsamps f^0 \textrm{ \flops}.
+\end{equation}
 
@@
 $\fftblock - \tmpsamps$ usable samples of output, the overlap-save method requires
-$$
-f^0 \cdot \frac{2 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock} \; \mathrm{\flops} \,.
-$$
+%
+\begin{equation}
+	\label{eq:fd-flops}
+	f^0 \cdot \frac{2 (\numtmps + 1) \lg \fftblock + 2 \numtmps}{1 - \tmpsamps/\fftblock} \textrm{ \flops}.
+\end{equation}
 
@@
 scaling, an entire single detector search for \NS--\NS\ signals in the
-1--3~$M_\odot$ component mass range could be implemented with
-$\sim$$10^{11}$~\flops.  Modern (ca. 2011) workstations can achieve peak
-computation rates up to $\sim$$10^{11}$~\flops.  In practice, we expect that a
-software implementation of \lloid\ will reach average computation rates that are
-perhaps a factor 10 less than this, $\sim$$10^{10}$~\flops\ per machine, due to
-non-floating point tasks including bookkeeping and thread synchronization.
-Given these considerations, we estimate that a full single-detector Advanced
-\LIGO\ search with \lloid\ will require about $\sim$$10$ machines.
+1--3~$M_\odot$ component mass range could be implemented with $(98544/657)\approx150$ times the cost, or $9.9 \times 10^{10}$~\flops.
 
@@
 Despite these subtleties, this estimate gave us $1.1 \times 10^{11}$~\flops,
-agreeing with the simple scaling argument above.  Therefore we believe that
-$\sim$$10$ machines is a realistic estimate of our computing requirements.
+agreeing with the simple scaling argument above.
+
+Modern (ca. 2011) workstations can achieve peak
+computation rates up to $\sim$$10^{11}$~\flops.  In practice, we expect that a
+software implementation of \lloid\ will reach average computation rates that are
+perhaps a factor 10 less than this, $\sim$$10^{10}$~\flops\ per machine, due to
+non-floating point tasks including bookkeeping and thread synchronization.
+Given these considerations, we estimate that a full single-detector Advanced
+\LIGO\ search with \lloid\ will require $\sim$$10$ machines.
 
 By comparison, using the conventional \TD\ method to achieve the same latency costs
-$\sim$$10^{13}$ \flops\ for this particular sub-bank, and so it would require
-a total of $\gtrsim$$10^{15}$ \flops\ or $\gtrsim$$10^5$ present-day machines
-to search the full parameter space.  Presently, the \LIGO\ Data Grid%
+$4.9 \times 10^{13}$~\flops\ for this particular sub-bank, and so simply scaling up by the factor of $150$ suggests that it would require $7.4 \times 10^{15}$~\flops\
+to search the full parameter space.
+To account for the varying sample rate and template duration across the parameter space, we can also directly calculate the cost for the full \TD\ method search using expression~(\ref{eq:td-flops}), resulting in $3.8 \times 10^{15}$~\flops, agreeing within an order of magnitude.
+This would require~$\gtrsim$$10^5$ current-day machines.
+Presently, the \LIGO\ Data Grid%
 \footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}} consists of
 only $\sim$$10^4$ machines, so direct \TD\ convolution is clearly impractical.
+
+The overlap-save \FD\ method is slightly more efficient than LLOID for this particular sub-bank, requiring $5.2 \times 10^8$~\flops.
+The scaling argument projects that a full \FD\ search would require $7.8 \times 10^{10}$~\flops.
+The direct calculation from expression~(\ref{eq:fd-flops}) gives $5.9 \times 10^{10}$~\flops, in order-of-magnitude agreement.
+It appears that in this application the conventional \FD\ search is scarcely a factor of 2 more computationally efficient than LLOID, but only at the price of thousands of seconds of latency.
 

### 2011-10-13 article.tex
Added some spacing to equations
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1402 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $$
-	y[k] = \sum_{n=0}^{N-1} h[n] x[k-n] = y_s[k] + y_n[k],
+	y[k] = \sum_{n=0}^{N-1} h[n] \, x[k-n] = y_s[k] + y_n[k],
 $$
@@
 $$
-\tilde{h}[n] = \tilde{s}^*[n] \tilde S^{-1}[n],
+\tilde{h}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n],
 $$
@@
 \label{eq:matched-filter-fd}
-\tilde{y}[n] = \tilde{s}^*[n] \tilde{S}^{-1}[n] \tilde{x}[n] \\
-\equiv \left(\tilde{S}^{-1/2}[n] \tilde{s}[n]\right)^* \left(\tilde{S}^{-1/2}[n] \tilde{x}[n] \right).
+\tilde{y}[n] = \tilde{s}^*[n] \, \tilde{S}^{-1}[n] \, \tilde{x}[n] \\
+\equiv \left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^* \left(\tilde{S}^{-1/2}[n] \, \tilde{x}[n] \right).
 \end{multline}
@@
 In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.
-Correspondingly, from this point on we shall use $h[k]$ to describe the \emph{whitened} templates, being the inverse DFT of $\left(\tilde{S}^{-1/2}[n] \tilde{s}[n]\right)^*$.
+Correspondingly, from this point on we shall use $h[k]$ to describe the \emph{whitened} templates, being the inverse DFT of $\left(\tilde{S}^{-1/2}[n] \, \tilde{s}[n]\right)^*$.
 
@@
 	\label{eq:SNRTD}
-	\rho_i [k] = \sum_{n=0}^{N-1} h_{i}[n] x [k-n].
+	\rho_i [k] = \sum_{n=0}^{N-1} h_{i}[n] \, x [k-n].
 \end{equation}

### 2011-10-14 article.tex
Addressed point 19
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1404 ccdd05d5-56c1-461a-bf98-85f6876523b4
Addressed point 2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1405 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 telescopes.  The typical latencies were 30 minutes \citep{HugheyGWPAW2011}.
-This was an important achievement, but too late to catch any prompt \EM\
-emission.  Since the \GW\ signal is in principle detectable even before the tidal
+This was an important achievement, but too late to catch any prompt optical flash
+or the onset of an on-axis optical afterglow.  Since the \GW\ signal is in principle detectable even before the tidal
 disruption, one might have the ambition of reporting \GW\ candidates not minutes
@@
 \GRB{}, allowing multi-wavelength observations of prompt emission. More
-frequently, low-latency alerts will be released after merger but will still
-yield extended X-ray tails and early afterglows.
+frequently, low-latency alerts will be released after merger but may still
+yield extended X-ray tails and early on-axis afterglows.
 

### 2011-10-15 article.tex
Addressed comment 9
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1406 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 It is unfeasible to search hundreds of square degrees for a prompt counterpart.
-For reference, LSST should see first light around the same time as
-Advanced \LIGO\ and will have an unparalleled field of view of 9.6~deg$^2$
-\citep{2008arXiv0805.2366I}.  However, 
+For comparison to some examples of modern ground-based wide-field survey instruments, the Palomar Transient Factory P48 ~\citep{2010SPIE.7735E.122L} has a $3.50 \times 2.31$~deg$^2$ field of view~(FOV); the Pan-STARRS P1~\citep{2002SPIE.4836..154K} has a 7~deg$^2$ FOV.
+Even the eagerly awaited LSST will have a FOV of 9.6~deg$^2$ \citep{2008arXiv0805.2366I}.
+However,
 it is possible to reduce the localization uncertainty by only looking at

### 2011-10-18 article.tex
Minor edit to section 3.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1407 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 where $y_s$ is the response of the filter to the signal alone and $y_n$ is the response of the signal to noise alone.
-The matched filter's coefficients maximize the ratio of the expectation of the filter's instantaneous response to the variance in the filter's output at all times:
+The matched filter's coefficients maximize the ratio of the expectation of the filter's instantaneous response to the variance in the filter's output:
 %
 $$
-\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]}{\var \left[ y[k] \right]} = \frac{y_s[h]}{\var \left[ y_n[k] \right]}.
+\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]}{\var \left[ y[k] \right]} = \frac{y_s[0]}{\var \left[ y_n[k] \right]}.
 $$

### 2011-10-18 article.tex
Another minor edit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1408 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 transformations combined reduce the number of floating point operations
-to the level that is competitive with the conventional high-latency \FD\
+to a level that is competitive with the conventional high-latency \FD\
 matched filter approach.  In the remainder of this section we describe the

### 2011-10-18 article.tex
Another minor edit
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1409 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 equation~(\ref{eq:svddecomp}), we define the early-warning output accumulated
-up to time slice $s$ using the recursion relation,
+up to time slice $s$ using the recurrence relation,
 %

### 2011-10-18 article.tex
More specific description of study in introduction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1411 ccdd05d5-56c1-461a-bf98-85f6876523b4
Shuffled around introduction
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1413 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 point operations per second is linear in the length of the templates, so it is
-prohibitively expensive for long templates.  Several efforts are under way to
+prohibitively expensive for long templates.
+
+Fortunately, the morphology of inspiral signals can be exploited to offset some
+of the computational complexity of known low-latency algorithms.  First, the signals
+evolve slowly in frequency, so that they can be broken into contiguous
+band-limited time intervals and processed at possibly lower sample rates.
+Second, inspiral filter banks consist of highly similar templates, admitting
+methods such as the singular value decomposition (\SVD)~\citep{Cannon:2010p10398}
+or the Gram-Schmidt process~\citep{rbf} to reduce the number of templates.
+
+Several efforts that exploit one or both of these properties are under way to
 develop low-latency \CBC\ search pipelines with tractable computing requirements.
@@
 
-Fortunately, the morphology of inspiral signals can be exploited to offset some
-of the computational complexity of known low-latency algorithms.  First, the signals
-evolve slowly in frequency, so that they can be broken into contiguous
-band-limited time intervals and processed at possibly lower sample rates.
-Second, inspiral filter banks consist of highly similar templates, admitting
-methods such as the singular value decomposition (\SVD)~\citep{Cannon:2010p10398}
-or the Gram-Schmidt process~\citep{rbf} to reduce the number of templates.  We
-will use both aspects to demonstrate that a very low latency detection statistic
+We will use both properties to demonstrate that a very low latency detection statistic
 is possible with current computing resources.  Assuming the other technical
@@
 
-The paper is organized as follows.  First, we discuss prospects for early-warning
-detection.  Then, we provide an overview of our novel method for detecting compact
-binary coalescence signals near real-time. We then describe a prototype
-implementation using open source signal processing software.  To validate our approach
-we present results of simulations.  We conclude with some remarks on what remains to
-prepare for the advanced detector era.
+The paper is organized as follows.  First, we discuss prospects for
+early-warning detection.  Then, we provide an overview of our novel
+method for detecting compact binary coalescence signals near real-time.
+We then describe a prototype implementation using open source signal
+processing software.  To validate our approach we present a case study
+focusing on a particular subset of the \NS--\NS\ parameter space.  We
+conclude with some remarks on what remains to prepare for the advanced
+detector era.
 

### 2011-11-15 article.tex
Addressed minor comment from reviewer
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1417 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 up to an arbitrary multiplicative constant.
-$\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively, here $\tilde{S}[n]$ is the power spectral density of $n[k]$.
+Here, $\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively; $\tilde{S}[n]$ is the power spectral density of $n[k]$.
 The discrete Fourier transform of the output is

### 2011-11-30 article.tex
Added appendix and some text referring to it in section 1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1418 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
+\usepackage{lscape}
 \usepackage[colorlinks]{hyperref}
@@
 prohibitively expensive for long templates.
+The computational challenges of low-latency \CBC\ searches are still more daunting for advanced detectors for which the inspiral signal remains in band for a large fraction of an hour~(see appendix~\ref{sec:low-frequency-cutoff}).
 
@@
 
+\appendix
+
+\section{\label{sec:low-frequency-cutoff}Low frequency cutoff for inspiral searches}
+
+Ground-based gravitational wave detectors are unavoidably affected at low frequencies by seismic and anthropogenic ground motion.
+The LIGO test masses are suspended from multiple-stage pendula, which attenuate ground motion down to the pole frequency.
+In the detector configuration in place during S6, seismic noise dominated the instrumental background below about 40~Hz.
+Considerable effort is being invested in improving seismic attenuation in Advanced LIGO using active and passive isolation~\citep{0264-9381-27-8-084006}, so that suspension thermal noise will dominate down to 10 to 15~Hz.
+Inspiral waveforms are chirps of infinite duration, but since an interferometric detector's noise diverges at this so-called `seismic wall,' templates for matched filter searches are truncated at a low frequency cutoff $f_\mathrm{low}$ in order to save computational overhead with negligible loss of SNR.
+
+The expected matched-filter SNR, integrated from $f_\mathrm{low}$ to $f_\mathrm{high}$, is given by equation~(\ref{eq:expected-snr}).
+The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the ISCO; for non-spinning systems, $f_\mathrm{ISCO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary.
+The choice of $f_\mathrm{low}$ is based on the fraction of the total SNR that is accumulated at frequencies above $f_\mathrm{low}$.
+To illustrate the relative contributions to the SNR at different frequencies for a (1.4,~1.4)~$\Msun$ binary, we normalized and plotted the integrand of equation~(\ref{eq:expected-snr}), the noise-weighted power spectral density of the inspiral waveform, in figure~\ref{fig:low-frequency-cutoff}b.
+This is the quantity
+%
+$$
+	\frac{1}{\rho^2}\frac{\mathrm{d}\rho^2}{\mathrm{d}f} = \frac{f^{-7/3}}{S(f)} \left( \int_0^{f_\mathrm{ISCO}} \frac{{f'}^{-7/3}}{S(f')} \, \mathrm{d}f' \right)^{-1},
+$$
+%
+which is normalized by the total SNR squared in order to put detectors with different absolute sensitivities on the same footing.
+We used several different noise power spectra: all of the envisioned Advanced LIGO configurations from~\citet{ALIGONoise}; the best achieved sensitivity at LIGO Hanford Observatory (LHO) in LIGO's fifth science run (S5), measured by \citet{S5InspiralRange}; and the best achieved sensitivity at LHO during S6, measured by \citet{S6InspiralRange}.
+(The noise spectra themselves are shown in figure~\ref{fig:low-frequency-cutoff}a.)  It is plain that during S5 and S6 the greatest contribution to the SNR was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the SNR is accumulated below 60~Hz.
+This information is presented in a complementary way in figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, interpreted as a fraction of the total ``available'' SNR,
+%
+$$
+	\rho_\mathrm{frac}(f_\mathrm{low}) = \sqrt{\left( \int_{f_\mathrm{low}}^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right) \left( \int_0^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right)^{-1}}.
+$$
+%
+Table~\ref{table:accum_snr} shows the fractional accumulated SNR for four selected low frequency cutoffs, 40~Hz, 30~Hz, 20~Hz, and 10~Hz.
+In S5 and S6, all of the SNR is accumulated above 40~Hz.
+For the `high frequency' Advanced LIGO configuration, almost 50\% of the SNR is accumulated \emph{below} 40~Hz.
+For the preferred final configuration, 'zero detuning, high power,' 86.1\% of the SNR is below 40~Hz, 93.2\% is below 30~Hz, and 98.1\% is below 20~Hz.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{rcccc}
+\hline\hline
+Noise model & 40 Hz & 30 Hz & 20 Hz & 10 Hz \\
+\hline
+LHO (best S5) & 100.0 & 100.0 & 100.0 & 100.0 \\
+LHO (best S6) & 100.0 & 100.0 & 100.0 & 100.0 \\
+high frequency & 53.3 & 80.1 & 97.6 & 100.0 \\
+no SRM & 87.8 & 95.1 & 98.7 & 100.0 \\
+BHBH 20$^\circ$ & 71.1 & 84.2 & 96.2 & 100.0 \\
+NSNS optimized & 91.5 & 96.3 & 99.0 & 100.0 \\
+zero detuning, low power & 67.9 & 80.0 & 93.5 & 100.0 \\
+zero detuning, high power & 86.1 & 93.2 & 98.1 & 100.0 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{table:accum_snr}Fractional accumulated SNR $\rho_\mathrm{frac}(f_\mathrm{low})$ for four selected low frequency cutoffs, $f_\mathrm{low}=40$~Hz, 30~Hz, 20~Hz, and 10~Hz.}
+\end{table}
+
+Since the gravitational wave amplitude is inversely proportional to the luminosity distance of the source, and the sensitive volume is proportional to distance cubed, the rate of detectable coalescences depends on the choice of low frequency cutoff.
+An inspiral search that is designed with a low frequency cutoff at the seismic wall would gain an increase in detection rate of $\rho_\mathrm{frac}^{-3}(f_\mathrm{low})$ relative to a search with a low frequency cutoff of $f_\mathrm{low}$.
+For a fractional accumulated SNR of 80\%, this would represent almost a twofold increase in the rate of detection.
+For $\rho_\mathrm{frac} = 90\%$, the detection rate would still be increased by as much as 37\%.
+Existing coalescing binary detection pipelines strive to sacrifice no more than 3\% of the available SNR; this forfeits less than a 10\% gain in detection rate.
+In order to satisfy this constraint, the low frequency cutoff would have to be placed below 20~Hz for all of the conceived Advanced LIGO configurations.
+
+The instantaneous gravitational wave frequency, given by equation~(\ref{eq:fgw}) is a power law function of time, so the amount of time for the gravitational wave frequency to evolve from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$ depends strongly on $f_\mathrm{low}$.
+The duration of a (1.4,~1.4)~$\Msun$ inspiral is show in figure~\ref{fig:low-frequency-cutoff}d.
+The inspiral takes only 25~s to evolve from 40~Hz to $f_\mathrm{ISCO}$, but takes 54~s to evolve from 30~Hz to $f_\mathrm{ISCO}$, 158~s from 20~Hz, and 1002~s from 10~Hz.
+
+\begin{figure}[b]
+	\includegraphics{f6a}%
+	\includegraphics{f6b}
+
+	\includegraphics{f6c}%
+	\includegraphics{f6d}
+
+	\includegraphics{f6e}
+
+	\caption{\label{fig:low-frequency-cutoff}Counterclockwise from top left:  (a)~Noise amplitude spectral density for a variety of Advanced LIGO noise models, S5, and S6.
+(b)~Normalized signal to noise per unit frequency, $(\mathrm{d}\rho^2/\mathrm{d}f)/\rho^2$, for a (1.4,~1.4)~$\Msun$ inspiral.
+(c)~Percentage of SNR that is accumulated from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, relative to SNR accumulated from $f_\mathrm{low} = 0\,\mathrm{Hz}$ to $f_\mathrm{ISCO}$.
+(d)~Amount of time for a \NS--\NS\ inspiral signal to evolve from frequency $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, as a function of $f_\mathrm{low}$.
+For figures (a) through (c), the line style indicates which noise model was used.}
+\end{figure}
+
 \bibliographystyle{apj}

### 2011-11-30 noisemodels.py

@@ -1,0 +1,37 @@
+"""
+Load power spectra for various LIGO noise estimates and measurements.
+"""
+__author__ = "Leo Singer <leo.singer@ligo.org>"
+__all__ = ('psds',)
+
+from numpy import sqrt, loadtxt
+import os.path
+
+asd_paths = (
+	('LHO (best S5)', 'data/max_H1-TMPLTBANK-871198828-2048.strainspec_asd_v4.txt'),
+	('LHO (best S6)', 'data/H1-TMPLTBANK-962268343-2048.strainspec.txt'),
+	('high freq', 'data/T0900288/High_Freq.txt'),
+	('no SRM', 'data/T0900288/NO_SRM.txt'),
+	('BHBH 20deg', 'data/T0900288/BHBH_20deg.txt'),
+	('NSNS optimized', 'data/T0900288/NSNS_Opt.txt'),
+	('zero det., low power', 'data/T0900288/ZERO_DET_low_P.txt'),
+	('zero det., high power', 'data/T0900288/ZERO_DET_high_P.txt'),
+)
+
+plotkwargs = {
+	'LHO (best S5)': {'color': '0.5', 'linestyle': '-', 'linewidth': 0.5},
+	'LHO (best S6)': {'color': 'k', 'linestyle': '-', 'linewidth': 0.5},
+	'high freq': {'color': 'k', 'linestyle': ':'},
+	'no SRM': {'color': 'k', 'linestyle': '-.'},
+	'NSNS optimized': {'color': 'k', 'linestyle': '--'},
+	'BHBH 20deg': {'color': '0.5', 'linestyle': '--'},
+	'zero det., low power': {'color': '0.5', 'linestyle': '-'},
+	'zero det., high power': {'color': 'k', 'linestyle': '-'}
+}
+
+asds = tuple( (key, loadtxt(os.path.join(os.path.dirname(__file__), value))) for key, value in asd_paths )
+
+# "LHO (best S6)" actually stores PSD, not ASD.
+for name, data in asds:
+	if name == 'LHO (best S6)':
+		data[:, 1] = sqrt(data[:, 1])

### 2011-11-30 plot_accum_snr.py

@@ -1,0 +1,75 @@
+#!/usr/bin/env python
+"""
+Plot fractional accumulated SNR versus low frequency cutoff for all noise models.
+Also generate tex table and print to stdout.
+"""
+
+import sys
+from noisemodels import asds, plotkwargs
+from pylab import *
+
+# Component mass 1 in M_sun
+m1 = 1.4
+
+# Component mass 2 in M_sun
+m2 = 1.4
+
+# Total mass in M_sun
+M = m1 + m2
+
+# ISCO frequency in Hz
+fISCO = 4400 / M
+
+fig_width = 3.35
+fig_height = 2.75
+fig = figure(figsize=(fig_width,fig_height))
+
+f_lows = range(40, 0, -10)
+
+print r'\begin{tabular}{r' + 'c'*len(f_lows) + '}'
+print r'\hline\hline'
+print 'Noise model',
+for f_low in f_lows:
+	print '&', f_low, 'Hz',
+print r'\\'
+print '\hline'
+
+for i, (name, data) in enumerate(asds):
+
+	# Unpack frequency and amplitue spectral density from array
+	f, s = data.T
+
+	# Compute size of frequency bins
+	df = diff(f)
+
+	# Mask out values that are less than fISCO or greater than 10 Hz
+	mask = (f <= fISCO) & (f >= 10)
+	f = f[mask]
+	s = s[mask]
+	df = df[mask[:-1]]
+
+	# Integrate SNR from f_high down to f_low
+	accum_snr = sqrt(cumsum((f**(-7./3) / s**2 * df)[::-1]))[::-1]
+	accum_snr /= accum_snr.max()
+	accum_snr *= 100
+
+	# Print out row for data table
+	print name,
+	for f_low in f_lows:
+		print '&', '%.1f' % (accum_snr[f.searchsorted(f_low)]),
+	print r'\\'
+
+	# Make plot
+	plot(f, accum_snr, 'k',
+		label=name, **plotkwargs[name])
+print '\hline'
+print r'\end{tabular}'
+	
+xlim(80., 0)
+ylim(50, 100)
+grid(linewidth=0.1)
+xlabel('low frequency cutoff $f_\mathrm{low}$ ($\mathrm{Hz}$)')
+ylabel('percent accumulated SNR')
+title(r'(b) Accumulated SNR vs. $f_\mathrm{low}$')
[…]

### 2011-11-30 plot_asds.py

@@ -1,0 +1,22 @@
+#!/usr/bin/env python
+
+import sys
+from noisemodels import asds, plotkwargs
+from pylab import *
+
+fig_width = 3.35
+fig_height = 2.75
+fig = figure(figsize=(fig_width,fig_height))
+for name, data in asds:
+	f, asd = data.T
+	loglog(f, asd, label=name, **plotkwargs[name])
+#legend(loc=(1.05,0.0))
+xlabel('frequency ($\mathrm{Hz}$)')
+ylabel(r'amplitude spectral density ($1/\sqrt{\mathrm{Hz}}$)')
+xlim(9, 3000)
+grid(which='minor', color='gray', linewidth=0.1)
+grid(linestyle='--', color='k', which='major', linewidth=0.1)
+ylim(1e-24, 3e-21)
+title('(c) LIGO noise models')
+plt.subplots_adjust(bottom=0.45/fig_height,top=1-0.25/fig_height,left=0.75/fig_width,right=(2.4+0.75)/fig_width)
+savefig(sys.argv[1], dpi=360)

### 2011-11-30 plot_inspiral_tf_relation.py

@@ -1,0 +1,55 @@
+#!/usr/bin/env python
+
+import sys
+from pylal.xlal import constants
+from pylab import *
+
+# Component mass 1 in M_sun
+m1 = 1.4
+
+# Component mass 2 in M_sun
+m2 = 1.4
+
+# Total mass in M_sun
+M = m1 + m2
+
+# Chirp mass
+Mchirp = (m1 * m2) ** 0.6 * M ** -0.2
+
+# ISCO frequency in Hz
+fISCO = 4400. / M
+
+def freq_for_time(t):
+	return (t / (5 * Mchirp * constants.LAL_MTSUN_SI)) ** (-3./8) / (8 * pi * Mchirp * constants.LAL_MTSUN_SI)
+
+def time_for_freq(f):
+	return 5 * (8 * pi) ** (-8./3) * (Mchirp * constants.LAL_MTSUN_SI) ** (-5./3) * f ** (-8./3)
+
+fig_width = 3.35
+fig_height = 2.75
+fig = figure(figsize=(fig_width,fig_height))
+f = arange(1., 100., 1.)
+loglog(f, time_for_freq(f) - time_for_freq(fISCO), 'k', linewidth=1)
+for f_low in arange(10., 50., 10.):
+	axvline(f_low, color='k', linewidth=1)
+	plot([f_low], time_for_freq(f_low), 'o', markersize=5, markeredgewidth=1, markerfacecolor='none', markeredgecolor='k')
+	annotate('%d s' % round(time_for_freq(f_low)), (f_low, time_for_freq(f_low)), xytext=(4, 8), textcoords='offset points', backgroundcolor='white')
+
+xlim(10, 100)
+ylim(10, 10000)
+grid(which='minor', linewidth=0.1)
+grid(which='major', linestyle='--', color='k')
+xlabel('low frequency cutoff $f_\mathrm{low}$ ($\mathrm{Hz}$)')
+ylabel('signal duration from $f=f_\mathrm{low}$ to $f_\mathrm{ISCO}$')
+
+
+#twiny()
+#xlim(1, 100)
+#ylim(10, 10000)
+#xscale('log')
+#yscale('log')
+xticks( (1, 10, 20, 30, 40, 100), ('1 Hz', '10 Hz', '20', '30', '40', '100'))
+
+title(r'(d) Inspiral duration vs. $f_\mathrm{low}$')
[…]

### 2011-11-30 plot_legend.py

@@ -1,0 +1,17 @@
+#!/usr/bin/env python
+
+import sys
+from noisemodels import asds, plotkwargs
+from pylab import *
+
+fig_width = 7.
+fig_height = 0.75
+fig = figure(figsize=(fig_width,fig_height))
+for name, data in asds:
+	plot([0, 1], [0, 1], label=name.replace('deg', '$^\circ$').replace('det.', 'detuning'), **plotkwargs[name])
+legend(loc=(0.,-1.5), ncol=3)
+ax = gca()
+ax.get_xaxis().set_visible(False)
+ax.get_yaxis().set_visible(False)
+plt.subplots_adjust(bottom=0.4,top=0.6,left=0.75/fig_width,right=(2.5+0.75)/fig_width)
+savefig(sys.argv[1])

### 2011-11-30 plot_weighted_asds.py

@@ -1,0 +1,49 @@
+#!/usr/bin/env python
+
+import sys
+from noisemodels import asds, plotkwargs
+from pylab import *
+
+# Component mass 1 in M_sun
+m1 = 1.4
+
+# Component mass 2 in M_sun
+m2 = 1.4
+
+# Total mass in M_sun
+M = m1 + m2
+
+# ISCO frequency in Hz
+fISCO = 4400 / M
+
+fig_width = 3.35
+fig_height = 2.75
+fig = figure(figsize=(fig_width,fig_height))
+
+for name, data in asds:
+
+	# Unpack frequency and amplitue spectral density from array
+	f, s = data.T
+
+	# Compute size of frequency bins
+	df = diff(f)
+
+	# Mask out values that are less than fISCO or greater than 10 Hz
+	mask = (f <= fISCO) & (f >= 10)
+	f = f[mask]
+	s = s[mask]
+	df = df[mask[:-1]]
+
+	weighted_psd = f**(-7./3) / (s * s)
+	weighted_psd /= sum(weighted_psd * df)
+
+	# Make plot
+	plot(f, weighted_psd, label=name, **plotkwargs[name])
+
+xlim(0, 300)
+grid(linewidth=0.1)
+xlabel('frequency ($\mathrm{Hz}$)')
+ylabel(r'normalized power spectral density')
+title('(a) Signal to noise per unit frequency')
+plt.subplots_adjust(bottom=0.45/fig_height,top=1-0.2/fig_height,left=0.75/fig_width,right=(2.4+0.75)/fig_width)
+savefig(sys.argv[1])

### 2011-11-30 article.tex
Addressed referee's comment 6
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1420 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Advanced \LIGO\ and Advanced Virgo.  This is the \emph{minimum} area; localization is best at high elevation
-from the plane containing the detectors and worst at zero elevation. As
-\citeauthor{Fairhurst2009} cautions, his formula also slightly
-\emph{overestimates} the minimum area for low SNR events because it
-makes an approximation that the timing errors are Gaussian distributed
-and that the sky location estimates are Gaussian distributed about the
-true location.  Nonetheless, given the large uncertainties in rates,
-\citeauthor{Fairhurst2009}'s approximations suffice for an order of
-magnitude estimate.
+from the plane containing the detectors and worst at zero elevation.
+\citeauthor{Fairhurst2009} also cautions that his Fisher matrix calculation fails to capture disconnected patches of probability, which occur prominently in networks of three detectors where there are generally two local maxima on opposite sides of the plane of the detectors.
+(\citeauthor{Fairhurst2009} also points out that even his approximation that the timing errors are Gaussian breaks down at low SNR.)  At most sky locations, the localization uncertainty estimated from timing is highly optimistic and will only suffice for an order of  magnitude estimate.
 Once per year, we expect to observe an

### 2011-11-30 plot_accum_snr.py
Fixed plot labels
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1423 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@ -70,6 +70,6 @@ ylim(50, 100)
 grid(linewidth=0.1)
 xlabel('low frequency cutoff $f_\mathrm{low}$ ($\mathrm{Hz}$)')
 ylabel('percent accumulated SNR')
-title(r'(b) Accumulated SNR vs. $f_\mathrm{low}$')
+title(r'(c) Accumulated SNR vs. $f_\mathrm{low}$')
 plt.subplots_adjust(bottom=0.45/fig_height,top=1-0.2/fig_height,left=0.75/fig_width,right=(2.4+0.75)/fig_width)
 savefig(sys.argv[1])

### 2011-11-30 plot_asds.py

@@ -17,6 +17,6 @@ xlim(9, 3000)
 grid(which='minor', color='gray', linewidth=0.1)
 grid(linestyle='--', color='k', which='major', linewidth=0.1)
 ylim(1e-24, 3e-21)
-title('(c) LIGO noise models')
+title('(a) LIGO noise models')
 plt.subplots_adjust(bottom=0.45/fig_height,top=1-0.25/fig_height,left=0.75/fig_width,right=(2.4+0.75)/fig_width)
 savefig(sys.argv[1], dpi=360)

### 2011-11-30 plot_weighted_asds.py

@@ -44,6 +44,6 @@ xlim(0, 300)
 grid(linewidth=0.1)
 xlabel('frequency ($\mathrm{Hz}$)')
 ylabel(r'normalized power spectral density')
-title('(a) Signal to noise per unit frequency')
+title('(b) Signal to noise per unit frequency')
 plt.subplots_adjust(bottom=0.45/fig_height,top=1-0.2/fig_height,left=0.75/fig_width,right=(2.4+0.75)/fig_width)
 savefig(sys.argv[1])

### 2011-11-30 article.tex
Removed extra space
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1422 ccdd05d5-56c1-461a-bf98-85f6876523b4
Not counterclockwise
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1425 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-	\caption{\label{fig:low-frequency-cutoff}Counterclockwise from top left:  (a)~Noise amplitude spectral density for a variety of Advanced LIGO noise models, S5, and S6.
+	\caption{\label{fig:low-frequency-cutoff}From top left:  (a)~Noise amplitude spectral density for a variety of Advanced LIGO noise models, S5, and S6.
 (b)~Normalized signal to noise per unit frequency, $(\mathrm{d}\rho^2/\mathrm{d}f)/\rho^2$, for a (1.4,~1.4)~$\Msun$ inspiral.

### 2011-11-30 article.tex
Refined reply to comment 6
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1426 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \citeauthor{Fairhurst2009} also cautions that his Fisher matrix calculation fails to capture disconnected patches of probability, which occur prominently in networks of three detectors where there are generally two local maxima on opposite sides of the plane of the detectors.
-(\citeauthor{Fairhurst2009} also points out that even his approximation that the timing errors are Gaussian breaks down at low SNR.)  At most sky locations, the localization uncertainty estimated from timing is highly optimistic and will only suffice for an order of magnitude estimate.
+Aside from the mirror degeneracy, characterizing the uncertainty region by the Fisher matrix alone tends to underestimate, rather than overestimate, the area for low SNR events, but this effect is generally more than compensated by the source being in an unfavorable sky location.
+For these reasons, the localization uncertainty estimated from timing is highly optimistic and will only suffice for an order of magnitude estimate.
 Once per year, we expect to observe an

### 2011-12-01 article.tex
Made Figure uppercase
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1429 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The choice of $f_\mathrm{low}$ is based on the fraction of the total SNR that is accumulated at frequencies above $f_\mathrm{low}$.
-To illustrate the relative contributions to the SNR at different frequencies for a (1.4,~1.4)~$\Msun$ binary, we normalized and plotted the integrand of equation~(\ref{eq:expected-snr}), the noise-weighted power spectral density of the inspiral waveform, in figure~\ref{fig:low-frequency-cutoff}b.
+To illustrate the relative contributions to the SNR at different frequencies for a (1.4,~1.4)~$\Msun$ binary, we normalized and plotted the integrand of equation~(\ref{eq:expected-snr}), the noise-weighted power spectral density of the inspiral waveform, in Figure~\ref{fig:low-frequency-cutoff}b.
 This is the quantity
@@
 We used several different noise power spectra: all of the envisioned Advanced LIGO configurations from~\citet{ALIGONoise}; the best achieved sensitivity at LIGO Hanford Observatory (LHO) in LIGO's fifth science run (S5), measured by \citet{S5InspiralRange}; and the best achieved sensitivity at LHO during S6, measured by \citet{S6InspiralRange}.
-(The noise spectra themselves are shown in figure~\ref{fig:low-frequency-cutoff}a.)  It is plain that during S5 and S6 the greatest contribution to the SNR was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the SNR is accumulated below 60~Hz.
-This information is presented in a complementary way in figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, interpreted as a fraction of the total ``available'' SNR,
+(The noise spectra themselves are shown in Figure~\ref{fig:low-frequency-cutoff}a.)  It is plain that during S5 and S6 the greatest contribution to the SNR was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the SNR is accumulated below 60~Hz.
+This information is presented in a complementary way in Figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, interpreted as a fraction of the total ``available'' SNR,
 %
@@
 The instantaneous gravitational wave frequency, given by equation~(\ref{eq:fgw}) is a power law function of time, so the amount of time for the gravitational wave frequency to evolve from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$ depends strongly on $f_\mathrm{low}$.
-The duration of a (1.4,~1.4)~$\Msun$ inspiral is show in figure~\ref{fig:low-frequency-cutoff}d.
+The duration of a (1.4,~1.4)~$\Msun$ inspiral is show in Figure~\ref{fig:low-frequency-cutoff}d.
 The inspiral takes only 25~s to evolve from 40~Hz to $f_\mathrm{ISCO}$, but takes 54~s to evolve from 30~Hz to $f_\mathrm{ISCO}$, 158~s from 20~Hz, and 1002~s from 10~Hz.
@@
 (d)~Amount of time for a \NS--\NS\ inspiral signal to evolve from frequency $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, as a function of $f_\mathrm{low}$.
-For figures (a) through (c), the line style indicates which noise model was used.}
+For Figures (a) through (c), the line style indicates which noise model was used.}
 \end{figure}

### 2012-01-05 article.tex
Addressed comments about aliasing and spectral leakage
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1434 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \usepackage{epstopdf}
-\usepackage{amsmath}
+\usepackage{amsmath,amsthm}
 \usepackage{stmaryrd} % For \shortuparrow and \shortdownarrow
 \usepackage{upgreek} % For \upmu, upright mu
+
+\newtheorem{thm}{Theorem}
 
@@
 \DeclareMathOperator{\var}{var}
+\DeclareMathOperator{\sinc}{sinc}
 
@@
 Final triggers are built from the reconstructed matched filter output.
-Another novel approach using networks of parallel, second order IIR filters is being explored
+Another novel approach using networks of parallel, second order infinite impulse response~(IIR) filters is being explored
 by \citet{shaunIIR, linqingIIR}.
@@
 
+\subsection{Other potential sources of \SNR\ loss}
+
+One possible source of \SNR\ loss for which we have not accounted is the leakage of sharp spectral features built into the whitening filter.
+In the \lloid\ algorithm, as with many other \GW\ search methods, the whitening procedure is treated as a data conditioning procedure that is separated from the matched filter stage.
+In this paper, we assume that the input to the filter bank is a whitened data stream --- the result of feeding the \GW\ antenna's strain time series through a filter that flattens and normalizes its spectrum --- and how the whitening filter is implemented is not described in detail.
+We elected to omit a detailed description of the whitening procedure since the focus here is on the implementation of a scalable inspiral filter bank.
+
+However, the inspiral templates themselves consist of the gravitational wave time series convolved with the impulse response of the whitening filter.
+As a consequence, the \lloid\ algorithm must faithfully replicate the effect of the whitening filter.
+Since in practice the noise spectra of ground-based \GW\ detectors contain both high-$Q$ lines at mechanical, electronic, and control resonances and a very sharp rolloff at the seismic wall, the frequency response of the \lloid\ filter bank must contain both high-$Q$ notches and a very abrupt high-pass filter.
+\fir\ filters with rapidly varying frequency responses tend to have long impulse responses and many coefficients.
+Since the \lloid\ basis filters have, by design, short impulse responses and very few coefficients, one might be concerned about spectral leakage contaminating the frequency response of the LLOID filter bank.
+
+The usual statement of the famous Nyquist-Shannon theorem, stated below as Theorem~\ref{thm:nyquist}, has a natural dual, \ref{thm:nyquist-dual}, that addresses the frequency resolution that can be achieved with a \fir\ filter of a given length.
+
+\begin{thm}
+	\label{thm:nyquist}
+	\citep[after][p. 518]{oppenheim1997signals}
+	Let $x(t)$ be a band-limited signal with continuous Fourier transform $\tilde{x}(f)$ such that $\tilde{x}(f') = 0 \; \forall \; f' : |f'| < f_M$.
+Then $x(t)$ is uniquely determined by its discrete samples $x(n/f_s)$, $n = 0, \pm 1, \pm 2, \dots$, if $f_M < f_s / 2$.
+\end{thm}
+
+\begin{thm}
+	\label{thm:nyquist-dual}
+	Let $x(t)$ be a compactly supported signal such that $x(t') = 0 \; \forall \; t' : |t'| < t_M$.
+Then its continuous Fourier transform $\tilde{x}(f)$ is uniquely determined by the discrete frequency components $\tilde{x}(n \, \Delta f)$, $n = 0, \pm 1, \pm 2, \dots$, if $t_M > 1 / (2 \, \Delta f)$.
+\end{thm}
+
+Another way of stating Theorem~\ref{thm:nyquist-dual} is that, provided $x(t)$ vanishes for $|t| > t_M$, where $t_M > 1 / (2 \, \Delta f)$, the continuous Fourier transform can be reconstructed at any frequency $f$ from a weighted sum of $\sinc$ functions centered at each of the discrete frequency components, namely,
+$$
+	\tilde{x}(f) \propto \sum_{n=-\infty}^{\infty} \tilde{x}\left(n \, \Delta f\right) \sinc \left[ 2 \pi (f - n \, \Delta f) t_M \right].
+$$
+Failure to meet this dual of the Nyquist-Shannon sampling criterion means that power from one of the discrete frequency components will ``leak'' into the rest of the spectrum through the side lobes of the $\sinc$ functions, producing a smeared out version of the desired frequency response.
+In the \GW\ data analysis literature, the synthesis of inspiral matched filters involves a step called \emph{inverse spectrum truncation} \citep[see][section VII]{findchirppaper} that fixes the number of coefficients based on the desired frequency resolution.
+
+In order to effectively flatten a line in the detector's noise power
+spectrum, the time scale of the templates must be at least as long as
+the damping time $\tau$ of the line, $\tau = 2 Q / \omega_0$, where $Q$ is the quality
+factor of the line and $w_0$ is the central angular frequency.  For example,
+in the S6 detector configuration known as ``Enhanced LIGO,'' the violin modes \citep{ELIGOSusp} had $Q \sim 10^5$ and $\omega_0 \sim (2 \pi) 340$~rad/s, for a coherence time $\tau \sim 10^2$~s.
+
+In our example template bank, many of the time slices are much shorter
+than this.  However, the impulse response of the LLOID filter bank is very nearly
+equal to the full templates themselves, and the full templates are much longer than many
+coherence times of the violin mode.  For this reason, we do not expect
+\lloid\ to be affected by sharp line features any more than the
+traditional \fft-based searches currently employed in the \GW\ field.  Future works ought to verify this reasonable expectation with numerical experiments.
+
+There could, in principle, be lines with coherence times many times
+longer than the template duration.  For example, the $Q$ of the violin
+modes may increase by orders of magnitude in Advanced \LIGO~\citep{ALIGOSusp}.  Also,
+there are certainly narrow lines that are non-stationary.  Both of these
+cases can be dealt with by preprocessing $h(t)$ with bandstop filters
+that attenuates the lines themselves but also conservative neighborhoods
+around them.  If such a bandstop was implemented as an \fir\ filter, it
+could be built into the time slices just as before.
+
+Another way to deal with line features with coherence times much longer than the templates would be to entrirely `factor' the whitening out of the \lloid\ filter bank.
+The data would be passed through the whitening filter twice and the inspiral templates would not depend on the detector's noise power spectral density at all.
+Any line features could be notched out in the whitening stage with IIR filters, which can achieve infinitely high $Q$ at just second order.
+
 \subsection{Lower bounds on computational cost and latency compared to other

### 2012-01-05 article.tex
Fixed some whitespace in referee reply
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1436 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 %
-where $D$ is the luminosity distance and $S(f)$ is the power spectral density of the detector noise.
+where $D$ is the luminosity distance and $S(f)$ is the one-sided power spectral density of the detector noise.
 $f_\mathrm{low}$ and $f_\mathrm{high}$ are low and high frequency limits of integration which may be chosen to extend across the entire bandwidth of the detector.
@@
 up to an arbitrary multiplicative constant.
-Here, $\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively; $\tilde{S}[n]$ is the power spectral density of $n[k]$.
-The discrete Fourier transform of the output is
+Here, $\tilde{h}[n]$, $\tilde{s}[n]$, and $\tilde{x}[n]$ are the discrete Fourier transforms (DFTs) of $h[k]$, $s[k]$, and $x[k]$, respectively; $\tilde{S}[n] = \expectation \left[ \tilde{n}[n] \tilde{n}^* [n] \right]$ is the folded, two-sided, discrete power spectrum of $n[k]$.
+It is related to the continuous, one-sided power spectral density $S(f)$ through
+$$
+	\tilde{S}[n] =
+	\begin{cases}
+		S(n f_s / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
+		\tilde{S}[N - n] & \textrm{otherwise},
+	\end{cases}
+$$
+where $N$ is the length of the filter and $f_s$ is the sample rate.
+(In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f_s / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
 %

### 2012-01-05 article.tex
Increased DCC version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1437 ccdd05d5-56c1-461a-bf98-85f6876523b4
Changed capitalization of FFT
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1439 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 much more computationally efficient to use the convolution theorem and the
-Fast Fourier Transform (\fft) to implement fast convolution in the frequency
+fast Fourier transform (\fft) to implement fast convolution in the frequency
 domain, requiring only
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v29}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v30}.
 

### 2012-01-05 article.tex
Added square to numerator in definition of signal to noise
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1440 ccdd05d5-56c1-461a-bf98-85f6876523b4
Rearranged clauses in sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1441 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $$
-\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]}{\var \left[ y[k] \right]} = \frac{y_s[0]}{\var \left[ y_n[k] \right]}.
+\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]^2}{\var \left[ y[k] \right]} = \frac{y_s[0]^2}{\var \left[ y_n[k] \right]}.
 $$
@@
 slices in a \TD\ analogue to the \FD\ decomposition employed by MBTA~\citep{Marion2004,Buskulic2010}.
-The application to GW data analysis is foreshadowed by an earlier \FD\ convolution algorithm based on splitting the impulse response of a filter into smaller blocks, proposed by \citet{gardner1995efficient}.
+The application to GW data analysis is foreshadowed by an earlier \FD\ convolution algorithm, proposed by \citet{gardner1995efficient}, based on splitting the impulse response of a filter into smaller blocks.
 We decompose each template

### 2012-01-05 article.tex
The authors of Cannon et al. -> Cannon et al.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1442 ccdd05d5-56c1-461a-bf98-85f6876523b4
Changed citation style
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1443 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 determined by the number $\numsvdtmps$ of basis templates that are kept in
-the approximation.  The authors of \citet{Cannon:2010p10398}
+the approximation.  \citet{Cannon:2010p10398}
 showed that highly accurate approximations of inspiral template banks could be
@@
 smaller than the original number of templates \numtmps\ and improves on the
-rank reduction demonstrated in \citep{Cannon:2010p10398} by nearly an order
+rank reduction demonstrated in \citet{Cannon:2010p10398} by nearly an order
 of magnitude.

### 2012-01-06 article.tex
Added clarification requested by Chad
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1446 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added another clarification to section 5.5
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1448 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 non-floating point tasks including bookkeeping and thread synchronization.
-Given these considerations, we estimate that a full single-detector Advanced
-\LIGO\ search with \lloid\ will require $\sim$$10$ machines.
+Given these considerations, we estimate that a full Advanced
+\LIGO, single-detector, \NS--\NS\ search with \lloid\ in will require $\sim$$10$ machines.
 
@@
 The direct calculation from expression~(\ref{eq:fd-flops}) gives $5.9 \times 10^{10}$~\flops, in order-of-magnitude agreement.
-It appears that in this application the conventional \FD\ search is scarcely a factor of 2 more computationally efficient than LLOID, but only at the price of thousands of seconds of latency.
+In this application, the conventional \FD\ search is scarcely a factor of 2 faster than LLOID while gaining only $0.3$\% in \SNR, but only at the price of thousands of seconds of latency.
 

### 2012-01-06 article.tex
Correction to appendix: above, not below
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1449 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 In S5 and S6, all of the SNR is accumulated above 40~Hz.
-For the `high frequency' Advanced LIGO configuration, almost 50\% of the SNR is accumulated \emph{below} 40~Hz.
-For the preferred final configuration, 'zero detuning, high power,' 86.1\% of the SNR is below 40~Hz, 93.2\% is below 30~Hz, and 98.1\% is below 20~Hz.
+For the `high frequency' Advanced LIGO configuration, scarcely \emph{half} of the SNR is accumulated above 40~Hz.
+For the preferred final configuration, 'zero detuning, high power,' 86.1\% of the SNR is above 40~Hz, 93.2\% is above 30~Hz, and 98.1\% is above 20~Hz.
 

### 2012-01-06 article.tex
Used typographers' quotes
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1450 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 For the `high frequency' Advanced LIGO configuration, scarcely \emph{half} of the SNR is accumulated above 40~Hz.
-For the preferred final configuration, 'zero detuning, high power,' 86.1\% of the SNR is above 40~Hz, 93.2\% is above 30~Hz, and 98.1\% is above 20~Hz.
+For the preferred final configuration, `zero detuning, high power,' 86.1\% of the SNR is above 40~Hz, 93.2\% is above 30~Hz, and 98.1\% is above 20~Hz.
 

### 2012-01-06 article.tex
Brought discussion of Q factor back to sampling theorem; mentioned weighted SVD
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1451 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-One possible source of \SNR\ loss for which we have not accounted is the leakage of sharp spectral features built into the whitening filter.
-In the \lloid\ algorithm, as with many other \GW\ search methods, the whitening procedure is treated as a data conditioning procedure that is separated from the matched filter stage.
-In this paper, we assume that the input to the filter bank is a whitened data stream --- the result of feeding the \GW\ antenna's strain time series through a filter that flattens and normalizes its spectrum --- and how the whitening filter is implemented is not described in detail.
+One possible source of \SNR\ loss for which we have not accounted is the leakage of sharp spectral features in the detector's noise spectrum due to the short durations of the time slices.
+In the \lloid\ algorithm, as with many other \GW\ search methods, whitening is treated as an entirely separate data conditioning stage.
+In this paper, we assume that the input to the filter bank is already whitened, having been passed through a filter that flattens and normalizes its spectrum.
 We elected to omit a detailed description of the whitening procedure since the focus here is on the implementation of a scalable inspiral filter bank.
@@
 the damping time $\tau$ of the line, $\tau = 2 Q / \omega_0$, where $Q$ is the quality
-factor of the line and $w_0$ is the central angular frequency.  For example,
+factor of the line and $w_0$ is the central angular frequency.
+To put this into the context of the sampling theorem, in order to resolve a notch with a particular $Q$ and $f_0$, a \fir\ filter must achieve a frequency resolution of $\Delta f \gtrsim \pi f_0 / Q$ and therefore its impulse response must last for at least a time $1/\Delta f = Q / \pi f_0$.
+For example,
 in the S6 detector configuration known as ``Enhanced LIGO,'' the violin modes \citep{ELIGOSusp} had $Q \sim 10^5$ and $\omega_0 \sim (2 \pi) 340$~rad/s, for a coherence time $\tau \sim 10^2$~s.
@@
 In our example template bank, many of the time slices are much shorter
-than this.  However, the impulse response of the LLOID filter bank is very nearly
-equal to the full templates themselves, and the full templates are much longer than many
-coherence times of the violin mode.  For this reason, we do not expect
-\lloid\ to be affected by sharp line features any more than the
-traditional \fft-based searches currently employed in the \GW\ field.  Future works ought to verify this reasonable expectation with numerical experiments.
+than this.  However, in summation the time slices have the same duration as the full templates themselves, and the full templates are much longer than many
+coherence times of the violin mode.  For this reason, we speculate that 
+\lloid\ should be just as robust to sharp line features as
+traditional \fft-based searches currently employed in the \GW\ field.
+Future works must verify this reasonable supposition with numerical experiments, including impulse response studies similar to the ones presented here but with detector noise power spectra containing lines with realistically high quality factors.
 
@@
 cases can be dealt with by preprocessing $h(t)$ with bandstop filters
-that attenuates the lines themselves but also conservative neighborhoods
-around them.  If such a bandstop was implemented as an \fir\ filter, it
-could be built into the time slices just as before.
-
-Another way to deal with line features with coherence times much longer than the templates would be to entrirely `factor' the whitening out of the \lloid\ filter bank.
-The data would be passed through the whitening filter twice and the inspiral templates would not depend on the detector's noise power spectral density at all.
+that attenuates the lines themselves but also conservatively large neighborhoods
+around them.  If such bandstops were implemented as an \fir\ filter, they
+could be built into the time slices without any difficulty.
+
+Perhaps another way to deal with line features with coherence times much longer than the templates would be to entrirely `factor' the whitening out of the \lloid\ filter bank.
 Any line features could be notched out in the whitening stage with IIR filters, which can achieve infinitely high $Q$ at just second order.
+If the detector data were passed through the whitening filter twice, then time-sliced filters need not depend on the detector's noise power spectral density at all.
+In such a variation on the \lloid\ method, the basis filters could be calculated from the \emph{weighted} \SVD~\citep[chapter 3.6]{WeightedSVD, jackson2003user} of the time-sliced templates, using the covariance of the detector noise as a weight matrix.
 

### 2012-01-10 article.tex
Some of Nick's comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1458 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Perhaps another way to deal with line features with coherence times much longer than the templates would be to entrirely `factor' the whitening out of the \lloid\ filter bank.
+Another way to deal with line features with coherence times much longer than the templates would be to entirely `factor' the whitening out of the \lloid\ filter bank.
 Any line features could be notched out in the whitening stage with IIR filters, which can achieve infinitely high $Q$ at just second order.

### 2012-01-10 article.tex
f_s -> f^0
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1459 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	\begin{cases}
-		S(n f_s / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
+		S(n f^0 / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
 		\tilde{S}[N - n] & \textrm{otherwise},
@@
 $$
-where $N$ is the length of the filter and $f_s$ is the sample rate.
-(In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f_s / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
+where $N$ is the length of the filter and $f^0$ is the sample rate.
+(In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f^0 / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
 %
@@
 	Let $x(t)$ be a band-limited signal with continuous Fourier transform $\tilde{x}(f)$ such that $\tilde{x}(f') = 0 \; \forall \; f' : |f'| < f_M$.
-Then $x(t)$ is uniquely determined by its discrete samples $x(n/f_s)$, $n = 0, \pm 1, \pm 2, \dots$, if $f_M < f_s / 2$.
+Then $x(t)$ is uniquely determined by its discrete samples $x(n/f^0)$, $n = 0, \pm 1, \pm 2, \dots$, if $f_M < f^0 / 2$.
 \end{thm}

### 2012-01-10 article.tex
Corrected definition of discrete PSD
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1461 ccdd05d5-56c1-461a-bf98-85f6876523b4
Addressed referee's comment 12
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1462 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	\begin{cases}
-		S(n f^0 / 2 N) / 2 & \textrm{if } 0 \leqslant n \leqslant N / 2 \\
+		S(n) & \textrm{if } n = 0 \textrm{ or } n = N / 2 \\
+		S(n f^0 / 2 N) / 2 & \textrm{if } 0 < n < N / 2 \\
 		\tilde{S}[N - n] & \textrm{otherwise},
@@
 evaluate the anti-aliasing filter only for the fraction $f^s / f^{s-1}$ of the
-samples that will not be deleted.  Consequently, an efficient decimator that
-requires only $2 N^\shortdownarrow f^{s-1}$ \flops\ is possible.
+samples that will not be deleted.  Consequently, an efficient decimator 
+requires only $2 N^\shortdownarrow f^{s-1}$ \flops.  (One common realization is an ingenious structure called a \emph{polyphase decimator}, described in chapter 1 of \citet{jovanovic2002multirate}.)
 
@@
 zero, it is possible to build an efficient interpolator that requires only
-$2 M N^\shortuparrow f^{s-1}$ \flops.
+$2 M N^\shortuparrow f^{s-1}$ \flops.  (Again, see \citet{jovanovic2002multirate} for a discussion of \emph{polyphase interpolation}.)
 

### 2012-01-22 article.tex
Added caveat to section 5.4
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1463 ccdd05d5-56c1-461a-bf98-85f6876523b4
Fixed typos in Nyquist theorem
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1465 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 	\citep[after][p. 518]{oppenheim1997signals}
-	Let $x(t)$ be a band-limited signal with continuous Fourier transform $\tilde{x}(f)$ such that $\tilde{x}(f') = 0 \; \forall \; f' : |f'| < f_M$.
-Then $x(t)$ is uniquely determined by its discrete samples $x(n/f^0)$, $n = 0, \pm 1, \pm 2, \dots$, if $f_M < f^0 / 2$.
+	Let $x(t)$ be a band-limited signal with continuous Fourier transform $\tilde{x}(f)$ such that $\tilde{x}(f') = 0 \; \forall \; f' : |f'| > f_M$.
+Then $x(t)$ is uniquely determined by its discrete samples $x(n/f^0)$, $n = 0, \pm 1, \pm 2, \dots$, if $f^0 > 2 f_M$.
 \end{thm}
@@
 	\label{thm:nyquist-dual}
-	Let $x(t)$ be a compactly supported signal such that $x(t') = 0 \; \forall \; t' : |t'| < t_M$.
-Then its continuous Fourier transform $\tilde{x}(f)$ is uniquely determined by the discrete frequency components $\tilde{x}(n \, \Delta f)$, $n = 0, \pm 1, \pm 2, \dots$, if $t_M > 1 / (2 \, \Delta f)$.
+	Let $x(t)$ be a compactly supported signal such that $x(t') = 0 \; \forall \; t' : |t'| > t_M$.
+Then its continuous Fourier transform $\tilde{x}(f)$ is uniquely determined by the discrete frequency components $\tilde{x}(n \, \Delta f)$, $n = 0, \pm 1, \pm 2, \dots$, if $\Delta f < 1 / (2 t_M)$.
 \end{thm}
 
-Another way of stating Theorem~\ref{thm:nyquist-dual} is that, provided $x(t)$ vanishes for $|t| > t_M$, where $t_M > 1 / (2 \, \Delta f)$, the continuous Fourier transform can be reconstructed at any frequency $f$ from a weighted sum of $\sinc$ functions centered at each of the discrete frequency components, namely,
+Another way of stating Theorem~\ref{thm:nyquist-dual} is that, provided $x(t)$ vanishes for $|t| > t_M$, where $t_M < 1 / (2 \, \Delta f)$, the continuous Fourier transform can be reconstructed at any frequency $f$ from a weighted sum of $\sinc$ functions centered at each of the discrete frequency components, namely,
 $$
@@
 five orders of magnitude faster than the conventional, direct \TD\ method.  However, the
-\FD\ method has a latency of over half of an hour, whereas the \lloid\ algorithm
-has no latency at all.
+\FD\ method has a latency of over half of an hour, whereas the \lloid\ method, with suitable design of the decimation and interpolation filters, has no more latency than the direct \TD\ method.
 %

### 2012-01-23 article.tex
Incremented DCC version
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1466 ccdd05d5-56c1-461a-bf98-85f6876523b4
Switched 'overestimate' with 'underestimate'
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1467 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \citeauthor{Fairhurst2009} also cautions that his Fisher matrix calculation fails to capture disconnected patches of probability, which occur prominently in networks of three detectors where there are generally two local maxima on opposite sides of the plane of the detectors.
-Aside from the mirror degeneracy, characterizing the uncertainty region by the Fisher matrix alone tends to underestimate, rather than overestimate, the area for low SNR events, but this effect is generally more than compensated by the source being in an unfavorable sky location.
+Aside from the mirror degeneracy, characterizing the uncertainty region by the Fisher matrix alone tends to overestimate, rather than underestimate, the area for low SNR events, but this effect is generally more than compensated by the source being in an unfavorable sky location.
 For these reasons, the localization uncertainty estimated from timing is highly optimistic and will only suffice for an order of magnitude estimate.
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v30}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v31}.
 

### 2012-01-23 article.tex
Rewrote paragraph on spectral leakage
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1468 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-Another way of stating Theorem~\ref{thm:nyquist-dual} is that, provided $x(t)$ vanishes for $|t| > t_M$, where $t_M < 1 / (2 \, \Delta f)$, the continuous Fourier transform can be reconstructed at any frequency $f$ from a weighted sum of $\sinc$ functions centered at each of the discrete frequency components, namely,
-$$
-	\tilde{x}(f) \propto \sum_{n=-\infty}^{\infty} \tilde{x}\left(n \, \Delta f\right) \sinc \left[ 2 \pi (f - n \, \Delta f) t_M \right].
-$$
-Failure to meet this dual of the Nyquist-Shannon sampling criterion means that power from one of the discrete frequency components will ``leak'' into the rest of the spectrum through the side lobes of the $\sinc$ functions, producing a smeared out version of the desired frequency response.
+Another way of stating Theorem~\ref{thm:nyquist-dual} is that, provided $x(t)$ is nonzero only for $|t| < 1 / (2 \, \Delta f)$, the continuous Fourier transform can be reconstructed at any frequency $f$ from a weighted sum of $\sinc$ functions centered at each of the discrete frequency components, namely,
+$$
+	\tilde{x}(f) \propto \sum_{n=-\infty}^{\infty} \tilde{x}\left(n \, \Delta f\right) \sinc \left[ \pi (f - n \, \Delta f) / \Delta f \right].
+$$
+Failure to meet the conditions of this dual of the sampling theorem results in spectral leakage.
+For a time-domain signal to capture spectral features that are the size of the central lobe of the $\sinc$ function, the signal must have a duration greater than $1/\Delta f$.
+If the signal $x(t)$ is truncated by sampling it for a shorter duration, then its Fourier transform becomes smeared out; conceptually, power ``leaks'' out into the side lobes of the $\sinc$ functions and washes away sharp spectral features.
 In the \GW\ data analysis literature, the synthesis of inspiral matched filters involves a step called \emph{inverse spectrum truncation} \citep[see][section VII]{findchirppaper} that fixes the number of coefficients based on the desired frequency resolution.

### 2012-01-23 article.tex
signal to noise -> (signal to noise)^2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1470 ccdd05d5-56c1-461a-bf98-85f6876523b4
v32
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1471 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 $$
-\textrm{signal to noise} = \frac{\expectation \left[ y[0] \right]^2}{\var \left[ y[k] \right]} = \frac{y_s[0]^2}{\var \left[ y_n[k] \right]}.
+(\textrm{signal to noise})^2 = \frac{\expectation \left[ y[0] \right]^2}{\var \left[ y[k] \right]} = \frac{y_s[0]^2}{\var \left[ y_n[k] \right]}.
 $$
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v31}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v32}.
 
@@
 For the preferred final configuration, `zero detuning, high power,' 86.1\% of the SNR is above 40~Hz, 93.2\% is above 30~Hz, and 98.1\% is above 20~Hz.
+(Since SNR accumulates in quadrature, this means, on the other hand, that under the `high frequency' configuration a template encompassing \emph{just the early inspiral} from 10 to 40~Hz would accumulate $\sqrt{1 - 0.533^2} \approx 84.6\%$ of the total SNR!
+In the `zero detuning, high power,' configuration, integration from 10 to 40~Hz alone would yield 50.9\% of the total SNR, from 10 to 30~Hz, 36.2\%, and from 10 to 20~Hz, 19.4\%.)
 
@@
 An inspiral search that is designed with a low frequency cutoff at the seismic wall would gain an increase in detection rate of $\rho_\mathrm{frac}^{-3}(f_\mathrm{low})$ relative to a search with a low frequency cutoff of $f_\mathrm{low}$.
-For a fractional accumulated SNR of 80\%, this would represent almost a twofold increase in the rate of detection.
-For $\rho_\mathrm{frac} = 90\%$, the detection rate would still be increased by as much as 37\%.
+This would represent almost a twofold increase in the rate of detection over a search with a fractional accumulated SNR of 80\%, and still a 37\% increase over a search with $\rho_\mathrm{frac} = 90\%$.
 Existing coalescing binary detection pipelines strive to sacrifice no more than 3\% of the available SNR; this forfeits less than a 10\% gain in detection rate.
-In order to satisfy this constraint, the low frequency cutoff would have to be placed below 20~Hz for all of the conceived Advanced LIGO configurations.
+In order to satisfy this constraint, the low frequency cutoff would have to be placed below 30~Hz for all of the conceived Advanced LIGO configurations.
 

### 2012-04-03 article.tex
Removed unused package
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1483 ccdd05d5-56c1-461a-bf98-85f6876523b4
ISCO -> LSO
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1484 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-\usepackage{lscape}
 \usepackage[colorlinks]{hyperref}
@@
 \newcommand{\GEO}{GEO600}%{{\sc geo600}}
-\newcommand{\ISCO}{ISCO}%{{\sc isco}}
+\newcommand{\LSO}{LSO}%{{\sc lso}}
 \newcommand{\SNR}{SNR}%{{\sc snr}}
@@
 is truncated at some prescribed time $t^0$, or equivalent frequency $f_\mathrm{high}$,
-often chosen to correspond to the innermost stable circular orbit (\ISCO). The beginning of the template is critically
+often chosen to correspond to the last stable orbit (\LSO). The beginning of the template is critically
 sampled at $2 f_\mathrm{low}$, but the end of the template is critically sampled at a
@@
 The expected matched-filter SNR, integrated from $f_\mathrm{low}$ to $f_\mathrm{high}$, is given by equation~(\ref{eq:expected-snr}).
-The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the ISCO; for non-spinning systems, $f_\mathrm{ISCO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary.
+The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the LSO; for non-spinning systems, $f_\mathrm{LSO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary.
 The choice of $f_\mathrm{low}$ is based on the fraction of the total SNR that is accumulated at frequencies above $f_\mathrm{low}$.
@@
 $$
-	\frac{1}{\rho^2}\frac{\mathrm{d}\rho^2}{\mathrm{d}f} = \frac{f^{-7/3}}{S(f)} \left( \int_0^{f_\mathrm{ISCO}} \frac{{f'}^{-7/3}}{S(f')} \, \mathrm{d}f' \right)^{-1},
+	\frac{1}{\rho^2}\frac{\mathrm{d}\rho^2}{\mathrm{d}f} = \frac{f^{-7/3}}{S(f)} \left( \int_0^{f_\mathrm{LSO}} \frac{{f'}^{-7/3}}{S(f')} \, \mathrm{d}f' \right)^{-1},
 $$
@@
 (The noise spectra themselves are shown in Figure~\ref{fig:low-frequency-cutoff}a.)  It is plain that during S5 and S6 the greatest contribution to the SNR was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the SNR is accumulated below 60~Hz.
-This information is presented in a complementary way in Figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, interpreted as a fraction of the total ``available'' SNR,
-%
-$$
-	\rho_\mathrm{frac}(f_\mathrm{low}) = \sqrt{\left( \int_{f_\mathrm{low}}^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right) \left( \int_0^{f_\mathrm{ISCO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right)^{-1}}.
+This information is presented in a complementary way in Figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{LSO}$, interpreted as a fraction of the total ``available'' SNR,
+%
+$$
+	\rho_\mathrm{frac}(f_\mathrm{low}) = \sqrt{\left( \int_{f_\mathrm{low}}^{f_\mathrm{LSO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right) \left( \int_0^{f_\mathrm{LSO}} \frac{{f}^{-7/3}}{S(f)} \, \mathrm{d}f \right)^{-1}}.
 $$
@@
 
-The instantaneous gravitational wave frequency, given by equation~(\ref{eq:fgw}) is a power law function of time, so the amount of time for the gravitational wave frequency to evolve from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$ depends strongly on $f_\mathrm{low}$.
+The instantaneous gravitational wave frequency, given by equation~(\ref{eq:fgw}) is a power law function of time, so the amount of time for the gravitational wave frequency to evolve from $f_\mathrm{low}$ to $f_\mathrm{LSO}$ depends strongly on $f_\mathrm{low}$.
 The duration of a (1.4,~1.4)~$\Msun$ inspiral is show in Figure~\ref{fig:low-frequency-cutoff}d.
-The inspiral takes only 25~s to evolve from 40~Hz to $f_\mathrm{ISCO}$, but takes 54~s to evolve from 30~Hz to $f_\mathrm{ISCO}$, 158~s from 20~Hz, and 1002~s from 10~Hz.
+The inspiral takes only 25~s to evolve from 40~Hz to $f_\mathrm{LSO}$, but takes 54~s to evolve from 30~Hz to $f_\mathrm{LSO}$, 158~s from 20~Hz, and 1002~s from 10~Hz.
 
@@
 (b)~Normalized signal to noise per unit frequency, $(\mathrm{d}\rho^2/\mathrm{d}f)/\rho^2$, for a (1.4,~1.4)~$\Msun$ inspiral.
-(c)~Percentage of SNR that is accumulated from $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, relative to SNR accumulated from $f_\mathrm{low} = 0\,\mathrm{Hz}$ to $f_\mathrm{ISCO}$.
-(d)~Amount of time for a \NS--\NS\ inspiral signal to evolve from frequency $f_\mathrm{low}$ to $f_\mathrm{ISCO}$, as a function of $f_\mathrm{low}$.
+(c)~Percentage of SNR that is accumulated from $f_\mathrm{low}$ to $f_\mathrm{LSO}$, relative to SNR accumulated from $f_\mathrm{low} = 0\,\mathrm{Hz}$ to $f_\mathrm{LSO}$.
+(d)~Amount of time for a \NS--\NS\ inspiral signal to evolve from frequency $f_\mathrm{low}$ to $f_\mathrm{LSO}$, as a function of $f_\mathrm{low}$.
 For Figures (a) through (c), the line style indicates which noise model was used.}

### 2012-04-03 article.tex
Increased DCC version number
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1486 ccdd05d5-56c1-461a-bf98-85f6876523b4
M -> Mt
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1488 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \begin{equation} \label{eq:fgw}
-	f(t) = \frac{1}{\pi \mathcal{M}}
-		\left[ \frac{5}{256}\frac{\mathcal{M}}{t} \right]^{3/8}
+	f(t) = \frac{1}{\pi \mathcal{M}_\mathrm{t}}
+		\left[ \frac{5}{256}\frac{\mathcal{M}_\mathrm{t}}{t} \right]^{3/8}
 \end{equation}
 %
-where $\mathcal{M} = M^{2/5} \mu^{3/5}$ is the chirp mass of the binary in units of time ($G
-M_\odot / c^3 \approx 5$~$\upmu$s), $M$ is the total mass, and $\mu$ is the reduced mass. 
+where $\mathcal{M}=M^{2/5} \mu^{3/5}$ is the chirp mass of the binary, $\mathcal{M}_\mathrm{t}=G \mathcal{M} / c^3$ is the chirp mass in units of time, $M$ is the total mass, and $\mu$ is the reduced mass.
 The expected value of the single detector \SNR\ for an optimally oriented (source at detector's zenith or nadir, orbital plane face-on) inspiral source is~\citep{Abadie:2010p10836}
@@
 	\rho =
-		\frac{\mathcal{M}^{5/6} c}{\pi^{2/3} D}
+		\frac{{\mathcal{M}_\mathrm{t}}^{5/6} c}{\pi^{2/3} D}
 		\sqrt{
@@
 
-This paper has \LIGO\ Document Number {LIGO-P0900004-v32}.
+This paper has \LIGO\ Document Number {LIGO-P0900004-v33}.
 

### 2012-04-03 article.tex
Corrected inappropriate reference
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1489 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 
-In the quadrupole approximation, the instantaneous frequency of the GW inspiral signal is related to the time $t$ relative to coalescence ~\citep{kidder1992, findchirppaper} through
+In the quadrupole approximation, the instantaneous frequency of the GW inspiral signal is related to the time $t$ relative to coalescence ~\citep[section 5.1 of][]{livrev12} through
 %

### 2012-04-03 article.tex
Redefined time slice boundaries in section 3.3.4
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1490 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Substituting approximate values for a template bank designed for component
-masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s,
-$f_\mathrm{low} = 10^1$~Hz, $t_\mathrm{high} = 1/1570$~s, $f_\mathrm{high} = 1570$~Hz,
-and $f^0 = (2)(1570)$~Hz, we find from equation~(\ref{eq:speedup}) that the
+masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $f_\mathrm{high} = f_\mathrm{ISCO} \approx 1570$~Hz, $f^0 = 2 f_\mathrm{ISCO}$, and $t_\mathrm{high} = {f_\mathrm{ISCO}}^{-1}$, we find from equation~(\ref{eq:speedup}) that the
 \lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the

### 2012-04-03 article.tex
Another reference to Living Review
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1491 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 The expected matched-filter SNR, integrated from $f_\mathrm{low}$ to $f_\mathrm{high}$, is given by equation~(\ref{eq:expected-snr}).
-The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the LSO; for non-spinning systems, $f_\mathrm{LSO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary.
+The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the LSO; for non-spinning systems, $f_\mathrm{LSO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary \citep[section 3.4.1 of][]{livrev12}.
 The choice of $f_\mathrm{low}$ is based on the fraction of the total SNR that is accumulated at frequencies above $f_\mathrm{low}$.

### 2012-04-03 article.tex
Added reference for TaylorF2
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1492 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added department affiliation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1493 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \altaffiltext{4}{Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot, Spain} 
-\altaffiltext{5}{University of Texas at Austin, Austin, TX, USA}
+\altaffiltext{5}{Department of Physics, University of Texas at Austin, Austin, TX, USA}
 \altaffiltext{6}{Center for Computational Relativity and Gravitation and School of Mathematical Sciences, Rochester Institute of Technology, Rochester, NY, USA}
@@
 from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO\ noise
-power spectrum~\citep{ALIGONoise}\footnote{\url{http://dcc.ligo.org/cgi-bin/DocDB/ShowDocument?docid=T0900288&version=3}}.  Templates are truncated at 10~Hz, where the projected
+power spectrum~\citep{ALIGONoise}\footnote{\url{http://dcc.ligo.org/cgi-bin/DocDB/ShowDocument?docid=T0900288&version=3}}.
+Waveforms are generated in the frequency domain in the stationary phase approximation at (post)$^{3.5}$-Newtonian order in phase and Newtonian order in amplitude \citep[the TaylorF2 waveforms described in][]{TaylorF2}.
+Templates are truncated at 10~Hz, where the projected
 sensitivity of Advanced \LIGO\ is interrupted by the ``seismic wall.''

### 2012-04-03 article.tex
Added department affiliation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1494 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added department affiliation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1495 ccdd05d5-56c1-461a-bf98-85f6876523b4
Corrected bibliography entry
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1499 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \altaffiltext{9}{Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
-\altaffiltext{10}{Leibniz Universit\"{a}t Hannover, Hannover, Germany}
-\altaffiltext{11}{Ryerson University, Toronto, ON, Canada}
+\altaffiltext{10}{Institut f\"{u}r Gravitationsphysik, Leibniz Universit\"{a}t Hannover, Hannover, Germany}
+\altaffiltext{11}{Department of Chemistry and Biology, Ryerson University, Toronto, ON, Canada}
 
@@
 were operating, several all-sky detection pipelines operated in a low-latency
-configuration to send astronomical alerts, namely Coherent WaveBurst (cWB), Omega, and Multi-Band Template Analysis (\mbta)~\citep{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
+configuration to send astronomical alerts, namely Coherent WaveBurst (cWB), Omega, and Multi-Band Template Analysis \citep[MBTA;][]{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
 cWB and Omega are both unmodeled searches for bursts based on time-frequency decomposition of the GW data.

### 2012-04-04 article.tex
Fixed redline comments
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1510 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 \usepackage{upgreek} % For \upmu, upright mu
+\usepackage{multirow}
 
@@
 \newcommand{\LSO}{LSO}%{{\sc lso}}
-\newcommand{\SNR}{SNR}%{{\sc snr}}
+\newcommand{\SNR}{S/N}%{{\sc snr}}
 \newcommand{\realtime}{real-time}
@@
 \newcommand{\fmax}{\ensuremath{f^0}}
-\newcommand{\flops}{flop/s}
+\newcommand{\flops}{flop~s$^{-1}$}
 \newcommand{\gstlal}{{\tt gstlal}}
@@
 
-\keywords{gamma ray burst: general --- gravitational waves --- methods: data analysis --- methods: numerical}
+\keywords{gamma-ray burst: general --- gravitational waves --- methods: data analysis --- methods: numerical}
 
 \begin{abstract}
-Rapid detection of compact binary coalescence with a network of advanced
+Rapid detection of compact binary coalescence (CBC) with a network of advanced
 gravitational-wave detectors will offer a unique opportunity for
@@
 community might make it possible to observe the onset of electromagnetic
-emission from compact binary coalescence.  We demonstrate a computationally
+emission from CBC.  We demonstrate a computationally
 practical filtering strategy that could produce early-warning triggers before
@@
 it possible to use \GW\ observations as early warning triggers for \EM\
-followup. In the first generation of ground-based laser interferometers, the
+follow-up. In the first generation of ground-based laser interferometers, the
 \GW\ community initiated a project to send alerts when potential \GW\
-transients were observed in order to trigger followup observations by \EM\
+transients were observed in order to trigger follow-up observations by \EM\
 telescopes.  The typical latencies were 30 minutes \citep{HugheyGWPAW2011}.
@@
 Compact binary coalescence (\CBC) is a plausible progenitor for most short
-gamma-ray bursts (short \GRB{}s) \citep{Lee:2005, nakar07}, but the
+gamma-ray bursts \citep[short \GRB{}s;][]{Lee:2005, nakar07}, but the
 association is not iron-clad \citep{2011ApJ...727..109V}. The tidally
@@
 several telescopes were already observing the afterglow of another \GRB\ in
-the same field of view \citep{2008Natur.455..183R}. The observed optical flashes
+the same field of view \citep[FOV;][]{2008Natur.455..183R}. The observed optical flashes
 peaked within tens of seconds and decayed quickly. For short \GRB\ energy
@@
 $130$~Mpc, the radio flux will peak around 9~GHz at $\sim$$5$~mJy, with emission
-in the $R$-band at $\sim$19~mag. Interestingly, roughly a quarter to half of
+in the $R$-band at $\sim$19~mag. Interestingly, roughly a quarter to half of the 
 observed short \GRB{}s also exhibit extended X-ray emission of $30$--$100$~s in
@@
 
-In October 2010, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its
+In 2010 October, \LIGO{}\footnote{\url{http://www.ligo.org/}} completed its
 sixth science run (S6) and Virgo\footnote{\url{http://www.ego-gw.it/}}
@@
 were operating, several all-sky detection pipelines operated in a low-latency
-configuration to send astronomical alerts, namely Coherent WaveBurst (cWB), Omega, and Multi-Band Template Analysis \citep[MBTA;][]{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
+configuration to send astronomical alerts, namely, Coherent WaveBurst (cWB), Omega, and Multi-Band Template Analysis \citep[MBTA;][]{HugheyGWPAW2011, S6lowlatency2, S6lowlatency3, S6lowlatency4}.
 cWB and Omega are both unmodeled searches for bursts based on time-frequency decomposition of the GW data.
@@
 Alerts were sent with latencies of 30--60 minutes, dominated by human vetting.
-Candidates were sent for \EM\ followup to several telescopes; Swift,
+Candidates were sent for \EM\ follow-up to several telescopes; \textit{Swift},
 LOFAR, ROTSE, TAROT, QUEST, SkyMapper,
@@
 gravitational radiation.  Currently, data are calibrated in blocks of 16~s.
-Within $\sim$1~min, data quality is assessed in order to create veto flags.
+Within $\sim$1~minute, data quality is assessed in order to create veto flags.
 These are both technical sources of latency that might be addressed with
@@
 to the human vetting process, this dominated the latency of the entire \EM\
-followup process.  Even if no other sources of latency existed, this trigger
+follow-up process.  Even if no other sources of latency existed, this trigger
 generation latency is too long to catch prompt or even extended emission.
 Low-latency trigger generation will become more challenging with advanced
-detectors because inspiral signals will stay in band up to ten times longer.  In
+detectors because inspiral signals will stay in band up to 10 times longer.  In
 this work, we will focus on reducing this source of latency.
@@
 low latency at the cost of low throughput.  The former are epitomized by the
-overlap-save algorithm for frequency domain (\FD) convolution, currently the
+overlap-save algorithm for frequency-domain (\FD) convolution, currently the
 preferred method in \GW\ searches.  The most obvious example of the latter is
@@
 prohibitively expensive for long templates.
-The computational challenges of low-latency \CBC\ searches are still more daunting for advanced detectors for which the inspiral signal remains in band for a large fraction of an hour~(see appendix~\ref{sec:low-frequency-cutoff}).
+The computational challenges of low-latency \CBC\ searches are still more daunting for advanced detectors for which the inspiral signal remains in band for a large fraction of an hour~(see the Appendix).
 
@@
 Final triggers are built from the reconstructed matched filter output.
-Another novel approach using networks of parallel, second order infinite impulse response~(IIR) filters is being explored
-by \citet{shaunIIR, linqingIIR}.
+Another novel approach using networks of parallel, second-order infinite impulse response~(IIR) filters is being explored
+by \citet{shaunIIR} and \citet{linqingIIR}.
 
@@
 early-warning detection.  Then, we provide an overview of our novel
-method for detecting compact binary coalescence signals near real-time.
+method for detecting CBC signals near real-time.
 We then describe a prototype implementation using open source signal
@@
 
-\section{Prospects for \earlywarning\ detection and \EM\ followup}
+\section{Prospects for \earlywarning\ detection and \EM\ follow-up}
 
@@
 before coalescence.  The heavy solid line corresponds to the most probable yearly rate
-estimate from~\citet{Abadie:2010p10836}.  The shaded region represents the 5\% to 95\% confidence interval
+estimate from~\citet{Abadie:2010p10836}.  The shaded region represents the 5\%--95\% confidence interval
 arising from substantial uncertainty in predicted event rates.}
@@
 Before the \GW\ signal leaves the detection band, we can imagine examining the
-signal to noise ratio (\SNR) accumulated up to that point and if it is
+signal-to-noise ratio (\SNR) accumulated up to that point and if it is
 already significant, release an alert immediately, trading \SNR\ and sky
@@
 	f(t) = \frac{1}{\pi \mathcal{M}_\mathrm{t}}
-		\left[ \frac{5}{256}\frac{\mathcal{M}_\mathrm{t}}{t} \right]^{3/8}
+		\left[ \frac{5}{256}\frac{\mathcal{M}_\mathrm{t}}{t} \right]^{3/8},
 \end{equation}
@@
 where $\mathcal{M}=M^{2/5} \mu^{3/5}$ is the chirp mass of the binary, $\mathcal{M}_\mathrm{t}=G \mathcal{M} / c^3$ is the chirp mass in units of time, $M$ is the total mass, and $\mu$ is the reduced mass.
-The expected value of the single detector \SNR\ for an optimally oriented (source at detector's zenith or nadir, orbital plane face-on) inspiral source is~\citep{Abadie:2010p10836}
+The expected value of the single-detector \SNR\ for an optimally oriented (source at detector's zenith or nadir, orbital plane face-on) inspiral source is~\citep{Abadie:2010p10836}
 %
@@
 			\frac{5}{6} \int_{f_\mathrm{low}}^{f_\mathrm{high}}
-			\frac{f^{-7/3}}{S(f)} \mathrm{d}f}
+			\frac{f^{-7/3}}{S(f)} \mathrm{d}f},
 \end{equation}
@@
 where $D$ is the luminosity distance and $S(f)$ is the one-sided power spectral density of the detector noise.
-$f_\mathrm{low}$ and $f_\mathrm{high}$ are low and high frequency limits of integration which may be chosen to extend across the entire bandwidth of the detector.
-If we want to trigger at a time $t$ before merger, then we must cut off the SNR integration at $f_\mathrm{high} = f(t)$ with $f(t)$ given by equation~(\ref{eq:fgw}) above.
+$f_\mathrm{low}$ and $f_\mathrm{high}$ are low- and high- frequency limits of integration which may be chosen to extend across the entire bandwidth of the detector.
+If we want to trigger at a time $t$ before merger, then we must cut off the SNR integration at $f_\mathrm{high} = f(t)$ with $f(t)$ given by Equation~(\ref{eq:fgw}) above.
 
@@
 \NS--\NS\ merger rates estimated in \citet{Abadie:2010p10836}.  The merger rates
-have substantial measurement uncertainty due to the small sample of known double pulsar systems that will merge within a Hubble time; they also have systematic uncertainty due sensitive dependence on the pulsar luminosity distribution function~\citep{KalogeraRates}.
+have substantial measurement uncertainty due to the small sample of known double pulsar systems that will merge within a Hubble time; they also have systematic uncertainty due to sensitive dependence on the pulsar luminosity distribution function~\citep{KalogeraRates}.
 The most probable estimates indicate that at a single-detector \SNR\ threshold of 8, we will
@@
 \begin{table}[h]
-\caption{\label{table:sky-localization-accuracy}Horizon distance, \SNR\ at
-merger, and area of 90\% confidence at selected times before merger for sources
-with expected detectability rates of 40, 10, 1, and 0.1~yr$^{-1}$.
-A dash (-----) signifies that the confidence area is omitted because at the indicated time the SNR would not have crossed the detection threshold of 8.}
+\caption{\label{table:sky-localization-accuracy}Horizon Distance, \SNR\ at
+Merger, and Area of 90\% Confidence at Selected Times Before Merger for Sources
+with Expected Detectability Rates of 40, 10, 1, and 0.1~yr$^{-1}$.}
 \begin{center}
@@
 \tableline\tableline
-rate & horiz. & final & \multicolumn{4}{c}{$A$(90\%) (deg$^2$)} \\
+Rate & Horizon & Final & \multicolumn{4}{c}{$A$(90\%) (deg$^2$)} \\
 \cline{4-7}
@@
 \end{tabular}
+\tablecomments{A dash (-----) signifies that the confidence area\\is omitted because at the indicated time the SNR would\\not have crossed the detection threshold of 8.}
 \end{center}
@@
 
-\EM\ followup requires estimating the location of the \GW\ source. The localization
+\EM\ follow-up requires estimating the location of the \GW\ source. The localization
 uncertainty can be estimated from the uncertainty in the time of arrival of the \GW{}s,
@@
 \citeauthor{Fairhurst2009} also cautions that his Fisher matrix calculation fails to capture disconnected patches of probability, which occur prominently in networks of three detectors where there are generally two local maxima on opposite sides of the plane of the detectors.
-Aside from the mirror degeneracy, characterizing the uncertainty region by the Fisher matrix alone tends to overestimate, rather than underestimate, the area for low SNR events, but this effect is generally more than compensated by the source being in an unfavorable sky location.
-For these reasons, the localization uncertainty estimated from timing is highly optimistic and will only suffice for an order of magnitude estimate.
+Aside from the mirror degeneracy, characterizing the uncertainty region by the Fisher matrix alone tends to overestimate, rather than underestimate, the area for low-\SNR\ events, but this effect is generally more than compensated by the source being in an unfavorable sky location.
+For these reasons, the localization uncertainty estimated from timing is highly optimistic and will only suffice for an order-of-magnitude estimate.
 Once per year, we expect to observe an
@@
 It is unfeasible to search hundreds of square degrees for a prompt counterpart.
-For comparison to some examples of modern ground-based wide-field survey instruments, the Palomar Transient Factory P48 ~\citep{2010SPIE.7735E.122L} has a $3.50 \times 2.31$~deg$^2$ field of view~(FOV); the Pan-STARRS P1~\citep{2002SPIE.4836..154K} has a 7~deg$^2$ FOV.
-Even the eagerly awaited LSST will have a FOV of 9.6~deg$^2$ \citep{2008arXiv0805.2366I}.
+For comparison to some examples of modern ground-based wide-field survey instruments, the Palomar Transient Factory P48 ~\citep{2010SPIE.7735E.122L} has a $3.50 \times 2.31$~deg$^2$ FOV; the Pan-STARRS P1~\citep{2002SPIE.4836..154K} has a 7~deg$^2$ FOV.
+Even the eagerly awaited LSST will have an FOV of 9.6~deg$^2$ \citep{2008arXiv0805.2366I}.
 However,
@@
 
-In this section we describe a decomposition of the \CBC\ signal space that
+In this section, we describe a decomposition of the \CBC\ signal space that
 reduces \TD\ filtering cost sufficiently to allow for the
@@
 banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.
-Suppose that the observed data $x[k]$ consists of a known, nominal signal $s[k]$ and additive, zero-mean noise $n[k]$,
+Suppose that the observed data $x[k]$ consists of a known, nominal signal $s[k]$, and additive, zero-mean noise $n[k]$
 %
@@
 where $N$ is the length of the filter and $f^0$ is the sample rate.
-(In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f^0 / 2$, or alternatively, that the data is low-pass filtered prior to matched filtering.)  The discrete Fourier transform of the output is
+(In order to satisfy the Nyquist-Shannon sampling criterion, it is assumed that the detector's continuous noise power spectral density $S(f)$ vanishes for all $f > f^0 / 2$, or alternatively, that the data are low-pass filtered prior to matched filtering.)  The DFT of the output is
 %
@@
 %
-The placement of parentheses in equation~(\ref{eq:matched-filter-fd}) emphasizes that the matched filter can be thought of as a cross-correlation of a whitened version of the data with a whitened version of the nominal signal.
+The placement of parentheses in Equation~(\ref{eq:matched-filter-fd}) emphasizes that the matched filter can be thought of as a cross-correlation of a whitened version of the data with a whitened version of the nominal signal.
 In this paper, we shall not describe the exact process by which the detector's noise power spectrum is estimated and deconvolved from the data; for the remainder of this paper we shall define $x[k]$ as the \emph{whitened} data stream.
@@
 
-Equation~(\ref{eq:SNRTD}) can be implemented in the time-domain as a bank of
+Equation~(\ref{eq:SNRTD}) can be implemented in the TD as a bank of
 finite impulse response (\fir) filters, requiring $\mathcal O(\numtmps
@@
 much more computationally efficient to use the convolution theorem and the
-fast Fourier transform (\fft) to implement fast convolution in the frequency
-domain, requiring only
+fast Fourier transform (\fft) to implement fast convolution in the FD, requiring only
 $\mathcal O(\numtmps \lg \tmpsamps)$ operations per sample but incurring
@@
 Here we describe a method for reducing the computational cost of a \TD\ search
-for compact binary coalescence.  We give a zero latency, real-time algorithm
+for CBC.  We give a zero latency, real-time algorithm
 that competes in terms of floating point operations per second with the
@@
 However, the resulting filters are still not
-orthogonal across the parameter space, and are in fact highly redundant.
+orthogonal across the parameter space and are in fact highly redundant.
 We use the \SVD\ to approximate the template bank by a set of orthogonal
@@
 transformations combined reduce the number of floating point operations
-to a level that is competitive with the conventional high-latency \FD\
-matched filter approach.  In the remainder of this section we describe the
+to a level that is competitive with the conventional high-latency \FD-matched filter approach.  In the remainder of this section we describe the
 \lloid\ algorithm in detail and provide some basic computational cost scaling.
@@
 The first step of our proposed method is to divide the templates into time
-slices in a \TD\ analogue to the \FD\ decomposition employed by MBTA~\citep{Marion2004,Buskulic2010}.
+slices in a \TD\ analog to the \FD\ decomposition employed by MBTA~\citep{Marion2004,Buskulic2010}.
 The application to GW data analysis is foreshadowed by an earlier \FD\ convolution algorithm, proposed by \citet{gardner1995efficient}, based on splitting the impulse response of a filter into smaller blocks.
@@
 For concreteness and simplicity, consider an inspiral waveform in the
-quadrapole approximation, for which the time-frequency relation is given by equation~(\ref{eq:fgw}).
+quadrupole approximation, for which the time-frequency relation is given by Equation~(\ref{eq:fgw}).
 This monotonic time-frequency relationship allows us
@@
 the bandwidth of the filters across the entire template bank can be significantly less
-than the full sample rate at which data is acquired.
+than the full sample rate at which data are acquired.
 
@@
 
-The time sliced templates can then be downsampled in each interval without
+The time-sliced templates can then be downsampled in each interval without
 aliasing, so we define them as
@@
 %
-We note that the time slice decomposition in equation~(\ref{eq:time-slices}) is
+We note that the time slice decomposition in Equation~(\ref{eq:time-slices}) is
 manifestly orthogonal since the time slices are disjoint in time.  In the next
-section we examine how to reduce the number of filters within each time slice
+section, we examine how to reduce the number of filters within each time slice
 via \SVD\ of the time-sliced templates.
@@
 particular minimal match.  A similar technique can be applied to the time-sliced
-templates as defined in equation~\ref{eq:time-sliced-templates} above.  The \SVD\
+templates as defined in Equation~(\ref{eq:time-sliced-templates}) above.  The \SVD\
 is a matrix factorization that takes the form
@@
 
-In the next section we describe how we form our early-warning detection
+In the next section, we describe how we form our early-warning detection
 statistic using the time slice decomposition and the \SVD.
@@
 rates with each of the time slices.  We will denote the decimated detector data
-streams using a superscript $^s$ to indicate the time slices to which they
+streams using a superscript ``$s$'' to indicate the time slices to which they
 correspond.  The operator $H^\shortdownarrow$ will represent the appropriate
@@
 From the combination of the time slice decomposition in
-equation~(\ref{eq:time-sliced-templates}) and the \SVD\ defined in
-equation~(\ref{eq:svddecomp}), we define the early-warning output accumulated
+Equation~(\ref{eq:time-sliced-templates}) and the \SVD\ defined in
+Equation~(\ref{eq:svddecomp}), we define the early-warning output accumulated
 up to time slice $s$ using the recurrence relation,
@@
 			\left(H^\uparrow \rho_i^{s+1}\right)[k]
-		}^\textrm{\clap{{\sc snr} from previous time slices}} \\
+		}^\textrm{\clap{S/N from previous time slices}} \\
 		% Plus ...
@@
 \protect\includegraphics{f3d} stand for \fir\ filters.  Sample
-rate decreases from the top of the diagram to the bottom.  In this diagram each
+rate decreases from the top of the diagram to the bottom.  In this diagram, each
 time slice contains three \fir\ filters that are linearly combined to produce
-four output channels.  In a typical pipeline the number of \fir\ filters is
+four output channels.  In a typical pipeline, the number of \fir\ filters is
 much less than the number of output channels.}
@@
 Ultimately, the latency of the entire \lloid\ algorithm is set by the decimation
-and interpolation filters because they are generally time-symmetric and slightly
+and interpolation filters because they are generally time symmetric and slightly
 acausal.  Fortunately, as long as the latency introduced by the decimation and interpolation
@@
 A decimation filter will be necessary to convert the 4096~Hz input signal $x[k] \equiv x^0[k]$ to the 512~Hz input $x^1[k]$, and an interpolation filter will be necessary to match the sample rates of the two early-warning outputs.
-In this example, as long as the decimation and interpolation filters are together acausal by less than $t^1 = 0.5$~s, the total SNR $\rho_i^0[k]$ will be available with a latency of zero samples.
+In this example, as long as the decimation and interpolation filters are together acausal by less than $t^1 = 0.5$~s, the total \SNR\ $\rho_i^0[k]$ will be available with a latency of zero samples.
 When zero latency is important, we may take this as a requirement for the decimation and interpolation filter kernels.
 
-In the next section we compute the expected computational cost scaling of this
+In the next section, we compute the expected computational cost scaling of this
 decomposition and compare it with the direct \TD\ implementation of
-\eqref{eq:SNRTD} and higher latency blockwise \FD\ methods.
+Equation~\eqref{eq:SNRTD} and higher latency blockwise \FD\ methods.
 
@@
 \begin{table}
-\caption{\label{tab:recap}Notation used to describe filters.}
+\caption{\label{tab:recap}Notation Used to Describe Filters.}
 \begin{center}
@@
 \tableline
-$f^s$		& sample rate in time slice $s$ \\
-\numtmps		& num. templates \\
-\tmpsamps	& num. samples per template \\
-\numslices	& num. time slices \\
-\numsvdtmps	& num. basis templates in time slice $s$ \\
-\slicessamps	& num. samples in decimated time slice $s$\\
-$N^\shortdownarrow$ & length of decimation filter \\
-$N^\shortuparrow$ & length of interpolation filter \\
+$f^s$		& Sample rate in time slice $s$ \\
+\numtmps		& Number of templates \\
+\tmpsamps	& Number of samples per template \\
+\numslices	& Number of time slices \\
+\numsvdtmps	& Number of basis templates in time slice $s$ \\
+\slicessamps	& Number of samples in decimated time slice $s$\\
+$N^\shortdownarrow$ & Length of decimation filter \\
+$N^\shortuparrow$ & Length of interpolation filter \\
 \tableline
@@
 samples that will not be deleted.  Consequently, an efficient decimator 
-requires only $2 N^\shortdownarrow f^{s-1}$ \flops.  (One common realization is an ingenious structure called a \emph{polyphase decimator}, described in chapter 1 of \citet{jovanovic2002multirate}.)
+requires only $2 N^\shortdownarrow f^{s-1}$ \flops.  (One common realization is an ingenious structure called a \emph{polyphase decimator}, described in Chapter 1 of \citet{jovanovic2002multirate}.)
 
@@
 slices at the same sample rate in order to keep the size of matrices that enter
-the singular value decomposition manageable.  Here we have assumed that the
+the SVD manageable.  Here we have assumed that the
 decimation filters are connected in parallel, converting from the base sample
@@
 evolves slowly enough in time, then we can use the time-frequency relationship
-of equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional, direct
+of Equation~(\ref{eq:fgw}) to estimate the speedup relative to the conventional, direct
 \TD\ method.  The reduction in \flops\ is approximately
@@
 so that we can approximate the time slice sample rate as twice the instantaneous
-gravitational wave frequency, $f^s \approx 2 f(t)$, and the number of samples in
+GW frequency, $f^s \approx 2 f(t)$, and the number of samples in
 the decimated time slice as the sample rate times an infinitesimally short time
 interval, $\slicessamps \approx 2 f(t) \, \mathrm{d}t$. The integral is
-evaluated using the power-law form of $f(t)$ from equation~(\ref{eq:fgw}).
+evaluated using the power-law form of $f(t)$ from Equation~(\ref{eq:fgw}).
 Substituting approximate values for a template bank designed for component
-masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $f_\mathrm{high} = f_\mathrm{ISCO} \approx 1570$~Hz, $f^0 = 2 f_\mathrm{ISCO}$, and $t_\mathrm{high} = {f_\mathrm{ISCO}}^{-1}$, we find from equation~(\ref{eq:speedup}) that the
+masses around (1.4, 1.4) $M_\odot$, $\alpha \approx 10^{-2}$, $t_\mathrm{low} = 10^3$~s, $f_\mathrm{low} = 10^1$~Hz, $f_\mathrm{high} = f_\mathrm{ISCO} \approx 1570$~Hz, $f^0 = 2 f_\mathrm{ISCO}$, and $t_\mathrm{high} = {f_\mathrm{ISCO}}^{-1}$, we find from Equation~(\ref{eq:speedup}) that the
 \lloid\ method requires only $\sim 10^{-6}$ times as many \flops\ as the
@@
 In this section we describe an implementation of the \lloid\ method described
-in section \ref{sec:method} suitable for rapid \GW\
+in Section \ref{sec:method} suitable for rapid \GW\
 searches for \CBC{}s.  The \lloid\ method requires several
-computations that can be completed before the analysis is underway.  Thus
+computations that can be completed before the analysis is underway.  Thus, 
 we divide the procedure into an offline planning stage and an
@@
 \SVD\ and is the approach considered in \citet{Cannon:2010p10398}.  Next, we choose
-time-slice boundaries as in equation~\eqref{eq:time-sliced-templates} such that all
+time slice boundaries as in Equation~\eqref{eq:time-sliced-templates} such that all
 of the templates within a sub-bank are sub-critically sampled at progressively lower
@@
 reconstruction matrix that maps them back to the original templates as
-described in equation~\eqref{eq:svddecomp}.  The downsampled basis templates,
-the reconstruction matrix, and the time-slice boundaries are all saved to disk.
+described in Equation~\eqref{eq:svddecomp}.  The downsampled basis templates,
+the reconstruction matrix, and the time slice boundaries are all saved to disk.
 
@@
 
-The \lloid\ algorithm is amenable to latency free, real-time implementation.
+The \lloid\ algorithm is amenable to latency-free, real-time implementation.
 However, a real-time search pipeline would require integration directly into the data acquisition and storage systems of the \LIGO\ observatories.
 A slightly more
-modest goal is to leverage existing low-latency, but not real-time, signal processing software in order to implement
+modest goal is to leverage existing low latency, but not real-time, signal processing software in order to implement
 the \lloid\ algorithm.
@@
 \gstreamer\ pipelines typically operate with very low (in some consumer
-applications, imperceptibly low) latency rather than in true real-time because
+applications, imperceptibly low) latency rather than in true real time because
 signals are partitioned into blocks of samples, or \emph{buffers}.  This affords
@@
 buffers sufficiently small.
-In any case, buffering is a reasonable strategy for low latency \LIGO\ data analysis because, as we previously remarked, the \LIGO\ data acquisition system has a granularity of $1/16$~s.
+In any case, buffering is a reasonable strategy for low-latency \LIGO\ data analysis because, as we previously remarked, the \LIGO\ data acquisition system has a granularity of $1/16$~s.
 
@@
 sensitivity of Advanced \LIGO\ is interrupted by the ``seismic wall.''
-This results in a grid of 98544 points, or
-$2 \times 98544 = 197088$~templates.  Then we create sub-banks by partitioning
+This results in a grid of 98,544 points, or
+$2 \times 98,544 = 197,088$~templates.  Then we create sub-banks by partitioning
 the parameter space by chirp mass.  Figure \ref{fig:tmpltbank} illustrates this
@@
 	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
-case study, consisting of component masses $m_1$, $m_2$, between 1 and 3~$M_\odot$, and
+case study, consisting of component masses $m_1$ and $m_2$, between 1 and 3~$M_\odot$, and
 chirp masses $\mathcal{M}$ between 1.1955 and 1.2045~$M_\odot$.}
 \end{figure}
-With this sub-bank we are able to construct an efficient time-slice decomposition
+With this sub-bank we are able to construct an efficient time slice decomposition
 that consists of 11 time slices with sample rates between 32 and 4096 Hz summarized
@@
 \begin{table*}
-\caption{\label{tab:time_slices} Filter design sub-bank of 1314 templates.  From
+\caption{\label{tab:time_slices} Filter Design Sub-Bank of 1314 Templates.}
+\begin{center}
+\begin{tabular}{crr@{,\,}lc*{6}{r}}
+\tableline\tableline
+\\ [-2ex]
+& $f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD\ tolerance)} \\% [1ex]
+\cline{6-11}
+\\[-2.5ex]
+& (Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline
+\multirow{11}{*}{\includegraphics{t3}} & 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\[0.5em]
+& 512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\[0.5em]
+& 256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\[0.5em]
+& 128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\[0.5em]
+& 64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\[0.5em]
+& 64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\[0.5em]
+& 64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\[0.5em]
+& 32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\[0.5em]
+& 32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\[0.5em]
+& 32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\[0.5em]
+& 32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\[0.5em]
+\tableline
+\end{tabular}
+\tablecomments{From
 left to right, this table shows the sample rate, time interval, number of
@@
 tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
-\begin{center}
-\begin{minipage}[c]{0.4\textwidth}
-\includegraphics{t3}
-\end{minipage}
-\begin{minipage}[c]{0.55\textwidth}
-\begin{tabular}{rr@{,\,}lc*{6}{r}}
-\tableline\tableline
-\\ [-2ex]
-$f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD\ tolerance)} \\% [1ex]
-\cline{5-10}
-\\[-2.5ex]
-(Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline
-4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
-512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
-256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
-128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
-64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
-64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
-64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
-32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
-32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
-32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
-32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
-\tableline
-\end{tabular}
-\end{minipage}
 \end{center}
@@
 the truncation of the \SVD\ at $L^s < M$ basis templates.  As remarked upon in
-\citet{Cannon:2010p10398} and section~\ref{sec:svd}, this effect is measured by
+\citet{Cannon:2010p10398} and Section~\ref{sec:svd}, this effect is measured by
 the \SVD\ tolerance.  The second comes from the limited bandwidth of the
@@
 minimum, maximum, and median mismatch are shown as functions of \SVD\ tolerance
-in Figure~\ref{fig:bw}a.  As the \SVD\ tolerance increases toward 1, the \SVD\
+in Figure~\ref{fig:bw}(a).  As the \SVD\ tolerance increases toward 1, the \SVD\
 becomes an exact matrix factorization, but the computational cost increases as
@@
 mismatch.  We use GStreamer's stock
-\texttt{audioresample} element, which provides a FIR decimation filter with
+\texttt{audioresample} element, which provides an FIR decimation filter with
 a Kaiser-windowed sinc function kernel.  The mismatch as a function of $N^\shortuparrow$ is shown in
-Figure~\ref{fig:bw}b.  The mismatch saturates at $\sim$$2 \times 10^{-4}$ with
+Figure~\ref{fig:bw}(b).  The mismatch saturates at $\sim$$2 \times 10^{-4}$ with
 $N^\shortuparrow = 64$.  We find that a filter length of 16 is sufficient to
@@
 
-However, the inspiral templates themselves consist of the gravitational wave time series convolved with the impulse response of the whitening filter.
+However, the inspiral templates themselves consist of the GW time series convolved with the impulse response of the whitening filter.
 As a consequence, the \lloid\ algorithm must faithfully replicate the effect of the whitening filter.
@@
 
-The usual statement of the famous Nyquist-Shannon theorem, stated below as Theorem~\ref{thm:nyquist}, has a natural dual, \ref{thm:nyquist-dual}, that addresses the frequency resolution that can be achieved with a \fir\ filter of a given length.
+The usual statement of the famous Nyquist-Shannon theorem, stated below as Theorem~\ref{thm:nyquist}, has a natural dual, Theorem~\ref{thm:nyquist-dual}, that addresses the frequency resolution that can be achieved with an \fir\ filter of a given length.
 
@@
 	\label{thm:nyquist}
-	\citep[after][p. 518]{oppenheim1997signals}
+	\citep[After][p. 518]{oppenheim1997signals}
 	Let $x(t)$ be a band-limited signal with continuous Fourier transform $\tilde{x}(f)$ such that $\tilde{x}(f') = 0 \; \forall \; f' : |f'| > f_M$.
-Then $x(t)$ is uniquely determined by its discrete samples $x(n/f^0)$, $n = 0, \pm 1, \pm 2, \dots$, if $f^0 > 2 f_M$.
+Then, $x(t)$ is uniquely determined by its discrete samples $x(n/f^0)$, $n = 0, \pm 1, \pm 2, \dots$, if $f^0 > 2 f_M$.
 \end{thm}
@@
 Failure to meet the conditions of this dual of the sampling theorem results in spectral leakage.
-For a time-domain signal to capture spectral features that are the size of the central lobe of the $\sinc$ function, the signal must have a duration greater than $1/\Delta f$.
+For a TD signal to capture spectral features that are the size of the central lobe of the $\sinc$ function, the signal must have a duration greater than $1/\Delta f$.
 If the signal $x(t)$ is truncated by sampling it for a shorter duration, then its Fourier transform becomes smeared out; conceptually, power ``leaks'' out into the side lobes of the $\sinc$ functions and washes away sharp spectral features.
-In the \GW\ data analysis literature, the synthesis of inspiral matched filters involves a step called \emph{inverse spectrum truncation} \citep[see][section VII]{findchirppaper} that fixes the number of coefficients based on the desired frequency resolution.
+In the \GW\ data analysis literature, the synthesis of inspiral matched filters involves a step called \emph{inverse spectrum truncation} \citep[see][Section VII]{findchirppaper} that fixes the number of coefficients based on the desired frequency resolution.
 
 In order to effectively flatten a line in the detector's noise power
-spectrum, the time scale of the templates must be at least as long as
+spectrum, the timescale of the templates must be at least as long as
 the damping time $\tau$ of the line, $\tau = 2 Q / \omega_0$, where $Q$ is the quality
 factor of the line and $w_0$ is the central angular frequency.
-To put this into the context of the sampling theorem, in order to resolve a notch with a particular $Q$ and $f_0$, a \fir\ filter must achieve a frequency resolution of $\Delta f \gtrsim \pi f_0 / Q$ and therefore its impulse response must last for at least a time $1/\Delta f = Q / \pi f_0$.
+To put this into the context of the sampling theorem, in order to resolve a notch with a particular $Q$ and $f_0$, an \fir\ filter must achieve a frequency resolution of $\Delta f \gtrsim \pi f_0 / Q$ and therefore its impulse response must last for at least a time $1/\Delta f = Q / \pi f_0$.
 For example,
-in the S6 detector configuration known as ``Enhanced LIGO,'' the violin modes \citep{ELIGOSusp} had $Q \sim 10^5$ and $\omega_0 \sim (2 \pi) 340$~rad/s, for a coherence time $\tau \sim 10^2$~s.
+in the S6 detector configuration known as ``Enhanced LIGO,'' the violin modes \citep{ELIGOSusp} had $Q \sim 10^5$ and $\omega_0 \sim (2 \pi) 340$~rad~s$^{-1}$, for a coherence time $\tau \sim 10^2$~s.
 
@@
 If the detector data were passed through the whitening filter twice, then time-sliced filters need not depend on the detector's noise power spectral density at all.
-In such a variation on the \lloid\ method, the basis filters could be calculated from the \emph{weighted} \SVD~\citep[chapter 3.6]{WeightedSVD, jackson2003user} of the time-sliced templates, using the covariance of the detector noise as a weight matrix.
+In such a variation on the \lloid\ method, the basis filters could be calculated from the \emph{weighted} \SVD~\citep[Chapter 3.6]{WeightedSVD, jackson2003user} of the time-sliced templates, using the covariance of the detector noise as a weight matrix.
 
@@
 \begin{table*}
-\caption{\label{table:flops}Computational cost in \flops\ and latency in seconds
-of the direct \TD\ method, the overlap-save \FD\ method, and \lloid.  Cost is given
-for both the sub-bank described in section~\ref{sec:bank-setup} and a full
-1--3~$M_\odot$ \NS--\NS\ search.  The last column gives the approximate number of machines per
-detector required for a full Advanced LIGO \NS--\NS\ search.}
+\caption{\label{table:flops}Computational Cost in Flop~s$^{-1}$ and Latency in Seconds
+of the Direct \TD\ Method, the Overlap-Save \FD\ Method, and \lloid.}
 \begin{center}
@@
 \tableline\tableline
-& \flops\ & & \flops\ & number of \\
-method & (sub-bank) & latency (s) & (\NS--\NS) & machines \\[0.1em]
+& Flop~s$^{-1}$ & & Flop~s$^{-1}$ & number of \\
+Method & (Sub-bank) & Latency (s) & (\NS--\NS) & Machines \\[0.1em]
 \tableline
-direct (\TD) & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$3.8\times10^5$ \\
-overlap-save (\FD) & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$5.9$ \\
+Direct (\TD) & $4.9\times10^{13}$ & 0 & $3.8\times10^{15}$ & $\sim$$3.8\times10^5$ \\
+Overlap-save (\FD) & $5.2\times10^8$ & $2\times10^3$ & $5.9\times10^{10}$ & $\sim$$5.9$ \\
 \lloid\ (theory) & $6.6\times10^8$ & 0 & $1.1 \times 10^{11}$ & $\sim$$11$ \\
@@
 \end{center}
+\tablecomments{Cost is given
+for both the sub-bank described in Section~\ref{sec:bank-setup} and a full
+1--3~$M_\odot$ \NS--\NS\ search.  The last column gives the approximate number of machines per
+detector required for a full Advanced LIGO \NS--\NS\ search.}
 \end{table*}
@@
 Assuming that other regions of the parameter space have similar computational
-scaling, an entire single detector search for \NS--\NS\ signals in the
+scaling, an entire single-detector search for \NS--\NS\ signals in the
 1--3~$M_\odot$ component mass range could be implemented with $(98544/657)\approx150$ times the cost, or $9.9 \times 10^{10}$~\flops.
@@
 tabulating the number of floating point operations using
-expression~(\ref{eq:lloid-flops}).  This should be a much more accurate measure
+Expression~(\ref{eq:lloid-flops}).  This should be a much more accurate measure
 because template length varies over the parameter space.  Lower chirp mass
@@
 The scaling argument projects that a full \FD\ search would require $7.8 \times 10^{10}$~\flops.
-The direct calculation from expression~(\ref{eq:fd-flops}) gives $5.9 \times 10^{10}$~\flops, in order-of-magnitude agreement.
-In this application, the conventional \FD\ search is scarcely a factor of 2 faster than LLOID while gaining only $0.3$\% in \SNR, but only at the price of thousands of seconds of latency.
+The direct calculation from Expression~(\ref{eq:fd-flops}) gives $5.9 \times 10^{10}$~\flops, in order-of-magnitude agreement.
+In this application, the conventional \FD\ search is scarcely a factor of two faster than LLOID while gaining only $0.3$\% in \SNR, but only at the price of thousands of seconds of latency.
 
@@
 and even early-warning detection of \GW{}s emitted during the coalescence
-of neutron stars and stellar-mass black holes.  It is one part of a complicated
+of \NS{}s and stellar-mass black holes.  It is one part of a complicated
 analysis and observation strategy that will unfortunately have other sources of
@@
 technical sources of \GW\ observation latency and encourage the possibility of
-even more rapid \EM\ followup observations to catch prompt emission in the
+even more rapid \EM\ follow-up observations to catch prompt emission in the
 advanced detector era.
@@
 latency free, real-time operation.  We anticipate requiring $\gtrsim$40
-modern multi-core computers to search for binary neutron stars using
-coincident \GW\ data from a 4-detector network.  In the future, additional
-computational savings could be acheived by conditionally reconstructing the
-\SNR\ time-series only during times when a composite detection statistic
+modern multi-core computers to search for binary \NS{}s using
+coincident \GW\ data from a four-detector network.  In the future, additional
+computational savings could be achieved by conditionally reconstructing the
+\SNR\ time series only during times when a composite detection statistic
 crosses a threshold~\citep{svd-compdetstat}.  However, the anticipated required
@@
 We have shown a prototype implementation of the \lloid\ algorithm using
-\gstreamer, an open source signal processing platform.  Although our prototype
+\gstreamer, an open-source signal processing platform.  Although our prototype
 already achieves latencies of less than one second, further fine tuning may
@@
 for advance detection, we have not yet explored data calibration and whitening,
-triggering, coincidence, and ranking of gravitational wave candidates in a
-framework that supports early \EM\ followup.  One might explore these and also
-using the time-slice decomposition and the \SVD\ to form low-latency
-signal-based vetoes (e.g.~$\chi^2$~statistics) that have been essential for
+triggering, coincidence, and ranking of GW candidates in a
+framework that supports early \EM\ follow-up.  One might explore these and also
+using the time slice decomposition and the \SVD\ to form low-latency
+signal-based vetoes (e.g.,~$\chi^2$~statistics) that have been essential for
 glitch rejection used in previous \GW\ \CBC\ searches.  These additional stages
@@
 rates and detector networks will be necessary in order to fully understand the
-prospects for early-warning detection, localization, and \EM\ followup using
+prospects for early-warning detection, localization, and \EM\ follow-up using
 the techniques we have described.
@@
 Massachusetts Institute of Technology with funding from the National Science
-Foundation and operates under cooperative agreement PHY-0107417.  CH would like
-to thank Ilya Mandel for many discussions about rate estimates and the prospects
-of early detection, Patrick Brady for countless fruitful conversations about low
-latency analysis methods, and John Zweizig for disscussions about \LIGO\ data
-acquisition.  NF would like to thank Alessandra Corsi and Larry Price for
-illuminating discussions on astronomical motivations.  LS would like thank Shaun
+Foundation and operates under cooperative agreement PHY-0107417.  C.H. thanks Ilya Mandel for many discussions about rate estimates and the prospects
+of early detection, Patrick Brady for countless fruitful conversations about low-latency analysis methods, and John Zweizig for discussions about \LIGO\ data
+acquisition.  N.F. thanks Alessandra Corsi and Larry Price for
+illuminating discussions on astronomical motivations.  L.S. thanks Shaun
 Hooper for productive conversations on signal processing.  This research is
 supported by the National Science Foundation through a Graduate Research
-Fellowship to LS and by the Perimeter Institute for Theoretical Physics through
-a fellowship to CH. DK is supported from the Max Planck Gesellschaft.
-MAF is supported by NSF grant PHY-0855494.
+Fellowship to L.S. and by the Perimeter Institute for Theoretical Physics through
+a fellowship to C.H. D.K. is supported from the Max Planck Gesellschaft.
+M.A.F. is supported by NSF Grant PHY-0855494.
 
@@
 
-Ground-based gravitational wave detectors are unavoidably affected at low frequencies by seismic and anthropogenic ground motion.
+Ground-based GW detectors are unavoidably affected at low frequencies by seismic and anthropogenic ground motion.
 The LIGO test masses are suspended from multiple-stage pendula, which attenuate ground motion down to the pole frequency.
 In the detector configuration in place during S6, seismic noise dominated the instrumental background below about 40~Hz.
-Considerable effort is being invested in improving seismic attenuation in Advanced LIGO using active and passive isolation~\citep{0264-9381-27-8-084006}, so that suspension thermal noise will dominate down to 10 to 15~Hz.
-Inspiral waveforms are chirps of infinite duration, but since an interferometric detector's noise diverges at this so-called `seismic wall,' templates for matched filter searches are truncated at a low frequency cutoff $f_\mathrm{low}$ in order to save computational overhead with negligible loss of SNR.
-
-The expected matched-filter SNR, integrated from $f_\mathrm{low}$ to $f_\mathrm{high}$, is given by equation~(\ref{eq:expected-snr}).
-The high frequency cutoff for the inspiral is frequently taken to be the gravitational wave frequency at the LSO; for non-spinning systems, $f_\mathrm{LSO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary \citep[section 3.4.1 of][]{livrev12}.
+Considerable effort is being invested in improving seismic attenuation in Advanced LIGO using active and passive isolation~\citep{0264-9381-27-8-084006}, so that suspension thermal noise will dominate down to 10--15~Hz.
+Inspiral waveforms are chirps of infinite duration, but since an interferometric detector's noise diverges at this so-called ``seismic wall,'' templates for matched filter searches are truncated at a low-frequency cutoff $f_\mathrm{low}$ in order to save computational overhead with negligible loss of SNR.
+
+The expected matched-filter SNR, integrated from $f_\mathrm{low}$ to $f_\mathrm{high}$, is given by Equation~(\ref{eq:expected-snr}).
+The high-frequency cutoff for the inspiral is frequently taken to be the GW frequency at the LSO; for non-spinning systems, $f_\mathrm{LSO} = 4400 (\Msun / M)$~Hz, where $M$ is the total mass of the binary \citep[section 3.4.1 of][]{livrev12}.
 The choice of $f_\mathrm{low}$ is based on the fraction of the total SNR that is accumulated at frequencies above $f_\mathrm{low}$.
-To illustrate the relative contributions to the SNR at different frequencies for a (1.4,~1.4)~$\Msun$ binary, we normalized and plotted the integrand of equation~(\ref{eq:expected-snr}), the noise-weighted power spectral density of the inspiral waveform, in Figure~\ref{fig:low-frequency-cutoff}b.
+To illustrate the relative contributions to the SNR at different frequencies for a (1.4,~1.4)~$\Msun$ binary, we normalized and plotted the integrand of Equation~(\ref{eq:expected-snr}), the noise-weighted power spectral density of the inspiral waveform, in Figure~\ref{fig:low-frequency-cutoff}(b).
 This is the quantity
@@
 which is normalized by the total SNR squared in order to put detectors with different absolute sensitivities on the same footing.
-We used several different noise power spectra: all of the envisioned Advanced LIGO configurations from~\citet{ALIGONoise}; the best achieved sensitivity at LIGO Hanford Observatory (LHO) in LIGO's fifth science run (S5), measured by \citet{S5InspiralRange}; and the best achieved sensitivity at LHO during S6, measured by \citet{S6InspiralRange}.
-(The noise spectra themselves are shown in Figure~\ref{fig:low-frequency-cutoff}a.)  It is plain that during S5 and S6 the greatest contribution to the SNR was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the SNR is accumulated below 60~Hz.
-This information is presented in a complementary way in Figure~\ref{fig:low-frequency-cutoff}c, as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{LSO}$, interpreted as a fraction of the total ``available'' SNR,
+We used several different noise power spectra: all of the envisioned Advanced LIGO configurations from~\citet{ALIGONoise}; the best-achieved sensitivity at LIGO Hanford Observatory (LHO) in LIGO's fifth science run (S5), measured by \citet{S5InspiralRange}; and the best-achieved sensitivity at LHO during S6, measured by \citet{S6InspiralRange}.
+(The noise spectra themselves are shown in Figure~\ref{fig:low-frequency-cutoff}(a).)  It is plain that during S5 and S6 the greatest contribution to the \SNR\ was between 100 and 150~Hz, but for all of the proposed Advanced LIGO configurations the bulk of the \SNR\ is accumulated below 60~Hz.
+This information is presented in a complementary way in Figure~\ref{fig:low-frequency-cutoff}(c), as the square root of the cumulative integral from $f_\mathrm{low}$ to $f_\mathrm{LSO}$, interpreted as a fraction of the total ``available'' \SNR,
 %
@@
 %
-Table~\ref{table:accum_snr} shows the fractional accumulated SNR for four selected low frequency cutoffs, 40~Hz, 30~Hz, 20~Hz, and 10~Hz.
-In S5 and S6, all of the SNR is accumulated above 40~Hz.
-For the `high frequency' Advanced LIGO configuration, scarcely \emph{half} of the SNR is accumulated above 40~Hz.
-For the preferred final configuration, `zero detuning, high power,' 86.1\% of the SNR is above 40~Hz, 93.2\% is above 30~Hz, and 98.1\% is above 20~Hz.
-(Since SNR accumulates in quadrature, this means, on the other hand, that under the `high frequency' configuration a template encompassing \emph{just the early inspiral} from 10 to 40~Hz would accumulate $\sqrt{1 - 0.533^2} \approx 84.6\%$ of the total SNR!
-In the `zero detuning, high power,' configuration, integration from 10 to 40~Hz alone would yield 50.9\% of the total SNR, from 10 to 30~Hz, 36.2\%, and from 10 to 20~Hz, 19.4\%.)
+Table~\ref{table:accum_snr} shows the fractional accumulated \SNR\ for four selected low-frequency cutoffs, 40~Hz, 30~Hz, 20~Hz, and 10~Hz.
+In S5 and S6, all of the \SNR\ is accumulated above 40~Hz.
+For the `high frequency' Advanced LIGO configuration, scarcely \emph{half} of the \SNR\ is accumulated above 40~Hz.
+For the preferred final configuration, `zero detuning, high power,' 86.1\% of the \SNR\ is above 40~Hz, 93.2\% is above 30~Hz, and 98.1\% is above 20~Hz.
+(Since \SNR\ accumulates in quadrature, this means, on the other hand, that under the `high frequency' configuration a template encompassing \emph{just the early inspiral} from 10 to 40~Hz would accumulate $\sqrt{1 - 0.533^2} \approx 84.6\%$ of the total \SNR!
+In the `zero detuning, high power,' configuration, integration from 10 to 40~Hz alone would yield 50.9\% of the total \SNR, from 10 to 30~Hz, 36.2\%, and from 10 to 20~Hz, 19.4\%.)
 
@@
 \begin{center}
+\caption{\label{table:accum_snr}Fractional Accumulated \SNR\ $\rho_\mathrm{frac}(f_\mathrm{low})$ for Four Selected Low Frequency Cutoffs, $f_\mathrm{low}=40$~Hz, 30~Hz, 20~Hz, and 10~Hz.}
 \begin{tabular}{rcccc}
@@
 LHO (best S6) & 100.0 & 100.0 & 100.0 & 100.0 \\
-high frequency & 53.3 & 80.1 & 97.6 & 100.0 \\
-no SRM & 87.8 & 95.1 & 98.7 & 100.0 \\
+High frequency & 53.3 & 80.1 & 97.6 & 100.0 \\
+No SRM & 87.8 & 95.1 & 98.7 & 100.0 \\
 BHBH 20$^\circ$ & 71.1 & 84.2 & 96.2 & 100.0 \\
 NSNS optimized & 91.5 & 96.3 & 99.0 & 100.0 \\
-zero detuning, low power & 67.9 & 80.0 & 93.5 & 100.0 \\
-zero detuning, high power & 86.1 & 93.2 & 98.1 & 100.0 \\
+Zero detuning, low power & 67.9 & 80.0 & 93.5 & 100.0 \\
+Zero detuning, high power & 86.1 & 93.2 & 98.1 & 100.0 \\
 \hline
@@
 \end{center}
-\caption{\label{table:accum_snr}Fractional accumulated SNR $\rho_\mathrm{frac}(f_\mathrm{low})$ for four selected low frequency cutoffs, $f_\mathrm{low}=40$~Hz, 30~Hz, 20~Hz, and 10~Hz.}
 \end{table}
 
-Since the gravitational wave amplitude is inversely proportional to the luminosity distance of the source, and the sensitive volume is proportional to distance cubed, the rate of detectable coalescences depends on the choice of low frequency cutoff.
-An inspiral search that is designed with a low frequency cutoff at the seismic wall would gain an increase in detection rate of $\rho_\mathrm{frac}^{-3}(f_\mathrm{low})$ relative to a search with a low frequency cutoff of $f_\mathrm{low}$.
-This would represent almost a twofold increase in the rate of detection over a search with a fractional accumulated SNR of 80\%, and still a 37\% increase over a search with $\rho_\mathrm{frac} = 90\%$.
-Existing coalescing binary detection pipelines strive to sacrifice no more than 3\% of the available SNR; this forfeits less than a 10\% gain in detection rate.
-In order to satisfy this constraint, the low frequency cutoff would have to be placed below 30~Hz for all of the conceived Advanced LIGO configurations.
-
-The instantaneous gravitational wave frequency, given by equation~(\ref{eq:fgw}) is a power law function of time, so the amount of time for the gravitational wave frequency to evolve from $f_\mathrm{low}$ to $f_\mathrm{LSO}$ depends strongly on $f_\mathrm{low}$.
-The duration of a (1.4,~1.4)~$\Msun$ inspiral is show in Figure~\ref{fig:low-frequency-cutoff}d.
+Since the GW amplitude is inversely proportional to the luminosity distance of the source, and the sensitive volume is proportional to distance cubed, the rate of detectable coalescences depends on the choice of low-frequency cutoff.
+An inspiral search that is designed with a low-frequency cutoff at the seismic wall would gain an increase in detection rate of $\rho_\mathrm{frac}^{-3}(f_\mathrm{low})$ relative to a search with a low-frequency cutoff of $f_\mathrm{low}$.
+This would represent almost a twofold increase in the rate of detection over a search with a fractional accumulated \SNR\ of 80\%, and still a 37\% increase over a search with $\rho_\mathrm{frac} = 90\%$.
+Existing coalescing binary detection pipelines strive to sacrifice no more than 3\% of the available \SNR; this forfeits less than a 10\% gain in detection rate.
+In order to satisfy this constraint, the low-frequency cutoff would have to be placed below 30~Hz for all of the conceived Advanced LIGO configurations.
+
+The instantaneous GW frequency, given by Equation~(\ref{eq:fgw}), is a power law function of time, so the amount of time for the GW frequency to evolve from $f_\mathrm{low}$ to $f_\mathrm{LSO}$ depends strongly on $f_\mathrm{low}$.
+The duration of a (1.4,~1.4)~$\Msun$ inspiral is show in Figure~\ref{fig:low-frequency-cutoff}(d).
 The inspiral takes only 25~s to evolve from 40~Hz to $f_\mathrm{LSO}$, but takes 54~s to evolve from 30~Hz to $f_\mathrm{LSO}$, 158~s from 20~Hz, and 1002~s from 10~Hz.
@@
 
-	\caption{\label{fig:low-frequency-cutoff}From top left:  (a)~Noise amplitude spectral density for a variety of Advanced LIGO noise models, S5, and S6.
-(b)~Normalized signal to noise per unit frequency, $(\mathrm{d}\rho^2/\mathrm{d}f)/\rho^2$, for a (1.4,~1.4)~$\Msun$ inspiral.
-(c)~Percentage of SNR that is accumulated from $f_\mathrm{low}$ to $f_\mathrm{LSO}$, relative to SNR accumulated from $f_\mathrm{low} = 0\,\mathrm{Hz}$ to $f_\mathrm{LSO}$.
+	\caption{\label{fig:low-frequency-cutoff}From top left:  (a)~noise amplitude spectral density for a variety of Advanced LIGO noise models, S5, and S6.
+(b)~Normalized signal-to-noise per unit frequency, $(\mathrm{d}\rho^2/\mathrm{d}f)/\rho^2$, for a (1.4,~1.4)~$\Msun$ inspiral.
+(c)~Percentage of \SNR\ that is accumulated from $f_\mathrm{low}$ to $f_\mathrm{LSO}$, relative to \SNR\ accumulated from $f_\mathrm{low} = 0\,\mathrm{Hz}$ to $f_\mathrm{LSO}$.
 (d)~Amount of time for a \NS--\NS\ inspiral signal to evolve from frequency $f_\mathrm{low}$ to $f_\mathrm{LSO}$, as a function of $f_\mathrm{low}$.
-For Figures (a) through (c), the line style indicates which noise model was used.}
+For (a)--(c), the line style indicates which noise model was used.}
 \end{figure}

### 2012-04-04 article.tex
Added commas
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1511 ccdd05d5-56c1-461a-bf98-85f6876523b4

@@
 Table~\ref{table:flops} shows that the \lloid\ method requires $6.6 \times 10^8$
-\flops\ to cover a sub-bank comprising 657 out of the total 98544 mass pairs.
+\flops\ to cover a sub-bank comprising 657 out of the total 98,544 mass pairs.
 Assuming that other regions of the parameter space have similar computational
 scaling, an entire single-detector search for \NS--\NS\ signals in the
-1--3~$M_\odot$ component mass range could be implemented with $(98544/657)\approx150$ times the cost, or $9.9 \times 10^{10}$~\flops.
+1--3~$M_\odot$ component mass range could be implemented with $(98,544/657)\approx150$ times the cost, or $9.9 \times 10^{10}$~\flops.
 

### 2014-05-06 article.tex
Write images as PDFs

@@
 \begin{figure}[h]
-\plotone{f1}
+\plotone{figures/snr_in_time}
 \caption{\label{fig:earlywarning}Expected number of \NS--\NS\
@@
 \begin{figure}[h]
-\plotone{f2}
+\plotone{figures/localization_uncertainty}
 \caption{\label{fig:sky-localization-accuracy}Area of the 90\% confidence
@@
 	\begin{center}
-		\includegraphics{f3}
+		\includegraphics{figures/lloid-diagram}
 		\caption{\label{fig:pipeline} Schematic of \lloid\ pipeline illustrating
 signal flow.  Circles with arrows represent interpolation
-\protect\includegraphics{f3a} or decimation
-\protect\includegraphics{f3b}.  Circles with plus
+\protect\includegraphics{figures/upsample-symbol} or decimation
+\protect\includegraphics{figures/downsample-symbol}.  Circles with plus
 signs represent summing junctions
-\protect\includegraphics{f3c}.  Squares
-\protect\includegraphics{f3d} stand for \fir\ filters.  Sample
+\protect\includegraphics{figures/adder-symbol}.  Squares
+\protect\includegraphics{figures/fir-symbol} stand for \fir\ filters.  Sample
 rate decreases from the top of the diagram to the bottom.  In this diagram, each
@@
 \begin{figure}[h]
-	\plotone{f4}
+	\plotone{figures/tmpltbank}
 	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
@@
 & (Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline
-\multirow{11}{*}{\includegraphics{t3}} & 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\[0.5em]
+\multirow{11}{*}{\includegraphics{figures/envelope}} & 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\[0.5em]
 & 512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\[0.5em]
@@
 		\begin{center}
-			\includegraphics{f5a}
+			\includegraphics{figures/bw}
 			(a) Mismatch versus \SVD\ tolerance
@@
 		\begin{center}
-			\includegraphics{f5b}
+			\includegraphics{figures/bw_resample}
 			(b) Mismatch versus $N^\shortuparrow$
@@
 \begin{figure}[b]
-	\includegraphics{f6a}%
-	\includegraphics{f6b}
-
-	\includegraphics{f6c}%
-	\includegraphics{f6d}
-
-	\includegraphics{f6e}
+	\includegraphics{figures/asds}%
+	\includegraphics{figures/weighted_asds}
+
+	\includegraphics{figures/accum_snr}%
+	\includegraphics{figures/inspiral_tf_relation}
+
+	\includegraphics{figures/psd_legend}
 

### 2014-05-06 figures/diagram.tex
Write images as PDFs

@@
 \tikzexternalize[force remake=true]
-\tikzset{external/system call={pdflatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"; pdftops -eps "\image".pdf "\image".eps}}
+\tikzset{external/system call={pdflatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"}}
 

### 2014-05-06 article.tex
Switch to emulateapj
Remove unnecessary macro

@@
-\documentclass[preprint2]{aastex}
+\documentclass[iop,apj]{emulateapj}
+\usepackage{apjfonts}
 
 \usepackage[colorlinks]{hyperref}
-\usepackage{epstopdf}
 \usepackage{amsmath,amsthm}
@@
 
-\newcommand{\earlywarning}{early-warning}
 \newcommand{\NS}{NS}
@@
 
-\title{Toward \earlywarning\ detection of gravitational waves from compact binary coalescence}
+\title{Toward early-warning detection of gravitational waves from compact binary coalescence}
 
@@
 
-\section{Prospects for \earlywarning\ detection and \EM\ follow-up}
+\section{Prospects for early-warning detection and \EM\ follow-up}
 
@@
 reduces \TD\ filtering cost sufficiently to allow for the
-possibility of \earlywarning\ detection with modest computing requirements.  We
+possibility of early-warning detection with modest computing requirements.  We
 expand on the ideas of \citet{Marion2004} and \citet{Buskulic2010} that describe a

### 2014-05-06 article.tex
Make two multiline equations single-line
Adjust spacing for emulateapj

@@
 %
-\begin{multline}
+\begin{equation}
 	\rho_i^s [k] =%
@@
 			\left(H^\uparrow \rho_i^{s+1}\right)[k]
-		}^\textrm{\clap{S/N from previous time slices}} \\
+		}^\textrm{\clap{S/N from previous time slices}}
 		% Plus ...
@@
 		}^\textrm{\clap{orthogonal {\sc fir} filters}} .
-\end{multline}
+\end{equation}
 %
@@
 %
-\begin{multline}
+\begin{equation}
 \label{eq:lloid-flops}
-\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s \\ + 2\sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} \left( N^\shortdownarrow f^0 + \numtmps N^\shortuparrow f^{s-1}\right)
-\end{multline}
+\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} \left( N^\shortdownarrow f^0 + \numtmps N^\shortuparrow f^{s-1}\right)
+\end{equation}
 %
@@
 \tableline\tableline
-\\ [-2ex]
+\\ [-1.5ex]
 & $f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD\ tolerance)} \\% [1ex]
 \cline{6-11}
-\\[-2.5ex]
-& (Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline
-\multirow{11}{*}{\includegraphics{figures/envelope}} & 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\[0.5em]
-& 512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\[0.5em]
-& 256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\[0.5em]
-& 128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\[0.5em]
-& 64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\[0.5em]
-& 64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\[0.5em]
-& 64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\[0.5em]
-& 32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\[0.5em]
-& 32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\[0.5em]
-& 32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\[0.5em]
-& 32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\[0.5em]
+\\[-2ex]
+& (Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline\\
+\multirow{11}{*}{\includegraphics{figures/envelope}} & 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\[1em]
+& 512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\[1em]
+& 256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\[1em]
+& 128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\[1em]
+& 64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\[1em]
+& 64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\[1em]
+& 64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\[1em]
+& 32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\[1em]
+& 32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\[1em]
+& 32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\[1em]
+& 32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\[1em]
 \tableline

### 2014-05-07 localization_uncertainty.py
Match fonts in figures

@@ -141,7 +141,7 @@ print r"\tableline"
 print r"\end{tabular}"
 
 
-fig = pylab.figure(figsize=(3,2))
+fig = pylab.figure(figsize=(3.5,2.5))
 ax = fig.add_subplot(1,1,1, adjustable='box')
 
 for rate in (40., 10., 1., 0.1):
@@ -156,7 +156,7 @@ pylab.xlim(0., 1000.)
 pylab.ylim(1e-1, 41253)
 ax.invert_xaxis()
 pylab.grid()
-pylab.ylabel(r'$A$(90%) (deg$^2$)')
+pylab.ylabel(r'$A$(90\%) (deg$^2$)')
 pylab.xlabel(r'time before coalescence, $t$ (s)')
-pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.2,right=0.85)
+pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.15,right=0.875)
 pylab.savefig(sys.argv[1])

### 2014-05-07 snr_in_time.py

@@ -44,7 +44,7 @@ def snr_to_num(fracsnr, number=40):
 	return number * (fracsnr)**3
 
 # Start a new figure
-fig = pylab.figure(figsize=(3,2))
+fig = pylab.figure(figsize=(3.5,2.5))
 
 markers = ['k-', 'k--', 'k:']
 
@@ -98,7 +98,7 @@ pylab.grid()
 #pylab.legend(loc='lower left')
 pylab.ylabel(r'detections yr$^{-1}$')
 pylab.xlabel(r'time before coalescence (s)')
-pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.2,right=0.95)
+pylab.subplots_adjust(bottom=0.2,top=0.95,left=0.15,right=0.95)
 pylab.xlim([1000, 0.])
 pylab.ylim([.1, 1000])
 pylab.savefig(sys.argv[1])

### 2014-05-07 plot_bank.py
Add missing script

@@ -1,0 +1,82 @@
+#!/usr/bin/env python
+
+from gstlal import lloidplots
+from glue.ligolw import utils, lsctables
+import numpy
+import pylab
+import matplotlib
+from matplotlib import ticker
+from mpl_toolkits.axes_grid1.inset_locator import zoomed_inset_axes
+from mpl_toolkits.axes_grid1.inset_locator import mark_inset
+import sys
+
+column1 = 'mchirp'
+column2 = 'mtotal'
+
+small_table = lsctables.table.get_table(
+	utils.load_filename('data/tmpltbank.xml'),
+	lsctables.SnglInspiralTable.tableName
+)
+n_small_templates = len(small_table)
+small_data = zip(small_table.get_column(column1), small_table.get_column(column2))
+
+# Create fill area (paint low chirp mass area black and skip drawing points there
+# in order to create a smaller PDF file
+
+@pylab.vectorize
+def soln(mchirp):
+	return pylab.roots([1., 0., -mchirp**(5./3), -1])[0]**3	
+
+fill_mtotal = pylab.linspace(2., 6., 50)
+fill_min_mchirp = 2 ** (-6./5) * fill_mtotal
+fill_max_mchirp = (fill_mtotal - 1) ** (3./5) * fill_mtotal ** (-1./5)
+#min_mchirp = 2. ** (-6./5) * 2.
+#max_big_mchirp = 5. ** (3./5) / 6. ** (1./5)
+#fill_mchirp = pylab.linspace(min_mchirp, max_big_mchirp, 50)
+#fill_min_mtotal = 2 ** (6./5) * fill_mchirp
+#fill_max_mtotal = soln(fill_mchirp)
+
+big_table = lsctables.table.get_table(
+	utils.load_filename('data/tmpltbank-pruned.xml'),
+	lsctables.SnglInspiralTable.tableName
+)
+n_big_templates = len(big_table)
+big_data = [x for x in zip(big_table.get_column(column1), big_table.get_column(column2)) if x not in small_data]
+#big_data2 = [x for x in big_data if x[0] >= max_big_mchirp]
+
+small_data = zip(*small_data)
+big_data = zip(*big_data)
+
+def foo():
[…]

### 2014-05-07 plot_bank.py
Adjust line and fill styles

@@ -48,7 +48,7 @@ small_data = zip(*small_data)
 big_data = zip(*big_data)
 
 def foo():
-	pylab.fill_betweenx(fill_mtotal, fill_min_mchirp, fill_max_mchirp, edgecolor='0.5', facecolor='0.5')
+	pylab.fill_betweenx(fill_mtotal, fill_min_mchirp, fill_max_mchirp, edgecolor='0.6', facecolor='0.6')
 	#pylab.plot(big_data[0], big_data[1], ',k', markersize=0.01)
 	pylab.plot(small_data[0], small_data[1], ',k')
 	#pylab.fill_between(fill_mchirp, fill_min_mtotal, fill_max_mtotal, edgecolor='none', facecolor='k')

### 2014-05-07 envelope.py
Update script

@@ -14,9 +14,9 @@ import sys
 # Place time slices
 
 # Generate plot
-from gstlal.svd_bank import read_bank
-bank = read_bank('data/svd_0_9.xml')
-pylab.figure(figsize=(2.6,2.5))
+from gstlal.svd_bank import read_banks
+bank, = read_banks('data/svd_0_9.xml', compat=True)
+pylab.figure(figsize=(3.5,2.5))
 ax = pylab.subplot(111)
 legend_artists = []
 legend_labels = []
@@ -38,5 +38,5 @@ pylab.xticks([-x.end for x in bank.bank_fragments if x.end > 10], [str(x.end) fo
 pylab.grid()
 pylab.xlabel('time relative to coalescence (s)')
 pylab.ylabel(r'strain amplitude')
-pylab.subplots_adjust(left=0.075, right=0.97, top=0.95, bottom=0.225)
+pylab.subplots_adjust(left=0.075, right=0.9, top=0.95, bottom=0.225)
 pylab.savefig(sys.argv[1])

### 2014-05-07 data/convert_svd_bank.py
Convert SVD bank files to new format

@@ -1,0 +1,50 @@
+#!/usr/bin/env python
+"""Convert legacy SVD bank files"""
+
+# Command line interface
+import argparse
+parser = argparse.ArgumentParser(description=__doc__)
+parser.add_argument(
+    'filename', metavar='FILENAME.xml[.gz]', nargs=1,
+    help='Name of file to modify in-place')
+args = parser.parse_args()
+
+# Late imports
+import os
+import glue.ligolw.utils
+import glue.ligolw.table
+import glue.ligolw.param
+import glue.ligolw.array
+import glue.ligolw.types
+import glue.ligolw.lsctables
+import numpy as np
+
+# Determine input filename
+filename, = args.filename
+_, ext = os.path.splitext(filename)
+gz = (ext == '.gz')
+
+# Read file, get root element
+xmldoc = glue.ligolw.utils.load_filename(filename, gz=gz)
+root, = xmldoc.childNodes
+
+# Add name to root element
+root.setAttribute(u'Name', u'gstlal_svd_bank_Bank')
+
+# Add empty sngl_inspiral table
+root.appendChild(glue.ligolw.lsctables.New(
+    glue.ligolw.lsctables.SnglInspiralTable))
+
+# Add empty bank_id param
+root.appendChild(glue.ligolw.param.new_param(
+    'bank_id', glue.ligolw.types.FromPyType[str], ' '))
+
+# Doesn't like empty logname
+logname = glue.ligolw.param.get_param(root, 'logname')
+logname.pcdata = ' '
+
+# Add empty autocorrelation_mask array
+root.appendChild(glue.ligolw.array.from_array(
+    'autocorrelation_mask', np.empty(0)))
+
+glue.ligolw.utils.write_filename(xmldoc, filename, gz=gz)

### 2014-05-07 envelope.py

@@ -15,7 +15,7 @@ import sys
 
 # Generate plot
 from gstlal.svd_bank import read_banks
-bank, = read_banks('data/svd_0_9.xml', compat=True)
+bank, = read_banks('data/svd_0_9.xml')
 pylab.figure(figsize=(3.5,2.5))
 ax = pylab.subplot(111)
 legend_artists = []
